### 2014-09-10 paper/CasillasFrank-Anticip.tex
first commit

\documentclass[authoryear, 12pt]{elsarticle}

\usepackage{graphicx}
% \usepackage{lineno}
\usepackage{amssymb}
\usepackage{wrapfig}
\usepackage{tipa}
\usepackage{float}

\journal{}

\begin{document}

\begin{frontmatter}

\title{Tracking and predicting turn structure during language acquisition}

\author[MPI]{Marisa Casillas}
% \address[StanfordLX]{Department of Linguistics, Stanford University}
\address[MPI]{Max Planck Institute for Psycholinguistics, Nijmegen}

\author[StanfordPSY]{Michael C. Frank}

\address[StanfordPSY]{Department of Psychology, Stanford University}

\begin{abstract}
Blah blah blah abstract needs to be written here.

\end{abstract}

\begin{keyword}
Turn-taking \sep Conversation \sep Development \sep Prosody \sep Lexical \sep Questions \sep Eye-tracking \sep Anticipation
%% MSC codes here, in the form: \MSC code \sep code
%% or \MSC[2008] code \sep code (2000 is the default)

\end{keyword}

\end{frontmatter}

% \linenumbers

\section{Introduction}
\label{sec:intro}

Spontaneous conversation is a universal context for using and learning language. Like other types of human interaction, it is organized at its core by the roles and goals of its participants. What sets conversation apart is its structure, however: sequences of interconnected, communicative actions that take place across alternating turns at talk. Sequential, turn-based structures in conversation are strikingly uniform across language communities and linguistic modalities. Turn-taking behaviors are also cross-culturally consistent in their basic features and the details of their implementation \citep{stivers2009, dingemanse2013, de-vosInPrep}. How does this ability develop?

Children participate in sequential coordination with their caregivers starting at three months of age---before they can rely on any linguistic cues in taking turns \citep{snow1977, jaffe2001, hilbrinkInPrep}. During their first year, infants know little about the lexical and syntactic systems of their language \citep[though they \textit{do} know a little about prosody;][]{christophe2001, bertoncini2013}. Of course, infant turn-taking is different from adult turn-taking in several ways. Infant turn-taking is heavily scaffolded by caregivers, has timing distinct from adult turn-taking, and lacks semantic content \citep{hilbrinkInPrep, jaffe2001, snow1977}. However, children's early, turn-structured social interactions are presumably a critical precursor to their conversational turn-taking. Early non-verbal interactions establish the protocol by which children come to use language with others.

% If we want to grasp children's linguistic input and practice, we need to understand their conversational skill development. And if we want to understand how children develop conversational skills, we need to take a close look at how turn-taking develops in tandem with language. 
In this study, we investigate when children begin to make predictions about upcoming turn structure, and how they integrate language into their predictions as they grow older. In what follows, we first review background about turn-taking and the state of current knowledge about adult turn prediction; we then discuss recent work on the development of turn-taking before turning to the details of our own study.

% For example, they show sensitivity to language-specific aspects of prosody in the first few days of life, and can use prosody for word segmentation and emotion recognition by 7--8 months \citep{jusczyk1999, johnson2001, grossmann2005}. In contrast, children don't know much about syntactic structure until their third year \citep{clark2009}.

\subsection{Turn-taking}

Turn-taking itself is not unique to conversation. Many other human activities are organized around sequential turns. Traffic in intersections and communication in computer networks both have characteristic turn-taking. Children's early games (e.g., give-and-take, peek-a-boo) have built-in, predictable turn structure \citep{ratner1978, ross1987}. Even monkeys take turns: Non-human primates such as marmosets and Campbell's monkeys vocalize contingently with each other in both natural and lab-controlled environments \citep{lemasson2011, takahashi2013}. In all these cases, turn-taking serves as a protocol for interaction, allowing the participants to coordinate with one another through sequences of contingent action.

Conversation distinguishes itself from non-conversational turn-taking behaviors, however, by the complexity of the turn sequencing involved. In the examples above (traffic, games, and monkeys) the set of sequence and action types is far more limited and predictable than what we find in everyday talk. For example, conversational turns come grouped into semantically-contingent sequences of action. The groups can span turn-by-turn exchanges (e.g., simple question--response, ``How are you?''--``Fine.'') or sequence-by-sequence exchanges (e.g., reciprocals, ``How are you?''--``Fine, and you?''--``Great!''). Sequences of action drive the conversation forward into the next, relevant sequences of talk (e.g., ''And you?''--``Great!''--``Why's that?''; \citealp{schegloff2007}). To take a turn, participants need to make predictions about what types of talk will be relevant next. In some cases, relevant next turns are somewhat obvious (e.g., question--response) while, in other cases, there are multiple relevant next actions to choose from, or no obvious next action at all (e.g., after a closed sequence).

Despite this complexity, conversational turn-taking is precise in its timing. Across a diverse sample of 10 languages, one study found a consistent average transition time of 0--200 msec at points of speaker switch \citep{stivers2009}. Current models of speech production suggest that it takes approximately 600 msec to produce a word, and even longer to produce a simple utterance \citep{levelt1989}. So, in order to achieve 200 msec turn transitions, speakers must begin formulating their response before the prior turn has ended. To formulate their response early on, speakers must track and anticipate what types of response might become relevant next. They also need to predict the content and form of upcoming speech so that they can launch their articulation at exactly the right moment. Prediction thus plays a key role in timely turn-taking.

\subsection{Adults' turn prediction}

Adults have a lot of information that could potentially help with their predictions about upcoming turn structure. Lexical, syntactic, and prosodic information (e.g., \textit{wh}- words, subject-auxiliary inversion, list intonation) can all inform addressees about upcoming linguistic structure \citep{de-ruiter2006, duncan1972, ford1996, bogelsUndRev}. Non-verbal cues (e.g., gaze, posture, and pointing) often appear at turn-boundaries as late indicators of an upcoming speaker switch \citep{rossano2009, stivers2010}. The sequential context of a turn can make it clear what will come next: answers after questions; thanks or denial after compliments, et cetera \citep{schegloff2007}. Prior work suggests that adult listeners primarily use lexicosyntactic information to accurately predict upcoming turn structure, while prosodic information is plays more of a secondary role \citep{de-ruiter2006}. For example, a mid-high ``continuation'' tone can cancel the possibility of a speaker switch at an otherwise syntactically- and pragmatically-complete turn point (e.g., at the start of a story-telling; \citealp{ford1996, bogelsUndRev}).

Recent experimental work on adult turn-taking has tried to tease apart which linguistic cues responders rely on in making online predictions about upcoming turn structure. De Ruiter and colleagues \citeyearpar{de-ruiter2006} asked participants to listen to snippets of spontaneous conversation and press a button whenever they anticipated that the current speaker was about to finish her turn. The snippets that participants listened to were controlled for the amount of linguistic information present; some snippets were normal, unmodified versions of the recording, but some had words and rhythm without pitch, pitch and rhythm but no words, neither words nor pitch, or were replaced entirely with a noise profile matched to the original recording. In the normal, unmodified snippets, participants' anticipatory button presses were highly accurate, falling an average of 180 msec before the phonetic end of the turns. Participants also made accurate predictions for the ``words and rhythm without pitch'' snippets; there was no significant difference compared to their performance in the normal  condition. But, in the snippets with no lexical information, participants' accuracy significantly declined. De Ruiter and colleagues \citeyearpar{de-ruiter2006} concluded that lexicosyntactic information is the primary linguistic signal listeners use for making online predictions about turn-end boundaries. They also found a significant, but secondary role for pitch: participants' accuracy in the pitch and rhythm (but no words) condition was still better than the neither-words-nor-pitch and the noise profile conditions. They concluded that pitch's importance for online prediction of turn-ends is restricted, and secondary to lexicosyntactic information.

Congruent evidence comes from studies varying the predictability of lexicosyntactic and pragmatic content. 
% Accurate prediction of linguistic and pragmatic content is essential to the mechanisms underlying adult turn taking. 
Adults anticipate turn ends better when they can more accurately predict exact words that will come next (\citealp{magyari2012}; see also \citealp{magyariUndRev}). They can also identify speech acts within the first word of an utterance \citep{gisladottirUndRev}, allowing them to start planning their response at the first moment possible. \citet*{bogelsmagyariInPrep} used a trivia set-up to get participants to answer questions in which the answer became known either early or late in the question (e.g., ``Which character, also called 007, appears in the famous movies?'' vs. ``Which character from the famous movies, is also called 007?''). Their participants showed a consistent neural response 500 msec after the answer became unambiguous, suggesting immediate response planning.

In another study investigating the relationship between lexicosyntactic and prosodic cues, \citet*{bogelsUndRev} used a combination of button-press and verbal responses.
 % to more explicitly test the role of prosodic cues in turn-boundary identification. 
They found that participants never verbally responded or pressed the ``anticipation'' button when hearing a syntactically-complete unit without a boundary tone. But utterance-level prosodic boundaries embedded in multi-utterance turns \textit{could} trick participants into anticipating a turn end 29\% of the time (see also de \citet{de-ruiter2006}:525). Their findings corroborate other corpus and experimental findings promoting a combination of cues (lexicosyntactic, prosodic, and pragmatic) as key for accurate turn-end prediction \citep{duncan1972, ford1996, hirvenkari2013}.

In sum, adults accurately and spontaneously make predictions about upcoming turn structure. They use a wide variety of cues in making their predictions and then act immediately upon them. Their predictive processing relies on a sophisticated body of knowledge about linguistic structure, non-verbal signals, and social actions. The acquisition of turn-taking skills must then be tied to children's knowledge about language, gaze, gesture, and social cues. But children's turn taking starts early in infancy, long before first words or gestures emerge; a primary role for lexicosyntactic cues thus doesn't fit well with children's pre-verbal turn-taking.

\subsection{Children's turn prediction}

\subsubsection{Observational studies}

The majority of work on children's early turn-taking has focused on observations of spontaneous interaction. Children's first turn-like structures appear as early as two to three months in proto-conversation with their caregivers. During proto-conversations, caregivers interact with their infants as if they were capable of making meaningful contributions; they take every look, vocalization, arm flail and burp as ``utterances'' in the joint discourse \citep{snow1977, jaffe2001}. Infants catch onto the structure of proto-conversations quickly. By three to four months they notice disturbances to the contingency of their caregivers' response and, in reaction, change the rate and quality of their vocalizations \citep{k-bloom1988, masataka1993}. Infants at this age also notice changes to social contingency outside of turn structure. In the Still Face paradigm, caregivers interact with their infants and then suddenly halt, taking on a neutral expression with a sustained gaze. When faced with this sudden disappearance of social contingency, infants three months and older try a range of methods to reinitiate the interaction, such as vocalization, reaching, and smiling before looking away or getting upset \citep{rochat1998, toda1993}.

The timing of children's responses to their caregivers' speech shows a non-linear pattern of fall-rise-fall from early infancy to middle childhood. Infants' turn timing at three months is often too early or too late: they start vocalizing in overlap on nearly 40\% of their caregivers' turns, and their non-overlapped vocalizations come after an average latency of 675 msec (adult average: 200 msec). Between four and nine months, children begin to reduce the number of turns happening overlap while also improving on their average response latency. But then children's response latencies slow down again, peaking at average gaps of 1100 msec at nine months, with only very gradual improvement after that \citep{hilbrinkInPrep}. While children's avoidance of overlap is nearly adult-like by nine months, the timing of their non-overlapped responses stays much longer than the 200 msec standard for the next several years \citep{ervin-tripp1979}.

The protracted development of children's timing may be attributable to their linguistic development: Taking turns on time is easier when the response is a simple vocalization rather than a linguistic utterance. Integrating language into the turn-taking system may be one major factor in children's delayed responses \citep{casillasUndRev}. If response planning (i.e., language production) is the primary hurdle in children's spontaneous turn taking, we should find evidence that children understand turn-taking behaviors before they are able to produce the behaviors themselves; this hypothesis has been more recently explored in experimental settings.

\subsubsection{Experimental studies}

Children begin to develop specific expectations about conversational behavior before they begin to speak. Sometime between four and six months, children begin to attend differently to face-to-face and back-to-back conversation; six-month-olds follow conversational speakers more with their gaze when at least one speaker is looking at the other \citep{augusti2010}. At ten months, infants expect people to look and talk at other people, and not to objects \citep{beier2012}. At twelve months infants expect to see responses to verbal (but not non-speech) utterances in face-to-face contexts \citep{thorgrimssonUndRev}.

There are mixed results regarding when children begin to anticipate turn structure in conversation. One study found that 12-month-olds make more predictive gaze shifts while watching human verbal conversation compared to object non-speech ``conversation'' \citep{bakker2011}, but another only found a similar effect at 36 months \citep{hofsten2009}. Neither of these studies had baselines to which the turn-relevant looking behavior could be compared, however. Such a baseline measurement is critical because there may be developmental differences in gaze shifting between conversational participants, even if such shifting is not related to turn structure. These developmental differences could produce artifactual changes in measures of turn-contingent shifting.

Keitel and colleagues \citeyearpar{keitel2013} addressed the random baseline issue in their study of 6-, 12-, 24-, and 36-month-olds, finding that children's anticipatory gaze frequency was only greater than chance for 36-month-olds and adults. Their study was the first to focus on the role of linguistic processing in children's turn predictions. They showed their participants two types of conversation videos: one normal and one with flattened pitch (i.e., with removed intonation contours), finding that only three-year-olds were affected by a lack of intonation. The adult control group made equal numbers of anticipatory looks in the videos, with and without intonation contours, consistent with prior adult findings \citep{de-ruiter2006}. Keitel and colleagues concluded that children's ability to predict upcoming turn structure relies on their ability to comprehend the stimuli (emerging around 3;0), especially with respect to semantic access. They also suggest that intonation takes a secondary role in turn prediction, but only \textit{after} children acquire more sophisticated, adult-like language comprehension systems (sometime after 3;0).

Although the Keitel et al. \citeyearpar{keitel2013} study constitutes a substantial advance over previous work, it nevertheless has several limitations. Because these limitations directly inform our own study design, we review them in some detail. First, their estimates of baseline gaze frequency were computed only during conversational turns (not pauses), while their estimates of anticipatory shifting were computed during pauses. Assuming that gaze switches happen most often during the gap between turns \citep{hirvenkari2013}, their choice of a ``random'' (baseline) gaze window would have maximized the possibility of finding a difference between ``random'' and anticipatory gaze frequencies.
% \footnote{If so, the null effect for children under 3;0 could be given a stronger interpretation and the significant effect for three-year-olds and adults could be given a weaker interpretation.} 
A stronger baseline would be to compare participants' looking behavior at turn transitions to their looking behavior during randomly-selected windows of time throughout the stimulus. We follow this approach in our work.

Second, the conversation stimuli they used were somewhat unusual. The average gap between turns was 900 msec, which is much longer than typical adult timing, where gaps are often around 200 msec \citep{stivers2009}. The speakers in the video were also asked to minimize their movements while performing a scripted, adult-directed conversation, which would have created a somewhat unnatural stimulus. In addition, in order to produce more naturalistic conversation, it would have been ideal to localize sources for the two voices in the video (i.e., to have the voices come out of separate left and right speakers). But both voices were recorded and played back on the same audio channel, which may have made it more difficult to distinguish the speakers. We again attempt to address these issues in our current study.

% A single-channel recording was most problematic for the no-intonation condition, where the two voices were flattened to have the same pitch. 
Despite these minor methodological issues, the Keitel et al. \citeyearpar{keitel2013} study still demonstrates intriguing age-based differences in children's ability to predict upcoming turn structure, and the results suggest that both semantic and prosodic development \textit{do} play a role in children's looking patterns. Our current work thus takes this paradigm as our starting point.

\subsection{The Current Study}

We report here on the role of linguistic processing in children's predictions about upcoming turn structure. In two eye-tracking experiments, we measured children's anticipatory gaze to upcoming responders while controlling for the amount of lexicosyntactic and prosodic information available. Experiment 1 uses a paradigm in which participants view video clips of naturalistic conversations across a number of different languages, so as to create conditions under which lexicosyntactic information was differentially available, but prosodic and non-linguistic cues were relatively constant. This paradigm showed minimal differences between the predictive looking behavior of preschoolers and adults. In Experiment 2, we created artificial (puppet) visual stimuli to better separate these cues from one another. In this more controlled paradigm, we found that children's predictive looking behavior improved from ages one to six, but that even one-year-olds made more anticipatory looks than would be expected by chance.

In both experiments, children consistently looked faster to responders after hearing questions, compared to non-questions. Both prosodic and lexicosyntactic information played a role in children's predictions about turn structure, but the two information sources were used differently at different ages. Our findings overall support an account in which predictive processes for turn-taking in conversation are present early but their integration with linguistic information takes substantial practice.

\section{Experiment 1}
\label{sec:exp1}

We recorded participants' eye movements as they watched six short videos of two-person (dyadic) conversation. Each video featured an improvised discourse in one of five languages (English, German, Hebrew, Japanese, and Korean). The participants, all native English speakers, were only expected to understand the two videos in English. We showed participants non-English videos to limit their access to lexical information while maintaining their access to other cues to turn boundaries (e.g., prosody, gaze, breath, phrase final lengthening). Using this method, we compared children and adult's anticipatory looks from the current speaker to the upcoming speaker at points of turn transition in English and non-English videos.

\subsection{Methods}
\label{sec:methods1}

\subsubsection{Participants}

We recruited 74 children between ages 3;0--6;0 and 11 undergraduate adults to participate in the study. Our child sample included 19 three-year-olds, 32 four-year-olds, and 23 five-year-olds, all enrolled in a local nursery school. All participants were native English speakers. Approximately one-third (N=25) of the children's parents and teachers reported that their child regularly heard a second (and sometimes third or further) language, but only one child frequently heard a language that was used in our non-English video stimuli, and we excluded his data from analyses. None of the adult participants reported fluency in a second language.

\begin{figure}[ht]
\begin{center}
\includegraphics[width=0.6\textwidth]{figures/FIG-FL-stim.png}
\end{center}
\caption{Example frame from a conversation video used in Experiment 1.} 
\label{speakers}
\end{figure}

\subsubsection{Materials}

\textit{Video-recordings}. We recorded pairs of talkers while they conversed in a sound-attenuated booth. Each talker was a native speaker of the language being recorded, and each talker pair was male-female. Using a Marantz PMD 660 solid state field recorder, we captured audio from two lapel microphones, one attached to each participant, while simultaneously recording video from the built-in camera of a MacBook. The talkers were volunteers, all members of the local post-graduate community, and were acquainted with their recording partner ahead of time. Each recording session began with a 20-minute warm-up period of spontaneous conversation during which the pair talked for five minutes on four topics (favorite foods, entertainment, hometown layout, and pets). Then we asked talkers to choose a new topic---one relevant to young children (e.g., riding a bike, eating breakfast)---and to improvise a dialogue on that topic. We asked them to speak as if they were on a children's television show in order to elicit child-directed speech toward each other. We recorded until the talkers achieved at least 30 seconds of uninterrupted discourse with enthusiastic, child-directed speech. Most talker pairs took less than five minutes to complete the task, usually by agreeing on a rough script at the start. We encouraged talkers to ask at least a few questions to each other during the improvisation. The resulting conversations were therefore not entirely spontaneous, but were as close as possible while still remaining child-oriented in topic, prosodic pattern, and lexicosyntactic construction.\footnote{All of the non-English talkers were fluent in English as a second language, and some fluently spoke a third or more language. We chose male-female pairs as a natural way of creating contrast between the two talker voices. See the videos here: NOTE: add link.}

After recording, we hand-aligned the audio and video files, and cropped each recording to the 30-second interval with the most turn activity. Because we recorded the conversations in stereo, the male and female voices came out of separate speakers during video playback. This gave each voice in the videos a localized source (from the left or right loudspeaker). We coded each turn transition in the videos for language group (English vs. non-English), inter-turn gap duration (in milliseconds), and speech act (question vs. non-question). The non-English stimuli were coded for speech act from a monolingual English-speaker's perspective, i.e., which turns ``sound like'' questions, and which don't. The coding for speech act in non-English videos was completed independently by two coders and then checked and discussed for agreement.

Because the conversational stimuli were recorded semi-spontaneously, the duration turn transitions and the number of speaker transitions in each video was variable. We measured the duration of each turn transition from the audio recording associated with each video. We excluded turn transitions longer than 550 msec and shorter than 0 msec (those with transitional overlap) from analysis.\footnote{Overlap occurs when an addressee begins a new turn before the current turn is finished. When overlap occurs, observers cannot switch their gaze in anticipation of the response because the response begins \textit{earlier} than expected; participants expect conversations to proceed with ``one speaker at a time'' \citep{sacks1974}. As such, they would still be fixated on the prior speaker when the overlap started, and then would have to switch their gaze \textit{reactively} to the responder.} This left approximately equal numbers of turn transitions available for analysis in the English (N$=$22) and non-English (N$=$21) videos. On average, the inter-turn gaps for English videos (mean$=$273 msec, median$=$291 msec) were slightly longer than for non-English videos (mean$=$226 msec, median$=$205 msec). The longer gaps in the English videos could give them a slight advantage: our definition of an ``anticipatory gaze shift'' includes shifts that are initiated during the gap between turns (Figure 2), so participants had slightly more time to make anticipatory shifts in the English videos.

Questions made up approximately half of the turn transitions in the English (54\%; N$=$12) and non-English (48\%; N$=$10) videos. In the English videos, inter-turn gaps were slightly shorter for questions (mean$=$271 msec, median$=$254 msec) than non-questions (mean$=$276 msec, median$=$300 msec). The opposite was true for the non-English videos: question transitions (mean$=$264 msec, median$=$257 msec) had longer inter-turn gaps than non-question transitions (mean$=$191 msec, median$=$168 msec). Thus non-question (i.e., declarative) transitions in non-English videos were the briefest type of turn transition in the stimuli on average.

\subsubsection{Procedure} 
Participants sat in front of an SMI 120Hz corneal reflection eye-tracker mounted beneath a large flatscreen display. The display and eye-tracker were secured to a table with an ergonomic arm that allowed the experimenter to position the whole apparatus at a comfortable height, approximately 60 cm from the viewer. We placed stereo speakers on the table, to the left and right of the display. Before the experiment started, we warned adult participants that they would see videos in several languages and that, though they weren't expected to understand the content of non-English videos, we \textit{would} ask them to answer general, non-language-based questions about the conversations. Then after each video we asked participants one of the following randomly-assigned questions: ``which speaker talked more?'', ``which speaker asked the most questions?'' ``which speaker seemed more friendly?'', and ``did the speakers' level of enthusiasm shift during the conversation?'' We also asked if the participants could understand any of what was said after each video. The participants responded verbally while an experimenter noted their responses.

Children were less inclined to simply sit and watch videos of conversation in languages they didn't speak, so we used a different procedure to keep them engaged. The experimenter started each session by asking the child about what languages he or she could speak, and about what other languages he or she had heard of. Then the experimenter expressed her own enthusiasm for learning about new languages, and invited the child to watch a video about ``new and different languages'' together. If the child agreed to watch, the experimenter and the child sat together in front of the display, with the child centered in front of the tracker and the experimenter off to the side. If the child began to look bored during video playback, the experimenter would talk during the filler videos, either commenting on the previous conversation (``That was a neat language!'') or giving the language name for the next conversation (``This next one is called Hebrew. Let's see what it's like.'') The experimenter's comments reinforced the video-watching as a joint task.

All participants (child and adult) completed a five-point calibration routine before the first video started. We used a dancing Elmo for the children's calibration image. During the study, participants watched all six 30-second conversation videos with 15--30 second filler videos in-between them (e.g., running puppies, singing muppets, flying bugs). The first and last conversations were in American English and the intervening conversations were Hebrew, Japanese, German, and Korean. The presentation order of the non-English videos was shuffled into four lists, which participants were assigned to randomly. The entire experiment, including instructions, took 10--15 minutes.

\subsubsection{Data preparation and coding}
To determine whether participants predicted upcoming turn transitions, we needed to define a set of criteria for what counted as an anticipatory gaze shift. Prior work using similar experimental procedures has found that adults and children make anticipatory gaze shifts to upcoming talkers within a wide time frame; the earliest shifts occur before the end of the prior turn, and the latest occur after the onset of the response turn, with most shifts occurring in the inter-turn gap (Keitel et al., 2013; Hirvenkari, 2013; Tice and Henetz, \citeyear{TiceHenetz11}). Following prior work, we measured how often our participants shifted their gaze from the prior to the upcoming speaker \textit{before} the shift in gaze could have been initiated in reaction to the onset of the speaker's response. In doing so, we assumed that it takes children 333 msec to plan an eye movement, following Fernald and colleagues' (2008) similar `looking while listening' method.

We checked each participant's gaze at each turn transition for three characteristics: (1) that the participant fixated on the prior speaker for at least 100 msec at the end of the prior turn, (2) that sometime thereafter the participant switched to fixate on the upcoming speaker for at least 100 ms, and (3) that the switch in gaze was initiated within the first 333 msec of the response turn, or earlier. These criteria guarantee that we only counted gaze shifts when: (1) participants were tracking the previous speaker, (2) switched their gaze to track the upcoming speaker, and (3) did so before they could have simply reacted to the onset of speech in the response. Under this assumption, a gaze shift that is initiated within the first 333 msec of the response (or earlier) was planned \textit{before} the child could react to the onset of speech itself. For adult participants we used 200 msec (instead of 333 msec) as the assumed time to plan an eye movement.

\begin{figure}[ht]
\begin{center}
\includegraphics[width=0.7\textwidth]{figures/FIG-AnticipCriteria.png}
\end{center}
\caption{Schematic summary of criteria for anticipatory gaze shifts from speaker A to speaker B during a turn transition. The criteria were the same for adults and children, except children were assumed to have a 333 msec window for planning eye movements, while adults were only given 200 msec.} 
\label{speakers}
\end{figure}

Using these criteria, we created a three-step algorithm for detecting anticipatory gaze switches. Consider, for example, when a child observes a turn transition between speaker A and speaker B (Figure 2). First, the child would need to fixate on A for at least 100 msec at the end of A's turn. But what is the relevant window for fixation? Consistent with the distribution of responses in explicit turn-boundary prediction tasks (de Ruiter et al., 2006), we counted any early switches as ``truly turn-related'' as long as they occurred in the final 333 msec of A's turn (200 msec for adults). To allow for these early switches, the window for fixating on A must start 433 msec before the end of A's turn (300 msec for adults). Second, given that the child fixated on A, we checked to see if the child also fixated on B. Because we allowed for early switches, the first possible fixation on B could start 333 msec before the end of A's turn. Assuming that it took the child 333 msec to plan an eye movement, the latest point when the gaze could have \textit{anticipated} speech would be 333 msec after B's turn began---a switch after this point could simply be a reaction to the onset of B's speech. Thus, the relevant window for fixating on B started 333 msec before the gap and ended 433 msec after it (200 msec and 300 ms, respectively, for adults). Success on the first two criteria jointly implied that the child made an anticipatory switch in gaze. Our third step was to estimate when the switch occurred. Starting at a point when we knew participants were fixated on A (from step 1), we incremented forward in time to the first gaze measurement where the child looked away from A. We recorded the look-away timestamp as the time of gaze launch to the upcoming speaker.

\textit{Random baseline analysis}. To ensure that participants' looking behavior was turn-related and not simply the result of random switching, we corrected for the probability of making an anticipatory switch at random. We estimated the baseline chance of making an anticipatory switch at each turn transition for each person. We ran all participants' eye-tracking data through switch identification with 100 randomly-shuffled versions of the original turn transition windows (Figure 3). For each of the 100 randomly-shuffled versions, we recorded and then averaged whether or not each participant made an anticipatory switch at each ``turn transition''. We then subtracted the random baseline value from each participant's actual switch result (1 or 0) for each turn transition to arrive at the final, baseline-corrected values. Subtracting the random baseline this way can give us an estimate of how often participants switched above chance across ages, conditions, and transition types. If participants make the same number of anticipatory switches, regardless of turn transition location, there would be no evidence that their switching behavior was linked to the conversation stimulus.

\begin{figure}[ht]
\begin{center}
\includegraphics[width=0.6\textwidth]{figures/FIG-ShuffledWindows.png}
\end{center}
\caption{Example of shuffling for five turn transition windows. The windows were +/-433 msec around the inter-turn gap for children, and from +/-300 msec for adults.} 
\label{shuffling}
\end{figure}

\subsection{Results and discussion}
\label{sec:results1}

Participants looked at the screen most of the time during video playback (81\% and 91\% on average for children and adults, respectively). They primarily kept their eyes on the person who was currently speaking in both English and non-English videos: they gazed at the current speaker between 38\% and 63\% of the time, looking back at the addressee between 15\% and 20\% of the time (Table 1). Even three-year-olds looked more at the current speaker than anything else, whether the videos were in a language they could understand or not. Children looked at the current speaker less than adults did during the non-English videos. Despite this, their looks to the addressee did not increase substantially in the non-English videos, indicating that their looks away were probably related to boredom rather than confusion about ongoing turn structure. Overall, participants' pattern of gaze to current speakers indicated that they performed basic turn tracking during the videos, regardless of language.

\begin{table}[H]
\begin{center}
  \begin{tabular}{llcccc}
    \hline
    Age group & Condition & Speaker & Addressee & Other onscreen & Offscreen\\ 
    \hline
    3 & English & 0.61 & 0.16 & 0.14 & 0.08 \\ 
    4 & English & 0.60 & 0.15 & 0.11 & 0.13 \\ 
    5 & English & 0.57 & 0.15 & 0.16 & 0.12 \\ 
    Adult & English & 0.63 & 0.16 & 0.16 & 0.05 \\ 
    3 & Non-English & 0.38 & 0.17 & 0.20 & 0.25 \\ 
    4 & Non-English & 0.43 & 0.19 & 0.21 & 0.18 \\ 
    5 & Non-English & 0.40 & 0.16 & 0.26 & 0.18 \\ 
    Adult & Non-English & 0.58 & 0.20 & 0.16 & 0.07 \\ 
%   Overall & & 0.49 & 0.17 & 0.18 & 0.16 \\
    \hline
  \end{tabular}
\end{center}
  \caption{Average proportion of gaze to the current speaker and addressee during periods of talk.}
\label{looktableexp1}
\end{table}

We identified anticipatory gaze switches for all 43 usable turn transitions, based on the criteria and algorithm outlined in Section 2.1.4. Because participants had greater access to lexical and syntactic information in English, we expected to see greater anticipation in the English videos than in the non-English videos. We also predicted that anticipation would be greater following questions compared to non-questions; questions have early cues to upcoming turn transition (e.g., \textit{wh-} words, subject-auxiliary inversion), and also make a next response immediately relevant. We also predicted that anticipatory looks would increase with development, along with children's increased linguistic competence.

To estimate the effects of linguistic condition and age on switching behavior, we corrected each participant's data by subtracting their random baseline estimations from their actual switching data for each turn transition. All comparisons below are made with the baseline-corrected values.

\begin{figure}[ht]
\begin{center}
\includegraphics[width=0.8\textwidth]{figures/FIG-randvsreal-FL.png}
\end{center}
\caption{Proportion anticipatory gaze switches made for actual (solid) and randomly-shuffled (dashed) turn transitions in each condition, and across ages. Error bars represent the standard error of the mean.} 
\label{randvsrealFL}
\end{figure}

Participants at all ages made anticipatory gaze switches more often than would be expected by chance, whether they could understand the language in the video or not (Figure 4). Anticipatory switch behavior in the non-English videos could not have been based on lexicosyntactic information, so both adults and children must have relied on other sources of information to make their predictions (e.g., prosody and non-verbal signals).

Overall, a greater proportion of participants made anticipatory switches in the English videos (0.17) compared to the non-English videos (0.09). But the tendency to make more anticipatory switches for English than non-English videos disappeared with age. While anticipatory switches increased with age for non-English videos, they decreased with age for English videos, resulting in nearly equal looking behavior for the two language conditions for adults  (Figure 5). Otherwise, age-based differences in predictive gaze shifts were minor.

\begin{figure}[ht]
\begin{center}
\includegraphics[width=0.5\textwidth]{figures/FIG-conditions-FL.png}
\end{center}
\caption{Proportion anticipatory gaze switches made for turn transitions in each language condition, and across ages. Error bars represent the standard error of the mean.} 
\label{conditionsFL}
\end{figure}

\begin{figure}[H]
\begin{center}
\includegraphics[width=0.8\textwidth]{figures/FIG-QvsNQ-FL.png}
\end{center}
\caption{Proportion anticipatory gaze switches made for question- (dashed) and non-question- (dotted) turn transitions in each condition, and across ages. Error bars represent the standard error of the mean.} 
\label{questionsFL}
\end{figure}

Participants also showed an advantage for questions. Children made anticipatory switches twice as often when they heard a question, compared to a non-question (18.7\% and 7.3\%). But the advantage for questions was not universal---it depended on language condition and age. The advantage for questions didn't emerge until age five in the non-English videos (Figure 6).

We confirmed these findings in two linear mixed-effects models of baseline-corrected anticipatory switches \citep{lme4, R}. We built one model each for children and adults. The child model included condition (English vs. non-English)\footnote{Because each non-English language was represented by a single stimulus, we cannot compute reliable cross-language differences. Gaze behavior might be best for languages that have more structural overlap English. English speakers can make predictions about the strength of upcoming Swedish prosodic boundaries nearly as well as Swedish speakers do, but Chinese speakers are at a disadvantage in the same task \cite{carlson2005}. We would need multiple items from each of the language to check for similarity effects of specific linguistic features.}, transition type (question vs. non-question), age, and duration of the inter-turn gap as predictors, with full interactions between condition, transition type, and age.  We included the duration of the inter-turn gap as a predictor since longer gaps also provide more opportunities to make anticipatory switches (Figure 2). We additionally included random effects of item (turn transition) and participant, with random slopes of condition, transition type, and their interaction for participants \citep{barr2013}. The adult model was exactly the same as the children's, excluding age-related effects.

Children's anticipatory gaze switches showed a main effect of language condition (\textit{$\beta$}=-0.448, \textit{SE}=0.126, \textit{t}=-3.56). As predicted, children made more anticipatory looks while they were watching videos that they could understand. In the English videos, children had access to lexicosyntactic information, which could have helped them make better predictions about upcoming turn structure. Although children performed better on the English videos, lexicosyntactic information was not exclusively responsible for their anticipatory switches; children still performed remarkably well when no lexical information was present at all, and even three-year-olds made anticipatory switches at an above-chance rate in the non-English videos.

Contrary to our expectation, there was no main effect of age (\textit{$\beta$}=-0.03, \textit{SE}=0.027, \textit{t}=-1.145). Children's anticipatory switches \textit{did} increase with age for the non-English videos, but it also decreased with age for the English videos, resulting in a significant interaction of age and language condition (\textit{$\beta$}=0.083, \textit{SE}=0.027, \textit{t}=3.063). We had expected that, as children got older, they would also show more anticipatory switches. But this is contradicted by decreased looks with age for English videos, which even extended to the adult data (Figure 5). One possible explanation is that older children and adults found the videos in English easy to comprehend, and so tracked the conversation less closely. Compared to three-year-olds, adults and older children spent more time looking away from the current speaker and addressee in the English conversations. In contrast, compared to three-year-olds, they spent more time looking at the current speaker and less time looking away in the non-English conversations (Table 1).

It is also possible that the timing of anticipatory gaze switches changes as children get older. That is, the moment at which observers chose to launch their gaze away from the current speaker might change with age. For example, children could have reacted more immediately to early cues in the turn (e.g., ``Which one...'') as they got older. If a change in timing were sufficiently large, we might not capture it in our analysis window, which was designed to capture switches close to the turn transition. To test for a change in switch timing, we would have to run new analyses on a larger set of turn transitions---ones with controlled early and late cues \citep[see, e.g.,][]{bogelsmagyariInPrep}. Within the current analysis windows there were some ``early switches'' (i.e., those launched before the end of the prior turn). But participants made approximately equal numbers of early switches, regardless of age and condition: 11--15\% (Figure 7). The lack of increased early switching with age might suggest that changes in timing are either larger than the analysis window, or are not at play.

\begin{figure}[ht]
\begin{center}
\includegraphics[width=0.8\textwidth]{figures/FIG-cumulative-FL.png}
\end{center}
\caption{Cumulative proportion of anticipatory gaze switches planned during the transition window, across conditions and ages. Age is indicated by hue (light to dark; darker = older). The solid line is the offset of the pre-gap turn, and the dashed line is the median onset of the post-gap turn in each condition. ``Early switches'' occur before 0.} 
\label{cumulativeFL}
\end{figure}

Children made more anticipatory switches after hearing questions than non-questions, but the advantage for questions was driven by age and linguistic access. There was no main effect of transition type alone (\textit{$\beta$}=-0.209, \textit{SE}=0.151, \textit{t}=-1.384). Rather, the model confirmed a two-way interaction of linguistic condition and transition type (\textit{$\beta$}=0.445, \textit{SE}=0.19, \textit{t}=2.344). Presumably, question effects were larger in the English videos than the non-English videos because there were more cues to questionhood in English (i.e., non-verbal, prosodic, \textit{and} lexicosyntactic). Children younger than five did not distinguish questions from non-questions in the non-English videos (Figure 6). The children's model confirmed a three-way interaction of age, linguistic condition, and transition type (\textit{$\beta$}=-0.091, \textit{SE}=0.042, \textit{t}=-2.183). Lexicosyntactic information might then be critical for young children to identify questions in conversation. Cues like subject-auxiliary inversion, and \textit{wh}-words occur early in the turn and are clear indicators that an answer will come next. Importantly, the turns marked as ``questions'' in the non-English data were not real questions, but turns that \textit{sounded like} questions to native English speakers. Question-like prosodic cues in the non-English videos could have been markedly different from question prosody in English, making it more difficult for three- and four-year-olds to identify ``questions'' in those conversations.

The increased switching after questions indicates that, so long as they have native lexical and prosodic cues, children as young as three spontaneously distinguish questions from other types of speech acts in third-party conversation. Children may treat questions differently from non-questions in conversation because they are predictably followed by an immediate response (i.e., the answer). Because the form of an answer is partially determined by its question (e.g., locative expressions for ``where'' questions), children could even be predicting what type of response they will hear, and look at the responder for confirmation. In contrast, non-questions do not usually require an immediate response, so it is not as easy to predict what will come next.

Inter-turn gap duration did not affect children's anticipatory switches (\textit{$\beta$}=0.202, \textit{SE}=0.151, \textit{t}=1.340), but it was the main factor driving adults' switches (\textit{$\beta$}=0.648, \textit{SE}=0.047, \textit{t}=13.855). The adult model also showed a marginal effect of transition type, with adults looking more after questions compared to non-questions (\textit{$\beta$}=-0.117, \textit{SE}=0.069, \textit{t}=-1.704). Language condition and its interaction with transition type were not significant. Language did not affect adults' anticipatory looking behavior, possibly because the animated, child-directed prosody and non-verbal cues made it possible for them to follow the interaction and track the current speaker easily.

A few results from Study 1 were unexpected. First, we found no significant main effect of age in the children's data, and we saw very little difference between adults and children (besides the age-related interactions). It appears that, especially in English, children were as good at tracking the current speaker and anticipating upcoming turns as adults were. To find out when children first begin making spontaneous anticipatory looks to responders, we will have to look at younger children. Second, children, but not adults, showed an advantage for lexicosyntactic information in their anticipatory switches. An advantage for lexicosyntactic information is consistent with prior work on adult turn taking \citep{de-ruiter2006}, so it's curious that the adults did not show a difference in anticipatory switching between videos they did and did not understand. As mentioned, they may have effectively used the non-verbal cues in the videos to track the interactions fluently and, consequently, found the interactions so easy to follow that they did not track the turn-taking closely. To identify which \textit{linguistic} cues adults and children rely on, we have to control the presence of non-verbal cues in the videos.

Finally, we had initially predicted that children would show early advantages for prosodic cues over syntactic ones because children begin developing language-specific knowledge about prosody long before lexicosyntax. We weren't able to directly test this prediction in Study 1 because we used a limited range of ages (three- to five-year-olds) and because we only controlled for lexicosyntactic information. Children had prosodic cues in both English and non-English videos, but the prosodic signal was non-native in the latter case. So, although using non-English videos was a natural way to control for lexical information, it also rendered us unable to assess prosody's role in participants' predictions. Relatedly, the advantage for questions was greater in the English videos than the non-English videos, lending the possibility that lexicosyntactic cues are essential for children in identifying questions during conversation. To test this idea, we would need to directly compare lexicosyntactic and prosodic cues in the participants' native language.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Experiment 2}
\label{sec:exp2}

We improved our design by using native-language stimuli, controlling for lexical \textit{and} prosodic information, eliminating non-verbal cues, and testing children from a wider age range. All of the videos in Study 2 were in the participants' native language (American English). To tease apart the role of lexical and prosodic information, we phonetically manipulated the speech signal for pitch, rhythm, and lexical access. By testing one- to six-year-olds we hoped to find the developmental onset of turn-predictive gaze. We also hoped to see whether the roles of prosody and lexicosyntax change as children get older.

Non-verbal cues in Study 1 (e.g., gaze and gesture) could have helped participants make predictions about upcoming turn structure  \citep{rossano2009, stivers2010}. Since our focus is on linguistic cues, we eliminated all gaze and gestural signals in Study 2 by replacing the videos of human actors with videos of puppets. Puppets are less realistic and expressive than human actors but, since we wanted to eliminate gestural and gaze cues, the puppets created a natural context for having somewhat motionless talkers in the videos. Additionally, the prosody-controlled condition included small but global changes to syllable duration that would have required complex video manipulation or precise re-enactment with human talkers, neither of which was feasible. For these reasons, we decided to substitute puppet videos for human videos in the final stimuli.

As in the first study, we recorded participants' eye movements as they watched six short videos of dyadic conversation, and then analyzed their anticipatory glances from the current speaker to the upcoming speaker at points of turn transition.

\subsection{Methods}
\label{sec:methods2}

\subsubsection{Participants}
We recruited 27 undergraduate adults and 129 children between ages 1;0--7;0 to participate in our study. We recruited our child participants from the Children's Discovery Museum in San Jose, California, targeting approximately 20 children for each of the six 1-year age groups (range$=$20--23). All participants were native English speakers, though some parents (N$=$27) reported that their child heard a second (and sometimes third) language at home. None of the adult participants reported fluency in a second language. We ran Study 2 at a children's museum because it gave us access to a more diverse age range of children from more variable socioeconomic backgrounds.

\begin{figure}[ht]
\begin{center}
\includegraphics[width=0.9\textwidth]{figures/FIG-EN-stim.png}
\end{center}
\caption{The six puppet pairs (and associated audio conditions). Each pair was linked to three distinct conversations from the same condition across the three experiment versions.}
\label{puppets}
\end{figure}

\subsubsection{Materials}
We created six short videos of improvised, child-friendly conversation. To eliminate non-verbal cues to turn transition and to control the types of linguistic information available in the stimuli we (1) recorded improvised conversations, (2) phonetically manipulated those recordings to limit the availability of prosodic and lexical information, and (3) recorded a new set of videos that featured puppets as talkers, using the manipulated audio as the puppets' speech.

\textit{Audio recordings}. The recording session was set up in the same way as the first study, but with a shorter warm up period (5--10 minutes) and a pre-determined topic for the child-friendly improvisation (`riding bikes', `pets', `breakfast', `birthday cake', `rainy days', or `the library'). All of the talkers were native English speakers, and were recorded in male-female pairs. As before, we asked talkers to speak ``as if they were on a children's television show'' and to ask at least a few questions during the improvisation. We cut each audio recording down to the 20-second interval with the most turn activity. The 20-second clips were then phonetically manipulated and used in the final video stimuli.

\textit{Audio Manipulation}. We created four versions of each audio clip: \textit{normal}, \textit{words only}, \textit{prosody only}, and \textit{no speech}. That is, one version with a full linguistic signal (\textit{normal}), and three with incomplete linguistic information (hereafter ``limited cue'' conditions). The \textit{normal} clips were the unmanipulated, original audio clips. The \textit{words only} clips were manipulated to have robot-like speech: we flattened the intonation contours to each talker's average pitch (F0) and we reset the duration of every nucleus and coda to each talker's average nucleus and coda duration.\footnote{We excluded hyper-lengthened words like [w\textipa{aU:}] `woooow!'. These were rare in the clips.} We made duration and pitch manipulations using PSOLA resynthesis in Praat \citep{Praat}. Thus, the \textit{words only} versions of the audio clips had no pitch or durational cues to upcoming turn boundaries, but did have intact lexicosyntactic cues. We created the \textit{prosody only} clips by low-pass filtering the original recording at 500 Hz with a 50 Hz Hanning window (following de Ruiter et al., 2006). This creates a ``muffled speech'' sound because low-pass filtering removes most of the phonetic information used to distinguish between phonemes. Thus, the \textit{prosody only} versions of the audio clips lacked lexical information, but retained their intonational and rhythmic cues to upcoming turn boundaries. The \textit{no speech} condition served as a non-linguistic baseline. For this condition, we replaced the original clip with multi-talker babble: eight different child-oriented conversations (\textit{not} including the original one), overlaid and cropped to the duration of the video. Thus, the \textit{no speech} audio clips lacked any linguistic information to upcoming turn boundaries---the only cue to turn-taking was the opening and closing of the puppets' mouths. Finally, because low-pass filtering removes significant acoustic energy, the \textit{prosody only} clips were much quieter than the other three conditions. Our last step was to downscale the intensity of videos from the three other conditions to match the volume of the \textit{prosody only} clips. We referred to the conditions as ``normal'', ``robot'', ``mermaid'', and ``birthday party'' speech when interacting with participants (Figure 8).

\textit{Video recordings}. We created puppet video recordings to match the manipulated 20-second audio clips. The puppets were minimally expressive; the experimenter could only control the opening and closing of their mouths; their head, eyes, arms, and body stayed still. Puppets were positioned looking forward to eliminate shared gaze as a cue to turn structure \citep{thorgrimssonUndRev}. We took care to match the puppets' mouth movements to the syllable onsets as closely as possible and avoided any mouth movement before the onset of a turn. We then hand-aligned the manipulated audio clips to the puppet video recordings.

We used three pairs of puppets used for the \textit{normal} condition---`red', `blue' and `yellow'---and one pair of puppets for each limited cue condition: ``robots'', ``merpeople'', and ``party-goers'' (Figure 8). We randomly assigned half of the conversation topics (`birthday cake', `pets', and `breakfast') to the \textit{normal} condition, and half to the limited cue conditions (`riding bikes', `rainy days', and `the library'). We then created three versions of the experiment, so that each of the six puppet pairs was associated with three different conversation topics across the different versions of the experiment. We ensured that the position of the talkers (left and right) was counterbalanced in each version by flipping the video and audio channels as needed.\footnote{See the videos here: FIX add link.}

The duration of turn transitions and the number of speaker changes across videos was variable because the conversations were recorded semi-spontaneously. We measured turn transitions from the audio recording of the \textit{normal}, \textit{words only}, and \textit{prosody only} conditions. There was no audio from the original conversation in the \textit{no speech} condition videos, so we measured turn transitions from the video recording, using ELAN video editing software \citep{ELAN}. There were 79 turn transitions for analysis, after excluding transitions longer than 550 msec and shorter than 0 msec. The remaining turn transitions were distributed evenly across transition types (questions N$=$47 and non-questions N$=$32) and conditions, keeping in mind that there were three \textit{normal} videos and only one limited cue video for each experiment version: \textit{normal} (N$=$36), \textit{words only} (N$=$13), \textit{prosody only} (N$=$12), and {no speech} (N$=$18). Question transitions (mean$=$358, median$=$405) were longer than non-question transitions (mean$=$296, median$=$288) on average, but gap duration was overall comparable across conditions: \textit{normal} (mean$=$333, median$=$337), \textit{words only} (mean$=$345, median$=$383), \textit{prosody only} (mean$=$320, median$=$336), and \textit{no words} (mean$=$324, median$=$328). The longer gaps for question transitions could give them an advantage because our anticipatory measure includes shifts initiated during the gap between turns (Figure 2).

\subsection{Procedure}
We used the same set up as in the first study. Each participant watched six puppet videos in random order, with five 15--30 second filler videos placed in-between (e.g., running puppies, moving balls, flying bugs). Three of the puppet videos had \textit{normal} audio while the other three had \textit{words only}, \textit{prosody only}, and \textit{no speech} audio. This study required no special instructions so the experimenter immediately began each session with calibration (same as before) and then stimulus presentation. The entire experiment took less than five minutes.

\textit{Data preparation, coding, and random baseline analysis}. We coded each turn transition for its linguistic condition (\textit{normal}, \textit{words only}, \textit{prosody only}, and \textit{no speech}) and transition type (question/non-question)\footnote{We coded \textit{wh-}questions as ``non-questions'' for the \textit{prosody only} videos. Prototypical \textit{wh-}questions are not marked with a final rising nuclear prosodic contour, but polar questions are \citep{hedberg2010}.}, identified anticipatory gaze switches to the upcoming speaker, and estimated a random baseline of anticipatory switches using the methods from Study 1.

\section{Results and discussion}
\label{sec:results2}

Participants' pattern of gaze indicated that they performed basic turn tracking across all ages and in all conditions. Participants looked at the screen most of the time during video playback (82\% and 86\% average for children and adults, respectively). Children and adults primarily kept their eyes on the person who was currently speaking: they gazed at the current speaker between 44\% and 69\% of the time, looking back at the addressee between 11\% and 14\% of the time (Table 2). They tracked the current speaker in every condition---even one-year-olds looked more at the current speaker than at anything else in the three limited cue conditions (40\% for \textit{words only}, 43\% for \textit{prosody only}, and 39\% for \textit{no speech}). There was a steady overall increase in looks to the current speaker with age and added linguistic information (Tables 2 and 3). Looks to the addressee also decreased with age, but the change was minimal.

\begin{table}[H]
\begin{center}
  \begin{tabular}{llcccc}
    \hline
    Age group & Speaker & Addressee & Other onscreen & Offscreen\\ 
    \hline
    1 & 0.44 & 0.14 & 0.23 & 0.19 \\ 
    2 & 0.50 & 0.13 & 0.24 & 0.14 \\ 
    3 & 0.47 & 0.12 & 0.25 & 0.16 \\ 
    4 & 0.48 & 0.11 & 0.29 & 0.12 \\ 
    5 & 0.54 & 0.11 & 0.20 & 0.14 \\ 
    6 & 0.60 & 0.12 & 0.18 & 0.10 \\
    Adult & 0.69 & 0.12 & 0.09 & 0.10 \\
%   Overall & 0.53 & 0.12 & 0.21 & 0.13 \\
    \hline
  \end{tabular}
\end{center}
  \caption{Average proportion of gaze to the current speaker and addressee during periods of talk across ages.}
\label{looktableexp2a}
\end{table}

\begin{table}[H]
\begin{center}
  \begin{tabular}{llcccc}
    \hline
    Condition & Speaker & Addressee & Other onscreen & Offscreen\\ 
    \hline
    Normal & 0.58 & 0.12 & 0.17 & 0.13 \\ 
    Words only & 0.54 & 0.11 & 0.24 & 0.10 \\ 
    Prosody only & 0.48 & 0.12 & 0.26 & 0.15 \\ 
    No speech & 0.44 & 0.13 & 0.26 & 0.18 \\
    \hline
  \end{tabular}
\end{center}
  \caption{Average proportion of gaze to the current speaker and addressee during periods of talk across conditions.}
\label{looktableexp2b}
\end{table}

% Average prop gaze to the current speaker across conditions for one-year-olds
%                           speaker addressee other-screen offscreen
%Normal   		  0.4737582 0.1366432    0.2003952 0.1892035
%Words only    0.3977261 0.1283417    0.3112005 0.1627317
%Prosody only 0.4290202 0.1611453    0.2558657 0.1539688
%No speech    0.3786925 0.1228615    0.2356423 0.2628037

We identified anticipatory gaze switches for all 79 usable turn transitions using the same criteria and algorithm as in Study 1. We expected to see the most anticipatory switches in the \textit{normal} condition and the least anticipatory switches in the \textit{no speech} condition because they contained the most and least linguistic information, respectively. We expected to replicate our finding that questions result in more anticipatory switches than non-questions, with the added hypothesis that the question effect is driven by lexicosyntactic cues. We also anticipated an overall increase in anticipatory switches with age. Finally, since the development of prosodic skills partially precedes the development of lexicosyntax, we expected to see more switches in the \textit{prosody only} condition compared to the \textit{words only} condition in the youngest age groups.

As in Study 1, we corrected each participant's gaze data by subtracting their random baseline from their actual switching for each turn transition. All comparisons below were made with the baseline-corrected values.

Participants at all ages made anticipatory gaze switches more often than would be expected by chance in almost all conditions, but sometimes by a small margin (Figure 9). For example, anticipatory switches in the \textit{no speech} videos were often only modestly above chance (with the exception of 5- and 6-year-olds). Further, 1- and 2-year-olds did not appear to reliably outperform chance conditions for the \textit{words only} videos.

\begin{figure}[ht]
\begin{center}
\includegraphics[width=0.99\textwidth]{figures/FIG-randvsreal-EN.png}
\end{center}
\caption{Proportion anticipatory gaze switches made for actual (solid) and randomly-shuffled (dashed) turn transitions in each condition, and across ages. Error bars represent the standard error of the mean.} 
\label{randvsrealEN}
\end{figure}

We observed that, overall, children and adults made the fewest anticipatory switches in the \textit{no speech} videos (9\%) and the most in the \textit{normal} videos (16\%), with the \textit{prosody only} and \textit{words only} videos falling in-between (14\% and 11\%, respectively; Figure 10). However, these condition-based differences were primarily driven by transition type effects (question vs. non-question).

\begin{figure}[H]
\begin{center}
\includegraphics[width=0.65\textwidth]{figures/FIG-conditions-EN.png}
\end{center}
\caption{Proportion anticipatory gaze switches made for turn transitions in each speech condition, and across ages. Error bars represent the standard error of the mean.} 
\label{conditionsEN}
\end{figure}

Participants made more anticipatory gaze switches following questions than non-questions, but only in the two lexical conditions (\textit{normal} and \textit{words only} speech). The advantage for questions was strongest in the \textit{normal} speech condition, where it emerged early in development. The question effect became larger with age in both lexical conditions. While children made small, sometimes non-linear, gains in their anticipatory switches for non-questions as they grew older, they made large gains in their anticipatory switches with age (especially between ages one and two; Figure 11). The increased question effect from age one to two indicates that, around this age, children begin to leverage lexicosyntactic cues in the speech signal to identify questions in ongoing conversation. In the lexical conditions, children had access to early cues to questionhood, such as \textit{wh}-words and subject-auxiliary inversion, to predict an upcoming speaker change. In the \textit{normal} condition, where the question effect was strongest, lexical and syntactic cues were reinforced by question prosody for polar questions.

\begin{figure}[ht]
\begin{center}
\includegraphics[width=0.99\textwidth]{figures/FIG-QvsNQ-EN.png}
\end{center}
\caption{Proportion anticipatory gaze switches made for question- (dashed) and non-question- (dotted) turn transitions in each condition, and across ages. Error bars represent the standard error of the mean.} 
\label{questionsEN}
\end{figure}

Notably, while children performed well with \textit{prosody only} speech from the start, they seemed to only reliably exceed chance in the \textit{words only} condition at ages 3;0 and older. We confirmed this pattern in two \textit{post-hoc} \textit{t}-tests comparing children's anticipatory switching rate against their estimated random switching rate before age 3;0. We found that children under 3;0 significantly exceeded chance for the \textit{prosody only} condition (\textit{t}(187.677)=-3.004, \textit{p}=.003), but not for the \textit{words only} condition (\textit{t}(229.828)=-0.8307, \textit{p}=.407), giving some support to the idea that there is an early advantage for prosody over lexicosyntax in children's turn prediction. All in all, children showed increases in anticipatory switching with age across all conditions. Adults reacted to the \textit{no speech} condition differently than the children did, tracking the interaction less closely than in the three linguistic conditions. They may have minimized switching in the \textit{no speech} condition because they realized that there was no conversation to follow.

We tested these findings in two linear mixed-effects models of the baseline-corrected anticipatory switches---one model each for children and adults. The child model included condition (\textit{normal}, \textit{words only}, \textit{prosody only}, and \textit{no speech}), transition type (question vs. non-question), age, and gap duration as fixed effects, with full interactions between condition, transition type, and age. The model also included random effects of item (turn transition) and participant, with random slopes of condition and transition type for participants. The adult model matched the child model, excluding age-related effects.

There was a significant effect of gap duration. Children had time to make more anticipatory switches in analysis windows with longer gaps (\textit{$\beta$}=0.294, \textit{SE}=0.091, \textit{t}=3.231). An effect of gap duration is expected since prior work using spontaneous anticipatory switching has found that most switches occur in the inter-turn gap (Keitel et al., 2013; Hirvenkari, 2013; Tice and Henetz, 2011).

Condition-related differences were driven by age and transition type (Figure 11). Consequently, the model of children's anticipatory switches revealed no significant main effects of condition, despite the large changes in the linguistic information available across conditions. One interpretation of this finding is that children do not need linguistic information to predict upcoming turns. For example, children still made some anticipatory switches in the \textit{no speech} condition, even though the only cue to turn taking was the alternating mouth movements of the conversing puppets. With sufficiently long inter-turn gaps, participants could simply switch their gaze to the responder when the prior talker finishes talking---a strategy that doesn't rely on linguistic information. But this interpretation is not compatible with our finding that children make more anticipatory looks after hearing questions, that they do so more often when lexical information is available, and that they do even more so as they age.

The model of children's data confirmed two three-way effects of age, condition, and transition type. Children made significantly more anticipatory switches with age in the \textit{normal} (\textit{$\beta$}=0.04, \textit{SE}=0.008, \textit{t}=5.168) and \textit{words only} conditions (\textit{$\beta$}=0.031, \textit{SE}=0.01, \textit{t}=3.170): they made a switch twice as often after hearing a question (16\%) than a non-question (8\%) in the \textit{words only} condition, and nearly four times as often in the \textit{normal} condition (19\% vs. 5\%). Since the question effect was limited to the lexical conditions, there was no main effect of transition type. There were also no further interactions of age and condition, or of transition type and condition. There was, however, a significant two-way interaction of transition type and age (\textit{$\beta$}=-0.031, \textit{SE}=0.01, \textit{t}=-3.18), in addition to the three-way interactions for the \textit{normal} and \textit{words only} conditions already mentioned.

Last but not least, the model confirmed a significant main effect of age in the children's data (\textit{$\beta$}=0.028, \textit{SE}=0.009, \textit{t}=3.018). Children made more anticipatory switches as they got older. Many of the increased anticipatory looks with age came from questions in the lexical conditions. Children also increased their anticipatory looking with age for non-questions and non-lexical conditions, but the gains were smaller and sometimes followed a non-linear trajectory (Figure 11). As in Study 1, early switches (planned before the end of the prior turn) did not show large increases with age, with adults making the fewest early switches in the three linguistic conditions (Figure 12).

\begin{figure}[ht]
\begin{center}
\includegraphics[width=0.99\textwidth]{figures/FIG-cumulative-EN.png}
\end{center}
\caption{Cumulative proportion of anticipatory gaze switches planned during the transition window, across conditions and ages. Age is indicated by hue (light to dark; darker = older). The solid line is the offset of the pre-gap turn, and the dashed line is the median onset of the post-gap turn in each condition. ``Early switches'' occur before 0.} 
\label{cumulativeEN}
\end{figure}

The advantage for questions over non-questions in these results reinforces our findings from Study 1. They suggest that even two-year-olds can reliably distinguish questions from other transition types in third-party conversation, especially when full linguistic information is available. Inspecting the data more closely, we found that 43\% of one-year-olds made more anticipatory switches after questions than non-questions in the \textit{normal} condition, and 24\% of them did so in the \textit{words only} condition. Meanwhile, 61\% of two-year-olds show the effect in the \textit{normal} condition, and 52\% of them do in the \textit{words only} condition, which is already close to the average for older children and adults (Table 4).

Our results suggest that the question effect emerges between ages one and two, but further testing is needed to establish whether this is the case. For example, the size and emergence of the question effect for different question types and question cues could differ (\textit{wh-} vs. \textit{yes-no}; early vs. late), but we do not have enough data to make those distinctions in the current analysis. Also, nearly 43\% of one-year-olds showed question effects in the \textit{normal} speech condition, which has the full range of linguistic cues children hear in everyday life. Future work with more fine-grained age samples and better controlled \textit{normal} speech might then find that children can identify questions in conversation even earlier.

\begin{table}[ht]
\begin{center}
  \begin{tabular}{ccc}
    \hline
    Age group & Normal speech & No speech \\ 
    \hline
    1 & 0.43 & 0.24 \\
    2 & 0.61 & 0.52 \\
    3 & 0.63 & 0.42 \\
    4 & 0.78 & 0.52 \\
    5 & 0.91 & 0.52 \\
    6 & 0.75 & 0.5 \\
    Adult & 0.67 & 0.57 \\
    \hline
  \end{tabular}
\end{center}
  \caption{Proportion of participants showing more anticipatory switches for questions than non-questions in the \textit{normal} and \textit{words only} conditions.}
\label{questioneffectexp2}
\end{table}

The model of adults' anticipatory gaze switches confirmed main effects of gap duration and linguistic condition. Longer gaps led to more anticipatory switches (\textit{$\beta$}=0.656, \textit{SE}=0.176, \textit{t}=3.726). Adults also made significantly more anticipatory switches in the \textit{normal} condition than in the \textit{no speech} baseline, indicating that linguistic access did affect their overall performance (\textit{$\beta$}=-0.211, \textit{SE}=0.085, \textit{t}=-2.486). There were no other main effects of linguistic condition. The model also showed a marginal two-way interaction of condition and transition type, with adults making more anticipatory gazes for questions in the \textit{normal} condition (\textit{$\beta$}=0.194, \textit{SE}=0.11, \textit{t}=1.773). Though the question effect was numerically present in both lexical conditions, it was only significant for adults in the \textit{normal} condition, in which they had both lexicosyntactic and prosodic cues to questionhood. This result suggests that prosodic cues still have a role to play in adult question recognition during third-party conversation. Since question effects were primarily limited to the \textit{normal} condition, there was no main effect of transition type in the adults' data.

The core aims of Study 2 were to gain better traction on the individual roles of prosody and lexicosyntax in children's turn predictions, and to expand our age range to capture more developmental change. We found that effects of linguistic processing and age primarily surface within the question effect: children and adults make more anticipatory switches after questions than non-questions, but do so more often when they have greater access to linguistic information, whether increased access came with age or with linguistic condition. Our results hint that lexical information is critical in identifying questions during conversation, especially for children. Even so, performance was consistently best in the \textit{normal} condition, in which participants had lexical and prosodic information.

By extending the age range of participants, we were able to see more of the developmental trajectory in children's predictions about turn structure. We saw that one-year-olds already made more anticipatory switches than would be expected by chance in two of the four conditions. As they got older, children demonstrated above-chance anticipatory switches in more conditions and made more anticipatory switches overall. Similar to Study 1, children between ages three and five behaved similarly to each other, and comparably to adults, showing small but steady gains in their anticipatory switches. Though the advantage for questions generally increased with age, the biggest developmental shift in our data occurred between one and two years, when the question effect began to reliably emerge in the two lexical conditions. It will be up to future work to disentangle which lexical (and prosodic) cues children begin to pick out in distinguishing questions from other speech acts, and when.

The results from Study 2 complement those of Study 1, but still leave some open questions. For example, we were surprised to find no main effect of linguistic condition. In fact, the only evidence that participants  \textit{used} linguistic information came from the question effects, which were limited to the two lexical conditions. Given prior work, it is surprising that taking away lexical and prosodic information didn't have a larger impact on participants' anticipatory looking. As in Study 1, most of the anticipatory switches were initiated during the inter-turn gap, indicating that participants were anticipating the start of the next response rather than the end of the ongoing response. In contrast, the button press responses in de Ruiter et al. (2006) cluster before the end of the pre-gap turn. A button-press method explicitly encourages participants to make an early response, but recent work has also found early predictions with spontaneous neural responses \citep{bogelsmagyariInPrep, gisladottirUndRev, magyariUndRev}. It could be that spontaneous anticipatory switches don't elicit early enough responses to see linguistically-driven differences in prediction. Questions make an immediate next response relevant---more so than declaratives---and may provide a natural motivation for encouraging earlier predictive responses. If so, it would explain why age and language-related differences are only visible within the question effect.

We found that the importance of lexicosyntactic and prosodic information for turn prediction differed, depending on age and transition type. While even one-year-olds made anticipatory switches at an above-chance rate in the \textit{normal} and \textit{prosody only} conditions, children didn't consistently surpass chance in the \textit{words only} condition until age three. This pattern of development suggests an early importance for prosodic information in children's turn anticipation. But the question effect, which emerges between ages one and two, appears to heavily rely on lexical information, and doesn't appear in the \textit{prosody only} or \textit{no speech} conditions. So although prosody may help children early on, it doesn't affect their looking behavior in distinguishing between questions and non-questions---a driving force in their later anticipatory looking patterns. However, for both adults and children, the question effect was strongest when both lexicosyntactic and prosodic information were present, suggesting that prosody still has an important role to play in turn prediction. In other words, the lexicosyntax alone was not equivalent to the full linguistic signal in participants' anticipatory looking (even for adults), as prior work might have predicted \citep{de-ruiter2006, magyari2012}.

We controlled linguistic information in Study 2 by using phonetically-manipulated speech \citep{de-ruiter2006}. While phonetic manipulation gave us complete control over the linguistic signal, it also created speech sounds that children don't usually hear in their natural environment. Many prior studies have used phonetically-altered speech with infants and young children \citep[cf.][]{jusczyk2000}, but almost none of them have done so in a conversational context. We found that the \textit{normal} speech condition showed the strongest anticipation effects over all, and we have attributed this to the fact that it gave participants full (lexicosyntactic and prosodic) linguistic information. However, it was also the only natural speech condition. Children could have had trouble processing the \textit{words only} and \textit{prosody only} conditions because they were unfamiliar, and not just because they had less linguistic information available. It would take a different type of stimulus to compare lexical and prosodic information without phonetic manipulation. For example, carefully scripted conversations could control for the presence or absence of prosodic and lexicosyntactic cues to turn transition with natural speech. Since the non-natural (limited cue) linguistic conditions did still elicit some differences in anticipation, children must have processed at least some linguistic information in the \textit{prosody only} and \textit{words only} conditions. Perhaps the advantage of the \textit{normal} condition was then not entirely due to its naturalness. We expect that future work with spontaneous conversational measures will still find an advantage for combined lexical and prosodic cues over lexical cues alone \citep[see also][]{duncan1972, ford1996, bogelsUndRev}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{General discussion}
\label{sec:gendisc}

Conversational turn-taking is a time-sensitive behavior that children begin to develop long before their first words. But, in early infancy, children have limited linguistic knowledge to help them to predict what speakers will say next when they will finish talking. As they acquire language, they also acquire the information needed to make accurate predictions about upcoming turn structure. But, until recently, we have had very little data on how children weave language into their already-existing turn-taking behaviors. Because early conversation plays a central role in children's language learning, we can not fully understand language acquisition without taking conversational skill into account. Conversation is the driving force behind what children say and hear, and participants have to know the rules of the game to play effectively---especially when multiple players are involved. Prediction allows ready speakers to get a word in at just the right moment. It also helps participants to comprehend  ongoing speech and catch onto potential misunderstandings via comparison to anticipated speech (e.g., `yes' or `no' in response to a polar question). Language is enormously helpful in making predictions of this type, and several studies have shown that adults rely on lexicosyntactic \citep{de-ruiter2006, magyari2012} and, to some extent, prosodic information \citep{ford1996, bogelsUndRev} in their turn-taking. In the current study, we have tried to tie together several loose ends concerning when and how children begin to use linguistic information in their predictions about upcoming turn structure.

In two studies of children's anticipatory gaze to upcoming speakers, we found evidence that turn prediction develops early in childhood, and integrates linguistic cues differently with age. Children under 3;0 showed an advantage for prosody over lexicosyntax, contrary to the typical finding for adults in prior work \citep{de-ruiter2006} and our own data. Children 2;0 and older showed a slight advantage for lexicosyntax in the form of a transition type effect: children made more anticipatory switches following questions than non-questions in the two lexical conditions. But, the question effect only significant in the \textit{normal} condition, in which children had both lexicosyntactic and prosodic information. Similarly, we found no evidence that, for children or adults, lexicosyntax alone was ``sufficient'' (or otherwise equal to full linguistic information) for spontaneous turn prediction \citep[pg. 531]{de-ruiter2006}.

Children begin to make predictions about upcoming turn structure early in development. The children in Study 1 were nearly as good as the adults when it came to tracking the current speaker and making anticipatory switches. In the \textit{normal} speech condition of Study 2, even one-year-olds robustly made more anticipatory switches than would be expected by chance, and some even showed the start of a question effect (Table 4). Young children are slow turn-takers in spontaneous conversation; children at 1;8--1;9 take an average of nearly 800 ms to respond to questions (compared to 200 ms for adult-adult conversation; \citealp{casillasUndRev, stivers2009}). So, despite the fact that children begin taking turns in infancy and spontaneously predict upcoming turn structure at age one, it takes them a long time to develop adult-like timing in conversation with others. This pattern of findings suggests that children's language production and response planning plays a key role in their response timing, slowing them down in real-time interaction.

In both studies we found a robust advantage for questions in participants' anticipatory looks to the upcoming speaker. Young children have a lot of experience in hearing and answering questions. A handful of studies estimate that questions make up approximately one third of the utterances children hear, before and after the onset of speech, and even into their preschool years \citep{fitneva2012, henning2005, shatz1979}\footnote{There is variation by individual and socioeconomic class \citep{hart1992}.}, many of which are ``test' questions early on---questions that the caregiver already has the answer to (e.g., ``What does a cat say?''). Caregivers use questions to get their young children's' attention and to ensure that information is in common ground, even if childrens' responses are non-verbal or infelicitous \citep{fitneva2012, snow1977}. Questions give children a way in to ongoing interaction. Because they initiate a two-part structure, questions give the addressee certainty about when to talk (next) and what to say (the answer). Children begin to make protoimperatives between 0;9 and 1;0, using gaze, gesture, and vocalization \citep{bates1975}. Children's use of protoimperatives at 0;9 could suggest that they attend to two-part interactive sequences (imperative--(non-)compliance) long before real question-answer sequences emerge. But, protodeclaratives emerge around the same time, and may even be a more frequent speech act type in children's production before 2;8 \cite{cameron-faulkner2014}. If so, there may be a disparity between how children attend different speech acts in comprehension and how they use them in speech production. The adult participants in our studies also show the question effect, suggesting that greater anticipation for question-answers persists through adulthood (see also Tice and Henetz, 2011).

It remains unclear from our results what it is about the questions that caused children to switch their gaze to the responder more often. Questions make it much more likely that a speaker switch will occur at the next possible point. They have both early and late syntactic cues to help distinguish them from other speech (e.g., subject-auxiliary inversion and a final rising tone). But other speech acts and request formats have these properties. Imperatives, compliments, and complaints all make a response from the addressee likely. Rhetorical and tag questions, on the other hand, can take a similar form to more prototypical question types, but don't make an answer immediately relevant. So though it is clear that children anticipate responses more often for questions than non-questions, their actual understanding of questionhood as a linguistically- or interactionally-marked speech act still needs further exploration.

One of our primary aims was to see how linguistic information affects children's predictions about upcoming turn structure. To test this we limited the speech signal in our conversations to eliminate lexicosyntactic or prosodic information, either by showing participants conversations in languages they didn't understand or by phonetically manipulating the audio. Prior work with adults found a consistent and critical role for lexicosyntax, leaving prosody in more of a secondary and supporting role \citep{de-ruiter2006, magyari2012}. Knowing that children know more about prosody than lexicosyntax early on, we thought it might be possible that young children would show an advantage for prosody instead. Our results suggest that there are roles for both prosody and lexicosyntax, but that children may put one or the other information source to use at different ages.

We found an early advantage for prosodic information, compared to lexicosyntactic information, in one- and two-year-olds. While children's anticipatory switching didn't exceed chance for lexicosyntactic information alone until age three, it was significantly greater than chance for prosodic information alone by age one. The early advantage for prosody was not driven by the advantage for questions: while \textit{some} one-year-olds in our study already made more switches after questions, most did not, and prosody alone did not elicit a large question effect, even with the older children or adults. Having learned very little lexicosyntax, children's prosodic knowledge was probably limited to basic prosodic contours and cues to prosodic boundaries, including lengthening and pauses \textit{pannekamp2006}. These cues could have been enough to enable anticipatory looking for young children. But prosodic signals are often continuous, highly variable, and distributed throughout the turn. They are also hitched to the lexicosyntactic form of the utterance, making it difficult to tease apart prosody from syntax in higher-level predictive processing.

Lexicosyntactic cues, on the other hand, are more often categorical cues and, in the case of questions in English, mark the speech act early in the utterance, and clearly, with ``do'', \textit{wh}-words, and subject-auxiliary inversion. Whereas prosodic cues to questionhood are only salient on \textit{yes/no} questions in English, lexicosyntactic markers were available on every instance of \textit{wh}- and \textit{yes/no} questions in our stimuli. Clear lexicosyntactic markers may have been the reason why we saw the greatest switches for questions in conditions where lexicosyntactic information was available. However, in both studies we have presented, the question effect was not limited by lexicosyntactic information alone. Older children and adults in Study 1 still showed a question effect in languages that they did not understand. In Study 2, the question effects were only significant in the speech with lexicosyntactic \textit{and} prosodic information, again as a function of age. The results suggest that children get better at identifying questions in the ongoing conversation as they get older, with major gains between ages one and two, but that prosody still plays an important role in helping them to do so.

These results are only partially consistent with prior work. De Ruiter and colleagues (2006) found that adults' accuracy in identifying upcoming turn-ends was not significantly affected when intonation was taken away. In their study of children and adults, Keitel and colleagues (2013) found the same for adults' anticipatory gaze switches, and only finding that intonation affected children's switches at 3;0 (with switches by younger children never exceeding chance). Ou results suggest that children's anticipatory switches emerge much earlier, and rely on prosody before they do lexicosyntax. Further, our adult control participants patterned similarly to the children: they showed question effects for languages they didn't speak in Study 1 and strongest question effects for lexicosyntactic information when it was combined with prosody in Study 2. Thus, neither the child nor the adult data support the idea that lexicosyntactic information is equivalent to full linguistic information in making predictions about upcoming turn structure \citep{de-ruiter2006}, at least in participants' spontaneous looking behavior.

The anticipatory switches participants make in our studies were spontaneous, and were made without a requirement for urgency or timeliness, as the responses are for a button-press measure \citep{de-ruiter2006, magyari2012}. Most of our participants initiated their anticipatory switches in the inter-turn gap, consistent with prior observer gaze measures (Keitel et al., 2013; Hirvenkari, 2013; Tice and Henetz, \citeyear{TiceHenetz11}). The timing of participants' anticipatory switches suggests that, while they watch a conversation, they move their eyes in anticipation of the upcoming turn (not the end of the ongoing turn). Furthermore, the fact that participants made most of their anticipatory switches following questions, compared to non-questions, suggests that observers aren't just targeting potential turn-ends with their predictions---they are targeting speaker switches. Although button press measures may be better able to tell us what participants \textit{can} do in predicting turn ends, they may also emphasize lexicosyntactic cues to a greater extent than we experience in normal conversation. If participants are primarily monitoring for speaker switch instead of turn ends in general, we need to more carefully consider how signals to impending speaker transition (verbal and non-verbal) are weighed against more generic signals to utterance structure in our cognitive-linguistic model of how people manage to take turns on time. Current models of adult turn-taking underline the role of linguistic prediction, but more recent work, including our own, may point some future work toward more shallow cue monitoring (e.g., for \textit{wh-}words) and inhibition \citep{bogelsUndRev}.

The studies we present here are not without their limitations. In Study 2, we tried to improve our method for controlling linguistic information by phonetically manipulating improvised dialogic speech. Although this allowed us to more closely follow prior work \cite{de-ruiter2006, keitel2013}, it also limits our interpretation of children's behavior. Low-pass filtered and robotic-like speech are not typical speech signals children hear, and so the limited cue conditions in Study 2 may have been a somewhat unfair comparison to the normal speech condition, which not only had full linguistic cues, but also was more typical for what children hear in their everyday life. An better alternative to explore for the future would be scripted speech, or speech carefully selected from a large corpus for specific linguistic properties. Relatedly, since we used improvised conversation recordings, we did not control how often the speakers switched back and forth, or how often a switch occurred after a pause in speech. It is possible that, to some extent, children could have formed expectations about the turn-structure on the basis of the frequent switching alone \citep[see, e.g., ][]{thorgrimssonUndRev}. If this were the only thing leading children's anticipatory gaze switches, we would have predicted more uniformity in looking behavior across the conditions and no major effects of linguistic signal, age, or transition type. But, another advantage of using scripted conversation would be the ability to control switch frequency to eliminate this issue. Finally, we carefully designed a scheme for identifying anticipatory turn switches in the data, building off of Keitel and colleagues' (2013) analyses. In both our study and theirs, anticipatory switches that happen very early in the turn are not counted. Instead, they would actually count as ``random'' switches in the baseline analyses. These same early switches are counted as non-random in other studies of children's gaze to conversational stimuli \citep{bakker2011, hofsten2009}. Since most studies, even those using a wider switching window \citep[e.g., ][]{hirvenkari2013} find that switches primarily happen close to or immediately after the inter-turn gap begins, the approach used here and in Keitel et al. (2013) probably captures most of the switching behavior. However, if early looks do indeed increase as children get older and have more access to early syntactic cues, it would be helpful to develop a second measure of very early anticipatory looks.

\bibliographystyle{elsarticle-harv}
\bibliography{anticip}

\end{document}

### 2014-09-10 analysis/E1/0-find.transitions.R

@@ -1,0 +1,193 @@
+# Returns a data frame of transition info (whether or not switches were made
+# and, if so, when they started) for all subjects and transitions
+grab.transition.info <- function(obs, fixation.window) {
+    # Preset the total number of rows needed
+    subs <- unique(obs$Subject)
+    gaps <- sort(as.integer(unique(obs$GapNum)))
+    N <- length(subs) * length(gaps)
+    # Make the empty data frame
+    proportion.data <- data.frame(Subject=rep(NA, N), Gap=rep(NA, N),
+        Switch=rep(NA, N), Start.Time=rep(NA, N))
+    # Iterate through data frame rows, filling in transition information for
+    # each subject and gap
+    df.row = 1
+    # Grab a subject's data
+    for (i in 1:length(subs)) {
+        sub.data <- subset(obs, Subject == subs[i])
+        # Grab data for a particular gap
+        for (j in 1:length(gaps)) {
+            # Find the transition information (is there a switch? where?)
+            gap.transition <- find.transition(subset(sub.data,
+                GapNum == gaps[j]), fixation.window)
+            # Add the result to the data frame
+            proportion.data[df.row,] <- c(as.character(subs[i]),
+                gaps[j], gap.transition[1], gap.transition[2])
+            # Move onto the next row in the ouput data frame
+            df.row = df.row + 1
+        }
+    }
+    return(proportion.data)
+}
+
+
+# Parent function, linking the results of fixation.check and
+# find.transition.start together
+find.transition <- function(obs, fixation.window) {
+    if(nrow(obs) < 1) {
+        return(list(NA,NA))
+    } 
+    # Dummy minstart
+    minstart <- "NA"
+    # Search for a 100 msec fixation on the origin speaker during
+    # the ANCHOR period
+    anchor <- fixation.check(subset(obs, Anchor.Window == "ANCHOR"), "ANCHOR",
+        fixation.window, minstart)
+    minstart <- anchor[2]
+    # Search for a 100 msec fixation on the response speaker during
+    # the TRANSITION period
+    transition <- fixation.check(subset(obs, Transition.Window == "TRANSITION"),
+        "TRANSITION", fixation.window, minstart)
+    # Return the results of the search
+    if(anchor[1] == "Y" & transition[1] == "Y") {
+    	# If you found fixations on the ANCHOR and TRANSITION, find the starting
+        # point of the transition
+        return(list(1, find.transition.start(obs, transition[2])))
+    }
[…]

### 2014-09-10 analysis/E1/0-helper.R

@@ -1,0 +1,89 @@
+# Reads in the raw tracker data files and converts some basic raw values for us.
+read.data <- function(path,file.name) {
+    # E.g., A-ADU-P01.txt (for condition A, ADULT, and participant 01)
+    name.parts <- unlist(strsplit(file.name, "[-_.]"))
+    data <- read.csv(paste(path,file.name,sep=""))
+    # Within the tracker data for this participant...
+    data <- within(data, {
+        L.POR.X..px. <- to.n(L.POR.X..px.)
+        R.POR.X..px. <- to.n(R.POR.X..px.)
+        L.POR.Y..px. <- to.n(L.POR.Y..px.)
+        R.POR.Y..px. <- to.n(R.POR.Y..px.)
+        # Average x and y positions
+        x.pos <- rowMeans(data[, c("L.POR.X..px.", "R.POR.X..px.")])
+        y.pos <- rowMeans(data[, c("L.POR.Y..px.", "R.POR.Y..px.")])
+        # Object.Hits are when the eyes are looking in an area of interest
+        L.looks <- L.Object.Hit == "LEFT" | R.Object.Hit == "LEFT"
+        R.looks <- L.Object.Hit == "RIGHT" | R.Object.Hit == "RIGHT"
+        # Events are what the eyes are doing (e.g., fixating, saccading, etc.)
+        Events <- L.Event.Info != "Blink" & L.Event.Info != "-" &
+            R.Event.Info != "Blink" & R.Event.Info != "-"
+        Saccades <- L.Event.Info == "Saccade" | R.Event.Info == "Saccade"
+        Fixations <- L.Event.Info == "Fixation" | R.Event.Info == "Fixation"
+        # Add a version/agegroup/subject column based on the file name
+        Condition <- rep(name.parts[1], nrow(data))
+        AgeGroup <- rep(name.parts[2], nrow(data))
+        Subject <- rep(paste(name.parts[1], name.parts[3], name.parts[2],
+            sep = "-"), nrow(data))
+        # Make the time into seconds
+        TimeSec <- round((Time - Time[1])/(1000000), 3)
+        MS.Increment <- c(0, diff(TimeSec))
+    })
+    return (data)
+}
+
+
+# Adds annotations to those measurements that are within the analysis windows
+# |--------A----|-|                     # Speaker A's turn
+#                      |-|----B-----|   # Speaker B's turn
+# Anchor:      |--|
+# Transition:   |---------|
+# Gap:            |----|
+#
[…]

### 2014-09-10 analysis/E1/0-make.random.windows.R

@@ -1,0 +1,178 @@
+# Takes in a vector of gap durations associated with each stimulus and returns a 
+# data frame of with the same set of gap durations, only shuffled at random 
+# within the time boundaries of the associated stimulus.
+find.rand.windows <- function(min, max, vecdur) {
+    N <- length(vecdur)
+    # Maximum start time for a gap is .433 msec before the end of the stimulus
+    give.win <- 0.433
+    # Create a new, empty data frame
+    new.windows <- data.frame(Onset=rep(NA, N), Offset=rep(NA, N),
+        Duration=rep(NA, N))
+    # Iterate through new.windows, filling in information for each gap duration
+    df.row = 1
+    for (i in 1:length(vecdur)) {
+        success=0
+        while(success < 1) {
+            dur = vecdur[i]
+            # Generate a random start value within the video stimulus
+            try.start <- runif(1, min + give.win, max - give.win - dur)
+            # Fill in the first value
+            if (is.numeric(new.windows[1,1]) == FALSE) {
+            	# Add the randomized gap to the data frame
+                new.windows[df.row,] <- c(try.start, try.start + dur, dur)
+                # Move onto the next row
+                df.row = df.row + 1
+                success = 1
+            }
+            # For the second row and beyond...
+            else {		
+                try.win <- c(try.start, try.start + dur)
+                # Check to see if the new proposed random gap overlaps with
+                # one we have already
+                used.win <- subset(new.windows[,1:2], Onset != "NA")
+                temp.win <- rbind(try.win, used.win)
+                # Orders the accepted and proposed random gaps by onset
+                # (crucial for the check.overlap function)
+                temp.win <- temp.win[order(temp.win$Onset),]
+                # If it doesn't, add it to the data frame
+                if (check.overlap(temp.win) == 0) {
+                    new.windows[df.row,] <- c(try.start, try.start + dur, dur)
+                    new.windows <- new.windows[order(new.windows$Onset),]
+                    df.row = df.row + 1
+                    success = 1
[…]

### 2014-09-10 analysis/E1/1-read.dataset.R

@@ -1,0 +1,26 @@
+rm(list = ls())
+source("0-useful.R")
+source("0-helper.R")
+
+raw.data.path <- "tracker_data/"
+info.path <- "info/"
+processed.data.path <- "processed_data/"
+
+files <- dir(raw.data.path, pattern="*.txt")
+all.data <- data.frame()
+
+# Loop through the files, grabbing the data we want
+for (file.name in files) {
+    # Print an update
+    print(file.name)
+    data <- read.data(raw.data.path, file.name)
+    # Take only the columns we want for further analysis
+    data <- subset(data, select = c("Time", "x.pos", "y.pos", "L.looks",
+        "R.looks", "Events", "Condition", "Subject", "AgeGroup", "TimeSec",
+        "MS.Increment", "Stimulus", "Saccades", "Fixations"))
+    # Bind it together
+    all.data <- rbind(all.data, data)
+}
+
+# Save the imported data
+write.csv(all.data, paste(processed.data.path,"imported.data.csv",sep=""))

### 2014-09-10 analysis/E1/2-prep.analysis.R

@@ -1,0 +1,174 @@
+rm(list = ls())
+source("0-useful.R")
+source("0-helper.R")
+source("0-find.transitions.R")
+
+raw.data.path <- "tracker_data/"
+info.path <- "info/"
+processed.data.path <- "processed_data/"
+
+# Read in supplemental data and prep
+vid.info <- read.csv(paste(info.path,"VideoSegmentInfo.csv",sep=""))
+turn.info <- read.csv(paste(info.path,"TurnInfo.csv",sep=""))
+subject.info <- read.csv(paste(info.path,"SubjectInfo.csv",sep=""))
+gap.info <- read.csv(paste(info.path,"GapInfo.csv",sep=""))
+
+# Round the time values for turns and gaps
+turn.info = within(turn.info, {
+  Onset = round(Onset, 3)
+  Offset = round(Offset, 3)
+})
+gap.info = within(gap.info, {
+  Onset = round(Onset, 3)
+  Offset = round(Offset, 3)
+})
+
+
+################################################################################
+# First we'll merge the gaze data with the information about the stimuli
+
+# Read in main datafile
+all.data <- read.csv(paste(processed.data.path,"imported.data.csv",sep=""))
+
+# Merge in condition sites, using onsets
+all.data$Onset <- rep(NA, nrow(all.data))
+for (cond in unique(all.data$Condition)) {
+    # For each TimeSecSeg measurement, find the onset of the segment
+    # (from vid.info) within which it lies
+    onsets <- vid.info$Onset[vid.info$Cond == cond]
+    # REMINDER: example use of findInterval:
+    # findInterval(c(2,5,7,8), c(4,5,6,8,9,10)) => [1] 0 2 3 4
+    # In this case it returns a list as long as all.data$TimeSecSeg, with
+    # the index of the video onset that the each time measurement falls within
+    # (e.g., t = 5, vid[2] and t=7, vid[3] in the above example)
+    onset.indices <- findInterval(all.data$TimeSec[all.data$Condition == cond],
+        onsets)
+    # Maps these values onto the subset of times it was derived from
+    all.data$Onset[all.data$Condition == cond] <- onsets[onset.indices]
+}
+# Add in vid.info data and re-order the main data frame
+all.data <- merge(all.data, vid.info, by = c("Condition", "Onset"))
+all.data <- all.data[order(all.data$Subject, all.data$TimeSec),]
+
+# Merge in turn sites, using onsets
+all.data$Onset <- rep(NA, nrow(all.data))
[…]

### 2014-09-10 analysis/E1/2-random.runs.R

@@ -1,0 +1,106 @@
+rm(list = ls())
+source("0-useful.R")
+source("0-helper.R")
+source("0-find.transitions.R")
+
+raw.data.path <- "tracker_data/"
+info.path <- "info/"
+processed.data.path <- "processed_data/"
+
+# Read in supplemental data and prep
+vid.info <- read.csv(paste(info.path,"VideoSegmentInfo.csv",sep=""))
+subject.info <- read.csv(paste(info.path,"SubjectInfo.csv",sep=""))
+
+all.data <- read.csv(paste(processed.data.path,"r.all.data.csv",sep=""))
+
+# Separate adult and child data for finding anticipatory switches
+# because adults have an assumed faster gaze planning time of 200 msec
+# compared to our assumed 333 msec for children
+all.data.A <- subset(all.data, AgeGroup == "ADU")
+all.data.C <- subset(all.data, AgeGroup == "CHI")
+
+# Anticipatory switch identification variables
+fixation.window = 0.1
+# could change utt settings to "reaction" rather than "anticipation" values
+utt1.overlap.A = 0.2
+utt2.overlap.A = 0.2
+utt1.overlap.C = 0.333
+utt2.overlap.C = 0.333
+
+# Create 100 versions of the analyses with the random gap info files we made
+# (using the random.runs.R script), and then run this loop...
+for  (i in 1:100) {
+    # Print an update
+    print(paste("Working on random version ",i," now...", sep=""))
+    # Start with a clean version of the data
+    r.all.data.A <- all.data.A
+    r.all.data.C <- all.data.C
+
+    # Read in one version of the random gaps
+    r.gap.info <- read.csv(paste(info.path,"GapInfoRandom",i,".csv", sep=""))
+    r.gap.info = within(r.gap.info, {
+        Onset = round(Onset, 3)
+        Offset = round(Offset, 3)
+    })
+    # Extract time windows for switch analysis and prepare new looking columns
+    print(paste("--Getting windows--"))
+    # For adults
+    r.target.windows.proport.A <- get.windows.categorical(r.all.data.A,
+        r.gap.info, utt1.overlap.A, utt2.overlap.A, fixation.window)
+    r.target.windows.proport.A$Looks.Origin <- as.integer(
+        r.target.windows.proport.A$Origin ==
+        r.target.windows.proport.A$Look.Dir)
+    r.target.windows.proport.A$Looks.Destination <- ifelse(((
[…]

### 2014-09-10 analysis/E1/3-plots.stats.analysis.R

@@ -1,0 +1,775 @@
+rm(list = ls())
+library(ggplot2)
+library(lme4)
+library(gridExtra)
+library(scales)  
+library(Hmisc)
+source("0-useful.R")
+
+raw.data.path <- "tracker_data/"
+info.path <- "info/"
+processed.data.path <- "processed_data/"
+plot.path <- "plots/"
+
+# Color and line palettes
+stoplightPalette <- c("dodgerblue3", "firebrick2")
+stoplightPalette2 <- c("gray26", "dodgerblue3", "firebrick2")
+stoplightPalette3 <- c("dodgerblue3", "firebrick2")
+lines <- c("solid", "dashed")
+lines2 <- c("longdash", "dotted")
+
+
+# Default setting for all.data in case you don't need the table values and
+# aren't actually loading all.data
+all.data <- 0
+
+# Takes forever to load, only uncomment if you want the table values for looks 
+# to current speaker (at the top)
+#all.data <- read.csv(paste(processed.data.path, "all.data.before.analysis.csv",
+#    sep=""))
+#all.data <- subset(all.data, Subject != "D-P01-CHI")
+
+
+# Read in gaze switch proportions
+# Actual data
+switch.final <- read.csv(paste(processed.data.path, "switch.final.csv", sep=""))
+sw.f.drop <- c("X", "X.1", "notes", "Ttype")
+switch.final <- switch.final[,!(names(switch.final) %in% sw.f.drop)]
+# Random baseline
+switch.final.r <- read.csv(paste(processed.data.path,
+    "random.runs.anticipatory.csv", sep=""))
+sw.f.r.drop <- c("X", "X.1", "X.2")
+switch.final.r <- switch.final.r[,!(names(switch.final.r) %in% sw.f.r.drop)]
+# Combine them
+switch.final <- switch.final[, order(names(switch.final))]
+switch.final.r <- switch.final.r[, order(names(switch.final.r))]
+switch.all <- rbind(switch.final, switch.final.r)
+# Exclude K-E bilingual child
+switch.all <- subset(switch.all, Subject != "D-P01-CHI")
+
+# Read in cumulative gaze switch data
[…]

### 2014-09-10 analysis/E2/0-find.transitions.R

@@ -1,0 +1,193 @@
+# Returns a data frame of transition info (whether or not switches were made
+# and, if so, when they started) for all subjects and transitions
+grab.transition.info <- function(obs, fixation.window) {
+    # Preset the total number of rows needed
+    subs <- unique(obs$Subject)
+    gaps <- sort(as.integer(unique(obs$GapNum)))
+    N <- length(subs) * length(gaps)
+    # Make the empty data frame
+    proportion.data <- data.frame(Subject=rep(NA, N), Gap=rep(NA, N),
+        Switch=rep(NA, N), Start.Time=rep(NA, N))
+    # Iterate through data frame rows, filling in transition information for
+    # each subject and gap
+    df.row = 1
+    # Grab a subject's data
+    for (i in 1:length(subs)) {
+        sub.data <- subset(obs, Subject == subs[i])
+        # Grab data for a particular gap
+        for (j in 1:length(gaps)) {
+            # Find the transition information (is there a switch? where?)
+            gap.transition <- find.transition(subset(sub.data,
+                GapNum == gaps[j]), fixation.window)
+            # Add the result to the data frame
+            proportion.data[df.row,] <- c(as.character(subs[i]),
+                gaps[j], gap.transition[1], gap.transition[2])
+            # Move onto the next row in the ouput data frame
+            df.row = df.row + 1
+        }
+    }
+    return(proportion.data)
+}
+
+
+# Parent function, linking the results of fixation.check and
+# find.transition.start together
+find.transition <- function(obs, fixation.window) {
+    if(nrow(obs) < 1) {
+        return(list(NA,NA))
+    } 
+    # Dummy minstart
+    minstart <- "NA"
+    # Search for a 100 msec fixation on the origin speaker during
+    # the ANCHOR period
+    anchor <- fixation.check(subset(obs, Anchor.Window == "ANCHOR"), "ANCHOR",
+        fixation.window, minstart)
+    minstart <- anchor[2]
+    # Search for a 100 msec fixation on the response speaker during
+    # the TRANSITION period
+    transition <- fixation.check(subset(obs, Transition.Window == "TRANSITION"),
+        "TRANSITION", fixation.window, minstart)
+    # Return the results of the search
+    if(anchor[1] == "Y" & transition[1] == "Y") {
+    	# If you found fixations on the ANCHOR and TRANSITION, find the starting
+        # point of the transition
+        return(list(1, find.transition.start(obs, transition[2])))
+    }
+    else {
[…]

### 2014-09-10 analysis/E2/0-helper.R

@@ -1,0 +1,102 @@
+# Reads in the raw tracker data files and converts some basic raw values for us.
+read.data <- function(path,file.name) {
+    # E.g., A-ADU-P01.txt (for condition A, ADULT, and participant 01)
+    name.parts <- unlist(strsplit(file.name, "[-_.]"))
+    data <- read.csv(paste(path,file.name,sep=""))
+    # Within the tracker data for this participant...
+    data <- within(data, {
+        L.POR.X..px. <- to.n(L.POR.X..px.)
+        R.POR.X..px. <- to.n(R.POR.X..px.)
+        L.POR.Y..px. <- to.n(L.POR.Y..px.)
+        R.POR.Y..px. <- to.n(R.POR.Y..px.)
+        # Average x and y positions
+        x.pos <- rowMeans(data[, c("L.POR.X..px.", "R.POR.X..px.")])
+        y.pos <- rowMeans(data[, c("L.POR.Y..px.", "R.POR.Y..px.")])
+        # Object.Hits are when the eyes are looking in an area of interest
+        F.looks <- L.Object.Hit == "FEMALE" | R.Object.Hit == "FEMALE"
+        M.looks <- L.Object.Hit == "MALE" | R.Object.Hit == "MALE"
+        # Events are what the eyes are doing (e.g., fixating, saccading, etc.)
+        Events <- L.Event.Info != "Blink" & L.Event.Info != "-" &
+            R.Event.Info != "Blink" & R.Event.Info != "-"
+        Saccades <- L.Event.Info == "Saccade" | R.Event.Info == "Saccade"
+        Fixations <- L.Event.Info == "Fixation" | R.Event.Info == "Fixation"
+        # Add a version/agegroup/subject column based on the file name
+        Version <- rep(name.parts[2], nrow(data))
+        AgeGroup <- rep(name.parts[1], nrow(data))
+        Subject <- rep(paste(name.parts[1], name.parts[2], name.parts[3],
+            sep = "-"), nrow(data))
+        # Make the time into seconds
+        TimeSec <- round((Time - Time[1])/(1000000), 3)
+        MS.Increment <- c(0, diff(TimeSec))
+        ### Add a column of times for each video segment
+        TimeSecSeg <- rep(NA, nrow(data))
+        videos = unique(data$Stimulus)
+        for (video in videos[videos != "-"]) {
+          curr.time = 0
+          sub.indx <- which(data$Stimulus == video)
+          MS.Increment[sub.indx[1]] = 0
[…]

### 2014-09-10 analysis/E2/0-make.random.windows.R

@@ -1,0 +1,175 @@
+# Takes in a vector of gap durations associated with each stimulus and returns a 
+# data frame of with the same set of gap durations, only shuffled at random 
+# within the time boundaries of the associated stimulus.
+find.rand.windows <- function(min, max, vecdur) {
+    N <- length(vecdur)
+    # Maximum start time for a gap is .433 msec before the end of the stimulus
+    give.win <- 0.433
+    # Create a new, empty data frame
+    new.windows <- data.frame(Onset=rep(NA, N), Offset=rep(NA, N),
+        Duration=rep(NA, N))
+    # Iterate through new.windows, filling in information for each gap duration
+    df.row = 1
+    for (i in 1:length(vecdur)) {
+        success=0
+        while(success < 1) {
+            dur = vecdur[i]
+            # Generate a random start value within the video stimulus
+            try.start <- runif(1, min + give.win, max - give.win - dur)
+            # Fill in the first value
+            if (is.numeric(new.windows[1,1]) == FALSE) {
+            	# Add the randomized gap to the data frame
+                new.windows[df.row,] <- c(try.start, try.start + dur, dur)
+                # Move onto the next row
+                df.row = df.row + 1
+                success = 1
+            }
+            # For the second row and beyond...
+            else {		
+                try.win <- c(try.start, try.start + dur)
+                # Check to see if the new proposed random gap overlaps with
+                # one we have already
+                used.win <- subset(new.windows[,1:2], Onset != "NA")
+                temp.win <- rbind(try.win, used.win)
+                # Orders the accepted and proposed random gaps by onset
+                # (crucial for the check.overlap function)
+                temp.win <- temp.win[order(temp.win$Onset),]
+                # If it doesn't, add it to the data frame
+                if (check.overlap(temp.win) == 0) {
+                    new.windows[df.row,] <- c(try.start, try.start + dur, dur)
+                    new.windows <- new.windows[order(new.windows$Onset),]
[…]

### 2014-09-10 analysis/E2/0-useful.R

@@ -1,0 +1,61 @@
+library(grid)
+library(ggplot2)
+library(bootstrap)
+library(lme4)
+library(stringr)
+library(plotrix)
+library(reshape)
+library(plyr)
+library(car)
+
+# Add some style elements for ggplot2
+plot.style <- theme(panel.grid.major = element_blank(),
+    panel.grid.minor = element_blank(),
+    axis.line = element_line(colour="black",size=.5),
+    axis.ticks = element_line(size=.5),
+    axis.title.x = element_text(vjust=-.5),
+    axis.title.y = element_text(angle=90,vjust=0.25),
+    panel.margin = unit(1.5,"lines"))
+
+theme_set(theme_bw())
+
+# Standard error of the mean
+sem <- function (x) {
+    sd(x) / sqrt(length(x))
+}
+
+na.mean <- function(x) {mean(x,na.rm=T)}
+
+to.n <- function(x) {
+    as.numeric(as.character(x))
+}
+
+# Number of unique subs
+n.unique <- function (x) {
+    length(unique(x))
+}
+
+# For bootstrapping 95% confidence intervals
+theta <- function(x,xdata) {mean(xdata[x])}
+ci.low <- function(x) {
+    mean(x) - quantile(bootstrap(1:length(x),1000,theta,x)$thetastar,.025)}
+ci.high <- function(x) {
+    quantile(bootstrap(1:length(x),1000,theta,x)$thetastar,.975) - mean(x)}
+ci95 <- function(x) {sem(x)*1.96}
+
+# For basic plots, add linear models with correlations
+lm.txt <- function (p1,p2,x=7.5,yoff=.05,lt=2,c="black",data=data) {
+    l <- lm(p2 ~ p1)
+    regLine(l,lty=lt,col=c)
+    cl <- coef(l)
+    text(x,cl[1] + cl[2] * x + yoff,
+        paste("r = ",sprintf("%2.2f",sqrt(summary(l)$r.squared)),
+        getstars(anova(l)$"Pr(>F)"[1]),sep=""), xpd="n")
+}
+
+getstars <- function(x) {
+    if (x > .1) {return("")}
+    if (x < .001) {return("***")}
+    if (x < .01) {return("**")}
+    if (x < .05) {return("*")}
+}

### 2014-09-10 analysis/E2/1-read.dataset.R

@@ -1,0 +1,29 @@
+rm(list = ls())
+source("0-useful.R")
+source("0-helper.R")
+
+raw.data.path <- "tracker_data/"
+info.path <- "info/"
+processed.data.path <- "processed_data/"
+
+files <- dir(raw.data.path, pattern="*.txt")
+all.data <- data.frame()
+
+# Loop through the files, grabbing the data we want
+for (file.name in files) {
+    # Print an update
+	print(file.name)
+	data <- read.data(raw.data.path, file.name)
+    # Take only the columns we want for further analysis
+	data <- subset(data, select = c("Time", "x.pos", "y.pos", "F.looks",
+	    "M.looks", "Events", "Version", "Subject", "TimeSec", "TimeSecSeg",
+	    "MS.Increment", "Stimulus", "Saccades", "Fixations"))
+    # Bind it together
+	all.data <- rbind(all.data, data)
+}
+
+# Rename one column
+colnames(all.data)[12] = "Segment"
+
+# Save the imported data
+write.csv(all.data, paste(processed.data.path,"imported.data.csv",sep=""))

### 2014-09-10 analysis/E2/2-prep.analysis.R

@@ -1,0 +1,198 @@
+rm(list = ls())
+source("0-useful.R")
+source("0-helper.R")
+source("0-find.transitions.R")
+
+raw.data.path <- "tracker_data/"
+info.path <- "info/"
+processed.data.path <- "processed_data/"
+
+# Read in supplemental data and prep
+vid.info <- read.csv(paste(info.path,"VideoSegmentInfo.csv",sep=""))
+turn.info <- read.csv(paste(info.path,"TurnInfo.csv",sep=""))
+subject.info <- read.csv(paste(info.path,"SubjectInfo.csv",sep=""))
+gap.info <- read.csv(paste(info.path,"GapInfo.csv",sep=""))
+
+# Round the time values for turns and gaps
+turn.info = within(turn.info, {
+  Onset = round(Onset, 3)
+  Offset = round(Offset, 3)
+})
+gap.info = within(gap.info, {
+  Onset = round(Onset, 3)
+  Offset = round(Offset, 3)
+})
+
+
+################################################################################
+# First we'll merge the gaze data with the information about the stimuli
+
+# Read in main datafile
+all.data <- read.csv(paste(processed.data.path,"imported.data.csv",sep=""))
+
+# Add in the video.info data, keep only target videos, and re-order the data
+all.data <- merge(all.data, vid.info, by = "Segment")
+all.data <- subset(all.data, StimulusGroup == "target")
+all.data <- all.data[order(all.data$Subject, all.data$Segment,
+    all.data$TimeSecSeg),]
+
+# Merge in turn sites, using onsets
+all.data$Onset <- rep(NA, nrow(all.data))
+for (vid in unique(all.data$Segment)) {
+    # For each TimeSecSeg measurement, find the onset of the segment
+    # (from turn.info) within which it lies
+    onsets <- turn.info$Onset[turn.info$Segment == vid]
+    # REMINDER: example use of findInterval:
+    # findInterval(c(2,5,7,8), c(4,5,6,8,9,10)) => [1] 0 2 3 4
+    # In this case it returns a list as long as all.data$TimeSecSeg, with
+    # the index of the video onset that the each time measurement falls within
+    # (e.g., t = 5, vid[2] and t=7, vid[3] in the above example)
+    vid.onset.indices <- findInterval(
+        all.data$TimeSecSeg[all.data$Segment == vid], onsets)
+    # Maps these values onto the subset of times it was derived from
[…]

### 2014-09-10 analysis/E2/2-random.runs.R

@@ -1,0 +1,177 @@
+rm(list = ls())
+source("0-useful.R")
+source("0-helper.R")
+source("0-find.transitions.R")
+
+raw.data.path <- "tracker_data/"
+info.path <- "info/"
+processed.data.path <- "processed_data/"
+
+# Read in supplemental data and prep
+vid.info <- read.csv(paste(info.path,"VideoSegmentInfo.csv",sep=""))
+subject.info <- read.csv(paste(info.path,"SubjectInfo.csv",sep=""))
+
+all.data <- read.csv(paste(processed.data.path,"r.all.data.csv",sep=""))
+
+# Create 100 versions of the analyses with the random gap info files we made
+# (using the random.runs.R script), then run this loop...
+for (i in 1:100) {
+    # Print an update
+    print(paste("Working on random version ",i," now...", sep=""))
+    # Start with a clean version of the data
+    r.all.data <- all.data
+
+    # Read in one version of the random gaps
+    r.gap.info <- read.csv(paste(info.path,"GapInfoRandom",i,".csv", sep=""))
+    r.gap.info = within(r.gap.info, {
+        Onset = round(Onset, 3)
+        Offset = round(Offset, 3)
+    })
+    
+    # Merge in gap sites, using new onset values
+    print(paste("--Merging info files--"))
+    r.all.data$Onset <- rep(NA, nrow(r.all.data))
+    for (vid in unique(r.all.data$Segment)) {
+        # For each TimeSecSeg measurement, find the onset of the utterance
+        # (from gap.info) within which it lies
+        onsets <- r.gap.info$Onset[r.gap.info$Segment == vid]
+        # REMINDER: example use of findInterval:
+        # findInterval(c(2,5,7,8), c(4,5,6,8,9,10)) => [1] 0 2 3 4
+        # In this case it returns a list as long as all.data$TimeSecSeg, with
+        # the index of the video onset that the each time measurement falls within
+        # (e.g., t = 5, vid[2] and t=7, vid[3] in the above example)
+        vid.gap.indices <- findInterval(
+            r.all.data$TimeSecSeg[r.all.data$Segment == vid], onsets)
+        # Maps these values onto the subset of times it was derived from
+        r.all.data$Onset[r.all.data$Segment == vid] <- onsets[vid.gap.indices]
+    }
+    # Add in gap.info data and re-order the main data frame
[…]

### 2014-09-10 analysis/E2/3-plots.stats.analysis.R

@@ -1,0 +1,829 @@
+rm(list = ls())
+library(ggplot2)
+library(lme4)
+library(gridExtra)
+library(scales)  
+library(Hmisc)
+source("0-useful.R")
+
+raw.data.path <- "tracker_data/"
+info.path <- "info/"
+processed.data.path <- "processed_data/"
+plot.path <- "plots/"
+
+# Color and line palettes
+stoplightPalette <- c("dodgerblue3", "forestgreen", "green3", "firebrick2")
+stoplightPalette2 <- c("gray26", "firebrick2", "forestgreen",
+    "dodgerblue3", "green3")
+stoplightPalette3 <- c("dodgerblue3", "forestgreen", "green3", "firebrick2")
+lines <- c("solid", "dashed")
+lines2 <- c("longdash", "dotted")
+
+
+# Default setting for all.data in case you don't need the table values and
+# aren't actually loading all.data
+all.data <- 0
+
+# Takes forever to load, only uncomment if you want the table values for looks 
+# to current speaker (at the top)
+#all.data <- read.csv(paste(processed.data.path, "all.data.before.analysis.csv",
+#    sep=""))
+
+# Read in gaze switch proportions
+# Actual data
+switch.final <- read.csv(paste(processed.data.path, "switch.final.csv", sep=""))
+sw.f.drop <- c("X", "X.1", "notes")
+switch.final <- switch.final[,!(names(switch.final) %in% sw.f.drop)]
+# Random baseline
+switch.final.r <- read.csv(paste(processed.data.path,
+    "random.runs.anticipatory.csv", sep=""))
+sw.f.r.drop <- c("X", "X.1", "X.2", "X.3")
+switch.final.r <- switch.final.r[,!(names(switch.final.r) %in% sw.f.r.drop)]
+# Combine them
+switch.final <- switch.final[, order(names(switch.final))]
+switch.final.r <- switch.final.r[, order(names(switch.final.r))]
+switch.all.coarse <- rbind(switch.final, switch.final.r)
+
+# Read in cumulative gaze switch data
+cumulative.data <- read.csv(paste(processed.data.path,
+    "cumulative.conditions.age.c.csv", sep=""))
+
+
[…]

### 2014-09-10 paper/CasillasFrank-Anticip.tex
mcf changes

@@
 
-\begin{figure}[ht]
+\begin{figure}[t]
 \begin{center}
@@
 \caption{Example frame from a conversation video used in Experiment 1.} 
-\label{speakers}
+\label{fig:speakers}
 \end{figure}
@@
 
-\textit{Video-recordings}.
-We recorded pairs of talkers while they conversed in a sound-attenuated booth.
+\textit{Video recordings}.
+We recorded pairs of talkers while they conversed in a sound-attenuated booth (see sample frame in Figure \ref{fig:speakers}).
 Each talker was a native speaker of the language being recorded, and each talker pair was male-female.
@@
 The talkers were volunteers, all members of the local post-graduate community, and were acquainted with their recording partner ahead of time.
+
 Each recording session began with a 20-minute warm-up period of spontaneous conversation during which the pair talked for five minutes on four topics (favorite foods, entertainment, hometown layout, and pets).
@@
 On average, the inter-turn gaps for English videos (mean$=$273 msec, median$=$291 msec) were slightly longer than for non-English videos (mean$=$226 msec, median$=$205 msec).
-The longer gaps in the English videos could give them a slight advantage: our definition of an ``anticipatory gaze shift'' includes shifts that are initiated during the gap between turns (Figure 2), so participants had slightly more time to make anticipatory shifts in the English videos.
+The longer gaps in the English videos could give them a slight advantage: our definition of an ``anticipatory gaze shift'' includes shifts that are initiated during the gap between turns (Figure \ref{fig:criterion}), so participants had slightly more time to make anticipatory shifts in the English videos.
 
@@
 We placed stereo speakers on the table, to the left and right of the display.
+
 Before the experiment started, we warned adult participants that they would see videos in several languages and that, though they weren't expected to understand the content of non-English videos, we \textit{would} ask them to answer general, non-language-based questions about the conversations.
@@
 If the child began to look bored during video playback, the experimenter would talk during the filler videos, either commenting on the previous conversation (``That was a neat language!'') or giving the language name for the next conversation (``This next one is called Hebrew.
-Let's see what it's like.'') The experimenter's comments reinforced the video-watching as a joint task.
+Let's see what it's like.'').
+The experimenter's comments reinforced the video-watching as a joint task.
 
@@
 \subsubsection{Data preparation and coding}
+\label{sec:algorithm}
+
 To determine whether participants predicted upcoming turn transitions, we needed to define a set of criteria for what counted as an anticipatory gaze shift.
@@
 The criteria were the same for adults and children, except children were assumed to have a 333 msec window for planning eye movements, while adults were only given 200 msec.}
-\label{speakers}
+\label{fig:criterion}
 \end{figure}
@@
 Using these criteria, we created a three-step algorithm for detecting anticipatory gaze switches.
-Consider, for example, when a child observes a turn transition between speaker A and speaker B (Figure 2).
+Consider, for example, when a child observes a turn transition between speaker A and speaker B (Figure \ref{fig:criterion}).
 First, the child would need to fixate on A for at least 100 msec at the end of A's turn.
@@
 
-\textit{Random baseline analysis}.
-To ensure that participants' looking behavior was turn-related and not simply the result of random switching, we corrected for the probability of making an anticipatory switch at random.
+\paragraph{Random baseline analysis} To ensure that participants' looking behavior was turn-related and not simply the result of random switching, we corrected for the probability of making an anticipatory switch at random.
 We estimated the baseline chance of making an anticipatory switch at each turn transition for each person.
-We ran all participants' eye-tracking data through switch identification with 100 randomly-shuffled versions of the original turn transition windows (Figure 3).
-For each of the 100 randomly-shuffled versions, we recorded and then averaged whether or not each participant made an anticipatory switch at each ``turn transition''.
-We then subtracted the random baseline value from each participant's actual switch result (1 or 0) for each turn transition to arrive at the final, baseline-corrected values.
+We ran all participants' eye-tracking data through switch identification with 100 randomly-shuffled versions of the original turn transition windows (Figure \ref{fig:shuffling}).
+For each of the 100 randomly-shuffled versions, we recorded and then averaged whether or not each participant made an anticipatory switch at each ``turn transition.'' We then subtracted the random baseline value from each participant's actual switch result (1 or 0) for each turn transition to arrive at the final, baseline-corrected values.
 Subtracting the random baseline this way can give us an estimate of how often participants switched above chance across ages, conditions, and transition types.
-If participants make the same number of anticipatory switches, regardless of turn transition location, there would be no evidence that their switching behavior was linked to the conversation stimulus.
+If participants made the same number of anticipatory switches, regardless of turn transition location, there would be no evidence that their switching behavior was linked to the conversation stimulus.
 
-\begin{figure}[ht]
+\begin{figure}[t]
 \begin{center}
@@
 \caption{Example of shuffling for five turn transition windows. The windows were +/-433 msec around the inter-turn gap for children, and from +/-300 msec for adults.} 
-\label{shuffling}
+\label{fig:shuffling}
 \end{figure}
@@
 
+\subsubsection{Descriptive analysis}
+
 Participants looked at the screen most of the time during video playback (81\% and 91\% on average for children and adults, respectively).
-They primarily kept their eyes on the person who was currently speaking in both English and non-English videos: they gazed at the current speaker between 38\% and 63\% of the time, looking back at the addressee between 15\% and 20\% of the time (Table 1).
+They primarily kept their eyes on the person who was currently speaking in both English and non-English videos: they gazed at the current speaker between 38\% and 63\% of the time, looking back at the addressee between 15\% and 20\% of the time (Table \ref{tab:e1_look}).
 Even three-year-olds looked more at the current speaker than anything else, whether the videos were in a language they could understand or not.
@@
 
-\begin{table}[H]
+\begin{table}[t]
 \begin{center}
@@
   \caption{Average proportion of gaze to the current speaker and addressee during periods of talk.}
-\label{looktableexp1}
+\label{tab:e1_look}
 \end{table}
 
-We identified anticipatory gaze switches for all 43 usable turn transitions, based on the criteria and algorithm outlined in Section 2.1.4.
+We identified anticipatory gaze switches for all 43 usable turn transitions, based on the criteria and algorithm outlined in Section \ref{sec:algorithm}.
 Because participants had greater access to lexical and syntactic information in English, we expected to see greater anticipation in the English videos than in the non-English videos.
@@
 Error bars represent the standard error of the mean.}
-\label{randvsrealFL}
+\label{fig:randvsrealFL}
 \end{figure}
 
-Participants at all ages made anticipatory gaze switches more often than would be expected by chance, whether they could understand the language in the video or not (Figure 4).
+Participants at all ages made anticipatory gaze switches more often than would be expected by chance, whether they could understand the language in the video or not (Figure \ref{fig:randvsrealFL}).
 Anticipatory switch behavior in the non-English videos could not have been based on lexicosyntactic information, so both adults and children must have relied on other sources of information to make their predictions (e.g., prosody and non-verbal signals).
@@
 But the tendency to make more anticipatory switches for English than non-English videos disappeared with age.
-While anticipatory switches increased with age for non-English videos, they decreased with age for English videos, resulting in nearly equal looking behavior for the two language conditions for adults  (Figure 5).
-Otherwise, age-based differences in predictive gaze shifts were minor.
+While anticipatory switches increased with age for non-English videos, they decreased with age for English videos, resulting in nearly equal looking behavior for the two language conditions for adults.
+Otherwise, age-based differences in predictive gaze shifts were minor (Figure \ref{fig:conditionsFL}).
 
-\begin{figure}[ht]
+\begin{figure}[t]
 \begin{center}
@@
 \end{center}
-\caption{Proportion anticipatory gaze switches made for turn transitions in each language condition, and across ages. Error bars represent the standard error of the mean.} 
-\label{conditionsFL}
+\caption{Baseline-corrected proportion anticipatory gaze switches made for turn transitions in each language condition, and across ages. Error bars represent the standard error of the mean.} 
+\label{fig:conditionsFL}
 \end{figure}
 
-\begin{figure}[H]
+\begin{figure}[t]
 \begin{center}
@@
 Error bars represent the standard error of the mean.}
-\label{questionsFL}
+\label{fig:questionsFL}
 \end{figure}
@@
 But the advantage for questions was not universal---it depended on language condition and age.
-The advantage for questions didn't emerge until age five in the non-English videos (Figure 6).
+The advantage for questions didn't emerge until age five in the non-English videos (Figure \ref{fig:questionsFL}).
 
-We confirmed these findings in two linear mixed-effects models of baseline-corrected anticipatory switches \citep{lme4, R}.
+\subsubsection{Statistical analysis}
+
+We quantified these findings using two linear mixed-effects models of baseline-corrected anticipatory switches \citep{lme4, R}.
 We built one model each for children and adults.
@@
 non-question), age, and duration of the inter-turn gap as predictors, with full interactions between condition, transition type, and age.
-We included the duration of the inter-turn gap as a predictor since longer gaps also provide more opportunities to make anticipatory switches (Figure 2).
+We included the duration of the inter-turn gap as a predictor since longer gaps also provide more opportunities to make anticipatory switches (Figure \ref{fig:criterion}).
 We additionally included random effects of item (turn transition) and participant, with random slopes of condition, transition type, and their interaction for participants \citep{barr2013}.
@@
 We had expected that, as children got older, they would also show more anticipatory switches.
-But this is contradicted by decreased looks with age for English videos, which even extended to the adult data (Figure 5).
+But this is contradicted by decreased looks with age for English videos, which even extended to the adult data.
 One possible explanation is that older children and adults found the videos in English easy to comprehend, and so tracked the conversation less closely.
 Compared to three-year-olds, adults and older children spent more time looking away from the current speaker and addressee in the English conversations.
-In contrast, compared to three-year-olds, they spent more time looking at the current speaker and less time looking away in the non-English conversations (Table 1).
+In contrast, compared to three-year-olds, they spent more time looking at the current speaker and less time looking away in the non-English conversations (Table \ref{tab:e1_look}).
 
@@
 Within the current analysis windows there were some ``early switches'' (i.e., those launched before the end of the prior turn).
-But participants made approximately equal numbers of early switches, regardless of age and condition: 11--15\% (Figure 7).
+But participants made approximately equal numbers of early switches, regardless of age and condition: 11--15\% (Figure \ref{fig:cumulativeFL}).
 The lack of increased early switching with age might suggest that changes in timing are either larger than the analysis window, or are not at play.
 
-\begin{figure}[ht]
+\begin{figure}[t]
 \begin{center}
@@
 ``Early switches'' occur before 0.}
-\label{cumulativeFL}
+\label{fig:cumulativeFL}
 \end{figure}
@@
 Presumably, question effects were larger in the English videos than the non-English videos because there were more cues to questionhood in English (i.e., non-verbal, prosodic, \textit{and} lexicosyntactic).
-Children younger than five did not distinguish questions from non-questions in the non-English videos (Figure 6).
+Children younger than five did not distinguish questions from non-questions in the non-English videos (Figure \ref{fig:questionsFL}).
 The children's model confirmed a three-way interaction of age, linguistic condition, and transition type (\textit{$\beta$}=-0.091, \textit{SE}=0.042, \textit{t}=-2.183).
@@
 
-A few results from Study 1 were unexpected.
+\subsubsection{Summary}
+
+Several results from Experiment 1 were unexpected.
 First, we found no significant main effect of age in the children's data, and we saw very little difference between adults and children (besides the age-related interactions).
 It appears that, especially in English, children were as good at tracking the current speaker and anticipating upcoming turns as adults were.
-To find out when children first begin making spontaneous anticipatory looks to responders, we will have to look at younger children.
+To find out when children first begin making spontaneous anticipatory looks to responders, younger participants are required.
+Experiment 2 thus includes a sample of one- and two-year-olds.
+
 Second, children, but not adults, showed an advantage for lexicosyntactic information in their anticipatory switches.
 An advantage for lexicosyntactic information is consistent with prior work on adult turn taking \citep{de-ruiter2006}, so it's curious that the adults did not show a difference in anticipatory switching between videos they did and did not understand.
-As mentioned, they may have effectively used the non-verbal cues in the videos to track the interactions fluently and, consequently, found the interactions so easy to follow that they did not track the turn-taking closely.
-To identify which \textit{linguistic} cues adults and children rely on, we have to control the presence of non-verbal cues in the videos.
+As mentioned above, it is likely that they used non-verbal cues to track the interactions and consequently found the videos so easy to follow that they did not track the turn-taking closely.
+To identify which \textit{linguistic} cues adults and children rely on, it is necessary to control the presence of non-verbal cues in the stimulus---Experiment 2 implements this control.
 
-Finally, we had initially predicted that children would show early advantages for prosodic cues over syntactic ones because children begin developing language-specific knowledge about prosody long before lexicosyntax.
-We weren't able to directly test this prediction in Study 1 because we used a limited range of ages (three- to five-year-olds) and because we only controlled for lexicosyntactic information.
-Children had prosodic cues in both English and non-English videos, but the prosodic signal was non-native in the latter case.
-So, although using non-English videos was a natural way to control for lexical information, it also rendered us unable to assess prosody's role in participants' predictions.
-Relatedly, the advantage for questions was greater in the English videos than the non-English videos, lending the possibility that lexicosyntactic cues are essential for children in identifying questions during conversation.
-To test this idea, we would need to directly compare lexicosyntactic and prosodic cues in the participants' native language.
+Finally, we had predicted that children would show early advantages for prosodic cues over syntactic ones because children begin developing language-specific knowledge about prosody long before lexicosyntax.
+We weren't able to directly test this prediction in Experiment 1 because we used a limited range of ages (three- to five-year-olds) and because we only controlled for lexicosyntactic information.
+Children had prosodic cues in both English and non-English videos, although the prosodic signal was non-native in the latter case.
+So, although using non-English videos was a natural way to control for lexical information, it did not allow for assessment of the role of prosody.
+Relatedly, the advantage for questions was greater in the English videos than the non-English videos, suggesting that lexicosyntactic cues are important for children in identifying questions during conversation.
+Again, to test this idea, we would need to directly compare lexicosyntactic and prosodic cues in the participants' native language.
+In sum, the current study lays the analytic groundwork for a method that allows for greater experimental control, which we introduce in Experiment 2.
 
@@
 We improved our design by using native-language stimuli, controlling for lexical \textit{and} prosodic information, eliminating non-verbal cues, and testing children from a wider age range.
-All of the videos in Study 2 were in the participants' native language (American English).
+All of the videos in Experiment 2 were in the participants' native language (American English).
 To tease apart the role of lexical and prosodic information, we phonetically manipulated the speech signal for pitch, rhythm, and lexical access.
 By testing one- to six-year-olds we hoped to find the developmental onset of turn-predictive gaze.
-We also hoped to see whether the roles of prosody and lexicosyntax change as children get older.
+We also hoped to measure changes in the relative roles of prosody and lexicosyntax across development.
 
-Non-verbal cues in Study 1 (e.g., gaze and gesture) could have helped participants make predictions about upcoming turn structure  \citep{rossano2009, stivers2010}.
-Since our focus is on linguistic cues, we eliminated all gaze and gestural signals in Study 2 by replacing the videos of human actors with videos of puppets.
+Non-verbal cues in Experiment 1 (e.g., gaze and gesture) could have helped participants make predictions about upcoming turn structure  \citep{rossano2009, stivers2010}.
+Since our focus is on linguistic cues, we eliminated all gaze and gestural signals in Experiment 2 by replacing the videos of human actors with videos of puppets.
 Puppets are less realistic and expressive than human actors but, since we wanted to eliminate gestural and gaze cues, the puppets created a natural context for having somewhat motionless talkers in the videos.
@@
 None of the adult participants reported fluency in a second language.
-We ran Study 2 at a children's museum because it gave us access to a more diverse age range of children from more variable socioeconomic backgrounds.
+We ran Experiment 2 at a local children's museum because it gave us access to children with a more diverse range of ages.
 
-\begin{figure}[ht]
+\begin{figure}[t]
 \begin{center}
@@
 \caption{The six puppet pairs (and associated audio conditions). Each pair was linked to three distinct conversations from the same condition across the three experiment versions.}
-\label{puppets}
+\label{fig:puppets}
 \end{figure}
@@
 \subsubsection{Materials}
-We created six short videos of improvised, child-friendly conversation.
+We created six short videos of improvised, child-friendly conversation (Figure \ref{fig:puppets}).
 To eliminate non-verbal cues to turn transition and to control the types of linguistic information available in the stimuli we (1) recorded improvised conversations, (2) phonetically manipulated those recordings to limit the availability of prosodic and lexical information, and (3) recorded a new set of videos that featured puppets as talkers, using the manipulated audio as the puppets' speech.
@@
 The \textit{normal} clips were the unmanipulated, original audio clips.
+
 The \textit{words only} clips were manipulated to have robot-like speech: we flattened the intonation contours to each talker's average pitch (F0) and we reset the duration of every nucleus and coda to each talker's average nucleus and coda duration.\footnote{We excluded hyper-lengthened words like [w\textipa{aU:}] `woooow!'.
@@
 Thus, the \textit{words only} versions of the audio clips had no pitch or durational cues to upcoming turn boundaries, but did have intact lexicosyntactic cues.
+
 We created the \textit{prosody only} clips by low-pass filtering the original recording at 500 Hz with a 50 Hz Hanning window (following de Ruiter et al., 2006).
-This creates a ``muffled speech'' sound because low-pass filtering removes most of the phonetic information used to distinguish between phonemes.
-Thus, the \textit{prosody only} versions of the audio clips lacked lexical information, but retained their intonational and rhythmic cues to upcoming turn boundaries.
+This manipulation creates a ``muffled speech'' sound because low-pass filtering removes most of the phonetic information used to distinguish between phonemes.
+The \textit{prosody only} versions of the audio clips lacked lexical information, but retained their intonational and rhythmic cues to upcoming turn boundaries.
+
 The \textit{no speech} condition served as a non-linguistic baseline.
@@
 Thus, the \textit{no speech} audio clips lacked any linguistic information to upcoming turn boundaries---the only cue to turn-taking was the opening and closing of the puppets' mouths.
+
 Finally, because low-pass filtering removes significant acoustic energy, the \textit{prosody only} clips were much quieter than the other three conditions.
 Our last step was to downscale the intensity of videos from the three other conditions to match the volume of the \textit{prosody only} clips.
-We referred to the conditions as ``normal'', ``robot'', ``mermaid'', and ``birthday party'' speech when interacting with participants (Figure 8).
+We referred to the conditions as ``normal'', ``robot'', ``mermaid'', and ``birthday party'' speech when interacting with participants.
 
@@
 There was no audio from the original conversation in the \textit{no speech} condition videos, so we measured turn transitions from the video recording, using ELAN video editing software \citep{ELAN}.
+
 There were 79 turn transitions for analysis, after excluding transitions longer than 550 msec and shorter than 0 msec.
@@
 Question transitions (mean$=$358, median$=$405) were longer than non-question transitions (mean$=$296, median$=$288) on average, but gap duration was overall comparable across conditions: \textit{normal} (mean$=$333, median$=$337), \textit{words only} (mean$=$345, median$=$383), \textit{prosody only} (mean$=$320, median$=$336), and \textit{no words} (mean$=$324, median$=$328).
-The longer gaps for question transitions could give them an advantage because our anticipatory measure includes shifts initiated during the gap between turns (Figure 2).
+The longer gaps for question transitions could give them an advantage because our anticipatory measure includes shifts initiated during the gap between turns (Figure \ref{fig:algorithm}).
 
 \subsection{Procedure}
-We used the same set up as in the first study.
+We used the same experimental apparatus and procedure as in the first study.
 Each participant watched six puppet videos in random order, with five 15--30 second filler videos placed in-between (e.g., running puppies, moving balls, flying bugs).
@@
 We coded each turn transition for its linguistic condition (\textit{normal}, \textit{words only}, \textit{prosody only}, and \textit{no speech}) and transition type (question/non-question)\footnote{We coded \textit{wh-}questions as ``non-questions'' for the \textit{prosody only} videos.
-Prototypical \textit{wh-}questions are not marked with a final rising nuclear prosodic contour, but polar questions are \citep{hedberg2010}.}, identified anticipatory gaze switches to the upcoming speaker, and estimated a random baseline of anticipatory switches using the methods from Study 1.
+Prototypical \textit{wh-}questions are not marked with a final rising nuclear prosodic contour, but polar questions are \citep{hedberg2010}.}, identified anticipatory gaze switches to the upcoming speaker, and estimated a random baseline of anticipatory switches using the methods from Experiment 1.
 
-\section{Results and discussion}
+\subsection{Results and discussion}
 \label{sec:results2}
 
+\subsubsection{Descriptive analysis}
 Participants' pattern of gaze indicated that they performed basic turn tracking across all ages and in all conditions.
@@
 They tracked the current speaker in every condition---even one-year-olds looked more at the current speaker than at anything else in the three limited cue conditions (40\% for \textit{words only}, 43\% for \textit{prosody only}, and 39\% for \textit{no speech}).
-There was a steady overall increase in looks to the current speaker with age and added linguistic information (Tables 2 and 3).
+There was a steady overall increase in looks to the current speaker with age and added linguistic information (Tables \ref{tab:look_e2} and \ref{tab:look_e2b}).
 Looks to the addressee also decreased with age, but the change was minimal.
 
-\begin{table}[H]
+\begin{table}[t]
 \begin{center}
@@
   \caption{Average proportion of gaze to the current speaker and addressee during periods of talk across ages.}
-\label{looktableexp2a}
+\label{tab:look_e2}
 \end{table}
 
-\begin{table}[H]
+\begin{table}[t]
 \begin{center}
@@
   \caption{Average proportion of gaze to the current speaker and addressee during periods of talk across conditions.}
-\label{looktableexp2b}
+\label{tab:look_e2b}
 \end{table}
@@
 
-We identified anticipatory gaze switches for all 79 usable turn transitions using the same criteria and algorithm as in Study 1.
+We identified anticipatory gaze switches for all 79 usable turn transitions using the same criteria and algorithm as in Experiment 1.
 We expected to see the most anticipatory switches in the \textit{normal} condition and the least anticipatory switches in the \textit{no speech} condition because they contained the most and least linguistic information, respectively.
@@
 
-As in Study 1, we corrected each participant's gaze data by subtracting their random baseline from their actual switching for each turn transition.
+As in Experiment 1, we corrected each participant's gaze data by subtracting their random baseline from their actual switching for each turn transition.
 All comparisons below were made with the baseline-corrected values.
 
-Participants at all ages made anticipatory gaze switches more often than would be expected by chance in almost all conditions, but sometimes by a small margin (Figure 9).
+Participants at all ages made anticipatory gaze switches more often than would be expected by chance in almost all conditions, but sometimes by a small margin (Figure \ref{fig:randvsrealEN}).
 For example, anticipatory switches in the \textit{no speech} videos were often only modestly above chance (with the exception of 5- and 6-year-olds).
@@
 \end{center}
-\caption{Proportion anticipatory gaze switches made for actual (solid) and randomly-shuffled (dashed) turn transitions in each condition, and across ages.
+\caption{Baseline-corrected proportion anticipatory gaze switches made for actual (solid) and randomly-shuffled (dashed) turn transitions in each condition, and across ages.
 Error bars represent the standard error of the mean.}
-\label{randvsrealEN}
+\label{fig:randvsrealEN}
 \end{figure}
 
-We observed that, overall, children and adults made the fewest anticipatory switches in the \textit{no speech} videos (9\%) and the most in the \textit{normal} videos (16\%), with the \textit{prosody only} and \textit{words only} videos falling in-between (14\% and 11\%, respectively; Figure 10).
+We observed that, overall, children and adults made the fewest anticipatory switches in the \textit{no speech} videos (9\%) and the most in the \textit{normal} videos (16\%), with the \textit{prosody only} and \textit{words only} videos falling in-between (14\% and 11\%, respectively; Figure \ref{fig:conditionsEN}).
 However, these condition-based differences were primarily driven by transition type effects (question vs.
@@
 
-\begin{figure}[H]
+\begin{figure}[t]
 \begin{center}
@@
 \caption{Proportion anticipatory gaze switches made for turn transitions in each speech condition, and across ages. Error bars represent the standard error of the mean.} 
-\label{conditionsEN}
+\label{fig:conditionsEN}
 \end{figure}
@@
 The question effect became larger with age in both lexical conditions.
-While children made small, sometimes non-linear, gains in their anticipatory switches for non-questions as they grew older, they made large gains in their anticipatory switches with age (especially between ages one and two; Figure 11).
+While children made small, sometimes non-linear, gains in their anticipatory switches for non-questions as they grew older, they made large gains in their anticipatory switches with age (especially between ages one and two; Figure \ref{fig:questionsEN}).
 The increased question effect from age one to two indicates that, around this age, children begin to leverage lexicosyntactic cues in the speech signal to identify questions in ongoing conversation.
@@
 \end{center}
-\caption{Proportion anticipatory gaze switches made for question- (dashed) and non-question- (dotted) turn transitions in each condition, and across ages.
+\caption{Baseline-corrected proportion anticipatory gaze switches made for question- (dashed) and non-question- (dotted) turn transitions in each condition, and across ages.
 Error bars represent the standard error of the mean.}
-\label{questionsEN}
+\label{fig:questionsEN}
 \end{figure}
@@
 They may have minimized switching in the \textit{no speech} condition because they realized that there was no conversation to follow.
+
+\subsubsection{Statistical analysis}
 
@@
 
-Condition-related differences were driven by age and transition type (Figure 11).
+Condition-related differences were driven by age and transition type (Figure \ref{fig:questionsEN}).
 Consequently, the model of children's anticipatory switches revealed no significant main effects of condition, despite the large changes in the linguistic information available across conditions.
@@
 Many of the increased anticipatory looks with age came from questions in the lexical conditions.
-Children also increased their anticipatory looking with age for non-questions and non-lexical conditions, but the gains were smaller and sometimes followed a non-linear trajectory (Figure 11).
-As in Study 1, early switches (planned before the end of the prior turn) did not show large increases with age, with adults making the fewest early switches in the three linguistic conditions (Figure 12).
+Children also increased their anticipatory looking with age for non-questions and non-lexical conditions, but the gains were smaller and sometimes followed a non-linear trajectory (Figure \ref{fig:questionsEN}).
+As in Study 1, early switches (planned before the end of the prior turn) did not show large increases with age, with adults making the fewest early switches in the three linguistic conditions (Figure \ref{fig:cumulativeEN}).
 
@@
 ``Early switches'' occur before 0.}
-\label{cumulativeEN}
+\label{fig:cumulativeEN}
 \end{figure}
@@
 Inspecting the data more closely, we found that 43\% of one-year-olds made more anticipatory switches after questions than non-questions in the \textit{normal} condition, and 24\% of them did so in the \textit{words only} condition.
-Meanwhile, 61\% of two-year-olds show the effect in the \textit{normal} condition, and 52\% of them do in the \textit{words only} condition, which is already close to the average for older children and adults (Table 4).
+Meanwhile, 61\% of two-year-olds show the effect in the \textit{normal} condition, and 52\% of them do in the \textit{words only} condition, which is already close to the average for older children and adults (Table \ref{tab:questioneffectexp2}).
 
@@
 
-\begin{table}[ht]
+\begin{table}[t]
 \begin{center}
@@
   \caption{Proportion of participants showing more anticipatory switches for questions than non-questions in the \textit{normal} and \textit{words only} conditions.}
-\label{questioneffectexp2}
+\label{tab:questioneffectexp2}
 \end{table}
@@
 
+\subsubsection{Summary}
+
 The core aims of Study 2 were to gain better traction on the individual roles of prosody and lexicosyntax in children's turn predictions, and to expand our age range to capture more developmental change.
-We found that effects of linguistic processing and age primarily surface within the question effect: children and adults make more anticipatory switches after questions than non-questions, but do so more often when they have greater access to linguistic information, whether increased access came with age or with linguistic condition.
+We found that effects of linguistic processing and age primarily surface within the question effect: Children and adults make more anticipatory switches after questions than non-questions, but do so more often when they have greater access to linguistic information, whether increased access came with age or with linguistic condition.
 Our results hint that lexical information is critical in identifying questions during conversation, especially for children.
@@
 As they got older, children demonstrated above-chance anticipatory switches in more conditions and made more anticipatory switches overall.
-Similar to Study 1, children between ages three and five behaved similarly to each other, and comparably to adults, showing small but steady gains in their anticipatory switches.
+Similar to Experiment 1, children between ages three and five behaved similarly to each other, and comparably to adults, showing small but steady gains in their anticipatory switches.
 Though the advantage for questions generally increased with age, the biggest developmental shift in our data occurred between one and two years, when the question effect began to reliably emerge in the two lexical conditions.
@@
 
-The results from Study 2 complement those of Study 1, but still leave some open questions.
+The results from Experiment 2 complement those of Experiment 1, but still leave some open questions.
 For example, we were surprised to find no main effect of linguistic condition.
@@
 Given prior work, it is surprising that taking away lexical and prosodic information didn't have a larger impact on participants' anticipatory looking.
-As in Study 1, most of the anticipatory switches were initiated during the inter-turn gap, indicating that participants were anticipating the start of the next response rather than the end of the ongoing response.
-In contrast, the button press responses in de Ruiter et al.
+As in Experiment 1, most of the anticipatory switches were initiated during the inter-turn gap, indicating that participants were anticipating the start of the next response rather than the end of the ongoing response.\footnote{In contrast, the button press responses in de Ruiter et al.
 (2006) cluster before the end of the pre-gap turn.
 A button-press method explicitly encourages participants to make an early response, but recent work has also found early predictions with spontaneous neural responses \citep{bogelsmagyariInPrep, gisladottirUndRev, magyariUndRev}.
-It could be that spontaneous anticipatory switches don't elicit early enough responses to see linguistically-driven differences in prediction.
+It could be that spontaneous anticipatory switches don't elicit early enough responses to see linguistically-driven differences in prediction.}
+
+We found that the importance of lexicosyntactic and prosodic information for turn prediction differed, depending on age and transition type.
 Questions make an immediate next response relevant---more so than declaratives---and may provide a natural motivation for encouraging earlier predictive responses.
@@
 
-We found that the importance of lexicosyntactic and prosodic information for turn prediction differed, depending on age and transition type.
-While even one-year-olds made anticipatory switches at an above-chance rate in the \textit{normal} and \textit{prosody only} conditions, children didn't consistently surpass chance in the \textit{words only} condition until age three.
+In addition, while even one-year-olds made anticipatory switches at an above-chance rate in the \textit{normal} and \textit{prosody only} conditions, children didn't consistently surpass chance in the \textit{words only} condition until age three.
 This pattern of development suggests an early importance for prosodic information in children's turn anticipation.
@@
 
-We controlled linguistic information in Study 2 by using phonetically-manipulated speech \citep{de-ruiter2006}.
+We controlled linguistic information in Experiment 2 by using phonetically-manipulated speech \citep{de-ruiter2006}.
 While phonetic manipulation gave us complete control over the linguistic signal, it also created speech sounds that children don't usually hear in their natural environment.
@@
 Children could have had trouble processing the \textit{words only} and \textit{prosody only} conditions because they were unfamiliar, and not just because they had less linguistic information available.
+
 It would take a different type of stimulus to compare lexical and prosodic information without phonetic manipulation.
 For example, carefully scripted conversations could control for the presence or absence of prosodic and lexicosyntactic cues to turn transition with natural speech.
-Since the non-natural (limited cue) linguistic conditions did still elicit some differences in anticipation, children must have processed at least some linguistic information in the \textit{prosody only} and \textit{words only} conditions.
+But since the non-natural (limited cue) linguistic conditions did still elicit some differences in anticipation, children must have processed at least some linguistic information in the \textit{prosody only} and \textit{words only} conditions.
 Perhaps the advantage of the \textit{normal} condition was then not entirely due to its naturalness.
@@
 
-\section{General discussion}
+\section{General Discussion}
 \label{sec:gendisc}
@@
 Conversational turn-taking is a time-sensitive behavior that children begin to develop long before their first words.
-But, in early infancy, children have limited linguistic knowledge to help them to predict what speakers will say next when they will finish talking.
+But in early infancy, children have limited linguistic knowledge to help them to predict what speakers will say next when they will finish talking.
 As they acquire language, they also acquire the information needed to make accurate predictions about upcoming turn structure.
-But, until recently, we have had very little data on how children weave language into their already-existing turn-taking behaviors.
-Because early conversation plays a central role in children's language learning, we can not fully understand language acquisition without taking conversational skill into account.
+Until recently, we have had very little data on how children weave language into their already-existing turn-taking behaviors.
+
+% Because early conversation plays a central role in children's language learning, we can not fully understand language acquisition without taking conversational skill into account.
 Conversation is the driving force behind what children say and hear, and participants have to know the rules of the game to play effectively---especially when multiple players are involved.
@@
 
-In two studies of children's anticipatory gaze to upcoming speakers, we found evidence that turn prediction develops early in childhood, and integrates linguistic cues differently with age.
+In two experiments investigating children's anticipatory gaze to upcoming speakers, we found evidence that turn prediction develops early in childhood, and that linguistic cues are integrated differently at different ages.
 Children under 3;0 showed an advantage for prosody over lexicosyntax, contrary to the typical finding for adults in prior work \citep{de-ruiter2006} and our own data.
-Children 2;0 and older showed a slight advantage for lexicosyntax in the form of a transition type effect: children made more anticipatory switches following questions than non-questions in the two lexical conditions.
-But, the question effect only significant in the \textit{normal} condition, in which children had both lexicosyntactic and prosodic information.
-Similarly, we found no evidence that, for children or adults, lexicosyntax alone was ``sufficient'' (or otherwise equal to full linguistic information) for spontaneous turn prediction \citep[pg.
-531]{de-ruiter2006}.
+Children 2;0 and older showed a slight advantage for lexicosyntax in the form of a transition type effect: children made more anticipatory switches following questions than non-questions in conditions where lexical information was present.
+% But, the question effect only significant in the \textit{normal} condition, in which children had both lexicosyntactic and prosodic information. 
+But we found no evidence that, for children or adults, lexicosyntax alone was ``sufficient'' (or otherwise equal to full linguistic information) for spontaneous turn prediction \citep[pg.
+531]{de-ruiter2006}: Children's performance was best in conditions when they had access to the full linguistic signal (including prosody).
 
-Children begin to make predictions about upcoming turn structure early in development.
-The children in Study 1 were nearly as good as the adults when it came to tracking the current speaker and making anticipatory switches.
-In the \textit{normal} speech condition of Study 2, even one-year-olds robustly made more anticipatory switches than would be expected by chance, and some even showed the start of a question effect (Table 4).
-Young children are slow turn-takers in spontaneous conversation; children at 1;8--1;9 take an average of nearly 800 ms to respond to questions (compared to 200 ms for adult-adult conversation; \citealp{casillasUndRev, stivers2009}).
+\subsection{Developmental implications}
+
+Children begin to make predictions about upcoming turn structure early in development, so our findings suggest that response planning plays a key role in explaining why they are so slow at turn-taking in real-time interaction.
+First, three- to five-year-old children in Experiment 1 were nearly as good as the adults when it came to tracking the current speaker and making anticipatory switches.
+And second, in the \textit{normal} speech condition of Experiment 2, even one-year-olds robustly made more anticipatory switches than would be expected by chance; some even showed the start of a question effect.
+Yet, as reviewed in the introduction, young children are slow turn-takers in spontaneous conversation.
 So, despite the fact that children begin taking turns in infancy and spontaneously predict upcoming turn structure at age one, it takes them a long time to develop adult-like timing in conversation with others.
-This pattern of findings suggests that children's language production and response planning plays a key role in their response timing, slowing them down in real-time interaction.
 
-In both studies we found a robust advantage for questions in participants' anticipatory looks to the upcoming speaker.
-Young children have a lot of experience in hearing and answering questions.
-A handful of studies estimate that questions make up approximately one third of the utterances children hear, before and after the onset of speech, and even into their preschool years \citep{fitneva2012, henning2005, shatz1979}\footnote{There is variation by individual and socioeconomic class \citep{hart1992}.}, many of which are ``test' questions early on---questions that the caregiver already has the answer to (e.g., ``What does a cat say?'').
+In both of our experiments we found a robust advantage for questions in participants' anticipatory looks to the upcoming speaker.
+This finding probably comes about because young children have a lot of experience in hearing and answering questions.
+Questions may make up approximately one third of the utterances children hear, before and after the onset of speech, and even into their preschool years \citep{fitneva2012, henning2005, shatz1979}\footnote{There is substantial variation in this proportion by individual and socioeconomic class \citep{hart1992}.}, many of which are ``test' questions early on---questions that the caregiver already has the answer to (e.g., ``What does a cat say?'').
 Caregivers use questions to get their young children's' attention and to ensure that information is in common ground, even if childrens' responses are non-verbal or infelicitous \citep{fitneva2012, snow1977}.
-Questions give children a way in to ongoing interaction.
+
+% I didn't really kno what to make of this -MCF
+% Questions also give children a way in to ongoing interaction.
 Because they initiate a two-part structure, questions give the addressee certainty about when to talk (next) and what to say (the answer).
@@
 
-It remains unclear from our results what it is about the questions that caused children to switch their gaze to the responder more often.
+What property of questions caused children to switch their gaze to the responder more often?
+We cannot definitively answer this question.
 Questions make it much more likely that a speaker switch will occur at the next possible point.
-They have both early and late syntactic cues to help distinguish them from other speech (e.g., subject-auxiliary inversion and a final rising tone).
+They also have both early and late syntactic cues to help distinguish them from other speech (e.g., subject-auxiliary inversion and a final rising tone).
 But other speech acts and request formats have these properties.
@@
 
-One of our primary aims was to see how linguistic information affects children's predictions about upcoming turn structure.
-To test this we limited the speech signal in our conversations to eliminate lexicosyntactic or prosodic information, either by showing participants conversations in languages they didn't understand or by phonetically manipulating the audio.
+A further aim of our study was to measure how linguistic information affects children's predictions about upcoming turn structure across development.
+To explore this question we limited the speech signal in our conversations to eliminate lexicosyntactic or prosodic information, either by showing participants conversations in languages they didn't understand or by phonetically manipulating the audio.
 Prior work with adults found a consistent and critical role for lexicosyntax, leaving prosody in more of a secondary and supporting role \citep{de-ruiter2006, magyari2012}.
 Knowing that children know more about prosody than lexicosyntax early on, we thought it might be possible that young children would show an advantage for prosody instead.
-Our results suggest that there are roles for both prosody and lexicosyntax, but that children may put one or the other information source to use at different ages.
+Our results suggest that there are roles for both prosody and lexicosyntax, but that there are different roles for these information sources at different ages.
 
@@
 However, in both studies we have presented, the question effect was not limited by lexicosyntactic information alone.
-Older children and adults in Study 1 still showed a question effect in languages that they did not understand.
-In Study 2, the question effects were only significant in the speech with lexicosyntactic \textit{and} prosodic information, again as a function of age.
+Older children and adults in Experiment 1 still showed a question effect in languages that they did not understand.
+In Experiment 2, the question effects were only reliable in the speech with lexicosyntactic \textit{and} prosodic information, again as a function of age.
 The results suggest that children get better at identifying questions in the ongoing conversation as they get older, with major gains between ages one and two, but that prosody still plays an important role in helping them to do so.
 
-These results are only partially consistent with prior work.
+\subsubsection{Relationship with prior work}
+
+Our findings are only partially consistent with those reported in prior work.
 De Ruiter and colleagues (2006) found that adults' accuracy in identifying upcoming turn-ends was not significantly affected when intonation was taken away.
 In their study of children and adults, Keitel and colleagues (2013) found the same for adults' anticipatory gaze switches, and only finding that intonation affected children's switches at 3;0 (with switches by younger children never exceeding chance).
-Ou results suggest that children's anticipatory switches emerge much earlier, and rely on prosody before they do lexicosyntax.
-Further, our adult control participants patterned similarly to the children: they showed question effects for languages they didn't speak in Study 1 and strongest question effects for lexicosyntactic information when it was combined with prosody in Study 2.
+Ou results, in contrast, suggest that children's anticipatory switches emerge much earlier, and rely on prosody before they do lexicosyntax.
+Further, our adult control participants patterned similarly to the children: they showed question effects for languages they didn't speak in Experiment 1 and strongest question effects for lexicosyntactic information when it was combined with prosody in Experiment 2.
 Thus, neither the child nor the adult data support the idea that lexicosyntactic information is equivalent to full linguistic information in making predictions about upcoming turn structure \citep{de-ruiter2006}, at least in participants' spontaneous looking behavior.
 
+Differences in experimental paradigm could begin to explain the disparity in our findings.
 The anticipatory switches participants make in our studies were spontaneous, and were made without a requirement for urgency or timeliness, as the responses are for a button-press measure \citep{de-ruiter2006, magyari2012}.
@@
 Although button press measures may be better able to tell us what participants \textit{can} do in predicting turn ends, they may also emphasize lexicosyntactic cues to a greater extent than we experience in normal conversation.
-If participants are primarily monitoring for speaker switch instead of turn ends in general, we need to more carefully consider how signals to impending speaker transition (verbal and non-verbal) are weighed against more generic signals to utterance structure in our cognitive-linguistic model of how people manage to take turns on time.
+
+% If participants are primarily monitoring for speaker switch instead of turn ends in general, we need to more carefully consider how signals to impending speaker transition (verbal and non-verbal) are weighed against more generic signals to utterance structure in our cognitive-linguistic model of how people manage to take turns on time.
 Current models of adult turn-taking underline the role of linguistic prediction, but more recent work, including our own, may point some future work toward more shallow cue monitoring (e.g., for \textit{wh-}words) and inhibition \citep{bogelsUndRev}.
 
-The studies we present here are not without their limitations.
-In Study 2, we tried to improve our method for controlling linguistic information by phonetically manipulating improvised dialogic speech.
-Although this allowed us to more closely follow prior work \cite{de-ruiter2006, keitel2013}, it also limits our interpretation of children's behavior.
+\subsection{Limitations}
+
+In Experiment 2, we tried to improve our method for controlling linguistic information by phonetically manipulating improvised dialogic speech.
+Although this decision allowed us to follow prior work more closely \cite{de-ruiter2006, keitel2013}, it also limits our interpretation of children's behavior.
 Low-pass filtered and robotic-like speech are not typical speech signals children hear, and so the limited cue conditions in Study 2 may have been a somewhat unfair comparison to the normal speech condition, which not only had full linguistic cues, but also was more typical for what children hear in their everyday life.
 An better alternative to explore for the future would be scripted speech, or speech carefully selected from a large corpus for specific linguistic properties.
+
 Relatedly, since we used improvised conversation recordings, we did not control how often the speakers switched back and forth, or how often a switch occurred after a pause in speech.
 It is possible that, to some extent, children could have formed expectations about the turn-structure on the basis of the frequent switching alone \citep[see, e.g., ][]{thorgrimssonUndRev}.
-If this were the only thing leading children's anticipatory gaze switches, we would have predicted more uniformity in looking behavior across the conditions and no major effects of linguistic signal, age, or transition type.
-But, another advantage of using scripted conversation would be the ability to control switch frequency to eliminate this issue.
+If this statistical expectation were the only thing leading children's anticipatory gaze switches, we would have predicted more uniformity in looking behavior across the conditions and no major effects of linguistic signal, age, or transition type, however, which we did not see.
+Nevertheless, another advantage of using scripted conversation would be the ability to control switch frequency and eliminate this issue.
+
 Finally, we carefully designed a scheme for identifying anticipatory turn switches in the data, building off of Keitel and colleagues' (2013) analyses.
@@
 (2013) probably captures most of the switching behavior.
-However, if early looks do indeed increase as children get older and have more access to early syntactic cues, it would be helpful to develop a second measure of very early anticipatory looks.
+But if early looks do indeed increase as children get older and have more access to early syntactic cues, it would be helpful to develop a second measure of very early anticipatory looks.
+
+\subsection{Conclusions}
+
+FILL ME IN
 

### 2014-09-11 paper/CasillasFrank-Anticip.tex
minor notes

@@
 
+SCAFFOLDING: 1. QUESTIONS 2. INFORMATION INTEGRATION
+
 Children begin to make predictions about upcoming turn structure early in development, so our findings suggest that response planning plays a key role in explaining why they are so slow at turn-taking in real-time interaction.
@@
 
-\subsubsection{Relationship with prior work}
+\subsection{Relationship with prior work}
 
@@
 Although button press measures may be better able to tell us what participants \textit{can} do in predicting turn ends, they may also emphasize lexicosyntactic cues to a greater extent than we experience in normal conversation.
+
+KEITEL??? STIMULUS DIFFERENCES? NUMBER OF TRIALS?  BUT WE REALLY DON'T KNOW. GENERAL CONSISTENCY - PROSODY DOES MATTER.
+
+END BY HIGHLIGHTING OUR CONTRIBUTION - ISOLATE LEXICOSYNTAX
 
@@
 
-FILL ME IN
+Because early conversation plays a central role in children's language learning, we can not fully understand language acquisition without taking conversational skill into account.
+Conversation is the driving force behind what children say and hear, and participants have to know the rules of the game to play effectively---especially when multiple players are involved.
+Prediction allows ready speakers to get a word in at just the right moment.
+It also helps participants to comprehend  ongoing speech and catch onto potential misunderstandings via comparison to anticipated speech (e.g., `yes' or `no' in response to a polar question).
+Language is enormously helpful in making predictions of this type, and several studies have shown that adults rely on lexicosyntactic \citep{de-ruiter2006, magyari2012} and, to some extent, prosodic information \citep{ford1996, bogelsUndRev} in their turn-taking.
+FINISH IT!
+% In the current study, we have tried to tie together several loose ends concerning when and how children begin to use linguistic information in their predictions about upcoming turn structure.
 

### 2014-09-19 paper/CasillasFrank-Anticip.tex
Some changes to text, mostly in discussion. Figure axis changes. Updates to references.

--- 2014-09-10 paper/CasillasFrank-Anticip.tex
@@
 
-\author[MPI]{Marisa Casillas}
+\author[MPI]{Marisa Casillas\corref{cor1}}
 % \address[StanfordLX]{Department of Linguistics, Stanford University}
 \address[MPI]{Max Planck Institute for Psycholinguistics, Nijmegen}
+\cortext[cor1]{Corresponding author}
 
@@
 \begin{abstract}
-Blah blah blah abstract needs to be written here.
-
+We investigate language acquisition through the lens of a fundamental conversational skill: turn taking.
+Children begin developing turn-taking skills in infancy, but take several years to assimilate language into their turn-taking systems.
+In two eye-tracking experiments with children ages 1--6, we measured participants' spontaneous predictions about upcoming speaker change while controlling the amount of linguistic information available.
+We found that children already predicted upcoming turns at age one, and integrated linguistic cues differently at different ages.
+Children under 3;0 showed a general advantage for prosody over lexicosyntax, contrary to the typical finding for adults.
+Children 2;0 and older showed more predictive switches for questions than non-questions, but only when lexical information was available.
+But, we found no evidence that lexicosyntax is sufficient for turn prediction---participants' performance was best overall with access to lexicosyntax and prosody together.
+Our results point to the importance of interaction in children’s real-time linguistic processing.
 \end{abstract}
@@
 Like other types of human interaction, it is organized at its core by the roles and goals of its participants.
-What sets conversation apart is its structure, however: sequences of interconnected, communicative actions that take place across alternating turns at talk.
+But what sets conversation apart is its structure: sequences of interconnected, communicative actions that take place across alternating turns at talk.
 Sequential, turn-based structures in conversation are strikingly uniform across language communities and linguistic modalities.
@@
 Children participate in sequential coordination with their caregivers starting at three months of age---before they can rely on any linguistic cues in taking turns \citep{snow1977, jaffe2001, hilbrinkInPrep}.
-During their first year, infants know little about the lexical and syntactic systems of their language \citep[though they \textit{do} know a little about prosody;][]{christophe2001, bertoncini2013}.
 Of course, infant turn-taking is different from adult turn-taking in several ways.
@@
 Early non-verbal interactions establish the protocol by which children come to use language with others.
-
-% If we want to grasp children's linguistic input and practice, we need to understand their conversational skill development.
-And if we want to understand how children develop conversational skills, we need to take a close look at how turn-taking develops in tandem with language.
-In this study, we investigate when children begin to make predictions about upcoming turn structure, and how they integrate language into their predictions as they grow older.
-In what follows, we first review background about turn-taking and the state of current knowledge about adult turn prediction; we then discuss recent work on the development of turn-taking before turning to the details of our own study.
-
-% For example, they show sensitivity to language-specific aspects of prosody in the first few days of life, and can use prosody for word segmentation and emotion recognition by 7--8 months \citep{jusczyk1999, johnson2001, grossmann2005}.
-In contrast, children don't know much about syntactic structure until their third year \citep{clark2009}.
+And as they acquire language, they also come to integrate it into the preverbal turn-taking systems.
+
+In this study, we investigate when children begin to make predictions about upcoming turn structure in conversation, and how they integrate language into their predictions as they grow older.
+In what follows, we first give a basic review of turn-taking research and the state of current knowledge about adult turn prediction.
+We then discuss recent work on the development of turn-taking skills before turning to the details of our own study.
 
@@
 Many other human activities are organized around sequential turns.
-Traffic in intersections and communication in computer networks both have characteristic turn-taking.
+Traffic intersections and computer network communication both use turn-taking systems.
 Children's early games (e.g., give-and-take, peek-a-boo) have built-in, predictable turn structure \citep{ratner1978, ross1987}.
@@
 
-Conversation distinguishes itself from non-conversational turn-taking behaviors, however, by the complexity of the turn sequencing involved.
+Conversation distinguishes itself from non-conversational turn-taking behaviors by the complexity of the turn sequencing involved.
 In the examples above (traffic, games, and monkeys) the set of sequence and action types is far more limited and predictable than what we find in everyday talk.
@@
 Despite this complexity, conversational turn-taking is precise in its timing.
-Across a diverse sample of 10 languages, one study found a consistent average transition time of 0--200 msec at points of speaker switch \citep{stivers2009}.
+Across a diverse sample of conversations in 10 languages, one study found a consistent average transition time of 0--200 msec at points of speaker switch \citep{stivers2009}.
 Current models of speech production suggest that it takes approximately 600 msec to produce a word, and even longer to produce a simple utterance \citep{levelt1989}.
-So, in order to achieve 200 msec turn transitions, speakers must begin formulating their response before the prior turn has ended.
+So in order to achieve 200 msec turn transitions, speakers must begin formulating their response before the prior turn has ended.
 To formulate their response early on, speakers must track and anticipate what types of response might become relevant next.
@@
 Adults have a lot of information that could potentially help with their predictions about upcoming turn structure.
-Lexical, syntactic, and prosodic information (e.g., \textit{wh}- words, subject-auxiliary inversion, list intonation) can all inform addressees about upcoming linguistic structure \citep{de-ruiter2006, duncan1972, ford1996, bogelsUndRev}.
+Lexical, syntactic, and prosodic information (e.g., \textit{wh}- words, subject-auxiliary inversion, list intonation) can all inform addressees about upcoming linguistic structure \citep{de-ruiter2006, duncan1972, ford1996, torreiraUndRev}.
 Non-verbal cues (e.g., gaze, posture, and pointing) often appear at turn-boundaries as late indicators of an upcoming speaker switch \citep{rossano2009, stivers2010}.
 The sequential context of a turn can make it clear what will come next: answers after questions; thanks or denial after compliments, et cetera \citep{schegloff2007}.
-Prior work suggests that adult listeners primarily use lexicosyntactic information to accurately predict upcoming turn structure, while prosodic information is plays more of a secondary role \citep{de-ruiter2006}.
-For example, a mid-high ``continuation'' tone can cancel the possibility of a speaker switch at an otherwise syntactically- and pragmatically-complete turn point (e.g., at the start of a story-telling; \citealp{ford1996, bogelsUndRev}).
+Prior work suggests that adult listeners primarily use lexicosyntactic information to accurately predict upcoming turn structure, while prosodic information plays more of a secondary role \citep{de-ruiter2006}.
+For example, a mid-high ``continuation'' tone can cancel the possibility of a speaker switch at an otherwise syntactically- and pragmatically-complete turn point (e.g., at the start of a story-telling; \citealp{ford1996, torreiraUndRev}).
 
 Recent experimental work on adult turn-taking has tried to tease apart which linguistic cues responders rely on in making online predictions about upcoming turn structure.
-De Ruiter and colleagues \citeyearpar{de-ruiter2006} asked participants to listen to snippets of spontaneous conversation and press a button whenever they anticipated that the current speaker was about to finish her turn.
-The snippets that participants listened to were controlled for the amount of linguistic information present; some snippets were normal, unmodified versions of the recording, but some had words and rhythm without pitch, pitch and rhythm but no words, neither words nor pitch, or were replaced entirely with a noise profile matched to the original recording.
+De Ruiter and colleagues \citeyearpar{de-ruiter2006} asked participants to listen to snippets of spontaneous conversation and to press a button whenever they anticipated that the current speaker was about to finish her turn.
+The snippets that participants listened to were controlled for the amount of linguistic information present; some snippets were normal, unmodified versions of the recording, but some had words and rhythm (but no pitch), pitch and rhythm (but no words), neither words nor pitch, or were replaced entirely with a noise profile matched to the original recording.
 In the normal, unmodified snippets, participants' anticipatory button presses were highly accurate, falling an average of 180 msec before the phonetic end of the turns.
-Participants also made accurate predictions for the ``words and rhythm without pitch'' snippets; there was no significant difference compared to their performance in the normal  condition.
+Participants also made accurate predictions for the snippets with words and rhythm (but no pitch); there was no significant difference compared to their performance in the normal condition.
 But, in the snippets with no lexical information, participants' accuracy significantly declined.
@@
 They also found a significant, but secondary role for pitch: participants' accuracy in the pitch and rhythm (but no words) condition was still better than the neither-words-nor-pitch and the noise profile conditions.
-They concluded that pitch's importance for online prediction of turn-ends is restricted, and secondary to lexicosyntactic information.
+They concluded that pitch's importance for online prediction of turn-ends is restricted, falling secondary to lexicosyntactic information.
 
 Congruent evidence comes from studies varying the predictability of lexicosyntactic and pragmatic content.
-% Accurate prediction of linguistic and pragmatic content is essential to the mechanisms underlying adult turn taking. 
-Adults anticipate turn ends better when they can more accurately predict exact words that will come next (\citealp{magyari2012}; see also \citealp{magyariUndRev}).
+Adults anticipate turn ends better when they can more accurately predict the exact words that will come next (\citealp{magyari2012}; see also \citealp{magyariUndRev}).
 They can also identify speech acts within the first word of an utterance \citep{gisladottirUndRev}, allowing them to start planning their response at the first moment possible.
-\citet*{bogelsmagyariInPrep} used a trivia set-up to get participants to answer questions in which the answer became known either early or late in the question (e.g., ``Which character, also called 007, appears in the famous movies?'' vs.
-``Which character from the famous movies, is also called 007?'').
+\citet*{bogelsmagyariInPrep} used a trivia quiz set-up to get participants to answer questions in which the answer became known either early or late in the question (e.g., ``Which character, also called 007, appears in the famous movies?'' vs.
+``Which character, from the famous movies, is also called 007?'').
 Their participants showed a consistent neural response 500 msec after the answer became unambiguous, suggesting immediate response planning.
 
-In another study investigating the relationship between lexicosyntactic and prosodic cues, \citet*{bogelsUndRev} used a combination of button-press and verbal responses.
- % to more explicitly test the role of prosodic cues in turn-boundary identification. 
-They found that participants never verbally responded or pressed the ``anticipation'' button when hearing a syntactically-complete unit without a boundary tone.
-But utterance-level prosodic boundaries embedded in multi-utterance turns \textit{could} trick participants into anticipating a turn end 29\% of the time (see also de \citet{de-ruiter2006}:525).
+In another study, \citet*{torreiraUndRev} used a combination of button-press and verbal responses to investigate the relationship between lexicosyntactic and prosodic cues in turn-boundary identification.
+They found that participants never verbally responded or pressed the ``turn-end'' button when hearing a syntactically-complete unit without a boundary tone.
+However, utterance-level prosodic boundaries embedded in multi-utterance turns \textit{could} trick participants into anticipating a turn end 29\% of the time (see also de \citet{de-ruiter2006}:525).
 Their findings corroborate other corpus and experimental findings promoting a combination of cues (lexicosyntactic, prosodic, and pragmatic) as key for accurate turn-end prediction \citep{duncan1972, ford1996, hirvenkari2013}.
@@
 The acquisition of turn-taking skills must then be tied to children's knowledge about language, gaze, gesture, and social cues.
-But children's turn taking starts early in infancy, long before first words or gestures emerge; a primary role for lexicosyntactic cues thus doesn't fit well with children's pre-verbal turn-taking.
+But children's turn taking starts early in infancy, long before first words or gestures emerge; a primary role for lexicosyntactic cues doesn't fit well with children's pre-verbal turn-taking.
 
@@
 
-The majority of work on children's early turn-taking has focused on observations of spontaneous interaction.
+The majority of work on children's early turn taking has focused on observations of spontaneous interaction.
 Children's first turn-like structures appear as early as two to three months in proto-conversation with their caregivers.
@@
 Between four and nine months, children begin to reduce the number of turns happening overlap while also improving on their average response latency.
-But then children's response latencies slow down again, peaking at average gaps of 1100 msec at nine months, with only very gradual improvement after that \citep{hilbrinkInPrep}.
+But then, later on, children's response latencies slow down again, peaking at average gaps of 1100 msec at nine months, with only very gradual improvement after that \citep{hilbrinkInPrep}.
 While children's avoidance of overlap is nearly adult-like by nine months, the timing of their non-overlapped responses stays much longer than the 200 msec standard for the next several years \citep{ervin-tripp1979}.
@@
 One study found that 12-month-olds make more predictive gaze shifts while watching human verbal conversation compared to object non-speech ``conversation'' \citep{bakker2011}, but another only found a similar effect at 36 months \citep{hofsten2009}.
-Neither of these studies had baselines to which the turn-relevant looking behavior could be compared, however.
-Such a baseline measurement is critical because there may be developmental differences in gaze shifting between conversational participants, even if such shifting is not related to turn structure.
-These developmental differences could produce artifactual changes in measures of turn-contingent shifting.
+However, neither of these two studies had baselines to which the turn-relevant looking behavior could be compared.
+A baseline measurement is critical because there may be developmental differences in gaze shifting between conversational participants, even if the shifting is not related to turn structure.
+Such developmental differences could produce artifactual changes in measures of turn-contingent shifting.
 
@@
 Their study was the first to focus on the role of linguistic processing in children's turn predictions.
-They showed their participants two types of conversation videos: one normal and one with flattened pitch (i.e., with removed intonation contours), finding that only three-year-olds were affected by a lack of intonation.
+They showed their participants two types of conversation videos: one normal and one with flattened pitch (i.e., with removed intonation contours), finding that only 36-month-olds were affected by a lack of intonation.
 The adult control group made equal numbers of anticipatory looks in the videos, with and without intonation contours, consistent with prior adult findings \citep{de-ruiter2006}.
@@
 Because these limitations directly inform our own study design, we review them in some detail.
-First, their estimates of baseline gaze frequency were computed only during conversational turns (not pauses), while their estimates of anticipatory shifting were computed during pauses.
-Assuming that gaze switches happen most often during the gap between turns \citep{hirvenkari2013}, their choice of a ``random'' (baseline) gaze window would have maximized the possibility of finding a difference between ``random'' and anticipatory gaze frequencies.
-% \footnote{If so, the null effect for children under 3;0 could be given a stronger interpretation and the significant effect for three-year-olds and adults could be given a weaker interpretation.} 
-A stronger baseline would be to compare participants' looking behavior at turn transitions to their looking behavior during randomly-selected windows of time throughout the stimulus.
-We follow this approach in our work.
+First, their estimates of baseline ``random'' gaze frequency were computed from switches that occurred while the speaker's turn was ongoing (i.e, during turns, and not during inter-turn gaps).
+Their estimates of anticipatory shifting frequency were computed from switches that occurred within the time window immediately surrounding inter-turn gaps (i.e., around inter-turn gaps, and primarily not during turns).
+Assuming that gaze switches happen most often during the inter-turn gap \citep{hirvenkari2013}, switches made during ongoing speech (their ``random'' baseline frequency) would be at a minimum.
+Therefore their ``random'' (baseline) frequency actually maximized the possibility of finding a difference between ``random'' and anticipatory gaze frequencies.
+A more conservative baseline would be to compare participants' looking behavior at turn transitions to their looking behavior during randomly-selected windows of time throughout the stimulus.
+We follow this latter approach in our work.
 
 Second, the conversation stimuli they used were somewhat unusual.
-The average gap between turns was 900 msec, which is much longer than typical adult timing, where gaps are often around 200 msec \citep{stivers2009}.
-The speakers in the video were also asked to minimize their movements while performing a scripted, adult-directed conversation, which would have created a somewhat unnatural stimulus.
+The average gap between turns was 900 msec, which is much longer than typical adult timing, where gaps average around 200 msec \citep{stivers2009}.
+The speakers in the videos were also asked to minimize their movements while performing a scripted, adult-directed conversation, which would have created a somewhat unnatural stimulus.
 In addition, in order to produce more naturalistic conversation, it would have been ideal to localize sources for the two voices in the video (i.e., to have the voices come out of separate left and right speakers).
 But both voices were recorded and played back on the same audio channel, which may have made it more difficult to distinguish the speakers.
-We again attempt to address these issues in our current study.
-
-% A single-channel recording was most problematic for the no-intonation condition, where the two voices were flattened to have the same pitch. 
+Again, we attempt to address these issues in our current study.
+
 Despite these minor methodological issues, the Keitel et al.
@@
 In two eye-tracking experiments, we measured children's anticipatory gaze to upcoming responders while controlling for the amount of lexicosyntactic and prosodic information available.
-Experiment 1 uses a paradigm in which participants view video clips of naturalistic conversations across a number of different languages, so as to create conditions under which lexicosyntactic information was differentially available, but prosodic and non-linguistic cues were relatively constant.
+Experiment 1 uses a paradigm in which English-speaking participants view video clips of naturalistic conversations from several different languages, to create conditions under which lexicosyntactic information was differentially available, but prosodic and non-linguistic cues were intact.
 This paradigm showed minimal differences between the predictive looking behavior of preschoolers and adults.
-In Experiment 2, we created artificial (puppet) visual stimuli to better separate these cues from one another.
+In Experiment 2, we created artificial (puppet) visual stimuli to better separate these linguistic cues from one another.
 In this more controlled paradigm, we found that children's predictive looking behavior improved from ages one to six, but that even one-year-olds made more anticipatory looks than would be expected by chance.
 
-In both experiments, children consistently looked faster to responders after hearing questions, compared to non-questions.
+In both experiments children consistently looked faster to responders after hearing questions, compared to non-questions.
 Both prosodic and lexicosyntactic information played a role in children's predictions about turn structure, but the two information sources were used differently at different ages.
-Our findings overall support an account in which predictive processes for turn-taking in conversation are present early but their integration with linguistic information takes substantial practice.
+Our findings overall support an account in which predictive processes for turn-taking in conversation are present early, but their integration with linguistic information takes substantial practice.
 
@@
 We chose male-female pairs as a natural way of creating contrast between the two talker voices.
-See the videos here: NOTE: add link.}
+See the videos here: NOTE: Link to be added.}
 
@@
 \end{center}
-\caption{Baseline-corrected proportion anticipatory gaze switches made for turn transitions in each language condition, and across ages. Error bars represent the standard error of the mean.} 
+\caption{Baseline-corrected proportion of anticipatory gaze switches made for turn transitions in each language condition, and across ages. Error bars represent the standard error of the mean.} 
 \label{fig:conditionsFL}
@@
 \end{center}
-\caption{Proportion anticipatory gaze switches made for question- (dashed) and non-question- (dotted) turn transitions in each condition, and across ages.
+\caption{Baseline-corrected proportion of anticipatory gaze switches made for question- (dashed) and non-question- (dotted) turn transitions in each condition, and across ages.
 Error bars represent the standard error of the mean.}
@@
 
-Children's anticipatory gaze switches showed a main effect of language condition (\textit{$\beta$}=-0.448, \textit{SE}=0.126, \textit{t}=-3.56).
-As predicted, children made more anticipatory looks while they were watching videos that they could understand.
+Children's anticipatory gaze switches showed a main effect of language condition (\textit{$\beta$}=-0.448, \textit{SE}=0.126, \textit{t}=-3.56, \textit{p}$<$.001).\footnote{We derive \textit{p}-values here by treating \textit{t} and \textit{z} (with a two-tailed assumption), as is justified for large data sets \citep{barr2013}.} As predicted, children made more anticipatory looks while they were watching videos that they could understand.
 In the English videos, children had access to lexicosyntactic information, which could have helped them make better predictions about upcoming turn structure.
@@
 
-Contrary to our expectation, there was no main effect of age (\textit{$\beta$}=-0.03, \textit{SE}=0.027, \textit{t}=-1.145).
-Children's anticipatory switches \textit{did} increase with age for the non-English videos, but it also decreased with age for the English videos, resulting in a significant interaction of age and language condition (\textit{$\beta$}=0.083, \textit{SE}=0.027, \textit{t}=3.063).
+Contrary to our expectation, there was no main effect of age (\textit{$\beta$}=-0.03, \textit{SE}=0.027, \textit{t}=-1.145, \textit{p}=.252).
+Children's anticipatory switches \textit{did} increase with age for the non-English videos, but it also decreased with age for the English videos, resulting in a significant interaction of age and language condition (\textit{$\beta$}=0.083, \textit{SE}=0.027, \textit{t}=3.063, \textit{p}$<$.001).
 We had expected that, as children got older, they would also show more anticipatory switches.
@@
 Children made more anticipatory switches after hearing questions than non-questions, but the advantage for questions was driven by age and linguistic access.
-There was no main effect of transition type alone (\textit{$\beta$}=-0.209, \textit{SE}=0.151, \textit{t}=-1.384).
-Rather, the model confirmed a two-way interaction of linguistic condition and transition type (\textit{$\beta$}=0.445, \textit{SE}=0.19, \textit{t}=2.344).
+There was no main effect of transition type alone (\textit{$\beta$}=-0.209, \textit{SE}=0.151, \textit{t}=-1.384, \textit{p}=.166).
+Rather, the model confirmed a two-way interaction of linguistic condition and transition type (\textit{$\beta$}=0.445, \textit{SE}=0.19, \textit{t}=2.344, \textit{p}$<$.02).
 Presumably, question effects were larger in the English videos than the non-English videos because there were more cues to questionhood in English (i.e., non-verbal, prosodic, \textit{and} lexicosyntactic).
 Children younger than five did not distinguish questions from non-questions in the non-English videos (Figure \ref{fig:questionsFL}).
-The children's model confirmed a three-way interaction of age, linguistic condition, and transition type (\textit{$\beta$}=-0.091, \textit{SE}=0.042, \textit{t}=-2.183).
+The children's model confirmed a three-way interaction of age, linguistic condition, and transition type (\textit{$\beta$}=-0.091, \textit{SE}=0.042, \textit{t}=-2.183, \textit{p}$<$.05).
 Lexicosyntactic information might then be critical for young children to identify questions in conversation.
@@
 
-Inter-turn gap duration did not affect children's anticipatory switches (\textit{$\beta$}=0.202, \textit{SE}=0.151, \textit{t}=1.340), but it was the main factor driving adults' switches (\textit{$\beta$}=0.648, \textit{SE}=0.047, \textit{t}=13.855).
-The adult model also showed a marginal effect of transition type, with adults looking more after questions compared to non-questions (\textit{$\beta$}=-0.117, \textit{SE}=0.069, \textit{t}=-1.704).
+Inter-turn gap duration did not affect children's anticipatory switches (\textit{$\beta$}=0.202, \textit{SE}=0.151, \textit{t}=1.340, \textit{p}=.180), but it was the main factor driving adults' switches (\textit{$\beta$}=0.648, \textit{SE}=0.047, \textit{t}=13.855, \textit{p}$<$.001).
+The adult model also showed a marginal effect of transition type, with adults looking more after questions compared to non-questions (\textit{$\beta$}=-0.117, \textit{SE}=0.069, \textit{t}=-1.704, \textit{p}=.088).
 Language condition and its interaction with transition type were not significant.
@@
 It appears that, especially in English, children were as good at tracking the current speaker and anticipating upcoming turns as adults were.
-To find out when children first begin making spontaneous anticipatory looks to responders, younger participants are required.
+To find out when children first begin making spontaneous anticipatory looks to responders, it is necessary to recruit younger participants.
 Experiment 2 thus includes a sample of one- and two-year-olds.
@@
 Children had prosodic cues in both English and non-English videos, although the prosodic signal was non-native in the latter case.
-So, although using non-English videos was a natural way to control for lexical information, it did not allow for assessment of the role of prosody.
+So, although using non-English videos was a natural way to control for lexical information, it did not allow us to assess the role of prosody.
 Relatedly, the advantage for questions was greater in the English videos than the non-English videos, suggesting that lexicosyntactic cues are important for children in identifying questions during conversation.
@@
 \subsubsection{Materials}
-We created six short videos of improvised, child-friendly conversation (Figure \ref{fig:puppets}).
+We created 18 short videos of improvised, child-friendly conversation (Figure \ref{fig:puppets}).
 To eliminate non-verbal cues to turn transition and to control the types of linguistic information available in the stimuli we (1) recorded improvised conversations, (2) phonetically manipulated those recordings to limit the availability of prosodic and lexical information, and (3) recorded a new set of videos that featured puppets as talkers, using the manipulated audio as the puppets' speech.
@@
 We randomly assigned half of the conversation topics (`birthday cake', `pets', and `breakfast') to the \textit{normal} condition, and half to the limited cue conditions (`riding bikes', `rainy days', and `the library').
-We then created three versions of the experiment, so that each of the six puppet pairs was associated with three different conversation topics across the different versions of the experiment.
-We ensured that the position of the talkers (left and right) was counterbalanced in each version by flipping the video and audio channels as needed.\footnote{See the videos here: FIX add link.}
+We then created three versions of the experiment, so that each of the six puppet pairs was associated with three different conversation topics across the different versions of the experiment (18 videos in total).
+We ensured that the position of the talkers (left and right) was counterbalanced in each version by flipping the video and audio channels as needed.\footnote{See the videos here: NOTE: Link to be added.}
 
@@
 Question transitions (mean$=$358, median$=$405) were longer than non-question transitions (mean$=$296, median$=$288) on average, but gap duration was overall comparable across conditions: \textit{normal} (mean$=$333, median$=$337), \textit{words only} (mean$=$345, median$=$383), \textit{prosody only} (mean$=$320, median$=$336), and \textit{no words} (mean$=$324, median$=$328).
-The longer gaps for question transitions could give them an advantage because our anticipatory measure includes shifts initiated during the gap between turns (Figure \ref{fig:algorithm}).
+The longer gaps for question transitions could give them an advantage because our anticipatory measure includes shifts initiated during the gap between turns (Figure \ref{fig:criterion}).
 
@@
 For example, anticipatory switches in the \textit{no speech} videos were often only modestly above chance (with the exception of 5- and 6-year-olds).
-Further, 1- and 2-year-olds did not appear to reliably outperform chance conditions for the \textit{words only} videos.
+Notably, while children performed well with \textit{prosody only} speech from the start, they seemed to only reliably exceed chance in the \textit{words only} condition at ages 3;0 and older, giving some support to the idea that there is an early advantage for prosody over lexicosyntax in children's turn prediction.
 
@@
 \end{center}
-\caption{Proportion anticipatory gaze switches made for turn transitions in each speech condition, and across ages. Error bars represent the standard error of the mean.} 
+\caption{Baseline-corrected proportion anticipatory gaze switches made for turn transitions in each speech condition, and across ages. Error bars represent the standard error of the mean.} 
 \label{fig:conditionsEN}
@@
 While children made small, sometimes non-linear, gains in their anticipatory switches for non-questions as they grew older, they made large gains in their anticipatory switches with age (especially between ages one and two; Figure \ref{fig:questionsEN}).
-The increased question effect from age one to two indicates that, around this age, children begin to leverage lexicosyntactic cues in the speech signal to identify questions in ongoing conversation.
+The increased switching for questions between ages one to two indicates that, around this age, children begin to leverage lexicosyntactic cues in the speech signal to identify questions in ongoing conversation.
 In the lexical conditions, children had access to early cues to questionhood, such as \textit{wh}-words and subject-auxiliary inversion, to predict an upcoming speaker change.
@@
 
-Notably, while children performed well with \textit{prosody only} speech from the start, they seemed to only reliably exceed chance in the \textit{words only} condition at ages 3;0 and older.
-We confirmed this pattern in two \textit{post-hoc} \textit{t}-tests comparing children's anticipatory switching rate against their estimated random switching rate before age 3;0.
-We found that children under 3;0 significantly exceeded chance for the \textit{prosody only} condition (\textit{t}(187.677)=-3.004, \textit{p}=.003), but not for the \textit{words only} condition (\textit{t}(229.828)=-0.8307, \textit{p}=.407), giving some support to the idea that there is an early advantage for prosody over lexicosyntax in children's turn prediction.
-All in all, children showed increases in anticipatory switching with age across all conditions.
+Children showed increased anticipatory switching with age across all conditions, making almost as many transitions as adults ages five and six.
 Adults reacted to the \textit{no speech} condition differently than the children did, tracking the interaction less closely than in the three linguistic conditions.
@@
 There was a significant effect of gap duration.
-Children had time to make more anticipatory switches in analysis windows with longer gaps (\textit{$\beta$}=0.294, \textit{SE}=0.091, \textit{t}=3.231).
+Children had time to make more anticipatory switches in analysis windows with longer gaps (\textit{$\beta$}=0.294, \textit{SE}=0.091, \textit{t}=3.231, \textit{p}$<$.01).
 An effect of gap duration is expected since prior work using spontaneous anticipatory switching has found that most switches occur in the inter-turn gap (Keitel et al., 2013; Hirvenkari, 2013; Tice and Henetz, 2011).
@@
 Condition-related differences were driven by age and transition type (Figure \ref{fig:questionsEN}).
-Consequently, the model of children's anticipatory switches revealed no significant main effects of condition, despite the large changes in the linguistic information available across conditions.
+Consequently, the model of children's anticipatory switches revealed no significant main effects of condition, despite large differences in the linguistic information available across conditions.
 One interpretation of this finding is that children do not need linguistic information to predict upcoming turns.
@@
 With sufficiently long inter-turn gaps, participants could simply switch their gaze to the responder when the prior talker finishes talking---a strategy that doesn't rely on linguistic information.
-But this interpretation is not compatible with our finding that children make more anticipatory looks after hearing questions, that they do so more often when lexical information is available, and that they do even more so as they age.
+But this interpretation is not compatible with our findings that children make more anticipatory looks after hearing questions, that they do so more often when lexical information is available, and that they do even more so as they age.
 
 The model of children's data confirmed two three-way effects of age, condition, and transition type.
-Children made significantly more anticipatory switches with age in the \textit{normal} (\textit{$\beta$}=0.04, \textit{SE}=0.008, \textit{t}=5.168) and \textit{words only} conditions (\textit{$\beta$}=0.031, \textit{SE}=0.01, \textit{t}=3.170): they made a switch twice as often after hearing a question (16\%) than a non-question (8\%) in the \textit{words only} condition, and nearly four times as often in the \textit{normal} condition (19\% vs.
+Children made significantly more anticipatory switches for questions with age in the \textit{normal} (\textit{$\beta$}=0.04, \textit{SE}=0.008, \textit{t}=5.168, \textit{p}$<$.001) and \textit{words only} conditions (\textit{$\beta$}=0.031, \textit{SE}=0.01, \textit{t}=3.170, \textit{p}$<$.01).
+Overall, they made a switch twice as often after hearing a question (16\%) than a non-question (8\%) in the \textit{words only} condition, and nearly four times as often in the \textit{normal} condition (19\% vs.
 5\%).
@@
 There were also no further interactions of age and condition, or of transition type and condition.
-There was, however, a significant two-way interaction of transition type and age (\textit{$\beta$}=-0.031, \textit{SE}=0.01, \textit{t}=-3.18), in addition to the three-way interactions for the \textit{normal} and \textit{words only} conditions already mentioned.
-
-Last but not least, the model confirmed a significant main effect of age in the children's data (\textit{$\beta$}=0.028, \textit{SE}=0.009, \textit{t}=3.018).
-Children made more anticipatory switches as they got older.
-Many of the increased anticipatory looks with age came from questions in the lexical conditions.
-Children also increased their anticipatory looking with age for non-questions and non-lexical conditions, but the gains were smaller and sometimes followed a non-linear trajectory (Figure \ref{fig:questionsEN}).
+There was, however, a significant two-way interaction of transition type and age (\textit{$\beta$}=-0.031, \textit{SE}=0.01, \textit{t}=-3.18, \textit{p}$<$.01) in addition to the three-way interactions for the \textit{normal} and \textit{words only} conditions already mentioned.
+
+Last but not least, we confirmed that children generally made more anticipatory switches as they got older: The model showed a significant main effect of age (\textit{$\beta$}=0.028, \textit{SE}=0.009, \textit{t}=3.018, \textit{p}$<$.01).
+Though the increase in anticipatory looks with age came, in large part, from questions effects in the lexical conditions, children also showed increased switches for non-questions and non-lexical conditions.
+The increases related to non-questions and non-lexical conditions were overall smaller and sometimes followed a non-linear trajectory (Figure \ref{fig:questionsEN}).
 As in Study 1, early switches (planned before the end of the prior turn) did not show large increases with age, with adults making the fewest early switches in the three linguistic conditions (Figure \ref{fig:cumulativeEN}).
@@
 \textit{yes-no}; early vs.
-late), but we do not have enough data to make those distinctions in the current analysis.
-Also, nearly 43\% of one-year-olds showed question effects in the \textit{normal} speech condition, which has the full range of linguistic cues children hear in everyday life.
+late in the utterance), but we do not have enough data to make those distinctions in the current analysis.
+Also, 43\% of one-year-olds already showed more switching for questions than non-questions in the \textit{normal} speech condition---the only condition with the full range of linguistic cues children hear in everyday life.
 Future work with more fine-grained age samples and better controlled \textit{normal} speech might then find that children can identify questions in conversation even earlier.
@@
     5 & 0.91 & 0.52 \\
-    6 & 0.75 & 0.5 \\
+    6 & 0.75 & 0.50 \\
     Adult & 0.67 & 0.57 \\
@@
 
+Finally, we confirmed that young children showed a benefit for prosody over lexicosyntax.
+We conducted two \textit{post-hoc} \textit{t}-tests for children under 3;0, comparing their anticipatory switching rate against their estimated random switching rate in the \textit{prosody only} and \textit{words only} conditions.
+We found that anticipatory gazes significantly exceeded chance for the \textit{prosody only} condition (\textit{t}(187.677)=-3.004, \textit{p}=.003), but not for the \textit{words only} condition (\textit{t}(229.828)=-0.8307, \textit{p}=.407).
+This finding suggests that there is an early advantage for prosody over lexicosyntax in children under 3;0.
+The advantage for prosody was not driven by a question effect; children and adults did not consistently show more switches after questions than non-questions in the \textit{prosody only} condition (Figure \ref{fig:questionsEN}).
+So, whereas the early advantage for prosody appeared for both transition types (question and non-question), the later-emerging advantage for lexicosyntax and prosody together was primarily limited to question transitions.
+This pattern of results may suggest a qualitative shift in the kind of predictions children make while observing a conversation, with linguistic cues (like prosody) playing different roles at different ages.
+
 The model of adults' anticipatory gaze switches confirmed main effects of gap duration and linguistic condition.
-Longer gaps led to more anticipatory switches (\textit{$\beta$}=0.656, \textit{SE}=0.176, \textit{t}=3.726).
-Adults also made significantly more anticipatory switches in the \textit{normal} condition than in the \textit{no speech} baseline, indicating that linguistic access did affect their overall performance (\textit{$\beta$}=-0.211, \textit{SE}=0.085, \textit{t}=-2.486).
+Longer gaps led to more anticipatory switches (\textit{$\beta$}=0.656, \textit{SE}=0.176, \textit{t}=3.726, \textit{p}$<$.001).
+Adults also made significantly more anticipatory switches in the \textit{normal} condition than in the \textit{no speech} baseline, indicating that linguistic access did affect their overall performance (\textit{$\beta$}=-0.211, \textit{SE}=0.085, \textit{t}=-2.486, \textit{p}$<$.05).
 There were no other main effects of linguistic condition.
-The model also showed a marginal two-way interaction of condition and transition type, with adults making more anticipatory gazes for questions in the \textit{normal} condition (\textit{$\beta$}=0.194, \textit{SE}=0.11, \textit{t}=1.773).
+The model also showed a marginal two-way interaction of condition and transition type, with adults making more anticipatory gazes for questions in the \textit{normal} condition (\textit{$\beta$}=0.194, \textit{SE}=0.11, \textit{t}=1.773, \textit{p}=.076).
 Though the question effect was numerically present in both lexical conditions, it was only significant for adults in the \textit{normal} condition, in which they had both lexicosyntactic and prosodic cues to questionhood.
@@
 Our results hint that lexical information is critical in identifying questions during conversation, especially for children.
-Even so, performance was consistently best in the \textit{normal} condition, in which participants had lexical and prosodic information.
+Even so, performance was consistently best in the \textit{normal} condition, in which participants had lexical and prosodic information together.
 
@@
 As they got older, children demonstrated above-chance anticipatory switches in more conditions and made more anticipatory switches overall.
-Similar to Experiment 1, children between ages three and five behaved similarly to each other, and comparably to adults, showing small but steady gains in their anticipatory switches.
+Similar to Experiment 1, children between ages three and five behaved similarly to each other, and comparably to adults, showing small but steady gains in their anticipatory switches with age.
 Though the advantage for questions generally increased with age, the biggest developmental shift in our data occurred between one and two years, when the question effect began to reliably emerge in the two lexical conditions.
@@
 For example, we were surprised to find no main effect of linguistic condition.
-In fact, the only evidence that participants \textit{used} linguistic information came from the question effects, which were limited to the two lexical conditions.
+In fact, the primary evidence that participants actually \textit{used} linguistic information came from the question effects, which were limited to the two lexical conditions.
 Given prior work, it is surprising that taking away lexical and prosodic information didn't have a larger impact on participants' anticipatory looking.
-As in Experiment 1, most of the anticipatory switches were initiated during the inter-turn gap, indicating that participants were anticipating the start of the next response rather than the end of the ongoing response.\footnote{In contrast, the button press responses in de Ruiter et al.
-(2006) cluster before the end of the pre-gap turn.
-A button-press method explicitly encourages participants to make an early response, but recent work has also found early predictions with spontaneous neural responses \citep{bogelsmagyariInPrep, gisladottirUndRev, magyariUndRev}.
-It could be that spontaneous anticipatory switches don't elicit early enough responses to see linguistically-driven differences in prediction.}
-
-We found that the importance of lexicosyntactic and prosodic information for turn prediction differed, depending on age and transition type.
-Questions make an immediate next response relevant---more so than declaratives---and may provide a natural motivation for encouraging earlier predictive responses.
-If so, it would explain why age and language-related differences are only visible within the question effect.
-
-In addition, while even one-year-olds made anticipatory switches at an above-chance rate in the \textit{normal} and \textit{prosody only} conditions, children didn't consistently surpass chance in the \textit{words only} condition until age three.
+We take this issue up further in the general discussion.
+
+We found that the importance of lexicosyntactic and prosodic information for turn prediction differed depending on age and transition type.
+While even one-year-olds made anticipatory switches at an above-chance rate in the \textit{normal} and \textit{prosody only} conditions, children didn't consistently surpass chance in the \textit{words only} condition until age three.
 This pattern of development suggests an early importance for prosodic information in children's turn anticipation.
-But the question effect, which emerges between ages one and two, appears to heavily rely on lexical information, and doesn't appear in the \textit{prosody only} or \textit{no speech} conditions.
-So although prosody may help children early on, it doesn't affect their looking behavior in distinguishing between questions and non-questions---a driving force in their later anticipatory looking patterns.
-However, for both adults and children, the question effect was strongest when both lexicosyntactic and prosodic information were present, suggesting that prosody still has an important role to play in turn prediction.
-In other words, the lexicosyntax alone was not equivalent to the full linguistic signal in participants' anticipatory looking (even for adults), as prior work might have predicted \citep{de-ruiter2006, magyari2012}.
+In contrast, the question effect---a primary driver of anticipatory switches in our data---relies heavily on lexical information.
+So, although prosody may help children early on, it does not, on its own, help children to distinguish questions from non-questions.
+Rather, for both adults and children, prosody's role in the question effect was an additive one: The question effect was strongest when children had access to lexical \textit{and} prosodic information.
+In short, prosody has different roles to play in turn prediction before and after the lexically-driven question effect emerges.
+Importantly, these results do not support a straightforward view of lexicosyntax as the primary linguistic cue for turn predictions (even for adults) \citep{de-ruiter2006, magyari2012}.
 
@@
 Perhaps the advantage of the \textit{normal} condition was then not entirely due to its naturalness.
-We expect that future work with spontaneous conversational measures will still find an advantage for combined lexical and prosodic cues over lexical cues alone \citep[see also][]{duncan1972, ford1996, bogelsUndRev}.
+We expect that future work with spontaneous conversational measures will still find an advantage for combined lexical and prosodic cues over lexical cues alone \citep[see also][]{duncan1972, ford1996, torreiraUndRev}.
 
@@
 
-Conversational turn-taking is a time-sensitive behavior that children begin to develop long before their first words.
-But in early infancy, children have limited linguistic knowledge to help them to predict what speakers will say next when they will finish talking.
+Children begin to develop conversational turn-taking skills long before their first words.
 As they acquire language, they also acquire the information needed to make accurate predictions about upcoming turn structure.
@@
 
-% Because early conversation plays a central role in children's language learning, we can not fully understand language acquisition without taking conversational skill into account.
-Conversation is the driving force behind what children say and hear, and participants have to know the rules of the game to play effectively---especially when multiple players are involved.
-Prediction allows ready speakers to get a word in at just the right moment.
-It also helps participants to comprehend  ongoing speech and catch onto potential misunderstandings via comparison to anticipated speech (e.g., `yes' or `no' in response to a polar question).
-Language is enormously helpful in making predictions of this type, and several studies have shown that adults rely on lexicosyntactic \citep{de-ruiter2006, magyari2012} and, to some extent, prosodic information \citep{ford1996, bogelsUndRev} in their turn-taking.
-In the current study, we have tried to tie together several loose ends concerning when and how children begin to use linguistic information in their predictions about upcoming turn structure.
-
 In two experiments investigating children's anticipatory gaze to upcoming speakers, we found evidence that turn prediction develops early in childhood, and that linguistic cues are integrated differently at different ages.
-Children under 3;0 showed an advantage for prosody over lexicosyntax, contrary to the typical finding for adults in prior work \citep{de-ruiter2006} and our own data.
-Children 2;0 and older showed a slight advantage for lexicosyntax in the form of a transition type effect: children made more anticipatory switches following questions than non-questions in conditions where lexical information was present.
-% But, the question effect only significant in the \textit{normal} condition, in which children had both lexicosyntactic and prosodic information. 
-But we found no evidence that, for children or adults, lexicosyntax alone was ``sufficient'' (or otherwise equal to full linguistic information) for spontaneous turn prediction \citep[pg.
-531]{de-ruiter2006}: Children's performance was best in conditions when they had access to the full linguistic signal (including prosody).
+Children under 3;0 showed an advantage for prosody over lexicosyntax, contrary to the typical finding for adults in prior work \citep{de-ruiter2006}.
+Children 2;0 and older also showed a slight advantage for lexicosyntax in the form of a transition type effect: children made more anticipatory switches following questions than non-questions in conditions where lexical information was present.
+We found no evidence that, for children or adults, lexicosyntax alone was ``sufficient'' (or otherwise equal) compared to full linguistic information) for spontaneous turn prediction \citep[pg.
+531]{de-ruiter2006}: participants' performance was best in conditions when they had access to the full linguistic signal.
 
@@
 
-Children begin to make predictions about upcoming turn structure early in development, so our findings suggest that response planning plays a key role in explaining why they are so slow at turn-taking in real-time interaction.
-First, three- to five-year-old children in Experiment 1 were nearly as good as the adults when it came to tracking the current speaker and making anticipatory switches.
-And second, in the \textit{normal} speech condition of Experiment 2, even one-year-olds robustly made more anticipatory switches than would be expected by chance; some even showed the start of a question effect.
-Yet, as reviewed in the introduction, young children are slow turn-takers in spontaneous conversation.
-So, despite the fact that children begin taking turns in infancy and spontaneously predict upcoming turn structure at age one, it takes them a long time to develop adult-like timing in conversation with others.
-
-In both of our experiments we found a robust advantage for questions in participants' anticipatory looks to the upcoming speaker.
-This finding probably comes about because young children have a lot of experience in hearing and answering questions.
-Questions may make up approximately one third of the utterances children hear, before and after the onset of speech, and even into their preschool years \citep{fitneva2012, henning2005, shatz1979}\footnote{There is substantial variation in this proportion by individual and socioeconomic class \citep{hart1992}.}, many of which are ``test' questions early on---questions that the caregiver already has the answer to (e.g., ``What does a cat say?'').
-Caregivers use questions to get their young children's' attention and to ensure that information is in common ground, even if childrens' responses are non-verbal or infelicitous \citep{fitneva2012, snow1977}.
-
-% I didn't really kno what to make of this -MCF
-% Questions also give children a way in to ongoing interaction.
-Because they initiate a two-part structure, questions give the addressee certainty about when to talk (next) and what to say (the answer).
-Children begin to make protoimperatives between 0;9 and 1;0, using gaze, gesture, and vocalization \citep{bates1975}.
-Children's use of protoimperatives at 0;9 could suggest that they attend to two-part interactive sequences (imperative--(non-)compliance) long before real question-answer sequences emerge.
-But, protodeclaratives emerge around the same time, and may even be a more frequent speech act type in children's production before 2;8 \cite{cameron-faulkner2014}.
-If so, there may be a disparity between how children attend different speech acts in comprehension and how they use them in speech production.
-The adult participants in our studies also show the question effect, suggesting that greater anticipation for question-answers persists through adulthood (see also Tice and Henetz, 2011).
-
-What property of questions caused children to switch their gaze to the responder more often?
-We cannot definitively answer this question.
-Questions make it much more likely that a speaker switch will occur at the next possible point.
-They also have both early and late syntactic cues to help distinguish them from other speech (e.g., subject-auxiliary inversion and a final rising tone).
-But other speech acts and request formats have these properties.
-Imperatives, compliments, and complaints all make a response from the addressee likely.
-Rhetorical and tag questions, on the other hand, can take a similar form to more prototypical question types, but don't make an answer immediately relevant.
-So though it is clear that children anticipate responses more often for questions than non-questions, their actual understanding of questionhood as a linguistically- or interactionally-marked speech act still needs further exploration.
-
-A further aim of our study was to measure how linguistic information affects children's predictions about upcoming turn structure across development.
-To explore this question we limited the speech signal in our conversations to eliminate lexicosyntactic or prosodic information, either by showing participants conversations in languages they didn't understand or by phonetically manipulating the audio.
-Prior work with adults found a consistent and critical role for lexicosyntax, leaving prosody in more of a secondary and supporting role \citep{de-ruiter2006, magyari2012}.
-Knowing that children know more about prosody than lexicosyntax early on, we thought it might be possible that young children would show an advantage for prosody instead.
-Our results suggest that there are roles for both prosody and lexicosyntax, but that there are different roles for these information sources at different ages.
-
-We found an early advantage for prosodic information, compared to lexicosyntactic information, in one- and two-year-olds.
-While children's anticipatory switching didn't exceed chance for lexicosyntactic information alone until age three, it was significantly greater than chance for prosodic information alone by age one.
-The early advantage for prosody was not driven by the advantage for questions: while \textit{some} one-year-olds in our study already made more switches after questions, most did not, and prosody alone did not elicit a large question effect, even with the older children or adults.
-Having learned very little lexicosyntax, children's prosodic knowledge was probably limited to basic prosodic contours and cues to prosodic boundaries, including lengthening and pauses \textit{pannekamp2006}.
-These cues could have been enough to enable anticipatory looking for young children.
-But prosodic signals are often continuous, highly variable, and distributed throughout the turn.
-They are also hitched to the lexicosyntactic form of the utterance, making it difficult to tease apart prosody from syntax in higher-level predictive processing.
-
-Lexicosyntactic cues, on the other hand, are more often categorical cues and, in the case of questions in English, mark the speech act early in the utterance, and clearly, with ``do'', \textit{wh}-words, and subject-auxiliary inversion.
-Whereas prosodic cues to questionhood are only salient on \textit{yes/no} questions in English, lexicosyntactic markers were available on every instance of \textit{wh}- and \textit{yes/no} questions in our stimuli.
-Clear lexicosyntactic markers may have been the reason why we saw the greatest switches for questions in conditions where lexicosyntactic information was available.
-However, in both studies we have presented, the question effect was not limited by lexicosyntactic information alone.
-Older children and adults in Experiment 1 still showed a question effect in languages that they did not understand.
-In Experiment 2, the question effects were only reliable in the speech with lexicosyntactic \textit{and} prosodic information, again as a function of age.
-The results suggest that children get better at identifying questions in the ongoing conversation as they get older, with major gains between ages one and two, but that prosody still plays an important role in helping them to do so.
-
-\subsubsection{Relationship with prior work}
+The primary aim of our study was to find out how linguistic information affects children's predictions about upcoming turn structure across development.
+Two primary developmental findings emerged from our data.
+First, participants' spontaneous predictions largely occurred during question--answer transitions.
+Second, children's use of linguistic signals in making predictions depended on their age, with an early advantage for prosody over lexicosyntax and a later (partial) advantage for lexicosyntax over prosody.
+
+\subsubsection{The advantage for questions}
+
+In both of our experiments we found a robust advantage for questions in participants' anticipatory switching.
+Children and adults both treated questions as if they had a special interactive status, and with good reason: Questions make a speaker switch immediately relevant, helping the listener to predict what will happen next (i.e., an answer from the addressee).
+Linguistic cues to questionhood often occur early in the utterance, making questions easy to identify early on and thereby giving observers time to plan a reaction (i.e., a gaze switch).
+Also, the form of a question can constrain the type of response that will come next (e.g., a location after a \textit{where} question), further increasing the predictability of the interactional sequence.
+
+Children get frequent and early experience with hearing and answering questions.
+Questions make up approximately one third of the utterances children hear, before and after the onset of speech, and even into their preschool years \citep{fitneva2012, henning2005, shatz1979}.\footnote{There is substantial variation in this proportion by individual and socioeconomic class \citep{hart1992}.} Many of the questions directed to young children are ``test' questions---questions that the caregiver already has the answer to (e.g., ``What does a cat say?'').
+Caregivers use questions to get their young children's attention and to ensure that information is in common ground, even if the responses are non-verbal or infelicitous \citep{fitneva2012, snow1977}.
+So, in addition to having a special interactive status, questions are a core characteristic of many caregiver-child interactions.
+
+We cannot definitively say with our current data what property of questions caused children to switch their gaze more often.
+It is true that questions make a speaker transition more likely and that questions have distinctive linguistic cues, both early and late, to identify themselves.
+But other speech acts and request formats have these same properties.
+Imperatives, compliments, and complaints all make a response from the addressee likely \citep{schegloff2007}.
+Rhetorical and tag questions, on the other hand, take a similar form to prototypical polar questions, but do not always require an answer.
+So, though it is clear that children anticipate responses more often for questions than non-questions, their actual understanding of questionhood as a linguistically- or interactionally-marked type of speech act still needs further exploration.
+
+\subsubsection{Linguistic cues for turn prediction}
+
+Prior work with adults has found a consistent and critical role for lexicosyntax in predicting upcoming turn structure, with a more secondary and supporting role for prosody \citep{de-ruiter2006, magyari2012}.
+Knowing that children comprehend more about prosody than lexicosyntax early on, we thought it possible that young children would instead show an advantage for prosody over lexicosyntax.
+Our results suggest that there are roles for both prosody and lexicosyntax in children's predictions about upcoming turn structure, but that there are different roles for these information sources at different ages.
+
+Early on, there was an advantage for prosody over lexicosyntax in children's anticipatory switches: Anticipatory switching didn't reliably exceed chance for lexicosyntactic information  until age three, but was already significantly greater than chance for prosodic information  at age one.
+This early advantage for prosody was not driven by the advantage for questions: While some one-year-olds in our study already made more switches after questions than non-questions, most did not, plus prosody alone did not elicit a robust question effect at any age (Figure \ref{fig:questionsEN}).
+
+Although children showed an initial advantage for prosody, their predictions before 3;0 were probably limited to basic prosodic contours and boundaries, and simple cues such as final lengthening and turn-final silence \citep{pannekamp2006}.
+More complex structural predictions would have required them to access lexicosyntactic information because higher-level structure in the prosodic signal is linked to the syntactic structure being used.
+The close link between prosodic and syntactic structure makes it difficult to tease apart how predictive processing in one domain (e.g., prosody) is distinct from the other (e.g., syntax) in turn prediction.
+We leave this challenge to future work.
+
+Children only showed an advantage for questions in the two lexical conditions, suggesting that lexicosyntax is a critical cue for question identification.
+The question effect emerged by age two, when children's syntactic comprehension and lexical inventories were still relatively limited \citep{clark2009}.
+That being said, syntactic and lexical markers to questionhood in English are frequent, categorical, and early-occurring in the utterance (e.g., ``do'', \textit{wh}-words, and subject-auxiliary inversion).
+Lexicosyntactic question cues were available on every instance of \textit{wh}- and \textit{yes/no} questions in our stimuli, whereas prosodic question cues were only salient on \textit{yes/no} questions.
+Thus, the unambiguous presence of lexicosyntactic cues on \textit{wh}- and \textit{yes/no} questions may have been one reason why we saw an advantage for questions in the lexical conditions, but not in the \textit{prosody only} condition.
+
+Importantly, the question effect was not driven purely by lexicosyntactic information.
+In both experiments, prosody still made a contribution to the advantage for questions, even though prosody alone was not enough to elicit a question effect.
+For example, older children and adults in Experiment 1 showed a question effect in the languages they did not understand---presumably they used prosodic and non-verbal information to distinguish questions from non-questions.
+In Experiment 2, the advantage for questions was biggest for speech with lexicosyntactic \textit{and} prosodic information together.
+All in all, children appear to improve at identifying questions in ongoing conversation by using lexicosyntactic cues as they get older.
+But, prosodic cues still aid in picking out questions from ongoing conversation.
+
+\subsection{Relationship with prior work}
 
@@
 De Ruiter and colleagues (2006) found that adults' accuracy in identifying upcoming turn-ends was not significantly affected when intonation was taken away.
-In their study of children and adults, Keitel and colleagues (2013) found the same for adults' anticipatory gaze switches, and only finding that intonation affected children's switches at 3;0 (with switches by younger children never exceeding chance).
-Ou results, in contrast, suggest that children's anticipatory switches emerge much earlier, and rely on prosody before they do lexicosyntax.
-Further, our adult control participants patterned similarly to the children: they showed question effects for languages they didn't speak in Experiment 1 and strongest question effects for lexicosyntactic information when it was combined with prosody in Experiment 2.
-Thus, neither the child nor the adult data support the idea that lexicosyntactic information is equivalent to full linguistic information in making predictions about upcoming turn structure \citep{de-ruiter2006}, at least in participants' spontaneous looking behavior.
-
-Differences in experimental paradigm could begin to explain the disparity in our findings.
-The anticipatory switches participants make in our studies were spontaneous, and were made without a requirement for urgency or timeliness, as the responses are for a button-press measure \citep{de-ruiter2006, magyari2012}.
-Most of our participants initiated their anticipatory switches in the inter-turn gap, consistent with prior observer gaze measures (Keitel et al., 2013; Hirvenkari, 2013; Tice and Henetz, \citeyear{TiceHenetz11}).
-The timing of participants' anticipatory switches suggests that, while they watch a conversation, they move their eyes in anticipation of the upcoming turn (not the end of the ongoing turn).
-Furthermore, the fact that participants made most of their anticipatory switches following questions, compared to non-questions, suggests that observers aren't just targeting potential turn-ends with their predictions---they are targeting speaker switches.
-Although button press measures may be better able to tell us what participants \textit{can} do in predicting turn ends, they may also emphasize lexicosyntactic cues to a greater extent than we experience in normal conversation.
-
-% If participants are primarily monitoring for speaker switch instead of turn ends in general, we need to more carefully consider how signals to impending speaker transition (verbal and non-verbal) are weighed against more generic signals to utterance structure in our cognitive-linguistic model of how people manage to take turns on time.
-Current models of adult turn-taking underline the role of linguistic prediction, but more recent work, including our own, may point some future work toward more shallow cue monitoring (e.g., for \textit{wh-}words) and inhibition \citep{bogelsUndRev}.
+Keitel and colleagues (2013) found the same for adults' anticipatory gaze switches, only finding that intonation affected children's switches at age 3;0 (with switches by younger children never exceeding chance).
+Our results, in contrast, suggest that children's anticipatory switches emerge much earlier, and rely on prosody before they do lexicosyntax.
+Furthermore, our adult control participants patterned similarly to the children: They showed question effects for languages they didn't speak (Experiment 1) and showed the strongest question effects for lexicosyntax and prosody together (Experiment 2).
+Thus, neither the child nor the adult data support the idea that lexicosyntactic information is equal to full linguistic information in making predictions about upcoming turn structure \citep{de-ruiter2006}, at least in participants' spontaneous looking behavior.
+
+Despite extreme differences in linguistic information across conditions, main effects of linguistic information were minimal in our experiments.
+But, using a button-press measure with similar stimuli, de Ruiter and colleagues (2006) found robust effects of linguistic condition.
+This disparity in findings could be partially explained by task differences.
+A button-press method explicitly encourages participants to make an early response, while observer gaze measures are purely passive.
+Participants' spontaneous anticipatory gaze switches may not elicit enough early responses to see significant linguistically-driven differences in prediction.
+Indeed, most switches happened during the inter-turn gap (Keitel et al., 2013; Hirvenkari, 2013; Tice and Henetz, \citeyear{TiceHenetz11}).
+Observers' gaze patterns suggest that they aren't targeting potential turn-ends with their predictions.
+Rather, they are targeting switches in speakership.
+So although button press measures may be better equipped to tell us what participants \textit{can} do in predicting turn ends, they may also emphasize early responding and, consequently, lexicosyntactic cues to a greater extent than we experience in everyday conversation.
+Within observer gaze experiments, questions may be the key to naturally elicit early responses.
+We have seen in the current study that questions are interactionally sufficient to motivate early predictions.
+
+Moving away from button press measures, Keitel and colleagues (2013) used observer gaze to investigate children's turn predictions, finding: (a) no evidence that children under 3;0 make switches above chance, (b) that three-year-olds' switches are significantly affected by flattened intonation, and (c) that adults' switches are not.
+Our current results are consistent with the Keitel et al.
+(2013) study with respect to the general finding that prosody matters for young children's turn predictions.
+But, our results also differ from theirs in a few key ways.
+
+First, we found that children begin making above-chance anticipatory switches much earlier than 3;0---our one-year-olds were already looking above chance in two of our four linguistic conditions in Experiment 2.
+This earlier emergence of anticipatory switching could be due to our more child-friendly stimuli, which we designed to be engaging and to mimic the kind of speech children often experience interactionally.
+Second, because we included more linguistic conditions to isolate the effects of prosody and lexicosyntax, we gained a broader perspective on how linguistic information changed children's predictions.
+Ultimately, the added conditions allowed us to find an early advantage for prosody and a later advantage for lexicosyntax.
+
+Third, we also found that anticipatory looking was primarily driven by question transitions, a pattern that has so far not been discussed in other observer gaze studies, on children or adults (Keitel et al., 2013; Hirvenkari, 2013; Tice and Henetz, \citeyear{TiceHenetz11}).
+The question effects in the current study were critical to understanding the role of lexicosyntactic information in participants' turn predictions.
+Finally, we found that our adult controls were affected by the absence of prosody (within the question effect) while Keitel et al.
+(2013) found that adults' were unaffected by flattened pitch.
+But, because we flattened pitch for individual speakers and controlled syllable duration, we cannot directly compare our findings to theirs.
+
+Our findings are consistent with observational work on children's turn taking.
+Young children are slow turn-takers in spontaneous conversation \citep{casillasUndRev} even though they begin taking turns in infancy \citep{hilbrinkInPrep, jaffe2001}.
+We found that children begin to spontaneously predict upcoming turn structure at age one.
+Children's gaze switches were also most common in the inter-turn gap, suggesting that they expected an immediate response from the addressee.
+Our results therefore support the idea of an early-developed competence for the timing of turn-taking.
+But because children do not themselves use adult-like conversational timing until much later \citep[age six; ][]{ervin-tripp1979}, our results also indirectly point toward response planning as a key factor contributing to children's slow responses in real-time interaction \citep{casillasUndRev}.
 
@@
 In Experiment 2, we tried to improve our method for controlling linguistic information by phonetically manipulating improvised dialogic speech.
-Although this decision allowed us to follow prior work more closely \cite{de-ruiter2006, keitel2013}, it also limits our interpretation of children's behavior.
-Low-pass filtered and robotic-like speech are not typical speech signals children hear, and so the limited cue conditions in Study 2 may have been a somewhat unfair comparison to the normal speech condition, which not only had full linguistic cues, but also was more typical for what children hear in their everyday life.
-An better alternative to explore for the future would be scripted speech, or speech carefully selected from a large corpus for specific linguistic properties.
+Although this decision allowed us to follow prior work more closely \citep{de-ruiter2006, keitel2013}, it also limits our interpretation of children's behavior.
+Low-pass filtered and robotic-like speech are not typical within children's speech environment, and so it may have been unfair to compare the limited cue conditions in Experiment 2 to the normal speech condition, which not only had full linguistic cues, but was also more typically present in children's speech environments.
+One alternative to explore for the future would be scripted speech, or speech that is carefully selected from a large corpus, given a set of specific linguistic constraints.
 
@@
 It is possible that, to some extent, children could have formed expectations about the turn-structure on the basis of the frequent switching alone \citep[see, e.g., ][]{thorgrimssonUndRev}.
-If this statistical expectation were the only thing leading children's anticipatory gaze switches, we would have predicted more uniformity in looking behavior across the conditions and no major effects of linguistic signal, age, or transition type, however, which we did not see.
+If this statistical expectation were the only thing leading children's anticipatory gaze switches, we would have predicted more uniformity in looking behavior across the conditions and no major effects of linguistic signal, age, or transition type.
+But that is now what we saw.
 Nevertheless, another advantage of using scripted conversation would be the ability to control switch frequency and eliminate this issue.
@@
 
-FILL ME IN
+Conversation plays a central role in children's language learning.
+It is the driving force behind what children say and what they hear.
+And yet many conversational skills, like turn taking, rely heavily on language.
+Adults use language to accurately predict turn structure in conversation, which facilitates their online comprehension and allows them to respond relevantly and on time.
+But young children do not have enough linguistic knowledge to take turns the way that adults do.
+Instead, they first acquire a preverbal turn-taking system for interaction, and then somehow assimilate language into that system as they acquire it.
+The simultaneous acquisition and integration of interactive and linguistic skills is no easy task, as evidenced by children's protracted development of turn-taking skills.
+In the current study we have investigated children's language acquisition through the lens of turn-taking, thereby bringing into focus children's application of linguistic knowledge in real-time interaction.
+We found that children acquire basic predictive skills early on, but that the integration of linguistic cues takes time and is primarily driven by sequences of action---in our case, by questions and their answers.
+Thus, even at the level of linguistic integration, children's attention during conversation is driven by goals from real-time interaction, the principle source and destination for their language learning.
+
+\section*{Acknowledgements}
+
+Thanks! (To be filled in, but must include NSF GRFP, ERC, Bing, and the CDM).
 

### 2014-09-19 analysis/E1/3-plots.stats.analysis.R

@@ -469,7 +469,7 @@ limits <- aes(ymax = m + se, ymin= m - se)
 
 p2 <- qplot(Age,m, group=Condition, ylim=c(-0.1,0.5),
     ymin= m - se, ymax= m + se, color=Condition,
-    xlab="Age (years)", ylab="Proportion gaze switches",
+    xlab="Age (years)", ylab="Baseline-corrected proportion switches",
     position=position_dodge(width=.1), data=df) +
     geom_line(size=2, position=position_dodge(width=.1)) +
     geom_pointrange(size=1, position=position_dodge(width=.1)) +
@@ -492,7 +492,7 @@ dev.off()
 
 p2bars <- qplot(Age,m, group=Condition, ylim=c(-0.1,0.5),
     ymin= m - se, ymax= m + se, fill=Condition,
-    xlab="Age (years)", ylab="Proportion gaze switches",
+    xlab="Age (years)", ylab="Baseline-corrected proportion switches",
     geom="bar", stat="identity", position=position_dodge(), data=df) +
     geom_errorbar(limits, position=position_dodge(width=0.9), width=0.25) +
     scale_fill_manual(name = "", values=stoplightPalette, labels = c("English  ", "Non-English")) +
@@ -592,7 +592,7 @@ limits <- aes(ymax = m + se, ymin= m - se)
 
 p4 <- qplot(Age,m,facets = . ~ Condition, group=Transition, ylim=c(-0.1,0.5),
     ymin= m - se, ymax= m + se, color=Condition, linetype=Transition,
-    xlab="Age (years)",ylab="Proportion gaze switches",
+    xlab="Age (years)",ylab="Baseline-corrected proportion switches",
     position=position_dodge(width=.1), data=df) +
     geom_line(size=2, position=position_dodge(width=.1)) +
     geom_pointrange(size=1, position=position_dodge(width=.1)) +

### 2014-09-19 analysis/E2/3-plots.stats.analysis.R

@@ -457,7 +457,7 @@ limits <- aes(ymax = m + se, ymin= m - se)
 
 p2 <- qplot(Age,m, group=Condition, ylim=c(-0.1,0.6),
     ymin= m - se, ymax= m + se, color=Condition,
-    xlab="Age (years)", ylab="Proportion gaze switches",
+    xlab="Age (years)", ylab="Baseline-corrected proportion switches",
     position=position_dodge(width=.1), data=df) +
 	geom_line(size=2, position=position_dodge(width=.1)) +
 	geom_pointrange(size=1, position=position_dodge(width=.1)) +
@@ -478,7 +478,7 @@ dev.off()
 
 p2bars <- qplot(Age,m, group=Condition, ylim=c(-0.1,0.6),
     ymin= m - se, ymax= m + se, fill=Condition,
-    xlab="Age (years)", ylab="Proportion gaze switches",
+    xlab="Age (years)", ylab="Baseline-corrected proportion switches",
     geom="bar", stat="identity", position=position_dodge(), data=df) +
     geom_errorbar(limits, position=position_dodge(width=0.9), width=0.25) +
     scale_fill_manual(name="", values=stoplightPalette) +
@@ -589,7 +589,7 @@ limits <- aes(ymax = m + se, ymin= m - se)
 
 p4 <- qplot(Age,m,facets = . ~ Condition, group=Transition,
     ylim=c(-0.1,0.6), ymin= m - se, ymax= m + se, color=Condition,
-    linetype=Transition, xlab="Age (years)",ylab="Proportion gaze switches",
+    linetype=Transition, xlab="Age (years)",ylab="Baseline-corrected proportion switches",
     position=position_dodge(width=.1), data=df) +
 	geom_line(size=2, position=position_dodge(width=.1)) +
 	geom_pointrange(size=1, position=position_dodge(width=.1)) +

### 2014-09-22 paper/CasillasFrank-Anticip.tex
minor tweaks

@@
 We investigate language acquisition through the lens of a fundamental conversational skill: turn taking.
-Children begin developing turn-taking skills in infancy, but take several years to assimilate language into their turn-taking systems.
-In two eye-tracking experiments with children ages 1--6, we measured participants' spontaneous predictions about upcoming speaker change while controlling the amount of linguistic information available.
-We found that children already predicted upcoming turns at age one, and integrated linguistic cues differently at different ages.
-Children under 3;0 showed a general advantage for prosody over lexicosyntax, contrary to the typical finding for adults.
-Children 2;0 and older showed more predictive switches for questions than non-questions, but only when lexical information was available.
-But, we found no evidence that lexicosyntax is sufficient for turn prediction---participants' performance was best overall with access to lexicosyntax and prosody together.
-Our results point to the importance of interaction in children’s real-time linguistic processing.
+Children begin developing turn-taking skills in infancy, but take several years to assimilate their growing knowledge of language with this ability.
+In two eye-tracking experiments with children across a wide developmental range, we measured spontaneous predictions about upcoming speaker change while controlling the amount of linguistic information available.
+We found that children already predicted upcoming turns at age one, but they integrated linguistic cues differently at different ages.
+Children under three showed a general advantage for prosody over lexicosyntax, contrary to the typical finding for adults.
+Children two and older showed more predictive switches for questions than non-questions, but only when lexical information was available.
+We found no evidence that lexicosyntax alone guides turn prediction---instead, participants' performance was best overall with access to lexicosyntax and prosody together.
+Our results point to the importance of interaction in children's real-time linguistic processing.
 \end{abstract}
@@
 They found that participants never verbally responded or pressed the ``turn-end'' button when hearing a syntactically-complete unit without a boundary tone.
-However, utterance-level prosodic boundaries embedded in multi-utterance turns \textit{could} trick participants into anticipating a turn end 29\% of the time (see also de \citet{de-ruiter2006}:525).
+Utterance-level prosodic boundaries embedded in multi-utterance turns \textit{could} trick participants into anticipating a turn end 29\% of the time, however (see also de \citet{de-ruiter2006}:525).
 Their findings corroborate other corpus and experimental findings promoting a combination of cues (lexicosyntactic, prosodic, and pragmatic) as key for accurate turn-end prediction \citep{duncan1972, ford1996, hirvenkari2013}.
@@
 One study found that 12-month-olds make more predictive gaze shifts while watching human verbal conversation compared to object non-speech ``conversation'' \citep{bakker2011}, but another only found a similar effect at 36 months \citep{hofsten2009}.
-However, neither of these two studies had baselines to which the turn-relevant looking behavior could be compared.
+Neither of these two studies had baselines to which the turn-relevant looking behavior could be compared, however.
 A baseline measurement is critical because there may be developmental differences in gaze shifting between conversational participants, even if the shifting is not related to turn structure.
@@
 Because these limitations directly inform our own study design, we review them in some detail.
-First, their estimates of baseline ``random'' gaze frequency were computed from switches that occurred while the speaker's turn was ongoing (i.e, during turns, and not during inter-turn gaps).
-Their estimates of anticipatory shifting frequency were computed from switches that occurred within the time window immediately surrounding inter-turn gaps (i.e., around inter-turn gaps, and primarily not during turns).
-Assuming that gaze switches happen most often during the inter-turn gap \citep{hirvenkari2013}, switches made during ongoing speech (their ``random'' baseline frequency) would be at a minimum.
-Therefore their ``random'' (baseline) frequency actually maximized the possibility of finding a difference between ``random'' and anticipatory gaze frequencies.
+First, their estimates of baseline gaze frequency (``random'' in their terminology) were computed from switches that occurred specifically while the speaker's turn was ongoing; in contrast, estimates of anticipatory shifting frequency were computed from switches that occurred within the time window immediately surrounding inter-turn gaps.
+Assuming that gaze switches happen most often during the inter-turn gap \citep{hirvenkari2013}, switches made during ongoing speech (their baseline) would be minimized.
+This baseline thus maximized the possibility of finding a difference between ``random'' and anticipatory gaze frequencies.
 A more conservative baseline would be to compare participants' looking behavior at turn transitions to their looking behavior during randomly-selected windows of time throughout the stimulus.
-We follow this latter approach in our work.
+We follow this conservative approach in our work.
 
@@
 
-Children showed increased anticipatory switching with age across all conditions, making almost as many transitions as adults ages five and six.
+Children showed increased anticipatory switching with age across all conditions, making almost as many transitions as adults by ages five and six.
 Adults reacted to the \textit{no speech} condition differently than the children did, tracking the interaction less closely than in the three linguistic conditions.
@@
 The core aims of Study 2 were to gain better traction on the individual roles of prosody and lexicosyntax in children's turn predictions, and to expand our age range to capture more developmental change.
-We found that effects of linguistic processing and age primarily surface within the question effect: Children and adults make more anticipatory switches after questions than non-questions, but do so more often when they have greater access to linguistic information, whether increased access came with age or with linguistic condition.
+We found that effects of linguistic processing and age were present in this dataset, but both were primarily found within the question effect.
+Children and adults made more anticipatory switches after questions than non-questions, but did so more often when they had greater access to linguistic information, whether increased access came with age or with linguistic condition.
 Our results hint that lexical information is critical in identifying questions during conversation, especially for children.
@@
 By extending the age range of participants, we were able to see more of the developmental trajectory in children's predictions about turn structure.
-We saw that one-year-olds already made more anticipatory switches than would be expected by chance in two of the four conditions.
+One-year-olds already made more anticipatory switches than would be expected by chance in two of the four conditions.
 As they got older, children demonstrated above-chance anticipatory switches in more conditions and made more anticipatory switches overall.
@@
 In two experiments investigating children's anticipatory gaze to upcoming speakers, we found evidence that turn prediction develops early in childhood, and that linguistic cues are integrated differently at different ages.
-Children under 3;0 showed an advantage for prosody over lexicosyntax, contrary to the typical finding for adults in prior work \citep{de-ruiter2006}.
-Children 2;0 and older also showed a slight advantage for lexicosyntax in the form of a transition type effect: children made more anticipatory switches following questions than non-questions in conditions where lexical information was present.
-We found no evidence that, for children or adults, lexicosyntax alone was ``sufficient'' (or otherwise equal) compared to full linguistic information) for spontaneous turn prediction \citep[pg.
-531]{de-ruiter2006}: participants' performance was best in conditions when they had access to the full linguistic signal.
+Children under three showed an advantage for prosody over lexicosyntax, contrary to the typical finding for adults in prior work \citep{de-ruiter2006}.
+Children two and older also showed a slight advantage for lexicosyntax in the form of a transition type effect: children made more anticipatory switches following questions than non-questions in conditions where lexical information was present.
+We found no evidence that, for children or adults, lexicosyntax alone was ``sufficient'' (equal to full linguistic information) for spontaneous turn prediction (cf.
+\citealp{de-ruiter2006}): Participants' performance was best in conditions when they had access to the full linguistic signal.
 
@@
 
-Although children showed an initial advantage for prosody, their predictions before 3;0 were probably limited to basic prosodic contours and boundaries, and simple cues such as final lengthening and turn-final silence \citep{pannekamp2006}.
+Although children showed an initial advantage for prosody, their predictions before age three were probably limited to basic prosodic contours and boundaries, and simple cues such as final lengthening and turn-final silence \citep{pannekamp2006}.
 More complex structural predictions would have required them to access lexicosyntactic information because higher-level structure in the prosodic signal is linked to the syntactic structure being used.
@@
 
-Thanks! (To be filled in, but must include NSF GRFP, ERC, Bing, and the CDM).
+We gratefully acknowledge the parents and children at Bing Nursery School, Children's Discovery Museum of San Jose.
+This work was supported by an ERC grant XYZ, an NSF GRFP to the first author, and a Merck Foundation Fellowship to the second author.
+An earlier version of these data and analyses were initially presented to conference audiences in \cite{casillas2012} and \cite{casillas2013}.
 

### 2014-09-29 paper/CasillasFrank-Anticip.tex
Merging in MCF's edits

@@
 Sequences of action drive the conversation forward into the next, relevant sequences of talk (e.g., ''And you?''--``Great!''--``Why's that?''; \citealp{schegloff2007}).
-To take a turn, participants need to make predictions about what types of talk will be relevant next.
+To take a turn, participants need to make predictions about what conversational content will be relevant next.
 In some cases, relevant next turns are somewhat obvious (e.g., question--response) while, in other cases, there are multiple relevant next actions to choose from, or no obvious next action at all (e.g., after a closed sequence).
 
-Despite this complexity, conversational turn-taking is precise in its timing.
+Despite this complexity, conversational turn-taking is often precise in its timing.
 Across a diverse sample of conversations in 10 languages, one study found a consistent average transition time of 0--200 msec at points of speaker switch \citep{stivers2009}.
-Current models of speech production suggest that it takes approximately 600 msec to produce a word, and even longer to produce a simple utterance \citep{levelt1989}.
+Experimental results and current models of speech production suggest that it takes approximately 600 msec to produce a content word, and even longer to produce a simple utterance \citep{griffin2000, levelt1989}.
 So in order to achieve 200 msec turn transitions, speakers must begin formulating their response before the prior turn has ended.
-To formulate their response early on, speakers must track and anticipate what types of response might become relevant next.
+Moreover, to formulate their response early on, speakers must track and anticipate what types of response might become relevant next.
 They also need to predict the content and form of upcoming speech so that they can launch their articulation at exactly the right moment.
-Prediction thus plays a key role in timely turn-taking.
+Prediction must thus play a key role in timely turn-taking.
 
 \subsection{Adults' turn prediction}
-
-Adults have a lot of information that could potentially help with their predictions about upcoming turn structure.
+% FIX: compress deR: make it clear about prosody issues, perhaps talk more clearly about torreira
+Adults have a lot of information at their disposal that could potentially help with their predictions about upcoming turn structure.
 Lexical, syntactic, and prosodic information (e.g., \textit{wh}- words, subject-auxiliary inversion, list intonation) can all inform addressees about upcoming linguistic structure \citep{de-ruiter2006, duncan1972, ford1996, torreiraUndRev}.
@@
 In another study, \citet*{torreiraUndRev} used a combination of button-press and verbal responses to investigate the relationship between lexicosyntactic and prosodic cues in turn-boundary identification.
-They found that participants never verbally responded or pressed the ``turn-end'' button when hearing a syntactically-complete unit without a boundary tone.
-Utterance-level prosodic boundaries embedded in multi-utterance turns \textit{could} trick participants into anticipating a turn end 29\% of the time, however (see also de \citet{de-ruiter2006}:525).
+They found that participants never verbally responded or pressed the ``turn-end'' button when hearing a syntactically-complete unit without an intonational phrase boundary.
+However, utterance-level prosodic boundaries embedded in multi-utterance turns could trick participants into anticipating a turn end 29\% of the time (see also de \citet{de-ruiter2006}:525).
 Their findings corroborate other corpus and experimental findings promoting a combination of cues (lexicosyntactic, prosodic, and pragmatic) as key for accurate turn-end prediction \citep{duncan1972, ford1996, hirvenkari2013}.
@@
 Their predictive processing relies on a sophisticated body of knowledge about linguistic structure, non-verbal signals, and social actions.
-The acquisition of turn-taking skills must then be tied to children's knowledge about language, gaze, gesture, and social cues.
+The later acquisition of verbal turn-taking skills must then be tied to children's knowledge about language, gaze, gesture, and social cues.
 But children's turn taking starts early in infancy, long before first words or gestures emerge; a primary role for lexicosyntactic cues doesn't fit well with children's pre-verbal turn-taking.
@@
 The timing of children's responses to their caregivers' speech shows a non-linear pattern of fall-rise-fall from early infancy to middle childhood.
-Infants' turn timing at three months is often too early or too late: they start vocalizing in overlap on nearly 40\% of their caregivers' turns, and their non-overlapped vocalizations come after an average latency of 675 msec (adult average: 200 msec).
+A recent study by Hilbrink et al.
+(in preparation) finds that infants' turn timing at three months is often too early or too late: they start vocalizing in overlap on nearly 40\% of their caregivers' turns, and their non-overlapped vocalizations come after an average latency of 675 msec (adult average: 200 msec).
 Between four and nine months, children begin to reduce the number of turns happening overlap while also improving on their average response latency.
@@
 
-Keitel and colleagues \citeyearpar{keitel2013} addressed the random baseline issue in their study of 6-, 12-, 24-, and 36-month-olds, finding that children's anticipatory gaze frequency was only greater than chance for 36-month-olds and adults.
+Keitel and colleagues \citeyearpar{keitel2013} addressed the random baseline issue in their study of 6-, 12-, 24-, and 36-month-olds.
+They asked participants to watch short videos of conversation, and tracked their eye movements at points of speaker change.
+They found that children's anticipatory gaze frequency was only greater than chance for 36-month-olds and adults.
 Their study was the first to focus on the role of linguistic processing in children's turn predictions.
-They showed their participants two types of conversation videos: one normal and one with flattened pitch (i.e., with removed intonation contours), finding that only 36-month-olds were affected by a lack of intonation.
+They showed their participants two types of conversation videos: one normal and one with flattened pitch (i.e., with flattened intonation contours), finding that only 36-month-olds were affected by a lack of intonation contours.
 The adult control group made equal numbers of anticipatory looks in the videos, with and without intonation contours, consistent with prior adult findings \citep{de-ruiter2006}.
@@
 The average gap between turns was 900 msec, which is much longer than typical adult timing, where gaps average around 200 msec \citep{stivers2009}.
-The speakers in the videos were also asked to minimize their movements while performing a scripted, adult-directed conversation, which would have created a somewhat unnatural stimulus.
+The speakers in the videos were also asked to minimize their movements while performing a scripted and adult-directed conversation, which would have created a somewhat unnatural stimulus.
 In addition, in order to produce more naturalistic conversation, it would have been ideal to localize sources for the two voices in the video (i.e., to have the voices come out of separate left and right speakers).
@@
 \subsection{The Current Study}
-
+% add bit about children's prosodic development and link this to the plans for the experiments
+% mention that we will be looking at anticipatory gaze switching
 We report here on the role of linguistic processing in children's predictions about upcoming turn structure.
@@
 The participants, all native English speakers, were only expected to understand the two videos in English.
-We showed participants non-English videos to limit their access to lexical information while maintaining their access to other cues to turn boundaries (e.g., prosody, gaze, breath, phrase final lengthening).
+We showed participants non-English videos to limit their access to lexical information while maintaining their access to other cues to turn boundaries (e.g., (non-native) prosody, gaze, breath, phrase final lengthening).
 Using this method, we compared children and adult's anticipatory looks from the current speaker to the upcoming speaker at points of turn transition in English and non-English videos.
@@
 Each talker was a native speaker of the language being recorded, and each talker pair was male-female.
-Using a Marantz PMD 660 solid state field recorder, we captured audio from two lapel microphones, one attached to each participant, while simultaneously recording video from the built-in camera of a MacBook.
+Using a Marantz PMD 660 solid state field recorder, we captured audio from two lapel microphones, one attached to each participant, while simultaneously recording video from the built-in camera of a MacBook laptop computer.
 The talkers were volunteers, all members of the local post-graduate community, and were acquainted with their recording partner ahead of time.
@@
 The non-English stimuli were coded for speech act from a monolingual English-speaker's perspective, i.e., which turns ``sound like'' questions, and which don't.
-The coding for speech act in non-English videos was completed independently by two coders and then checked and discussed for agreement.
-
-Because the conversational stimuli were recorded semi-spontaneously, the duration turn transitions and the number of speaker transitions in each video was variable.
+
+Because the conversational stimuli were recorded semi-spontaneously, the duration of turn transitions and the number of speaker transitions in each video was variable.
 We measured the duration of each turn transition from the audio recording associated with each video.
@@
 
+% check over again to make as clear as possible
 Using these criteria, we created a three-step algorithm for detecting anticipatory gaze switches.
@@
 
+% check over again to make as clear as possible
 \paragraph{Random baseline analysis} To ensure that participants' looking behavior was turn-related and not simply the result of random switching, we corrected for the probability of making an anticipatory switch at random.
@@
 We identified anticipatory gaze switches for all 43 usable turn transitions, based on the criteria and algorithm outlined in Section \ref{sec:algorithm}.
-Because participants had greater access to lexical and syntactic information in English, we expected to see greater anticipation in the English videos than in the non-English videos.
+Because participants had greater access to linguistic information in English, we expected to see greater anticipation in the English videos than in the non-English videos.
 We also predicted that anticipation would be greater following questions compared to non-questions; questions have early cues to upcoming turn transition (e.g., \textit{wh-} words, subject-auxiliary inversion), and also make a next response immediately relevant.
 We also predicted that anticipatory looks would increase with development, along with children's increased linguistic competence.
-
-To estimate the effects of linguistic condition and age on switching behavior, we corrected each participant's data by subtracting their random baseline estimations from their actual switching data for each turn transition.
-All comparisons below are made with the baseline-corrected values.
 
@@
 
+To estimate the effects of linguistic condition and age on switching behavior, we estimated each participants' random baseline switching behavior for each turn transition.
 Participants at all ages made anticipatory gaze switches more often than would be expected by chance, whether they could understand the language in the video or not (Figure \ref{fig:randvsrealFL}).
-Anticipatory switch behavior in the non-English videos could not have been based on lexicosyntactic information, so both adults and children must have relied on other sources of information to make their predictions (e.g., prosody and non-verbal signals).
+Anticipatory switch behavior in the non-English videos could not have been based on lexicosyntactic information, so both adults and children must have relied on other sources of information to make their predictions (e.g., non-native prosodic cues and non-verbal signals).
+We then corrected each participant's data by subtracting their random baseline estimations from their actual switching data for each turn transition.
+All comparisons below are made with the baseline-corrected values.
 
@@
 
-Participants also showed an advantage for questions.
+Participants also showed an advantage for questions (Figure \ref{fig:questionsFL}).
 Children made anticipatory switches twice as often when they heard a question, compared to a non-question (18.7\% and 7.3\%).
 But the advantage for questions was not universal---it depended on language condition and age.
-The advantage for questions didn't emerge until age five in the non-English videos (Figure \ref{fig:questionsFL}).
+The advantage for questions didn't emerge until age five in the non-English videos .
 
@@
 
-We quantified these findings using two linear mixed-effects models of baseline-corrected anticipatory switches \citep{lme4, R}.
+We quantified these findings using two linear mixed-effects models of participants' baseline-corrected anticipatory switches \citep{lme4, R}.
 We built one model each for children and adults.
@@
 
-Children's anticipatory gaze switches showed a main effect of language condition (\textit{$\beta$}=-0.448, \textit{SE}=0.126, \textit{t}=-3.56, \textit{p}$<$.001).\footnote{We derive \textit{p}-values here by treating \textit{t} and \textit{z} (with a two-tailed assumption), as is justified for large data sets \citep{barr2013}.} As predicted, children made more anticipatory looks while they were watching videos that they could understand.
+Children's anticipatory gaze switches showed a main effect of language condition (\textit{$\beta$}=-0.448, \textit{SE}=0.126, \textit{t}=-3.56, \textit{p}$<$.001).\footnote{We derive \textit{p}-values here by treating \textit{t} as \textit{z} (with a two-tailed assumption), as is justified for large data sets \citep{barr2013}.} As predicted, children made more anticipatory looks while they were watching videos that they could understand.
 In the English videos, children had access to lexicosyntactic information, which could have helped them make better predictions about upcoming turn structure.
@@
 Again, to test this idea, we would need to directly compare lexicosyntactic and prosodic cues in the participants' native language.
-In sum, the current study lays the analytic groundwork for a method that allows for greater experimental control, which we introduce in Experiment 2.
+In sum, Experiment 1 lays the analytic groundwork for a method that allows for greater experimental control, which we introduce in Experiment 2.
 
@@
 All of the videos in Experiment 2 were in the participants' native language (American English).
-To tease apart the role of lexical and prosodic information, we phonetically manipulated the speech signal for pitch, rhythm, and lexical access.
+To tease apart the role of lexical and prosodic information, we phonetically manipulated the speech signal for pitch, syllable duration, and lexical access.
 By testing one- to six-year-olds we hoped to find the developmental onset of turn-predictive gaze.
@@
 Since our focus is on linguistic cues, we eliminated all gaze and gestural signals in Experiment 2 by replacing the videos of human actors with videos of puppets.
-Puppets are less realistic and expressive than human actors but, since we wanted to eliminate gestural and gaze cues, the puppets created a natural context for having somewhat motionless talkers in the videos.
+Puppets are less realistic and expressive than human actors, but they create a natural context for having somewhat motionless talkers in the videos (thereby allowing us to eliminate gestural and gaze cues).
 Additionally, the prosody-controlled condition included small but global changes to syllable duration that would have required complex video manipulation or precise re-enactment with human talkers, neither of which was feasible.
@@
 We took care to match the puppets' mouth movements to the syllable onsets as closely as possible and avoided any mouth movement before the onset of a turn.
-We then hand-aligned the manipulated audio clips to the puppet video recordings.
+We then added the manipulated audio clips to the puppet video recordings by hand.
 
@@
 The remaining turn transitions were distributed evenly across transition types (questions N$=$47 and non-questions N$=$32) and conditions, keeping in mind that there were three \textit{normal} videos and only one limited cue video for each experiment version: \textit{normal} (N$=$36), \textit{words only} (N$=$13), \textit{prosody only} (N$=$12), and {no speech} (N$=$18).
-Question transitions (mean$=$358, median$=$405) were longer than non-question transitions (mean$=$296, median$=$288) on average, but gap duration was overall comparable across conditions: \textit{normal} (mean$=$333, median$=$337), \textit{words only} (mean$=$345, median$=$383), \textit{prosody only} (mean$=$320, median$=$336), and \textit{no words} (mean$=$324, median$=$328).
+Inter-turn gaps for questions (mean$=$358, median$=$405) were longer than those for non-questions (mean$=$296, median$=$288) on average, but gap duration was overall comparable across conditions: \textit{normal} (mean$=$333, median$=$337), \textit{words only} (mean$=$345, median$=$383), \textit{prosody only} (mean$=$320, median$=$336), and \textit{no words} (mean$=$324, median$=$328).
 The longer gaps for question transitions could give them an advantage because our anticipatory measure includes shifts initiated during the gap between turns (Figure \ref{fig:criterion}).
@@
 
+% address this!!
 Participants at all ages made anticipatory gaze switches more often than would be expected by chance in almost all conditions, but sometimes by a small margin (Figure \ref{fig:randvsrealEN}).
@@
 \end{figure}
-
+% evidence for integration of diff cues?
 Participants made more anticipatory gaze switches following questions than non-questions, but only in the two lexical conditions (\textit{normal} and \textit{words only} speech).
@@
 
-We tested these findings in two linear mixed-effects models of the baseline-corrected anticipatory switches---one model each for children and adults.
+We tested these findings in two linear mixed-effects models of participants' baseline-corrected anticipatory switches---one model each for children and adults.
 The child model included condition (\textit{normal}, \textit{words only}, \textit{prosody only}, and \textit{no speech}), transition type (question vs.
@@
 
+% no main effects is kind of meaningless
 Condition-related differences were driven by age and transition type (Figure \ref{fig:questionsEN}).
@@
 By extending the age range of participants, we were able to see more of the developmental trajectory in children's predictions about turn structure.
-One-year-olds already made more anticipatory switches than would be expected by chance in two of the four conditions.
+We saw that one-year-olds already made more anticipatory switches than would be expected by chance in two of the four conditions.
 As they got older, children demonstrated above-chance anticipatory switches in more conditions and made more anticipatory switches overall.
@@
 Though the advantage for questions generally increased with age, the biggest developmental shift in our data occurred between one and two years, when the question effect began to reliably emerge in the two lexical conditions.
-It will be up to future work to disentangle which lexical (and prosodic) cues children begin to pick out in distinguishing questions from other speech acts, and when.
+It will be up to future work to disentangle which lexical and prosodic cues children begin to pick out in distinguishing questions from other speech acts, and when.
 
@@
 Importantly, these results do not support a straightforward view of lexicosyntax as the primary linguistic cue for turn predictions (even for adults) \citep{de-ruiter2006, magyari2012}.
-
+% and in agreement w/ torr, other corpus
+
+%mention again how speech manic was diff from dR
 We controlled linguistic information in Experiment 2 by using phonetically-manipulated speech \citep{de-ruiter2006}.
@@
 Children could have had trouble processing the \textit{words only} and \textit{prosody only} conditions because they were unfamiliar, and not just because they had less linguistic information available.
-
+%maybe move to GD?
 It would take a different type of stimulus to compare lexical and prosodic information without phonetic manipulation.
@@
 \label{sec:gendisc}
-
+%Refs again after 1st sentence
 Children begin to develop conversational turn-taking skills long before their first words.
@@
 
+%"typical" is problematic; also contrast global vs. local findings for pros, lex
 In two experiments investigating children's anticipatory gaze to upcoming speakers, we found evidence that turn prediction develops early in childhood, and that linguistic cues are integrated differently at different ages.
-Children under three showed an advantage for prosody over lexicosyntax, contrary to the typical finding for adults in prior work \citep{de-ruiter2006}.
-Children two and older also showed a slight advantage for lexicosyntax in the form of a transition type effect: children made more anticipatory switches following questions than non-questions in conditions where lexical information was present.
-We found no evidence that, for children or adults, lexicosyntax alone was ``sufficient'' (equal to full linguistic information) for spontaneous turn prediction (cf.
-\citealp{de-ruiter2006}): Participants' performance was best in conditions when they had access to the full linguistic signal.
+Children under 3;0 showed an advantage for prosody over lexicosyntax, contrary to the typical finding for adults in prior work \citep{de-ruiter2006}.
+Children 2;0 and older also showed a slight advantage for lexicosyntax in the form of a transition type effect: children made more anticipatory switches following questions than non-questions in conditions where lexical information was present.
+We found no evidence that, for children or adults, lexicosyntax alone was ``sufficient'' (equal to full linguistic information) for spontaneous turn prediction \citep[pg.
+531]{de-ruiter2006}: Participants' performance was best in conditions when they had access to the full linguistic signal.
 
@@
 \subsubsection{The advantage for questions}
-
+% remind what advantage was
 In both of our experiments we found a robust advantage for questions in participants' anticipatory switching.
@@
 Also, the form of a question can constrain the type of response that will come next (e.g., a location after a \textit{where} question), further increasing the predictability of the interactional sequence.
-
+% potentially makes more sense w/ pros intro
 Children get frequent and early experience with hearing and answering questions.
@@
 \subsubsection{Linguistic cues for turn prediction}
-
+% mention adult counter evidence again, elaborate on prosody a little (and add refs after "knowing..." sentence)
 Prior work with adults has found a consistent and critical role for lexicosyntax in predicting upcoming turn structure, with a more secondary and supporting role for prosody \citep{de-ruiter2006, magyari2012}.
@@
 More complex structural predictions would have required them to access lexicosyntactic information because higher-level structure in the prosodic signal is linked to the syntactic structure being used.
-The close link between prosodic and syntactic structure makes it difficult to tease apart how predictive processing in one domain (e.g., prosody) is distinct from the other (e.g., syntax) in turn prediction.
-We leave this challenge to future work.
 
@@
 
-\subsection{Relationship with prior work}
-
+The close link between prosodic and syntactic structure makes it difficult to tease apart how predictive processing in one domain (e.g., prosody) is distinct from the other (e.g., syntax) in turn prediction.
+We leave this challenge to future work.
+
+\subsection{What's new}
+% mention again that our prosody manipulation included durational changes
 Our findings are only partially consistent with those reported in prior work.
-De Ruiter and colleagues (2006) found that adults' accuracy in identifying upcoming turn-ends was not significantly affected when intonation was taken away.
-Keitel and colleagues (2013) found the same for adults' anticipatory gaze switches, only finding that intonation affected children's switches at age 3;0 (with switches by younger children never exceeding chance).
+De Ruiter and colleagues (2006) found that adults' accuracy in identifying upcoming turn-ends was not significantly affected when intonation contours were flattened.
+Keitel and colleagues (2013) found the same for adults' anticipatory gaze switches, only finding that intonation flattening affected children's switches at age 3;0 (with switches by younger children never exceeding chance).
 Our results, in contrast, suggest that children's anticipatory switches emerge much earlier, and rely on prosody before they do lexicosyntax.
@@
 
+%include subheadings to look at diffs w/ adult vs. chi literature
 Despite extreme differences in linguistic information across conditions, main effects of linguistic information were minimal in our experiments.
@@
 
+% move to the front of this section
 Our findings are consistent with observational work on children's turn taking.
@@
 One alternative to explore for the future would be scripted speech, or speech that is carefully selected from a large corpus, given a set of specific linguistic constraints.
+% or cross-splicing
 
@@
 If this statistical expectation were the only thing leading children's anticipatory gaze switches, we would have predicted more uniformity in looking behavior across the conditions and no major effects of linguistic signal, age, or transition type.
-But that is now what we saw.
+But that is not what we saw.
 Nevertheless, another advantage of using scripted conversation would be the ability to control switch frequency and eliminate this issue.
@@
 Thus, even at the level of linguistic integration, children's attention during conversation is driven by goals from real-time interaction, the principle source and destination for their language learning.
+% increase link between findings and last sentence, or don't make the jump (can just leave out).
 

### 2014-09-30 paper/CasillasFrank-Anticip.tex
Edits based on pre-submission feedback from Francisco, Steve, and the FLA group.

@@
 
-\author[MPI]{Marisa Casillas\corref{cor1}}
-% \address[StanfordLX]{Department of Linguistics, Stanford University}
+\author[MPI,StanfordLX]{Marisa Casillas\corref{cor1}}
 \address[MPI]{Max Planck Institute for Psycholinguistics, Nijmegen}
+\address[StanfordLX]{Department of Linguistics, Stanford University}
 \cortext[cor1]{Corresponding author}
@@
 We found that children already predicted upcoming turns at age one, but they integrated linguistic cues differently at different ages.
-Children under three showed a general advantage for prosody over lexicosyntax, contrary to the typical finding for adults.
+Children under three showed a general advantage for prosody over lexicosyntax, contrary to prior findings for adults.
 Children two and older showed more predictive switches for questions than non-questions, but only when lexical information was available.
 We found no evidence that lexicosyntax alone guides turn prediction---instead, participants' performance was best overall with access to lexicosyntax and prosody together.
-Our results point to the importance of interaction in children's real-time linguistic processing.
+Our results point to the importance of studying children's linguistic processing in the context of conversation.
 \end{abstract}
@@
 \begin{keyword}
-Turn-taking \sep Conversation \sep Development \sep Prosody \sep Lexical \sep Questions \sep Eye-tracking \sep Anticipation
+Turn taking \sep Conversation \sep Development \sep Prosody \sep Lexical \sep Questions \sep Eye-tracking \sep Anticipation
 %% MSC codes here, in the form: \MSC code \sep code
@@
 
-% \linenumbers
+%\linenumbers
 
@@
 
-Children participate in sequential coordination with their caregivers starting at three months of age---before they can rely on any linguistic cues in taking turns \citep{snow1977, jaffe2001, hilbrinkInPrep}.
-Of course, infant turn-taking is different from adult turn-taking in several ways.
-Infant turn-taking is heavily scaffolded by caregivers, has timing distinct from adult turn-taking, and lacks semantic content \citep{hilbrinkInPrep, jaffe2001, snow1977}.
-However, children's early, turn-structured social interactions are presumably a critical precursor to their conversational turn-taking.
-Early non-verbal interactions establish the protocol by which children come to use language with others.
+Children participate in sequential coordination with their caregivers starting at three months of age---before they can rely on any linguistic cues in taking turns \citep[see, among others, ][]{bateson1975, snow1977, jaffe2001, hilbrinkInPrep}.
+Of course, infant turn taking is different from adult turn taking in several ways.
+Infant turn taking is heavily scaffolded by caregivers, has distinct timing in comparison to adult turn taking, and lacks semantic content \citep{hilbrinkInPrep, jaffe2001}.
+However, children's early, turn-structured social interactions are presumably a critical precursor to their conversational turn taking.
+Along these lines, early non-verbal interactions establish the protocol by which children come to use language with others.
 And as they acquire language, they also come to integrate it into the preverbal turn-taking systems.
@@
 
-\subsection{Turn-taking}
-
-Turn-taking itself is not unique to conversation.
-Many other human activities are organized around sequential turns.
+\subsection{Turn taking}
+
+Turn taking itself is not unique to conversation.
+Many other human activities are organized around sequential turns at action.
 Traffic intersections and computer network communication both use turn-taking systems.
@@
 Even monkeys take turns: Non-human primates such as marmosets and Campbell's monkeys vocalize contingently with each other in both natural and lab-controlled environments \citep{lemasson2011, takahashi2013}.
-In all these cases, turn-taking serves as a protocol for interaction, allowing the participants to coordinate with one another through sequences of contingent action.
+In all these cases, turn taking serves as a protocol for interaction, allowing the participants to coordinate with one another through sequences of contingent action.
 
@@
 
-Despite this complexity, conversational turn-taking is often precise in its timing.
+Despite this complexity, conversational turn taking is often precise in its timing.
 Across a diverse sample of conversations in 10 languages, one study found a consistent average transition time of 0--200 msec at points of speaker switch \citep{stivers2009}.
 Experimental results and current models of speech production suggest that it takes approximately 600 msec to produce a content word, and even longer to produce a simple utterance \citep{griffin2000, levelt1989}.
-So in order to achieve 200 msec turn transitions, speakers must begin formulating their response before the prior turn has ended.
+So in order to achieve 200 msec turn transitions, speakers must begin formulating their response before the prior turn has ended \citep{levinson2013}.
 Moreover, to formulate their response early on, speakers must track and anticipate what types of response might become relevant next.
 They also need to predict the content and form of upcoming speech so that they can launch their articulation at exactly the right moment.
-Prediction must thus play a key role in timely turn-taking.
+Prediction thus plays a key role in timely turn taking.
 
 \subsection{Adults' turn prediction}
-% FIX: compress deR: make it clear about prosody issues, perhaps talk more clearly about torreira
-Adults have a lot of information at their disposal that could potentially help with their predictions about upcoming turn structure.
-Lexical, syntactic, and prosodic information (e.g., \textit{wh}- words, subject-auxiliary inversion, list intonation) can all inform addressees about upcoming linguistic structure \citep{de-ruiter2006, duncan1972, ford1996, torreiraUndRev}.
-Non-verbal cues (e.g., gaze, posture, and pointing) often appear at turn-boundaries as late indicators of an upcoming speaker switch \citep{rossano2009, stivers2010}.
-The sequential context of a turn can make it clear what will come next: answers after questions; thanks or denial after compliments, et cetera \citep{schegloff2007}.
-Prior work suggests that adult listeners primarily use lexicosyntactic information to accurately predict upcoming turn structure, while prosodic information plays more of a secondary role \citep{de-ruiter2006}.
-For example, a mid-high ``continuation'' tone can cancel the possibility of a speaker switch at an otherwise syntactically- and pragmatically-complete turn point (e.g., at the start of a story-telling; \citealp{ford1996, torreiraUndRev}).
-
-Recent experimental work on adult turn-taking has tried to tease apart which linguistic cues responders rely on in making online predictions about upcoming turn structure.
+
+Adults have a lot of information at their disposal to help make accurate predictions about upcoming turn content.
+Lexical, syntactic, and prosodic information (e.g., \textit{wh}- words, subject-auxiliary inversion, and list intonation) can all inform addressees about upcoming linguistic structure \citep{de-ruiter2006, duncan1972, ford1996, torreiraUndRev}.
+Non-verbal cues (e.g., gaze, posture, and pointing) often appear at turn-boundaries and can sometimes act as late indicators of an upcoming speaker switch \citep{rossano2009, stivers2010}.
+Additionally, the sequential context of a turn can make it clear what will come next: answers after questions; thanks or denial after compliments, et cetera \citep{schegloff2007}.
+
+Prior work suggests that adult listeners primarily use lexicosyntactic information to accurately predict upcoming turn structure \citep{de-ruiter2006}.
 De Ruiter and colleagues \citeyearpar{de-ruiter2006} asked participants to listen to snippets of spontaneous conversation and to press a button whenever they anticipated that the current speaker was about to finish her turn.
-The snippets that participants listened to were controlled for the amount of linguistic information present; some snippets were normal, unmodified versions of the recording, but some had words and rhythm (but no pitch), pitch and rhythm (but no words), neither words nor pitch, or were replaced entirely with a noise profile matched to the original recording.
-In the normal, unmodified snippets, participants' anticipatory button presses were highly accurate, falling an average of 180 msec before the phonetic end of the turns.
-Participants also made accurate predictions for the snippets with words and rhythm (but no pitch); there was no significant difference compared to their performance in the normal condition.
-But, in the snippets with no lexical information, participants' accuracy significantly declined.
-De Ruiter and colleagues \citeyearpar{de-ruiter2006} concluded that lexicosyntactic information is the primary linguistic signal listeners use for making online predictions about turn-end boundaries.
-They also found a significant, but secondary role for pitch: participants' accuracy in the pitch and rhythm (but no words) condition was still better than the neither-words-nor-pitch and the noise profile conditions.
-They concluded that pitch's importance for online prediction of turn-ends is restricted, falling secondary to lexicosyntactic information.
-
-Congruent evidence comes from studies varying the predictability of lexicosyntactic and pragmatic content.
-Adults anticipate turn ends better when they can more accurately predict the exact words that will come next (\citealp{magyari2012}; see also \citealp{magyariUndRev}).
-They can also identify speech acts within the first word of an utterance \citep{gisladottirUndRev}, allowing them to start planning their response at the first moment possible.
-\citet*{bogelsmagyariInPrep} used a trivia quiz set-up to get participants to answer questions in which the answer became known either early or late in the question (e.g., ``Which character, also called 007, appears in the famous movies?'' vs.
-``Which character, from the famous movies, is also called 007?'').
-Their participants showed a consistent neural response 500 msec after the answer became unambiguous, suggesting immediate response planning.
-
-In another study, \citet*{torreiraUndRev} used a combination of button-press and verbal responses to investigate the relationship between lexicosyntactic and prosodic cues in turn-boundary identification.
-They found that participants never verbally responded or pressed the ``turn-end'' button when hearing a syntactically-complete unit without an intonational phrase boundary.
-However, utterance-level prosodic boundaries embedded in multi-utterance turns could trick participants into anticipating a turn end 29\% of the time (see also de \citet{de-ruiter2006}:525).
-Their findings corroborate other corpus and experimental findings promoting a combination of cues (lexicosyntactic, prosodic, and pragmatic) as key for accurate turn-end prediction \citep{duncan1972, ford1996, hirvenkari2013}.
+The speech snippets were controlled for the amount of linguistic information present; some were normal, but others had flattened pitch, low-pass filtered speech, or further manipulations.
+De Ruiter and colleagues found that, with pitch-flattened speech, the timing of participants' button responses was comparable to their timing with the full linguistic signal.
+But, when no lexical information was available, participants' responses were significantly earlier than the full linguistic signal.
+The authors concluded that lexicosyntactic information\footnote{The ``lexicosyntactic'' condition only included flattened pitch and so was not exclusively lexicosyntactic---the speech would still have residual prosodic structure, including syllable duration and intensity.} was necessary and possibly sufficient for turn-end projection, while intonation was neither necessary nor sufficient.
+Congruent evidence comes from studies varying the predictability of lexicosyntactic and pragmatic content: Adults anticipate turn ends better when they can more accurately predict the exact words that will come next (\citealp{magyari2012}; see also \citealp{magyariUndRev}).
+They can also identify speech acts within the first word of an utterance \citep{gisladottirUndRev}, allowing them to start planning their response at the first moment possible \citep{bogelsmagyariInPrep}.
+
+The role of prosody for adult turn-prediction is still a matter of debate.
+De Ruiter and colleagues' (2006) experiment only tested the role of intonation, which is only a partial index of prosody.
+Prosodic structure is also tied closely to the syntax of an utterance, and so the two linguistic signals are difficult to control independently \citep{ford1996}.
+\citet*{torreiraUndRev} used a combination of button-press and verbal responses to investigate the relationship between lexicosyntactic and prosodic cues in turn-end prediction.
+Critically, their stimuli were cross-spliced so that each item had full prosodic cues to accompany the lexicosyntax.
+Because of the splicing, they were able to create items that had syntactically-complete units with no intonational phrase boundary at the end.
+Participants never verbally responded or pressed the ``turn-end'' button when hearing a syntactically-complete phrase without an intonational phrase boundary.
+And when intonational phrase boundaries were embedded in multi-utterance turns, participants were tricked into pressing the ``turn-end'' button 29\% of the time.
+Their results suggest that listeners actually do rely on prosodic cues to execute a response (see also de \citet{de-ruiter2006}:525).
+These experimental findings corroborate other corpus and experimental work promoting a combination of cues (lexicosyntactic, prosodic, and pragmatic) as key for accurate turn-end prediction \citep{duncan1972, ford1996, hirvenkari2013}.
 
 In sum, adults accurately and spontaneously make predictions about upcoming turn structure.
-They use a wide variety of cues in making their predictions and then act immediately upon them.
-Their predictive processing relies on a sophisticated body of knowledge about linguistic structure, non-verbal signals, and social actions.
-The later acquisition of verbal turn-taking skills must then be tied to children's knowledge about language, gaze, gesture, and social cues.
-But children's turn taking starts early in infancy, long before first words or gestures emerge; a primary role for lexicosyntactic cues doesn't fit well with children's pre-verbal turn-taking.
+Their predictions rely on a sophisticated body of knowledge about linguistic structure, non-verbal signals, and social actions.
+Knowing this, we could expect that children's acquisition of turn-taking skills is closely tied to their knowledge about language, gaze, gesture, and social cues.
+But children's turn taking starts early in infancy, long before first words or gestures emerge.
+So a primary role for lexicosyntactic cues doesn't fit well with children's pre-verbal turn taking.
 
@@
 The majority of work on children's early turn taking has focused on observations of spontaneous interaction.
-Children's first turn-like structures appear as early as two to three months in proto-conversation with their caregivers.
-During proto-conversations, caregivers interact with their infants as if they were capable of making meaningful contributions; they take every look, vocalization, arm flail and burp as ``utterances'' in the joint discourse \citep{snow1977, jaffe2001}.
+Children's first turn-like structures appear as early as two to three months in proto-conversation with their caregivers \citep{bruner1975, bruner1985}.
+During proto-conversations, caregivers interact with their infants as if they were capable of making meaningful contributions; they take every look, vocalization, arm flail and burp as ``utterances'' in the joint discourse \citep{bateson1975, snow1977, jaffe2001}.
 Infants catch onto the structure of proto-conversations quickly.
@@
 A recent study by Hilbrink et al.
-(in preparation) finds that infants' turn timing at three months is often too early or too late: they start vocalizing in overlap on nearly 40\% of their caregivers' turns, and their non-overlapped vocalizations come after an average latency of 675 msec (adult average: 200 msec).
-Between four and nine months, children begin to reduce the number of turns happening overlap while also improving on their average response latency.
+(in preparation) finds that infants' turn timing at three months is often too early or too late: they start vocalizing in overlap on nearly 40\% of their caregivers' turns, and their non-overlapped vocalizations come after an average inter-turn silent gap of 675 msec (adult average: 200 msec).
+Between four and nine months, children begin to reduce the number of turns happening in overlap while also improving on their average response latency.
 But then, later on, children's response latencies slow down again, peaking at average gaps of 1100 msec at nine months, with only very gradual improvement after that \citep{hilbrinkInPrep}.
-While children's avoidance of overlap is nearly adult-like by nine months, the timing of their non-overlapped responses stays much longer than the 200 msec standard for the next several years \citep{ervin-tripp1979}.
+While children's avoidance of overlap is nearly adult-like by nine months, the timing of their non-overlapped responses stays much longer than the 200 msec standard for the next several years \citep{garvey1984, ervin-tripp1979}.
 
@@
 Integrating language into the turn-taking system may be one major factor in children's delayed responses \citep{casillasUndRev}.
-If response planning (i.e., language production) is the primary hurdle in children's spontaneous turn taking, we should find evidence that children understand turn-taking behaviors before they are able to produce the behaviors themselves; this hypothesis has been more recently explored in experimental settings.
+If response planning (i.e., language production) is the primary hurdle in children's spontaneous turn taking, we should find evidence that children understand turn-taking behaviors before they are able to produce the behaviors themselves; this hypothesis has been recently explored in experimental settings.
 
@@
 There are mixed results regarding when children begin to anticipate turn structure in conversation.
-One study found that 12-month-olds make more predictive gaze shifts while watching human verbal conversation compared to object non-speech ``conversation'' \citep{bakker2011}, but another only found a similar effect at 36 months \citep{hofsten2009}.
+One study found that 12-month-olds make more predictive gaze shifts to a responder while watching human verbal conversation compared to conversation-like interactions with objects \citep{bakker2011}, but another only found a similar effect at 36 months \citep{hofsten2009}.
 Neither of these two studies had baselines to which the turn-relevant looking behavior could be compared, however.
@@
 The adult control group made equal numbers of anticipatory looks in the videos, with and without intonation contours, consistent with prior adult findings \citep{de-ruiter2006}.
-Keitel and colleagues concluded that children's ability to predict upcoming turn structure relies on their ability to comprehend the stimuli (emerging around 3;0), especially with respect to semantic access.
-They also suggest that intonation takes a secondary role in turn prediction, but only \textit{after} children acquire more sophisticated, adult-like language comprehension systems (sometime after 3;0).
+Keitel and colleagues concluded that children's ability to predict upcoming turn structure relies on their ability to comprehend the stimuli (emerging around 36 months), especially with respect to semantic access.
+They also suggest that intonation takes a secondary role in turn prediction, but only \textit{after} children acquire more sophisticated, adult-like language comprehension systems (sometime after 36 months).
 
@@
 First, their estimates of baseline gaze frequency (``random'' in their terminology) were computed from switches that occurred specifically while the speaker's turn was ongoing; in contrast, estimates of anticipatory shifting frequency were computed from switches that occurred within the time window immediately surrounding inter-turn gaps.
-Assuming that gaze switches happen most often during the inter-turn gap \citep{hirvenkari2013}, switches made during ongoing speech (their baseline) would be minimized.
+Assuming that gaze switches happen most often during the inter-turn gap \citep{hirvenkari2013}, the switches made during ongoing speech (their baseline) would be minimized.
 This baseline thus maximized the possibility of finding a difference between ``random'' and anticipatory gaze frequencies.
@@
 In addition, in order to produce more naturalistic conversation, it would have been ideal to localize sources for the two voices in the video (i.e., to have the voices come out of separate left and right speakers).
-But both voices were recorded and played back on the same audio channel, which may have made it more difficult to distinguish the speakers.
+But both voices were recorded and played back on the same audio channel, which may have made it more difficult to distinguish the two talkers.
 Again, we attempt to address these issues in our current study.
@@
 Despite these minor methodological issues, the Keitel et al.
-\citeyearpar{keitel2013} study still demonstrates intriguing age-based differences in children's ability to predict upcoming turn structure, and the results suggest that both semantic and prosodic development \textit{do} play a role in children's looking patterns.
-Our current work thus takes this paradigm as our starting point.
+\citeyearpar{keitel2013} study still demonstrates intriguing age-based differences in children's ability to predict upcoming turn structure, and the results suggest that both semantic and intonational development \textit{do} play a role in children's looking patterns.
+Our current work thus takes this paradigm as our starting point.\footnote{See also \citet{casillas2012, casillas2013}.}
+
+\subsubsection{Prosodic development}
+The roles of prosody and lexicosyntax in children's turn predictions are currently unknown, but children understand more about prosody early in life.
+Children begin to acquire prosody in the womb, and can distinguish their native language's rhythm type from others (e.g., syllable-timed vs.
+stress-timed) 2--5 days after birth \citep{mehler1988, moon1993, nazzi2003}.
+Beginning between four and five months, infants prefer pauses in speech to be inserted at prosodic boundaries, and by 6 months they can start using prosodic markers to pick out sub-clausal syntactic units \citep{jusczyk1995, soderstrom2003}.
+They show preference for the typical stress patterns of their native language over others by 6--9 months (e.g., iambic vs.
+trochaic), and can use prosodic information to segment the speech stream into smaller chunks from 8 months onward \citep{johnson2001, jusczyk1993, morgan1995}.
+In comparison, children show only a limited lexical inventory at six months, and begin to recognize function words just before their first birthdays, with syntactic categorization beginning around 14 months \citep{bergelson2013, shi2010}.
+Two-month-olds notice changes in word order, but this ability appears to rely on prosodic cueing \citep{mandel1996}.
+Generally speaking then, our current knowledge about children's linguistic development points to a possible early advantage for prosody in children's turn-taking predictions.
 
 \subsection{The Current Study}
-% add bit about children's prosodic development and link this to the plans for the experiments
-% mention that we will be looking at anticipatory gaze switching
 We report here on the role of linguistic processing in children's predictions about upcoming turn structure.
+We focus in particular on how children use prosodic and lexicosyntactic information to make their predictions.
+Prior work has focused mainly on lexicosyntax and intonation, and not on prosody proper (\citealp{de-ruiter2006}; \citealp{keitel2013}, but see \citealp{torreiraUndRev}), even though infants seem to acquire the basic rhythmic properties of the prosodic signal first \citep{mehler1988, moon1993, nazzi2003}.
+
 In two eye-tracking experiments, we measured children's anticipatory gaze to upcoming responders while controlling for the amount of lexicosyntactic and prosodic information available.
-Experiment 1 uses a paradigm in which English-speaking participants view video clips of naturalistic conversations from several different languages, to create conditions under which lexicosyntactic information was differentially available, but prosodic and non-linguistic cues were intact.
+Experiment 1 uses a paradigm in which English-speaking participants view video clips of naturalistic conversations from several different languages in order to create conditions under which lexicosyntactic information was differentially available, but prosodic and non-linguistic cues were intact.
 This paradigm showed minimal differences between the predictive looking behavior of preschoolers and adults.
@@
 Both prosodic and lexicosyntactic information played a role in children's predictions about turn structure, but the two information sources were used differently at different ages.
-Our findings overall support an account in which predictive processes for turn-taking in conversation are present early, but their integration with linguistic information takes substantial practice.
+Our findings overall support an account in which predictive processes for turn taking in conversation are present early, but their integration with linguistic information takes substantial practice.
 
@@
 
-We recruited 74 children between ages 3;0--6;0 and 11 undergraduate adults to participate in the study.
+We recruited 74 children between ages 3;0--5;11 and 11 undergraduate adults to participate in the study.
 Our child sample included 19 three-year-olds, 32 four-year-olds, and 23 five-year-olds, all enrolled in a local nursery school.
@@
 We chose male-female pairs as a natural way of creating contrast between the two talker voices.
-See the videos here: NOTE: Link to be added.}
-
-After recording, we hand-aligned the audio and video files, and cropped each recording to the 30-second interval with the most turn activity.
+See an example run-though of the videos here: www.youtube.com/channel/UCGEZQcM9t8Zfjqqi\_B1Q5Sw.}
+
+After recording, we combined the audio and video files by hand, and cropped each recording to the 30-second interval with the most turn activity.
 Because we recorded the conversations in stereo, the male and female voices came out of separate speakers during video playback.
@@
 We measured the duration of each turn transition from the audio recording associated with each video.
-We excluded turn transitions longer than 550 msec and shorter than 0 msec (those with transitional overlap) from analysis.\footnote{Overlap occurs when an addressee begins a new turn before the current turn is finished.
-When overlap occurs, observers cannot switch their gaze in anticipation of the response because the response begins \textit{earlier} than expected; participants expect conversations to proceed with ``one speaker at a time'' \citep{sacks1974}.
+We excluded turn transitions longer than 550 msec and shorter than 0 msec (those with transitional overlap) from analysis.\footnote{Overlap occurs when a responder begins a new turn before the current turn is finished.
+When overlap occurs, observers cannot switch their gaze in anticipation of the response because the response begins earlier than expected; participants expect conversations to proceed with ``one speaker at a time'' \citep{sacks1974}.
 As such, they would still be fixated on the prior speaker when the overlap started, and then would have to switch their gaze \textit{reactively} to the responder.} This left approximately equal numbers of turn transitions available for analysis in the English (N$=$22) and non-English (N$=$21) videos.
 On average, the inter-turn gaps for English videos (mean$=$273 msec, median$=$291 msec) were slightly longer than for non-English videos (mean$=$226 msec, median$=$205 msec).
-The longer gaps in the English videos could give them a slight advantage: our definition of an ``anticipatory gaze shift'' includes shifts that are initiated during the gap between turns (Figure \ref{fig:criterion}), so participants had slightly more time to make anticipatory shifts in the English videos.
+The longer gaps in the English videos could give them a slight advantage: Our definition of an ``anticipatory gaze shift'' includes shifts that are initiated during the gap between turns (Figure \ref{fig:criterion}), so participants had slightly more time to make anticipatory shifts in the English videos.
 
@@
 Before the experiment started, we warned adult participants that they would see videos in several languages and that, though they weren't expected to understand the content of non-English videos, we \textit{would} ask them to answer general, non-language-based questions about the conversations.
-Then after each video we asked participants one of the following randomly-assigned questions: ``which speaker talked more?'', ``which speaker asked the most questions?'' ``which speaker seemed more friendly?'', and ``did the speakers' level of enthusiasm shift during the conversation?'' We also asked if the participants could understand any of what was said after each video.
+Then after each video we asked participants one of the following randomly-assigned questions: ``which speaker talked more?'', ``which speaker asked the most questions?'', ``which speaker seemed more friendly?'', and ``did the speakers' level of enthusiasm shift during the conversation?'' We also asked if the participants could understand any of what was said after each video.
 The participants responded verbally while an experimenter noted their responses.
@@
 Following prior work, we measured how often our participants shifted their gaze from the prior to the upcoming speaker \textit{before} the shift in gaze could have been initiated in reaction to the onset of the speaker's response.
-In doing so, we assumed that it takes children 333 msec to plan an eye movement, following Fernald and colleagues' (2008) similar `looking while listening' method.
-
-We checked each participant's gaze at each turn transition for three characteristics: (1) that the participant fixated on the prior speaker for at least 100 msec at the end of the prior turn, (2) that sometime thereafter the participant switched to fixate on the upcoming speaker for at least 100 ms, and (3) that the switch in gaze was initiated within the first 333 msec of the response turn, or earlier.
+In doing so, we assumed that it takes children 333 msec to plan an eye movement, following Fernald and colleagues' \citeyearpar{fernald2008} similar `looking while listening' method.
+We assumed that it takes adults 200 msec to plan an eye movement, following standards from adult anticipatory processing studies \citep[e.g., ][]{kamide2003}.
+
+We checked each participant's gaze at each turn transition for three characteristics (Figure \ref{fig:criterion}): (1) that the participant fixated on the prior speaker for at least 100 msec at the end of the prior turn, (2) that sometime thereafter the participant switched to fixate on the upcoming speaker for at least 100 ms, and (3) that the switch in gaze was initiated within the first 333 msec of the response turn, or earlier.
 These criteria guarantee that we only counted gaze shifts when: (1) participants were tracking the previous speaker, (2) switched their gaze to track the upcoming speaker, and (3) did so before they could have simply reacted to the onset of speech in the response.
-Under this assumption, a gaze shift that is initiated within the first 333 msec of the response (or earlier) was planned \textit{before} the child could react to the onset of speech itself.
-For adult participants we used 200 msec (instead of 333 msec) as the assumed time to plan an eye movement.
+Under this assumption, a gaze shift that was initiated within the first 333 msec of the response (or earlier) was planned \textit{before} the child could react to the onset of speech itself.
+
+As mentioned, most anticipatory switches happen in the inter-turn gap, but we also allowed anticipatory gaze switches that occurred in the final syllables of the prior turn.
+Early switches are consistent with the distribution of responses in explicit turn-boundary prediction tasks.
+For example, in a button press task, adult participants anticipate turn ends approximately 200 msec in advance of the turn's end, and anticipatory responses to pitch-flattened stimuli come even earlier \citep{de-ruiter2006}.
+We therefore allowed switches to occur as early as 333 msec before the end of the prior turn for children, and 200 msec before the end of the prior turn for adults.
+For very early and very late switches, our requirement for 100 msec of fixation on each speaker would sometimes extend outside of the transition window boundaries (333 msec before and after the inter-turn gap).
+The maximally available fixation window was 100 msec before and after the earliest and latest possible switch point (433 msec before and after the inter-turn gap).
+We did not count switches made during the fixation window as anticipatory.
+We \textit{did} count switches made during the inter-turn gap.
+The period of time from the beginning of the possible fixation window on the prior speaker to the end of the possible fixation window on the responder was our total analysis window (433 msec $+$ the inter-turn gap $+$ 433 msec), which we use to create the random baselines described below.
+For adult participants we used 200 msec (instead of 333 msec) as the assumed time to plan an eye movement in defining all analysis windows.
 
@@
 
-% check over again to make as clear as possible
-Using these criteria, we created a three-step algorithm for detecting anticipatory gaze switches.
-Consider, for example, when a child observes a turn transition between speaker A and speaker B (Figure \ref{fig:criterion}).
-First, the child would need to fixate on A for at least 100 msec at the end of A's turn.
-But what is the relevant window for fixation?
-Consistent with the distribution of responses in explicit turn-boundary prediction tasks (de Ruiter et al., 2006), we counted any early switches as ``truly turn-related'' as long as they occurred in the final 333 msec of A's turn (200 msec for adults).
-To allow for these early switches, the window for fixating on A must start 433 msec before the end of A's turn (300 msec for adults).
-Second, given that the child fixated on A, we checked to see if the child also fixated on B.
-Because we allowed for early switches, the first possible fixation on B could start 333 msec before the end of A's turn.
-Assuming that it took the child 333 msec to plan an eye movement, the latest point when the gaze could have \textit{anticipated} speech would be 333 msec after B's turn began---a switch after this point could simply be a reaction to the onset of B's speech.
-Thus, the relevant window for fixating on B started 333 msec before the gap and ended 433 msec after it (200 msec and 300 ms, respectively, for adults).
-Success on the first two criteria jointly implied that the child made an anticipatory switch in gaze.
-Our third step was to estimate when the switch occurred.
-Starting at a point when we knew participants were fixated on A (from step 1), we incremented forward in time to the first gaze measurement where the child looked away from A.
-We recorded the look-away timestamp as the time of gaze launch to the upcoming speaker.
-
-% check over again to make as clear as possible
 \paragraph{Random baseline analysis} To ensure that participants' looking behavior was turn-related and not simply the result of random switching, we corrected for the probability of making an anticipatory switch at random.
-We estimated the baseline chance of making an anticipatory switch at each turn transition for each person.
-We ran all participants' eye-tracking data through switch identification with 100 randomly-shuffled versions of the original turn transition windows (Figure \ref{fig:shuffling}).
-For each of the 100 randomly-shuffled versions, we recorded and then averaged whether or not each participant made an anticipatory switch at each ``turn transition.'' We then subtracted the random baseline value from each participant's actual switch result (1 or 0) for each turn transition to arrive at the final, baseline-corrected values.
-Subtracting the random baseline this way can give us an estimate of how often participants switched above chance across ages, conditions, and transition types.
+We estimated the baseline chance of making an anticipatory switch at each turn transition for each participant.
+We ran each participants' eye-tracking signal through our switch identification procedure \ref{sec:algorithm} with 100 randomly-shuffled versions of the analysis windows (Figure \ref{fig:shuffling}).
+For each of the 100 randomly-shuffled versions, we recorded whether the participant made a ``switch'' or not.
+We then averaged participants' switches from all 100 randomly-shuffled versions of the analysis windows, yielding a value between 0 and 1 for each turn transition for each participant.
+We then subtracted the random baseline value from each participant's actual switch result (1 or 0) for each turn transition to arrive at the final, baseline-corrected values.
+Subtracting the random baseline this way gives us an estimate of how often participants switched above chance across ages, conditions, and transition types.
 If participants made the same number of anticipatory switches, regardless of turn transition location, there would be no evidence that their switching behavior was linked to the conversation stimulus.
@@
 \end{center}
-\caption{Example of shuffling for five turn transition windows. The windows were +/-433 msec around the inter-turn gap for children, and from +/-300 msec for adults.} 
+\caption{Example of shuffling for five turn transition analysis windows. The windows were $\pm$433 msec around the inter-turn gap for children, and from $\pm$300 msec for adults.} 
 \label{fig:shuffling}
@@
 Participants looked at the screen most of the time during video playback (81\% and 91\% on average for children and adults, respectively).
-They primarily kept their eyes on the person who was currently speaking in both English and non-English videos: they gazed at the current speaker between 38\% and 63\% of the time, looking back at the addressee between 15\% and 20\% of the time (Table \ref{tab:e1_look}).
+They primarily kept their eyes on the person who was currently speaking in both English and non-English videos: They gazed at the current speaker between 38\% and 63\% of the time, looking back at the addressee between 15\% and 20\% of the time (Table \ref{tab:e1_look}).
 Even three-year-olds looked more at the current speaker than anything else, whether the videos were in a language they could understand or not.
@@
 
-We identified anticipatory gaze switches for all 43 usable turn transitions, based on the criteria and algorithm outlined in Section \ref{sec:algorithm}.
+We identified anticipatory gaze switches for all 43 usable turn transitions, based on the criteria outlined in Section \ref{sec:algorithm}.
 Because participants had greater access to linguistic information in English, we expected to see greater anticipation in the English videos than in the non-English videos.
@@
 \end{center}
-\caption{Proportion anticipatory gaze switches made for actual (solid) and randomly-shuffled (dashed) turn transitions in each condition, and across ages.
+\caption{Proportion of anticipatory gaze switches made for actual (solid) and randomly-shuffled (dashed) turn transitions in each condition, and across ages.
 Error bars represent the standard error of the mean.}
@@
 Anticipatory switch behavior in the non-English videos could not have been based on lexicosyntactic information, so both adults and children must have relied on other sources of information to make their predictions (e.g., non-native prosodic cues and non-verbal signals).
-We then corrected each participant's data by subtracting their random baseline estimations from their actual switching data for each turn transition.
-All comparisons below are made with the baseline-corrected values.
+For all further analyses we then corrected each participant's data by subtracting their random baseline estimations from their actual switching data for each turn transition.
 
 Overall, a greater proportion of participants made anticipatory switches in the English videos (0.17) compared to the non-English videos (0.09).
-But the tendency to make more anticipatory switches for English than non-English videos disappeared with age.
+But the tendency to make more anticipatory switches for English than non-English videos disappeared with age (Figure \ref{fig:conditionsFL}).
 While anticipatory switches increased with age for non-English videos, they decreased with age for English videos, resulting in nearly equal looking behavior for the two language conditions for adults.
-Otherwise, age-based differences in predictive gaze shifts were minor (Figure \ref{fig:conditionsFL}).
+Otherwise, age-based differences in predictive gaze shifts were minor.
 
@@
 But the advantage for questions was not universal---it depended on language condition and age.
-The advantage for questions didn't emerge until age five in the non-English videos .
+The advantage for questions didn't emerge until age five in the non-English videos.
 
@@
 non-English)\footnote{Because each non-English language was represented by a single stimulus, we cannot compute reliable cross-language differences.
-Gaze behavior might be best for languages that have more structural overlap English.
-English speakers can make predictions about the strength of upcoming Swedish prosodic boundaries nearly as well as Swedish speakers do, but Chinese speakers are at a disadvantage in the same task \cite{carlson2005}.
+Gaze behavior might be best for languages that have the most structural overlap with participants' native language: English speakers can make predictions about the strength of upcoming Swedish prosodic boundaries nearly as well as Swedish speakers do, but Chinese speakers are at a disadvantage in the same task \cite{carlson2005}.
 We would need multiple items from each of the language to check for similarity effects of specific linguistic features.}, transition type (question vs.
@@
 \caption{Cumulative proportion of anticipatory gaze switches planned during the transition window, across conditions and ages.
-Age is indicated by hue (light to dark; darker = older).
+Age is indicated by hue (darker = older).
 The solid line is the offset of the pre-gap turn, and the dashed line is the median onset of the post-gap turn in each condition.
@@
 An advantage for lexicosyntactic information is consistent with prior work on adult turn taking \citep{de-ruiter2006}, so it's curious that the adults did not show a difference in anticipatory switching between videos they did and did not understand.
-As mentioned above, it is likely that they used non-verbal cues to track the interactions and consequently found the videos so easy to follow that they did not track the turn-taking closely.
-To identify which \textit{linguistic} cues adults and children rely on, it is necessary to control the presence of non-verbal cues in the stimulus---Experiment 2 implements this control.
+As mentioned above, it is likely that they used non-verbal cues to track the interactions and consequently found the videos so easy to follow that they did not track the turn taking closely.
+To identify which linguistic cues adults and children rely on, it is necessary to control the presence of non-verbal cues in the stimulus---Experiment 2 implements this control.
 
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
 \section{Experiment 2}
@@
 \textit{Audio recordings}.
-The recording session was set up in the same way as the first study, but with a shorter warm up period (5--10 minutes) and a pre-determined topic for the child-friendly improvisation (`riding bikes', `pets', `breakfast', `birthday cake', `rainy days', or `the library').
+The recording session was set up in the same way as the first experiment, but with a shorter warm up period (5--10 minutes) and a pre-determined topic for the child-friendly improvisation (`riding bikes', `pets', `breakfast', `birthday cake', `rainy days', or `the library').
 All of the talkers were native English speakers, and were recorded in male-female pairs.
@@
 These were rare in the clips.} We made duration and pitch manipulations using PSOLA resynthesis in Praat \citep{Praat}.
-Thus, the \textit{words only} versions of the audio clips had no pitch or durational cues to upcoming turn boundaries, but did have intact lexicosyntactic cues.
+Thus, the \textit{words only} versions of the audio clips had no pitch or durational cues to upcoming turn boundaries, but did have intact lexicosyntactic cues (and residual phonetic correlates of prosody, like intensity).
 
@@
 The \textit{no speech} condition served as a non-linguistic baseline.
-For this condition, we replaced the original clip with multi-talker babble: eight different child-oriented conversations (\textit{not} including the original one), overlaid and cropped to the duration of the video.
-Thus, the \textit{no speech} audio clips lacked any linguistic information to upcoming turn boundaries---the only cue to turn-taking was the opening and closing of the puppets' mouths.
+For this condition, we replaced the original clip with multi-talker babble: Eight different child-oriented conversations (not including the original one), overlaid and cropped to the duration of the video.
+Thus, the \textit{no speech} audio clips lacked any linguistic information to upcoming turn boundaries---the only cue to turn taking was the opening and closing of the puppets' mouths.
 
@@
 Puppets were positioned looking forward to eliminate shared gaze as a cue to turn structure \citep{thorgrimssonUndRev}.
-We took care to match the puppets' mouth movements to the syllable onsets as closely as possible and avoided any mouth movement before the onset of a turn.
+We took care to match the puppets' mouth movements to the syllable onsets as closely as possible, and avoided any mouth movement before the onset of a turn.
 We then added the manipulated audio clips to the puppet video recordings by hand.
@@
 We then created three versions of the experiment, so that each of the six puppet pairs was associated with three different conversation topics across the different versions of the experiment (18 videos in total).
-We ensured that the position of the talkers (left and right) was counterbalanced in each version by flipping the video and audio channels as needed.\footnote{See the videos here: NOTE: Link to be added.}
+We ensured that the position of the talkers (left and right) was counterbalanced in each version by flipping the video and audio channels as needed.\footnote{See the videos here: www.youtube.com/channel/UCGEZQcM9t8Zfjqqi\_B1Q5Sw.}
 
@@
 We coded each turn transition for its linguistic condition (\textit{normal}, \textit{words only}, \textit{prosody only}, and \textit{no speech}) and transition type (question/non-question)\footnote{We coded \textit{wh-}questions as ``non-questions'' for the \textit{prosody only} videos.
-Prototypical \textit{wh-}questions are not marked with a final rising nuclear prosodic contour, but polar questions are \citep{hedberg2010}.}, identified anticipatory gaze switches to the upcoming speaker, and estimated a random baseline of anticipatory switches using the methods from Experiment 1.
+Polar questions had a final rising prosodic contour, but \textit{wh-}questions did not  \citep{hedberg2010}.}, identified anticipatory gaze switches to the upcoming speaker, and estimated a random baseline of anticipatory switches using the methods from Experiment 1.
 
@@
 
-We identified anticipatory gaze switches for all 79 usable turn transitions using the same criteria and algorithm as in Experiment 1.
+We identified anticipatory gaze switches for all 79 usable turn transitions using the same criteria as in Experiment 1.
 We expected to see the most anticipatory switches in the \textit{normal} condition and the least anticipatory switches in the \textit{no speech} condition because they contained the most and least linguistic information, respectively.
@@
 
-As in Experiment 1, we corrected each participant's gaze data by subtracting their random baseline from their actual switching for each turn transition.
-All comparisons below were made with the baseline-corrected values.
-
-% address this!!
-Participants at all ages made anticipatory gaze switches more often than would be expected by chance in almost all conditions, but sometimes by a small margin (Figure \ref{fig:randvsrealEN}).
-For example, anticipatory switches in the \textit{no speech} videos were often only modestly above chance (with the exception of 5- and 6-year-olds).
-Notably, while children performed well with \textit{prosody only} speech from the start, they seemed to only reliably exceed chance in the \textit{words only} condition at ages 3;0 and older, giving some support to the idea that there is an early advantage for prosody over lexicosyntax in children's turn prediction.
+We found that participants at all ages made anticipatory gaze switches more often than would be expected by chance in almost all conditions, but sometimes by a small margin (Figure \ref{fig:randvsrealEN}).
+For example, anticipatory switches in the \textit{no speech} videos were often modestly above chance (with the exception of 5- and 6-year-olds).
+One interpretation of this finding is that children do not use linguistic information to predict upcoming turns---the only cue to turn taking in the \textit{no speech} condition was the alternating mouth movements of the conversing puppets.
+With sufficiently long inter-turn gaps, participants could simply switch their gaze to the responder when the prior talker finishes talking---a strategy that doesn't rely on linguistic information.
+But this interpretation is not compatible with our the effects of linguistic information that do emerge in the data: an early advantage for prosody and a later advantage for lexical question cues.
+
+We saw a generalize early advantage for prosody over lexicosyntax in children's spontaneous turn predictions.
+Notably, while children performed well with \textit{prosody only} speech from the start, they seemed to only reliably exceed chance in the \textit{words only} condition at ages three and older.
+By age three, the general advantage for prosody over lexicosyntax seems to disappear with more equal rates of switching overall for the two partial linguistic conditions.
 
@@
 \end{center}
-\caption{Baseline-corrected proportion anticipatory gaze switches made for actual (solid) and randomly-shuffled (dashed) turn transitions in each condition, and across ages.
+\caption{Baseline-corrected proportion of anticipatory gaze switches made for actual (solid) and randomly-shuffled (dashed) turn transitions in each condition, and across ages.
 Error bars represent the standard error of the mean.}
@@
 \end{center}
-\caption{Baseline-corrected proportion anticipatory gaze switches made for turn transitions in each speech condition, and across ages. Error bars represent the standard error of the mean.} 
+\caption{Baseline-corrected proportion of anticipatory gaze switches made for turn transitions in each speech condition, and across ages. Error bars represent the standard error of the mean.} 
 \label{fig:conditionsEN}
 \end{figure}
-% evidence for integration of diff cues?
-Participants made more anticipatory gaze switches following questions than non-questions, but only in the two lexical conditions (\textit{normal} and \textit{words only} speech).
+
+As in Experiment 1, we made condition, age and transition type comparisons based on baselin-corrected switching values; we corrected each participant's gaze data by subtracting their random baseline from their actual switching for each turn transition.
+
+We found that participants made more anticipatory gaze switches following questions than non-questions, but only in the two lexical conditions (\textit{normal} and \textit{words only} speech).
 The advantage for questions was strongest in the \textit{normal} speech condition, where it emerged early in development.
@@
 \end{center}
-\caption{Baseline-corrected proportion anticipatory gaze switches made for question- (dashed) and non-question- (dotted) turn transitions in each condition, and across ages.
+\caption{Baseline-corrected proportion of anticipatory gaze switches made for question- (dashed) and non-question- (dotted) turn transitions in each condition, and across ages.
 Error bars represent the standard error of the mean.}
@@
 We tested these findings in two linear mixed-effects models of participants' baseline-corrected anticipatory switches---one model each for children and adults.
-The child model included condition (\textit{normal}, \textit{words only}, \textit{prosody only}, and \textit{no speech}), transition type (question vs.
+The child model included condition (\textit{normal}, \textit{words only}, \textit{prosody only}, and \textit{no speech}; \textit{normal} in the intercept), transition type (question vs.
 non-question), age, and gap duration as fixed effects, with full interactions between condition, transition type, and age.
@@
 
-% no main effects is kind of meaningless
-Condition-related differences were driven by age and transition type (Figure \ref{fig:questionsEN}).
-Consequently, the model of children's anticipatory switches revealed no significant main effects of condition, despite large differences in the linguistic information available across conditions.
-One interpretation of this finding is that children do not need linguistic information to predict upcoming turns.
-For example, children still made some anticipatory switches in the \textit{no speech} condition, even though the only cue to turn taking was the alternating mouth movements of the conversing puppets.
-With sufficiently long inter-turn gaps, participants could simply switch their gaze to the responder when the prior talker finishes talking---a strategy that doesn't rely on linguistic information.
-But this interpretation is not compatible with our findings that children make more anticipatory looks after hearing questions, that they do so more often when lexical information is available, and that they do even more so as they age.
-
+The only effects of linguistic condition derived from interactions with age and transition type, despite large differences in the linguistic information available across conditions.
+(Figure \ref{fig:questionsEN}).
 The model of children's data confirmed two three-way effects of age, condition, and transition type.
@@
 The increases related to non-questions and non-lexical conditions were overall smaller and sometimes followed a non-linear trajectory (Figure \ref{fig:questionsEN}).
-As in Study 1, early switches (planned before the end of the prior turn) did not show large increases with age, with adults making the fewest early switches in the three linguistic conditions (Figure \ref{fig:cumulativeEN}).
+As in Experiment 1, early switches (planned before the end of the prior turn) did not show large increases with age, with adults making the fewest early switches in the three linguistic conditions (Figure \ref{fig:cumulativeEN}).
 
@@
 
-The advantage for questions over non-questions in these results reinforces our findings from Study 1.
+The advantage for questions over non-questions in these results reinforces our findings from Experiment 1.
 They suggest that even two-year-olds can reliably distinguish questions from other transition types in third-party conversation, especially when full linguistic information is available.
@@
 Our results suggest that the question effect emerges between ages one and two, but further testing is needed to establish whether this is the case.
-For example, the size and emergence of the question effect for different question types and question cues could differ (\textit{wh-} vs.
+For example, the size and emergence of the question effect for various question types and question cues could differ (\textit{wh-} vs.
 \textit{yes-no}; early vs.
@@
 \end{center}
-  \caption{Proportion of participants showing more anticipatory switches for questions than non-questions in the \textit{normal} and \textit{words only} conditions.}
+  \caption{Proportion of participants showing numerically more anticipatory switches for questions than non-questions in the \textit{normal} and \textit{words only} conditions.}
 \label{tab:questioneffectexp2}
@@
 We found that anticipatory gazes significantly exceeded chance for the \textit{prosody only} condition (\textit{t}(187.677)=-3.004, \textit{p}=.003), but not for the \textit{words only} condition (\textit{t}(229.828)=-0.8307, \textit{p}=.407).
-This finding suggests that there is an early advantage for prosody over lexicosyntax in children under 3;0.
+This finding suggests that there is indeed an early advantage for prosody over lexicosyntax in children under 3;0.
 The advantage for prosody was not driven by a question effect; children and adults did not consistently show more switches after questions than non-questions in the \textit{prosody only} condition (Figure \ref{fig:questionsEN}).
@@
 
-The core aims of Study 2 were to gain better traction on the individual roles of prosody and lexicosyntax in children's turn predictions, and to expand our age range to capture more developmental change.
+The core aims of Experiment 2 were to gain better traction on the individual roles of prosody and lexicosyntax in children's turn predictions, and to expand our age range to capture more developmental change.
 We found that effects of linguistic processing and age were present in this dataset, but both were primarily found within the question effect.
-Children and adults made more anticipatory switches after questions than non-questions, but did so more often when they had greater access to linguistic information, whether increased access came with age or with linguistic condition.
+Children and adults made more anticipatory switches after questions than non-questions, but did so more often when they had greater access to linguistic information, whether increased access came from age or with speech condition.
 Our results hint that lexical information is critical in identifying questions during conversation, especially for children.
-Even so, performance was consistently best in the \textit{normal} condition, in which participants had lexical and prosodic information together.
+Even so, performance was consistently best in the \textit{normal} condition, in which participants could integrate lexical and prosodic information together.
 
@@
 As they got older, children demonstrated above-chance anticipatory switches in more conditions and made more anticipatory switches overall.
-Similar to Experiment 1, children between ages three and five behaved similarly to each other, and comparably to adults, showing small but steady gains in their anticipatory switches with age.
+Like in Experiment 1, children between ages three and five behaved similarly to each other, and comparably to adults, showing small but steady gains in their anticipatory switches with age.
 Though the advantage for questions generally increased with age, the biggest developmental shift in our data occurred between one and two years, when the question effect began to reliably emerge in the two lexical conditions.
@@
 The results from Experiment 2 complement those of Experiment 1, but still leave some open questions.
-For example, we were surprised to find no main effect of linguistic condition.
+For example, we were surprised to find no independent effects of linguistic condition outside of age and transition type interactions.
 In fact, the primary evidence that participants actually \textit{used} linguistic information came from the question effects, which were limited to the two lexical conditions.
@@
 This pattern of development suggests an early importance for prosodic information in children's turn anticipation.
-In contrast, the question effect---a primary driver of anticipatory switches in our data---relies heavily on lexical information.
+In contrast, the question effect---a primary driver of anticipatory switches in our data---seems to rely on lexical information, with some room for further support from prosody.
 So, although prosody may help children early on, it does not, on its own, help children to distinguish questions from non-questions.
@@
 In short, prosody has different roles to play in turn prediction before and after the lexically-driven question effect emerges.
-Importantly, these results do not support a straightforward view of lexicosyntax as the primary linguistic cue for turn predictions (even for adults) \citep{de-ruiter2006, magyari2012}.
-% and in agreement w/ torr, other corpus
-
-%mention again how speech manic was diff from dR
-We controlled linguistic information in Experiment 2 by using phonetically-manipulated speech \citep{de-ruiter2006}.
-While phonetic manipulation gave us complete control over the linguistic signal, it also created speech sounds that children don't usually hear in their natural environment.
+Importantly, these results do not support a straightforward view of lexicosyntax as the primary linguistic cue for turn predictions \citep[even for adults; ][]{de-ruiter2006, magyari2012}, but do support accounts of multiple cue integration \citep{duncan1972, ford1996, hirvenkari2013, torreiraUndRev}.
+
+We controlled linguistic information in Experiment 2 by using phonetically-manipulated speech, extending speech manipulations used in prior work \citep{de-ruiter2006, keitel2013}.
+While phonetic manipulation gave us control over the linguistic signal, it also created speech sounds that children don't usually hear in their natural environment.
 Many prior studies have used phonetically-altered speech with infants and young children \citep[cf.][]{jusczyk2000}, but almost none of them have done so in a conversational context.
@@
 Children could have had trouble processing the \textit{words only} and \textit{prosody only} conditions because they were unfamiliar, and not just because they had less linguistic information available.
-%maybe move to GD?
+
 It would take a different type of stimulus to compare lexical and prosodic information without phonetic manipulation.
-For example, carefully scripted conversations could control for the presence or absence of prosodic and lexicosyntactic cues to turn transition with natural speech.
+For example, carefully scripted or cross-spliced conversations could control for the presence or absence of prosodic and lexicosyntactic cues to turn transition with natural speech.
 But since the non-natural (limited cue) linguistic conditions did still elicit some differences in anticipation, children must have processed at least some linguistic information in the \textit{prosody only} and \textit{words only} conditions.
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
 \section{General Discussion}
 \label{sec:gendisc}
-%Refs again after 1st sentence
-Children begin to develop conversational turn-taking skills long before their first words.
+Children begin to develop conversational turn-taking skills long before their first words \citep{bateson1975, snow1977, jaffe2001, hilbrinkInPrep}.
 As they acquire language, they also acquire the information needed to make accurate predictions about upcoming turn structure.
@@
 
-%"typical" is problematic; also contrast global vs. local findings for pros, lex
 In two experiments investigating children's anticipatory gaze to upcoming speakers, we found evidence that turn prediction develops early in childhood, and that linguistic cues are integrated differently at different ages.
-Children under 3;0 showed an advantage for prosody over lexicosyntax, contrary to the typical finding for adults in prior work \citep{de-ruiter2006}.
-Children 2;0 and older also showed a slight advantage for lexicosyntax in the form of a transition type effect: children made more anticipatory switches following questions than non-questions in conditions where lexical information was present.
+Children under three showed a general advantage for prosody over lexicosyntax, contrary to findings for adults in prior work \citep{de-ruiter2006}.
+Children two and older also showed a slight advantage for lexicosyntax, but only in the form of a transition type effect: they made more anticipatory switches following questions than non-questions in conditions where lexical information was present.
+Two-year-olds thus appeared to be at a developmental transition point, still showing an overall advantage for prosody, but beginning to show an advantage for lexicosyntax within the domain of question-answer sequences.
 We found no evidence that, for children or adults, lexicosyntax alone was ``sufficient'' (equal to full linguistic information) for spontaneous turn prediction \citep[pg.
@@
 \subsubsection{The advantage for questions}
-% remind what advantage was
-In both of our experiments we found a robust advantage for questions in participants' anticipatory switching.
+
+In both of our experiments we found a robust advantage for questions in participants' anticipatory switching because participants made more switches after questions than non-questions.
 Children and adults both treated questions as if they had a special interactive status, and with good reason: Questions make a speaker switch immediately relevant, helping the listener to predict what will happen next (i.e., an answer from the addressee).
@@
 Also, the form of a question can constrain the type of response that will come next (e.g., a location after a \textit{where} question), further increasing the predictability of the interactional sequence.
-% potentially makes more sense w/ pros intro
+
 Children get frequent and early experience with hearing and answering questions.
 Questions make up approximately one third of the utterances children hear, before and after the onset of speech, and even into their preschool years \citep{fitneva2012, henning2005, shatz1979}.\footnote{There is substantial variation in this proportion by individual and socioeconomic class \citep{hart1992}.} Many of the questions directed to young children are ``test' questions---questions that the caregiver already has the answer to (e.g., ``What does a cat say?'').
-Caregivers use questions to get their young children's attention and to ensure that information is in common ground, even if the responses are non-verbal or infelicitous \citep{fitneva2012, snow1977}.
+Caregivers use questions to get their young children's attention and to ensure that information is in common ground, even if the responses are non-verbal or infelicitous \citep{bruner1985, fitneva2012, snow1977}.
 So, in addition to having a special interactive status, questions are a core characteristic of many caregiver-child interactions.
@@
 \subsubsection{Linguistic cues for turn prediction}
-% mention adult counter evidence again, elaborate on prosody a little (and add refs after "knowing..." sentence)
-Prior work with adults has found a consistent and critical role for lexicosyntax in predicting upcoming turn structure, with a more secondary and supporting role for prosody \citep{de-ruiter2006, magyari2012}.
-Knowing that children comprehend more about prosody than lexicosyntax early on, we thought it possible that young children would instead show an advantage for prosody over lexicosyntax.
+
+Prior work with adults has found a consistent and critical role for lexicosyntax in predicting upcoming turn structure \citep{de-ruiter2006, magyari2012}, with more debate over the role of prosody \citep{duncan1972, ford1996, torreiraUndRev}.
+Knowing that children comprehend more about prosody than lexicosyntax early on (see introduction; also see \citealp{speer2009} for a review), we thought it possible that young children would instead show an advantage for prosody over lexicosyntax.
 Our results suggest that there are roles for both prosody and lexicosyntax in children's predictions about upcoming turn structure, but that there are different roles for these information sources at different ages.
@@
 
-Children only showed an advantage for questions in the two lexical conditions, suggesting that lexicosyntax is a critical cue for question identification.
-The question effect emerged by age two, when children's syntactic comprehension and lexical inventories were still relatively limited \citep{clark2009}.
+In both experiments, even young children showed an advantage for questions in conditions where they had lexicosyntax, suggesting that lexicosyntax is a critical cue for question identification.
+The question effect emerged by age two, when children's syntactic comprehension and lexical inventories are still relatively limited \citep{clark2009}.
 That being said, syntactic and lexical markers to questionhood in English are frequent, categorical, and early-occurring in the utterance (e.g., ``do'', \textit{wh}-words, and subject-auxiliary inversion).
@@
 
-\subsection{What's new}
-% mention again that our prosody manipulation included durational changes
-Our findings are only partially consistent with those reported in prior work.
+\subsection{Contributions of the current work}
+
+Our findings are consistent with observational work on children's turn taking.
+Young children are slow turn-takers in spontaneous conversation \citep{casillasUndRev} even though they begin taking turns in infancy \citep{hilbrinkInPrep, jaffe2001}.
+We found that children begin to spontaneously predict upcoming turn structure at age one.
+Children's gaze switches were also most common in the inter-turn gap, suggesting that they expected an immediate response from the addressee.
+Our results therefore support the idea of an early-developed competence for the timing of turn taking.
+But because children do not themselves use adult-like conversational timing until much later \citep[age six; ][]{ervin-tripp1979}, our results also indirectly point toward response planning as a key factor contributing to children's slow responses in real-time interaction \citep{casillasUndRev}.
+
+Our findings are only partially consistent with prior experimental work on turn taking.
 De Ruiter and colleagues (2006) found that adults' accuracy in identifying upcoming turn-ends was not significantly affected when intonation contours were flattened.
@@
 Thus, neither the child nor the adult data support the idea that lexicosyntactic information is equal to full linguistic information in making predictions about upcoming turn structure \citep{de-ruiter2006}, at least in participants' spontaneous looking behavior.
-
-%include subheadings to look at diffs w/ adult vs. chi literature
-Despite extreme differences in linguistic information across conditions, main effects of linguistic information were minimal in our experiments.
+This may partially be due to the fact that we controlled syllable duration in addition to intonation, thereby eliminating two of the most salient cues to prosodic structure.
+
+Despite extreme differences in linguistic information across conditions, independent effects of linguistic information (outside of interactions) were minimal in our experiments.
 But, using a button-press measure with similar stimuli, de Ruiter and colleagues (2006) found robust effects of linguistic condition.
@@
 A button-press method explicitly encourages participants to make an early response, while observer gaze measures are purely passive.
-Participants' spontaneous anticipatory gaze switches may not elicit enough early responses to see significant linguistically-driven differences in prediction.
+Participants' spontaneous anticipatory gaze switches may not elicit enough early responses to see large linguistically-driven differences in prediction.
 Indeed, most switches happened during the inter-turn gap (Keitel et al., 2013; Hirvenkari, 2013; Tice and Henetz, \citeyear{TiceHenetz11}).
@@
 Rather, they are targeting switches in speakership.
-So although button press measures may be better equipped to tell us what participants \textit{can} do in predicting turn ends, they may also emphasize early responding and, consequently, lexicosyntactic cues to a greater extent than we experience in everyday conversation.
-Within observer gaze experiments, questions may be the key to naturally elicit early responses.
+So although button press measures may be better equipped to tell us what participants \textit{can} do in predicting turn ends, they may also emphasize early responding (and consequently, lexicosyntactic cues) to a greater extent than we experience in everyday conversation.
+Within observer gaze experiments, questions may be the key to naturally eliciting early responses.
 We have seen in the current study that questions are interactionally sufficient to motivate early predictions.
@@
 Moving away from button press measures, Keitel and colleagues (2013) used observer gaze to investigate children's turn predictions, finding: (a) no evidence that children under 3;0 make switches above chance, (b) that three-year-olds' switches are significantly affected by flattened intonation, and (c) that adults' switches are not.
-Our current results are consistent with the Keitel et al.
-(2013) study with respect to the general finding that prosody matters for young children's turn predictions.
-But, our results also differ from theirs in a few key ways.
+Our current results are consistent with the general finding from their study, that prosody matters for young children's turn predictions.
+But our results also differ from theirs in a few important ways.
 
@@
 
-Third, we also found that anticipatory looking was primarily driven by question transitions, a pattern that has so far not been discussed in other observer gaze studies, on children or adults (Keitel et al., 2013; Hirvenkari, 2013; Tice and Henetz, \citeyear{TiceHenetz11}).
+Third, we found that anticipatory looking was primarily driven by question transitions, a pattern that has so far not been discussed in other observer gaze studies, on children or adults (Keitel et al., 2013; Hirvenkari, 2013; Tice and Henetz, \citeyear{TiceHenetz11}).
 The question effects in the current study were critical to understanding the role of lexicosyntactic information in participants' turn predictions.
@@
 (2013) found that adults' were unaffected by flattened pitch.
-But, because we flattened pitch for individual speakers and controlled syllable duration, we cannot directly compare our findings to theirs.
-
-% move to the front of this section
-Our findings are consistent with observational work on children's turn taking.
-Young children are slow turn-takers in spontaneous conversation \citep{casillasUndRev} even though they begin taking turns in infancy \citep{hilbrinkInPrep, jaffe2001}.
-We found that children begin to spontaneously predict upcoming turn structure at age one.
-Children's gaze switches were also most common in the inter-turn gap, suggesting that they expected an immediate response from the addressee.
-Our results therefore support the idea of an early-developed competence for the timing of turn-taking.
-But because children do not themselves use adult-like conversational timing until much later \citep[age six; ][]{ervin-tripp1979}, our results also indirectly point toward response planning as a key factor contributing to children's slow responses in real-time interaction \citep{casillasUndRev}.
+But again, because we flattened pitch for individual speakers and controlled syllable duration, we are not able to directly compare our findings to theirs.
 
@@
 Although this decision allowed us to follow prior work more closely \citep{de-ruiter2006, keitel2013}, it also limits our interpretation of children's behavior.
-Low-pass filtered and robotic-like speech are not typical within children's speech environment, and so it may have been unfair to compare the limited cue conditions in Experiment 2 to the normal speech condition, which not only had full linguistic cues, but was also more typically present in children's speech environments.
-One alternative to explore for the future would be scripted speech, or speech that is carefully selected from a large corpus, given a set of specific linguistic constraints.
-% or cross-splicing
+Low-pass filtered and robotic-like speech are not typical within children's speech environment, and so it may have been unfair to compare the limited cue conditions in Experiment 2 to the normal speech condition, which not only had full linguistic cues, but is also more typically present in children's speech environments.
+One alternative to explore for the future would be scripted speech, cross-spliced speech, or speech that is carefully selected from a large corpus, given a set of specific linguistic constraints.
 
 Relatedly, since we used improvised conversation recordings, we did not control how often the speakers switched back and forth, or how often a switch occurred after a pause in speech.
-It is possible that, to some extent, children could have formed expectations about the turn-structure on the basis of the frequent switching alone \citep[see, e.g., ][]{thorgrimssonUndRev}.
+It is possible that, to some extent, children could have formed expectations about the turn structure on the basis of the frequent switching alone \citep[see, e.g., ][]{thorgrimssonUndRev}.
 If this statistical expectation were the only thing leading children's anticipatory gaze switches, we would have predicted more uniformity in looking behavior across the conditions and no major effects of linguistic signal, age, or transition type.
@@
 Since most studies, even those using a wider switching window \citep[e.g., ][]{hirvenkari2013} find that switches primarily happen close to or immediately after the inter-turn gap begins, the approach used here and in Keitel et al.
-(2013) probably captures most of the switching behavior.
+(2013) probably captures most of participants' turn-relevant switching behavior.
 But if early looks do indeed increase as children get older and have more access to early syntactic cues, it would be helpful to develop a second measure of very early anticipatory looks.
@@
 The simultaneous acquisition and integration of interactive and linguistic skills is no easy task, as evidenced by children's protracted development of turn-taking skills.
-In the current study we have investigated children's language acquisition through the lens of turn-taking, thereby bringing into focus children's application of linguistic knowledge in real-time interaction.
+In the current study we have investigated children's language acquisition through the lens of turn taking, thereby bringing into focus children's application of linguistic knowledge in real-time interaction.
 We found that children acquire basic predictive skills early on, but that the integration of linguistic cues takes time and is primarily driven by sequences of action---in our case, by questions and their answers.
-Thus, even at the level of linguistic integration, children's attention during conversation is driven by goals from real-time interaction, the principle source and destination for their language learning.
-% increase link between findings and last sentence, or don't make the jump (can just leave out).
 
@@
 
-We gratefully acknowledge the parents and children at Bing Nursery School, Children's Discovery Museum of San Jose.
-This work was supported by an ERC grant XYZ, an NSF GRFP to the first author, and a Merck Foundation Fellowship to the second author.
-An earlier version of these data and analyses were initially presented to conference audiences in \cite{casillas2012} and \cite{casillas2013}.
+We gratefully acknowledge the parents and children at Bing Nursery School and the Children's Discovery Museum of San Jose.
+This work was supported by an ERC Advanced Grant to Stephen C.
+Levinson (269484-INTERACT), NSF graduate research and dissertation improvement fellowships to the first author, and a Merck Foundation fellowship to the second author.
+Earlier versions of these data and analyses were presented to conference audiences \citep{casillas2012, casillas2013}.
+We also thank Tania Henetz, Francisco Torreira, Stephen C.
+Levinson, Eve V.
+Clark, and the First Language Acquisition group at Radboud University for their feedback on earlier versions of this work.
 

### 2014-09-30 paper/CasillasFrank-Anticip.tex
ref order, experiment > study, age expressions

@@
 \usepackage{graphicx}
-% \usepackage{lineno}
+\usepackage{lineno}
 \usepackage{amssymb}
@@
 
-%\linenumbers
+\linenumbers
 
@@
 Sequential, turn-based structures in conversation are strikingly uniform across language communities and linguistic modalities.
-Turn-taking behaviors are also cross-culturally consistent in their basic features and the details of their implementation \citep{stivers2009, dingemanse2013, de-vosInPrep}.
+Turn-taking behaviors are also cross-culturally consistent in their basic features and the details of their implementation \citep{de-vosInPrep, dingemanse2013, stivers2009}.
 How does this ability develop?
 
-Children participate in sequential coordination with their caregivers starting at three months of age---before they can rely on any linguistic cues in taking turns \citep[see, among others, ][]{bateson1975, snow1977, jaffe2001, hilbrinkInPrep}.
+Children participate in sequential coordination with their caregivers starting at three months of age---before they can rely on any linguistic cues in taking turns \citep[see, among others, ][]{bateson1975, hilbrinkInPrep, jaffe2001, snow1977}.
 Of course, infant turn taking is different from adult turn taking in several ways.
@@
 Children's first turn-like structures appear as early as two to three months in proto-conversation with their caregivers \citep{bruner1975, bruner1985}.
-During proto-conversations, caregivers interact with their infants as if they were capable of making meaningful contributions; they take every look, vocalization, arm flail and burp as ``utterances'' in the joint discourse \citep{bateson1975, snow1977, jaffe2001}.
+During proto-conversations, caregivers interact with their infants as if they were capable of making meaningful contributions; they take every look, vocalization, arm flail and burp as ``utterances'' in the joint discourse \citep{bateson1975, jaffe2001, snow1977}.
 Infants catch onto the structure of proto-conversations quickly.
@@
 
-We recruited 74 children between ages 3;0--5;11 and 11 undergraduate adults to participate in the study.
+We recruited 74 children between ages 3;0--5;11 and 11 undergraduate adults to participate in the experiment.
 Our child sample included 19 three-year-olds, 32 four-year-olds, and 23 five-year-olds, all enrolled in a local nursery school.
@@
 We used a dancing Elmo for the children's calibration image.
-During the study, participants watched all six 30-second conversation videos with 15--30 second filler videos in-between them (e.g., running puppies, singing muppets, flying bugs).
+During the experiment, participants watched all six 30-second conversation videos with 15--30 second filler videos in-between them (e.g., running puppies, singing muppets, flying bugs).
 The first and last conversations were in American English and the intervening conversations were Hebrew, Japanese, German, and Korean.
@@
 
-As in the first study, we recorded participants' eye movements as they watched six short videos of dyadic conversation, and then analyzed their anticipatory glances from the current speaker to the upcoming speaker at points of turn transition.
+As in the first experiment, we recorded participants' eye movements as they watched six short videos of dyadic conversation, and then analyzed their anticipatory glances from the current speaker to the upcoming speaker at points of turn transition.
 
@@
 \subsubsection{Participants}
-We recruited 27 undergraduate adults and 129 children between ages 1;0--7;0 to participate in our study.
+We recruited 27 undergraduate adults and 129 children between ages 1;0--6;11 to participate in our experiment.
 We recruited our child participants from the Children's Discovery Museum in San Jose, California, targeting approximately 20 children for each of the six 1-year age groups (range$=$20--23).
@@
 \subsection{Procedure}
-We used the same experimental apparatus and procedure as in the first study.
+We used the same experimental apparatus and procedure as in the first experiment.
 Each participant watched six puppet videos in random order, with five 15--30 second filler videos placed in-between (e.g., running puppies, moving balls, flying bugs).
 Three of the puppet videos had \textit{normal} audio while the other three had \textit{words only}, \textit{prosody only}, and \textit{no speech} audio.
-This study required no special instructions so the experimenter immediately began each session with calibration (same as before) and then stimulus presentation.
+This experiment required no special instructions so the experimenter immediately began each session with calibration (same as before) and then stimulus presentation.
 The entire experiment took less than five minutes.
@@
 Finally, we confirmed that young children showed a benefit for prosody over lexicosyntax.
-We conducted two \textit{post-hoc} \textit{t}-tests for children under 3;0, comparing their anticipatory switching rate against their estimated random switching rate in the \textit{prosody only} and \textit{words only} conditions.
+We conducted two \textit{post-hoc} \textit{t}-tests for children under three, comparing their anticipatory switching rate against their estimated random switching rate in the \textit{prosody only} and \textit{words only} conditions.
 We found that anticipatory gazes significantly exceeded chance for the \textit{prosody only} condition (\textit{t}(187.677)=-3.004, \textit{p}=.003), but not for the \textit{words only} condition (\textit{t}(229.828)=-0.8307, \textit{p}=.407).
-This finding suggests that there is indeed an early advantage for prosody over lexicosyntax in children under 3;0.
+This finding suggests that there is indeed an early advantage for prosody over lexicosyntax in children under three.
 The advantage for prosody was not driven by a question effect; children and adults did not consistently show more switches after questions than non-questions in the \textit{prosody only} condition (Figure \ref{fig:questionsEN}).
@@
 \label{sec:gendisc}
-Children begin to develop conversational turn-taking skills long before their first words \citep{bateson1975, snow1977, jaffe2001, hilbrinkInPrep}.
+Children begin to develop conversational turn-taking skills long before their first words \citep{bateson1975, hilbrinkInPrep, jaffe2001, snow1977}.
 As they acquire language, they also acquire the information needed to make accurate predictions about upcoming turn structure.
@@
 De Ruiter and colleagues (2006) found that adults' accuracy in identifying upcoming turn-ends was not significantly affected when intonation contours were flattened.
-Keitel and colleagues (2013) found the same for adults' anticipatory gaze switches, only finding that intonation flattening affected children's switches at age 3;0 (with switches by younger children never exceeding chance).
+Keitel and colleagues (2013) found the same for adults' anticipatory gaze switches, only finding that intonation flattening affected children's switches at 36 months (with switches by younger children never exceeding chance).
 Our results, in contrast, suggest that children's anticipatory switches emerge much earlier, and rely on prosody before they do lexicosyntax.
@@
 
-Moving away from button press measures, Keitel and colleagues (2013) used observer gaze to investigate children's turn predictions, finding: (a) no evidence that children under 3;0 make switches above chance, (b) that three-year-olds' switches are significantly affected by flattened intonation, and (c) that adults' switches are not.
+Moving away from button press measures, Keitel and colleagues (2013) used observer gaze to investigate children's turn predictions, finding: (a) no evidence that children under three make switches above chance, (b) that three-year-olds' switches are significantly affected by flattened intonation, and (c) that adults' switches are not.
 Our current results are consistent with the general finding from their study, that prosody matters for young children's turn predictions.
@@
 
-First, we found that children begin making above-chance anticipatory switches much earlier than 3;0---our one-year-olds were already looking above chance in two of our four linguistic conditions in Experiment 2.
+First, we found that children begin making above-chance anticipatory switches much earlier than three---our one-year-olds were already looking above chance in two of our four linguistic conditions in Experiment 2.
 This earlier emergence of anticipatory switching could be due to our more child-friendly stimuli, which we designed to be engaging and to mimic the kind of speech children often experience interactionally.

### 2015-05-01 paper/CasillasFrank-Anticip-Edits.tex

--- 2014-09-30 paper/CasillasFrank-Anticip.tex
@@
 \usepackage{float}
+\usepackage{setspace}
+
+%\doublespacing
 
@@
 
-\title{Tracking and predicting turn structure during language acquisition}
+\title{The development of children's ability to track and predict turn structure in conversation}
 
-\author[MPI,StanfordLX]{Marisa Casillas\corref{cor1}}
-\address[MPI]{Max Planck Institute for Psycholinguistics, Nijmegen}
-\address[StanfordLX]{Department of Linguistics, Stanford University}
-\cortext[cor1]{Corresponding author}
+\author[MPI]{Marisa Casillas\corref{cor1}}
+\address[MPI]{Max Planck Institute for Psycholinguistics}
+%\address[StanfordLX]{Department of Linguistics, Stanford University}
+\cortext[cor1]{Corresponding author. Address: Max Planck Institute for Psycholinguistics, Wundtlaan 1, 6525XD, Nijmegen, The Netherlands. \textit{E-mail address:} marisa.casillas@mpi.nl.}
+%\newline \newline Address M.C. Frank: Stanford University, Jordan Hall Rm. 420-278, 450 Serra Mall, Stanford, CA 94305. \textit{E-mail address:} mcfrank@stanford.edu
 
 \author[StanfordPSY]{Michael C. Frank}
-
 \address[StanfordPSY]{Department of Psychology, Stanford University}
@@
 Children begin developing turn-taking skills in infancy, but take several years to assimilate their growing knowledge of language with this ability.
-In two eye-tracking experiments with children across a wide developmental range, we measured spontaneous predictions about upcoming speaker change while controlling the amount of linguistic information available.
+In two eye-tracking experiments with children across a wide developmental range (1;0--6;11), we measured spontaneous predictions about upcoming speaker change while controlling the amount of linguistic information available.
 We found that children already predicted upcoming turns at age one, but they integrated linguistic cues differently at different ages.
@@
 We found no evidence that lexicosyntax alone guides turn prediction---instead, participants' performance was best overall with access to lexicosyntax and prosody together.
-Our results point to the importance of studying children's linguistic processing in the context of conversation.
 \end{abstract}
@@
 Spontaneous conversation is a universal context for using and learning language.
-Like other types of human interaction, it is organized at its core by the roles and goals of its participants.
-But what sets conversation apart is its structure: sequences of interconnected, communicative actions that take place across alternating turns at talk.
-Sequential, turn-based structures in conversation are strikingly uniform across language communities and linguistic modalities.
-Turn-taking behaviors are also cross-culturally consistent in their basic features and the details of their implementation \citep{de-vosInPrep, dingemanse2013, stivers2009}.
-How does this ability develop?
+Like other types of human interaction (e.g., driving, playing a sport, duetting), it is organized at its core by the roles and goals of its participants.
+But what sets it apart from other modes of interaction is its structure: sequences of interconnected, communicative acts that take place within alternating turns at talk.
+Turn-based structures in conversation are strikingly consistent across linguistic communities and  modalities \citep{de-vosInPrep, dingemanse2013, stivers2009}.
+In particular, the timing of transitions from one speaker to the next is consistently rapid, averaging between 0 and 200 msec.
+To achieve such rapid turn transitions, addressees must engage in predictive language comprehension and early response planning, motivating the idea that linguistic and interactional processing are deeply entwined in real-time language use, at least for adult speakers \citep{levinson2013}.
 
-Children participate in sequential coordination with their caregivers starting at three months of age---before they can rely on any linguistic cues in taking turns \citep[see, among others, ][]{bateson1975, hilbrinkInPrep, jaffe2001, snow1977}.
-Of course, infant turn taking is different from adult turn taking in several ways.
-Infant turn taking is heavily scaffolded by caregivers, has distinct timing in comparison to adult turn taking, and lacks semantic content \citep{hilbrinkInPrep, jaffe2001}.
-However, children's early, turn-structured social interactions are presumably a critical precursor to their conversational turn taking.
-Along these lines, early non-verbal interactions establish the protocol by which children come to use language with others.
-And as they acquire language, they also come to integrate it into the preverbal turn-taking systems.
+Infants and young children manage to participate in their everyday interactions with quite limited linguistic repertoires---children begin taking turns (of a sort) long before their first words emerge \citep[age 3--4 months: see, e.g.,][]{bateson1975, hilbrinkUR, jaffe2001, snow1977} and, once they start using language, it takes a few years before they can give verbal responses with rapid turn transitions \citep{casillas2015, garvey1984, ervin-tripp1979}.
+Little is known about when and how children's linguistic knowledge begins to influence their  interactional behavior (as it does for adults).
+One idea is that they start out by using non-linguistic cues to make predictions about turn-taking behavior, and then later integrate linguistic cues as they acquire them.
+In the current study we investigate when children begin to make predictions about upcoming turn structure, and how they integrate language into their predictions as they grow older.
 
-In this study, we investigate when children begin to make predictions about upcoming turn structure in conversation, and how they integrate language into their predictions as they grow older.
-In what follows, we first give a basic review of turn-taking research and the state of current knowledge about adult turn prediction.
-We then discuss recent work on the development of turn-taking skills before turning to the details of our own study.
-
-\subsection{Turn taking}
-
-Turn taking itself is not unique to conversation.
-Many other human activities are organized around sequential turns at action.
-Traffic intersections and computer network communication both use turn-taking systems.
-Children's early games (e.g., give-and-take, peek-a-boo) have built-in, predictable turn structure \citep{ratner1978, ross1987}.
-Even monkeys take turns: Non-human primates such as marmosets and Campbell's monkeys vocalize contingently with each other in both natural and lab-controlled environments \citep{lemasson2011, takahashi2013}.
-In all these cases, turn taking serves as a protocol for interaction, allowing the participants to coordinate with one another through sequences of contingent action.
-
-Conversation distinguishes itself from non-conversational turn-taking behaviors by the complexity of the turn sequencing involved.
-In the examples above (traffic, games, and monkeys) the set of sequence and action types is far more limited and predictable than what we find in everyday talk.
-For example, conversational turns come grouped into semantically-contingent sequences of action.
-The groups can span turn-by-turn exchanges (e.g., simple question--response, ``How are you?''--``Fine.'') or sequence-by-sequence exchanges (e.g., reciprocals, ``How are you?''--``Fine, and you?''--``Great!'').
-Sequences of action drive the conversation forward into the next, relevant sequences of talk (e.g., ''And you?''--``Great!''--``Why's that?''; \citealp{schegloff2007}).
-To take a turn, participants need to make predictions about what conversational content will be relevant next.
-In some cases, relevant next turns are somewhat obvious (e.g., question--response) while, in other cases, there are multiple relevant next actions to choose from, or no obvious next action at all (e.g., after a closed sequence).
-
-Despite this complexity, conversational turn taking is often precise in its timing.
-Across a diverse sample of conversations in 10 languages, one study found a consistent average transition time of 0--200 msec at points of speaker switch \citep{stivers2009}.
-Experimental results and current models of speech production suggest that it takes approximately 600 msec to produce a content word, and even longer to produce a simple utterance \citep{griffin2000, levelt1989}.
-So in order to achieve 200 msec turn transitions, speakers must begin formulating their response before the prior turn has ended \citep{levinson2013}.
-Moreover, to formulate their response early on, speakers must track and anticipate what types of response might become relevant next.
-They also need to predict the content and form of upcoming speech so that they can launch their articulation at exactly the right moment.
-Prediction thus plays a key role in timely turn taking.
-
-\subsection{Adults' turn prediction}
-
-Adults have a lot of information at their disposal to help make accurate predictions about upcoming turn content.
-Lexical, syntactic, and prosodic information (e.g., \textit{wh}- words, subject-auxiliary inversion, and list intonation) can all inform addressees about upcoming linguistic structure \citep{de-ruiter2006, duncan1972, ford1996, torreiraUndRev}.
-Non-verbal cues (e.g., gaze, posture, and pointing) often appear at turn-boundaries and can sometimes act as late indicators of an upcoming speaker switch \citep{rossano2009, stivers2010}.
-Additionally, the sequential context of a turn can make it clear what will come next: answers after questions; thanks or denial after compliments, et cetera \citep{schegloff2007}.
+\subsection{Adults turn prediction}
 
@@
 De Ruiter and colleagues \citeyearpar{de-ruiter2006} asked participants to listen to snippets of spontaneous conversation and to press a button whenever they anticipated that the current speaker was about to finish her turn.
-The speech snippets were controlled for the amount of linguistic information present; some were normal, but others had flattened pitch, low-pass filtered speech, or further manipulations.
-De Ruiter and colleagues found that, with pitch-flattened speech, the timing of participants' button responses was comparable to their timing with the full linguistic signal.
-But, when no lexical information was available, participants' responses were significantly earlier than the full linguistic signal.
+The speech snippets were controlled for the amount of linguistic information present; some were normal, but others had flattened pitch (no intonation), low-pass filtered speech (no lexico-syntactic cues), or further manipulations.
+The timing of participants' button responses with pitch-flattened speech was comparable to their timing with the full linguistic signal but, when no lexical information was available, was significantly earlier than responses for the full linguistic signal.
 The authors concluded that lexicosyntactic information\footnote{The ``lexicosyntactic'' condition only included flattened pitch and so was not exclusively lexicosyntactic---the speech would still have residual prosodic structure, including syllable duration and intensity.} was necessary and possibly sufficient for turn-end projection, while intonation was neither necessary nor sufficient.
 Congruent evidence comes from studies varying the predictability of lexicosyntactic and pragmatic content: Adults anticipate turn ends better when they can more accurately predict the exact words that will come next (\citealp{magyari2012}; see also \citealp{magyariUndRev}).
-They can also identify speech acts within the first word of an utterance \citep{gisladottirUndRev}, allowing them to start planning their response at the first moment possible \citep{bogelsmagyariInPrep}.
 
-The role of prosody for adult turn-prediction is still a matter of debate.
-De Ruiter and colleagues' (2006) experiment only tested the role of intonation, which is only a partial index of prosody.
-Prosodic structure is also tied closely to the syntax of an utterance, and so the two linguistic signals are difficult to control independently \citep{ford1996}.
-\citet*{torreiraUndRev} used a combination of button-press and verbal responses to investigate the relationship between lexicosyntactic and prosodic cues in turn-end prediction.
-Critically, their stimuli were cross-spliced so that each item had full prosodic cues to accompany the lexicosyntax.
-Because of the splicing, they were able to create items that had syntactically-complete units with no intonational phrase boundary at the end.
-Participants never verbally responded or pressed the ``turn-end'' button when hearing a syntactically-complete phrase without an intonational phrase boundary.
-And when intonational phrase boundaries were embedded in multi-utterance turns, participants were tricked into pressing the ``turn-end'' button 29\% of the time.
-Their results suggest that listeners actually do rely on prosodic cues to execute a response (see also de \citet{de-ruiter2006}:525).
-These experimental findings corroborate other corpus and experimental work promoting a combination of cues (lexicosyntactic, prosodic, and pragmatic) as key for accurate turn-end prediction \citep{duncan1972, ford1996, hirvenkari2013}.
+The role of prosody (rhythm, intensity, and intonation) for adult turn-prediction is still a matter of debate.
+Prosodic structure is closely tied to the syntax of an utterance, making the two linguistic signals difficult to control independently \citep{ford1996}.
+\citet*{torreiraUndRev} measured the turn-end predictive effects of prosody and syntax by splicing together separate segments of speech to create syntactically-complete stimuli with and without intonational phrase boundaries at the end.
+In their study, participants never responded (verbally or with a ``turn-end'' button-press)  without hearing an intonational phrase boundary.
+On the other hand, when they embedded intonational phrase boundaries in multi-utterance turns, participants were tricked into pressing the ``turn-end'' button on 29\% of trials.
+Listeners then \textit{do} rely on prosodic cues to predict upcoming turn structure (see also \citet{de-ruiter2006}:525), perhaps using them in combination with other linguistic cues (lexicosyntactic, and pragmatic) for greater accuracy in response timing \citep{duncan1972, ford1996, levinson2013, hirvenkari2013}.
 
-In sum, adults accurately and spontaneously make predictions about upcoming turn structure.
-Their predictions rely on a sophisticated body of knowledge about linguistic structure, non-verbal signals, and social actions.
-Knowing this, we could expect that children's acquisition of turn-taking skills is closely tied to their knowledge about language, gaze, gesture, and social cues.
-But children's turn taking starts early in infancy, long before first words or gestures emerge.
-So a primary role for lexicosyntactic cues doesn't fit well with children's pre-verbal turn taking.
+Taken together, these corpus-based and experimental findings on human conversation suggest that linguistic cues may be critical to smooth turn transitions in adult conversation.
+If so, we need an account of how children learn to integrate linguistic cues into their interactive turn taking over the course of language acquisition.
 
-\subsection{Children's turn prediction}
+\subsection{Child turn prediction}
 
@@
 
-The majority of work on children's early turn taking has focused on observations of spontaneous interaction.
-Children's first turn-like structures appear as early as two to three months in proto-conversation with their caregivers \citep{bruner1975, bruner1985}.
-During proto-conversations, caregivers interact with their infants as if they were capable of making meaningful contributions; they take every look, vocalization, arm flail and burp as ``utterances'' in the joint discourse \citep{bateson1975, jaffe2001, snow1977}.
+Children's first turn-like structures appear as early as two to three months after birth, in proto-conversation with their caregivers \citep{bruner1975, bruner1985}.
+During proto-conversations, caregivers interact with their infants as if they were capable of making meaningful contributions; they take every look, vocalization, arm flail, and burp as ``utterances'' in the joint discourse \citep{bateson1975, jaffe2001, snow1977}.
 Infants catch onto the structure of proto-conversations quickly.
-By three to four months they notice disturbances to the contingency of their caregivers' response and, in reaction, change the rate and quality of their vocalizations \citep{k-bloom1988, masataka1993}.
-Infants at this age also notice changes to social contingency outside of turn structure.
-In the Still Face paradigm, caregivers interact with their infants and then suddenly halt, taking on a neutral expression with a sustained gaze.
-When faced with this sudden disappearance of social contingency, infants three months and older try a range of methods to reinitiate the interaction, such as vocalization, reaching, and smiling before looking away or getting upset \citep{rochat1998, toda1993}.
+By three to four months they notice disturbances to the contingency of caregiver responses and, in reaction, change the rate and quality of their vocalizations \citep{k-bloom1988, masataka1993}.
 
-The timing of children's responses to their caregivers' speech shows a non-linear pattern of fall-rise-fall from early infancy to middle childhood.
-A recent study by Hilbrink et al.
-(in preparation) finds that infants' turn timing at three months is often too early or too late: they start vocalizing in overlap on nearly 40\% of their caregivers' turns, and their non-overlapped vocalizations come after an average inter-turn silent gap of 675 msec (adult average: 200 msec).
-Between four and nine months, children begin to reduce the number of turns happening in overlap while also improving on their average response latency.
-But then, later on, children's response latencies slow down again, peaking at average gaps of 1100 msec at nine months, with only very gradual improvement after that \citep{hilbrinkInPrep}.
-While children's avoidance of overlap is nearly adult-like by nine months, the timing of their non-overlapped responses stays much longer than the 200 msec standard for the next several years \citep{garvey1984, ervin-tripp1979}.
+The timing of children's responses to their caregivers' speech shows a non-linear pattern from early infancy to middle childhood.
+Infants' turn timing at three months is often too early or too late; they overlap their vocalizations with caregiver speech nearly 40\% of the time and come in after inter-turn silent gaps that average around 675 msec (adults: 200 msec; \citealp{hilbrinkUR}).
+Between four and nine months children reduce the number of turns happening in overlap.
+At the same time, children's silent gaps temporarily begin to get shorter, before then getting longer again, peaking at an average duration of 1100 msec at nine months.
+After nine months, children's gap durations continue getting shorter again, but very gradually \citep{casillas2015, hilbrinkUR}; while children's avoidance of overlap is nearly adult-like by nine months, the timing of their non-overlapped responses stays much longer than the 200 msec standard for the next several years \citep{garvey1984, ervin-tripp1979}.
 
-The protracted development of children's timing may be attributable to their linguistic development: Taking turns on time is easier when the response is a simple vocalization rather than a linguistic utterance.
-Integrating language into the turn-taking system may be one major factor in children's delayed responses \citep{casillasUndRev}.
-If response planning (i.e., language production) is the primary hurdle in children's spontaneous turn taking, we should find evidence that children understand turn-taking behaviors before they are able to produce the behaviors themselves; this hypothesis has been recently explored in experimental settings.
+The gradual development of children's turn timing after age nine months may be attributable to linguistic development; taking turns on time is easier when the response is a simple vocalization rather than a linguistic utterance.
+If response planning (i.e., language production) is the primary hurdle in children's spontaneous turn taking, children might understand turn-taking behaviors before they are able to produce the behaviors themselves; this hypothesis has been explored recently in experimental contexts.
 
@@
 
-Children begin to develop specific expectations about conversational behavior before they begin to speak.
-Sometime between four and six months, children begin to attend differently to face-to-face and back-to-back conversation; six-month-olds follow conversational speakers more with their gaze when at least one speaker is looking at the other \citep{augusti2010}.
+Children develop specific expectations about conversational behavior before they begin to speak.
+For example, sometime between four and six months, children begin attending differently to face-to-face and back-to-back conversation; six-month-olds follow conversational speakers more with their gaze when at least one speaker is looking at the other \citep{augusti2010}.
 At ten months, infants expect people to look and talk at other people, and not to objects \citep{beier2012}.
-At twelve months infants expect to see responses to verbal (but not non-speech) utterances in face-to-face contexts \citep{thorgrimssonUndRev}.
 
 There are mixed results regarding when children begin to anticipate turn structure in conversation.
-One study found that 12-month-olds make more predictive gaze shifts to a responder while watching human verbal conversation compared to conversation-like interactions with objects \citep{bakker2011}, but another only found a similar effect at 36 months \citep{hofsten2009}.
-Neither of these two studies had baselines to which the turn-relevant looking behavior could be compared, however.
-A baseline measurement is critical because there may be developmental differences in gaze shifting between conversational participants, even if the shifting is not related to turn structure.
-Such developmental differences could produce artifactual changes in measures of turn-contingent shifting.
+One study found that 12-month-olds make more predictive gaze shifts to a responder while watching human verbal conversation compared to ``conversation'' between two objects \citep{bakker2011}, but another study only found a similar effect at 36 months \citep{hofsten2009}.
+Neither study compared children's looking behavior to a random baseline, even though baseline estimates can help account for non-turn-contingent developmental differences in gaze shifting.
+A third study, testing turn-transition prediction with 6-, 12-, 24-, and 36-month-olds (and adult controls), found that children's rate of anticipatory gaze switching to the responder was only greater than chance for 36-month-olds and adults \citep{keitel2013}.
+In their study, participants saw two types of conversation videos: one with pitch-flattened speech and one with normal speech.
+The 36-month-olds, but not the adults, were affected by a lack of intonation contours, leading the authors to conclude that children's ability to predict upcoming turn structure relies on their ability to retrieve lexicosemantic meaning from the stimuli (emerging around 36 months).
+They suggest that, once children acquire sophisticated, adult-like language comprehension (sometime after 36 months),  intonation takes a secondary role in turn prediction (but also see \citealp{keitel2015}), consistent with other findings on adult turn-end prediction \citep{ruiter2006}.
 
-Keitel and colleagues \citeyearpar{keitel2013} addressed the random baseline issue in their study of 6-, 12-, 24-, and 36-month-olds.
-They asked participants to watch short videos of conversation, and tracked their eye movements at points of speaker change.
-They found that children's anticipatory gaze frequency was only greater than chance for 36-month-olds and adults.
-Their study was the first to focus on the role of linguistic processing in children's turn predictions.
-They showed their participants two types of conversation videos: one normal and one with flattened pitch (i.e., with flattened intonation contours), finding that only 36-month-olds were affected by a lack of intonation contours.
-The adult control group made equal numbers of anticipatory looks in the videos, with and without intonation contours, consistent with prior adult findings \citep{de-ruiter2006}.
-Keitel and colleagues concluded that children's ability to predict upcoming turn structure relies on their ability to comprehend the stimuli (emerging around 36 months), especially with respect to semantic access.
-They also suggest that intonation takes a secondary role in turn prediction, but only \textit{after} children acquire more sophisticated, adult-like language comprehension systems (sometime after 36 months).
+An early role for intonation is sensible given that children begin acquiring prosodic knowledge early in infancy.
+Children can distinguish their native language's rhythm type from others (e.g., syllable-timed vs.
+stress-timed) 2--5 days after birth \citep{mehler1988, moon1993, nazzi2003} and, before 8 months, they can already identify prosodic boundaries and use them to parse information in the speech stream \citep{johnson2001, jusczyk1993, jusczyk1995, morgan1995, soderstrom2003}.
+In comparison, 8-month-olds still have quite small lexical inventories and, even at 12--14 months, children have just begun to work out syntactic categories \citep{bergelson2013, shi2010}.
+Two-month-olds notice changes in word order, but this ability appears to rely on prosodic cueing \citep{mandel1996}.
+Generally speaking then, we might predict from children's linguistic development that children might use prosodic cues in their turn-taking predictions before they use lexicosyntactic cues.
+Keitel et al.'s \citeyearpar{keitel2013, keitel2015} findings support this idea, but suggest that children's ability to make predictions about upcoming turn structure only emerges at 36 months.
 
 Although the Keitel et al.
-\citeyearpar{keitel2013} study constitutes a substantial advance over previous work, it nevertheless has several limitations.
-Because these limitations directly inform our own study design, we review them in some detail.
-First, their estimates of baseline gaze frequency (``random'' in their terminology) were computed from switches that occurred specifically while the speaker's turn was ongoing; in contrast, estimates of anticipatory shifting frequency were computed from switches that occurred within the time window immediately surrounding inter-turn gaps.
-Assuming that gaze switches happen most often during the inter-turn gap \citep{hirvenkari2013}, the switches made during ongoing speech (their baseline) would be minimized.
-This baseline thus maximized the possibility of finding a difference between ``random'' and anticipatory gaze frequencies.
-A more conservative baseline would be to compare participants' looking behavior at turn transitions to their looking behavior during randomly-selected windows of time throughout the stimulus.
-We follow this conservative approach in our work.
-
-Second, the conversation stimuli they used were somewhat unusual.
-The average gap between turns was 900 msec, which is much longer than typical adult timing, where gaps average around 200 msec \citep{stivers2009}.
-The speakers in the videos were also asked to minimize their movements while performing a scripted and adult-directed conversation, which would have created a somewhat unnatural stimulus.
-In addition, in order to produce more naturalistic conversation, it would have been ideal to localize sources for the two voices in the video (i.e., to have the voices come out of separate left and right speakers).
-But both voices were recorded and played back on the same audio channel, which may have made it more difficult to distinguish the two talkers.
-Again, we attempt to address these issues in our current study.
-
-Despite these minor methodological issues, the Keitel et al.
-\citeyearpar{keitel2013} study still demonstrates intriguing age-based differences in children's ability to predict upcoming turn structure, and the results suggest that both semantic and intonational development \textit{do} play a role in children's looking patterns.
-Our current work thus takes this paradigm as our starting point.\footnote{See also \citet{casillas2012, casillas2013}.}
-
-\subsubsection{Prosodic development}
-The roles of prosody and lexicosyntax in children's turn predictions are currently unknown, but children understand more about prosody early in life.
-Children begin to acquire prosody in the womb, and can distinguish their native language's rhythm type from others (e.g., syllable-timed vs.
-stress-timed) 2--5 days after birth \citep{mehler1988, moon1993, nazzi2003}.
-Beginning between four and five months, infants prefer pauses in speech to be inserted at prosodic boundaries, and by 6 months they can start using prosodic markers to pick out sub-clausal syntactic units \citep{jusczyk1995, soderstrom2003}.
-They show preference for the typical stress patterns of their native language over others by 6--9 months (e.g., iambic vs.
-trochaic), and can use prosodic information to segment the speech stream into smaller chunks from 8 months onward \citep{johnson2001, jusczyk1993, morgan1995}.
-In comparison, children show only a limited lexical inventory at six months, and begin to recognize function words just before their first birthdays, with syntactic categorization beginning around 14 months \citep{bergelson2013, shi2010}.
-Two-month-olds notice changes in word order, but this ability appears to rely on prosodic cueing \citep{mandel1996}.
-Generally speaking then, our current knowledge about children's linguistic development points to a possible early advantage for prosody in children's turn-taking predictions.
+\citeyearpar{keitel2013, keitel2015} studies constitutes a substantial advance over previous work, it has several limitations.
+First, the estimates of ``random'' baseline shift frequency were computed exclusively from switches that occurred during the speaker's turn whereas estimates of anticipatory shifting frequency were computed exclusively from switches that occurred at points of speaker switch.
+Assuming that gaze switches happen most often during at points of speaker switch \citep{hirvenkari2013}, this analysis maximizes the probability of finding a difference between ``random'' and anticipatory switch frequencies.
+A truly random baseline would instead compare switches at turn transitions to switches at randomly-selected time points throughout the stimuli.
+Second, the conversation stimuli were somewhat unusual: the inter-turn gaps were longer than is typical for conversation ($~$900 msec), the conversations were scripted and adult-directed, and the speakers were recorded and played back on the same audio channel, making the two voices harder for viewers to distinguish.
+Despite these methodological issues, the studies demonstrate intriguing age-based differences in children's ability to predict upcoming turn structure.
+The results suggest that both semantic and intonational development \textit{do} play a role in children's looking patterns.
+Our current work takes this paradigm as our starting point.\footnote{See also \citet{casillas2012, casillas2013}.}
 
 \subsection{The Current Study}
-We report here on the role of linguistic processing in children's predictions about upcoming turn structure.
-We focus in particular on how children use prosodic and lexicosyntactic information to make their predictions.
-Prior work has focused mainly on lexicosyntax and intonation, and not on prosody proper (\citealp{de-ruiter2006}; \citealp{keitel2013}, but see \citealp{torreiraUndRev}), even though infants seem to acquire the basic rhythmic properties of the prosodic signal first \citep{mehler1988, moon1993, nazzi2003}.
-
+We report on the role of linguistic processing in children's predictions about upcoming turn structure.
 In two eye-tracking experiments, we measured children's anticipatory gaze to upcoming responders while controlling for the amount of lexicosyntactic and prosodic information available.
-Experiment 1 uses a paradigm in which English-speaking participants view video clips of naturalistic conversations from several different languages in order to create conditions under which lexicosyntactic information was differentially available, but prosodic and non-linguistic cues were intact.
+In Experiment 1 we showed English-speaking 3--5-year-olds and adults conversation video clips in several different languages to control for the presence of lexicosyntactic information while leaving prosodic and non-linguistic cues intact.
 This paradigm showed minimal differences between the predictive looking behavior of preschoolers and adults.
-In Experiment 2, we created artificial (puppet) visual stimuli to better separate these linguistic cues from one another.
+In Experiment 2, we used artificial (puppet) visual stimuli and acoustic manipulation in English-only videos to control the presence of prosodic and lexical cues in the stimuli while eliminating predictive visual cues to turn taking (e.g., gaze and gesture).
 In this more controlled paradigm, we found that children's predictive looking behavior improved from ages one to six, but that even one-year-olds made more anticipatory looks than would be expected by chance.
 
-In both experiments children consistently looked faster to responders after hearing questions, compared to non-questions.
+%In both experiments children consistently looked faster to responders after hearing questions, compared to non-questions.
 Both prosodic and lexicosyntactic information played a role in children's predictions about turn structure, but the two information sources were used differently at different ages.
-Our findings overall support an account in which predictive processes for turn taking in conversation are present early, but their integration with linguistic information takes substantial practice.
+Our findings generally support an account in which predictive processes for turn taking in conversation are present early, but their integration with linguistic information is gradual.
 
@@
 \label{sec:gendisc}
-Children begin to develop conversational turn-taking skills long before their first words \citep{bateson1975, hilbrinkInPrep, jaffe2001, snow1977}.
+Children begin to develop conversational turn-taking skills long before their first words \citep{bateson1975, hilbrinkUR, jaffe2001, snow1977}.
 As they acquire language, they also acquire the information needed to make accurate predictions about upcoming turn structure.
@@
 Our findings are consistent with observational work on children's turn taking.
-Young children are slow turn-takers in spontaneous conversation \citep{casillasUndRev} even though they begin taking turns in infancy \citep{hilbrinkInPrep, jaffe2001}.
+Young children are slow turn-takers in spontaneous conversation \citep{casillas2015} even though they begin taking turns in infancy \citep{hilbrinkUR, jaffe2001}.
 We found that children begin to spontaneously predict upcoming turn structure at age one.
@@
 Our results therefore support the idea of an early-developed competence for the timing of turn taking.
-But because children do not themselves use adult-like conversational timing until much later \citep[age six; ][]{ervin-tripp1979}, our results also indirectly point toward response planning as a key factor contributing to children's slow responses in real-time interaction \citep{casillasUndRev}.
+But because children do not themselves use adult-like conversational timing until much later \citep[age six; ][]{ervin-tripp1979}, our results also indirectly point toward response planning as a key factor contributing to children's slow responses in real-time interaction \citep{casillas2015}.
 
@@
 \bibliographystyle{elsarticle-harv}
-\bibliography{anticip}
+\bibliography{anticip-Edits}
 

### 2015-05-01 analysis/E1/0-find.transitions.R
optimization

@@ -1,193 +1,0 @@
-# Returns a data frame of transition info (whether or not switches were made
-# and, if so, when they started) for all subjects and transitions
-grab.transition.info <- function(obs, fixation.window) {
-    # Preset the total number of rows needed
-    subs <- unique(obs$Subject)
-    gaps <- sort(as.integer(unique(obs$GapNum)))
-    N <- length(subs) * length(gaps)
-    # Make the empty data frame
-    proportion.data <- data.frame(Subject=rep(NA, N), Gap=rep(NA, N),
-        Switch=rep(NA, N), Start.Time=rep(NA, N))
-    # Iterate through data frame rows, filling in transition information for
-    # each subject and gap
-    df.row = 1
-    # Grab a subject's data
-    for (i in 1:length(subs)) {
-        sub.data <- subset(obs, Subject == subs[i])
-        # Grab data for a particular gap
-        for (j in 1:length(gaps)) {
-            # Find the transition information (is there a switch? where?)
-            gap.transition <- find.transition(subset(sub.data,
-                GapNum == gaps[j]), fixation.window)
-            # Add the result to the data frame
-            proportion.data[df.row,] <- c(as.character(subs[i]),
-                gaps[j], gap.transition[1], gap.transition[2])
-            # Move onto the next row in the ouput data frame
-            df.row = df.row + 1
-        }
-    }
-    return(proportion.data)
-}
-
-
-# Parent function, linking the results of fixation.check and
-# find.transition.start together
-find.transition <- function(obs, fixation.window) {
-    if(nrow(obs) < 1) {
-        return(list(NA,NA))
-    } 
-    # Dummy minstart
-    minstart <- "NA"
-    # Search for a 100 msec fixation on the origin speaker during
-    # the ANCHOR period
-    anchor <- fixation.check(subset(obs, Anchor.Window == "ANCHOR"), "ANCHOR",
-        fixation.window, minstart)
-    minstart <- anchor[2]
-    # Search for a 100 msec fixation on the response speaker during
-    # the TRANSITION period
-    transition <- fixation.check(subset(obs, Transition.Window == "TRANSITION"),
-        "TRANSITION", fixation.window, minstart)
-    # Return the results of the search
-    if(anchor[1] == "Y" & transition[1] == "Y") {
-    	# If you found fixations on the ANCHOR and TRANSITION, find the starting
-        # point of the transition
-        return(list(1, find.transition.start(obs, transition[2])))
-    }
[…]

### 2015-05-01 analysis/E1/0-helper.R

@@ -41,49 +41,356 @@ read.data <- function(path,file.name) {
 # Gap:            |----|
 #
 get.windows.categorical <- function(obs, gap.info, utt1.window,
-    utt2.window, fixation.window) {
-    # Get rid of gaps over 550 ms (marked as 0 in the spreadsheet)
-    gaps <- subset(gap.info, Gap != "na" & Gap != "0")
-    df.out = data.frame()
-    for (i in 1:nrow(gaps)) {
-        gap <- gaps[i,]
-        rows <- subset(obs,
-            Condition == toString(gap$Condition) & 
-            (TimeSec >= gap$Onset - utt1.window - fixation.window) & 
-            (TimeSec <= gap$Offset + utt2.window + fixation.window))           
-        if (nrow(rows) == 0) {
-    	    next
-        }
-        # Set the anchor window measurements to "ANCHOR"
-        anchor.window.idx <- which(
-            (rows$TimeSec >= gap$Onset - utt1.window - fixation.window) &
-            rows$TimeSec <= gap$Onset)
-        rows$Anchor.Window <- rep(NA, nrow(rows))
-        rows$Anchor.Window[anchor.window.idx] <- "ANCHOR"
-        # Set the transition window measurements to "TRANSITION"
-        transition.window.idx <- which(
-            (rows$TimeSec >= gap$Onset - utt1.window) &
-            (rows$TimeSec <= gap$Offset + utt2.window + fixation.window))
-        rows$Transition.Window <- rep(NA, nrow(rows))
-        rows$Transition.Window[transition.window.idx] <- "TRANSITION"
-        # Set the gap window measurements to "GAP"
-        gap.window.idx <- which(rows$TimeSec >= gap$Onset &
-            rows$TimeSec <= gap$Offset)
-        rows$Gap.Window <- rep(NA, nrow(rows))
-        rows$Gap.Window[gap.window.idx] <- "GAP"
-        # Set the origin to the previous speaker at that gap
-        rows$Origin <- rep(as.character(gap$SpeakerPrev), nrow(rows))
-        # Set the QS status to the transition type (question or non-question)
-        rows$QS <- rep(as.character(gap$Type), nrow(rows))
-        # Set the gap number
-        rows$GapNum <- rep(toString(gap$Gap), nrow(rows))
-        # Create a "cumulative" time to plot the analysis window from t=0    
-        minTime <- min(rows$TimeSec)
-        rows$Cumulative.Time <- round((rows$TimeSec - minTime), 3) 
[…]

### 2015-05-01 analysis/E1/0-useful.R

@@ -7,6 +7,8 @@ library(plotrix)
 library(reshape)
 library(plyr)
 library(car)
+library(data.table)
+library(bit64)
 
 # Add some style elements for ggplot2
 plot.style <- theme(panel.grid.major = element_blank(),

### 2015-05-01 analysis/E1/1-read.dataset.R

@@ -2,9 +2,9 @@ rm(list = ls())
 source("0-useful.R")
 source("0-helper.R")
 
-raw.data.path <- "tracker_data/"
-info.path <- "info/"
-processed.data.path <- "processed_data/"
+raw.data.path <- "../../data/E1/tracker_data/"
+info.path <- "../../data/E1/info/"
+processed.data.path <- "../../data/E1/processed_data/"
 
 files <- dir(raw.data.path, pattern="*.txt")
 all.data <- data.frame()

### 2015-05-01 analysis/E1/2-prep.analysis.R

@@ -3,9 +3,9 @@ source("0-useful.R")
 source("0-helper.R")
 source("0-find.transitions.R")
 
-raw.data.path <- "tracker_data/"
-info.path <- "info/"
-processed.data.path <- "processed_data/"
+raw.data.path <- "../../data/E1/tracker_data/"
+info.path <- "../../data/E1/info/"
+processed.data.path <- "../../data/E1/processed_data/"
 
 # Read in supplemental data and prep
 vid.info <- read.csv(paste(info.path,"VideoSegmentInfo.csv",sep=""))
@@ -28,7 +28,7 @@ gap.info = within(gap.info, {
 # First we'll merge the gaze data with the information about the stimuli
 
 # Read in main datafile
-all.data <- read.csv(paste(processed.data.path,"imported.data.csv",sep=""))
+all.data <- fread(paste(processed.data.path,"imported.data.csv",sep=""))
 
 # Merge in condition sites, using onsets
 all.data$Onset <- rep(NA, nrow(all.data))

### 2015-05-01 analysis/E1/2-random.runs.R

@@ -1,17 +1,24 @@
 rm(list = ls())
 source("0-useful.R")
 source("0-helper.R")
-source("0-find.transitions.R")
+# source("0-find.transitions.R")
 
-raw.data.path <- "tracker_data/"
-info.path <- "info/"
-processed.data.path <- "processed_data/"
+#### BENCHMARKING LIBRARIES
+# library(devtools)
+# devtools::install_github("hadley/lineprof")
+library(lineprof)
+library(microbenchmark)
+
+# back to other things
+raw.data.path <- "../../data/E1/tracker_data/"
+info.path <- "../../data/E1/info/"
+processed.data.path <- "../../data/E1/processed_data/"
 
 # Read in supplemental data and prep
-vid.info <- read.csv(paste(info.path,"VideoSegmentInfo.csv",sep=""))
-subject.info <- read.csv(paste(info.path,"SubjectInfo.csv",sep=""))
+vid.info <- fread(paste(info.path,"VideoSegmentInfo.csv",sep=""))
+subject.info <- fread(paste(info.path,"SubjectInfo.csv",sep=""))
 
-all.data <- read.csv(paste(processed.data.path,"r.all.data.csv",sep=""))
+all.data <- fread(paste(processed.data.path,"r.all.data.csv",sep=""))
 
 # Separate adult and child data for finding anticipatory switches
 # because adults have an assumed faster gaze planning time of 200 msec
@@ -27,69 +34,91 @@ utt2.overlap.A = 0.2
 utt1.overlap.C = 0.333
 utt2.overlap.C = 0.333
 
-# Create 100 versions of the analyses with the random gap info files we made
-# (using the random.runs.R script), and then run this loop...
-for  (i in 1:100) {
-    # Print an update
-    print(paste("Working on random version ",i," now...", sep=""))
-    # Start with a clean version of the data
-    r.all.data.A <- all.data.A
-    r.all.data.C <- all.data.C
+###
+source("2-replicate_source.R")
+l <- lineprof(make.replicate(all.data.A, all.data.C, i, info.path, processed.data.path))
+microbenchmark(make.replicate(all.data.A, all.data.C, i, info.path, processed.data.path),
+               times = 1)
 
-    # Read in one version of the random gaps
-    r.gap.info <- read.csv(paste(info.path,"GapInfoRandom",i,".csv", sep=""))
-    r.gap.info = within(r.gap.info, {
-        Onset = round(Onset, 3)
-        Offset = round(Offset, 3)
-    })
[…]

### 2015-05-01 analysis/E1/3-plots.stats.analysis.R

@@ -6,9 +6,9 @@ library(scales)
 library(Hmisc)
 source("0-useful.R")
 
-raw.data.path <- "tracker_data/"
-info.path <- "info/"
-processed.data.path <- "processed_data/"
+raw.data.path <- "../../data/E1/tracker_data/"
+info.path <- "../../data/E1/info/"
+processed.data.path <- "../../data/E1/processed_data/"
 plot.path <- "plots/"
 
 # Color and line palettes
@@ -32,7 +32,7 @@ all.data <- 0
 
 # Read in gaze switch proportions
 # Actual data
-switch.final <- read.csv(paste(processed.data.path, "switch.final.csv", sep=""))
+switch.final <- fread(paste(processed.data.path, "switch.final.csv", sep=""))
 sw.f.drop <- c("X", "X.1", "notes", "Ttype")
 switch.final <- switch.final[,!(names(switch.final) %in% sw.f.drop)]
 # Random baseline

### 2015-05-01 analysis/E1/2-replicate_source.R
optim 2

@@ -1,0 +1,51 @@
+# Create version of the analysis with the random gap info files we made
+# (using the random.runs.R script)
+
+make.replicate <- function (all.data.A, all.data.C, i,
+                            info.path, processed.data.path) {
+
+  source("0-useful.R")
+  source("0-helper.R")
+  
+  # Read in one version of the random gaps
+  # note - casting to numeric here because of the NA being lower-case in the files 
+  r.gap.info <- read.csv(paste(info.path,"GapInfoRandom",i,".csv", sep="")) %>%
+    mutate(Gap = as.numeric(as.character(r.gap.info$Gap)), 
+           Onset = round(Onset, 3), 
+           Offset = round(Offset, 3))
+  
+  # Extract time windows for switch analysis and prepare new looking columns
+  r.target.windows.proport.A <- get.windows.categorical2(all.data.A,  # For adults
+                                                         r.gap.info, 
+                                                         utt1.overlap.A, 
+                                                         utt2.overlap.A, 
+                                                         fixation.window) %>%
+    mutate(Looks.Origin = Origin == Look.Dir, 
+           Looks.Destination = ifelse(((Origin == "L" & Look.Dir == "R") |
+                                          (Origin == "R" & Look.Dir == "L")), 1, 0))
+  
+  r.target.windows.proport.C <- get.windows.categorical2(all.data.C, # For children
+                                                         r.gap.info, 
+                                                         utt1.overlap.C, 
+                                                         utt2.overlap.C, 
+                                                         fixation.window) %>%
+    mutate(Looks.Origin = Origin == Look.Dir,
+           Looks.Destination = ifelse(((Origin == "L" & Look.Dir == "R") |
+                                         (Origin == "R" & Look.Dir == "L")), 1, 0))
+  
+  # Identify anticipatory switches in the data and combine it again
+  r.trans.info <- bind_rows(grab.transition.info2(r.target.windows.proport.A,
+                                          fixation.window), 
+                            grab.transition.info2(r.target.windows.proport.C,
+                                          fixation.window))
+
+  # Merge gap and subject info back in for analysis
+  switch.final <- r.trans.info %>% 
+    left_join(filter(r.gap.info, Condition == "A")) %>%
+    left_join(subject.info, by = "Subject") %>%
+    arrange(Subject, Gap) %>%
+    mutate(SampleType = "RANDOM")
+
+  write.csv(switch.final, paste(processed.data.path,
+                                "switch.final.optim.",i,".csv", sep=""))
+}

### 2015-05-03 analysis/E1/0-helper.R
more optimization, fixed bug in finding starts of turns in random code

@@ -126,8 +126,7 @@ get.gap2 <- function(condition, onset, offset, speakerprev, type, gap,
            QS = type,
            Gap = gap,
            Cumulative.Time = round(TimeSec - min(TimeSec),3))
-}  
-
+}   
 
 ## get gap original function
 get.gap <- function(condition, onset, offset, speakerprev, type, gap,
@@ -246,14 +245,17 @@ grab.transition.info2 <- function(obs, fixation.window) {
            transition.filter = as.numeric(stats::filter(idt, filter = rep(1, 12), 
                                                         side = 1)))  %>%
     summarise(Anchor = any(anchor.filter > .09, na.rm = TRUE),
-              Transition = any(transition.filter > .09, na.rm = TRUE),
-              Switch = Anchor & Transition, 
-              # find start time, which is last look to origin before transition.
+              Transition = any(transition.filter > .09, na.rm = TRUE),              
+              Switch = Anchor & Transition,               
+              # first look to destination is corrected by 100ms for filter
+              Transition.Time = Cumulative.Time[transition.filter > .09 & 
+                                                  !is.na(transition.filter)][1] - .1,              
+              # last look to origin before first look to transition
               Start = ifelse(Switch, 
                              max(Cumulative.Time[Looks.Origin & 
-                                                   (transition.filter < .09 | 
-                                                      is.na(transition.filter))], 
-                                 na.rm=TRUE), 0)) %>%
+                                                   Cumulative.Time < Transition.Time], 
+                                 na.rm=TRUE),
+                             0)) %>%
     arrange(Gap) # arrange nicely, not necessary
 }
 

### 2015-05-03 analysis/E1/2-random.runs.R

@@ -106,10 +106,10 @@ library(doParallel)
 start <- Sys.time()
 
 #setup parallel backend 
-cl<-makeCluster(4)
+cl<-makeCluster(4, method="FORK")
 registerDoParallel(cl, cores = 4)
 
-foreach (i = 1:4) %dopar% {
+foreach (i = 1:4, .packages = "dplyr") %dopar% {
   make.replicate(all.data.A, all.data.C, i, info.path, processed.data.path) 
 }
 

### 2015-11-20 analysis/E1/0-helper.R
Updated analyses. Too many changes to list.

@@ -5,9 +5,10 @@ library(bootstrap)
 library(lme4)
 library(stringr)
 library(plotrix)
-library(reshape)
+library(reshape2)
 library(plyr)
 library(car)
+
 
 ################################################################################
 # Misc helper functions
@@ -156,4 +157,95 @@ smoothLD <- function(dt) {
 	}
 	# Return the smoothed data
 	return(dt)
+}
+
+
+################################################################################
+# Random run plot functions
+################################################################################
+
+combine.runs <- function(DT, filelist, name) {
+	curr.row <- 1
+	for (file in filelist) {
+		# Read in a file
+		data <- fread(paste(processed.data.path, file, sep=""))
+		# Add the data to the big data table
+		DT[curr.row, names(DT) := data]
+		curr.row <- curr.row + 1
+	}
+	setkeyv(DT, "Run")
+	# Write it out
+	write.csv(DT, paste(processed.data.path,
+		name, ".random.runs.csv",sep=""), row.names=FALSE)
+	# Return it
+	return(DT)
+}
+
+get95ths <- function(meltdf, tails, origdf) {
+	variables <- unique(meltdf$variable)
+	if (tails == 1) {
+		idx95 <- round((nrow(origdf) - 1)*.95)
+		ninetyfifths <- data.frame(
+			variable=variables,
+			tvalue=rep(0,length(variables)))
+		for (v in variables) {
+			ninetyfifths$tvalue[which(ninetyfifths$variable == v)] <-
+				sort(subset(meltdf, variable == v)$tvalue)[idx95]
+		}
+	}
+	else if (tails == 2) {
+		idx95.l <- round((nrow(origdf) - 1)*.025)
+		idx95.u <- round((nrow(origdf) - 1)*.975)
+		ninetyfifths <- data.frame(
+			variable=variables,
+			tvalue.l=rep(0,length(variables)),
+			tvalue.u=rep(0,length(variables)))
+		for (v in variables) {
+			ninetyfifths$tvalue.l[which(ninetyfifths$variable == v)] <-
+				sort(subset(meltdf, variable == v)$tvalue)[idx95.l]
+			ninetyfifths$tvalue.u[which(ninetyfifths$variable == v)] <-
+				sort(subset(meltdf, variable == v)$tvalue)[idx95.u]
+		}
+
+	}
+	return(ninetyfifths)
+}
+
+getPercentile <- function(realvals, randvals, tails) {
[…]

### 2015-11-20 analysis/E1/0-make.random.windows.R

@@ -1,178 +1,0 @@
-# Takes in a vector of gap durations associated with each stimulus and returns a 
-# data frame of with the same set of gap durations, only shuffled at random 
-# within the time boundaries of the associated stimulus.
-find.rand.windows <- function(min, max, vecdur) {
-    N <- length(vecdur)
-    # Maximum start time for a gap is .433 msec before the end of the stimulus
-    give.win <- 0.433
-    # Create a new, empty data frame
-    new.windows <- data.frame(Onset=rep(NA, N), Offset=rep(NA, N),
-        Duration=rep(NA, N))
-    # Iterate through new.windows, filling in information for each gap duration
-    df.row = 1
-    for (i in 1:length(vecdur)) {
-        success=0
-        while(success < 1) {
-            dur = vecdur[i]
-            # Generate a random start value within the video stimulus
-            try.start <- runif(1, min + give.win, max - give.win - dur)
-            # Fill in the first value
-            if (is.numeric(new.windows[1,1]) == FALSE) {
-            	# Add the randomized gap to the data frame
-                new.windows[df.row,] <- c(try.start, try.start + dur, dur)
-                # Move onto the next row
-                df.row = df.row + 1
-                success = 1
-            }
-            # For the second row and beyond...
-            else {		
-                try.win <- c(try.start, try.start + dur)
-                # Check to see if the new proposed random gap overlaps with
-                # one we have already
-                used.win <- subset(new.windows[,1:2], Onset != "NA")
-                temp.win <- rbind(try.win, used.win)
-                # Orders the accepted and proposed random gaps by onset
-                # (crucial for the check.overlap function)
-                temp.win <- temp.win[order(temp.win$Onset),]
-                # If it doesn't, add it to the data frame
-                if (check.overlap(temp.win) == 0) {
-                    new.windows[df.row,] <- c(try.start, try.start + dur, dur)
-                    new.windows <- new.windows[order(new.windows$Onset),]
-                    df.row = df.row + 1
-                    success = 1
[…]

### 2015-11-20 analysis/E1/0-useful.R

@@ -1,63 +1,0 @@
-library(grid)
-library(ggplot2)
-library(bootstrap)
-library(lme4)
-library(stringr)
-library(plotrix)
-library(reshape)
-library(plyr)
-library(car)
-library(data.table)
-library(bit64)
-
-# Add some style elements for ggplot2
-plot.style <- theme(panel.grid.major = element_blank(),
-    panel.grid.minor = element_blank(),
-    axis.line = element_line(colour="black",size=.5),
-    axis.ticks = element_line(size=.5),
-    axis.title.x = element_text(vjust=-.5),
-    axis.title.y = element_text(angle=90,vjust=0.25),
-    panel.margin = unit(1.5,"lines"))
-
-theme_set(theme_bw())
-
-# Standard error of the mean
-sem <- function (x) {
-    sd(x) / sqrt(length(x))
-}
-
-na.mean <- function(x) {mean(x,na.rm=T)}
-
-to.n <- function(x) {
-    as.numeric(as.character(x))
-}
-
-# Number of unique subs
-n.unique <- function (x) {
-    length(unique(x))
-}
-
-# For bootstrapping 95% confidence intervals
-theta <- function(x,xdata) {mean(xdata[x])}
-ci.low <- function(x) {
-    mean(x) - quantile(bootstrap(1:length(x),1000,theta,x)$thetastar,.025)}
-ci.high <- function(x) {
-    quantile(bootstrap(1:length(x),1000,theta,x)$thetastar,.975) - mean(x)}
-ci95 <- function(x) {sem(x)*1.96}
-
-# For basic plots, add linear models with correlations
-lm.txt <- function (p1,p2,x=7.5,yoff=.05,lt=2,c="black",data=data) {
-    l <- lm(p2 ~ p1)
-    regLine(l,lty=lt,col=c)
-    cl <- coef(l)
-    text(x,cl[1] + cl[2] * x + yoff,
-        paste("r = ",sprintf("%2.2f",sqrt(summary(l)$r.squared)),
-        getstars(anova(l)$"Pr(>F)"[1]),sep=""), xpd="n")
-}
-
-getstars <- function(x) {
-    if (x > .1) {return("")}
-    if (x < .001) {return("***")}
-    if (x < .01) {return("**")}
-    if (x < .05) {return("*")}
-}

### 2015-11-20 analysis/E1/1-read.dataset.R

@@ -1,26 +1,0 @@
-rm(list = ls())
-source("0-useful.R")
-source("0-helper.R")
-
-raw.data.path <- "../../data/E1/tracker_data/"
-info.path <- "../../data/E1/info/"
-processed.data.path <- "../../data/E1/processed_data/"
-
-files <- dir(raw.data.path, pattern="*.txt")
-all.data <- data.frame()
-
-# Loop through the files, grabbing the data we want
-for (file.name in files) {
-    # Print an update
-    print(file.name)
-    data <- read.data(raw.data.path, file.name)
-    # Take only the columns we want for further analysis
-    data <- subset(data, select = c("Time", "x.pos", "y.pos", "L.looks",
-        "R.looks", "Events", "Condition", "Subject", "AgeGroup", "TimeSec",
-        "MS.Increment", "Stimulus", "Saccades", "Fixations"))
-    # Bind it together
-    all.data <- rbind(all.data, data)
-}
-
-# Save the imported data
-write.csv(all.data, paste(processed.data.path,"imported.data.csv",sep=""))

### 2015-11-20 analysis/E1/2-find.switches.R

@@ -39,7 +39,7 @@ all.data.C5 <- filter(all.data, Age == "5")
 # Anticipatory switch identification variables:
 fix.window <- 0.1
 # Saccadic RTs set for...
-saccadeRTs <- c(0.278, 0.252, 0.226, 0.2)
+saccadeRTs <- c(0.2, 0.2, 0.2, 0.2)
 # 3-year-olds
 utt.overlap.C3 = saccadeRTs[1]
 # 4-year-olds

### 2015-11-20 analysis/E1/2-prep.analysis.R

@@ -1,174 +1,0 @@
-rm(list = ls())
-source("0-useful.R")
-source("0-helper.R")
-source("0-find.transitions.R")
-
-raw.data.path <- "../../data/E1/tracker_data/"
-info.path <- "../../data/E1/info/"
-processed.data.path <- "../../data/E1/processed_data/"
-
-# Read in supplemental data and prep
-vid.info <- read.csv(paste(info.path,"VideoSegmentInfo.csv",sep=""))
-turn.info <- read.csv(paste(info.path,"TurnInfo.csv",sep=""))
-subject.info <- read.csv(paste(info.path,"SubjectInfo.csv",sep=""))
-gap.info <- read.csv(paste(info.path,"GapInfo.csv",sep=""))
-
-# Round the time values for turns and gaps
-turn.info = within(turn.info, {
-  Onset = round(Onset, 3)
-  Offset = round(Offset, 3)
-})
-gap.info = within(gap.info, {
-  Onset = round(Onset, 3)
-  Offset = round(Offset, 3)
-})
-
-
-################################################################################
-# First we'll merge the gaze data with the information about the stimuli
-
-# Read in main datafile
-all.data <- fread(paste(processed.data.path,"imported.data.csv",sep=""))
-
-# Merge in condition sites, using onsets
-all.data$Onset <- rep(NA, nrow(all.data))
-for (cond in unique(all.data$Condition)) {
-    # For each TimeSecSeg measurement, find the onset of the segment
-    # (from vid.info) within which it lies
-    onsets <- vid.info$Onset[vid.info$Cond == cond]
-    # REMINDER: example use of findInterval:
-    # findInterval(c(2,5,7,8), c(4,5,6,8,9,10)) => [1] 0 2 3 4
-    # In this case it returns a list as long as all.data$TimeSecSeg, with
-    # the index of the video onset that the each time measurement falls within
-    # (e.g., t = 5, vid[2] and t=7, vid[3] in the above example)
-    onset.indices <- findInterval(all.data$TimeSec[all.data$Condition == cond],
-        onsets)
-    # Maps these values onto the subset of times it was derived from
-    all.data$Onset[all.data$Condition == cond] <- onsets[onset.indices]
-}
-# Add in vid.info data and re-order the main data frame
-all.data <- merge(all.data, vid.info, by = c("Condition", "Onset"))
-all.data <- all.data[order(all.data$Subject, all.data$TimeSec),]
-
-# Merge in turn sites, using onsets
[…]

### 2015-11-20 analysis/E1/2-random.runs.R

@@ -1,135 +1,0 @@
-rm(list = ls())
-source("0-useful.R")
-source("0-helper.R")
-# source("0-find.transitions.R")
-
-#### BENCHMARKING LIBRARIES
-# library(devtools)
-# devtools::install_github("hadley/lineprof")
-library(lineprof)
-library(microbenchmark)
-
-# back to other things
-raw.data.path <- "../../data/E1/tracker_data/"
-info.path <- "../../data/E1/info/"
-processed.data.path <- "../../data/E1/processed_data/"
-
-# Read in supplemental data and prep
-vid.info <- fread(paste(info.path,"VideoSegmentInfo.csv",sep=""))
-subject.info <- fread(paste(info.path,"SubjectInfo.csv",sep=""))
-
-all.data <- fread(paste(processed.data.path,"r.all.data.csv",sep=""))
-
-# Separate adult and child data for finding anticipatory switches
-# because adults have an assumed faster gaze planning time of 200 msec
-# compared to our assumed 333 msec for children
-all.data.A <- subset(all.data, AgeGroup == "ADU")
-all.data.C <- subset(all.data, AgeGroup == "CHI")
-
-# Anticipatory switch identification variables
-fixation.window = 0.1
-# could change utt settings to "reaction" rather than "anticipation" values
-utt1.overlap.A = 0.2
-utt2.overlap.A = 0.2
-utt1.overlap.C = 0.333
-utt2.overlap.C = 0.333
-
-###
-source("2-replicate_source.R")
-l <- lineprof(make.replicate(all.data.A, all.data.C, i, info.path, processed.data.path))
-microbenchmark(make.replicate(all.data.A, all.data.C, i, info.path, processed.data.path),
-               times = 1)
-
-# time     alloc   release     dups                     ref                                    src
-# 1   0.007     0.716     0.000      317 2-replicate_source.R#14 make.replicate/read.csv               
-# 2   8.069  4157.738  4008.718  5843236 2-replicate_source.R#22 make.replicate/get.windows.categorical
-# 3   0.001     0.127     0.000      266 2-replicate_source.R#24 make.replicate/==                     
-# 4   0.017     5.131     0.000     2334 2-replicate_source.R#24 make.replicate/$<-                    
-# 5   0.022     0.922     0.000     1217 2-replicate_source.R#27 make.replicate/ifelse                 
[…]

### 2015-11-20 analysis/E1/2-replicate_source.R

@@ -1,51 +1,0 @@
-# Create version of the analysis with the random gap info files we made
-# (using the random.runs.R script)
-
-make.replicate <- function (all.data.A, all.data.C, i,
-                            info.path, processed.data.path) {
-
-  source("0-useful.R")
-  source("0-helper.R")
-  
-  # Read in one version of the random gaps
-  # note - casting to numeric here because of the NA being lower-case in the files 
-  r.gap.info <- read.csv(paste(info.path,"GapInfoRandom",i,".csv", sep="")) %>%
-    mutate(Gap = as.numeric(as.character(r.gap.info$Gap)), 
-           Onset = round(Onset, 3), 
-           Offset = round(Offset, 3))
-  
-  # Extract time windows for switch analysis and prepare new looking columns
-  r.target.windows.proport.A <- get.windows.categorical2(all.data.A,  # For adults
-                                                         r.gap.info, 
-                                                         utt1.overlap.A, 
-                                                         utt2.overlap.A, 
-                                                         fixation.window) %>%
-    mutate(Looks.Origin = Origin == Look.Dir, 
-           Looks.Destination = ifelse(((Origin == "L" & Look.Dir == "R") |
-                                          (Origin == "R" & Look.Dir == "L")), 1, 0))
-  
-  r.target.windows.proport.C <- get.windows.categorical2(all.data.C, # For children
-                                                         r.gap.info, 
-                                                         utt1.overlap.C, 
-                                                         utt2.overlap.C, 
-                                                         fixation.window) %>%
-    mutate(Looks.Origin = Origin == Look.Dir,
-           Looks.Destination = ifelse(((Origin == "L" & Look.Dir == "R") |
-                                         (Origin == "R" & Look.Dir == "L")), 1, 0))
-  
-  # Identify anticipatory switches in the data and combine it again
-  r.trans.info <- bind_rows(grab.transition.info2(r.target.windows.proport.A,
-                                          fixation.window), 
-                            grab.transition.info2(r.target.windows.proport.C,
-                                          fixation.window))
-
-  # Merge gap and subject info back in for analysis
-  switch.final <- r.trans.info %>% 
-    left_join(filter(r.gap.info, Condition == "A")) %>%
-    left_join(subject.info, by = "Subject") %>%
-    arrange(Subject, Gap) %>%
-    mutate(SampleType = "RANDOM")
-
-  write.csv(switch.final, paste(processed.data.path,
-                                "switch.final.optim.",i,".csv", sep=""))
-}

### 2015-11-20 analysis/E1/3-plots.stats.analysis.R

@@ -34,17 +34,17 @@ files <- dir(processed.data.path, pattern="switch.final.rand.[0-9]+.csv")
 N <- length(files) * 4000 # > max length of a switch.final.rand csv file
 switch.final.r <- data.table(
 	Subject = rep("", N),
-	Gap = rep(0, N),
-	Switch = rep(0, N),
+	Gap = rep(as.integer(0), N),
+	Switch = rep(FALSE, N),
 	Start = rep(0, N),
-	Run = rep(0,N))
+	Run = rep(as.integer(0),N))
 # Fill the data table with the random runs
 curr.row <- 1
 for (file in files) {
 	# Read in a file
 	data <- fread(paste(processed.data.path, file, sep=""))
 	# Record the random run number
-	data[, Run := unlist(strsplit(file, "[.]"))[4]]
+	data[, Run := as.integer(unlist(strsplit(file, "[.]"))[4])]
 	# Add the data to the big data table
 	curr.rows <- curr.row:(curr.row + nrow(data) - 1)
 	switch.final.r[curr.rows, names(switch.final.r) := data]
@@ -78,6 +78,9 @@ switch.all <- switch.all %>%
         inner_join(gap.info, by = "Gap") %>%
         inner_join(sub.info, by = "Subject")
 
+# Write combined data
+write.csv(switch.all, paste(processed.data.path,
+	"SwitchAll.csv", sep=""), row.names=FALSE)
 
 ################################################################################
 # Create plots
@@ -139,6 +142,71 @@ p1 <- qplot(Age,m,facets = . ~ Language, group=Sample, ylim=c(-0.1,0.6),
 png(paste(plot.path, "samples-by-lang-groups.png", sep=""),
     width=1100,height=600,units="px", bg = "transparent")
 print(p1)
+dev.off()    
+################################################################################
+
+
+###### DIFF BETWEEN RANDOM AND TRANSITION w/ TYPE  #############################
+mean.agg <- aggregate(Switch ~ LgGroup + Sample + Type + Age, switch.all, mean)
+langs <- mean.agg$LgGroup
+ages <- mean.agg$Age
+samples <- mean.agg$Sample
+types <- mean.agg$Type
+means <- mean.agg$Switch
+errs <- aggregate(Switch ~ LgGroup + Sample + Type + Age, switch.all, sem)$Switch
+linecolors <- c(
+	"Random (both)", "Random (both)",
+	"English", "Non-English",
[…]

### 2015-11-20 analysis/E1/4-model-functions.R

@@ -1,0 +1,87 @@
+source("0-helper.R")
+library(dplyr)
+library(ggplot2)
+library(lme4)
+library(gridExtra)
+library(scales)  
+library(Hmisc)
+library(data.table)
+
+################################################################################
+# Read in data in processed_data/ path and set plotting variables
+################################################################################
+processed.data.path <- "processed_data/" #"../../data/E1/processed_data/"
+switch.all <- fread(paste(processed.data.path,"SwitchAll.csv", sep=""))
+chi.coefs <- fread(paste(processed.data.path,"chi.coefs-0.csv", sep=""))
+adu.coefs <- fread(paste(processed.data.path,"adu.coefs-0.csv", sep=""))
+
+run.models <- function(ns) {
+    # Split the data into adults and children for separate models
+    switch.rand <- filter(switch.all, Sample == "Random") %>%
+                   select(-Sample)
+                   
+    switch.rand.C <- filter(switch.rand, Age != "A") %>%
+                     mutate(Age = as.numeric(Age))
+    switch.rand.A <- filter(switch.rand, Age == "A")
+
+    withCallingHandlers({
+        for(i in ns){
+            w <- length(warnings())
+            errors <- ""
+	    	# Update
+    	    print(i)
+
+    		# Retrieve the random gap info data for this particular run
+	    	rand.data.C <- switch.rand.C[Run == i,]
+	    	chi.coefs.r <- chi.coefs[0,]
+
+    		# Models
+	    	# Children
+			chi.max.r <- glmer(Switch ~ Age * LgGroup * Type + Duration + 
+			    (Type * LgGroup|Subject) + (1|Gap), data=rand.data.C,
+			    family=binomial, control = glmerControl(optimizer="bobyqa"))
+    		chi.coefs.r <- rbind(chi.coefs.r,
+    			cbind(data.frame(t(sapply(fixef(chi.max.r),c))),
+    			data.frame(Sample = "random", Error = errors, Run = i)))
+    		write.csv(chi.coefs.r, paste(
+				processed.data.path, "chi.coefs.r-", i, ".csv", sep=""),
+				row.names=FALSE)		
+        }
+    }, warning = function(w){
+	    if (errors == "") {
+	        errors <<- w$message
+        }
+        invokeRestart("muffleWarning")
+    })
+
+    withCallingHandlers({
+        for(i in ns){
[…]

### 2015-11-20 analysis/E1/5-compare.R

@@ -1,0 +1,264 @@
+rm(list=ls())
+source("0-helper.R")
+
+################################################################################
+# Read in data in processed_data/ path and set plotting variables
+################################################################################
+# Set paths
+processed.data.path <- "processed_data/" #"../../data/E1/processed_data/"
+plot.path <- "plots/" #"plots/"
+
+# Read in all the random runs and combine them
+adu.r.files <- dir(processed.data.path, pattern="adu.coefs.r-[0-9]+.csv")
+chi.r.files <- dir(processed.data.path, pattern="chi.coefs.r-[0-9]+.csv")
+
+adu.r.model <- read.csv(paste(processed.data.path, adu.r.files[1], sep=""))
+chi.r.model <- read.csv(paste(processed.data.path, chi.r.files[1], sep=""))
+adu.model <- data.table(read.csv(paste(processed.data.path, pattern="adu.coefs-0.csv", sep="")))
+chi.model <- data.table(read.csv(paste(processed.data.path, pattern="chi.coefs-0.csv", sep="")))
+
+# Initialize an empty data table
+aduN <- length(adu.r.files)+1
+chiN <- length(chi.r.files)+1
+
+aduDT <- data.table(
+	matrix(0, aduN, length(adu.r.model),
+	dimnames=list(c(), colnames(adu.r.model))))
+aduDT$Sample <- rep("", aduN)
+aduDT$Error <- rep("", aduN)
+
+chiDT <- data.table(
+	matrix(0, chiN, length(chi.r.model),
+	dimnames=list(c(), colnames(chi.r.model))))
+chiDT$Sample <- rep("", chiN)
+chiDT$Error <- rep("", chiN)
+
+aduruns <- combine.runs(aduDT, adu.r.files, "adu")
+adu.model$Sample <- as.character(adu.model$Sample) 
+adu.model$Error <- as.character(adu.model$Error) 
+adu.model$Run <- as.numeric(adu.model$Run)
+aduruns[1, names(aduruns) := adu.model]
+aduruns.all <- aduruns
+aduruns <- subset(aduruns, Error == "" | Sample == "real")
+nrow(aduruns)/nrow(aduruns.all)
+
+chiruns <- combine.runs(chiDT, chi.r.files, "chi")
[…]

### 2015-11-20 analysis/E1/_RUNME.R

@@ -5,37 +5,10 @@ source("2-find.switches.R")
 get.switches(1, FALSE)
 
 # Find switches in randomly-shuffled gap versions
-rand.runs <- 1:5
-make.rand.gaps(rand.runs)
-get.switches(rand.runs, TRUE)
+# rand.runs <- 1:5000
+# make.rand.gaps(rand.runs)
+# get.switches(rand.runs, TRUE)
 
 source("3-plots.stats.analysis.R")
+source("4-model-functions.R")
 run.models(rand.runs)
-compare.models()
-
-#### BENCHMARKING LIBRARIES
-# library(devtools)
-# devtools::install_github("hadley/lineprof")
-# library(lineprof)
-# library(microbenchmark)
-
-# ###########
-# l <- lineprof(get.switches(1, FALSE))
-# microbenchmark(get.switches(1, FALSE), times = 1)
-
-# ########### Middy did not try to fix parallelization yet
-# library(foreach)
-# library(doParallel)
-
-# start <- Sys.time()
-
-# #setup parallel backend 
-# cl<-makeCluster(4, method="FORK")
-# registerDoParallel(cl, cores = 4)
-
-# foreach (i = 1:4, .packages = "dplyr") %dopar% {
-  # make.replicate(all.data.A, all.data.C, i, info.path, processed.data.path) 
-# }
-
-# stopCluster(cl)
-# stoptime <- Sys.time() - start

### 2015-11-20 analysis/E2/0-find.transitions.R

@@ -1,176 +1,0 @@
-################################################################################
-# Returns a data frame of transition info (whether or not switches were made
-# and, if so, when they started) for all subjects and transitions in obs
-grab.transition.info <- function(obs, fixation.window) {
-    # Preset the total number of rows needed
-    subs <- unique(obs$Subject)
-    gaps <- sort(as.integer(unique(obs$GapNum)))
-    N <- length(subs) * length(gaps)
-    # Make an empty data frame
-    proportion.data <- data.table(
-    	Subject=rep("", N),
-    	Gap=rep(0, N),
-        Switch=rep(0, N),
-        Start.Time=rep(0, N))
-    # Iterate through data frame rows, filling in
-    # transition information for each subject and gap
-    df.row <- 1
-    # Grab a subject's data
-    for (sub in subs) {
-        sub.data <- subset(obs, Subject == sub)
-        # Grab data for a particular gap
-        sub.gaps <- unique(sub.data$GapNum)
-        for (gap in sub.gaps) {
-            # Find the transition information. Returns two parts:
-            # 1 - Was there a switch in the analysis window
-            # 2 - If so, when in the window did the switch occur?
-            gap.transition <- find.transition(subset(sub.data,
-                GapNum == gap), fixation.window)
-            # Add the transition info to the data frame
-            proportion.data[df.row,] <- c(as.character(sub),
-                gap, gap.transition[1], gap.transition[2])
-            # Move onto the next row in the ouput data frame
-            df.row = df.row + 1
-        }
-    }
-    # Return the output data frame,
-    # minus any remaining empty rows
-    return(proportion.data[1:df.row])
-}
-
-################################################################################
-# Called by grab.transition.info
-# Parent function to fixation.check
-# Sets up data to check for fixations and switch start times
-# Returns the result to grab.transition.info
-find.transition <- function(obs, fixation.window) {
-	# In case there are no eye tracking measurements in this
-	# particular analysis window
-    if (nrow(obs) == 0) {
-        return(list(NA,NA))
-    }
-    
-    # Dummy minstart; will set later
-    minstart <- "NA"
-    
-    # Find the subset of the total analysis window that is just
-    # linked to the Anchor window
-    anchorobs <- subset(obs, Anchor.Window == "ANCHOR")
-    # If there is none, exit without success
[…]

### 2015-11-20 analysis/E2/0-helper.R

@@ -5,7 +5,7 @@ library(bootstrap)
 library(lme4)
 library(stringr)
 library(plotrix)
-library(reshape)
+library(reshape2)
 library(plyr)
 library(car)
 
@@ -165,4 +165,95 @@ smoothLD <- function(dt) {
 	}
 	# Return the smoothed data
 	return(dt)
+}
+
+
+################################################################################
+# Random run plot functions
+################################################################################
+
+combine.runs <- function(DT, filelist, name) {
+	curr.row <- 1
+	for (file in filelist) {
+		# Read in a file
+		data <- fread(paste(processed.data.path, file, sep=""))
+		# Add the data to the big data table
+		DT[curr.row, names(DT) := data]
+		curr.row <- curr.row + 1
+	}
+	setkeyv(DT, "Run")
+	# Write it out
+	write.csv(DT, paste(processed.data.path,
+		name, ".random.runs.csv",sep=""), row.names=FALSE)
+	# Return it
+	return(DT)
+}
+
+get95ths <- function(meltdf, tails, origdf) {
+	variables <- unique(meltdf$variable)
+	if (tails == 1) {
+		idx95 <- round((nrow(origdf) - 1)*.95)
+		ninetyfifths <- data.frame(
+			variable=variables,
+			tvalue=rep(0,length(variables)))
+		for (v in variables) {
+			ninetyfifths$tvalue[which(ninetyfifths$variable == v)] <-
+				sort(subset(meltdf, variable == v)$tvalue)[idx95]
+		}
+	}
+	else if (tails == 2) {
+		idx95.l <- round((nrow(origdf) - 1)*.025)
+		idx95.u <- round((nrow(origdf) - 1)*.975)
+		ninetyfifths <- data.frame(
+			variable=variables,
+			tvalue.l=rep(0,length(variables)),
+			tvalue.u=rep(0,length(variables)))
+		for (v in variables) {
+			ninetyfifths$tvalue.l[which(ninetyfifths$variable == v)] <-
+				sort(subset(meltdf, variable == v)$tvalue)[idx95.l]
+			ninetyfifths$tvalue.u[which(ninetyfifths$variable == v)] <-
+				sort(subset(meltdf, variable == v)$tvalue)[idx95.u]
+		}
+
+	}
+	return(ninetyfifths)
+}
+
+getPercentile <- function(realvals, randvals, tails) {
+	variables <- unique(randvals$variable)
+	if (tails == 1) {
[…]

### 2015-11-20 analysis/E2/2-find.switches.R

@@ -42,7 +42,7 @@ all.data.C6 <- filter(all.data, Age == 6)
 # Anticipatory switch identification variables:
 fix.window <- 0.1
 # Saccadic RTs set for...
-saccadeRTs <- c(0.33, 0.304, 0.278, 0.252, 0.226, 0.2, 0.2)
+saccadeRTs <- c(0.2, 0.2, 0.2, 0.2, 0.2, 0.2, 0.2)
 # 1-year-olds
 utt.overlap.C1 = saccadeRTs[1]
 # 2-year-olds

### 2015-11-20 analysis/E2/3-plots.stats.analysis.R

@@ -19,6 +19,7 @@ plot.path <- "plots/" #"plots/"
 # Read in supplemental data and prepped data
 gap.info <- fread(paste(info.path,"GapInfo.csv", sep=""))
 subject.info <- fread(paste(info.path,"SubjectInfo.csv",sep=""))
+vid.info <- fread(paste(info.path,"VideoSegmentInfo.csv",sep=""))
 
 # Color and line palettes
 stoplightPalette <- c("dodgerblue3", "forestgreen",
@@ -36,10 +37,10 @@ files <- dir(processed.data.path, pattern="switch.final.rand.[0-9]+.csv")
 N <- length(files) * 4000 # > max length of a switch.final.rand csv file
 switch.final.r <- data.table(
 	Subject = rep("", N),
-	Gap = rep(0, N),
-	Switch = rep(0, N),
+	Gap = rep(as.integer(0), N),
+	Switch = rep(FALSE, N),
 	Start = rep(0, N),
-	Run = rep(0,N))
+	Run = rep(as.integer(0),N))
 # Fill the data table with the random runs
 curr.row <- 1
 for (file in files) {
@@ -82,6 +83,10 @@ sub.info <- select(subject.info, Subject, Age) %>%
 switch.all <- switch.all %>%
         inner_join(gap.info, by = "Gap") %>%
         inner_join(sub.info, by = "Subject")
+
+# Write combined data
+write.csv(switch.all, paste(processed.data.path,
+	"SwitchAll.csv", sep=""), row.names=FALSE)
 
 
 ################################################################################
@@ -185,7 +190,7 @@ df$Condition <- factor(df$Condition, levels=c("Normal speech", "Prosody only",
 
 limits <- aes(ymax = m + se, ymin= m - se)
 
-p3 <- qplot(Age,m,facets = . ~ Condition, group=Transition, ylim=c(-0.1,0.6),
+p2 <- qplot(Age,m,facets = . ~ Condition, group=Transition, ylim=c(-0.1,0.6),
     ymin= m - se, ymax= m + se, color=Condition, linetype=Transition,
     xlab="Age (years)",ylab="Proportion gaze switches",
     position=position_dodge(width=.1), data=df) +
@@ -210,7 +215,7 @@ p3 <- qplot(Age,m,facets = . ~ Condition, group=Transition, ylim=c(-0.1,0.6),
     plot.margin=unit(c(10,10,10,10),"mm"))
[…]

### 2015-11-20 analysis/E2/4-boredom.hypoth.R

@@ -1,297 +1,0 @@
-# library(grid)
-# library(ggplot2)
-# library(bootstrap)
-# library(lme4)
-# library(stringr)
-# library(plotrix)
-# library(reshape)
-# library(plyr)
-# library(car)
-# library(gridExtra)
-# library(scales)  
-# library(Hmisc)
-# library(moments)
-rm(list = ls())
-source("0-helper.R")
-library(data.table)
-library(bit64)
-
-################################################################################
-################################################################################
-info.path <- "info/"
-processed.data.path <- "processed_data_boredom/"
-plot.path <- "plots/"
-
-colorPalette <- c("black", "gray")
-
-# Read in supplemental data and prep
-vid.info <- fread(paste(info.path, "VideoSegmentInfo.csv", sep=""))
-#vid.info$VideoDuration <- NULL
-turn.info <- fread(paste(info.path, "TurnInfo.csv", sep=""))
-subject.info <- fread(paste(info.path, "SubjectInfo.csv", sep=""))
-
-# Round the time values for turns
-turn.info = within(turn.info, {
-  Onset = round(Onset, 5)
-  Offset = round(Offset, 5)
-})
-
-# Read in main datafile
-all.data <- fread("processed_data/imported.data.csv")
-
-# Merge in the stimulus (video) information
-setkey(all.data, "Segment")
-setkey(vid.info, "Segment")
-all.data <- vid.info[all.data]
-# Subset the data to only include target stimuli
-all.data <- all.data[StimulusGroup == "target"]
-
-# Collapse male-looks and female-looks into a single look-direction column
-# F and M are each TRUE or FALSE (coerced to 1 or 0)
-all.data[, Look.Dir := ifelse(
-	F.looks > M.looks, "F", ifelse(
-	M.looks > F.looks, "M", "0"))]
-
-# Write out normal data up until now
-write.csv(all.data,
-    paste(processed.data.path, "all.data.before.analysis.bh.csv", sep=""),
-    row.names = FALSE)
-
-# Re-order the data table
-setkeyv(all.data, c("Subject", "Segment", "TimeSec"))
-
-all.data <- smoothLD(all.data)
-
-# Write out the prepped data for input into the transition analysis
-write.csv(all.data,
-    paste(processed.data.path, "prepped.data.bh.csv", sep=""),
-    row.names=FALSE)
-
[…]

### 2015-11-20 analysis/E2/4-model-functions.R

@@ -1,0 +1,154 @@
+source("0-helper.R")
+library(dplyr)
+library(ggplot2)
+library(lme4)
+library(gridExtra)
+library(scales)  
+library(Hmisc)
+library(data.table)
+
+################################################################################
+# Read in data in processed_data/ path and set plotting variables
+################################################################################
+processed.data.path <- "processed_data/" #"../../data/E1/processed_data/"
+switch.all <- fread(paste(processed.data.path,"SwitchAll.csv",sep=""))
+chi.coef <- fread(paste(processed.data.path,"chi.coefs-0.csv",sep=""))
+chi.coef.l <- fread(paste(processed.data.path,"chi.coefs.l-0.csv",sep=""))
+adu.coef <- fread(paste(processed.data.path,"adu.coefs-0.csv",sep=""))
+adu.coef.l <- fread(paste(processed.data.path,"adu.coefs.l-0.csv",sep=""))
+
+run.models <- function(ns) {
+    # Split the data into adults and children for separate models
+    switch.rand <- filter(switch.all, Sample == "Random") %>%
+                   select(-Sample) %>%
+				   mutate(Condition = as.factor(Condition),
+		   		          Type = as.factor(Type))
+
+	switch.rand$Cond2 <- ifelse((switch.rand$Condition == "normal" |
+		switch.rand$Condition == "robot"),"lexical", "non-lexical")
+
+	switch.rand.C <- filter(switch.rand, Age != 21) %>%
+         			        mutate(Age = as.numeric(Age))
+	switch.rand.A <- filter(switch.rand, Age == 21)
+
+    withCallingHandlers({
+        for(i in ns){
+            w <- length(warnings())
+            errors <- ""
+	    	# Update
+    	    print(i)
+
+    		# Retrieve the random gap info data for this particular run
+	    	rand.data.C <- switch.rand.C[Run == i,]
+	    	chi.coef.r <- chi.coef[0,]
+
+    		# Models
+	    	# Children
+			chi.max.r <- glmer(Switch ~ Age * Condition * Type + Duration + 
+ 			   (Type|Subject) + (1|Gap), data=rand.data.C,
+ 			   family=binomial, control = glmerControl(optimizer="bobyqa"))
+    		chi.coef.r <- rbind(chi.coef.r,
[…]

### 2015-11-20 analysis/E2/5-compare.R

@@ -1,0 +1,253 @@
+rm(list=ls())
+source("0-helper.R")
+
+################################################################################
+# Read in data in processed_data/ path and set plotting variables
+################################################################################
+# Set paths
+processed.data.path <- "processed_data/" #"../../data/E1/processed_data/"
+plot.path <- "plots/" #"plots/"
+
+# Read in all the random runs and combine them
+adu.r.files <- dir(processed.data.path, pattern="adu.coefs.r-[0-9]+.csv")
+chi.r.files <- dir(processed.data.path, pattern="chi.coefs.r-[0-9]+.csv")
+adu.l.r.files <- dir(processed.data.path, pattern="adu.coefs.l.r-[0-9]+.csv")
+chi.l.r.files <- dir(processed.data.path, pattern="chi.coefs.l.r-[0-9]+.csv")
+
+adu.r.model <- read.csv(paste(processed.data.path, adu.r.files[1], sep=""))
+chi.r.model <- read.csv(paste(processed.data.path, chi.r.files[1], sep=""))
+adu.model <- data.table(read.csv(paste(processed.data.path, pattern="adu.coefs-0.csv", sep="")))
+chi.model <- data.table(read.csv(paste(processed.data.path, pattern="chi.coefs-0.csv", sep="")))
+
+# Initialize an empty data table
+aduN <- length(adu.r.files)+1
+chiN <- length(chi.r.files)+1
+
+aduDT <- data.table(
+	matrix(0, aduN, length(adu.r.model),
+	dimnames=list(c(), colnames(adu.r.model))))
+aduDT$Sample <- rep("", aduN)
+aduDT$Error <- rep("", aduN)
+
+chiDT <- data.table(
+	matrix(0, chiN, length(chi.r.model),
+	dimnames=list(c(), colnames(chi.r.model))))
+chiDT$Sample <- rep("", chiN)
+chiDT$Error <- rep("", chiN)
+
+aduruns <- combine.runs(aduDT, adu.r.files, "adu")
+adu.model$Sample <- as.character(adu.model$Sample) 
+adu.model$Error <- as.character(adu.model$Error) 
+adu.model$Run <- as.numeric(adu.model$Run)
+aduruns[1, names(aduruns) := adu.model]
+aduruns.all <- aduruns
[…]

### 2015-11-20 analysis/E2/_RUNME.R

@@ -5,10 +5,9 @@ source("2-find.switches.R")
 get.switches(1, FALSE)
 
 # Find switches in randomly-shuffled gap versions
-rand.runs <- 1:5
-make.rand.gaps(rand.runs)
-get.switches(rand.runs, TRUE)
+# rand.runs <- 1:5000
+# make.rand.gaps(rand.runs)
+# get.switches(rand.runs, TRUE)
 
 source("3-plots.stats.analysis.R")
-run.models(rand.runs)
-compare.models()
+source("4-model-functions.R")

### 2015-11-20 paper/CasillasFrank-Anticip.tex
Submitted version to JML (1 Oct 2014).
optimization
Resubmission changes in progress

@@
 \usepackage{graphicx}
-\usepackage{lineno}
+% \usepackage{lineno}
 \usepackage{amssymb}
 \usepackage{wrapfig}
+\usepackage{subfig}
 \usepackage{tipa}
@@
 
-\title{Tracking and predicting turn structure during language acquisition}
-
-\author[MPI,StanfordLX]{Marisa Casillas\corref{cor1}}
+\title{The development of children's ability to track and predict turn structure in conversation}
+
+\author[MPI]{Marisa Casillas\corref{cor1}}
 \address[MPI]{Max Planck Institute for Psycholinguistics, Nijmegen}
-\address[StanfordLX]{Department of Linguistics, Stanford University}
+%\address[StanfordLX]{Department of Linguistics, Stanford University}
 \cortext[cor1]{Corresponding author}
@@
 \begin{abstract}
-We investigate language acquisition through the lens of a fundamental conversational skill: turn taking.
-Children begin developing turn-taking skills in infancy, but take several years to assimilate their growing knowledge of language with this ability.
-In two eye-tracking experiments with children across a wide developmental range, we measured spontaneous predictions about upcoming speaker change while controlling the amount of linguistic information available.
-We found that children already predicted upcoming turns at age one, but they integrated linguistic cues differently at different ages.
+We investigate language acquisition through the lens of a fundamental conversational skill: Turn taking.
+Children begin developing turn-taking skills in infancy, but take several years to assimilate their growing knowledge of language into their turn-taking behavior.
+In two eye-tracking experiments, with children across a wide developmental range, we measured spontaneous predictions about upcoming speaker change while controlling the amount of linguistic information available.
+We found that children already predicted upcoming turns at age one, but that they integrated linguistic cues differently at different ages.
 Children under three showed a general advantage for prosody over lexicosyntax, contrary to prior findings for adults.
@@
 
-\linenumbers
+% \linenumbers
 
@@
 Like other types of human interaction, it is organized at its core by the roles and goals of its participants.
-But what sets conversation apart is its structure: sequences of interconnected, communicative actions that take place across alternating turns at talk.
+But what sets conversation apart is its structure: Sequences of interconnected, communicative actions that take place across alternating turns at talk.
 Sequential, turn-based structures in conversation are strikingly uniform across language communities and linguistic modalities.
-Turn-taking behaviors are also cross-culturally consistent in their basic features and the details of their implementation \citep{de-vosInPrep, dingemanse2013, stivers2009}.
+Turn-taking behaviors are also cross-culturally consistent in their basic features and the details of their implementation \citep{de-vos2015, dingemanse2013, stivers2009}.
 How does this ability develop?
 
-Children participate in sequential coordination with their caregivers starting at three months of age---before they can rely on any linguistic cues in taking turns \citep[see, among others, ][]{bateson1975, hilbrinkInPrep, jaffe2001, snow1977}.
+Children participate in sequential coordination with their caregivers starting at three months of age---before they can rely on any linguistic cues in taking turns \citep[see, among others, ][]{bateson1975, hilbrink2015, jaffe2001, snow1977}.
 Of course, infant turn taking is different from adult turn taking in several ways.
-Infant turn taking is heavily scaffolded by caregivers, has distinct timing in comparison to adult turn taking, and lacks semantic content \citep{hilbrinkInPrep, jaffe2001}.
+Infant turn taking is heavily scaffolded by caregivers, has distinct timing in comparison to adult turn taking, and lacks semantic content \citep{hilbrink2015, jaffe2001}.
 However, children's early, turn-structured social interactions are presumably a critical precursor to their conversational turn taking.
@@
 To take a turn, participants need to make predictions about what conversational content will be relevant next.
-In some cases, relevant next turns are somewhat obvious (e.g., question--response) while, in other cases, there are multiple relevant next actions to choose from, or no obvious next action at all (e.g., after a closed sequence).
+In some cases, relevant next turns are somewhat obvious (e.g., question--response) while, in other cases, there are multiple relevant next actions to choose from, or no obvious next action at all (e.g., after a closing).
 
 Despite this complexity, conversational turn taking is often precise in its timing.
-Across a diverse sample of conversations in 10 languages, one study found a consistent average transition time of 0--200 msec at points of speaker switch \citep{stivers2009}.
+Across a diverse sample of conversations in 10 languages, one study found a consistent average turn transition time of 0--200 msec at points of speaker switch \citep{stivers2009}.
 Experimental results and current models of speech production suggest that it takes approximately 600 msec to produce a content word, and even longer to produce a simple utterance \citep{griffin2000, levelt1989}.
@@
 Adults have a lot of information at their disposal to help make accurate predictions about upcoming turn content.
-Lexical, syntactic, and prosodic information (e.g., \textit{wh}- words, subject-auxiliary inversion, and list intonation) can all inform addressees about upcoming linguistic structure \citep{de-ruiter2006, duncan1972, ford1996, torreiraUndRev}.
+Lexical, syntactic, and prosodic information (e.g., \textit{wh}- words, subject-auxiliary inversion, and list intonation) can all inform addressees about upcoming linguistic structure \citep{de-ruiter2006, duncan1972, ford1996, torreira2015}.
 Non-verbal cues (e.g., gaze, posture, and pointing) often appear at turn-boundaries and can sometimes act as late indicators of an upcoming speaker switch \citep{rossano2009, stivers2010}.
-Additionally, the sequential context of a turn can make it clear what will come next: answers after questions; thanks or denial after compliments, et cetera \citep{schegloff2007}.
+Additionally, the sequential context of a turn can make it clear what will come next: Answers after questions; thanks or denial after compliments, et cetera \citep{schegloff2007}.
 
 Prior work suggests that adult listeners primarily use lexicosyntactic information to accurately predict upcoming turn structure \citep{de-ruiter2006}.
-De Ruiter and colleagues \citeyearpar{de-ruiter2006} asked participants to listen to snippets of spontaneous conversation and to press a button whenever they anticipated that the current speaker was about to finish her turn.
+De Ruiter and colleagues \citeyearpar{de-ruiter2006} asked participants to listen to snippets of spontaneous conversation and to press a button whenever they anticipated that the current speaker was about to finish his or her turn.
 The speech snippets were controlled for the amount of linguistic information present; some were normal, but others had flattened pitch, low-pass filtered speech, or further manipulations.
 De Ruiter and colleagues found that, with pitch-flattened speech, the timing of participants' button responses was comparable to their timing with the full linguistic signal.
-But, when no lexical information was available, participants' responses were significantly earlier than the full linguistic signal.
+But, when no lexical information was available, participants' responses were significantly earlier.
 The authors concluded that lexicosyntactic information\footnote{The ``lexicosyntactic'' condition only included flattened pitch and so was not exclusively lexicosyntactic---the speech would still have residual prosodic structure, including syllable duration and intensity.} was necessary and possibly sufficient for turn-end projection, while intonation was neither necessary nor sufficient.
-Congruent evidence comes from studies varying the predictability of lexicosyntactic and pragmatic content: Adults anticipate turn ends better when they can more accurately predict the exact words that will come next (\citealp{magyari2012}; see also \citealp{magyariUndRev}).
-They can also identify speech acts within the first word of an utterance \citep{gisladottirUndRev}, allowing them to start planning their response at the first moment possible \citep{bogelsmagyariInPrep}.
+Congruent evidence comes from studies varying the predictability of lexicosyntactic and pragmatic content: Adults anticipate turn ends better when they can more accurately predict the exact words that will come next (\citealp{magyari2012}; see also \citealp{magyari2014}).
+They can also identify speech acts within the first word of an utterance \citep{gisladottir2015}, allowing them to start planning their response at the first moment possible \citep{bogels2015}.
 
 The role of prosody for adult turn-prediction is still a matter of debate.
-De Ruiter and colleagues' (2006) experiment only tested the role of intonation, which is only a partial index of prosody.
+De Ruiter and colleagues' (2006) experiment focused on the role of intonation, which is only a partial index of prosody.
 Prosodic structure is also tied closely to the syntax of an utterance, and so the two linguistic signals are difficult to control independently \citep{ford1996}.
-\citet*{torreiraUndRev} used a combination of button-press and verbal responses to investigate the relationship between lexicosyntactic and prosodic cues in turn-end prediction.
+\citet*{torreira2015} used a combination of button-press and verbal responses to investigate the relationship between lexicosyntactic and prosodic cues in turn-end prediction.
 Critically, their stimuli were cross-spliced so that each item had full prosodic cues to accompany the lexicosyntax.
@@
 Knowing this, we could expect that children's acquisition of turn-taking skills is closely tied to their knowledge about language, gaze, gesture, and social cues.
-But children's turn taking starts early in infancy, long before first words or gestures emerge.
+But children's turn taking starts early in infancy, long before their first words or gestures emerge.
 So a primary role for lexicosyntactic cues doesn't fit well with children's pre-verbal turn taking.
@@
 Children's first turn-like structures appear as early as two to three months in proto-conversation with their caregivers \citep{bruner1975, bruner1985}.
-During proto-conversations, caregivers interact with their infants as if they were capable of making meaningful contributions; they take every look, vocalization, arm flail and burp as ``utterances'' in the joint discourse \citep{bateson1975, jaffe2001, snow1977}.
+During proto-conversations, caregivers interact with their infants as if they were capable of making meaningful contributions; they take every look, vocalization, arm flail, and burp as ``utterances'' in the joint discourse \citep{bateson1975, jaffe2001, snow1977}.
 Infants catch onto the structure of proto-conversations quickly.
@@
 A recent study by Hilbrink et al.
-(in preparation) finds that infants' turn timing at three months is often too early or too late: they start vocalizing in overlap on nearly 40\% of their caregivers' turns, and their non-overlapped vocalizations come after an average inter-turn silent gap of 675 msec (adult average: 200 msec).
+(2015) finds that infants' turn timing at three months is often too early or too late: They start vocalizing in overlap on $~$40\% of their caregivers' turns, and their non-overlapped vocalizations come after an average inter-turn silent gap of 350--900 msec (adult average: 200 msec).
 Between four and nine months, children begin to reduce the number of turns happening in overlap while also improving on their average response latency.
-But then, later on, children's response latencies slow down again, peaking at average gaps of 1100 msec at nine months, with only very gradual improvement after that \citep{hilbrinkInPrep}.
-While children's avoidance of overlap is nearly adult-like by nine months, the timing of their non-overlapped responses stays much longer than the 200 msec standard for the next several years \citep{garvey1984, ervin-tripp1979}.
+But then, later on, children's response latencies slow down again, peaking at average gaps of more than 1000 msec at nine months, with only very gradual improvement after that \citep{hilbrink2015}.
+While children's avoidance of overlap is nearly adult-like by nine months, the timing of their non-overlapped responses stays much longer than the 200 msec standard for the next few years \citep{garvey1984, ervin-tripp1979}.
 
 The protracted development of children's timing may be attributable to their linguistic development: Taking turns on time is easier when the response is a simple vocalization rather than a linguistic utterance.
-Integrating language into the turn-taking system may be one major factor in children's delayed responses \citep{casillasUndRev}.
+Integrating language into the turn-taking system may be one major factor in children's delayed responses \citep{casillas2016}.
 If response planning (i.e., language production) is the primary hurdle in children's spontaneous turn taking, we should find evidence that children understand turn-taking behaviors before they are able to produce the behaviors themselves; this hypothesis has been recently explored in experimental settings.
@@
 At ten months, infants expect people to look and talk at other people, and not to objects \citep{beier2012}.
-At twelve months infants expect to see responses to verbal (but not non-speech) utterances in face-to-face contexts \citep{thorgrimssonUndRev}.
+At twelve months infants expect to see responses to verbal (but not non-speech) utterances in face-to-face contexts \citep{thorgrimsson2015}.
 
@@
 One study found that 12-month-olds make more predictive gaze shifts to a responder while watching human verbal conversation compared to conversation-like interactions with objects \citep{bakker2011}, but another only found a similar effect at 36 months \citep{hofsten2009}.
-Neither of these two studies had baselines to which the turn-relevant looking behavior could be compared, however.
+However, neither of these two studies had baselines to which the turn-relevant looking behavior could be compared.
 A baseline measurement is critical because there may be developmental differences in gaze shifting between conversational participants, even if the shifting is not related to turn structure.
@@
 Their study was the first to focus on the role of linguistic processing in children's turn predictions.
-They showed their participants two types of conversation videos: one normal and one with flattened pitch (i.e., with flattened intonation contours), finding that only 36-month-olds were affected by a lack of intonation contours.
+They showed their participants two types of conversation videos: One normal and one with flattened pitch (i.e., with flattened intonation contours), finding that only 36-month-olds were affected by a lack of intonation contours.
 The adult control group made equal numbers of anticipatory looks in the videos, with and without intonation contours, consistent with prior adult findings \citep{de-ruiter2006}.
@@
 Although the Keitel et al.
-\citeyearpar{keitel2013} study constitutes a substantial advance over previous work, it nevertheless has several limitations.
+\citeyearpar{keitel2013} study constitutes a substantial advance over previous work, it has its own limitations.
 Because these limitations directly inform our own study design, we review them in some detail.
-First, their estimates of baseline gaze frequency (``random'' in their terminology) were computed from switches that occurred specifically while the speaker's turn was ongoing; in contrast, estimates of anticipatory shifting frequency were computed from switches that occurred within the time window immediately surrounding inter-turn gaps.
-Assuming that gaze switches happen most often during the inter-turn gap \citep{hirvenkari2013}, the switches made during ongoing speech (their baseline) would be minimized.
-This baseline thus maximized the possibility of finding a difference between ``random'' and anticipatory gaze frequencies.
+First, their estimates of baseline gaze frequency (``random'' in their terminology) were not random.
+Instead, they used gaze switches during ongoing speech as a baseline, during which switching is least likely to occur \citep{hirvenkari2013} and thereby maximizing their chances of finding a difference between gaze frequency at turn transitions and their baseline rate.
 A more conservative baseline would be to compare participants' looking behavior at turn transitions to their looking behavior during randomly-selected windows of time throughout the stimulus.
@@
 The speakers in the videos were also asked to minimize their movements while performing a scripted and adult-directed conversation, which would have created a somewhat unnatural stimulus.
-In addition, in order to produce more naturalistic conversation, it would have been ideal to localize sources for the two voices in the video (i.e., to have the voices come out of separate left and right speakers).
+Additionally, in order to produce more naturalistic conversation, it would have been ideal to localize the sound sources for the two voices in the video (i.e., to have the voices come out of separate left and right speakers).
 But both voices were recorded and played back on the same audio channel, which may have made it more difficult to distinguish the two talkers.
@@
 \subsubsection{Prosodic development}
-The roles of prosody and lexicosyntax in children's turn predictions are currently unknown, but children understand more about prosody early in life.
+The roles of prosody and lexicosyntax in children's turn predictions are currently unknown, but children understand more about prosody than lexicosyntax early in life.
 Children begin to acquire prosody in the womb, and can distinguish their native language's rhythm type from others (e.g., syllable-timed vs.
@@
 In comparison, children show only a limited lexical inventory at six months, and begin to recognize function words just before their first birthdays, with syntactic categorization beginning around 14 months \citep{bergelson2013, shi2010}.
-Two-month-olds notice changes in word order, but this ability appears to rely on prosodic cueing \citep{mandel1996}.
+Two-month-olds also notice changes in word order, but this ability appears to rely on prosodic cueing \citep{mandel1996}.
 Generally speaking then, our current knowledge about children's linguistic development points to a possible early advantage for prosody in children's turn-taking predictions.
@@
 We focus in particular on how children use prosodic and lexicosyntactic information to make their predictions.
-Prior work has focused mainly on lexicosyntax and intonation, and not on prosody proper (\citealp{de-ruiter2006}; \citealp{keitel2013}, but see \citealp{torreiraUndRev}), even though infants seem to acquire the basic rhythmic properties of the prosodic signal first \citep{mehler1988, moon1993, nazzi2003}.
+Prior work has focused mainly on lexicosyntax and intonation, and not on prosody proper (\citealp{de-ruiter2006}; \citealp{keitel2013}, but see \citealp{torreira2015}), even though infants seem to acquire the basic rhythmic properties of the prosodic signal first \citep{mehler1988, moon1993, nazzi2003}.
 
 In two eye-tracking experiments, we measured children's anticipatory gaze to upcoming responders while controlling for the amount of lexicosyntactic and prosodic information available.
-Experiment 1 uses a paradigm in which English-speaking participants view video clips of naturalistic conversations from several different languages in order to create conditions under which lexicosyntactic information was differentially available, but prosodic and non-linguistic cues were intact.
-This paradigm showed minimal differences between the predictive looking behavior of preschoolers and adults.
-In Experiment 2, we created artificial (puppet) visual stimuli to better separate these linguistic cues from one another.
+In Experiment 1, English-speaking participants viewed video clips of naturalistic conversation from several different languages.
+We used multiple languages to control for the presence of lexicosyntactic information while keeping prosodic and non-linguistic cues intact.
+The results showed minimal differences between the predictive looking behavior of preschoolers and adults.
+In Experiment 2, we created artificial (puppet) visual scenes, enabling us to separately control lexicosyntactic and prosodic cues in the conversational stimuli.
 In this more controlled paradigm, we found that children's predictive looking behavior improved from ages one to six, but that even one-year-olds made more anticipatory looks than would be expected by chance.
@@
 
-We recorded participants' eye movements as they watched six short videos of two-person (dyadic) conversation.
-Each video featured an improvised discourse in one of five languages (English, German, Hebrew, Japanese, and Korean).
+We recorded participants' eye movements as they watched six short videos of two-person (dyadic) conversation interspersed with attention-getting filler videos.
+Each conversation video featured an improvised discourse in one of five languages (English, German, Hebrew, Japanese, and Korean); participants saw two videos in English and one in every other language.
 The participants, all native English speakers, were only expected to understand the two videos in English.
@@
 Using a Marantz PMD 660 solid state field recorder, we captured audio from two lapel microphones, one attached to each participant, while simultaneously recording video from the built-in camera of a MacBook laptop computer.
-The talkers were volunteers, all members of the local post-graduate community, and were acquainted with their recording partner ahead of time.
+The talkers were volunteers and were acquainted with their recording partner ahead of time.
 
@@
 This gave each voice in the videos a localized source (from the left or right loudspeaker).
-We coded each turn transition in the videos for language group (English vs.
+We coded each turn transition in the videos for language condition (English vs.
 non-English), inter-turn gap duration (in milliseconds), and speech act (question vs.
 non-question).
-The non-English stimuli were coded for speech act from a monolingual English-speaker's perspective, i.e., which turns ``sound like'' questions, and which don't.
+The non-English stimuli were coded for speech act from a monolingual English-speaker's perspective, i.e., which turns ``sound like'' questions, and which don't: we asked five native American English speakers to listen to the audio signal for each turn and judge whether it sounded like a question.
+We then coded turns with at least 80\% ``yes'' responses as questions.
 
@@
 We measured the duration of each turn transition from the audio recording associated with each video.
-We excluded turn transitions longer than 550 msec and shorter than 0 msec (those with transitional overlap) from analysis.\footnote{Overlap occurs when a responder begins a new turn before the current turn is finished.
-When overlap occurs, observers cannot switch their gaze in anticipation of the response because the response begins earlier than expected; participants expect conversations to proceed with ``one speaker at a time'' \citep{sacks1974}.
-As such, they would still be fixated on the prior speaker when the overlap started, and then would have to switch their gaze \textit{reactively} to the responder.} This left approximately equal numbers of turn transitions available for analysis in the English (N$=$22) and non-English (N$=$21) videos.
-On average, the inter-turn gaps for English videos (mean$=$273 msec, median$=$291 msec) were slightly longer than for non-English videos (mean$=$226 msec, median$=$205 msec).
+We excluded turn transitions longer than 550 msec and shorter than 90 msec, including overlapped transitions, from analysis.\footnote{Overlap occurs when a responder begins a new turn before the current turn is finished.
+When overlap occurs, observers cannot switch their gaze in anticipation of the response because the response began earlier than expected; participants expect conversations to proceed with ``one speaker at a time'' \citep{sacks1974}.
+As such, they would still be fixated on the prior speaker when the overlap started, and then would have to switch their gaze \textit{reactively} to the responder.} This left approximately equal numbers of turn transitions available for analysis in the English (N$=$20) and non-English (N$=$16) videos.
+On average, the inter-turn gaps for English videos (mean$=$318, median$=$302, stdv$=$112 msec) were slightly longer than for non-English videos (mean$=$286, median$=$251, stdv$=$122 msec).
 The longer gaps in the English videos could give them a slight advantage: Our definition of an ``anticipatory gaze shift'' includes shifts that are initiated during the gap between turns (Figure \ref{fig:criterion}), so participants had slightly more time to make anticipatory shifts in the English videos.
 
-Questions made up approximately half of the turn transitions in the English (54\%; N$=$12) and non-English (48\%; N$=$10) videos.
-In the English videos, inter-turn gaps were slightly shorter for questions (mean$=$271 msec, median$=$254 msec) than non-questions (mean$=$276 msec, median$=$300 msec).
-The opposite was true for the non-English videos: question transitions (mean$=$264 msec, median$=$257 msec) had longer inter-turn gaps than non-question transitions (mean$=$191 msec, median$=$168 msec).
-Thus non-question (i.e., declarative) transitions in non-English videos were the briefest type of turn transition in the stimuli on average.
+Questions made up exactly half of the turn transitions in the English (N$=$10) and non-English (N$=$8) videos.
+In the English videos, inter-turn gaps were slightly shorter for questions (mean$=$310, median$=$293, stdev$=$112 msec) than non-questions (mean$=$325, median$=$315, stdv$=$118 msec).
+Non-English videos did not show a large difference in transition time for questions (mean$=$270, median$=$257, stdv$=$116 msec) and non-questions (mean$=$302, median$=$252, stdv$=$134 msec).
 
@@
 Before the experiment started, we warned adult participants that they would see videos in several languages and that, though they weren't expected to understand the content of non-English videos, we \textit{would} ask them to answer general, non-language-based questions about the conversations.
-Then after each video we asked participants one of the following randomly-assigned questions: ``which speaker talked more?'', ``which speaker asked the most questions?'', ``which speaker seemed more friendly?'', and ``did the speakers' level of enthusiasm shift during the conversation?'' We also asked if the participants could understand any of what was said after each video.
+Then after each video we asked participants one of the following randomly-assigned questions: ``Which speaker talked more?'', ``Which speaker asked the most questions?'', ``Which speaker seemed more friendly?'', and ``Did the speakers' level of enthusiasm shift during the conversation?'' We also asked if the participants could understand any of what was said after each video.
 The participants responded verbally while an experimenter noted their responses.
 
-Children were less inclined to simply sit and watch videos of conversation in languages they didn't speak, so we used a different procedure to keep them engaged.
-The experimenter started each session by asking the child about what languages he or she could speak, and about what other languages he or she had heard of.
+Children were less inclined to simply sit and watch videos of conversation in languages they didn't speak, so we used a different procedure to keep them engaged: The experimenter started each session by asking the child about what languages he or she could speak, and about what other languages he or she had heard of.
 Then the experimenter expressed her own enthusiasm for learning about new languages, and invited the child to watch a video about ``new and different languages'' together.
 If the child agreed to watch, the experimenter and the child sat together in front of the display, with the child centered in front of the tracker and the experimenter off to the side.
-If the child began to look bored during video playback, the experimenter would talk during the filler videos, either commenting on the previous conversation (``That was a neat language!'') or giving the language name for the next conversation (``This next one is called Hebrew.
-Let's see what it's like.'').
-The experimenter's comments reinforced the video-watching as a joint task.
+Each conversation video was preceded and followed by a 15--30 second attention-getting filler video (e.g., running puppies, singing muppets, flying bugs).
+If the child began to look bored, the experimenter would talk during the fillers, either commenting on the previous conversation (``That was a neat language!'') or giving the language name for the next conversation (``This next one is called Hebrew.
+Let's see what it's like.'') The experimenter's comments reinforced the video-watching as a joint task.
 
@@
 We used a dancing Elmo for the children's calibration image.
-During the experiment, participants watched all six 30-second conversation videos with 15--30 second filler videos in-between them (e.g., running puppies, singing muppets, flying bugs).
+During the experiment, participants watched all six 30-second conversation videos.
 The first and last conversations were in American English and the intervening conversations were Hebrew, Japanese, German, and Korean.
@@
 Following prior work, we measured how often our participants shifted their gaze from the prior to the upcoming speaker \textit{before} the shift in gaze could have been initiated in reaction to the onset of the speaker's response.
-In doing so, we assumed that it takes children 333 msec to plan an eye movement, following Fernald and colleagues' \citeyearpar{fernald2008} similar `looking while listening' method.
-We assumed that it takes adults 200 msec to plan an eye movement, following standards from adult anticipatory processing studies \citep[e.g., ][]{kamide2003}.
-
-We checked each participant's gaze at each turn transition for three characteristics (Figure \ref{fig:criterion}): (1) that the participant fixated on the prior speaker for at least 100 msec at the end of the prior turn, (2) that sometime thereafter the participant switched to fixate on the upcoming speaker for at least 100 ms, and (3) that the switch in gaze was initiated within the first 333 msec of the response turn, or earlier.
-These criteria guarantee that we only counted gaze shifts when: (1) participants were tracking the previous speaker, (2) switched their gaze to track the upcoming speaker, and (3) did so before they could have simply reacted to the onset of speech in the response.
-Under this assumption, a gaze shift that was initiated within the first 333 msec of the response (or earlier) was planned \textit{before} the child could react to the onset of speech itself.
+In doing so, we assumed that it takes participants 200 msec to plan an eye movement, following standards from adult anticipatory processing studies \citep[e.g., ][]{kamide2003}.
+
+We checked each participant's gaze at each turn transition for three characteristics (Figure \ref{fig:criterion}): (1) That the participant fixated on the prior speaker for at least 100 msec at the end of the prior turn, (2) that sometime thereafter the participant switched to fixate on the upcoming speaker for at least 100 ms, and (3) that the switch in gaze was initiated within the first 200 msec of the response turn, or earlier.
+These criteria guarantee that we only counted gaze shifts when: (1) Participants were tracking the previous speaker, (2) switched their gaze to track the upcoming speaker, and (3) did so before they could have simply reacted to the onset of speech in the response.
+Under this assumption, a gaze shift that was initiated within the first 200 msec of the response (or earlier) was planned \textit{before} the child could react to the onset of speech itself.
 
@@
 For example, in a button press task, adult participants anticipate turn ends approximately 200 msec in advance of the turn's end, and anticipatory responses to pitch-flattened stimuli come even earlier \citep{de-ruiter2006}.
-We therefore allowed switches to occur as early as 333 msec before the end of the prior turn for children, and 200 msec before the end of the prior turn for adults.
-For very early and very late switches, our requirement for 100 msec of fixation on each speaker would sometimes extend outside of the transition window boundaries (333 msec before and after the inter-turn gap).
-The maximally available fixation window was 100 msec before and after the earliest and latest possible switch point (433 msec before and after the inter-turn gap).
+We therefore allowed switches to occur as early as 200 msec before the end of the prior turn.
+For very early and very late switches, our requirement for 100 msec of fixation on each speaker would sometimes extend outside of the transition window boundaries (200 msec before and after the inter-turn gap).
+The maximally available fixation window was 100 msec before and after the earliest and latest possible switch point (300 msec before and after the inter-turn gap).
 We did not count switches made during the fixation window as anticipatory.
 We \textit{did} count switches made during the inter-turn gap.
-The period of time from the beginning of the possible fixation window on the prior speaker to the end of the possible fixation window on the responder was our total analysis window (433 msec $+$ the inter-turn gap $+$ 433 msec), which we use to create the random baselines described below.
-For adult participants we used 200 msec (instead of 333 msec) as the assumed time to plan an eye movement in defining all analysis windows.
+The period of time from the beginning of the possible fixation window on the prior speaker to the end of the possible fixation window on the responder was our total analysis window (300 msec $+$ the inter-turn gap $+$ 300 msec).
 
@@
 \end{center}
-\caption{Schematic summary of criteria for anticipatory gaze shifts from speaker A to speaker B during a turn transition.
-The criteria were the same for adults and children, except children were assumed to have a 333 msec window for planning eye movements, while adults were only given 200 msec.}
+\caption{Schematic summary of criteria for anticipatory gaze shifts from speaker A to speaker B during a turn transition.} 
 \label{fig:criterion}
@@
 
-\paragraph{Random baseline analysis} To ensure that participants' looking behavior was turn-related and not simply the result of random switching, we corrected for the probability of making an anticipatory switch at random.
-We estimated the baseline chance of making an anticipatory switch at each turn transition for each participant.
-We ran each participants' eye-tracking signal through our switch identification procedure \ref{sec:algorithm} with 100 randomly-shuffled versions of the analysis windows (Figure \ref{fig:shuffling}).
-For each of the 100 randomly-shuffled versions, we recorded whether the participant made a ``switch'' or not.
-We then averaged participants' switches from all 100 randomly-shuffled versions of the analysis windows, yielding a value between 0 and 1 for each turn transition for each participant.
-We then subtracted the random baseline value from each participant's actual switch result (1 or 0) for each turn transition to arrive at the final, baseline-corrected values.
-Subtracting the random baseline this way gives us an estimate of how often participants switched above chance across ages, conditions, and transition types.
-If participants made the same number of anticipatory switches, regardless of turn transition location, there would be no evidence that their switching behavior was linked to the conversation stimulus.
-
-\begin{figure}[t]
-\begin{center}
-\includegraphics[width=0.6\textwidth]{figures/FIG-ShuffledWindows.png}
-\end{center}
-\caption{Example of shuffling for five turn transition analysis windows. The windows were $\pm$433 msec around the inter-turn gap for children, and from $\pm$300 msec for adults.} 
-\label{fig:shuffling}
-\end{figure}
+\paragraph{Predictions}
+We expected participants to show greater anticipation in the English videos than in the non-English videos because of their increased access to linguistic information in English.
+We also predicted that anticipation would be greater following questions compared to non-questions; questions have early cues to upcoming turn transition (e.g., \textit{wh-} words, subject-auxiliary inversion), and also make a next response immediately relevant.
+Our third prediction was that anticipatory looks would increase with development, along with children's increased linguistic competence.
 
@@
 \label{sec:results1}
-
-\subsubsection{Descriptive analysis}
 
@@
 \end{table}
-
-We identified anticipatory gaze switches for all 43 usable turn transitions, based on the criteria outlined in Section \ref{sec:algorithm}.
-Because participants had greater access to linguistic information in English, we expected to see greater anticipation in the English videos than in the non-English videos.
-We also predicted that anticipation would be greater following questions compared to non-questions; questions have early cues to upcoming turn transition (e.g., \textit{wh-} words, subject-auxiliary inversion), and also make a next response immediately relevant.
-We also predicted that anticipatory looks would increase with development, along with children's increased linguistic competence.
-
-\begin{figure}[t]
-\begin{center}
-\includegraphics[width=0.8\textwidth]{figures/FIG-randvsreal-FL.png}
-\end{center}
-\caption{Proportion of anticipatory gaze switches made for actual (solid) and randomly-shuffled (dashed) turn transitions in each condition, and across ages.
+%
+%\begin{figure}[t]
+%\begin{center}
+%\includegraphics[width=0.8\textwidth]{figures/FIG-randvsreal-FL.png}
+%\end{center}
+%\caption{Proportion of anticipatory gaze switches made for actual (solid) and randomly-shuffled (dashed) turn transitions in each condition, and across ages.
 Error bars represent the standard error of the mean.}
-\label{fig:randvsrealFL}
-\end{figure}
-
-To estimate the effects of linguistic condition and age on switching behavior, we estimated each participants' random baseline switching behavior for each turn transition.
+%\label{fig:randvsrealFL}
+%\end{figure}
+%
+%To estimate the effects of linguistic condition and age on switching behavior, we estimated each participants' random baseline switching behavior for each turn transition.
 Participants at all ages made anticipatory gaze switches more often than would be expected by chance, whether they could understand the language in the video or not (Figure \ref{fig:randvsrealFL}).
@@
 For all further analyses we then corrected each participant's data by subtracting their random baseline estimations from their actual switching data for each turn transition.
-
-Overall, a greater proportion of participants made anticipatory switches in the English videos (0.17) compared to the non-English videos (0.09).
+%
+%Overall, a greater proportion of participants made anticipatory switches in the English videos (0.17) compared to the non-English videos (0.09).
 But the tendency to make more anticipatory switches for English than non-English videos disappeared with age (Figure \ref{fig:conditionsFL}).
@@
 Otherwise, age-based differences in predictive gaze shifts were minor.
-
-\begin{figure}[t]
-\begin{center}
-\includegraphics[width=0.5\textwidth]{figures/FIG-conditions-FL.png}
-\end{center}
-\caption{Baseline-corrected proportion of anticipatory gaze switches made for turn transitions in each language condition, and across ages. Error bars represent the standard error of the mean.} 
-\label{fig:conditionsFL}
-\end{figure}
-
-\begin{figure}[t]
-\begin{center}
-\includegraphics[width=0.8\textwidth]{figures/FIG-QvsNQ-FL.png}
-\end{center}
-\caption{Baseline-corrected proportion of anticipatory gaze switches made for question- (dashed) and non-question- (dotted) turn transitions in each condition, and across ages.
+%
+%\begin{figure}[t]
+%\begin{center}
+%\includegraphics[width=0.5\textwidth]{figures/FIG-conditions-FL.png}
+%\end{center}
+%\caption{Baseline-corrected proportion of anticipatory gaze switches made for turn transitions in each language condition, and across ages. Error bars represent the standard error of the mean.} 
+%\label{fig:conditionsFL}
+%\end{figure}
+%
+%\begin{figure}[t]
+%\begin{center}
+%\includegraphics[width=0.8\textwidth]{figures/FIG-QvsNQ-FL.png}
+%\end{center}
+%\caption{Baseline-corrected proportion of anticipatory gaze switches made for question- (dashed) and non-question- (dotted) turn transitions in each condition, and across ages.
 Error bars represent the standard error of the mean.}
-\label{fig:questionsFL}
-\end{figure}
-
-Participants also showed an advantage for questions (Figure \ref{fig:questionsFL}).
+%\label{fig:questionsFL}
+%\end{figure}
+%
+%Participants also showed an advantage for questions (Figure \ref{fig:questionsFL}).
 Children made anticipatory switches twice as often when they heard a question, compared to a non-question (18.7\% and 7.3\%).
@@
 The advantage for questions didn't emerge until age five in the non-English videos.
-
-\subsubsection{Statistical analysis}
-
-We quantified these findings using two linear mixed-effects models of participants' baseline-corrected anticipatory switches \citep{lme4, R}.
+%
+
+\subsubsection{Statistical models}
+\label{sec:models1}
+
+We identified anticipatory gaze switches for all 36 usable turn transitions, based on the criteria outlined in Section \ref{sec:algorithm}, and analyzed them for effects of language, transition type, and age with two mixed-effects logistic regressions \citep{lme4, R}.
 We built one model each for children and adults.
+We modeled children and adults separately because effects of age are only pertinent to the children's data.
 The child model included condition (English vs.
-non-English)\footnote{Because each non-English language was represented by a single stimulus, we cannot compute reliable cross-language differences.
-Gaze behavior might be best for languages that have the most structural overlap with participants' native language: English speakers can make predictions about the strength of upcoming Swedish prosodic boundaries nearly as well as Swedish speakers do, but Chinese speakers are at a disadvantage in the same task \cite{carlson2005}.
-We would need multiple items from each of the language to check for similarity effects of specific linguistic features.}, transition type (question vs.
-non-question), age, and duration of the inter-turn gap as predictors, with full interactions between condition, transition type, and age.
+non-English)\footnote{Because each non-English language was represented by a single stimulus, we cannot treat individual languages as factors.
+Gaze behavior might be best for non-native languages that have the most structural overlap with participants' native language: English speakers can make predictions about the strength of upcoming Swedish prosodic boundaries nearly as well as Swedish speakers do, but Chinese speakers are at a disadvantage in the same task \citep{carlson2005}.
+We would need multiple items from each of the languages to check for similarity effects of specific linguistic features.}, transition type (question vs.
+non-question), age (3, 4, 5), and duration of the inter-turn gap (seconds, e.g., 0.441) as predictors, with full interactions between condition, transition type, and age.
 We included the duration of the inter-turn gap as a predictor since longer gaps also provide more opportunities to make anticipatory switches (Figure \ref{fig:criterion}).
-We additionally included random effects of item (turn transition) and participant, with random slopes of condition, transition type, and their interaction for participants \citep{barr2013}.
-The adult model was exactly the same as the children's, excluding age-related effects.
-
-Children's anticipatory gaze switches showed a main effect of language condition (\textit{$\beta$}=-0.448, \textit{SE}=0.126, \textit{t}=-3.56, \textit{p}$<$.001).\footnote{We derive \textit{p}-values here by treating \textit{t} as \textit{z} (with a two-tailed assumption), as is justified for large data sets \citep{barr2013}.} As predicted, children made more anticipatory looks while they were watching videos that they could understand.
-In the English videos, children had access to lexicosyntactic information, which could have helped them make better predictions about upcoming turn structure.
-Although children performed better on the English videos, lexicosyntactic information was not exclusively responsible for their anticipatory switches; children still performed remarkably well when no lexical information was present at all, and even three-year-olds made anticipatory switches at an above-chance rate in the non-English videos.
-
-Contrary to our expectation, there was no main effect of age (\textit{$\beta$}=-0.03, \textit{SE}=0.027, \textit{t}=-1.145, \textit{p}=.252).
-Children's anticipatory switches \textit{did} increase with age for the non-English videos, but it also decreased with age for the English videos, resulting in a significant interaction of age and language condition (\textit{$\beta$}=0.083, \textit{SE}=0.027, \textit{t}=3.063, \textit{p}$<$.001).
-We had expected that, as children got older, they would also show more anticipatory switches.
-But this is contradicted by decreased looks with age for English videos, which even extended to the adult data.
-One possible explanation is that older children and adults found the videos in English easy to comprehend, and so tracked the conversation less closely.
-Compared to three-year-olds, adults and older children spent more time looking away from the current speaker and addressee in the English conversations.
-In contrast, compared to three-year-olds, they spent more time looking at the current speaker and less time looking away in the non-English conversations (Table \ref{tab:e1_look}).
-
-It is also possible that the timing of anticipatory gaze switches changes as children get older.
-That is, the moment at which observers chose to launch their gaze away from the current speaker might change with age.
-For example, children could have reacted more immediately to early cues in the turn (e.g., ``Which one...'') as they got older.
-If a change in timing were sufficiently large, we might not capture it in our analysis window, which was designed to capture switches close to the turn transition.
-To test for a change in switch timing, we would have to run new analyses on a larger set of turn transitions---ones with controlled early and late cues \citep[see, e.g.,][]{bogelsmagyariInPrep}.
-Within the current analysis windows there were some ``early switches'' (i.e., those launched before the end of the prior turn).
-But participants made approximately equal numbers of early switches, regardless of age and condition: 11--15\% (Figure \ref{fig:cumulativeFL}).
-The lack of increased early switching with age might suggest that changes in timing are either larger than the analysis window, or are not at play.
-
-\begin{figure}[t]
+We additionally included random effects of item (turn transition) and participant, with random slopes of condition, transition type, and their interaction for participants \citep{barr2013}.\footnote{The models we report are all qualitatively \textit{unchanged} by the exclusion of their random slopes.
+We have left the random slopes in because of minor participant-level variation in the predictors modeled.} The adult model included condition, transition type, duration, and their interactions as predictors with participant and item included as random effects and random slopes of condition, transition type, and their interaction for participant.
+
+\begin{table}[t]
 \begin{center}
-\includegraphics[width=0.8\textwidth]{figures/FIG-cumulative-FL.png}
+  \begin{tabular}{lcccc}
+  \textbf{\textit{Children}} &&&& \\
+    \hline
+                   & Estimate & Std. Error & z value & Pr($>$$|$z$|$)    \\
+    \hline
+(Intercept)   &      -0.96146 &   0.84901 & -1.132 & 0.257446 \\    
+Age             &    -0.18268 &   0.17507 & -1.043 & 0.296725   \\  
+LgGroupNE &          -3.29347 &   0.96045 & -3.429 & 0.000606 \\
+TypeS          &     -1.10129 &   0.86494 & -1.273 & 0.202925     \\
+Duration      &       3.40169 &   1.22826 &  2.770 & 0.005614  \\
+Age:LgGroupNE &       0.52065 &   0.21190 &  2.457 & 0.014008 \\  
+Age:TypeS          & -0.01628 &   0.19437 & -0.084 & 0.933232     \\
+LgGroupNE:TypeS &     2.68166 &   1.35016 &  1.986 & 0.047013 \\  
+Age:LgGroupNE:TypeS & -0.45632 &   0.30163 & -1.513 & 0.130315  \\   
+    \hline
+  &&&& \\
+  \textbf{\textit{Adults}} &&&& \\
+    \hline
+                       &  Estimate & Std. Error & z value & Pr($>$$|$z$|$) \\    
+    \hline
+(Intercept)       &        -0.1966 &    0.6942 & -0.283 & 0.776988     \\
+LgGroupNE    &             -0.8812 &    0.9602 & -0.918 & 0.358754   \\  
+TypeS             &        -4.4953 &    1.3139 & -3.421 & 0.000623  \\
+Duration          &        -1.1227 &    1.9880 & -0.565 & 0.572238     \\
+LgGroupNE:TypeS &           3.2972 &    1.6101 &  2.048 & 0.040581  \\  
+LgGroupNE:Duration &        1.3626 &    3.0077 &  0.453 & 0.650527     \\
+TypeS:Duration          &  10.5107 &    3.3459 &  3.141 & 0.001682  \\
+LgGroupNE:TypeS:Duration & -6.3156 &    4.4926 & -1.406 0.159790 \\    
+    \hline
+  \end{tabular}
 \end{center}
-\caption{Cumulative proportion of anticipatory gaze switches planned during the transition window, across conditions and ages.
-Age is indicated by hue (darker = older).
-The solid line is the offset of the pre-gap turn, and the dashed line is the median onset of the post-gap turn in each condition.
-``Early switches'' occur before 0.}
-\label{fig:cumulativeFL}
+  \caption{Model output for children and adults' anticipatory gaze switches.}
+\label{tab:E1-models}
+\end{table}
+
+Children's anticipatory gaze switches showed effects of language condition (\textit{$\beta$}=-3.29, \textit{SE}=0.961, \textit{t}=-3.43, \textit{p}$<$.001) and gap duration (\textit{$\beta$}=3.4, \textit{SE}=1.229, \textit{t}=2.77, \textit{p}$<$.01) with additional effects of a language condition-transition type interaction (\textit{$\beta$}=2.68, \textit{SE}=1.35, \textit{t}=1.99, \textit{p}$<$.05) and an age-language condition interaction (\textit{$\beta$}=0.52, \textit{SE}=0.212, \textit{t}=2.46, \textit{p}$<$.05).
+There were no significant effects of age or transition type alone (\textit{$\beta$}=-0.18, \textit{SE}=0.175, \textit{t}=-1.04, \textit{p}$=$.3 and \textit{$\beta$}=-1.10, \textit{SE}=0.865, \textit{t}=-1.27, \textit{p}$=$.2, respectively).
+
+Adults' anticipatory gaze switches shows an effect of transition type (\textit{$\beta$}=-4.5, \textit{SE}=1.314, \textit{t}=-3.42, \textit{p}$<$.001) and significant interactions between language condition and transition type (\textit{$\beta$}=3.3, \textit{SE}=1.61, \textit{t}=2.05, \textit{p}$<$.05) and transition type and gap duration (\textit{$\beta$}=10.51, \textit{SE}=3.346, \textit{t}=3.141, \textit{p}$<$.01).
+
+\begin{figure}[h]
+\begin{center}
+\includegraphics[width=0.8\textwidth]{figures/FIG-ShuffledWindows.png}
+\end{center}
+\caption{Example of shuffling for five turn transition analysis windows. The windows were $\pm$300 msec around the inter-turn gap.} 
+\label{fig:shuffling}
 \end{figure}
 
-Children made more anticipatory switches after hearing questions than non-questions, but the advantage for questions was driven by age and linguistic access.
-There was no main effect of transition type alone (\textit{$\beta$}=-0.209, \textit{SE}=0.151, \textit{t}=-1.384, \textit{p}=.166).
-Rather, the model confirmed a two-way interaction of linguistic condition and transition type (\textit{$\beta$}=0.445, \textit{SE}=0.19, \textit{t}=2.344, \textit{p}$<$.02).
-Presumably, question effects were larger in the English videos than the non-English videos because there were more cues to questionhood in English (i.e., non-verbal, prosodic, \textit{and} lexicosyntactic).
-Children younger than five did not distinguish questions from non-questions in the non-English videos (Figure \ref{fig:questionsFL}).
-The children's model confirmed a three-way interaction of age, linguistic condition, and transition type (\textit{$\beta$}=-0.091, \textit{SE}=0.042, \textit{t}=-2.183, \textit{p}$<$.05).
-Lexicosyntactic information might then be critical for young children to identify questions in conversation.
-Cues like subject-auxiliary inversion, and \textit{wh}-words occur early in the turn and are clear indicators that an answer will come next.
-Importantly, the turns marked as ``questions'' in the non-English data were not real questions, but turns that \textit{sounded like} questions to native English speakers.
-Question-like prosodic cues in the non-English videos could have been markedly different from question prosody in English, making it more difficult for three- and four-year-olds to identify ``questions'' in those conversations.
-
-The increased switching after questions indicates that, so long as they have native lexical and prosodic cues, children as young as three spontaneously distinguish questions from other types of speech acts in third-party conversation.
-Children may treat questions differently from non-questions in conversation because they are predictably followed by an immediate response (i.e., the answer).
-Because the form of an answer is partially determined by its question (e.g., locative expressions for ``where'' questions), children could even be predicting what type of response they will hear, and look at the responder for confirmation.
-In contrast, non-questions do not usually require an immediate response, so it is not as easy to predict what will come next.
-
-Inter-turn gap duration did not affect children's anticipatory switches (\textit{$\beta$}=0.202, \textit{SE}=0.151, \textit{t}=1.340, \textit{p}=.180), but it was the main factor driving adults' switches (\textit{$\beta$}=0.648, \textit{SE}=0.047, \textit{t}=13.855, \textit{p}$<$.001).
-The adult model also showed a marginal effect of transition type, with adults looking more after questions compared to non-questions (\textit{$\beta$}=-0.117, \textit{SE}=0.069, \textit{t}=-1.704, \textit{p}=.088).
-Language condition and its interaction with transition type were not significant.
-Language did not affect adults' anticipatory looking behavior, possibly because the animated, child-directed prosody and non-verbal cues made it possible for them to follow the interaction and track the current speaker easily.
+\subsubsection{Random baseline comparison}
+
+We estimated the probability that these patterns were the result of random looking by running the same regression models on participants' real eye-tracking data, only this time calculating their anticipatory gaze switches with respect to randomly permuted turn transitions (Figure \ref{fig:shuffling}).
+This involved: (1) randomizing the order and temporal placement of the analysis windows (Figure \ref{fig:criterion}) for each stimulus, thereby randomly redistributing the windows of analysis across the eye-tracking signal for each stimulus, then (2) re-running each participant's eye tracking data through switch identification (described in \ref{sec:algorithm}) with the randomly permuted analysis windows, and finally (3) entering the resulting anticipatory gazes into the same statistical model we used with the original data (above).
+Properties of each turn transition (e.g., prior speaker identity, transition type, duration, etc.) stayed constant, even though its onset and offset time within the eye-tracking signal was shuffled.
+
+In effect, this procedure de-links participants' gaze data from the turn structure in the original stimulus, thereby allowing us to compare turn-related (original) and non-turn-related (randomly permuted) looking behavior from the same data.
+The random permutations represent average anticipatory gaze rate over all possible starting points; a random baseline.
+If we anticipatory gaze rates turned out to the same for the original and randomly permuted versions of the data, we would have no evidence for turn-relevant gaze switching.
+Further, by running both data sets through the identical statistical models, we can estimate how likely it is that predictor effects in the original data (e.g., the effect of language condition) arose from random looking.
+
+We completed this random baseline procedure on 5,000 permutations of the original turn transition analysis windows and then compared the beta estimates from the original data models (above) to the distribution of beta estimates for the models of the 5,000 randomly permuted datasets.
+We estimated whether each effect in our original statistical models differed from chance by calculating how many random beta estimates were smaller than the original beta estimate for each predictor, using the absolute value of all beta estimates for a two-tailed test.
+For example, children's original ``language condition'' beta estimate was $|$-3.29$|$, which is greater than 96.9\% of $|$beta estimates$|$ from the models of randomly-permuted data.
+This leads us to conclude that that the effect of language condition in the original model is unlikely to be the result of random gaze shifting.
+We excluded the output of random-permutation models that did not converge to remove unreliable \textit{$\beta$} estimates from our percentile calculations below (22.5\% and 24.4\% of models for children and adults, respectively).
+
+\begin{figure}[h]
+  \centering
+  \subfloat[Children]{\includegraphics[width=0.49\textwidth]{figures/E1-chi-randrun-dist.png}\label{fig:E1-ChiBetas}}
+  \hfill
+  \subfloat[Adults]{\includegraphics[width=0.49\textwidth]{figures/E1-adu-randrun-dist.png}\label{fig:E1-AduBetas}}
+  \caption{Random-permutation and original $|$beta estimates$|$ for predictors of children and adults' anticipatory gaze rates.
+Gray dots $=$ random model estimates, White dots $=$ original model estimates, Triangles $=$ 95th percentile for each beta estimate distribution.}
+\end{figure}
+
+Most of the significant predictors from models of the original, turn-related data can not be explained by random looking.
+The children's data showed strong evidence of differentiation between the random and original beta estimates for the effect of language condition ($>$96.9\% of random beta estimates) and the age-language condition interaction ($>$93.9\%), with somewhat weaker support for the language condition-transition type interaction ($>$85.1\%), and no support for the effect of gap duration ($>$73\%; Figure \ref{fig:E1-ChiBetas}).
+The adult data showed moderate evidence for the effect of transition type ($>$88.6\%), weak evidence for the interaction of transition type and duration ($>$80.6\%), and no evidence for the interaction of language condition and transition type ($>$59.8\%; Figure \ref{fig:E1-AduBetas}).
+
+\begin{figure}[h]
+\begin{center}
+\includegraphics[width=0.8\textwidth]{figures/E1-samples-by-lang-groups-trans-types.png}
+\end{center}
+\caption{Anticipatory gaze rates across language condition and transition type for the real (red and blue) and randomly permuted (gray) data.} 
+\label{fig:E1-randvsreal}
+\end{figure}
+
+% Statistical model comparing random to real data at each age.
 
@@
 
-Several results from Experiment 1 were unexpected.
-First, we found no significant main effect of age in the children's data, and we saw very little difference between adults and children (besides the age-related interactions).
+%Several results from Experiment 1 were unexpected.
+First, we found no significant effect of age in the children's data, and we saw very little difference between adults and children (besides the age-related interactions).
 It appears that, especially in English, children were as good at tracking the current speaker and anticipating upcoming turns as adults were.
@@
 
-Second, children, but not adults, showed an advantage for lexicosyntactic information in their anticipatory switches.
+%Second, children, but not adults, showed an advantage for lexicosyntactic information in their anticipatory switches.
 An advantage for lexicosyntactic information is consistent with prior work on adult turn taking \citep{de-ruiter2006}, so it's curious that the adults did not show a difference in anticipatory switching between videos they did and did not understand.
@@
 
-Finally, we had predicted that children would show early advantages for prosodic cues over syntactic ones because children begin developing language-specific knowledge about prosody long before lexicosyntax.
+%Finally, we had predicted that children would show early advantages for prosodic cues over syntactic ones because children begin developing language-specific knowledge about prosody long before lexicosyntax.
 We weren't able to directly test this prediction in Experiment 1 because we used a limited range of ages (three- to five-year-olds) and because we only controlled for lexicosyntactic information.
@@
 In sum, Experiment 1 lays the analytic groundwork for a method that allows for greater experimental control, which we introduce in Experiment 2.
+
+%% E1 DISCUSSION LEFTOVER TEXT
+
+%Language did not affect adults' anticipatory looking behavior, possibly because the animated, child-directed prosody and non-verbal cues made it possible for them to follow the interaction and track the current speaker easily.
+% As predicted, children made more anticipatory looks while they were watching videos that they could understand.
+In the English videos, children had access to lexicosyntactic information, which could have helped them make better predictions about upcoming turn structure.
+Although children performed better on the English videos, lexicosyntactic information was not exclusively responsible for their anticipatory switches; children still performed remarkably well when no lexical information was present at all, and even three-year-olds made anticipatory switches at an above-chance rate in the non-English videos.
+%Contrary to our expectation, there was no effect of age.
+Children's anticipatory switches \textit{did} increase with age for the non-English videos, but it also decreased with age for the English videos, resulting in a significant interaction of age and language condition.
+We had expected that, as children got older, they would also show more anticipatory switches.
+But this is contradicted by decreased looks with age for English videos, which even extended to the adult data.
+One possible explanation is that older children and adults found the videos in English easy to comprehend, and so tracked the conversation less closely.
+Compared to three-year-olds, adults and older children spent more time looking away from the current speaker and addressee in the English conversations.
+In contrast, compared to three-year-olds, they spent more time looking at the current speaker and less time looking away in the non-English conversations (Table \ref{tab:e1_look}).
+%It is also possible that the timing of anticipatory gaze switches changes as children get older.
+That is, the moment at which observers chose to launch their gaze away from the current speaker might change with age.
+For example, children could have reacted more immediately to early cues in the turn (e.g., ``Which one...'') as they got older.
+If a change in timing were sufficiently large, we might not capture it in our analysis window, which was designed to capture switches close to the turn transition.
+To test for a change in switch timing, we would have to run new analyses on a larger set of turn transitions---ones with controlled early and late cues \citep[see, e.g.,][]{bogels2015}.
+Within the current analysis windows there were some ``early switches'' (i.e., those launched before the end of the prior turn).
+But participants made approximately equal numbers of early switches, regardless of age and condition: 11--15\% (Figure \ref{fig:cumulativeFL}).
+The lack of increased early switching with age might suggest that changes in timing are either larger than the analysis window, or are not at play.
+%Children made more anticipatory switches after hearing questions than non-questions, but the advantage for questions was driven by age and linguistic access.
+There was no effect of transition type alone.
+Rather, the model confirmed a two-way interaction of linguistic condition and transition type.
+Presumably, question effects were larger in the English videos than the non-English videos because there were more cues to questionhood in English (i.e., non-verbal, prosodic, \textit{and} lexicosyntactic).
+Children younger than five did not distinguish questions from non-questions in the non-English videos (Figure \ref{fig:questionsFL}).
+The children's model confirmed a three-way interaction of age, linguistic condition, and transition type.
+Lexicosyntactic information might then be critical for young children to identify questions in conversation.
+Cues like subject-auxiliary inversion, and \textit{wh}-words occur early in the turn and are clear indicators that an answer will come next.
+Importantly, the turns marked as ``questions'' in the non-English data were not real questions, but turns that \textit{sounded like} questions to native English speakers.
+Question-like prosodic cues in the non-English videos could have been markedly different from question prosody in English, making it more difficult for three- and four-year-olds to identify ``questions'' in those conversations.
+%The increased switching after questions indicates that, so long as they have native lexical and prosodic cues, children as young as three spontaneously distinguish questions from other types of speech acts in third-party conversation.
+Children may treat questions differently from non-questions in conversation because they are predictably followed by an immediate response (i.e., the answer).
+Because the form of an answer is partially determined by its question (e.g., locative expressions for ``where'' questions), children could even be predicting what type of response they will hear, and look at the responder for confirmation.
+In contrast, non-questions do not usually require an immediate response, so it is not as easy to predict what will come next.
 
@@
 \textit{Audio Manipulation}.
-We created four versions of each audio clip: \textit{normal}, \textit{words only}, \textit{prosody only}, and \textit{no speech}.
+We created four versions of each audio clip: \textit{Normal}, \textit{words only}, \textit{prosody only}, and \textit{no speech}.
 That is, one version with a full linguistic signal (\textit{normal}), and three with incomplete linguistic information (hereafter ``limited cue'' conditions).
@@
 
-The \textit{words only} clips were manipulated to have robot-like speech: we flattened the intonation contours to each talker's average pitch (F0) and we reset the duration of every nucleus and coda to each talker's average nucleus and coda duration.\footnote{We excluded hyper-lengthened words like [w\textipa{aU:}] `woooow!'.
+The \textit{words only} clips were manipulated to have robot-like speech: We flattened the intonation contours to each talker's average pitch (F0) and we reset the duration of every nucleus and coda to each talker's average nucleus and coda duration.\footnote{We excluded hyper-lengthened words like [w\textipa{aU:}] `woooow!'.
 These were rare in the clips.} We made duration and pitch manipulations using PSOLA resynthesis in Praat \citep{Praat}.
@@
 The puppets were minimally expressive; the experimenter could only control the opening and closing of their mouths; their head, eyes, arms, and body stayed still.
-Puppets were positioned looking forward to eliminate shared gaze as a cue to turn structure \citep{thorgrimssonUndRev}.
+Puppets were positioned looking forward to eliminate shared gaze as a cue to turn structure \citep{thorgrimsson2015}.
 We took care to match the puppets' mouth movements to the syllable onsets as closely as possible, and avoided any mouth movement before the onset of a turn.
@@
 
-We used three pairs of puppets used for the \textit{normal} condition---`red', `blue' and `yellow'---and one pair of puppets for each limited cue condition: ``robots'', ``merpeople'', and ``party-goers'' (Figure 8).
+We used three pairs of puppets used for the \textit{normal} condition---`red', `blue' and `yellow'---and one pair of puppets for each limited cue condition: ``Robots'', ``merpeople'', and ``party-goers'' (Figure 8).
 We randomly assigned half of the conversation topics (`birthday cake', `pets', and `breakfast') to the \textit{normal} condition, and half to the limited cue conditions (`riding bikes', `rainy days', and `the library').
@@
 There were 79 turn transitions for analysis, after excluding transitions longer than 550 msec and shorter than 0 msec.
-The remaining turn transitions were distributed evenly across transition types (questions N$=$47 and non-questions N$=$32) and conditions, keeping in mind that there were three \textit{normal} videos and only one limited cue video for each experiment version: \textit{normal} (N$=$36), \textit{words only} (N$=$13), \textit{prosody only} (N$=$12), and {no speech} (N$=$18).
-Inter-turn gaps for questions (mean$=$358, median$=$405) were longer than those for non-questions (mean$=$296, median$=$288) on average, but gap duration was overall comparable across conditions: \textit{normal} (mean$=$333, median$=$337), \textit{words only} (mean$=$345, median$=$383), \textit{prosody only} (mean$=$320, median$=$336), and \textit{no words} (mean$=$324, median$=$328).
+The remaining turn transitions were distributed evenly across transition types (questions N$=$47 and non-questions N$=$32) and conditions, keeping in mind that there were three \textit{normal} videos and only one limited cue video for each experiment version: \textit{Normal} (N$=$36), \textit{words only} (N$=$13), \textit{prosody only} (N$=$12), and {no speech} (N$=$18).
+Inter-turn gaps for questions (mean$=$358, median$=$405) were longer than those for non-questions (mean$=$296, median$=$288) on average, but gap duration was overall comparable across conditions: \textit{Normal} (mean$=$333, median$=$337), \textit{words only} (mean$=$345, median$=$383), \textit{prosody only} (mean$=$320, median$=$336), and \textit{no words} (mean$=$324, median$=$328).
 The longer gaps for question transitions could give them an advantage because our anticipatory measure includes shifts initiated during the gap between turns (Figure \ref{fig:criterion}).
@@
 
-\subsubsection{Descriptive analysis}
-Participants' pattern of gaze indicated that they performed basic turn tracking across all ages and in all conditions.
+%\subsubsection{Descriptive analysis}
+%Participants' pattern of gaze indicated that they performed basic turn tracking across all ages and in all conditions.
 Participants looked at the screen most of the time during video playback (82\% and 86\% average for children and adults, respectively).
-Children and adults primarily kept their eyes on the person who was currently speaking: they gazed at the current speaker between 44\% and 69\% of the time, looking back at the addressee between 11\% and 14\% of the time (Table 2).
+Children and adults primarily kept their eyes on the person who was currently speaking: They gazed at the current speaker between 44\% and 69\% of the time, looking back at the addressee between 11\% and 14\% of the time (Table 2).
 They tracked the current speaker in every condition---even one-year-olds looked more at the current speaker than at anything else in the three limited cue conditions (40\% for \textit{words only}, 43\% for \textit{prosody only}, and 39\% for \textit{no speech}).
@@
 Looks to the addressee also decreased with age, but the change was minimal.
-
-\begin{table}[t]
-\begin{center}
-  \begin{tabular}{llcccc}
-    \hline
-    Age group & Speaker & Addressee & Other onscreen & Offscreen\\ 
-    \hline
-    1 & 0.44 & 0.14 & 0.23 & 0.19 \\ 
-    2 & 0.50 & 0.13 & 0.24 & 0.14 \\ 
-    3 & 0.47 & 0.12 & 0.25 & 0.16 \\ 
-    4 & 0.48 & 0.11 & 0.29 & 0.12 \\ 
-    5 & 0.54 & 0.11 & 0.20 & 0.14 \\ 
-    6 & 0.60 & 0.12 & 0.18 & 0.10 \\
-    Adult & 0.69 & 0.12 & 0.09 & 0.10 \\
-%   Overall & 0.53 & 0.12 & 0.21 & 0.13 \\
-    \hline
-  \end{tabular}
-\end{center}
-  \caption{Average proportion of gaze to the current speaker and addressee during periods of talk across ages.}
-\label{tab:look_e2}
-\end{table}
-
-\begin{table}[t]
-\begin{center}
-  \begin{tabular}{llcccc}
-    \hline
-    Condition & Speaker & Addressee & Other onscreen & Offscreen\\ 
-    \hline
-    Normal & 0.58 & 0.12 & 0.17 & 0.13 \\ 
-    Words only & 0.54 & 0.11 & 0.24 & 0.10 \\ 
-    Prosody only & 0.48 & 0.12 & 0.26 & 0.15 \\ 
-    No speech & 0.44 & 0.13 & 0.26 & 0.18 \\
-    \hline
-  \end{tabular}
-\end{center}
-  \caption{Average proportion of gaze to the current speaker and addressee during periods of talk across conditions.}
-\label{tab:look_e2b}
-\end{table}
-
-% Average prop gaze to the current speaker across conditions for one-year-olds
-%                           speaker addressee other-screen offscreen
-%Normal   		  0.4737582 0.1366432    0.2003952 0.1892035
-%Words only    0.3977261 0.1283417    0.3112005 0.1627317
-%Prosody only 0.4290202 0.1611453    0.2558657 0.1539688
-%No speech    0.3786925 0.1228615    0.2356423 0.2628037
-
-We identified anticipatory gaze switches for all 79 usable turn transitions using the same criteria as in Experiment 1.
+%
+%\begin{table}[t]
+%\begin{center}
+%  \begin{tabular}{llcccc}
+%    \hline
+%    Age group & Speaker & Addressee & Other onscreen & Offscreen\\ 
+%    \hline
+%    1 & 0.44 & 0.14 & 0.23 & 0.19 \\ 
+%    2 & 0.50 & 0.13 & 0.24 & 0.14 \\ 
+%    3 & 0.47 & 0.12 & 0.25 & 0.16 \\ 
+%    4 & 0.48 & 0.11 & 0.29 & 0.12 \\ 
+%    5 & 0.54 & 0.11 & 0.20 & 0.14 \\ 
+%    6 & 0.60 & 0.12 & 0.18 & 0.10 \\
+%    Adult & 0.69 & 0.12 & 0.09 & 0.10 \\
+%%   Overall & 0.53 & 0.12 & 0.21 & 0.13 \\
+%    \hline
+%  \end{tabular}
+%\end{center}
+%  \caption{Average proportion of gaze to the current speaker and addressee during periods of talk across ages.}
+%\label{tab:look_e2}
+%\end{table}
+%
+%\begin{table}[t]
+%\begin{center}
+%  \begin{tabular}{llcccc}
+%    \hline
+%    Condition & Speaker & Addressee & Other onscreen & Offscreen\\ 
+%    \hline
+%    Normal & 0.58 & 0.12 & 0.17 & 0.13 \\ 
+%    Words only & 0.54 & 0.11 & 0.24 & 0.10 \\ 
+%    Prosody only & 0.48 & 0.12 & 0.26 & 0.15 \\ 
+%    No speech & 0.44 & 0.13 & 0.26 & 0.18 \\
+%    \hline
+%  \end{tabular}
+%\end{center}
+%  \caption{Average proportion of gaze to the current speaker and addressee during periods of talk across conditions.}
+%\label{tab:look_e2b}
+%\end{table}
+%
+%% Average prop gaze to the current speaker across conditions for one-year-olds
+%%                           speaker addressee other-screen offscreen
+%%Normal   		  0.4737582 0.1366432    0.2003952 0.1892035
+%%Words only    0.3977261 0.1283417    0.3112005 0.1627317
+%%Prosody only 0.4290202 0.1611453    0.2558657 0.1539688
+%%No speech    0.3786925 0.1228615    0.2356423 0.2628037
+%
+%We identified anticipatory gaze switches for all 79 usable turn transitions using the same criteria as in Experiment 1.
 We expected to see the most anticipatory switches in the \textit{normal} condition and the least anticipatory switches in the \textit{no speech} condition because they contained the most and least linguistic information, respectively.
@@
 Finally, since the development of prosodic skills partially precedes the development of lexicosyntax, we expected to see more switches in the \textit{prosody only} condition compared to the \textit{words only} condition in the youngest age groups.
-
-We found that participants at all ages made anticipatory gaze switches more often than would be expected by chance in almost all conditions, but sometimes by a small margin (Figure \ref{fig:randvsrealEN}).
+%
+%We found that participants at all ages made anticipatory gaze switches more often than would be expected by chance in almost all conditions, but sometimes by a small margin (Figure \ref{fig:randvsrealEN}).
 For example, anticipatory switches in the \textit{no speech} videos were often modestly above chance (with the exception of 5- and 6-year-olds).
@@
 With sufficiently long inter-turn gaps, participants could simply switch their gaze to the responder when the prior talker finishes talking---a strategy that doesn't rely on linguistic information.
-But this interpretation is not compatible with our the effects of linguistic information that do emerge in the data: an early advantage for prosody and a later advantage for lexical question cues.
-
-We saw a generalize early advantage for prosody over lexicosyntax in children's spontaneous turn predictions.
+But this interpretation is not compatible with our the effects of linguistic information that do emerge in the data: An early advantage for prosody and a later advantage for lexical question cues.
+%
+%We saw a generalize early advantage for prosody over lexicosyntax in children's spontaneous turn predictions.
 Notably, while children performed well with \textit{prosody only} speech from the start, they seemed to only reliably exceed chance in the \textit{words only} condition at ages three and older.
 By age three, the general advantage for prosody over lexicosyntax seems to disappear with more equal rates of switching overall for the two partial linguistic conditions.
-
-\begin{figure}[t]
-\begin{center}
-\includegraphics[width=0.99\textwidth]{figures/FIG-randvsreal-EN.png}
-\end{center}
-\caption{Baseline-corrected proportion of anticipatory gaze switches made for actual (solid) and randomly-shuffled (dashed) turn transitions in each condition, and across ages.
+%
+%\begin{figure}[t]
+%\begin{center}
+%\includegraphics[width=0.99\textwidth]{figures/FIG-randvsreal-EN.png}
+%\end{center}
+%\caption{Baseline-corrected proportion of anticipatory gaze switches made for actual (solid) and randomly-shuffled (dashed) turn transitions in each condition, and across ages.
 Error bars represent the standard error of the mean.}
-\label{fig:randvsrealEN}
-\end{figure}
-
-We observed that, overall, children and adults made the fewest anticipatory switches in the \textit{no speech} videos (9\%) and the most in the \textit{normal} videos (16\%), with the \textit{prosody only} and \textit{words only} videos falling in-between (14\% and 11\%, respectively; Figure \ref{fig:conditionsEN}).
+%\label{fig:randvsrealEN}
+%\end{figure}
+%
+%We observed that, overall, children and adults made the fewest anticipatory switches in the \textit{no speech} videos (9\%) and the most in the \textit{normal} videos (16\%), with the \textit{prosody only} and \textit{words only} videos falling in-between (14\% and 11\%, respectively; Figure \ref{fig:conditionsEN}).
 However, these condition-based differences were primarily driven by transition type effects (question vs.
 non-question).
-
-\begin{figure}[t]
-\begin{center}
-\includegraphics[width=0.65\textwidth]{figures/FIG-conditions-EN.png}
-\end{center}
-\caption{Baseline-corrected proportion of anticipatory gaze switches made for turn transitions in each speech condition, and across ages. Error bars represent the standard error of the mean.} 
-\label{fig:conditionsEN}
-\end{figure}
-
-As in Experiment 1, we made condition, age and transition type comparisons based on baselin-corrected switching values; we corrected each participant's gaze data by subtracting their random baseline from their actual switching for each turn transition.
-
-We found that participants made more anticipatory gaze switches following questions than non-questions, but only in the two lexical conditions (\textit{normal} and \textit{words only} speech).
+%
+%\begin{figure}[t]
+%\begin{center}
+%\includegraphics[width=0.65\textwidth]{figures/FIG-conditions-EN.png}
+%\end{center}
+%\caption{Baseline-corrected proportion of anticipatory gaze switches made for turn transitions in each speech condition, and across ages. Error bars represent the standard error of the mean.} 
+%\label{fig:conditionsEN}
+%\end{figure}
+%
+%As in Experiment 1, we made condition, age and transition type comparisons based on baselin-corrected switching values; we corrected each participant's gaze data by subtracting their random baseline from their actual switching for each turn transition.
+%
+%We found that participants made more anticipatory gaze switches following questions than non-questions, but only in the two lexical conditions (\textit{normal} and \textit{words only} speech).
 The advantage for questions was strongest in the \textit{normal} speech condition, where it emerged early in development.
@@
 In the \textit{normal} condition, where the question effect was strongest, lexical and syntactic cues were reinforced by question prosody for polar questions.
-
-\begin{figure}[t]
-\begin{center}
-\includegraphics[width=0.99\textwidth]{figures/FIG-QvsNQ-EN.png}
-\end{center}
-\caption{Baseline-corrected proportion of anticipatory gaze switches made for question- (dashed) and non-question- (dotted) turn transitions in each condition, and across ages.
+%
+%\begin{figure}[t]
+%\begin{center}
+%\includegraphics[width=0.99\textwidth]{figures/FIG-QvsNQ-EN.png}
+%\end{center}
+%\caption{Baseline-corrected proportion of anticipatory gaze switches made for question- (dashed) and non-question- (dotted) turn transitions in each condition, and across ages.
 Error bars represent the standard error of the mean.}
-\label{fig:questionsEN}
-\end{figure}
-
-Children showed increased anticipatory switching with age across all conditions, making almost as many transitions as adults by ages five and six.
+%\label{fig:questionsEN}
+%\end{figure}
+%
+%Children showed increased anticipatory switching with age across all conditions, making almost as many transitions as adults by ages five and six.
 Adults reacted to the \textit{no speech} condition differently than the children did, tracking the interaction less closely than in the three linguistic conditions.
 They may have minimized switching in the \textit{no speech} condition because they realized that there was no conversation to follow.
-
-\subsubsection{Statistical analysis}
-
-We tested these findings in two linear mixed-effects models of participants' baseline-corrected anticipatory switches---one model each for children and adults.
+%
+%\subsubsection{Statistical analysis}
+%
+%We tested these findings in two linear mixed-effects models of participants' baseline-corrected anticipatory switches---one model each for children and adults.
 The child model included condition (\textit{normal}, \textit{words only}, \textit{prosody only}, and \textit{no speech}; \textit{normal} in the intercept), transition type (question vs.
@@
 The adult model matched the child model, excluding age-related effects.
-
-There was a significant effect of gap duration.
+%
+%There was a significant effect of gap duration.
 Children had time to make more anticipatory switches in analysis windows with longer gaps (\textit{$\beta$}=0.294, \textit{SE}=0.091, \textit{t}=3.231, \textit{p}$<$.01).
 An effect of gap duration is expected since prior work using spontaneous anticipatory switching has found that most switches occur in the inter-turn gap (Keitel et al., 2013; Hirvenkari, 2013; Tice and Henetz, 2011).
-
-The only effects of linguistic condition derived from interactions with age and transition type, despite large differences in the linguistic information available across conditions.
+%
+%The only effects of linguistic condition derived from interactions with age and transition type, despite large differences in the linguistic information available across conditions.
 (Figure \ref{fig:questionsEN}).
@@
 5\%).
-Since the question effect was limited to the lexical conditions, there was no main effect of transition type.
+Since the question effect was limited to the lexical conditions, there was no effect of transition type.
 There were also no further interactions of age and condition, or of transition type and condition.
 There was, however, a significant two-way interaction of transition type and age (\textit{$\beta$}=-0.031, \textit{SE}=0.01, \textit{t}=-3.18, \textit{p}$<$.01) in addition to the three-way interactions for the \textit{normal} and \textit{words only} conditions already mentioned.
-
-Last but not least, we confirmed that children generally made more anticipatory switches as they got older: The model showed a significant main effect of age (\textit{$\beta$}=0.028, \textit{SE}=0.009, \textit{t}=3.018, \textit{p}$<$.01).
+%
+%Last but not least, we confirmed that children generally made more anticipatory switches as they got older: The model showed a significant effect of age (\textit{$\beta$}=0.028, \textit{SE}=0.009, \textit{t}=3.018, \textit{p}$<$.01).
 Though the increase in anticipatory looks with age came, in large part, from questions effects in the lexical conditions, children also showed increased switches for non-questions and non-lexical conditions.
@@
 As in Experiment 1, early switches (planned before the end of the prior turn) did not show large increases with age, with adults making the fewest early switches in the three linguistic conditions (Figure \ref{fig:cumulativeEN}).
-
-\begin{figure}[ht]
-\begin{center}
-\includegraphics[width=0.99\textwidth]{figures/FIG-cumulative-EN.png}
-\end{center}
-\caption{Cumulative proportion of anticipatory gaze switches planned during the transition window, across conditions and ages.
+%
+%\begin{figure}[ht]
+%\begin{center}
+%\includegraphics[width=0.99\textwidth]{figures/FIG-cumulative-EN.png}
+%\end{center}
+%\caption{Cumulative proportion of anticipatory gaze switches planned during the transition window, across conditions and ages.
 Age is indicated by hue (darker = older).
@@
 ``Early switches'' occur before 0.}
-\label{fig:cumulativeEN}
-\end{figure}
-
-The advantage for questions over non-questions in these results reinforces our findings from Experiment 1.
+%\label{fig:cumulativeEN}
+%\end{figure}
+%
+%The advantage for questions over non-questions in these results reinforces our findings from Experiment 1.
 They suggest that even two-year-olds can reliably distinguish questions from other transition types in third-party conversation, especially when full linguistic information is available.
@@
 Meanwhile, 61\% of two-year-olds show the effect in the \textit{normal} condition, and 52\% of them do in the \textit{words only} condition, which is already close to the average for older children and adults (Table \ref{tab:questioneffectexp2}).
-
-Our results suggest that the question effect emerges between ages one and two, but further testing is needed to establish whether this is the case.
+%
+%Our results suggest that the question effect emerges between ages one and two, but further testing is needed to establish whether this is the case.
 For example, the size and emergence of the question effect for various question types and question cues could differ (\textit{wh-} vs.
@@
 Future work with more fine-grained age samples and better controlled \textit{normal} speech might then find that children can identify questions in conversation even earlier.
-
-\begin{table}[t]
-\begin{center}
-  \begin{tabular}{ccc}
-    \hline
-    Age group & Normal speech & No speech \\ 
-    \hline
-    1 & 0.43 & 0.24 \\
-    2 & 0.61 & 0.52 \\
-    3 & 0.63 & 0.42 \\
-    4 & 0.78 & 0.52 \\
-    5 & 0.91 & 0.52 \\
-    6 & 0.75 & 0.50 \\
-    Adult & 0.67 & 0.57 \\
-    \hline
-  \end{tabular}
-\end{center}
-  \caption{Proportion of participants showing numerically more anticipatory switches for questions than non-questions in the \textit{normal} and \textit{words only} conditions.}
-\label{tab:questioneffectexp2}
-\end{table}
-
-Finally, we confirmed that young children showed a benefit for prosody over lexicosyntax.
+%
+%\begin{table}[t]
+%\begin{center}
+%  \begin{tabular}{ccc}
+%    \hline
+%    Age group & Normal speech & No speech \\ 
+%    \hline
+%    1 & 0.43 & 0.24 \\
+%    2 & 0.61 & 0.52 \\
+%    3 & 0.63 & 0.42 \\
+%    4 & 0.78 & 0.52 \\
+%    5 & 0.91 & 0.52 \\
+%    6 & 0.75 & 0.50 \\
+%    Adult & 0.67 & 0.57 \\
+%    \hline
+%  \end{tabular}
+%\end{center}
+%  \caption{Proportion of participants showing numerically more anticipatory switches for questions than non-questions in the \textit{normal} and \textit{words only} conditions.}
+%\label{tab:questioneffectexp2}
+%\end{table}
+%
+%Finally, we confirmed that young children showed a benefit for prosody over lexicosyntax.
 We conducted two \textit{post-hoc} \textit{t}-tests for children under three, comparing their anticipatory switching rate against their estimated random switching rate in the \textit{prosody only} and \textit{words only} conditions.
@@
 This pattern of results may suggest a qualitative shift in the kind of predictions children make while observing a conversation, with linguistic cues (like prosody) playing different roles at different ages.
-
-The model of adults' anticipatory gaze switches confirmed main effects of gap duration and linguistic condition.
+%
+%The model of adults' anticipatory gaze switches confirmed effects of gap duration and linguistic condition.
 Longer gaps led to more anticipatory switches (\textit{$\beta$}=0.656, \textit{SE}=0.176, \textit{t}=3.726, \textit{p}$<$.001).
 Adults also made significantly more anticipatory switches in the \textit{normal} condition than in the \textit{no speech} baseline, indicating that linguistic access did affect their overall performance (\textit{$\beta$}=-0.211, \textit{SE}=0.085, \textit{t}=-2.486, \textit{p}$<$.05).
-There were no other main effects of linguistic condition.
+There were no other effects of linguistic condition.
 The model also showed a marginal two-way interaction of condition and transition type, with adults making more anticipatory gazes for questions in the \textit{normal} condition (\textit{$\beta$}=0.194, \textit{SE}=0.11, \textit{t}=1.773, \textit{p}=.076).
@@
 This result suggests that prosodic cues still have a role to play in adult question recognition during third-party conversation.
-Since question effects were primarily limited to the \textit{normal} condition, there was no main effect of transition type in the adults' data.
+Since question effects were primarily limited to the \textit{normal} condition, there was no effect of transition type in the adults' data.
 
@@
 
-The core aims of Experiment 2 were to gain better traction on the individual roles of prosody and lexicosyntax in children's turn predictions, and to expand our age range to capture more developmental change.
+%The core aims of Experiment 2 were to gain better traction on the individual roles of prosody and lexicosyntax in children's turn predictions, and to expand our age range to capture more developmental change.
 We found that effects of linguistic processing and age were present in this dataset, but both were primarily found within the question effect.
@@
 Even so, performance was consistently best in the \textit{normal} condition, in which participants could integrate lexical and prosodic information together.
-
-By extending the age range of participants, we were able to see more of the developmental trajectory in children's predictions about turn structure.
+%
+%By extending the age range of participants, we were able to see more of the developmental trajectory in children's predictions about turn structure.
 We saw that one-year-olds already made more anticipatory switches than would be expected by chance in two of the four conditions.
@@
 It will be up to future work to disentangle which lexical and prosodic cues children begin to pick out in distinguishing questions from other speech acts, and when.
-
-The results from Experiment 2 complement those of Experiment 1, but still leave some open questions.
+%
+%The results from Experiment 2 complement those of Experiment 1, but still leave some open questions.
 For example, we were surprised to find no independent effects of linguistic condition outside of age and transition type interactions.
@@
 We take this issue up further in the general discussion.
-
-We found that the importance of lexicosyntactic and prosodic information for turn prediction differed depending on age and transition type.
+%
+%We found that the importance of lexicosyntactic and prosodic information for turn prediction differed depending on age and transition type.
 While even one-year-olds made anticipatory switches at an above-chance rate in the \textit{normal} and \textit{prosody only} conditions, children didn't consistently surpass chance in the \textit{words only} condition until age three.
@@
 In short, prosody has different roles to play in turn prediction before and after the lexically-driven question effect emerges.
-Importantly, these results do not support a straightforward view of lexicosyntax as the primary linguistic cue for turn predictions \citep[even for adults; ][]{de-ruiter2006, magyari2012}, but do support accounts of multiple cue integration \citep{duncan1972, ford1996, hirvenkari2013, torreiraUndRev}.
-
-We controlled linguistic information in Experiment 2 by using phonetically-manipulated speech, extending speech manipulations used in prior work \citep{de-ruiter2006, keitel2013}.
+Importantly, these results do not support a straightforward view of lexicosyntax as the primary linguistic cue for turn predictions \citep[even for adults; ][]{de-ruiter2006, magyari2012}, but do support accounts of multiple cue integration \citep{duncan1972, ford1996, hirvenkari2013, torreira2015}.
+%
+%We controlled linguistic information in Experiment 2 by using phonetically-manipulated speech, extending speech manipulations used in prior work \citep{de-ruiter2006, keitel2013}.
 While phonetic manipulation gave us control over the linguistic signal, it also created speech sounds that children don't usually hear in their natural environment.
@@
 Children could have had trouble processing the \textit{words only} and \textit{prosody only} conditions because they were unfamiliar, and not just because they had less linguistic information available.
-
-It would take a different type of stimulus to compare lexical and prosodic information without phonetic manipulation.
+%
+%It would take a different type of stimulus to compare lexical and prosodic information without phonetic manipulation.
 For example, carefully scripted or cross-spliced conversations could control for the presence or absence of prosodic and lexicosyntactic cues to turn transition with natural speech.
@@
 Perhaps the advantage of the \textit{normal} condition was then not entirely due to its naturalness.
-We expect that future work with spontaneous conversational measures will still find an advantage for combined lexical and prosodic cues over lexical cues alone \citep[see also][]{duncan1972, ford1996, torreiraUndRev}.
+We expect that future work with spontaneous conversational measures will still find an advantage for combined lexical and prosodic cues over lexical cues alone \citep[see also][]{duncan1972, ford1996, torreira2015}.
 
@@
 \label{sec:gendisc}
-Children begin to develop conversational turn-taking skills long before their first words \citep{bateson1975, hilbrinkInPrep, jaffe2001, snow1977}.
+%Children begin to develop conversational turn-taking skills long before their first words \citep{bateson1975, hilbrink2015, jaffe2001, snow1977}.
 As they acquire language, they also acquire the information needed to make accurate predictions about upcoming turn structure.
 Until recently, we have had very little data on how children weave language into their already-existing turn-taking behaviors.
-
-In two experiments investigating children's anticipatory gaze to upcoming speakers, we found evidence that turn prediction develops early in childhood, and that linguistic cues are integrated differently at different ages.
+%
+%In two experiments investigating children's anticipatory gaze to upcoming speakers, we found evidence that turn prediction develops early in childhood, and that linguistic cues are integrated differently at different ages.
 Children under three showed a general advantage for prosody over lexicosyntax, contrary to findings for adults in prior work \citep{de-ruiter2006}.
-Children two and older also showed a slight advantage for lexicosyntax, but only in the form of a transition type effect: they made more anticipatory switches following questions than non-questions in conditions where lexical information was present.
+Children two and older also showed a slight advantage for lexicosyntax, but only in the form of a transition type effect: They made more anticipatory switches following questions than non-questions in conditions where lexical information was present.
 Two-year-olds thus appeared to be at a developmental transition point, still showing an overall advantage for prosody, but beginning to show an advantage for lexicosyntax within the domain of question-answer sequences.
@@
 531]{de-ruiter2006}: Participants' performance was best in conditions when they had access to the full linguistic signal.
-
-\subsection{Developmental implications}
-
-The primary aim of our study was to find out how linguistic information affects children's predictions about upcoming turn structure across development.
+%
+%\subsection{Developmental implications}
+%
+%The primary aim of our study was to find out how linguistic information affects children's predictions about upcoming turn structure across development.
 Two primary developmental findings emerged from our data.
@@
 Second, children's use of linguistic signals in making predictions depended on their age, with an early advantage for prosody over lexicosyntax and a later (partial) advantage for lexicosyntax over prosody.
-
-\subsubsection{The advantage for questions}
-
-In both of our experiments we found a robust advantage for questions in participants' anticipatory switching because participants made more switches after questions than non-questions.
+%
+%\subsubsection{The advantage for questions}
+%
+%In both of our experiments we found a robust advantage for questions in participants' anticipatory switching because participants made more switches after questions than non-questions.
 Children and adults both treated questions as if they had a special interactive status, and with good reason: Questions make a speaker switch immediately relevant, helping the listener to predict what will happen next (i.e., an answer from the addressee).
@@
 Also, the form of a question can constrain the type of response that will come next (e.g., a location after a \textit{where} question), further increasing the predictability of the interactional sequence.
-
-Children get frequent and early experience with hearing and answering questions.
+%
+%Children get frequent and early experience with hearing and answering questions.
 Questions make up approximately one third of the utterances children hear, before and after the onset of speech, and even into their preschool years \citep{fitneva2012, henning2005, shatz1979}.\footnote{There is substantial variation in this proportion by individual and socioeconomic class \citep{hart1992}.} Many of the questions directed to young children are ``test' questions---questions that the caregiver already has the answer to (e.g., ``What does a cat say?'').
@@
 So, in addition to having a special interactive status, questions are a core characteristic of many caregiver-child interactions.
-
-We cannot definitively say with our current data what property of questions caused children to switch their gaze more often.
+%
+%We cannot definitively say with our current data what property of questions caused children to switch their gaze more often.
 It is true that questions make a speaker transition more likely and that questions have distinctive linguistic cues, both early and late, to identify themselves.
@@
 So, though it is clear that children anticipate responses more often for questions than non-questions, their actual understanding of questionhood as a linguistically- or interactionally-marked type of speech act still needs further exploration.
-
-\subsubsection{Linguistic cues for turn prediction}
-
-Prior work with adults has found a consistent and critical role for lexicosyntax in predicting upcoming turn structure \citep{de-ruiter2006, magyari2012}, with more debate over the role of prosody \citep{duncan1972, ford1996, torreiraUndRev}.
+%
+%\subsubsection{Linguistic cues for turn prediction}
+%
+%Prior work with adults has found a consistent and critical role for lexicosyntax in predicting upcoming turn structure \citep{de-ruiter2006, magyari2012}, with more debate over the role of prosody \citep{duncan1972, ford1996, torreira2015}.
 Knowing that children comprehend more about prosody than lexicosyntax early on (see introduction; also see \citealp{speer2009} for a review), we thought it possible that young children would instead show an advantage for prosody over lexicosyntax.
 Our results suggest that there are roles for both prosody and lexicosyntax in children's predictions about upcoming turn structure, but that there are different roles for these information sources at different ages.
-
-Early on, there was an advantage for prosody over lexicosyntax in children's anticipatory switches: Anticipatory switching didn't reliably exceed chance for lexicosyntactic information  until age three, but was already significantly greater than chance for prosodic information  at age one.
+%
+%Early on, there was an advantage for prosody over lexicosyntax in children's anticipatory switches: Anticipatory switching didn't reliably exceed chance for lexicosyntactic information  until age three, but was already significantly greater than chance for prosodic information  at age one.
 This early advantage for prosody was not driven by the advantage for questions: While some one-year-olds in our study already made more switches after questions than non-questions, most did not, plus prosody alone did not elicit a robust question effect at any age (Figure \ref{fig:questionsEN}).
-
-Although children showed an initial advantage for prosody, their predictions before age three were probably limited to basic prosodic contours and boundaries, and simple cues such as final lengthening and turn-final silence \citep{pannekamp2006}.
+%
+%Although children showed an initial advantage for prosody, their predictions before age three were probably limited to basic prosodic contours and boundaries, and simple cues such as final lengthening and turn-final silence \citep{pannekamp2006}.
 More complex structural predictions would have required them to access lexicosyntactic information because higher-level structure in the prosodic signal is linked to the syntactic structure being used.
-
-In both experiments, even young children showed an advantage for questions in conditions where they had lexicosyntax, suggesting that lexicosyntax is a critical cue for question identification.
+%
+%In both experiments, even young children showed an advantage for questions in conditions where they had lexicosyntax, suggesting that lexicosyntax is a critical cue for question identification.
 The question effect emerged by age two, when children's syntactic comprehension and lexical inventories are still relatively limited \citep{clark2009}.
@@
 Thus, the unambiguous presence of lexicosyntactic cues on \textit{wh}- and \textit{yes/no} questions may have been one reason why we saw an advantage for questions in the lexical conditions, but not in the \textit{prosody only} condition.
-
-Importantly, the question effect was not driven purely by lexicosyntactic information.
+%
+%Importantly, the question effect was not driven purely by lexicosyntactic information.
 In both experiments, prosody still made a contribution to the advantage for questions, even though prosody alone was not enough to elicit a question effect.
@@
 But, prosodic cues still aid in picking out questions from ongoing conversation.
-
-The close link between prosodic and syntactic structure makes it difficult to tease apart how predictive processing in one domain (e.g., prosody) is distinct from the other (e.g., syntax) in turn prediction.
+%
+%The close link between prosodic and syntactic structure makes it difficult to tease apart how predictive processing in one domain (e.g., prosody) is distinct from the other (e.g., syntax) in turn prediction.
 We leave this challenge to future work.
-
-\subsection{Contributions of the current work}
-
-Our findings are consistent with observational work on children's turn taking.
-Young children are slow turn-takers in spontaneous conversation \citep{casillasUndRev} even though they begin taking turns in infancy \citep{hilbrinkInPrep, jaffe2001}.
+%
+%\subsection{Contributions of the current work}
+%
+%Our findings are consistent with observational work on children's turn taking.
+Young children are slow turn-takers in spontaneous conversation \citep{casillas2016} even though they begin taking turns in infancy \citep{hilbrink2015, jaffe2001}.
 We found that children begin to spontaneously predict upcoming turn structure at age one.
@@
 Our results therefore support the idea of an early-developed competence for the timing of turn taking.
-But because children do not themselves use adult-like conversational timing until much later \citep[age six; ][]{ervin-tripp1979}, our results also indirectly point toward response planning as a key factor contributing to children's slow responses in real-time interaction \citep{casillasUndRev}.
-
-Our findings are only partially consistent with prior experimental work on turn taking.
+But because children do not themselves use adult-like conversational timing until much later \citep[age six; ][]{ervin-tripp1979}, our results also indirectly point toward response planning as a key factor contributing to children's slow responses in real-time interaction \citep{casillas2016}.
+%
+%Our findings are only partially consistent with prior experimental work on turn taking.
 De Ruiter and colleagues (2006) found that adults' accuracy in identifying upcoming turn-ends was not significantly affected when intonation contours were flattened.
@@
 This may partially be due to the fact that we controlled syllable duration in addition to intonation, thereby eliminating two of the most salient cues to prosodic structure.
-
-Despite extreme differences in linguistic information across conditions, independent effects of linguistic information (outside of interactions) were minimal in our experiments.
+%
+%Despite extreme differences in linguistic information across conditions, independent effects of linguistic information (outside of interactions) were minimal in our experiments.
 But, using a button-press measure with similar stimuli, de Ruiter and colleagues (2006) found robust effects of linguistic condition.
@@
 We have seen in the current study that questions are interactionally sufficient to motivate early predictions.
-
-Moving away from button press measures, Keitel and colleagues (2013) used observer gaze to investigate children's turn predictions, finding: (a) no evidence that children under three make switches above chance, (b) that three-year-olds' switches are significantly affected by flattened intonation, and (c) that adults' switches are not.
+%
+%Moving away from button press measures, Keitel and colleagues (2013) used observer gaze to investigate children's turn predictions, finding: (a) No evidence that children under three make switches above chance, (b) that three-year-olds' switches are significantly affected by flattened intonation, and (c) that adults' switches are not.
 Our current results are consistent with the general finding from their study, that prosody matters for young children's turn predictions.
 But our results also differ from theirs in a few important ways.
-
-First, we found that children begin making above-chance anticipatory switches much earlier than three---our one-year-olds were already looking above chance in two of our four linguistic conditions in Experiment 2.
+%
+%First, we found that children begin making above-chance anticipatory switches much earlier than three---our one-year-olds were already looking above chance in two of our four linguistic conditions in Experiment 2.
 This earlier emergence of anticipatory switching could be due to our more child-friendly stimuli, which we designed to be engaging and to mimic the kind of speech children often experience interactionally.
@@
 Ultimately, the added conditions allowed us to find an early advantage for prosody and a later advantage for lexicosyntax.
-
-Third, we found that anticipatory looking was primarily driven by question transitions, a pattern that has so far not been discussed in other observer gaze studies, on children or adults (Keitel et al., 2013; Hirvenkari, 2013; Tice and Henetz, \citeyear{TiceHenetz11}).
+%
+%Third, we found that anticipatory looking was primarily driven by question transitions, a pattern that has so far not been discussed in other observer gaze studies, on children or adults (Keitel et al., 2013; Hirvenkari, 2013; Tice and Henetz, \citeyear{TiceHenetz11}).
 The question effects in the current study were critical to understanding the role of lexicosyntactic information in participants' turn predictions.
@@
 But again, because we flattened pitch for individual speakers and controlled syllable duration, we are not able to directly compare our findings to theirs.
-
-\subsection{Limitations}
-
-In Experiment 2, we tried to improve our method for controlling linguistic information by phonetically manipulating improvised dialogic speech.
+%
+%\subsection{Limitations}
+%
+%In Experiment 2, we tried to improve our method for controlling linguistic information by phonetically manipulating improvised dialogic speech.
 Although this decision allowed us to follow prior work more closely \citep{de-ruiter2006, keitel2013}, it also limits our interpretation of children's behavior.
@@
 One alternative to explore for the future would be scripted speech, cross-spliced speech, or speech that is carefully selected from a large corpus, given a set of specific linguistic constraints.
-
-Relatedly, since we used improvised conversation recordings, we did not control how often the speakers switched back and forth, or how often a switch occurred after a pause in speech.
-It is possible that, to some extent, children could have formed expectations about the turn structure on the basis of the frequent switching alone \citep[see, e.g., ][]{thorgrimssonUndRev}.
+%
+%Relatedly, since we used improvised conversation recordings, we did not control how often the speakers switched back and forth, or how often a switch occurred after a pause in speech.
+It is possible that, to some extent, children could have formed expectations about the turn structure on the basis of the frequent switching alone \citep[see, e.g., ][]{thorgrimsson2015}.
 If this statistical expectation were the only thing leading children's anticipatory gaze switches, we would have predicted more uniformity in looking behavior across the conditions and no major effects of linguistic signal, age, or transition type.
@@
 Nevertheless, another advantage of using scripted conversation would be the ability to control switch frequency and eliminate this issue.
-
-Finally, we carefully designed a scheme for identifying anticipatory turn switches in the data, building off of Keitel and colleagues' (2013) analyses.
+%
+%Finally, we carefully designed a scheme for identifying anticipatory turn switches in the data, building off of Keitel and colleagues' (2013) analyses.
 In both our study and theirs, anticipatory switches that happen very early in the turn are not counted.
@@
 But if early looks do indeed increase as children get older and have more access to early syntactic cues, it would be helpful to develop a second measure of very early anticipatory looks.
-
-\subsection{Conclusions}
-
-Conversation plays a central role in children's language learning.
+%
+%\subsection{Conclusions}
+%
+%Conversation plays a central role in children's language learning.
 It is the driving force behind what children say and what they hear.

### 2015-11-28 paper/CasillasFrank-Anticip.tex
Small changes to randrun analysis/plots; paper updates in progress

@@
 We chose male-female pairs as a natural way of creating contrast between the two talker voices.
-See an example run-though of the videos here: www.youtube.com/channel/UCGEZQcM9t8Zfjqqi\_B1Q5Sw.}
+See an example run-though of the videos here: }
+
+%% PART UNFINISHED: link for example video
 
@@
 
-\subsection{Results and discussion}
+\subsection{Results}
 \label{sec:results1}
@@
 \end{table}
-%
-%\begin{figure}[t]
-%\begin{center}
-%\includegraphics[width=0.8\textwidth]{figures/FIG-randvsreal-FL.png}
-%\end{center}
-%\caption{Proportion of anticipatory gaze switches made for actual (solid) and randomly-shuffled (dashed) turn transitions in each condition, and across ages.
-Error bars represent the standard error of the mean.}
-%\label{fig:randvsrealFL}
-%\end{figure}
-%
-%To estimate the effects of linguistic condition and age on switching behavior, we estimated each participants' random baseline switching behavior for each turn transition.
-Participants at all ages made anticipatory gaze switches more often than would be expected by chance, whether they could understand the language in the video or not (Figure \ref{fig:randvsrealFL}).
-Anticipatory switch behavior in the non-English videos could not have been based on lexicosyntactic information, so both adults and children must have relied on other sources of information to make their predictions (e.g., non-native prosodic cues and non-verbal signals).
-For all further analyses we then corrected each participant's data by subtracting their random baseline estimations from their actual switching data for each turn transition.
-%
-%Overall, a greater proportion of participants made anticipatory switches in the English videos (0.17) compared to the non-English videos (0.09).
-But the tendency to make more anticipatory switches for English than non-English videos disappeared with age (Figure \ref{fig:conditionsFL}).
-While anticipatory switches increased with age for non-English videos, they decreased with age for English videos, resulting in nearly equal looking behavior for the two language conditions for adults.
-Otherwise, age-based differences in predictive gaze shifts were minor.
-%
-%\begin{figure}[t]
-%\begin{center}
-%\includegraphics[width=0.5\textwidth]{figures/FIG-conditions-FL.png}
-%\end{center}
-%\caption{Baseline-corrected proportion of anticipatory gaze switches made for turn transitions in each language condition, and across ages. Error bars represent the standard error of the mean.} 
-%\label{fig:conditionsFL}
-%\end{figure}
-%
-%\begin{figure}[t]
-%\begin{center}
-%\includegraphics[width=0.8\textwidth]{figures/FIG-QvsNQ-FL.png}
-%\end{center}
-%\caption{Baseline-corrected proportion of anticipatory gaze switches made for question- (dashed) and non-question- (dotted) turn transitions in each condition, and across ages.
-Error bars represent the standard error of the mean.}
-%\label{fig:questionsFL}
-%\end{figure}
-%
-%Participants also showed an advantage for questions (Figure \ref{fig:questionsFL}).
-Children made anticipatory switches twice as often when they heard a question, compared to a non-question (18.7\% and 7.3\%).
-But the advantage for questions was not universal---it depended on language condition and age.
-The advantage for questions didn't emerge until age five in the non-English videos.
-%
+
+\begin{table}[h!]
+\begin{center}
+  \begin{tabular}{lcccl}
+  \textbf{\textit{Children}} &&&& \\
+    \hline
+                   & Estimate & Std. Error & z value & Pr($>$$|$z$|$)    \\
+    \hline
+    (Intercept)   &      -0.96146 &   0.84901 & -1.132 & 0.257446 \\    
+    Age             &    -0.18268 &   0.17507 & -1.043 & 0.296725   \\  
+    LgCond$=$\textit{non-English} &          -3.29347 &   0.96045 & -3.429 & 0.000606 *** \\
+    Type$=$\textit{non-Question}          &     -1.10129 &   0.86494 & -1.273 & 0.202925     \\
+    Duration      &       3.40169 &   1.22826 &  2.770 & 0.005614  **\\
+    Age*LgCond$=$\textit{non-English} &       0.52065 &   0.21190 &  2.457 & 0.014008 ** \\  
+    Age*TypeS$=$\textit{non-Question}          & -0.01628 &   0.19437 & -0.084 & 0.933232     \\
+    LgCond$=$\textit{non-English}* &     2.68166 &   1.35016 &  1.986 & 0.047013 * \\
+    Type$=$\textit{non-Question} &&&& \\
+    Age*LgCond$=$\textit{non-English}* & -0.45632 &   0.30163 & -1.513 & 0.130315  \\
+    Type$=$\textit{non-Question} &&&& \\
+    \hline
+  &&&& \\
+  \textbf{\textit{Adults}} &&&& \\
+    \hline
+                       &  Estimate & Std. Error & z value & Pr($>$$|$z$|$) \\    
+    \hline
+    (Intercept)       &        -0.1966 &    0.6942 & -0.283 & 0.776988     \\
+    LgCond$=$\textit{non-English}    &             -0.8812 &    0.9602 & -0.918 & 0.358754   \\  
+    Type$=$\textit{non-Question}             &        -4.4953 &    1.3139 & -3.421 & 0.000623 *** \\
+    Duration          &        -1.1227 &    1.9880 & -0.565 & 0.572238     \\
+    LgCond$=$\textit{non-English}* &           3.2972 &    1.6101 &  2.048 & 0.040581 * \\ 
+    Type$=$\textit{non-Question} &&&& \\
+    LgCond$=$\textit{non-English}* &        1.3626 &    3.0077 &  0.453 & 0.650527     \\
+    Duration &&&& \\
+    Type$=$\textit{non-Question}*          &  10.5107 &    3.3459 &  3.141 & 0.001682  ** \\
+    Duration &&&& \\
+    LgCond$=$\textit{non-English}* & -6.3156 &    4.4926 & -1.406 & 0.159790 \\    
+    Type$=$\textit{non-Question}* &&&& \\
+    Duration &&&& \\
+    \hline
+  \end{tabular}
+\end{center}
+  \caption{Model output for children and adults' anticipatory gaze switches.}
+\label{tab:E1-models}
+\end{table}
 
@@
 
-\begin{table}[t]
-\begin{center}
-  \begin{tabular}{lcccc}
-  \textbf{\textit{Children}} &&&& \\
-    \hline
-                   & Estimate & Std. Error & z value & Pr($>$$|$z$|$)    \\
-    \hline
-(Intercept)   &      -0.96146 &   0.84901 & -1.132 & 0.257446 \\    
-Age             &    -0.18268 &   0.17507 & -1.043 & 0.296725   \\  
-LgGroupNE &          -3.29347 &   0.96045 & -3.429 & 0.000606 \\
-TypeS          &     -1.10129 &   0.86494 & -1.273 & 0.202925     \\
-Duration      &       3.40169 &   1.22826 &  2.770 & 0.005614  \\
-Age:LgGroupNE &       0.52065 &   0.21190 &  2.457 & 0.014008 \\  
-Age:TypeS          & -0.01628 &   0.19437 & -0.084 & 0.933232     \\
-LgGroupNE:TypeS &     2.68166 &   1.35016 &  1.986 & 0.047013 \\  
-Age:LgGroupNE:TypeS & -0.45632 &   0.30163 & -1.513 & 0.130315  \\   
-    \hline
-  &&&& \\
-  \textbf{\textit{Adults}} &&&& \\
-    \hline
-                       &  Estimate & Std. Error & z value & Pr($>$$|$z$|$) \\    
-    \hline
-(Intercept)       &        -0.1966 &    0.6942 & -0.283 & 0.776988     \\
-LgGroupNE    &             -0.8812 &    0.9602 & -0.918 & 0.358754   \\  
-TypeS             &        -4.4953 &    1.3139 & -3.421 & 0.000623  \\
-Duration          &        -1.1227 &    1.9880 & -0.565 & 0.572238     \\
-LgGroupNE:TypeS &           3.2972 &    1.6101 &  2.048 & 0.040581  \\  
-LgGroupNE:Duration &        1.3626 &    3.0077 &  0.453 & 0.650527     \\
-TypeS:Duration          &  10.5107 &    3.3459 &  3.141 & 0.001682  \\
-LgGroupNE:TypeS:Duration & -6.3156 &    4.4926 & -1.406 0.159790 \\    
-    \hline
-  \end{tabular}
-\end{center}
-  \caption{Model output for children and adults' anticipatory gaze switches.}
-\label{tab:E1-models}
-\end{table}
-
-Children's anticipatory gaze switches showed effects of language condition (\textit{$\beta$}=-3.29, \textit{SE}=0.961, \textit{t}=-3.43, \textit{p}$<$.001) and gap duration (\textit{$\beta$}=3.4, \textit{SE}=1.229, \textit{t}=2.77, \textit{p}$<$.01) with additional effects of a language condition-transition type interaction (\textit{$\beta$}=2.68, \textit{SE}=1.35, \textit{t}=1.99, \textit{p}$<$.05) and an age-language condition interaction (\textit{$\beta$}=0.52, \textit{SE}=0.212, \textit{t}=2.46, \textit{p}$<$.05).
+Children's anticipatory gaze switches showed effects of language condition (\textit{$\beta$}=-3.29, \textit{SE}=0.961, \textit{t}=-3.43, \textit{p}$<$.001) and gap duration (\textit{$\beta$}=3.4, \textit{SE}=1.229, \textit{t}=2.77, \textit{p}$<$.01) with additional effects of an age-by-language condition interaction (\textit{$\beta$}=0.52, \textit{SE}=0.212, \textit{t}=2.46, \textit{p}$<$.05) and a language condition-by-transition type interaction (\textit{$\beta$}=2.68, \textit{SE}=1.35, \textit{t}=1.99, \textit{p}$<$.05).
 There were no significant effects of age or transition type alone (\textit{$\beta$}=-0.18, \textit{SE}=0.175, \textit{t}=-1.04, \textit{p}$=$.3 and \textit{$\beta$}=-1.10, \textit{SE}=0.865, \textit{t}=-1.27, \textit{p}$=$.2, respectively).
@@
 
-\begin{figure}[h]
+\subsubsection{Random baseline comparison}
+\label{sec:randbaseline1}
+
+We estimated the probability that these patterns were the result of random looking by running the same regression models on participants' real eye-tracking data, only this time calculating their anticipatory gaze switches with respect to randomly permuted turn transition windows.
+This process involved: (1) randomizing the order and temporal placement of the analysis windows within each stimulus (Figure \ref{fig:shuffling}; the analysis window is defined in Figure \ref{fig:criterion}), thereby randomly redistributing the analysis windows across the eye-tracking signal, (2) re-running each participant's eye tracking data through switch identification (described in \ref{sec:algorithm}), but this time using the randomly permuted analysis windows, and (3) modeling the anticipatory gazes from the randomly permuted data with the same statistical models we used for the original data (Section \ref{sec:models1}; Table \ref{tab:E1-models}).
+Importantly, although the onset time of each transition was shuffled within the eye-tracking signal, the other intrinsic properties of each turn transition (e.g., prior speaker identity, transition type, gap duration, language condition, etc.) stayed constant across each randomly permuted version of the data.
+
+\begin{figure}[ht!]
 \begin{center}
@@
 
-\subsubsection{Random baseline comparison}
-
-We estimated the probability that these patterns were the result of random looking by running the same regression models on participants' real eye-tracking data, only this time calculating their anticipatory gaze switches with respect to randomly permuted turn transitions (Figure \ref{fig:shuffling}).
-This involved: (1) randomizing the order and temporal placement of the analysis windows (Figure \ref{fig:criterion}) for each stimulus, thereby randomly redistributing the windows of analysis across the eye-tracking signal for each stimulus, then (2) re-running each participant's eye tracking data through switch identification (described in \ref{sec:algorithm}) with the randomly permuted analysis windows, and finally (3) entering the resulting anticipatory gazes into the same statistical model we used with the original data (above).
-Properties of each turn transition (e.g., prior speaker identity, transition type, duration, etc.) stayed constant, even though its onset and offset time within the eye-tracking signal was shuffled.
-
-In effect, this procedure de-links participants' gaze data from the turn structure in the original stimulus, thereby allowing us to compare turn-related (original) and non-turn-related (randomly permuted) looking behavior from the same data.
-The random permutations represent average anticipatory gaze rate over all possible starting points; a random baseline.
-If we anticipatory gaze rates turned out to the same for the original and randomly permuted versions of the data, we would have no evidence for turn-relevant gaze switching.
-Further, by running both data sets through the identical statistical models, we can estimate how likely it is that predictor effects in the original data (e.g., the effect of language condition) arose from random looking.
-
-We completed this random baseline procedure on 5,000 permutations of the original turn transition analysis windows and then compared the beta estimates from the original data models (above) to the distribution of beta estimates for the models of the 5,000 randomly permuted datasets.
-We estimated whether each effect in our original statistical models differed from chance by calculating how many random beta estimates were smaller than the original beta estimate for each predictor, using the absolute value of all beta estimates for a two-tailed test.
-For example, children's original ``language condition'' beta estimate was $|$-3.29$|$, which is greater than 96.9\% of $|$beta estimates$|$ from the models of randomly-permuted data.
-This leads us to conclude that that the effect of language condition in the original model is unlikely to be the result of random gaze shifting.
-We excluded the output of random-permutation models that did not converge to remove unreliable \textit{$\beta$} estimates from our percentile calculations below (22.5\% and 24.4\% of models for children and adults, respectively).
+This procedure effectively de-links participants' gaze data from the turn structure in the original stimulus, thereby allowing us to compare turn-related (original) and non-turn-related (randomly permuted) looking behavior using the same eye movements.
+The resulting anticipatory gazes from the randomly permuted analysis windows represent an average anticipatory gaze rate over all possible starting points: a random baseline.
+By running the real and randomly permuted data sets through identical statistical models, we can also estimate how likely it is that predictor effects in the original data (e.g., the effect of language condition; Table \ref{tab:E1-models}) arose from random looking.
+
+We completed this baseline procedure on 5,000 random permutations of the original turn transition analysis windows and compared the \textit{t}-values from each predictor in the original models (Table \ref{tab:E1-models}) to the distribution of \textit{t}-values for each predictor in the 5,000 models of the randomly permuted datasets.\footnote{We report \textit{t}-values rather than beta estimates because the standard errors in the randomly permuted data models were much higher than for the original data.
+For those interested, plots of the beta and standard error distributions are available in the Appendix.} We could then test whether significant effects from the original statistical models differed from the random baseline by calculating the proportion of random data \textit{t}-values exceeded by the original \textit{t}-value for each predictor, using the absolute value of all \textit{t}-values for a two-tailed test.
+For example, children's original ``language condition'' \textit{t}-value was $|$3.429$|$, which is greater than 99.9\% of all $|$\textit{t}-value$|$ estimates from the randomly-permuted data models (i.e., \textit{p} $=$ .001).
+This leads us to conclude that that the effect of language condition in the original model was highly unlikely to be the result of random gaze shifting.
+
+\begin{figure}[h!]
+  \centering
+  \subfloat[Children]{\includegraphics[width=0.49\textwidth]{figures/E1-chi-randrun-t-vals-absolute.png}\label{fig:E1-ChiTs}}
+  \hfill
+  \subfloat[Adults]{\includegraphics[width=0.49\textwidth]{figures/E1-adu-randrun-t-vals-absolute.png}\label{fig:E1-AduTs}}
+  \caption{Random-permutation and original $|$\textit{t}-values$|$ for predictors of children and adults' anticipatory gaze rates.
+Gray dots $=$ random model estimates, White dots $=$ original model estimates, Triangles $=$ 95th percentile for each \textit{t}-value distribution.}
+\end{figure}
+
+% PART UNFINISHED (potentially)
+% We excluded the output of random-permutation models that did not converge to remove unreliable model results from our percentile calculations below (XX\% and XX\% of models for children and adults, respectively).
+
+Our baseline analyses revealed that none of the significant predictors from models of the original, turn-related data can be explained by random looking.
+The children's data showed strong evidence of differentiation from the randomly permuted data for all four significant effects in the original model (Table \ref{tab:E1-models}: Children):  the original \textit{t}-values for language condition, gap duration, the age-language condition interaction, and the language condition-transition type interaction were all greater than 95\% of \textit{t}-values for the randomly permuted data (99.9\%, 95.5\%, 99.4\%, and 96\%, respectively; Figure \ref{fig:E1-ChiTs}).
+Similarly, the adults' data showed significant differentiation from the randomly permuted data for two of the three originally significant predictors---transition type and the transition type-gap duration interaction (greater than 99.9\% and 99.7\% of random \textit{t}-values, respectively)---with marginal differentiation for the interaction of language condition and transition type (greater than 94.7\% of random \textit{t}-values; Figure \ref{fig:E1-AduTs}).
+The effects of language condition and transition type for the real and randomly permuted data can also be observed in Figure \ref{fig:E1-randvsreal}.
 
 \begin{figure}[h]
-  \centering
-  \subfloat[Children]{\includegraphics[width=0.49\textwidth]{figures/E1-chi-randrun-dist.png}\label{fig:E1-ChiBetas}}
-  \hfill
-  \subfloat[Adults]{\includegraphics[width=0.49\textwidth]{figures/E1-adu-randrun-dist.png}\label{fig:E1-AduBetas}}
-  \caption{Random-permutation and original $|$beta estimates$|$ for predictors of children and adults' anticipatory gaze rates.
-Gray dots $=$ random model estimates, White dots $=$ original model estimates, Triangles $=$ 95th percentile for each beta estimate distribution.}
-\end{figure}
-
-Most of the significant predictors from models of the original, turn-related data can not be explained by random looking.
-The children's data showed strong evidence of differentiation between the random and original beta estimates for the effect of language condition ($>$96.9\% of random beta estimates) and the age-language condition interaction ($>$93.9\%), with somewhat weaker support for the language condition-transition type interaction ($>$85.1\%), and no support for the effect of gap duration ($>$73\%; Figure \ref{fig:E1-ChiBetas}).
-The adult data showed moderate evidence for the effect of transition type ($>$88.6\%), weak evidence for the interaction of transition type and duration ($>$80.6\%), and no evidence for the interaction of language condition and transition type ($>$59.8\%; Figure \ref{fig:E1-AduBetas}).
-
-\begin{figure}[h]
 \begin{center}
@@
 \end{center}
-\caption{Anticipatory gaze rates across language condition and transition type for the real (red and blue) and randomly permuted (gray) data.} 
+\caption{Anticipatory gaze rates across language condition and transition type for the real (red and blue) and randomly permuted (gray) data. Vertical bars represent the standard error.} 
 \label{fig:E1-randvsreal}
@@
 
-% Statistical model comparing random to real data at each age.
+\begin{figure}[h!]
+\begin{center}
+\includegraphics[width=0.8\textwidth]{figures/E1-child-randvsreal-ttest-agebylg.png}
+\end{center}
+\caption{Pairwise comparisons of the language condition effect across ages for the original data (black dots) and the 5,000 randomly permuted datasets (distribution).} 
+\label{fig:E1-lgageinteraction}
+\end{figure}
+
+\paragraph{Developmental effects}
+The model of the children's data revealed a significant interaction of age and language condition (tab:E1-models) that was highly unlikely to have derived from random looking (Figure \ref{fig:E1-randvsreal}).
+To further explore this effect, we compared the average effect of language condition across all ages: we extracted the average difference score for the two language conditions (English minus non-English) for each subject, computing an overall average for each random permutation of the data.
+For each random permutation, we then made pairwise comparisons of the average difference scores across ages.
+Figure \ref{fig:E1-lgageinteraction} plots the real-data difference scores against the random-data difference score distribution for each pairwise age comparison, showing that 3- and 4-year olds were affected equally by language condition, but that 5-year-olds affected less than both 3- and 4-year-olds (with 99.52\% and 99.96\% of difference scores greater than the randomly permuted data, respectively, i.e., differences of \textit{p} $<$ .01 and \textit{p} $<$ .001).
+
+\subsection{Discussion}
+\label{sec:discussion1}
+
+Children and adults spontaneously tracked the turn structure of the conversations, making anticipatory gaze switches at an above-chance rate across all ages and conditions (Table \ref{tab:e1_look} ; Figure \ref{fig:E1-randvsreal}).
+Children's anticipatory gaze rates were affected by language condition, transition type, age, and gap duration (Table \ref{tab:E1-models}), none of which could be explained by a baseline of random gaze switching (Figure \ref{fig:E1-ChiTs}).
+
+Language condition (English vs.
+non-English) affected children's anticipations in two ways (Table \ref{tab:E1-models};  Figure \ref{fig:E1-randvsreal}).
+First, children made more anticipatory switches overall in English videos, compared to non-English videos.
+This effect suggests that lexical access is important for children's ability to anticipate upcoming turn structure; children had no lexical access to the speech in the non-English videos, though they did have access to (non-native) prosodic cues and non-verbal behavior, consistent with prior work on turn-end prediction in adults (\citealp{de-ruiter2006}; \citealp{magyari2012}) and children \citep{keitel2013}.
+Second, children systematically made more anticipatory switches after hearing a question compared to a non-question, but only in the English condition, suggesting that, when children have access to lexical cues, they are more likely to make an anticipatory gaze switch when they can expect an immediate response from the addressee (i.e., an answer).
+In itself, this explanation also suggests that children are picking up on lexical cues to questionhood in anticipating upcoming answers (e.g., subject-auxiliary inversion, \textit{wh}-words, etc.).
+
+Children's anticipatory gaze switches were also affected by their age, but only in the non-English videos: 3- and 4-year-olds made many more anticipatory switches when watching videos in English compared to non-English, but this effect of language condition had attenuated significantly by age 5 (Table \ref{tab:E1-models};  Figure \ref{fig:E1-randvsreal}; Figure \ref{fig:E1-lgageinteraction}).
+This interaction suggests that the 5-year-olds were able to leverage anticipatory cues in the non-English videos in a way that 3- and 4-year-olds could not, possibly by shifting more attention to the non-native prosodic or non-verbal cues.
+Prior work on children's turn-structure anticipation proposed that children's turn-end predictions center on lexicosyntactic structure (and not, e.g., prosody) as they get older.
+The current results suggest more flexibility in children's predictions; when they do not have access to linguistic structure they are more likely to rely on alternative cues as they get older.
+
+Finally, children showed an effect of gap duration (Table \ref{tab:E1-models}).
+This effect is straightforward: longer gaps resulted in longer analysis windows, yielding more time for children to make an anticipatory gaze.
+
+Adults' anticipatory gaze rates were also affected by transition type, language condition, and gap duration (Table \ref{tab:E1-models}), none of which could be easily explained by a baseline of random gaze switching (Figure \ref{fig:E1-AduTs}).
+Like children, adults made more anticipatory switches after hearing questions compared to non-questions, suggesting that anticipation mattered more to them when an immediate response was expected.
+Also like children, the advantage for questions was driven by lexical access such that adults must have relied on lexicosyntactic cues to questionhood in picking out turns that potentially require an immediate response, though this effect was only marginally significant compared to the distribution of randomly permuted data (\textit{p} $=$ .053; Figure \ref{fig:E1-AduTs}).
+Finally, adults' anticipation rates were also affected by gap duration, but more so for questions than non-questions (Table \ref{tab:E1-models}).
+This interaction suggests that adults were less likely overall to make anticipatory switches at non-questions (as is evident for adults and children in Figure \ref{fig:E1-randvsreal}), and so did not benefit from extra time to do so compared to long gaps for questions.
 
 \subsubsection{Summary}
-
-%Several results from Experiment 1 were unexpected.
-First, we found no significant effect of age in the children's data, and we saw very little difference between adults and children (besides the age-related interactions).
-It appears that, especially in English, children were as good at tracking the current speaker and anticipating upcoming turns as adults were.
-To find out when children first begin making spontaneous anticipatory looks to responders, it is necessary to recruit younger participants.
-Experiment 2 thus includes a sample of one- and two-year-olds.
-
-%Second, children, but not adults, showed an advantage for lexicosyntactic information in their anticipatory switches.
-An advantage for lexicosyntactic information is consistent with prior work on adult turn taking \citep{de-ruiter2006}, so it's curious that the adults did not show a difference in anticipatory switching between videos they did and did not understand.
-As mentioned above, it is likely that they used non-verbal cues to track the interactions and consequently found the videos so easy to follow that they did not track the turn taking closely.
-To identify which linguistic cues adults and children rely on, it is necessary to control the presence of non-verbal cues in the stimulus---Experiment 2 implements this control.
-
-%Finally, we had predicted that children would show early advantages for prosodic cues over syntactic ones because children begin developing language-specific knowledge about prosody long before lexicosyntax.
-We weren't able to directly test this prediction in Experiment 1 because we used a limited range of ages (three- to five-year-olds) and because we only controlled for lexicosyntactic information.
-Children had prosodic cues in both English and non-English videos, although the prosodic signal was non-native in the latter case.
-So, although using non-English videos was a natural way to control for lexical information, it did not allow us to assess the role of prosody.
-Relatedly, the advantage for questions was greater in the English videos than the non-English videos, suggesting that lexicosyntactic cues are important for children in identifying questions during conversation.
-Again, to test this idea, we would need to directly compare lexicosyntactic and prosodic cues in the participants' native language.
+\label{sec:summary1}
+
+Children and adults' predictions alike were benefited by access to lexical information (English) and speech act status (questionhood), suggesting that linguistic cues, particularly lexical ones, facilitate their spontaneous predictions about upcoming turn structure through the identification of turns that will have immediate responses from the addressee.
+Children's anticipatory gaze rates for questions and non-questions in English was stable across ages and comparable to adult behavior (Figure \ref{fig:E1-randvsreal}), suggesting that they can identify questions in native stimuli with adult-like competence by age three.
+Although participants' ability to recognize questions was facilitated by lexical access (i.e., English vs.
+non-English), the prosody in the non-English videos was non-native, and so the experimental design can not conclusively show which linguistic cues children relied on in the English videos to identify question turns.
+
+Relatedly, though lexical access clearly facilitated participants' anticipatory gaze rate, it was not necessary for participants, especially adults, in order to exceed chance switching rates (Figure \ref{fig:E1-randvsreal}), suggesting that participants use non-lexical cues (e.g., prosody, non-verbal behavior) to make anticipatory eye movements at least some of the time.
+In the case of adults there was no overall effect of language condition, only an interaction between language condition and transition type: adults spontaneously anticipated upcoming speakers whether they had lexical access or not, but when given lexical access, focused their anticipations on turns with immediate responses (questions).
+
+Interestingly, adults and children both were strongly affected by transition type, in that they made more anticipatory switches after hearing questions, compared to non-questions.
+We had hypothesized that a projected response might affect participants' predictions but it in fact seems to be driving many of the effects in our data.
+Even in the English videos, when participants had full access to linguistic cues, their rates of anticipation were relatively low---in fact, comparable to the non-English videos---unless the turn was a question (Figure \ref{fig:E1-randvsreal}).
+Prior work using online, metalinguistic tasks has shows that participants can use linguistic cues to accurately predict upcoming turn ends.
+The current results suggest that, in their spontaneous predictions, both children and adults monitor the linguistic structure of unfolding turns for cues to upcoming responses.
+Because the form of an answer is partially determined by its question (e.g., locative expressions for ``where'' questions), participants could even be predicting what type of response they will hear, and so they are looking anticipatorily at the responder for confirmation of their prediction.
+This interpretation is initial, but in-line with work showing early response planning and speech act recognition \citep{gisladottir2015, bogels2015}.
+
+Overall, children and adults behaved relatively similarly and our language manipulation (English vs.
+non-English) was too coarse to comment on when children begin to use different types of native linguistic cues (e.g., prosody vs.
+lexicosyntax); we would instead need to directly compare lexicosyntactic and prosodic cues in the participants' native language, controlling for the presence of non-verbal cues.
+To see the emergence of anticipatory gaze switching we would also need to include younger children since participants already reliably made anticipatory gaze switches at age three.
 In sum, Experiment 1 lays the analytic groundwork for a method that allows for greater experimental control, which we introduce in Experiment 2.
-
-%% E1 DISCUSSION LEFTOVER TEXT
-
-%Language did not affect adults' anticipatory looking behavior, possibly because the animated, child-directed prosody and non-verbal cues made it possible for them to follow the interaction and track the current speaker easily.
-% As predicted, children made more anticipatory looks while they were watching videos that they could understand.
-In the English videos, children had access to lexicosyntactic information, which could have helped them make better predictions about upcoming turn structure.
-Although children performed better on the English videos, lexicosyntactic information was not exclusively responsible for their anticipatory switches; children still performed remarkably well when no lexical information was present at all, and even three-year-olds made anticipatory switches at an above-chance rate in the non-English videos.
-%Contrary to our expectation, there was no effect of age.
-Children's anticipatory switches \textit{did} increase with age for the non-English videos, but it also decreased with age for the English videos, resulting in a significant interaction of age and language condition.
-We had expected that, as children got older, they would also show more anticipatory switches.
-But this is contradicted by decreased looks with age for English videos, which even extended to the adult data.
-One possible explanation is that older children and adults found the videos in English easy to comprehend, and so tracked the conversation less closely.
-Compared to three-year-olds, adults and older children spent more time looking away from the current speaker and addressee in the English conversations.
-In contrast, compared to three-year-olds, they spent more time looking at the current speaker and less time looking away in the non-English conversations (Table \ref{tab:e1_look}).
-%It is also possible that the timing of anticipatory gaze switches changes as children get older.
-That is, the moment at which observers chose to launch their gaze away from the current speaker might change with age.
-For example, children could have reacted more immediately to early cues in the turn (e.g., ``Which one...'') as they got older.
-If a change in timing were sufficiently large, we might not capture it in our analysis window, which was designed to capture switches close to the turn transition.
-To test for a change in switch timing, we would have to run new analyses on a larger set of turn transitions---ones with controlled early and late cues \citep[see, e.g.,][]{bogels2015}.
-Within the current analysis windows there were some ``early switches'' (i.e., those launched before the end of the prior turn).
-But participants made approximately equal numbers of early switches, regardless of age and condition: 11--15\% (Figure \ref{fig:cumulativeFL}).
-The lack of increased early switching with age might suggest that changes in timing are either larger than the analysis window, or are not at play.
-%Children made more anticipatory switches after hearing questions than non-questions, but the advantage for questions was driven by age and linguistic access.
-There was no effect of transition type alone.
-Rather, the model confirmed a two-way interaction of linguistic condition and transition type.
-Presumably, question effects were larger in the English videos than the non-English videos because there were more cues to questionhood in English (i.e., non-verbal, prosodic, \textit{and} lexicosyntactic).
-Children younger than five did not distinguish questions from non-questions in the non-English videos (Figure \ref{fig:questionsFL}).
-The children's model confirmed a three-way interaction of age, linguistic condition, and transition type.
-Lexicosyntactic information might then be critical for young children to identify questions in conversation.
-Cues like subject-auxiliary inversion, and \textit{wh}-words occur early in the turn and are clear indicators that an answer will come next.
-Importantly, the turns marked as ``questions'' in the non-English data were not real questions, but turns that \textit{sounded like} questions to native English speakers.
-Question-like prosodic cues in the non-English videos could have been markedly different from question prosody in English, making it more difficult for three- and four-year-olds to identify ``questions'' in those conversations.
-%The increased switching after questions indicates that, so long as they have native lexical and prosodic cues, children as young as three spontaneously distinguish questions from other types of speech acts in third-party conversation.
-Children may treat questions differently from non-questions in conversation because they are predictably followed by an immediate response (i.e., the answer).
-Because the form of an answer is partially determined by its question (e.g., locative expressions for ``where'' questions), children could even be predicting what type of response they will hear, and look at the responder for confirmation.
-In contrast, non-questions do not usually require an immediate response, so it is not as easy to predict what will come next.
 
@@
 
-We improved our design by using native-language stimuli, controlling for lexical \textit{and} prosodic information, eliminating non-verbal cues, and testing children from a wider age range.
+We improved our design by using native-language stimuli, controlling for lexical and prosodic information, eliminating non-verbal cues, and testing children from a wider age range.
 All of the videos in Experiment 2 were in the participants' native language (American English).
@@
 We created 18 short videos of improvised, child-friendly conversation (Figure \ref{fig:puppets}).
-To eliminate non-verbal cues to turn transition and to control the types of linguistic information available in the stimuli we (1) recorded improvised conversations, (2) phonetically manipulated those recordings to limit the availability of prosodic and lexical information, and (3) recorded a new set of videos that featured puppets as talkers, using the manipulated audio as the puppets' speech.
+To eliminate non-verbal cues to turn transition and to control the types of linguistic information available in the stimuli we first audio-recorded improvised conversations, then phonetically manipulated those recordings to limit the availability of prosodic and lexical information, and finally recorded video to accompany the manipulated audio, featuring puppets as talkers.
 
@@
 These were rare in the clips.} We made duration and pitch manipulations using PSOLA resynthesis in Praat \citep{Praat}.
-Thus, the \textit{words only} versions of the audio clips had no pitch or durational cues to upcoming turn boundaries, but did have intact lexicosyntactic cues (and residual phonetic correlates of prosody, like intensity).
+Thus, the \textit{words only} versions of the audio clips had no pitch or durational cues to upcoming turn boundaries, but did have intact lexicosyntactic cues (and residual phonetic correlates of prosody, e.g., intensity).
 
 We created the \textit{prosody only} clips by low-pass filtering the original recording at 500 Hz with a 50 Hz Hanning window (following de Ruiter et al., 2006).
-This manipulation creates a ``muffled speech'' sound because low-pass filtering removes most of the phonetic information used to distinguish between phonemes.
+This manipulation creates a ``muffled speech'' effect because low-pass filtering removes most of the phonetic information used to distinguish between phonemes.
 The \textit{prosody only} versions of the audio clips lacked lexical information, but retained their intonational and rhythmic cues to upcoming turn boundaries.
@@
 The \textit{no speech} condition served as a non-linguistic baseline.
-For this condition, we replaced the original clip with multi-talker babble: Eight different child-oriented conversations (not including the original one), overlaid and cropped to the duration of the video.
+For this condition, we replaced the original clip with multi-talker babble: We overlaid different child-oriented conversations (not including the original one), and then cropped the result to the duration of the original video.
 Thus, the \textit{no speech} audio clips lacked any linguistic information to upcoming turn boundaries---the only cue to turn taking was the opening and closing of the puppets' mouths.
@@
 Finally, because low-pass filtering removes significant acoustic energy, the \textit{prosody only} clips were much quieter than the other three conditions.
-Our last step was to downscale the intensity of videos from the three other conditions to match the volume of the \textit{prosody only} clips.
+Our last step was to downscale the intensity of the audio tracks in the three other conditions to match the volume of the \textit{prosody only} clips.
 We referred to the conditions as ``normal'', ``robot'', ``mermaid'', and ``birthday party'' speech when interacting with participants.
@@
 Puppets were positioned looking forward to eliminate shared gaze as a cue to turn structure \citep{thorgrimsson2015}.
-We took care to match the puppets' mouth movements to the syllable onsets as closely as possible, and avoided any mouth movement before the onset of a turn.
+We took care to match the puppets' mouth movements to the syllable onsets as closely as possible, specifically avoiding any mouth movement before the onset of a turn.
 We then added the manipulated audio clips to the puppet video recordings by hand.
 
-We used three pairs of puppets used for the \textit{normal} condition---`red', `blue' and `yellow'---and one pair of puppets for each limited cue condition: ``Robots'', ``merpeople'', and ``party-goers'' (Figure 8).
+We used three pairs of puppets used for the \textit{normal} condition---`red', `blue' and `yellow'---and one pair of puppets for each limited cue condition: ``robots'', ``merpeople'', and ``party-goers'' (Figure 8).
 We randomly assigned half of the conversation topics (`birthday cake', `pets', and `breakfast') to the \textit{normal} condition, and half to the limited cue conditions (`riding bikes', `rainy days', and `the library').
 We then created three versions of the experiment, so that each of the six puppet pairs was associated with three different conversation topics across the different versions of the experiment (18 videos in total).
-We ensured that the position of the talkers (left and right) was counterbalanced in each version by flipping the video and audio channels as needed.\footnote{See the videos here: www.youtube.com/channel/UCGEZQcM9t8Zfjqqi\_B1Q5Sw.}
+We ensured that the position of the talkers (left and right) was counterbalanced in each version by flipping the video and audio channels as needed.
+
+% PART UNFINISHED: can add a link at the end of the prior paragraph to video examples
 
@@
 
-There were 79 turn transitions for analysis, after excluding transitions longer than 550 msec and shorter than 0 msec.
-The remaining turn transitions were distributed evenly across transition types (questions N$=$47 and non-questions N$=$32) and conditions, keeping in mind that there were three \textit{normal} videos and only one limited cue video for each experiment version: \textit{Normal} (N$=$36), \textit{words only} (N$=$13), \textit{prosody only} (N$=$12), and {no speech} (N$=$18).
-Inter-turn gaps for questions (mean$=$358, median$=$405) were longer than those for non-questions (mean$=$296, median$=$288) on average, but gap duration was overall comparable across conditions: \textit{Normal} (mean$=$333, median$=$337), \textit{words only} (mean$=$345, median$=$383), \textit{prosody only} (mean$=$320, median$=$336), and \textit{no words} (mean$=$324, median$=$328).
+% PART UNFINISHED: update gap information
+There were 85 turn transitions for analysis after excluding transitions longer than 550 msec and shorter than 90 msec.
+The remaining turn transitions had slightly more questions than non-question (N$=$50 and N$=$35, respectively), with transitions distributed somewhat evenly across conditions (keeping in mind that there were three \textit{normal} videos and only one limited cue video for each experiment version): \textit{Normal} (N$=$36), \textit{words only} (N$=$13), \textit{prosody only} (N$=$17), and {no speech} (N$=$19).
+Inter-turn gaps for questions (mean$=$365, median$=$427) were longer than those for non-questions (mean$=$302, median$=$323) on average, but gap duration was overall comparable across conditions: \textit{Normal} (mean$=$334, median$=$321), \textit{words only} (mean$=$347, median$=$369), \textit{prosody only} (mean$=$365, median$=$369), and \textit{no words} (mean$=$319, median$=$329).
 The longer gaps for question transitions could give them an advantage because our anticipatory measure includes shifts initiated during the gap between turns (Figure \ref{fig:criterion}).
@@
 
-\textit{Data preparation, coding, and random baseline analysis}.
+\subsubsection{Data preparation and coding}
 We coded each turn transition for its linguistic condition (\textit{normal}, \textit{words only}, \textit{prosody only}, and \textit{no speech}) and transition type (question/non-question)\footnote{We coded \textit{wh-}questions as ``non-questions'' for the \textit{prosody only} videos.
-Polar questions had a final rising prosodic contour, but \textit{wh-}questions did not  \citep{hedberg2010}.}, identified anticipatory gaze switches to the upcoming speaker, and estimated a random baseline of anticipatory switches using the methods from Experiment 1.
-
-\subsection{Results and discussion}
+Polar questions had a final rising prosodic contour, but \textit{wh-}questions did not  \citep{hedberg2010}.} and identified anticipatory gaze switches to the upcoming speaker using the methods from Experiment 1.
+
+\subsection{Results}
 \label{sec:results2}
 
-%\subsubsection{Descriptive analysis}
-%Participants' pattern of gaze indicated that they performed basic turn tracking across all ages and in all conditions.
+Participants' pattern of gaze indicated that they performed basic turn tracking across all ages and in all conditions.
 Participants looked at the screen most of the time during video playback (82\% and 86\% average for children and adults, respectively).
@@
 Looks to the addressee also decreased with age, but the change was minimal.
-%
-%\begin{table}[t]
-%\begin{center}
-%  \begin{tabular}{llcccc}
-%    \hline
-%    Age group & Speaker & Addressee & Other onscreen & Offscreen\\ 
-%    \hline
-%    1 & 0.44 & 0.14 & 0.23 & 0.19 \\ 
-%    2 & 0.50 & 0.13 & 0.24 & 0.14 \\ 
-%    3 & 0.47 & 0.12 & 0.25 & 0.16 \\ 
-%    4 & 0.48 & 0.11 & 0.29 & 0.12 \\ 
-%    5 & 0.54 & 0.11 & 0.20 & 0.14 \\ 
-%    6 & 0.60 & 0.12 & 0.18 & 0.10 \\
-%    Adult & 0.69 & 0.12 & 0.09 & 0.10 \\
-%%   Overall & 0.53 & 0.12 & 0.21 & 0.13 \\
-%    \hline
-%  \end{tabular}
-%\end{center}
-%  \caption{Average proportion of gaze to the current speaker and addressee during periods of talk across ages.}
-%\label{tab:look_e2}
-%\end{table}
-%
-%\begin{table}[t]
-%\begin{center}
-%  \begin{tabular}{llcccc}
-%    \hline
-%    Condition & Speaker & Addressee & Other onscreen & Offscreen\\ 
-%    \hline
-%    Normal & 0.58 & 0.12 & 0.17 & 0.13 \\ 
-%    Words only & 0.54 & 0.11 & 0.24 & 0.10 \\ 
-%    Prosody only & 0.48 & 0.12 & 0.26 & 0.15 \\ 
-%    No speech & 0.44 & 0.13 & 0.26 & 0.18 \\
-%    \hline
-%  \end{tabular}
-%\end{center}
-%  \caption{Average proportion of gaze to the current speaker and addressee during periods of talk across conditions.}
-%\label{tab:look_e2b}
-%\end{table}
-%
-%% Average prop gaze to the current speaker across conditions for one-year-olds
-%%                           speaker addressee other-screen offscreen
-%%Normal   		  0.4737582 0.1366432    0.2003952 0.1892035
-%%Words only    0.3977261 0.1283417    0.3112005 0.1627317
-%%Prosody only 0.4290202 0.1611453    0.2558657 0.1539688
-%%No speech    0.3786925 0.1228615    0.2356423 0.2628037
-%
+
+\begin{table}[h]
+\begin{center}
+  \begin{tabular}{llcccc}
+    \hline
+    Age group & Speaker & Addressee & Other onscreen & Offscreen\\ 
+    \hline
+    1 & 0.44 & 0.14 & 0.23 & 0.19 \\ 
+    2 & 0.50 & 0.13 & 0.24 & 0.14 \\ 
+    3 & 0.47 & 0.12 & 0.25 & 0.16 \\ 
+    4 & 0.48 & 0.11 & 0.29 & 0.12 \\ 
+    5 & 0.54 & 0.11 & 0.20 & 0.14 \\ 
+    6 & 0.60 & 0.12 & 0.18 & 0.10 \\
+    Adult & 0.69 & 0.12 & 0.09 & 0.10 \\
+%   Overall & 0.53 & 0.12 & 0.21 & 0.13 \\
+    \hline
+  \end{tabular}
+\end{center}
+  \caption{Average proportion of gaze to the current speaker and addressee during periods of talk across ages.}
+\label{tab:look_e2}
+\end{table}
+
+\begin{table}[h]
+\begin{center}
+  \begin{tabular}{llcccc}
+    \hline
+    Condition & Speaker & Addressee & Other onscreen & Offscreen\\ 
+    \hline
+    Normal & 0.58 & 0.12 & 0.17 & 0.13 \\ 
+    Words only & 0.54 & 0.11 & 0.24 & 0.10 \\ 
+    Prosody only & 0.48 & 0.12 & 0.26 & 0.15 \\ 
+    No speech & 0.44 & 0.13 & 0.26 & 0.18 \\
+    \hline
+  \end{tabular}
+\end{center}
+  \caption{Average proportion of gaze to the current speaker and addressee during periods of talk across conditions.}
+\label{tab:look_e2b}
+\end{table}
+
+% Average prop gaze to the current speaker across conditions for one-year-olds
+%                           speaker addressee other-screen offscreen
+%Normal   		  0.4737582 0.1366432    0.2003952 0.1892035
+%Words only    0.3977261 0.1283417    0.3112005 0.1627317
+%Prosody only 0.4290202 0.1611453    0.2558657 0.1539688
+%No speech    0.3786925 0.1228615    0.2356423 0.2628037
+
+\begin{table}[h!]
+\begin{center}
+  \begin{tabular}{lcccl}
+  \textbf{\textit{Children}} &&&& \\
+    \hline
+                       & Estimate & Std. Error & z value & Pr($>$$|$z$|$)    \\
+    \hline
+    (Intercept)                & -3.57414  &   0.48576 &  -7.358 & 1.87e-13 ***    \\
+    Age                         & 0.02543 &    0.10260  &  0.248   & 0.8042        \\
+    Condition$=$\textit{muffled}         &   -1.63429  &   0.86390 &  -1.892   & 0.0585 .      \\
+    Condition$=$\textit{normal}            &  0.36710 &    0.43296 &   0.848  &  0.3965        \\
+    Condition$=$\textit{robot}            &  -0.26741 &    0.59071 &  -0.453  &  0.6508        \\
+    Type$=$\textit{non-Question}                      & -0.81873  &   0.59985 &  -1.365  &  0.1723        \\
+    Duration                   &  4.17672  &   0.62446  &  6.689 & 2.25e-11 ***    \\
+    Age*Condition$=$\textit{muffled}       &   0.39317 &     0.18907 &   2.080  &  0.0376 *      \\
+    Age*Condition$=$\textit{normal}       &   0.12919 &    0.10227  &  1.263  &  0.2065        \\
+    Age*Condition$=$\textit{robot}           &   0.13740  &   0.13568   & 1.013  &  0.3112        \\
+    Age*Type$=$\textit{non-Question}                &    0.15116   &  0.13643  &  1.108  &  0.2679        \\
+    Condition$=$\textit{muffled}*     &  1.77190  &   1.24864  &  1.419 &   0.1559        \\
+    Type$=$\textit{non-Question} &&&& \\
+    Condition$=$\textit{normal}*     &   0.91059  &   0.72095  &  1.263 &   0.2066        \\
+    Type$=$\textit{non-Question} &&&& \\
+    Condition$=$\textit{robot}*     &   -1.02193  &   1.01227  & -1.010  &  0.3127        \\
+    Type$=$\textit{non-Question} &&&& \\
+    Age*Condition$=$\textit{muffled}* &  -0.47057 &    0.28703 &  -1.639 &   0.1011        \\
+    Type$=$\textit{non-Question} &&&& \\
+    Age*Condition$=$\textit{normal}* &   -0.37542  &   0.16963  & -2.213 &   0.0269 *      \\
+    Type$=$\textit{non-Question} &&&& \\
+    Age*Condition$=$\textit{robot}*  &   0.08946  &   0.22349   & 0.400 &   0.6890        \\
+    Type$=$\textit{non-Question} &&&& \\
+    \hline
+  &&&& \\
+  \textbf{\textit{Adults}} &&&& \\
+    \hline
+                       &  Estimate & Std. Error & z value & Pr($>$$|$z$|$) \\    
+    \hline
+    (Intercept)           &  -3.4557   &   0.7199 &  -4.800 & 1.58e-06 *** \\ 
+    Condition$=$\textit{muffled}     &     0.3349  &    0.8965  &  0.374 & 0.708692     \\ 
+    Condition$=$\textit{normal}    &      1.2556   &   0.5633  &  2.229 & 0.025805 *   \\ 
+    Condition$=$\textit{robot}       &     1.5938  &    0.7208 &   2.211 & 0.027023 *   \\ 
+    Type$=$\textit{non-Question}                  &   0.4292  &    0.6089  &  0.705 & 0.480916     \\ 
+    Duration             &     4.7500   &   1.2480  &  3.806 & 0.000141 *** \\ 
+    Condition$=$\textit{muffled}*  &  0.6627  &    1.2138 &   0.546 & 0.585108     \\ 
+    Type$=$\textit{non-Question} &&&& \\
+    Condition$=$\textit{normal}* &   -0.9452  &    0.7631 &  -1.239 & 0.215475     \\ 
+    Type$=$\textit{non-Question} &&&& \\
+    Condition$=$\textit{robot}*   &  -1.1265  &    0.9109  & -1.237 & 0.216201     \\ 
+    Type$=$\textit{non-Question} &&&& \\
+    \hline
+  \end{tabular}
+\end{center}
+  \caption{Model output for children and adults' anticipatory gaze switches.}
+\label{tab:E2-models}
+\end{table}
+
+\subsubsection{Statistical models}
+\label{sec:models2}
+
+We identified anticipatory gaze switches for all 85 usable turn transitions, and analyzed them for effects of language condition, transition type, and age with two mixed-effects logistic regressions \citep{lme4, R}.
+We again built separate models for children and adults because effects of age were only pertinent to the children's data.
+The child model included condition (normal/prosody only/words only/no speech; with no speech in the intercept), transition type (question vs.
+non-question), age (1, 2, 3, 4, 5, 6), and duration of the inter-turn gap (in seconds) as predictors, with full interactions between language condition, transition type, and age.
+We again included the duration of the inter-turn gap as a control predictor and added random effects of item (turn transition) and participant, with random slopes of transition type for participants \citep{barr2013}.
+The adult model included condition, transition type, their interactions, and duration as a control predictor, with participant and item included as random effects and random slopes of condition and transition type.
+
+Children's anticipatory gaze switches showed an effect of gap duration (\textit{$\beta$}=4.18, \textit{SE}=0.624, \textit{t}=6.689, \textit{p}$<$.001), a two-way interaction of age and language condition (for prosody only speech compared to the no speech intercept; \textit{$\beta$}=0.393, \textit{SE}=0.189, \textit{t}=2.08, \textit{p}$<$.05), and a three-way interaction of age, transition type, and language condition (for normal speech compared to the no speech intercept; \textit{$\beta$}=-0.375, \textit{SE}=0.17, \textit{t}=-2.213, \textit{p}$<$.05).
+There were no significant effects of age or transition type alone (Table \ref{sec:models}), with only a marginal effect of language condition (for prosody only compared to the no speech intercept; \textit{$\beta$}=-1.634, \textit{SE}=0.864, \textit{t}=-1.89, \textit{p}$=$.06)
+
+Adults' anticipatory gaze switches showed effects of gap duration (\textit{$\beta$}=4.75, \textit{SE}=1.248, \textit{t}=3.806, \textit{p}$<$.001) and language condition (for normal speech \textit{$\beta$}=1.256, \textit{SE}=0.563, \textit{t}=2.229, \textit{p}$<$05.
+and words only speech \textit{$\beta$}=1.594, \textit{SE}=0.721, \textit{t}=2.211, \textit{p}$<$.05  compared to the no speech intercept).
+There were no effects of transition type (\textit{$\beta$}=0.429, \textit{SE}=0.609, \textit{t}=0.705, \textit{p}$=$.48).
+
+\begin{figure}[h!]
+  \centering
+  \subfloat[Children]{\includegraphics[width=0.49\textwidth]{figures/E2-chi-randrun-t-vals-absolute.png}\label{fig:E2-ChiTs}}
+  \hfill
+  \subfloat[Adults]{\includegraphics[width=0.49\textwidth]{figures/E2-adu-randrun-t-vals-absolute.png}\label{fig:E2-AduTs}}
+  \caption{Random-permutation and original $|$\textit{t}-values$|$ for predictors of children and adults' anticipatory gaze rates.
+Gray dots $=$ random model estimates, White dots $=$ original model estimates, Triangles $=$ 95th percentile for each \textit{t}-value distribution.}
+\end{figure}
+
+\subsubsection{Random baseline comparison}
+\label{sec:randbaseline2}
+
+Using the same technique described in experiment 1 (Section \ref{sec:randbaseline1}), we created and modeled 5,000 random permutations of participants' anticipatory gaze.
+Our baseline analyses revealed that none of the significant predictors from models of the original, turn-related data (Table \ref{tab:E2-models}: Children) can be explained by random looking.
+In the children's data, the original \textit{t}-values for language condition (prosody only), gap duration, the two-way interaction of age and language condition (prosody only) and the three-way interaction of age, transition type, and language condition (normal only) were all greater than 95\% of the randomly permuted \textit{t}-value (96.2\%, 100\%, 95.1\%, and 95.1\%, respectively; Figure \ref{fig:E2-ChiTs}).
+Similarly, the adults' data showed significant differentiation from the randomly permuted data for all originally significant predictors: language condition for normal and words only speech and gap duration (greater than 96.8\%, 98.7\%, and 100\% of random \textit{t}-values, respectively; Figure \ref{fig:E2-AduTs}).
+The effects of language condition and transition type for the real and randomly permuted data can also be observed in Figure \ref{fig:E1-randvsreal}.
+
+% PART UNFINISHED (potentially)
+% We excluded the output of random-permutation models that did not converge to remove unreliable model results from our percentile calculations below (XX\% and XX\% of models for children and adults, respectively).
+
+\begin{figure}[h]
+\begin{center}
+\includegraphics[width=0.9\textwidth]{figures/E2-samples-by-lang-groups-trans-types.png}
+\end{center}
+\caption{Anticipatory gaze rates across language condition and transition type for the real (blue, dark green, light green, and red) and randomly permuted (gray) data.
+Vertical bars represent the standard error.}
+\label{fig:E2-randvsreal}
+\end{figure}
+
+\begin{figure}[h!]
+\begin{center}
+\includegraphics[width=0.9\textwidth]{figures/E2-child-randvsreal-ttest-muffledages.png}
+\end{center}
+\caption{Significant pairwise comparisons of the prosody only-no speech linguisitic condition effect, across ages, for the original data (black dots) and the 5,000 randomly permuted datasets (distribution).}
+\label{fig:E2-lgageinteraction}
+\end{figure}
+
+\begin{figure}[h!]
+\begin{center}
+\includegraphics[width=0.9\textwidth]{figures/E2-child-randvsreal-ttest-normaltypesages.png}
+\end{center}
+\caption{Significant pairwise comparisons of the normal speech-no speech language condition effect for questions status, across ages, for the original data (black dots) and the 5,000 randomly permuted datasets (distribution).}
+\label{fig:E2-lgagetypeinteraction}
+\end{figure}
+
+\paragraph{Developmental effects}
+The model of the children's data revealed two significant interactions with age, neither of which derived from random looking (Figure \ref{fig:E2-randvsreal}).
+The first was a significant interaction of age and language condition (for prosody only compared to the no speech baseline), suggesting a different age effect between the two linguistic conditions.
+As in Experiment 1, we further explored the age effect by extracting the average difference score over subjects for the interaction in each random permutation of the data, making pairwise comparisons between the six age groups.
+Figure \ref{fig:E2-lgageinteraction} shows the comparisons that demonstrate a significant difference in age for the no speech baseline vs.
+the prosody only condition, revealing that anticipatory gaze in the prosody only condition significantly improves with age, especially at ages 5 and 6, compared to the no-speech baseline (all with difference scores greater than 95\% of the random data scores; \textit{p} $<$ .05).
+
+The second interaction with age was a three-way interaction of age, transition type, and language condition (for normal speech compared to the no speech baseline).
+We again created pairwise comparisons of the average difference scores for the interaction in each random permutation of the data, with significant differences shown in Figure \ref{fig:E2-lgagetypeinteraction}.
+These pairwise comparisons showed that the effect of transition type in the normal speech condition became larger with age, with significant improvements by age 4 over ages 1 and 2 (99.9\% and 98.86\%, respectively), by age 5 over age 4 (97.54\%), and by age 6 over ages 1, 2, and 5 (99.5\%, 97.36\%, and 95.04\%), all equating to significant differences from chance (\textit{p} $<$ .05).
+
+\subsection{Discussion}
+\label{sec:discussion2}
+
+\subsubsection{Summary}
+\label{sec:summary2}
+
 %We identified anticipatory gaze switches for all 79 usable turn transitions using the same criteria as in Experiment 1.
@@
 
-\subsubsection{Summary}
-
 %The core aims of Experiment 2 were to gain better traction on the individual roles of prosody and lexicosyntax in children's turn predictions, and to expand our age range to capture more developmental change.
@@
 
+% PART UNFINISHED: APPENDIX
+
 \end{document}

### 2015-11-28 analysis/E1/0-helper.R

@@ -13,6 +13,10 @@ library(car)
 ################################################################################
 # Misc helper functions
 ################################################################################
+# Stat summary functions
+tstat <- function(m) fixef(m)/sqrt(diag(vcov(m)))
+SEstat <- function(m) sqrt(diag(vcov(m)))
+
 # Add some style elements for ggplot2
 plot.style <- theme(panel.grid.major = element_blank(),
     panel.grid.minor = element_blank(),
@@ -163,89 +167,159 @@ smoothLD <- function(dt) {
 ################################################################################
 # Random run plot functions
 ################################################################################
-
-combine.runs <- function(DT, filelist, name) {
-	curr.row <- 1
-	for (file in filelist) {
-		# Read in a file
-		data <- fread(paste(processed.data.path, file, sep=""))
-		# Add the data to the big data table
-		DT[curr.row, names(DT) := data]
-		curr.row <- curr.row + 1
+combine.runs <- function(betas, ses, ts, fits,
+	modelfiles, infofiles, realm, reali) {
+	# Random run info
+	for (file in modelfiles) {
+		data <- fread(paste(model.path, file, sep=""))
+		run <- as.numeric(unlist(strsplit(
+			file, "[.-]"))[5])
+		# betas
+		insert <- data.frame(t(data$B))
+		colnames(insert) <- data$Predictor
+		betas[Run == run, (data$Predictor) := insert]
+		# ses
+		insert <- data.frame(t(data$SE))
+		colnames(insert) <- data$Predictor
+		ses[Run == run, (data$Predictor) := insert]
+		# ts
+		insert <- data.frame(t(data$t))
+		colnames(insert) <- data$Predictor
+		ts[Run == run, (data$Predictor) := insert]
 	}
-	setkeyv(DT, "Run")
-	# Write it out
-	write.csv(DT, paste(processed.data.path,
-		name, ".random.runs.csv",sep=""), row.names=FALSE)
-	# Return it
-	return(DT)
+	for (file in infofiles) {
+		data <- fread(paste(model.path, file, sep=""))
+		run <- as.numeric(unlist(strsplit(
+			file, "[.-]"))[5])
[…]

### 2015-11-28 analysis/E1/3-plots.stats.analysis.R

@@ -14,12 +14,12 @@ library(data.table)
 raw.data.path <- "tracker_data/" #"../../data/E1/tracker_data/"
 info.path <- "info/" # "../../data/E1/info/"
 processed.data.path <- "processed_data/" #"../../data/E1/processed_data/"
+model.path <- "processed_data/models/"
 plot.path <- "plots/" #"plots/"
 
 # Read in supplemental data and prepped data
 gap.info <- fread(paste(info.path,"GapInfo.csv", sep=""))
 subject.info <- fread(paste(info.path,"SubjectInfo.csv",sep=""))
-
 
 # Color and line palettes
 stoplightPalette <- c("dodgerblue3", "firebrick2")
@@ -290,53 +290,380 @@ adu.max <- glmer(Switch ~ LgGroup * Type * Duration +
 #		- Random slopes maximized but don't make qualitative difference
 
 # Export the model values for use in the random runs
-chi.coefs <- data.frame(t(sapply(fixef(chi.max),c)))
-chi.coefs$Sample <- "real"
-chi.coefs$Error <- ""
-chi.coefs$Run <- 0
-
-adu.coefs <- data.frame(t(sapply(fixef(adu.max),c)))
-adu.coefs$Sample <- "real"
-adu.coefs$Error <- ""
-adu.coefs$Run <- 0
-
-# Model results
-write.csv(summary(chi.max)$coefficients, paste(
-	processed.data.path, "chi.model.csv", sep=""),
+ll.cm <- logLik(chi.max)
+chi.max.m <- data.frame(
+	B = fixef(chi.max),
+	SE = SEstat(chi.max),
+	t = tstat(chi.max))
+chi.max.m <- cbind(Predictor = rownames(chi.max.m), chi.max.m)
+rownames(chi.max.m) <- NULL
+chi.max.i <- data.frame(
+	LogLik = head(ll.cm), AIC = AIC(ll.cm),
+	NumObs = as.numeric(unlist(attributes(ll.cm))[1]),
+	Sample = "real", Error = "", Run = 0)
+chi.max.r <- resid(chi.max)
+write.csv(chi.max.m, paste(
+	model.path, "chi.max.m-0.csv", sep=""),
 	row.names=FALSE)
-write.csv(summary(adu.max)$coefficients, paste(
-	processed.data.path, "adu.model.csv", sep=""),
+write.csv(chi.max.i, paste(
[…]

### 2015-11-28 analysis/E1/4-model-functions.R

@@ -11,6 +11,7 @@ library(data.table)
 # Read in data in processed_data/ path and set plotting variables
 ################################################################################
 processed.data.path <- "processed_data/" #"../../data/E1/processed_data/"
+model.path <- "processed_data/models/"
 switch.all <- fread(paste(processed.data.path,"SwitchAll.csv", sep=""))
 chi.coefs <- fread(paste(processed.data.path,"chi.coefs-0.csv", sep=""))
 adu.coefs <- fread(paste(processed.data.path,"adu.coefs-0.csv", sep=""))
@@ -33,19 +34,38 @@ run.models <- function(ns) {
 
     		# Retrieve the random gap info data for this particular run
 	    	rand.data.C <- switch.rand.C[Run == i,]
-	    	chi.coefs.r <- chi.coefs[0,]
 
-    		# Models
-	    	# Children
 			chi.max.r <- glmer(Switch ~ Age * LgGroup * Type + Duration + 
 			    (Type * LgGroup|Subject) + (1|Gap), data=rand.data.C,
 			    family=binomial, control = glmerControl(optimizer="bobyqa"))
-    		chi.coefs.r <- rbind(chi.coefs.r,
-    			cbind(data.frame(t(sapply(fixef(chi.max.r),c))),
-    			data.frame(Sample = "random", Error = errors, Run = i)))
-    		write.csv(chi.coefs.r, paste(
-				processed.data.path, "chi.coefs.r-", i, ".csv", sep=""),
-				row.names=FALSE)		
+			ll.cmr <- logLik(chi.max.r)
+			    
+			chi.max.r.m <- data.frame(
+				B = fixef(chi.max.r),
+				SE = SEstat(chi.max.r),
+				t = tstat(chi.max.r))
+			chi.max.r.m <- cbind(Predictor = rownames(chi.max.r.m), chi.max.r.m)
+			rownames(chi.max.r.m) <- NULL
+				
+			chi.max.r.i <- data.frame(
+				LogLik = head(ll.cmr), AIC = AIC(ll.cmr),
+				NumObs = as.numeric(unlist(attributes(ll.cmr))[1]),
+				Sample = "random", Error = errors, Run = i)
+				
+			chi.max.r.r <- resid(chi.max.r)
+			 
+    		write.csv(chi.max.r.m, paste(
[…]

### 2015-11-28 analysis/E1/5-compare.R

@@ -6,259 +6,773 @@ source("0-helper.R")
 ################################################################################
 # Set paths
 processed.data.path <- "processed_data/" #"../../data/E1/processed_data/"
+model.path <- "processed_data/models/" #"../../data/E1/processed_data/"
 plot.path <- "plots/" #"plots/"
 
 # Read in all the random runs and combine them
-adu.r.files <- dir(processed.data.path, pattern="adu.coefs.r-[0-9]+.csv")
-chi.r.files <- dir(processed.data.path, pattern="chi.coefs.r-[0-9]+.csv")
+adu.r.i.files <- dir(model.path, pattern="adu.max.r.i-[0-9]+.csv$")
+adu.r.m.files <- dir(model.path, pattern="adu.max.r.m-[0-9]+.csv$")
+chi.r.i.files <- dir(model.path, pattern="chi.max.r.i-[0-9]+.csv$")
+chi.r.m.files <- dir(model.path, pattern="chi.max.r.m-[0-9]+.csv$")
 
-adu.r.model <- read.csv(paste(processed.data.path, adu.r.files[1], sep=""))
-chi.r.model <- read.csv(paste(processed.data.path, chi.r.files[1], sep=""))
-adu.model <- data.table(read.csv(paste(processed.data.path, pattern="adu.coefs-0.csv", sep="")))
-chi.model <- data.table(read.csv(paste(processed.data.path, pattern="chi.coefs-0.csv", sep="")))
+adu.model.i <- read.csv(paste(model.path, "adu.max.i-0.csv", sep=""))
+adu.model.m <- read.csv(paste(model.path, "adu.max.m-0.csv", sep=""))
+chi.model.i <- read.csv(paste(model.path, "chi.max.i-0.csv", sep=""))
+chi.model.m <- read.csv(paste(model.path, "chi.max.m-0.csv", sep=""))
 
-# Initialize an empty data table
-aduN <- length(adu.r.files)+1
-chiN <- length(chi.r.files)+1
+### Initialize empty data tables
+# Model output tables
+aduvars <- as.character(adu.model.m$Predictor)
+chivars <- as.character(chi.model.m$Predictor)
+aruns <- c(0:length(adu.r.m.files))
+cruns <- c(0:length(chi.r.m.files))
+aduB <- data.table(Run = aruns)
[…]

### 2015-11-28 analysis/E1/_RUNME.R

@@ -5,10 +5,39 @@ source("2-find.switches.R")
 get.switches(1, FALSE)
 
 # Find switches in randomly-shuffled gap versions
-# rand.runs <- 1:5000
-# make.rand.gaps(rand.runs)
-# get.switches(rand.runs, TRUE)
+rand.runs <- 301:5000
+make.rand.gaps(rand.runs)
+get.switches(rand.runs, TRUE)
 
 source("3-plots.stats.analysis.R")
 source("4-model-functions.R")
 run.models(rand.runs)
+
+compare.models()
+
+#### BENCHMARKING LIBRARIES
+# library(devtools)
+# devtools::install_github("hadley/lineprof")
+# library(lineprof)
+# library(microbenchmark)
+
+# ###########
+# l <- lineprof(get.switches(1, FALSE))
+# microbenchmark(get.switches(1, FALSE), times = 1)
+
+# ########### Middy did not try to fix parallelization yet
+# library(foreach)
+# library(doParallel)
+
+# start <- Sys.time()
+
+# #setup parallel backend 
+# cl<-makeCluster(4, method="FORK")
+# registerDoParallel(cl, cores = 4)
+
+# foreach (i = 1:4, .packages = "dplyr") %dopar% {
+  # make.replicate(all.data.A, all.data.C, i, info.path, processed.data.path) 
+# }
+
+# stopCluster(cl)
+# stoptime <- Sys.time() - start

### 2015-11-28 analysis/E2/0-helper.R

@@ -12,6 +12,10 @@ library(car)
 ################################################################################
 # Misc helper functions (from Mike)
 ################################################################################
+# Stat summary functions
+tstat <- function(m) fixef(m)/sqrt(diag(vcov(m)))
+SEstat <- function(m) sqrt(diag(vcov(m)))
+
 # Add some style elements for ggplot2
 plot.style <- theme(panel.grid.major = element_blank(),
     panel.grid.minor = element_blank(),
@@ -171,89 +175,159 @@ smoothLD <- function(dt) {
 ################################################################################
 # Random run plot functions
 ################################################################################
-
-combine.runs <- function(DT, filelist, name) {
-	curr.row <- 1
-	for (file in filelist) {
-		# Read in a file
-		data <- fread(paste(processed.data.path, file, sep=""))
-		# Add the data to the big data table
-		DT[curr.row, names(DT) := data]
-		curr.row <- curr.row + 1
+combine.runs <- function(betas, ses, ts, fits,
+	modelfiles, infofiles, realm, reali) {
+	# Random run info
+	for (file in modelfiles) {
+		data <- fread(paste(model.path, file, sep=""))
+		run <- as.numeric(unlist(strsplit(
+			file, "[.-]"))[5])
+		# betas
+		insert <- data.frame(t(data$B))
+		colnames(insert) <- data$Predictor
+		betas[Run == run, (data$Predictor) := insert]
+		# ses
+		insert <- data.frame(t(data$SE))
+		colnames(insert) <- data$Predictor
+		ses[Run == run, (data$Predictor) := insert]
+		# ts
+		insert <- data.frame(t(data$t))
+		colnames(insert) <- data$Predictor
+		ts[Run == run, (data$Predictor) := insert]
 	}
-	setkeyv(DT, "Run")
-	# Write it out
-	write.csv(DT, paste(processed.data.path,
-		name, ".random.runs.csv",sep=""), row.names=FALSE)
-	# Return it
-	return(DT)
+	for (file in infofiles) {
+		data <- fread(paste(model.path, file, sep=""))
+		run <- as.numeric(unlist(strsplit(
+			file, "[.-]"))[5])
[…]

### 2015-11-28 analysis/E2/3-plots.stats.analysis.R

@@ -14,6 +14,7 @@ library(data.table)
 raw.data.path <- "tracker_data/" #"../../data/E1/tracker_data/"
 info.path <- "info/" # "../../data/E1/info/"
 processed.data.path <- "processed_data/" #"../../data/E1/processed_data/"
+model.path <- "processed_data/models/"
 plot.path <- "plots/" #"plots/"
 
 # Read in supplemental data and prepped data
@@ -161,6 +162,74 @@ dev.off()
 ################################################################################
 
 
+###### DIFF BETWEEN RANDOM AND TRANSITION w/ TYPE  #############################
+mean.agg <- aggregate(Switch ~ Condition + Sample + Type + Age, switch.all, mean)
+langs <- mean.agg$Condition
+ages <- mean.agg$Age
+ages[ages == "21"] <- "A"
+samples <- mean.agg$Sample
+types <- mean.agg$Type
+means <- mean.agg$Switch
+errs <- aggregate(Switch ~ Condition + Sample + Type + Age, switch.all, sem)$Switch
+linecolors <- rep(c(
+	"Random (all)", "Random (all)",
+	"Random (all)", "Random (all)",
+	"No speech", "Prosody only",
+	"Normal speech", "Words only"),14)
+
+df <- data.frame(
+    Age = factor(ages),
+    Sample = factor(samples),
+    Type = factor(types),
+    Condition = factor(langs),
+    m = means,
+    se = errs,
+    ConditionL = factor(linecolors)
+)
+df$Sample <- factor(df$Sample, levels=rev(levels(df$Sample)))
+df$Condition <- factor(df$Condition, labels=c("No speech", "Prosody only",
+    "Normal speech", "Words only"))
+df$Type <- factor(df$Type, labels=c("Question", "Non-Question"))
+df$Condition <- factor(df$Condition, levels=c("Normal speech", "Prosody only",
+    "Words only", "No speech"))
+df$ConditionL <- factor(df$ConditionL, levels=c("Normal speech", "Prosody only",
+    "Words only", "No speech", "Random (all)"))
+
+limits <- aes(ymax = m + se, ymin= m - se)
+dodge <- position_dodge(width=0.9)
+
+p2 <- qplot(Age,m,facets = . ~ Condition, group=interaction(Sample,Type),
[…]

### 2015-11-28 analysis/E2/4-model-functions.R

@@ -11,11 +11,10 @@ library(data.table)
 # Read in data in processed_data/ path and set plotting variables
 ################################################################################
 processed.data.path <- "processed_data/" #"../../data/E1/processed_data/"
+model.path <- "processed_data/models/"
 switch.all <- fread(paste(processed.data.path,"SwitchAll.csv",sep=""))
 chi.coef <- fread(paste(processed.data.path,"chi.coefs-0.csv",sep=""))
-chi.coef.l <- fread(paste(processed.data.path,"chi.coefs.l-0.csv",sep=""))
 adu.coef <- fread(paste(processed.data.path,"adu.coefs-0.csv",sep=""))
-adu.coef.l <- fread(paste(processed.data.path,"adu.coefs.l-0.csv",sep=""))
 
 run.models <- function(ns) {
     # Split the data into adults and children for separate models
@@ -23,9 +22,6 @@ run.models <- function(ns) {
                    select(-Sample) %>%
 				   mutate(Condition = as.factor(Condition),
 		   		          Type = as.factor(Type))
-
-	switch.rand$Cond2 <- ifelse((switch.rand$Condition == "normal" |
-		switch.rand$Condition == "robot"),"lexical", "non-lexical")
 
 	switch.rand.C <- filter(switch.rand, Age != 21) %>%
          			        mutate(Age = as.numeric(Age))
@@ -40,49 +36,38 @@ run.models <- function(ns) {
 
     		# Retrieve the random gap info data for this particular run
 	    	rand.data.C <- switch.rand.C[Run == i,]
-	    	chi.coef.r <- chi.coef[0,]
 
-    		# Models
-	    	# Children
 			chi.max.r <- glmer(Switch ~ Age * Condition * Type + Duration + 
  			   (Type|Subject) + (1|Gap), data=rand.data.C,
  			   family=binomial, control = glmerControl(optimizer="bobyqa"))
-    		chi.coef.r <- rbind(chi.coef.r,
-    			cbind(data.frame(t(sapply(fixef(chi.max.r),c))),
-    			data.frame(Sample = "random", Error = errors, Run = i)))
-    		write.csv(chi.coef.r, paste(
-				processed.data.path, "chi.coefs.r-", i, ".csv", sep=""),
[…]

### 2015-11-28 analysis/E2/5-compare.R

@@ -6,119 +6,686 @@ source("0-helper.R")
 ################################################################################
 # Set paths
 processed.data.path <- "processed_data/" #"../../data/E1/processed_data/"
+model.path <- "processed_data/models/" #"../../data/E1/processed_data/"
 plot.path <- "plots/" #"plots/"
 
 # Read in all the random runs and combine them
-adu.r.files <- dir(processed.data.path, pattern="adu.coefs.r-[0-9]+.csv")
-chi.r.files <- dir(processed.data.path, pattern="chi.coefs.r-[0-9]+.csv")
-adu.l.r.files <- dir(processed.data.path, pattern="adu.coefs.l.r-[0-9]+.csv")
-chi.l.r.files <- dir(processed.data.path, pattern="chi.coefs.l.r-[0-9]+.csv")
+adu.r.i.files <- dir(model.path, pattern="adu.max.r.i-[0-9]+.csv$")
+adu.r.m.files <- dir(model.path, pattern="adu.max.r.m-[0-9]+.csv$")
+chi.r.i.files <- dir(model.path, pattern="chi.max.r.i-[0-9]+.csv$")
+chi.r.m.files <- dir(model.path, pattern="chi.max.r.m-[0-9]+.csv$")
 
-adu.r.model <- read.csv(paste(processed.data.path, adu.r.files[1], sep=""))
-chi.r.model <- read.csv(paste(processed.data.path, chi.r.files[1], sep=""))
-adu.model <- data.table(read.csv(paste(processed.data.path, pattern="adu.coefs-0.csv", sep="")))
-chi.model <- data.table(read.csv(paste(processed.data.path, pattern="chi.coefs-0.csv", sep="")))
+adu.model.i <- read.csv(paste(model.path, "adu.max.i-0.csv", sep=""))
+adu.model.m <- read.csv(paste(model.path, "adu.max.m-0.csv", sep=""))
+chi.model.i <- read.csv(paste(model.path, "chi.max.i-0.csv", sep=""))
+chi.model.m <- read.csv(paste(model.path, "chi.max.m-0.csv", sep=""))
 
-# Initialize an empty data table
-aduN <- length(adu.r.files)+1
-chiN <- length(chi.r.files)+1
+### Initialize empty data tables
+# Model output tables
+aduvars <- as.character(adu.model.m$Predictor)
[…]

### 2015-11-28 analysis/E2/Extras.R

@@ -1,0 +1,98 @@
+switch.rand <- filter(switch.all, Sample != "Transition")
+switch.real <- filter(switch.all, Sample == "Transition") %>%
+               select(-Sample)
+switch.real2 <- filter(switch.all, Sample == "Transition")
+randswitchrate <- aggregate(Switch ~ Run, switch.rand, mean)
+realswitchrate <- aggregate(Switch ~ Run, switch.real2, mean)
+switch.rand.C <- filter(switch.rand, Age != "A") %>%
+                 mutate(Age = as.numeric(Age))
+switch.rand.A <- filter(switch.rand, Age == "A")
+
+
+ggplot(randswitchrate, aes(x=Switch)) + geom_density() +
+	geom_point(data=realswitchrate, aes(x=Switch, y=0), size=4, color="black") +
+	geom_point(data=realswitchrate, aes(x=Switch, y=0), size=3, color="red") +
+	ggtitle("Anticipatory switches overall\n")
+
+# Set paths
+processed.data.path <- "processed_data/" #"../../data/E1/processed_data/"
+plot.path <- "plots/" #"plots/"
+
+# Read in all the random runs and combine them
+adu.r.files <- dir(processed.data.path, pattern="adu.coefs.r-[0-9]+.csv")
+chi.r.files <- dir(processed.data.path, pattern="chi.coefs.r-[0-9]+.csv")
+
+adu.r.model <- read.csv(paste(processed.data.path, adu.r.files[1], sep=""))
+chi.r.model <- read.csv(paste(processed.data.path, chi.r.files[1], sep=""))
+adu.model <- data.table(read.csv(paste(processed.data.path, pattern="adu.coefs-0.csv", sep="")))
+chi.model <- data.table(read.csv(paste(processed.data.path, pattern="chi.coefs-0.csv", sep="")))
+
+# Initialize an empty data table
+aduN <- length(adu.r.files)+1
+chiN <- length(chi.r.files)+1
+
+aduDT <- data.table(
+	matrix(0, aduN, length(adu.r.model),
+	dimnames=list(c(), colnames(adu.r.model))))
+aduDT$Sample <- rep("", aduN)
+aduDT$Error <- rep("", aduN)
+
[…]

### 2015-11-28 analysis/E2/_RUNME.R

@@ -5,9 +5,10 @@ source("2-find.switches.R")
 get.switches(1, FALSE)
 
 # Find switches in randomly-shuffled gap versions
-# rand.runs <- 1:5000
-# make.rand.gaps(rand.runs)
-# get.switches(rand.runs, TRUE)
+rand.runs <- 1:5000
+make.rand.gaps(rand.runs)
+get.switches(rand.runs, TRUE)
 
 source("3-plots.stats.analysis.R")
 source("4-model-functions.R")
+compare.models()

### 2015-11-28 paper/CasillasFrank-Anticip.tex
E2 summary draft added

@@
 The resulting conversations were therefore not entirely spontaneous, but were as close as possible while still remaining child-oriented in topic, prosodic pattern, and lexicosyntactic construction.\footnote{All of the non-English talkers were fluent in English as a second language, and some fluently spoke a third or more language.
-We chose male-female pairs as a natural way of creating contrast between the two talker voices.
-See an example run-though of the videos here: }
+We chose male-female pairs as a natural way of creating contrast between the two talker voices.}
 
@@
 Despite this, their looks to the addressee did not increase substantially in the non-English videos, indicating that their looks away were probably related to boredom rather than confusion about ongoing turn structure.
-Overall, participants' pattern of gaze to current speakers indicated that they performed basic turn tracking during the videos, regardless of language.
+Overall, participants' pattern of gaze to current speakers demonstrated that they performed basic turn tracking during the videos, regardless of language.
 
@@
 We included the duration of the inter-turn gap as a predictor since longer gaps also provide more opportunities to make anticipatory switches (Figure \ref{fig:criterion}).
-We additionally included random effects of item (turn transition) and participant, with random slopes of condition, transition type, and their interaction for participants \citep{barr2013}.\footnote{The models we report are all qualitatively \textit{unchanged} by the exclusion of their random slopes.
+We additionally included random effects of item (turn transition) and participant, with random slopes of condition, transition type, and their interaction for participants \citep{barr2013}.\footnote{The models we report are all qualitatively unchanged by the exclusion of their random slopes.
 We have left the random slopes in because of minor participant-level variation in the predictors modeled.} The adult model included condition, transition type, duration, and their interactions as predictors with participant and item included as random effects and random slopes of condition, transition type, and their interaction for participant.
@@
 We estimated the probability that these patterns were the result of random looking by running the same regression models on participants' real eye-tracking data, only this time calculating their anticipatory gaze switches with respect to randomly permuted turn transition windows.
-This process involved: (1) randomizing the order and temporal placement of the analysis windows within each stimulus (Figure \ref{fig:shuffling}; the analysis window is defined in Figure \ref{fig:criterion}), thereby randomly redistributing the analysis windows across the eye-tracking signal, (2) re-running each participant's eye tracking data through switch identification (described in \ref{sec:algorithm}), but this time using the randomly permuted analysis windows, and (3) modeling the anticipatory gazes from the randomly permuted data with the same statistical models we used for the original data (Section \ref{sec:models1}; Table \ref{tab:E1-models}).
-Importantly, although the onset time of each transition was shuffled within the eye-tracking signal, the other intrinsic properties of each turn transition (e.g., prior speaker identity, transition type, gap duration, language condition, etc.) stayed constant across each randomly permuted version of the data.
+This process involved: (1) randomizing the order and temporal placement of the analysis windows within each stimulus (Figure \ref{fig:shuffling}; ``analysis window'' is defined in Figure \ref{fig:criterion}), thereby randomly redistributing the analysis windows across the eye-tracking signal, (2) re-running each participant's eye tracking data through switch identification (described in \ref{sec:algorithm}), this time using the randomly permuted analysis windows, and (3) modeling the anticipatory gazes from the randomly permuted data with the same statistical models we used for the original data (Section \ref{sec:models1}; Table \ref{tab:E1-models}).
+Importantly, although the onset time of each transition was shuffled within the eye-tracking signal, the other intrinsic properties of each turn transition (e.g., prior speaker identity, transition type, gap duration, language condition, etc.) stayed constant across each random permutation.
 
@@
 \paragraph{Developmental effects}
-The model of the children's data revealed a significant interaction of age and language condition (tab:E1-models) that was highly unlikely to have derived from random looking (Figure \ref{fig:E1-randvsreal}).
+The model of the children's data revealed a significant interaction of age and language condition (Table \ref{tab:E1-models}) that was highly unlikely to have derived from random looking (Figure \ref{fig:E1-randvsreal}).
 To further explore this effect, we compared the average effect of language condition across all ages: we extracted the average difference score for the two language conditions (English minus non-English) for each subject, computing an overall average for each random permutation of the data.
@@
 
-Children and adults spontaneously tracked the turn structure of the conversations, making anticipatory gaze switches at an above-chance rate across all ages and conditions (Table \ref{tab:e1_look} ; Figure \ref{fig:E1-randvsreal}).
+Children and adults spontaneously tracked the turn structure of the conversations, making anticipatory gaze switches at an above-chance rate across all ages and conditions (Table \ref{tab:e1_look}; Figure \ref{fig:E1-randvsreal}).
 Children's anticipatory gaze rates were affected by language condition, transition type, age, and gap duration (Table \ref{tab:E1-models}), none of which could be explained by a baseline of random gaze switching (Figure \ref{fig:E1-ChiTs}).
@@
 This effect suggests that lexical access is important for children's ability to anticipate upcoming turn structure; children had no lexical access to the speech in the non-English videos, though they did have access to (non-native) prosodic cues and non-verbal behavior, consistent with prior work on turn-end prediction in adults (\citealp{de-ruiter2006}; \citealp{magyari2012}) and children \citep{keitel2013}.
-Second, children systematically made more anticipatory switches after hearing a question compared to a non-question, but only in the English condition, suggesting that, when children have access to lexical cues, they are more likely to make an anticipatory gaze switch when they can expect an immediate response from the addressee (i.e., an answer).
-In itself, this explanation also suggests that children are picking up on lexical cues to questionhood in anticipating upcoming answers (e.g., subject-auxiliary inversion, \textit{wh}-words, etc.).
+Second, children systematically made more anticipatory switches after hearing a question compared to a non-question, but only in the English condition, suggesting that, when children have access to lexical cues, they are more likely to make an anticipatory gaze switch if they can expect an immediate response from the addressee.
+If so, then children's attention to lexical cues for turn taking may primarily be in monitoring the signal for cues to questionhood (e.g., subject-auxiliary inversion, \textit{wh}-words, etc.).
 
@@
 This interaction suggests that the 5-year-olds were able to leverage anticipatory cues in the non-English videos in a way that 3- and 4-year-olds could not, possibly by shifting more attention to the non-native prosodic or non-verbal cues.
-Prior work on children's turn-structure anticipation proposed that children's turn-end predictions center on lexicosyntactic structure (and not, e.g., prosody) as they get older.
-The current results suggest more flexibility in children's predictions; when they do not have access to linguistic structure they are more likely to rely on alternative cues as they get older.
+Prior work on children's turn-structure anticipation proposed that children's turn-end predictions rely primarily on lexicosyntactic structure (and not, e.g., prosody) as they get older \citep{keitel2013}.
+The current results suggest more flexibility in children's predictions; when they do not have access to lexical information, older children are likely to find alternative cues to turn taking behavior.
 
@@
 Like children, adults made more anticipatory switches after hearing questions compared to non-questions, suggesting that anticipation mattered more to them when an immediate response was expected.
-Also like children, the advantage for questions was driven by lexical access such that adults must have relied on lexicosyntactic cues to questionhood in picking out turns that potentially require an immediate response, though this effect was only marginally significant compared to the distribution of randomly permuted data (\textit{p} $=$ .053; Figure \ref{fig:E1-AduTs}).
+Also like children, the advantage for questions was driven by lexical access such that adults must have relied on lexicosyntactic cues to questionhood in picking out turns that potentially require an immediate response, though this effect was only marginally divergent from the distribution of randomly permuted data (\textit{p} $=$ .053; Figure \ref{fig:E1-AduTs}).
 Finally, adults' anticipation rates were also affected by gap duration, but more so for questions than non-questions (Table \ref{tab:E1-models}).
@@
 
-Children and adults' predictions alike were benefited by access to lexical information (English) and speech act status (questionhood), suggesting that linguistic cues, particularly lexical ones, facilitate their spontaneous predictions about upcoming turn structure through the identification of turns that will have immediate responses from the addressee.
+Children and adults' predictions alike were benefited by access to lexical information (English) and speech act status (questionhood), suggesting that linguistic cues, particularly lexical ones, facilitate their spontaneous predictions about upcoming turn structure through the identification of turns with immediate responses.
 Children's anticipatory gaze rates for questions and non-questions in English was stable across ages and comparable to adult behavior (Figure \ref{fig:E1-randvsreal}), suggesting that they can identify questions in native stimuli with adult-like competence by age three.
@@
 non-English), the prosody in the non-English videos was non-native, and so the experimental design can not conclusively show which linguistic cues children relied on in the English videos to identify question turns.
-
-Relatedly, though lexical access clearly facilitated participants' anticipatory gaze rate, it was not necessary for participants, especially adults, in order to exceed chance switching rates (Figure \ref{fig:E1-randvsreal}), suggesting that participants use non-lexical cues (e.g., prosody, non-verbal behavior) to make anticipatory eye movements at least some of the time.
-In the case of adults there was no overall effect of language condition, only an interaction between language condition and transition type: adults spontaneously anticipated upcoming speakers whether they had lexical access or not, but when given lexical access, focused their anticipations on turns with immediate responses (questions).
+Relatedly, though lexical access clearly facilitated participants' anticipatory gaze rate, it was not necessary for participants---especially adults---in order to exceed chance switching rates (Figure \ref{fig:E1-randvsreal}), suggesting that participants use non-lexical cues (e.g., prosody, non-verbal behavior) to make anticipatory eye movements at least some of the time.
 
 Interestingly, adults and children both were strongly affected by transition type, in that they made more anticipatory switches after hearing questions, compared to non-questions.
-We had hypothesized that a projected response might affect participants' predictions but it in fact seems to be driving many of the effects in our data.
 Even in the English videos, when participants had full access to linguistic cues, their rates of anticipation were relatively low---in fact, comparable to the non-English videos---unless the turn was a question (Figure \ref{fig:E1-randvsreal}).
 Prior work using online, metalinguistic tasks has shows that participants can use linguistic cues to accurately predict upcoming turn ends.
-The current results suggest that, in their spontaneous predictions, both children and adults monitor the linguistic structure of unfolding turns for cues to upcoming responses.
-Because the form of an answer is partially determined by its question (e.g., locative expressions for ``where'' questions), participants could even be predicting what type of response they will hear, and so they are looking anticipatorily at the responder for confirmation of their prediction.
-This interpretation is initial, but in-line with work showing early response planning and speech act recognition \citep{gisladottir2015, bogels2015}.
-
-Overall, children and adults behaved relatively similarly and our language manipulation (English vs.
+The current results suggest that, in their spontaneous predictions about third-party conversation, both children and adults monitor the linguistic structure of unfolding turns for cues to upcoming responses.
+
+Children and adults generally behaved relatively similarly in this first experiment and our language manipulation (English vs.
 non-English) was too coarse to comment on when children begin to use different types of native linguistic cues (e.g., prosody vs.
 lexicosyntax); we would instead need to directly compare lexicosyntactic and prosodic cues in the participants' native language, controlling for the presence of non-verbal cues.
-To see the emergence of anticipatory gaze switching we would also need to include younger children since participants already reliably made anticipatory gaze switches at age three.
-In sum, Experiment 1 lays the analytic groundwork for a method that allows for greater experimental control, which we introduce in Experiment 2.
+To see the emergence of anticipatory gaze switching we would also need to include younger children, since participants already reliably made anticipatory gaze switches at age three.
+Experiment 1 thus lays the analytic groundwork for a method that allows for greater experimental control, which we introduce in Experiment 2.
 
@@
 We again built separate models for children and adults because effects of age were only pertinent to the children's data.
-The child model included condition (normal/prosody only/words only/no speech; with no speech in the intercept), transition type (question vs.
+The child model included condition (normal/prosody only/words only/no speech; with no speech as the reference level), transition type (question vs.
 non-question), age (1, 2, 3, 4, 5, 6), and duration of the inter-turn gap (in seconds) as predictors, with full interactions between language condition, transition type, and age.
@@
 
-Children's anticipatory gaze switches showed an effect of gap duration (\textit{$\beta$}=4.18, \textit{SE}=0.624, \textit{t}=6.689, \textit{p}$<$.001), a two-way interaction of age and language condition (for prosody only speech compared to the no speech intercept; \textit{$\beta$}=0.393, \textit{SE}=0.189, \textit{t}=2.08, \textit{p}$<$.05), and a three-way interaction of age, transition type, and language condition (for normal speech compared to the no speech intercept; \textit{$\beta$}=-0.375, \textit{SE}=0.17, \textit{t}=-2.213, \textit{p}$<$.05).
-There were no significant effects of age or transition type alone (Table \ref{sec:models}), with only a marginal effect of language condition (for prosody only compared to the no speech intercept; \textit{$\beta$}=-1.634, \textit{SE}=0.864, \textit{t}=-1.89, \textit{p}$=$.06)
+Children's anticipatory gaze switches showed an effect of gap duration (\textit{$\beta$}=4.18, \textit{SE}=0.624, \textit{t}=6.689, \textit{p}$<$.001), a two-way interaction of age and language condition (for prosody only speech compared to the no speech reference level; \textit{$\beta$}=0.393, \textit{SE}=0.189, \textit{t}=2.08, \textit{p}$<$.05), and a three-way interaction of age, transition type, and language condition (for normal speech compared to the no speech reference level; \textit{$\beta$}=-0.375, \textit{SE}=0.17, \textit{t}=-2.213, \textit{p}$<$.05).
+There were no significant effects of age or transition type alone (Table \ref{sec:models2}), with only a marginal effect of language condition (for prosody only compared to the no speech reference level; \textit{$\beta$}=-1.634, \textit{SE}=0.864, \textit{t}=-1.89, \textit{p}$=$.06)
 
 Adults' anticipatory gaze switches showed effects of gap duration (\textit{$\beta$}=4.75, \textit{SE}=1.248, \textit{t}=3.806, \textit{p}$<$.001) and language condition (for normal speech \textit{$\beta$}=1.256, \textit{SE}=0.563, \textit{t}=2.229, \textit{p}$<$05.
-and words only speech \textit{$\beta$}=1.594, \textit{SE}=0.721, \textit{t}=2.211, \textit{p}$<$.05  compared to the no speech intercept).
+and words only speech \textit{$\beta$}=1.594, \textit{SE}=0.721, \textit{t}=2.211, \textit{p}$<$.05  compared to the no speech reference level).
 There were no effects of transition type (\textit{$\beta$}=0.429, \textit{SE}=0.609, \textit{t}=0.705, \textit{p}$=$.48).
@@
 Our baseline analyses revealed that none of the significant predictors from models of the original, turn-related data (Table \ref{tab:E2-models}: Children) can be explained by random looking.
-In the children's data, the original \textit{t}-values for language condition (prosody only), gap duration, the two-way interaction of age and language condition (prosody only) and the three-way interaction of age, transition type, and language condition (normal only) were all greater than 95\% of the randomly permuted \textit{t}-value (96.2\%, 100\%, 95.1\%, and 95.1\%, respectively; Figure \ref{fig:E2-ChiTs}).
-Similarly, the adults' data showed significant differentiation from the randomly permuted data for all originally significant predictors: language condition for normal and words only speech and gap duration (greater than 96.8\%, 98.7\%, and 100\% of random \textit{t}-values, respectively; Figure \ref{fig:E2-AduTs}).
+In the children's data, the original model's \textit{t}-values for language condition (prosody only), gap duration, the two-way interaction of age and language condition (prosody only) and the three-way interaction of age, transition type, and language condition (normal speech) were all greater than 95\% of the randomly permuted \textit{t}-values (96.2\%, 100\%, 95.1\%, and 95.1\%, respectively; Figure \ref{fig:E2-ChiTs}).
+Similarly, the adults' data showed significant differentiation from the randomly permuted data for all originally significant predictors: gap duration and language condition for normal speech and words-only speech (greater than 100\%, 96.8\%, and 98.7\% of random \textit{t}-values, respectively; Figure \ref{fig:E2-AduTs}).
 The effects of language condition and transition type for the real and randomly permuted data can also be observed in Figure \ref{fig:E1-randvsreal}.
@@
 \paragraph{Developmental effects}
-The model of the children's data revealed two significant interactions with age, neither of which derived from random looking (Figure \ref{fig:E2-randvsreal}).
-The first was a significant interaction of age and language condition (for prosody only compared to the no speech baseline), suggesting a different age effect between the two linguistic conditions.
-As in Experiment 1, we further explored the age effect by extracting the average difference score over subjects for the interaction in each random permutation of the data, making pairwise comparisons between the six age groups.
+The model of the children's data revealed two significant interactions with age, neither of which derived from random looking (Table \ref{tab:E2-models}; Figure \ref{fig:E2-randvsreal}).
+The first was a significant interaction of age and language condition (for prosody only compared to the no speech reference level), suggesting a different age effect between the two linguistic conditions.
+As in Experiment 1, we further explored the age effect by extracting the average difference score over subjects for this age-language condition interaction in each random permutation of the data, making pairwise comparisons between the six age groups.
 Figure \ref{fig:E2-lgageinteraction} shows the comparisons that demonstrate a significant difference in age for the no speech baseline vs.
-the prosody only condition, revealing that anticipatory gaze in the prosody only condition significantly improves with age, especially at ages 5 and 6, compared to the no-speech baseline (all with difference scores greater than 95\% of the random data scores; \textit{p} $<$ .05).
+the prosody only condition, revealing that anticipatory gaze in the prosody only condition significantly improves with age, especially at ages 5 and 6 (all with difference scores greater than 95\% of the random data scores; \textit{p} $<$ .05).
 
 The second interaction with age was a three-way interaction of age, transition type, and language condition (for normal speech compared to the no speech baseline).
-We again created pairwise comparisons of the average difference scores for the interaction in each random permutation of the data, with significant differences shown in Figure \ref{fig:E2-lgagetypeinteraction}.
-These pairwise comparisons showed that the effect of transition type in the normal speech condition became larger with age, with significant improvements by age 4 over ages 1 and 2 (99.9\% and 98.86\%, respectively), by age 5 over age 4 (97.54\%), and by age 6 over ages 1, 2, and 5 (99.5\%, 97.36\%, and 95.04\%), all equating to significant differences from chance (\textit{p} $<$ .05).
+We again created pairwise comparisons of the average difference scores for this age-transition type-language condition interaction in each random permutation of the data, with significant differences shown in Figure \ref{fig:E2-lgagetypeinteraction}.
+These pairwise comparisons showed that the effect of transition type in the normal speech condition became larger with age, with significant improvements by age 4 over ages 1 and 2 (99.9\% and 98.86\%, respectively), by age 5 over age 4 (97.54\%), and by age 6 over ages 1, 2, and 5 (99.5\%, 97.36\%, and 95.04\%), all significantly different from chance (\textit{p} $<$ .05).
 
@@
 \label{sec:discussion2}
+As in Experiment 1, children and adults spontaneously tracked the turn structure of the conversations, making anticipatory gaze switches at an above-chance rate across all ages but not in all conditions (Table \ref{tab:e2_look}; Figure \ref{fig:E2-randvsreal}).
+In the normal speech condition, however, when children had access to full linguistic information, they made more anticipatory gaze switches than expected by chance even at age one.
+Children's anticipatory gaze rates were affected by gap duration, plus interactions of language condition, transition type and age (Table \ref{tab:E2-models}), none of which could be explained by a baseline of random gaze switching (Figure \ref{fig:E2-ChiTs}).
+
+Two of the three linguistic conditions affected children's anticipatory gaze switches, compared to the no-speech reference level: prosody-only speech and normal speech.
+Prosody only speech resulted in fewer anticipatory gazes overall, compared to the no-speech baseline, though the effect was only marginal (\textit{p}$=$.06).
+However, prosody-only speech \textit{did} significantly differ from the no speech condition in its interaction with age: whereas age gains were negligible in the no speech condition, 5- and 6-year-olds in the prosody-only condition showed significantly more anticipatory gaze switches than younger children (Figure \ref{fig:E2-lgageinteraction}), going from below- and at-chance anticipatory gaze rates to being well above chance (Figure \ref{fig:E2-randvsreal}).
+
+The normal speech condition showed a significantly more gains in anticipatory gaze for questions with age compared to the no-speech condition.
+As in Experiment 1, children consistently made more anticipatory gazes after hearing questions when they had access to lexical material.
+But now, with a larger age span in Experiment 2, we can begin to see a developmental path for the question effect.
+At least for normal speech, greater anticipatory looking for questions is not present from the start: 4-year-olds are the first to make significant gains on 1- and 2-year-olds, but then there are further significant gains at age 5, and again age 6 (Figure \ref{fig:E2-lgagetypeinteraction}).
+While children at ages five and six showed adult-like differentiation of questions and non-questions in the normal speech condition, 1-year-olds had nearly identical switch rates for the two transition types.
+This suggests that the participants' tendency to make more anticipatory switches for questions emerges after age one and continues developing, with rapid gains in ages three through age six.
+Finally, children showed a straightforward effect of gap duration (Table \ref{tab:E2-models}), as in Experiment 1.
+
+Adults' anticipatory gaze rates were affected by gap duration and two of the language conditions (Table \ref{tab:E2-models}), none of which could be explained by a baseline of random gaze switching (Figure \ref{fig:E2-AduTs}).
+Adults made more anticipatory switches overall for the normal speech and words-only conditions compared to the no-speech condition, falling in-line with past work showing that adults primarily use lexical information in making predictions about upcoming turn structure \citep{de-ruiter2006}.
+Though adults did make more anticipatory switches for questions than non-questions on average in the two lexical conditions (Figure \ref{fig:E2-randvsreal}), the effect of transition type was not significant in either (Table \ref{tab:E2-models}), unlike Experiment 1.
+Like children, adults also showed a straightforward effect of gap duration.
 
@@
 \label{sec:gendisc}
+
+% CAN USE LATER: Because the form of an answer is partially determined by its question (e.g., locative expressions for ``where'' questions), participants could even be predicting what type of response they will hear, and so they are looking anticipatorily at the responder for confirmation of their prediction.
+This interpretation is initial, but in-line with work showing early response planning and speech act recognition \citep{gisladottir2015, bogels2015}.
+
 %Children begin to develop conversational turn-taking skills long before their first words \citep{bateson1975, hilbrink2015, jaffe2001, snow1977}.

### 2015-11-28 paper/CasillasFrank-Anticip.tex
completed summary for E2

@@
 
-%We identified anticipatory gaze switches for all 79 usable turn transitions using the same criteria as in Experiment 1.
-We expected to see the most anticipatory switches in the \textit{normal} condition and the least anticipatory switches in the \textit{no speech} condition because they contained the most and least linguistic information, respectively.
-We expected to replicate our finding that questions result in more anticipatory switches than non-questions, with the added hypothesis that the question effect is driven by lexicosyntactic cues.
-We also anticipated an overall increase in anticipatory switches with age.
+Children and adults both showed more anticipatory gaze switches with increased linguistic information, but only for a subset of the linguistic conditions and transition types.
+We had expected to see the most anticipatory switches in the \textit{normal} condition and the least anticipatory switches in the \textit{no speech} condition because they contained the most and least linguistic information, respectively.
+We had also expected to replicate our finding from Experiment 1 that questions result in more anticipatory switches than non-questions, with the added hypothesis that the question effect is driven by lexicosyntactic cues.
+We additionally anticipated an overall increase in anticipatory switches with age.
 Finally, since the development of prosodic skills partially precedes the development of lexicosyntax, we expected to see more switches in the \textit{prosody only} condition compared to the \textit{words only} condition in the youngest age groups.
-%
-%We found that participants at all ages made anticipatory gaze switches more often than would be expected by chance in almost all conditions, but sometimes by a small margin (Figure \ref{fig:randvsrealEN}).
-For example, anticipatory switches in the \textit{no speech} videos were often modestly above chance (with the exception of 5- and 6-year-olds).
-One interpretation of this finding is that children do not use linguistic information to predict upcoming turns---the only cue to turn taking in the \textit{no speech} condition was the alternating mouth movements of the conversing puppets.
-With sufficiently long inter-turn gaps, participants could simply switch their gaze to the responder when the prior talker finishes talking---a strategy that doesn't rely on linguistic information.
-But this interpretation is not compatible with our the effects of linguistic information that do emerge in the data: An early advantage for prosody and a later advantage for lexical question cues.
-%
-%We saw a generalize early advantage for prosody over lexicosyntax in children's spontaneous turn predictions.
-Notably, while children performed well with \textit{prosody only} speech from the start, they seemed to only reliably exceed chance in the \textit{words only} condition at ages three and older.
-By age three, the general advantage for prosody over lexicosyntax seems to disappear with more equal rates of switching overall for the two partial linguistic conditions.
-%
-%\begin{figure}[t]
-%\begin{center}
-%\includegraphics[width=0.99\textwidth]{figures/FIG-randvsreal-EN.png}
-%\end{center}
-%\caption{Baseline-corrected proportion of anticipatory gaze switches made for actual (solid) and randomly-shuffled (dashed) turn transitions in each condition, and across ages.
-Error bars represent the standard error of the mean.}
-%\label{fig:randvsrealEN}
-%\end{figure}
-%
-%We observed that, overall, children and adults made the fewest anticipatory switches in the \textit{no speech} videos (9\%) and the most in the \textit{normal} videos (16\%), with the \textit{prosody only} and \textit{words only} videos falling in-between (14\% and 11\%, respectively; Figure \ref{fig:conditionsEN}).
-However, these condition-based differences were primarily driven by transition type effects (question vs.
-non-question).
-%
-%\begin{figure}[t]
-%\begin{center}
-%\includegraphics[width=0.65\textwidth]{figures/FIG-conditions-EN.png}
-%\end{center}
-%\caption{Baseline-corrected proportion of anticipatory gaze switches made for turn transitions in each speech condition, and across ages. Error bars represent the standard error of the mean.} 
-%\label{fig:conditionsEN}
-%\end{figure}
-%
-%As in Experiment 1, we made condition, age and transition type comparisons based on baselin-corrected switching values; we corrected each participant's gaze data by subtracting their random baseline from their actual switching for each turn transition.
-%
-%We found that participants made more anticipatory gaze switches following questions than non-questions, but only in the two lexical conditions (\textit{normal} and \textit{words only} speech).
-The advantage for questions was strongest in the \textit{normal} speech condition, where it emerged early in development.
-The question effect became larger with age in both lexical conditions.
-While children made small, sometimes non-linear, gains in their anticipatory switches for non-questions as they grew older, they made large gains in their anticipatory switches with age (especially between ages one and two; Figure \ref{fig:questionsEN}).
-The increased switching for questions between ages one to two indicates that, around this age, children begin to leverage lexicosyntactic cues in the speech signal to identify questions in ongoing conversation.
-In the lexical conditions, children had access to early cues to questionhood, such as \textit{wh}-words and subject-auxiliary inversion, to predict an upcoming speaker change.
-In the \textit{normal} condition, where the question effect was strongest, lexical and syntactic cues were reinforced by question prosody for polar questions.
-%
-%\begin{figure}[t]
-%\begin{center}
-%\includegraphics[width=0.99\textwidth]{figures/FIG-QvsNQ-EN.png}
-%\end{center}
-%\caption{Baseline-corrected proportion of anticipatory gaze switches made for question- (dashed) and non-question- (dotted) turn transitions in each condition, and across ages.
-Error bars represent the standard error of the mean.}
-%\label{fig:questionsEN}
-%\end{figure}
-%
-%Children showed increased anticipatory switching with age across all conditions, making almost as many transitions as adults by ages five and six.
-Adults reacted to the \textit{no speech} condition differently than the children did, tracking the interaction less closely than in the three linguistic conditions.
-They may have minimized switching in the \textit{no speech} condition because they realized that there was no conversation to follow.
-%
-%\subsubsection{Statistical analysis}
-%
-%We tested these findings in two linear mixed-effects models of participants' baseline-corrected anticipatory switches---one model each for children and adults.
-The child model included condition (\textit{normal}, \textit{words only}, \textit{prosody only}, and \textit{no speech}; \textit{normal} in the intercept), transition type (question vs.
-non-question), age, and gap duration as fixed effects, with full interactions between condition, transition type, and age.
-The model also included random effects of item (turn transition) and participant, with random slopes of condition and transition type for participants.
-The adult model matched the child model, excluding age-related effects.
-%
-%There was a significant effect of gap duration.
-Children had time to make more anticipatory switches in analysis windows with longer gaps (\textit{$\beta$}=0.294, \textit{SE}=0.091, \textit{t}=3.231, \textit{p}$<$.01).
-An effect of gap duration is expected since prior work using spontaneous anticipatory switching has found that most switches occur in the inter-turn gap (Keitel et al., 2013; Hirvenkari, 2013; Tice and Henetz, 2011).
-%
-%The only effects of linguistic condition derived from interactions with age and transition type, despite large differences in the linguistic information available across conditions.
-(Figure \ref{fig:questionsEN}).
-The model of children's data confirmed two three-way effects of age, condition, and transition type.
-Children made significantly more anticipatory switches for questions with age in the \textit{normal} (\textit{$\beta$}=0.04, \textit{SE}=0.008, \textit{t}=5.168, \textit{p}$<$.001) and \textit{words only} conditions (\textit{$\beta$}=0.031, \textit{SE}=0.01, \textit{t}=3.170, \textit{p}$<$.01).
-Overall, they made a switch twice as often after hearing a question (16\%) than a non-question (8\%) in the \textit{words only} condition, and nearly four times as often in the \textit{normal} condition (19\% vs.
-5\%).
-Since the question effect was limited to the lexical conditions, there was no effect of transition type.
-There were also no further interactions of age and condition, or of transition type and condition.
-There was, however, a significant two-way interaction of transition type and age (\textit{$\beta$}=-0.031, \textit{SE}=0.01, \textit{t}=-3.18, \textit{p}$<$.01) in addition to the three-way interactions for the \textit{normal} and \textit{words only} conditions already mentioned.
-%
-%Last but not least, we confirmed that children generally made more anticipatory switches as they got older: The model showed a significant effect of age (\textit{$\beta$}=0.028, \textit{SE}=0.009, \textit{t}=3.018, \textit{p}$<$.01).
-Though the increase in anticipatory looks with age came, in large part, from questions effects in the lexical conditions, children also showed increased switches for non-questions and non-lexical conditions.
-The increases related to non-questions and non-lexical conditions were overall smaller and sometimes followed a non-linear trajectory (Figure \ref{fig:questionsEN}).
-As in Experiment 1, early switches (planned before the end of the prior turn) did not show large increases with age, with adults making the fewest early switches in the three linguistic conditions (Figure \ref{fig:cumulativeEN}).
-%
-%\begin{figure}[ht]
-%\begin{center}
-%\includegraphics[width=0.99\textwidth]{figures/FIG-cumulative-EN.png}
-%\end{center}
-%\caption{Cumulative proportion of anticipatory gaze switches planned during the transition window, across conditions and ages.
-Age is indicated by hue (darker = older).
-The solid line is the offset of the pre-gap turn, and the dashed line is the median onset of the post-gap turn in each condition.
-``Early switches'' occur before 0.}
-%\label{fig:cumulativeEN}
-%\end{figure}
-%
-%The advantage for questions over non-questions in these results reinforces our findings from Experiment 1.
-They suggest that even two-year-olds can reliably distinguish questions from other transition types in third-party conversation, especially when full linguistic information is available.
-Inspecting the data more closely, we found that 43\% of one-year-olds made more anticipatory switches after questions than non-questions in the \textit{normal} condition, and 24\% of them did so in the \textit{words only} condition.
-Meanwhile, 61\% of two-year-olds show the effect in the \textit{normal} condition, and 52\% of them do in the \textit{words only} condition, which is already close to the average for older children and adults (Table \ref{tab:questioneffectexp2}).
-%
+
+In fact, children and adults did show more anticipatory gaze switching in the normal condition compared to the no-speech condition, but for children this effect only emerged in the form of anticipations following question turns, which increased with age faster than it did with no linguistic cues at all (i.e., in the no-speech condition).
+Adults also showed significantly higher anticipatory switch rates in the words only condition but no effects of transition type, alone or within linguistic conditions.
+Taken together, these results partly replicate the findings from Experiment 1: participants make more anticipatory switches when they have access to lexical information and, when they do, tend to make more anticipatory switches for questions compared to non-questions.
+
+We had anticipated significant gains in anticipatory switching with age, but children only showed significant developmental increases in the prosody only condition and the normal condition (for question transitions).
+Rather than showing an early advantage for prosody over lexical information, children did not show significant improvement (or even above-chance performance) until age five.
+In contrast, their anticipatory gaze switches were, on average, already above chance at age one in the normal condition (with both lexical and prosodic information) and at age three in the words only condition (with lexical information).
+These findings do not support an early role for prosody in children's spontaneous predictions about upcoming turn structure.
+On the contrary, their predictions were best when lexical information was present, especially when following a question.
+
+However, these results do not support the idea that lexical information is sufficient for children (as is proposed for adults; \citealp{de-ruiter2006}).
+On the contrary, children showed significant gains in the normal speech condition, but \textit{not} in the words only condition.
+Notably, the normal speech condition, in addition to having both lexical and prosodic information, is also the one most likely to occur in our participants' daily lives (compared to muffled or robotic speech) and therefore may have an extra advantage over the other conditions.
+
+Finally, on average, participants at all ages made anticipatory gaze switches more often than would be expected by chance in the \textit{no speech} condition.
+One interpretation of this finding is that participants were not using linguistic information to predict upcoming turns at all---the only cue to turn taking in the \textit{no speech} condition was the alternating mouth movements of the conversing puppets.
+But this interpretation is not compatible with the effects of linguistic information that do emerge in the adult and child data: An advantage for lexical cues shaped by transition type and age.
+
+The core aims of Experiment 2 were to gain better traction on the individual roles of prosody and lexicosyntax in children's turn predictions, and to expand our age range to capture more developmental change.
+We found that effects of linguistic processing and age \textit{were} present in the dataset, but were primarily related to the question effect, and primarily occur in the normal speech condition.
+The relation between prosody and lexicosyntax in young children's anticipations is still not clear: The results of the prosody only manipulation suggest that children do not reliably use low-pass filtered speech to anticipate upcoming turn structure until age five, but on the other hand, the words only condition did not show significant overall differences from the no-speech baseline, including its interactions with age or transition type.
+If anything, the results do not support the idea that there is a strong early advantage for prosody in turn predictions.
+
+\section{General Discussion}
+\label{sec:gendisc}
+
+% CAN USE LATER: Because the form of an answer is partially determined by its question (e.g., locative expressions for ``where'' questions), participants could even be predicting what type of response they will hear, and so they are looking anticipatorily at the responder for confirmation of their prediction.
+This interpretation is initial, but in-line with work showing early response planning and speech act recognition \citep{gisladottir2015, bogels2015}.
+
+%Though children's anticipatory gazes did show some gains with age, adults on average still made more anticipatory switches than children in all conditions, even compared to 6-year-olds.
+Curiously, child and adult performance were much more similar by ages 3--5 in Experiment 1.
+This may indicate that children rely more in non-verbal cues in anticipating turn transitions or, alternatively, that adults are better at flexibly adapting to the turn-relevant cues present at any moment.
+
 %Our results suggest that the question effect emerges between ages one and two, but further testing is needed to establish whether this is the case.
@@
 Future work with more fine-grained age samples and better controlled \textit{normal} speech might then find that children can identify questions in conversation even earlier.
-%
-%\begin{table}[t]
-%\begin{center}
-%  \begin{tabular}{ccc}
-%    \hline
-%    Age group & Normal speech & No speech \\ 
-%    \hline
-%    1 & 0.43 & 0.24 \\
-%    2 & 0.61 & 0.52 \\
-%    3 & 0.63 & 0.42 \\
-%    4 & 0.78 & 0.52 \\
-%    5 & 0.91 & 0.52 \\
-%    6 & 0.75 & 0.50 \\
-%    Adult & 0.67 & 0.57 \\
-%    \hline
-%  \end{tabular}
-%\end{center}
-%  \caption{Proportion of participants showing numerically more anticipatory switches for questions than non-questions in the \textit{normal} and \textit{words only} conditions.}
-%\label{tab:questioneffectexp2}
-%\end{table}
-%
-%Finally, we confirmed that young children showed a benefit for prosody over lexicosyntax.
-We conducted two \textit{post-hoc} \textit{t}-tests for children under three, comparing their anticipatory switching rate against their estimated random switching rate in the \textit{prosody only} and \textit{words only} conditions.
-We found that anticipatory gazes significantly exceeded chance for the \textit{prosody only} condition (\textit{t}(187.677)=-3.004, \textit{p}=.003), but not for the \textit{words only} condition (\textit{t}(229.828)=-0.8307, \textit{p}=.407).
-This finding suggests that there is indeed an early advantage for prosody over lexicosyntax in children under three.
-The advantage for prosody was not driven by a question effect; children and adults did not consistently show more switches after questions than non-questions in the \textit{prosody only} condition (Figure \ref{fig:questionsEN}).
-So, whereas the early advantage for prosody appeared for both transition types (question and non-question), the later-emerging advantage for lexicosyntax and prosody together was primarily limited to question transitions.
-This pattern of results may suggest a qualitative shift in the kind of predictions children make while observing a conversation, with linguistic cues (like prosody) playing different roles at different ages.
-%
-%The model of adults' anticipatory gaze switches confirmed effects of gap duration and linguistic condition.
-Longer gaps led to more anticipatory switches (\textit{$\beta$}=0.656, \textit{SE}=0.176, \textit{t}=3.726, \textit{p}$<$.001).
-Adults also made significantly more anticipatory switches in the \textit{normal} condition than in the \textit{no speech} baseline, indicating that linguistic access did affect their overall performance (\textit{$\beta$}=-0.211, \textit{SE}=0.085, \textit{t}=-2.486, \textit{p}$<$.05).
-There were no other effects of linguistic condition.
-The model also showed a marginal two-way interaction of condition and transition type, with adults making more anticipatory gazes for questions in the \textit{normal} condition (\textit{$\beta$}=0.194, \textit{SE}=0.11, \textit{t}=1.773, \textit{p}=.076).
-Though the question effect was numerically present in both lexical conditions, it was only significant for adults in the \textit{normal} condition, in which they had both lexicosyntactic and prosodic cues to questionhood.
-This result suggests that prosodic cues still have a role to play in adult question recognition during third-party conversation.
-Since question effects were primarily limited to the \textit{normal} condition, there was no effect of transition type in the adults' data.
-
-%The core aims of Experiment 2 were to gain better traction on the individual roles of prosody and lexicosyntax in children's turn predictions, and to expand our age range to capture more developmental change.
-We found that effects of linguistic processing and age were present in this dataset, but both were primarily found within the question effect.
-Children and adults made more anticipatory switches after questions than non-questions, but did so more often when they had greater access to linguistic information, whether increased access came from age or with speech condition.
-Our results hint that lexical information is critical in identifying questions during conversation, especially for children.
-Even so, performance was consistently best in the \textit{normal} condition, in which participants could integrate lexical and prosodic information together.
-%
-%By extending the age range of participants, we were able to see more of the developmental trajectory in children's predictions about turn structure.
-We saw that one-year-olds already made more anticipatory switches than would be expected by chance in two of the four conditions.
-As they got older, children demonstrated above-chance anticipatory switches in more conditions and made more anticipatory switches overall.
-Like in Experiment 1, children between ages three and five behaved similarly to each other, and comparably to adults, showing small but steady gains in their anticipatory switches with age.
-Though the advantage for questions generally increased with age, the biggest developmental shift in our data occurred between one and two years, when the question effect began to reliably emerge in the two lexical conditions.
-It will be up to future work to disentangle which lexical and prosodic cues children begin to pick out in distinguishing questions from other speech acts, and when.
-%
-%The results from Experiment 2 complement those of Experiment 1, but still leave some open questions.
-For example, we were surprised to find no independent effects of linguistic condition outside of age and transition type interactions.
-In fact, the primary evidence that participants actually \textit{used} linguistic information came from the question effects, which were limited to the two lexical conditions.
-Given prior work, it is surprising that taking away lexical and prosodic information didn't have a larger impact on participants' anticipatory looking.
-We take this issue up further in the general discussion.
-%
-%We found that the importance of lexicosyntactic and prosodic information for turn prediction differed depending on age and transition type.
-While even one-year-olds made anticipatory switches at an above-chance rate in the \textit{normal} and \textit{prosody only} conditions, children didn't consistently surpass chance in the \textit{words only} condition until age three.
-This pattern of development suggests an early importance for prosodic information in children's turn anticipation.
-In contrast, the question effect---a primary driver of anticipatory switches in our data---seems to rely on lexical information, with some room for further support from prosody.
-So, although prosody may help children early on, it does not, on its own, help children to distinguish questions from non-questions.
-Rather, for both adults and children, prosody's role in the question effect was an additive one: The question effect was strongest when children had access to lexical \textit{and} prosodic information.
-In short, prosody has different roles to play in turn prediction before and after the lexically-driven question effect emerges.
-Importantly, these results do not support a straightforward view of lexicosyntax as the primary linguistic cue for turn predictions \citep[even for adults; ][]{de-ruiter2006, magyari2012}, but do support accounts of multiple cue integration \citep{duncan1972, ford1996, hirvenkari2013, torreira2015}.
-%
+
 %We controlled linguistic information in Experiment 2 by using phonetically-manipulated speech, extending speech manipulations used in prior work \citep{de-ruiter2006, keitel2013}.
@@
 
-\section{General Discussion}
-\label{sec:gendisc}
-
-% CAN USE LATER: Because the form of an answer is partially determined by its question (e.g., locative expressions for ``where'' questions), participants could even be predicting what type of response they will hear, and so they are looking anticipatorily at the responder for confirmation of their prediction.
-This interpretation is initial, but in-line with work showing early response planning and speech act recognition \citep{gisladottir2015, bogels2015}.
+%%%%%%%
 

### 2015-11-28 paper/CasillasFrank-Anticip.tex
some additions to the general discussion, not yet complete

@@
 \usepackage{float}
+\usepackage{longtable}
 
@@
 
-\begin{table}[h!]
-\begin{center}
-  \begin{tabular}{lcccl}
+\subsubsection{Statistical models}
+\label{sec:models2}
+
+We identified anticipatory gaze switches for all 85 usable turn transitions, and analyzed them for effects of language condition, transition type, and age with two mixed-effects logistic regressions \citep{lme4, R}.
+We again built separate models for children and adults because effects of age were only pertinent to the children's data.
+The child model included condition (normal/prosody only/words only/no speech; with no speech as the reference level), transition type (question vs.
+non-question), age (1, 2, 3, 4, 5, 6), and duration of the inter-turn gap (in seconds) as predictors, with full interactions between language condition, transition type, and age.
+We again included the duration of the inter-turn gap as a control predictor and added random effects of item (turn transition) and participant, with random slopes of transition type for participants \citep{barr2013}.
+The adult model included condition, transition type, their interactions, and duration as a control predictor, with participant and item included as random effects and random slopes of condition and transition type.
+
+\begin{longtable}{lcccl}
   \textbf{\textit{Children}} &&&& \\
@@
     \hline
-  \end{tabular}
-\end{center}
   \caption{Model output for children and adults' anticipatory gaze switches.}
 \label{tab:E2-models}
-\end{table}
-
-\subsubsection{Statistical models}
-\label{sec:models2}
-
-We identified anticipatory gaze switches for all 85 usable turn transitions, and analyzed them for effects of language condition, transition type, and age with two mixed-effects logistic regressions \citep{lme4, R}.
-We again built separate models for children and adults because effects of age were only pertinent to the children's data.
-The child model included condition (normal/prosody only/words only/no speech; with no speech as the reference level), transition type (question vs.
-non-question), age (1, 2, 3, 4, 5, 6), and duration of the inter-turn gap (in seconds) as predictors, with full interactions between language condition, transition type, and age.
-We again included the duration of the inter-turn gap as a control predictor and added random effects of item (turn transition) and participant, with random slopes of transition type for participants \citep{barr2013}.
-The adult model included condition, transition type, their interactions, and duration as a control predictor, with participant and item included as random effects and random slopes of condition and transition type.
+\end{longtable}
 
@@
 
-% CAN USE LATER: Because the form of an answer is partially determined by its question (e.g., locative expressions for ``where'' questions), participants could even be predicting what type of response they will hear, and so they are looking anticipatorily at the responder for confirmation of their prediction.
-This interpretation is initial, but in-line with work showing early response planning and speech act recognition \citep{gisladottir2015, bogels2015}.
-
-%Though children's anticipatory gazes did show some gains with age, adults on average still made more anticipatory switches than children in all conditions, even compared to 6-year-olds.
-Curiously, child and adult performance were much more similar by ages 3--5 in Experiment 1.
-This may indicate that children rely more in non-verbal cues in anticipating turn transitions or, alternatively, that adults are better at flexibly adapting to the turn-relevant cues present at any moment.
-
-%Our results suggest that the question effect emerges between ages one and two, but further testing is needed to establish whether this is the case.
+Children begin to develop conversational turn-taking skills long before their first words \citep{bateson1975, hilbrink2015, jaffe2001, snow1977}.
+As they acquire language, they also acquire the information needed to make accurate predictions about upcoming turn structure.
+Until recently, we have had very little data on how children weave language into their already-existing turn-taking behaviors.
+
+In two experiments investigating children's anticipatory gaze to upcoming speakers, we found evidence that turn prediction develops early in childhood and that spontaneous predictions are primarily driven by participants' expectation of an immediate response in the next turn.
+In making predictions about upcoming turn structure, children used a combination lexical and prosodic cues; neither prosodic nor lexical cues alone were sufficient to support increased anticipatory gaze.
+We also found no early advantage for prosody over lexicosyntax, and instead found that children were unable to make above-chance anticipatory gazes in the prosody only condition until age five.
+We discuss these findings with respect to the role of linguistic cues in predictions about upcoming turn structure, the importance of questions in spontaneous predictions about conversation, and children's developing competence as conversationalists.
+
+\subsection{Predicting turn structure with linguistic cues}
+
+Prior work with adults has found a consistent and critical role for lexicosyntax in predicting upcoming turn structure \citep{de-ruiter2006, magyari2012}, with the role of prosody still under debate \citep{duncan1972, ford1996, torreira2015}.
+Knowing that children comprehend more about prosody than lexicosyntax early on (see introduction; also see \citealp{speer2009} for a review), we thought it possible that young children would instead show an advantage for prosody over lexicosyntax.
+Our results suggest that, on the contrary, when presented with \textit{only} prosodic information, children's spontaneous predictions about upcoming turn structure are limited until age five.
+Do children instead rely on lexicosyntax, as adults are proposed to do?
+
+We found no evidence that, for children, lexicosyntax alone is ``sufficient'' (equal to full linguistic information) for spontaneous turn prediction \citep[pg.
+531]{de-ruiter2006}: In both experiments, children's performance was best in conditions when they had access to the full linguistic signal.
+Adults on the other hand, showed significant gains in anticipatory gaze switching in both conditions with lexical cues, compared to a no speech baseline.
+
+Participants appeared to use linguistic cues to make more anticipatory gaze switches in both experiments.
+Questions are often marked both prosodically and lexicosyntactically for their speech act status.
+The close link between prosodic and syntactic structure makes it difficult to tease apart how predictive processing in one domain is distinct from the other in turn prediction.
+That being said, compared to prosodic contours (e.g., final rising intonation), lexicosyntactic cues like \textut{wh}-words, \textit{do}-insertion, and subject-auxiliary inversion are frequent, categorical, and early-occurring in the utterance.
+Children may therefore have an easier time picking out and interpreting lexical cues to questionhood on the fly.
+Question turns started yielding significantly more anticipatory gazes by age 3--4 in the normal speech condition of Experiment 2, by which time children frequently hear and use a variety of polar \textit{wh}-questions \citep{clark2009}.
+Furthermore, while lexicosyntactic question cues were available on every instance of \textit{wh}- and \textit{yes/no} questions in our stimuli, prosodic question cues were only salient on \textit{yes/no} questions and, even then, the mapping of prosodic contour to speech act (e.g., high final rises for polar questions) is far from one-to-one \citep{hedberg2010}.
+
+Finally, we found that our adult controls were affected by the absence of prosody (within the question effect) while Keitel et al.
+(2013) found that adults' were unaffected by flattened pitch.
+But again, because we flattened pitch for individual speakers and controlled syllable duration, we are not able to directly compare our findings to theirs.
+
+% WHAT CUES ARE THEY USING?
+
+\subsubsection{The question effect}
+% What exactly is the Q effect? Could it exist for non-Qs?
+% When does the Q effect emerge and exactly what cues is it relying on?
+% Are they predicting the form of the answer?
+
+Third, we found that anticipatory looking was primarily driven by question transitions, a pattern that has so far not been discussed in other observer gaze studies, on children or adults (Keitel et al., 2013; Hirvenkari, 2013; Tice and Henetz, \citeyear{TiceHenetz11}).
+The question effects in the current study were critical to understanding the role of lexicosyntactic information in participants' turn predictions.
+
+In both of our experiments we found a robust advantage for questions in participants' anticipatory switching because participants made more switches after questions than non-questions.
+Children and adults both treated questions as if they had a special interactive status, and with good reason: Questions make a speaker switch immediately relevant, helping the listener to predict what will happen next (i.e., an answer from the addressee).
+Linguistic cues to questionhood often occur early in the utterance, making questions easy to identify early on and thereby giving observers time to plan a reaction (i.e., a gaze switch).
+Also, the form of a question can constrain the type of response that will come next (e.g., a location after a \textit{where} question), further increasing the predictability of the interactional sequence.
+
+Children get frequent and early experience with hearing and answering questions.
+Questions make up approximately one third of the utterances children hear, before and after the onset of speech, and even into their preschool years \citep{fitneva2012, henning2005, shatz1979}.\footnote{There is substantial variation in this proportion by individual and socioeconomic class \citep{hart1992}.} Many of the questions directed to young children are ``test' questions---questions that the caregiver already has the answer to (e.g., ``What does a cat say?'').
+Caregivers use questions to get their young children's attention and to ensure that information is in common ground, even if the responses are non-verbal or infelicitous \citep{bruner1985, fitneva2012, snow1977}.
+So, in addition to having a special interactive status, questions are a core characteristic of many caregiver-child interactions.
+
+We cannot definitively say with our current data what property of questions caused children to switch their gaze more often.
+It is true that questions make a speaker transition more likely and that questions have distinctive linguistic cues, both early and late, to identify themselves.
+But other speech acts and request formats have these same properties.
+Imperatives, compliments, and complaints all make a response from the addressee likely \citep{schegloff2007}.
+Rhetorical and tag questions, on the other hand, take a similar form to prototypical polar questions, but do not always require an answer.
+So, though it is clear that children anticipate responses more often for questions than non-questions, their actual understanding of questionhood as a linguistically- or interactionally-marked type of speech act still needs further exploration.
+
+Our results suggest that the question effect emerges between ages one and two, but further testing is needed to establish whether this is the case.
 For example, the size and emergence of the question effect for various question types and question cues could differ (\textit{wh-} vs.
@@
 
-%We controlled linguistic information in Experiment 2 by using phonetically-manipulated speech, extending speech manipulations used in prior work \citep{de-ruiter2006, keitel2013}.
+Because the form of an answer is partially determined by its question (e.g., locative expressions for ``where'' questions), participants could even be predicting what type of response they will hear, and so they are looking anticipatorily at the responder for confirmation of their prediction.
+This interpretation is initial, but in-line with work showing early response planning and speech act recognition \citep{gisladottir2015, bogels2015}.
+
+\subsubsection{Early competence for turn taking?}
+% When does children's performance become adult-like?
+
+Though children's anticipatory gazes did show some gains with age, adults on average still made more anticipatory switches than children in all conditions, even compared to 6-year-olds.
+Curiously, child and adult performance were much more similar by ages 3--5 in Experiment 1.
+This may indicate that children rely more in non-verbal cues in anticipating turn transitions or, alternatively, that adults are better at flexibly adapting to the turn-relevant cues present at any moment.
+
+First, we found that children begin making above-chance anticipatory switches much earlier than three---our one-year-olds were already looking above chance in two of our four linguistic conditions in Experiment 2.
+This earlier emergence of anticipatory switching could be due to our more child-friendly stimuli, which we designed to be engaging and to mimic the kind of speech children often experience interactionally.
+Second, because we included more linguistic conditions to isolate the effects of prosody and lexicosyntax, we gained a broader perspective on how linguistic information changed children's predictions.
+Ultimately, the added conditions allowed us to find an early advantage for prosody and a later advantage for lexicosyntax.
+
+Our findings are consistent with observational work on children's turn taking.
+Young children are slow turn-takers in spontaneous conversation \citep{casillas2016} even though they begin taking turns in infancy \citep{hilbrink2015, jaffe2001}.
+We found that children begin to spontaneously predict upcoming turn structure at age one.
+Children's gaze switches were also most common in the inter-turn gap, suggesting that they expected an immediate response from the addressee.
+Our results therefore support the idea of an early-developed competence for the timing of turn taking.
+But because children do not themselves use adult-like conversational timing until much later \citep[age six; ][]{ervin-tripp1979}, our results also indirectly point toward response planning as a key factor contributing to children's slow responses in real-time interaction \citep{casillas2016}.
+
+\subsection{Limitations}
+% assumptions about saccadic planning time
+% There is a problem with manipulated speech
+
+We controlled linguistic information in Experiment 2 by using phonetically-manipulated speech, extending speech manipulations used in prior work \citep{de-ruiter2006, keitel2013}.
 While phonetic manipulation gave us control over the linguistic signal, it also created speech sounds that children don't usually hear in their natural environment.
@@
 Children could have had trouble processing the \textit{words only} and \textit{prosody only} conditions because they were unfamiliar, and not just because they had less linguistic information available.
-%
-%It would take a different type of stimulus to compare lexical and prosodic information without phonetic manipulation.
+
+It would take a different type of stimulus to compare lexical and prosodic information without phonetic manipulation.
 For example, carefully scripted or cross-spliced conversations could control for the presence or absence of prosodic and lexicosyntactic cues to turn transition with natural speech.
@@
 
-%%%%%%%
-
-%Children begin to develop conversational turn-taking skills long before their first words \citep{bateson1975, hilbrink2015, jaffe2001, snow1977}.
-As they acquire language, they also acquire the information needed to make accurate predictions about upcoming turn structure.
-Until recently, we have had very little data on how children weave language into their already-existing turn-taking behaviors.
-%
-%In two experiments investigating children's anticipatory gaze to upcoming speakers, we found evidence that turn prediction develops early in childhood, and that linguistic cues are integrated differently at different ages.
-Children under three showed a general advantage for prosody over lexicosyntax, contrary to findings for adults in prior work \citep{de-ruiter2006}.
-Children two and older also showed a slight advantage for lexicosyntax, but only in the form of a transition type effect: They made more anticipatory switches following questions than non-questions in conditions where lexical information was present.
-Two-year-olds thus appeared to be at a developmental transition point, still showing an overall advantage for prosody, but beginning to show an advantage for lexicosyntax within the domain of question-answer sequences.
-We found no evidence that, for children or adults, lexicosyntax alone was ``sufficient'' (equal to full linguistic information) for spontaneous turn prediction \citep[pg.
-531]{de-ruiter2006}: Participants' performance was best in conditions when they had access to the full linguistic signal.
-%
-%\subsection{Developmental implications}
-%
-%The primary aim of our study was to find out how linguistic information affects children's predictions about upcoming turn structure across development.
-Two primary developmental findings emerged from our data.
-First, participants' spontaneous predictions largely occurred during question--answer transitions.
-Second, children's use of linguistic signals in making predictions depended on their age, with an early advantage for prosody over lexicosyntax and a later (partial) advantage for lexicosyntax over prosody.
-%
-%\subsubsection{The advantage for questions}
-%
-%In both of our experiments we found a robust advantage for questions in participants' anticipatory switching because participants made more switches after questions than non-questions.
-Children and adults both treated questions as if they had a special interactive status, and with good reason: Questions make a speaker switch immediately relevant, helping the listener to predict what will happen next (i.e., an answer from the addressee).
-Linguistic cues to questionhood often occur early in the utterance, making questions easy to identify early on and thereby giving observers time to plan a reaction (i.e., a gaze switch).
-Also, the form of a question can constrain the type of response that will come next (e.g., a location after a \textit{where} question), further increasing the predictability of the interactional sequence.
-%
-%Children get frequent and early experience with hearing and answering questions.
-Questions make up approximately one third of the utterances children hear, before and after the onset of speech, and even into their preschool years \citep{fitneva2012, henning2005, shatz1979}.\footnote{There is substantial variation in this proportion by individual and socioeconomic class \citep{hart1992}.} Many of the questions directed to young children are ``test' questions---questions that the caregiver already has the answer to (e.g., ``What does a cat say?'').
-Caregivers use questions to get their young children's attention and to ensure that information is in common ground, even if the responses are non-verbal or infelicitous \citep{bruner1985, fitneva2012, snow1977}.
-So, in addition to having a special interactive status, questions are a core characteristic of many caregiver-child interactions.
-%
-%We cannot definitively say with our current data what property of questions caused children to switch their gaze more often.
-It is true that questions make a speaker transition more likely and that questions have distinctive linguistic cues, both early and late, to identify themselves.
-But other speech acts and request formats have these same properties.
-Imperatives, compliments, and complaints all make a response from the addressee likely \citep{schegloff2007}.
-Rhetorical and tag questions, on the other hand, take a similar form to prototypical polar questions, but do not always require an answer.
-So, though it is clear that children anticipate responses more often for questions than non-questions, their actual understanding of questionhood as a linguistically- or interactionally-marked type of speech act still needs further exploration.
-%
-%\subsubsection{Linguistic cues for turn prediction}
-%
-%Prior work with adults has found a consistent and critical role for lexicosyntax in predicting upcoming turn structure \citep{de-ruiter2006, magyari2012}, with more debate over the role of prosody \citep{duncan1972, ford1996, torreira2015}.
-Knowing that children comprehend more about prosody than lexicosyntax early on (see introduction; also see \citealp{speer2009} for a review), we thought it possible that young children would instead show an advantage for prosody over lexicosyntax.
-Our results suggest that there are roles for both prosody and lexicosyntax in children's predictions about upcoming turn structure, but that there are different roles for these information sources at different ages.
-%
-%Early on, there was an advantage for prosody over lexicosyntax in children's anticipatory switches: Anticipatory switching didn't reliably exceed chance for lexicosyntactic information  until age three, but was already significantly greater than chance for prosodic information  at age one.
-This early advantage for prosody was not driven by the advantage for questions: While some one-year-olds in our study already made more switches after questions than non-questions, most did not, plus prosody alone did not elicit a robust question effect at any age (Figure \ref{fig:questionsEN}).
-%
-%Although children showed an initial advantage for prosody, their predictions before age three were probably limited to basic prosodic contours and boundaries, and simple cues such as final lengthening and turn-final silence \citep{pannekamp2006}.
-More complex structural predictions would have required them to access lexicosyntactic information because higher-level structure in the prosodic signal is linked to the syntactic structure being used.
-%
-%In both experiments, even young children showed an advantage for questions in conditions where they had lexicosyntax, suggesting that lexicosyntax is a critical cue for question identification.
-The question effect emerged by age two, when children's syntactic comprehension and lexical inventories are still relatively limited \citep{clark2009}.
-That being said, syntactic and lexical markers to questionhood in English are frequent, categorical, and early-occurring in the utterance (e.g., ``do'', \textit{wh}-words, and subject-auxiliary inversion).
-Lexicosyntactic question cues were available on every instance of \textit{wh}- and \textit{yes/no} questions in our stimuli, whereas prosodic question cues were only salient on \textit{yes/no} questions.
-Thus, the unambiguous presence of lexicosyntactic cues on \textit{wh}- and \textit{yes/no} questions may have been one reason why we saw an advantage for questions in the lexical conditions, but not in the \textit{prosody only} condition.
-%
-%Importantly, the question effect was not driven purely by lexicosyntactic information.
-In both experiments, prosody still made a contribution to the advantage for questions, even though prosody alone was not enough to elicit a question effect.
-For example, older children and adults in Experiment 1 showed a question effect in the languages they did not understand---presumably they used prosodic and non-verbal information to distinguish questions from non-questions.
-In Experiment 2, the advantage for questions was biggest for speech with lexicosyntactic \textit{and} prosodic information together.
-All in all, children appear to improve at identifying questions in ongoing conversation by using lexicosyntactic cues as they get older.
-But, prosodic cues still aid in picking out questions from ongoing conversation.
-%
-%The close link between prosodic and syntactic structure makes it difficult to tease apart how predictive processing in one domain (e.g., prosody) is distinct from the other (e.g., syntax) in turn prediction.
-We leave this challenge to future work.
-%
-%\subsection{Contributions of the current work}
-%
-%Our findings are consistent with observational work on children's turn taking.
-Young children are slow turn-takers in spontaneous conversation \citep{casillas2016} even though they begin taking turns in infancy \citep{hilbrink2015, jaffe2001}.
-We found that children begin to spontaneously predict upcoming turn structure at age one.
-Children's gaze switches were also most common in the inter-turn gap, suggesting that they expected an immediate response from the addressee.
-Our results therefore support the idea of an early-developed competence for the timing of turn taking.
-But because children do not themselves use adult-like conversational timing until much later \citep[age six; ][]{ervin-tripp1979}, our results also indirectly point toward response planning as a key factor contributing to children's slow responses in real-time interaction \citep{casillas2016}.
-%
-%Our findings are only partially consistent with prior experimental work on turn taking.
-De Ruiter and colleagues (2006) found that adults' accuracy in identifying upcoming turn-ends was not significantly affected when intonation contours were flattened.
-Keitel and colleagues (2013) found the same for adults' anticipatory gaze switches, only finding that intonation flattening affected children's switches at 36 months (with switches by younger children never exceeding chance).
-Our results, in contrast, suggest that children's anticipatory switches emerge much earlier, and rely on prosody before they do lexicosyntax.
-Furthermore, our adult control participants patterned similarly to the children: They showed question effects for languages they didn't speak (Experiment 1) and showed the strongest question effects for lexicosyntax and prosody together (Experiment 2).
-Thus, neither the child nor the adult data support the idea that lexicosyntactic information is equal to full linguistic information in making predictions about upcoming turn structure \citep{de-ruiter2006}, at least in participants' spontaneous looking behavior.
-This may partially be due to the fact that we controlled syllable duration in addition to intonation, thereby eliminating two of the most salient cues to prosodic structure.
-%
-%Despite extreme differences in linguistic information across conditions, independent effects of linguistic information (outside of interactions) were minimal in our experiments.
+In Experiment 2, we tried to improve our method for controlling linguistic information by phonetically manipulating improvised dialogic speech.
+Although this decision allowed us to follow prior work more closely \citep{de-ruiter2006, keitel2013}, it also limits our interpretation of children's behavior.
+Low-pass filtered and robotic-like speech are not typical within children's speech environment, and so it may have been unfair to compare the limited cue conditions in Experiment 2 to the normal speech condition, which not only had full linguistic cues, but is also more typically present in children's speech environments.
+One alternative to explore for the future would be scripted speech, cross-spliced speech, or speech that is carefully selected from a large corpus, given a set of specific linguistic constraints.
+
+Relatedly, since we used improvised conversation recordings, we did not control how often the speakers switched back and forth, or how often a switch occurred after a pause in speech.
+It is possible that, to some extent, children could have formed expectations about the turn structure on the basis of the frequent switching alone \citep[see, e.g., ][]{thorgrimsson2015}.
+If this statistical expectation were the only thing leading children's anticipatory gaze switches, we would have predicted more uniformity in looking behavior across the conditions and no major effects of linguistic signal, age, or transition type.
+But that is not what we saw.
+Nevertheless, another advantage of using scripted conversation would be the ability to control switch frequency and eliminate this issue.
+
+Finally, we carefully designed a scheme for identifying anticipatory turn switches in the data, building off of Keitel and colleagues' (2013) analyses.
+In both our study and theirs, anticipatory switches that happen very early in the turn are not counted.
+Instead, they would actually count as ``random'' switches in the baseline analyses.
+These same early switches are counted as non-random in other studies of children's gaze to conversational stimuli \citep{bakker2011, hofsten2009}.
+Since most studies, even those using a wider switching window \citep[e.g., ][]{hirvenkari2013} find that switches primarily happen close to or immediately after the inter-turn gap begins, the approach used here and in Keitel et al.
+(2013) probably captures most of participants' turn-relevant switching behavior.
+But if early looks do indeed increase as children get older and have more access to early syntactic cues, it would be helpful to develop a second measure of very early anticipatory looks.
+
+Despite extreme differences in linguistic information across conditions, independent effects of linguistic information (outside of interactions) were minimal in our experiments.
 But, using a button-press measure with similar stimuli, de Ruiter and colleagues (2006) found robust effects of linguistic condition.
@@
 We have seen in the current study that questions are interactionally sufficient to motivate early predictions.
-%
-%Moving away from button press measures, Keitel and colleagues (2013) used observer gaze to investigate children's turn predictions, finding: (a) No evidence that children under three make switches above chance, (b) that three-year-olds' switches are significantly affected by flattened intonation, and (c) that adults' switches are not.
-Our current results are consistent with the general finding from their study, that prosody matters for young children's turn predictions.
-But our results also differ from theirs in a few important ways.
-%
-%First, we found that children begin making above-chance anticipatory switches much earlier than three---our one-year-olds were already looking above chance in two of our four linguistic conditions in Experiment 2.
-This earlier emergence of anticipatory switching could be due to our more child-friendly stimuli, which we designed to be engaging and to mimic the kind of speech children often experience interactionally.
-Second, because we included more linguistic conditions to isolate the effects of prosody and lexicosyntax, we gained a broader perspective on how linguistic information changed children's predictions.
-Ultimately, the added conditions allowed us to find an early advantage for prosody and a later advantage for lexicosyntax.
-%
-%Third, we found that anticipatory looking was primarily driven by question transitions, a pattern that has so far not been discussed in other observer gaze studies, on children or adults (Keitel et al., 2013; Hirvenkari, 2013; Tice and Henetz, \citeyear{TiceHenetz11}).
-The question effects in the current study were critical to understanding the role of lexicosyntactic information in participants' turn predictions.
-Finally, we found that our adult controls were affected by the absence of prosody (within the question effect) while Keitel et al.
-(2013) found that adults' were unaffected by flattened pitch.
-But again, because we flattened pitch for individual speakers and controlled syllable duration, we are not able to directly compare our findings to theirs.
-%
-%\subsection{Limitations}
-%
-%In Experiment 2, we tried to improve our method for controlling linguistic information by phonetically manipulating improvised dialogic speech.
-Although this decision allowed us to follow prior work more closely \citep{de-ruiter2006, keitel2013}, it also limits our interpretation of children's behavior.
-Low-pass filtered and robotic-like speech are not typical within children's speech environment, and so it may have been unfair to compare the limited cue conditions in Experiment 2 to the normal speech condition, which not only had full linguistic cues, but is also more typically present in children's speech environments.
-One alternative to explore for the future would be scripted speech, cross-spliced speech, or speech that is carefully selected from a large corpus, given a set of specific linguistic constraints.
-%
-%Relatedly, since we used improvised conversation recordings, we did not control how often the speakers switched back and forth, or how often a switch occurred after a pause in speech.
-It is possible that, to some extent, children could have formed expectations about the turn structure on the basis of the frequent switching alone \citep[see, e.g., ][]{thorgrimsson2015}.
-If this statistical expectation were the only thing leading children's anticipatory gaze switches, we would have predicted more uniformity in looking behavior across the conditions and no major effects of linguistic signal, age, or transition type.
-But that is not what we saw.
-Nevertheless, another advantage of using scripted conversation would be the ability to control switch frequency and eliminate this issue.
-%
-%Finally, we carefully designed a scheme for identifying anticipatory turn switches in the data, building off of Keitel and colleagues' (2013) analyses.
-In both our study and theirs, anticipatory switches that happen very early in the turn are not counted.
-Instead, they would actually count as ``random'' switches in the baseline analyses.
-These same early switches are counted as non-random in other studies of children's gaze to conversational stimuli \citep{bakker2011, hofsten2009}.
-Since most studies, even those using a wider switching window \citep[e.g., ][]{hirvenkari2013} find that switches primarily happen close to or immediately after the inter-turn gap begins, the approach used here and in Keitel et al.
-(2013) probably captures most of participants' turn-relevant switching behavior.
-But if early looks do indeed increase as children get older and have more access to early syntactic cues, it would be helpful to develop a second measure of very early anticipatory looks.
-%
-%\subsection{Conclusions}
+
+\subsection{Conclusions}
 %

### 2015-11-30 paper/CasillasFrank-Anticip.tex
full draft of paper

@@
 \begin{abstract}
-We investigate language acquisition through the lens of a fundamental conversational skill: Turn taking.
-Children begin developing turn-taking skills in infancy, but take several years to assimilate their growing knowledge of language into their turn-taking behavior.
-In two eye-tracking experiments, with children across a wide developmental range, we measured spontaneous predictions about upcoming speaker change while controlling the amount of linguistic information available.
-We found that children already predicted upcoming turns at age one, but that they integrated linguistic cues differently at different ages.
-Children under three showed a general advantage for prosody over lexicosyntax, contrary to prior findings for adults.
-Children two and older showed more predictive switches for questions than non-questions, but only when lexical information was available.
-We found no evidence that lexicosyntax alone guides turn prediction---instead, participants' performance was best overall with access to lexicosyntax and prosody together.
-Our results point to the importance of studying children's linguistic processing in the context of conversation.
+Children begin developing turn-taking skills in infancy but take several years to assimilate their growing knowledge of language into their turn-taking behavior.
+In two eye-tracking experiments, we measured children's anticipatory gaze to upcoming responders while controlling linguistic cues to upcoming turn structure.
+In Experiment 1, we showed English and non-English conversations to English-speaking participants, finding minimal differences between the predictive looking behavior of preschoolers and adults.
+In Experiment 2, we phonetically controlled lexicosyntactic and prosodic cues in English-only speech, finding that children's predictive looking behavior improved from ages one to six, but that even one-year-olds made more anticipatory looks than would be expected by chance.
+In both experiments, children and adults anticipated more often after hearing questions.
+Like adults, prosody alone did not improve childrenÕs predictive gaze shifts.
+But, unlike adults, lexical cues alone were also not sufficient to improve prediction---children's performance was best overall with access to lexicosyntax and prosody together.
+Our findings support an account in which turn prediction emerges in infancy, but takes several years before becoming fully integrated with linguistic processing.
 \end{abstract}
@@
 
-% \linenumbers
+\linenumbers
 
@@
 Along these lines, early non-verbal interactions establish the protocol by which children come to use language with others.
-And as they acquire language, they also come to integrate it into the preverbal turn-taking systems.
+And as they acquire language, they also come to integrate it into their preverbal turn-taking systems.
 
@@
 To take a turn, participants need to make predictions about what conversational content will be relevant next.
-In some cases, relevant next turns are somewhat obvious (e.g., question--response) while, in other cases, there are multiple relevant next actions to choose from, or no obvious next action at all (e.g., after a closing).
+In some cases, relevant next turns are somewhat obvious (e.g., question--response) while, in other cases, there are multiple relevant next actions to choose from or no obvious next action at all (e.g., after a closing).
 
@@
 Non-verbal cues (e.g., gaze, posture, and pointing) often appear at turn-boundaries and can sometimes act as late indicators of an upcoming speaker switch \citep{rossano2009, stivers2010}.
-Additionally, the sequential context of a turn can make it clear what will come next: Answers after questions; thanks or denial after compliments, et cetera \citep{schegloff2007}.
+Additionally, the sequential context of a turn can make it clear what will come next: Answers after questions, thanks or denial after compliments, et cetera \citep{schegloff2007}.
 
@@
 De Ruiter and colleagues found that, with pitch-flattened speech, the timing of participants' button responses was comparable to their timing with the full linguistic signal.
-But, when no lexical information was available, participants' responses were significantly earlier.
+But when no lexical information was available, participants' responses were significantly earlier.
 The authors concluded that lexicosyntactic information\footnote{The ``lexicosyntactic'' condition only included flattened pitch and so was not exclusively lexicosyntactic---the speech would still have residual prosodic structure, including syllable duration and intensity.} was necessary and possibly sufficient for turn-end projection, while intonation was neither necessary nor sufficient.
@@
 
-The role of prosody for adult turn-prediction is still a matter of debate.
+The role of prosody for adult turn prediction is still a matter of debate.
 De Ruiter and colleagues' (2006) experiment focused on the role of intonation, which is only a partial index of prosody.
@@
 But then, later on, children's response latencies slow down again, peaking at average gaps of more than 1000 msec at nine months, with only very gradual improvement after that \citep{hilbrink2015}.
-While children's avoidance of overlap is nearly adult-like by nine months, the timing of their non-overlapped responses stays much longer than the 200 msec standard for the next few years \citep{garvey1984, ervin-tripp1979}.
+While children's avoidance of overlap is nearly adult-like by nine months, the timing of their non-overlapped responses stays much longer than the 200 msec standard for the next few years \citep{casillas2016, garvey1984, ervin-tripp1979}.
 
@@
 Keitel and colleagues \citeyearpar{keitel2013} addressed the random baseline issue in their study of 6-, 12-, 24-, and 36-month-olds.
-They asked participants to watch short videos of conversation, and tracked their eye movements at points of speaker change.
+They asked participants to watch short videos of conversation and tracked their eye movements at points of speaker change.
 They found that children's anticipatory gaze frequency was only greater than chance for 36-month-olds and adults.
@@
 Instead, they used gaze switches during ongoing speech as a baseline, during which switching is least likely to occur \citep{hirvenkari2013} and thereby maximizing their chances of finding a difference between gaze frequency at turn transitions and their baseline rate.
-A more conservative baseline would be to compare participants' looking behavior at turn transitions to their looking behavior during randomly-selected windows of time throughout the stimulus.
+A more conservative baseline would be to compare participants' looking behavior at turn transitions to their looking behavior during randomly selected windows of time throughout the stimulus, including turn transitions.
 We follow this conservative approach in our work.
@@
 \citeyearpar{keitel2013} study still demonstrates intriguing age-based differences in children's ability to predict upcoming turn structure, and the results suggest that both semantic and intonational development \textit{do} play a role in children's looking patterns.
-Our current work thus takes this paradigm as our starting point.\footnote{See also \citet{casillas2012, casillas2013}.}
+Our current work thus takes this paradigm as a starting point.\footnote{See also \citet{casillas2012, casillas2013}.}
 
@@
 
-\subsection{The Current Study}
-We report here on the role of linguistic processing in children's predictions about upcoming turn structure.
-We focus in particular on how children use prosodic and lexicosyntactic information to make their predictions.
+We report here on the role of linguistic processing in children's predictions about upcoming turn structure, in particular on how children use prosodic and lexicosyntactic information to make their predictions.
 Prior work has focused mainly on lexicosyntax and intonation, and not on prosody proper (\citealp{de-ruiter2006}; \citealp{keitel2013}, but see \citealp{torreira2015}), even though infants seem to acquire the basic rhythmic properties of the prosodic signal first \citep{mehler1988, moon1993, nazzi2003}.
-
-In two eye-tracking experiments, we measured children's anticipatory gaze to upcoming responders while controlling for the amount of lexicosyntactic and prosodic information available.
-In Experiment 1, English-speaking participants viewed video clips of naturalistic conversation from several different languages.
-We used multiple languages to control for the presence of lexicosyntactic information while keeping prosodic and non-linguistic cues intact.
-The results showed minimal differences between the predictive looking behavior of preschoolers and adults.
-In Experiment 2, we created artificial (puppet) visual scenes, enabling us to separately control lexicosyntactic and prosodic cues in the conversational stimuli.
-In this more controlled paradigm, we found that children's predictive looking behavior improved from ages one to six, but that even one-year-olds made more anticipatory looks than would be expected by chance.
-
-In both experiments children consistently looked faster to responders after hearing questions, compared to non-questions.
-Both prosodic and lexicosyntactic information played a role in children's predictions about turn structure, but the two information sources were used differently at different ages.
-Our findings overall support an account in which predictive processes for turn taking in conversation are present early, but their integration with linguistic information takes substantial practice.
 
@@
 We encouraged talkers to ask at least a few questions to each other during the improvisation.
-The resulting conversations were therefore not entirely spontaneous, but were as close as possible while still remaining child-oriented in topic, prosodic pattern, and lexicosyntactic construction.\footnote{All of the non-English talkers were fluent in English as a second language, and some fluently spoke a third or more language.
+The resulting conversations were therefore not entirely spontaneous, but were as close as possible while still remaining child-oriented in topic, prosodic pattern, and lexicosyntactic construction.\footnote{All of the non-English talkers were fluent in English as a second language, and some fluently spoke three or more languages.
 We chose male-female pairs as a natural way of creating contrast between the two talker voices.}
-
-%% PART UNFINISHED: link for example video
 
@@
 As such, they would still be fixated on the prior speaker when the overlap started, and then would have to switch their gaze \textit{reactively} to the responder.} This left approximately equal numbers of turn transitions available for analysis in the English (N$=$20) and non-English (N$=$16) videos.
-On average, the inter-turn gaps for English videos (mean$=$318, median$=$302, stdv$=$112 msec) were slightly longer than for non-English videos (mean$=$286, median$=$251, stdv$=$122 msec).
+On average, the inter-turn gaps for English videos (mean$=$318, median$=$302, stdev$=$112 msec) were slightly longer than for non-English videos (mean$=$286, median$=$251, stdev$=$122 msec).
 The longer gaps in the English videos could give them a slight advantage: Our definition of an ``anticipatory gaze shift'' includes shifts that are initiated during the gap between turns (Figure \ref{fig:criterion}), so participants had slightly more time to make anticipatory shifts in the English videos.
@@
 Questions made up exactly half of the turn transitions in the English (N$=$10) and non-English (N$=$8) videos.
-In the English videos, inter-turn gaps were slightly shorter for questions (mean$=$310, median$=$293, stdev$=$112 msec) than non-questions (mean$=$325, median$=$315, stdv$=$118 msec).
-Non-English videos did not show a large difference in transition time for questions (mean$=$270, median$=$257, stdv$=$116 msec) and non-questions (mean$=$302, median$=$252, stdv$=$134 msec).
+In the English videos, inter-turn gaps were slightly shorter for questions (mean$=$310, median$=$293, stdev$=$112 msec) than non-questions (mean$=$325, median$=$315, stdev$=$118 msec).
+Non-English videos did not show a large difference in transition time for questions (mean$=$270, median$=$257, stdev$=$116 msec) and non-questions (mean$=$302, median$=$252, stdev$=$134 msec).
 
@@
 \begin{table}[h!]
+  \begin{scriptsize}
 \begin{center}
@@
     LgCond$=$\textit{non-English}* &     2.68166 &   1.35016 &  1.986 & 0.047013 * \\
-    Type$=$\textit{non-Question} &&&& \\
+    \hspace*{5mm} Type$=$\textit{non-Question} &&&& \\
     Age*LgCond$=$\textit{non-English}* & -0.45632 &   0.30163 & -1.513 & 0.130315  \\
-    Type$=$\textit{non-Question} &&&& \\
+    \hspace*{5mm} Type$=$\textit{non-Question} &&&& \\
     \hline
@@
     LgCond$=$\textit{non-English}* &        1.3626 &    3.0077 &  0.453 & 0.650527     \\
-    Duration &&&& \\
+    \hspace*{5mm} Duration &&&& \\
     Type$=$\textit{non-Question}*          &  10.5107 &    3.3459 &  3.141 & 0.001682  ** \\
-    Duration &&&& \\
+    \hspace*{5mm} Duration &&&& \\
     LgCond$=$\textit{non-English}* & -6.3156 &    4.4926 & -1.406 & 0.159790 \\    
     Type$=$\textit{non-Question}* &&&& \\
-    Duration &&&& \\
+    \hspace*{5mm} Duration &&&& \\
     \hline
@@
 \end{center}
+  \end{scriptsize}
   \caption{Model output for children and adults' anticipatory gaze switches.}
@@
 
-% PART UNFINISHED (potentially)
-% We excluded the output of random-permutation models that did not converge to remove unreliable model results from our percentile calculations below (XX\% and XX\% of models for children and adults, respectively).
-
 Our baseline analyses revealed that none of the significant predictors from models of the original, turn-related data can be explained by random looking.
@@
 The effects of language condition and transition type for the real and randomly permuted data can also be observed in Figure \ref{fig:E1-randvsreal}.
+We excluded the output of random-permutation models that did not ultimately converge to remove unreliable model results from our percentile calculations below (78\% and 76\% of models for children and adults, respectively).
 
@@
 
-% PART UNFINISHED: can add a link at the end of the prior paragraph to video examples
-
 The duration of turn transitions and the number of speaker changes across videos was variable because the conversations were recorded semi-spontaneously.
@@
 
-% PART UNFINISHED: update gap information
 There were 85 turn transitions for analysis after excluding transitions longer than 550 msec and shorter than 90 msec.
@@
 
+\begin{scriptsize}
 \begin{longtable}{lcccl}
@@
     Age                         & 0.02543 &    0.10260  &  0.248   & 0.8042        \\
-    Condition$=$\textit{muffled}         &   -1.63429  &   0.86390 &  -1.892   & 0.0585 .      \\
-    Condition$=$\textit{normal}            &  0.36710 &    0.43296 &   0.848  &  0.3965        \\
-    Condition$=$\textit{robot}            &  -0.26741 &    0.59071 &  -0.453  &  0.6508        \\
     Type$=$\textit{non-Question}                      & -0.81873  &   0.59985 &  -1.365  &  0.1723        \\
     Duration                   &  4.17672  &   0.62446  &  6.689 & 2.25e-11 ***    \\
+    Age*Type$=$\textit{non-Question}                &    0.15116   &  0.13643  &  1.108  &  0.2679        \\
+	\hline
+    Condition$=$\textit{normal}            &  0.36710 &    0.43296 &   0.848  &  0.3965        \\
+    Age*Condition$=$\textit{normal}       &   0.12919 &    0.10227  &  1.263  &  0.2065        \\
+    Condition$=$\textit{normal}*     &   0.91059  &   0.72095  &  1.263 &   0.2066        \\
+    \hspace*{5mm} Type$=$\textit{non-Question} &&&& \\
+    Age*Condition$=$\textit{normal}* &   -0.37542  &   0.16963  & -2.213 &   0.0269 *      \\
+    \hspace*{5mm} Type$=$\textit{non-Question} &&&& \\
+	\hline
+    Condition$=$\textit{muffled}         &   -1.63429  &   0.86390 &  -1.892   & 0.0585 .      \\
     Age*Condition$=$\textit{muffled}       &   0.39317 &     0.18907 &   2.080  &  0.0376 *      \\
-    Age*Condition$=$\textit{normal}       &   0.12919 &    0.10227  &  1.263  &  0.2065        \\
+    Condition$=$\textit{muffled}*     &  1.77190  &   1.24864  &  1.419 &   0.1559        \\
+    \hspace*{5mm} Type$=$\textit{non-Question} &&&& \\
+    Age*Condition$=$\textit{muffled}* &  -0.47057 &    0.28703 &  -1.639 &   0.1011        \\
+    \hspace*{5mm} Type$=$\textit{non-Question} &&&& \\
+	\hline
+    Condition$=$\textit{robot}            &  -0.26741 &    0.59071 &  -0.453  &  0.6508        \\
     Age*Condition$=$\textit{robot}           &   0.13740  &   0.13568   & 1.013  &  0.3112        \\
-    Age*Type$=$\textit{non-Question}                &    0.15116   &  0.13643  &  1.108  &  0.2679        \\
-    Condition$=$\textit{muffled}*     &  1.77190  &   1.24864  &  1.419 &   0.1559        \\
-    Type$=$\textit{non-Question} &&&& \\
-    Condition$=$\textit{normal}*     &   0.91059  &   0.72095  &  1.263 &   0.2066        \\
-    Type$=$\textit{non-Question} &&&& \\
     Condition$=$\textit{robot}*     &   -1.02193  &   1.01227  & -1.010  &  0.3127        \\
-    Type$=$\textit{non-Question} &&&& \\
-    Age*Condition$=$\textit{muffled}* &  -0.47057 &    0.28703 &  -1.639 &   0.1011        \\
-    Type$=$\textit{non-Question} &&&& \\
-    Age*Condition$=$\textit{normal}* &   -0.37542  &   0.16963  & -2.213 &   0.0269 *      \\
-    Type$=$\textit{non-Question} &&&& \\
+    \hspace*{5mm} Type$=$\textit{non-Question} &&&& \\
     Age*Condition$=$\textit{robot}*  &   0.08946  &   0.22349   & 0.400 &   0.6890        \\
-    Type$=$\textit{non-Question} &&&& \\
+    \hspace*{5mm} Type$=$\textit{non-Question} &&&& \\
+
     \hline
@@
     (Intercept)           &  -3.4557   &   0.7199 &  -4.800 & 1.58e-06 *** \\ 
-    Condition$=$\textit{muffled}     &     0.3349  &    0.8965  &  0.374 & 0.708692     \\ 
-    Condition$=$\textit{normal}    &      1.2556   &   0.5633  &  2.229 & 0.025805 *   \\ 
-    Condition$=$\textit{robot}       &     1.5938  &    0.7208 &   2.211 & 0.027023 *   \\ 
     Type$=$\textit{non-Question}                  &   0.4292  &    0.6089  &  0.705 & 0.480916     \\ 
     Duration             &     4.7500   &   1.2480  &  3.806 & 0.000141 *** \\ 
+	\hline
+    Condition$=$\textit{normal}    &      1.2556   &   0.5633  &  2.229 & 0.025805 *   \\ 
+    Condition$=$\textit{normal}* &   -0.9452  &    0.7631 &  -1.239 & 0.215475     \\ 
+    \hspace*{5mm} Type$=$\textit{non-Question} &&&& \\
+	\hline
+    Condition$=$\textit{muffled}     &     0.3349  &    0.8965  &  0.374 & 0.708692     \\ 
     Condition$=$\textit{muffled}*  &  0.6627  &    1.2138 &   0.546 & 0.585108     \\ 
-    Type$=$\textit{non-Question} &&&& \\
-    Condition$=$\textit{normal}* &   -0.9452  &    0.7631 &  -1.239 & 0.215475     \\ 
-    Type$=$\textit{non-Question} &&&& \\
+    \hspace*{5mm} Type$=$\textit{non-Question} &&&& \\
+	\hline
+    Condition$=$\textit{robot}       &     1.5938  &    0.7208 &   2.211 & 0.027023 *   \\ 
     Condition$=$\textit{robot}*   &  -1.1265  &    0.9109  & -1.237 & 0.216201     \\ 
-    Type$=$\textit{non-Question} &&&& \\
-    \hline
+    \hspace*{5mm} Type$=$\textit{non-Question} &&&& \\
+    \hline
+    &&&& \\
   \caption{Model output for children and adults' anticipatory gaze switches.}
@@
 \end{longtable}
+\end{scriptsize}
 
@@
 The effects of language condition and transition type for the real and randomly permuted data can also be observed in Figure \ref{fig:E1-randvsreal}.
-
-% PART UNFINISHED (potentially)
-% We excluded the output of random-permutation models that did not converge to remove unreliable model results from our percentile calculations below (XX\% and XX\% of models for children and adults, respectively).
+We excluded the output of random-permutation models that did not ultimately converge to remove unreliable model results from our percentile calculations below (69\% and 70\% of models for children and adults, respectively).
 
@@
 The close link between prosodic and syntactic structure makes it difficult to tease apart how predictive processing in one domain is distinct from the other in turn prediction.
-That being said, compared to prosodic contours (e.g., final rising intonation), lexicosyntactic cues like \textut{wh}-words, \textit{do}-insertion, and subject-auxiliary inversion are frequent, categorical, and early-occurring in the utterance.
+That being said, compared to prosodic contours (e.g., final rising intonation), lexicosyntactic cues like \textit{wh}-words, \textit{do}-insertion, and subject-auxiliary inversion are frequent, categorical, and early-occurring in the utterance.
 Children may therefore have an easier time picking out and interpreting lexical cues to questionhood on the fly.
 Question turns started yielding significantly more anticipatory gazes by age 3--4 in the normal speech condition of Experiment 2, by which time children frequently hear and use a variety of polar \textit{wh}-questions \citep{clark2009}.
-Furthermore, while lexicosyntactic question cues were available on every instance of \textit{wh}- and \textit{yes/no} questions in our stimuli, prosodic question cues were only salient on \textit{yes/no} questions and, even then, the mapping of prosodic contour to speech act (e.g., high final rises for polar questions) is far from one-to-one \citep{hedberg2010}.
-
-Finally, we found that our adult controls were affected by the absence of prosody (within the question effect) while Keitel et al.
-(2013) found that adults' were unaffected by flattened pitch.
-But again, because we flattened pitch for individual speakers and controlled syllable duration, we are not able to directly compare our findings to theirs.
-
-% WHAT CUES ARE THEY USING?
+Furthermore, while lexicosyntactic question cues were available on every instance of \textit{wh}- and \textit{yes/no} questions in our stimuli, prosodic question cues were only salient on \textit{yes/no} questions and, even then, the mapping of prosodic contour to speech act (e.g., high final rises for polar questions) is far from one-to-one.
 
 \subsubsection{The question effect}
-% What exactly is the Q effect? Could it exist for non-Qs?
-% When does the Q effect emerge and exactly what cues is it relying on?
-% Are they predicting the form of the answer?
-
-Third, we found that anticipatory looking was primarily driven by question transitions, a pattern that has so far not been discussed in other observer gaze studies, on children or adults (Keitel et al., 2013; Hirvenkari, 2013; Tice and Henetz, \citeyear{TiceHenetz11}).
-The question effects in the current study were critical to understanding the role of lexicosyntactic information in participants' turn predictions.
-
-In both of our experiments we found a robust advantage for questions in participants' anticipatory switching because participants made more switches after questions than non-questions.
-Children and adults both treated questions as if they had a special interactive status, and with good reason: Questions make a speaker switch immediately relevant, helping the listener to predict what will happen next (i.e., an answer from the addressee).
-Linguistic cues to questionhood often occur early in the utterance, making questions easy to identify early on and thereby giving observers time to plan a reaction (i.e., a gaze switch).
-Also, the form of a question can constrain the type of response that will come next (e.g., a location after a \textit{where} question), further increasing the predictability of the interactional sequence.
-
-Children get frequent and early experience with hearing and answering questions.
-Questions make up approximately one third of the utterances children hear, before and after the onset of speech, and even into their preschool years \citep{fitneva2012, henning2005, shatz1979}.\footnote{There is substantial variation in this proportion by individual and socioeconomic class \citep{hart1992}.} Many of the questions directed to young children are ``test' questions---questions that the caregiver already has the answer to (e.g., ``What does a cat say?'').
-Caregivers use questions to get their young children's attention and to ensure that information is in common ground, even if the responses are non-verbal or infelicitous \citep{bruner1985, fitneva2012, snow1977}.
-So, in addition to having a special interactive status, questions are a core characteristic of many caregiver-child interactions.
-
-We cannot definitively say with our current data what property of questions caused children to switch their gaze more often.
-It is true that questions make a speaker transition more likely and that questions have distinctive linguistic cues, both early and late, to identify themselves.
-But other speech acts and request formats have these same properties.
-Imperatives, compliments, and complaints all make a response from the addressee likely \citep{schegloff2007}.
-Rhetorical and tag questions, on the other hand, take a similar form to prototypical polar questions, but do not always require an answer.
-So, though it is clear that children anticipate responses more often for questions than non-questions, their actual understanding of questionhood as a linguistically- or interactionally-marked type of speech act still needs further exploration.
-
-Our results suggest that the question effect emerges between ages one and two, but further testing is needed to establish whether this is the case.
-For example, the size and emergence of the question effect for various question types and question cues could differ (\textit{wh-} vs.
-\textit{yes-no}; early vs.
-late in the utterance), but we do not have enough data to make those distinctions in the current analysis.
-Also, 43\% of one-year-olds already showed more switching for questions than non-questions in the \textit{normal} speech condition---the only condition with the full range of linguistic cues children hear in everyday life.
-Future work with more fine-grained age samples and better controlled \textit{normal} speech might then find that children can identify questions in conversation even earlier.
-
-Because the form of an answer is partially determined by its question (e.g., locative expressions for ``where'' questions), participants could even be predicting what type of response they will hear, and so they are looking anticipatorily at the responder for confirmation of their prediction.
-This interpretation is initial, but in-line with work showing early response planning and speech act recognition \citep{gisladottir2015, bogels2015}.
+
+In both experiments, anticipatory looking was primarily driven by question transitions, a pattern that had not been previously reported in other observer gaze studies, on children or adults (Keitel et al., 2013; Hirvenkari, 2013; Tice and Henetz, \citeyear{TiceHenetz11}).
+Questions make a speaker switch immediately relevant, helping the listener to predict with high certainty what will happen next (i.e., an answer from the addressee).
+As mentioned, linguistic cues to questionhood also often occur early in the utterance, giving observers time to plan a gaze switch.
+Because the form of a question can constrain the type of response that will come next (e.g., a location after a \textit{where} question), questions can even help listeners predict specific upcoming content in the next turn.
+
+Our results suggest that question turns start significantly affecting children's predictions between ages three and four, but further testing with finer-grained age samples with stimuli focused on specific communicative acts and linguistic cues is needed to better sketch out the developmental trajectory of this effect.
+For example, the effect size and age of emergence might differ by question type (e.g., \textit{wh-} vs.
+\textit{yes-no}) or the location of question-identifying cues within the unfolding utterance (early vs.
+late), but we would not be able to see it given the current design.
+
+Prior work on children's acquisition of questions indicates that they may already have some understanding about question-answer  sequences by the time they begin to speak: Questions make up approximately one third of the utterances children hear, before and after the onset of speech, and even into their preschool years, even though the types and complexity of questions change throughout development \citep{casillas2016, fitneva2012, henning2005, shatz1979}.\footnote{There is substantial variation question frequency by individual and socioeconomic class \citep{hart1992}.} For the first few years, many of the questions directed to children are ``test' questions---questions that the caregiver already has the answer to (e.g., ``What does a cat say?''), but this changes as children get older.
+Questions help caregivers to get their young children's attention and to ensure that information is in common ground, even if the responses are non-verbal or infelicitous \citep{bruner1985, fitneva2012, snow1977}.
+So, in addition to having a special interactive status (for adults and children alike), questions are a core characteristic of many caregiver-child interactions, motivating a general benefit for questions in turn structure anticipation.
+
+All that being said, our current data do not tell us what it is about questions that makes children and adults more likely to anticipatorily switch their gaze to addressees.
+Other request formats, such as imperatives, compliments, and complaints make a response from the addressee highly likely in the next turn \citep{schegloff2007}.
+Rhetorical and tag questions, on the other hand, take a similar form to prototypical polar questions, but often do not require an answer.
+So, though it is clear that participants anticipated responses more often for questions than non-questions, we do not yet know whether their predictive action is limited to turns formatted as questions or is generally applicable to turn structures that project an immediate response from the addressee.
+
+Much recent work on prediction during turn taking has focused on participants' use of linguistic cues to predict the end of the current turn \citep{torreira2015, magyari2012, de-ruiter2006, ford1996, duncan1972}, in some cases finding that lexical information was sufficient for prediction.
+Our current results suggest that, spontaneous predictions are instead driven by predictions about what is \textit{beyond} the end of the current turn---that questions are sufficient for prediction.
+
+To integrate these results and understand how listeners make predictions for turn taking, it is crucial to account for the participants' role in the interaction.
+The results we present here are based on predictions about third-party conversation, which enables participants to follow interactions with no chance of actually participating.
+Although recent work has shown that similar anticipatory eye gazes do occur in spontaneous conversation \citep{holler2015}, we do not yet know if the same question advantage occurs, or which linguistic cues seem to drive it.
+It is possible that participants track conversation for cues to upcoming opportunities/obligations to speak and, when one is found, focus more attention on predicting the precise timing of the current turn's end.
+To answer these questions we will need to innovate first-person prediction measures that can be used in real-time interaction.
 
 \subsubsection{Early competence for turn taking?}
-% When does children's performance become adult-like?
-
-Though children's anticipatory gazes did show some gains with age, adults on average still made more anticipatory switches than children in all conditions, even compared to 6-year-olds.
-Curiously, child and adult performance were much more similar by ages 3--5 in Experiment 1.
+
+One of the core aims of our study was to test whether children show an early competence for turn taking, as is proposed by studies of spontaneous mother-infant interaction and theories about the mechanisms underlying human interaction \citep{hilbrink2015, levinson2006}.
+Although children and adults' gaze patterns were quite similar in Experiment 1, children in Experiment two showed change with age in their anticipatory looking, even in the most prototypical speech condition (``normal'' speech), and 6-year-olds still did not achieve adult-like levels of anticipatory looking on average.
 This may indicate that children rely more in non-verbal cues in anticipating turn transitions or, alternatively, that adults are better at flexibly adapting to the turn-relevant cues present at any moment.
-
-First, we found that children begin making above-chance anticipatory switches much earlier than three---our one-year-olds were already looking above chance in two of our four linguistic conditions in Experiment 2.
-This earlier emergence of anticipatory switching could be due to our more child-friendly stimuli, which we designed to be engaging and to mimic the kind of speech children often experience interactionally.
-Second, because we included more linguistic conditions to isolate the effects of prosody and lexicosyntax, we gained a broader perspective on how linguistic information changed children's predictions.
-Ultimately, the added conditions allowed us to find an early advantage for prosody and a later advantage for lexicosyntax.
-
-Our findings are consistent with observational work on children's turn taking.
-Young children are slow turn-takers in spontaneous conversation \citep{casillas2016} even though they begin taking turns in infancy \citep{hilbrink2015, jaffe2001}.
-We found that children begin to spontaneously predict upcoming turn structure at age one.
-Children's gaze switches were also most common in the inter-turn gap, suggesting that they expected an immediate response from the addressee.
-Our results therefore support the idea of an early-developed competence for the timing of turn taking.
-But because children do not themselves use adult-like conversational timing until much later \citep[age six; ][]{ervin-tripp1979}, our results also indirectly point toward response planning as a key factor contributing to children's slow responses in real-time interaction \citep{casillas2016}.
-
-\subsection{Limitations}
-% assumptions about saccadic planning time
-% There is a problem with manipulated speech
-
-We controlled linguistic information in Experiment 2 by using phonetically-manipulated speech, extending speech manipulations used in prior work \citep{de-ruiter2006, keitel2013}.
-While phonetic manipulation gave us control over the linguistic signal, it also created speech sounds that children don't usually hear in their natural environment.
+When they \textit{did} have full linguistic information, children on average still made above-chance anticipatory gazes to responders, suggesting that at least some turn-taking competence for third-party conversation is present in infancy, though it is a weak effect.
+
+Taken together, the data suggest that turn-taking skills do begin to emerge in infancy, but that adult-like competence is not achieved until much later, mirroring results from children's spontaneous interactions with their caregivers (infants: \citealp{hilbrink2015, jaffe2001, snow1977}; older children: \citealp{casillas2016, garvey1984, ervin-tripp1979}).
+It is possible, however, that first-person measures of children's predictions would show more frequent turn structure anticipations at younger ages.
+
+\subsection{Limitations and future work}
+
+Although Experiments 1 and 2 have offered new findings regarding the relation to speech act in online turn predictions and the emergence of those predictions in the first six years, there remain at least two major limitations to our work: speech naturalness and participant role.
+
+Following prior work \citep{de-ruiter2006, keitel2013}, we used phonetically manipulated speech in Experiment 2, resulting in speech sounds that children don't usually hear in their natural environment.
 Many prior studies have used phonetically-altered speech with infants and young children \citep[cf.][]{jusczyk2000}, but almost none of them have done so in a conversational context.
-We found that the \textit{normal} speech condition showed the strongest anticipation effects over all, and we have attributed this to the fact that it gave participants full (lexicosyntactic and prosodic) linguistic information.
-However, it was also the only natural speech condition.
 Children could have had trouble processing the \textit{words only} and \textit{prosody only} conditions because they were unfamiliar, and not just because they had less linguistic information available.
-
-It would take a different type of stimulus to compare lexical and prosodic information without phonetic manipulation.
-For example, carefully scripted or cross-spliced conversations could control for the presence or absence of prosodic and lexicosyntactic cues to turn transition with natural speech.
-But since the non-natural (limited cue) linguistic conditions did still elicit some differences in anticipation, children must have processed at least some linguistic information in the \textit{prosody only} and \textit{words only} conditions.
-Perhaps the advantage of the \textit{normal} condition was then not entirely due to its naturalness.
-We expect that future work with spontaneous conversational measures will still find an advantage for combined lexical and prosodic cues over lexical cues alone \citep[see also][]{duncan1972, ford1996, torreira2015}.
-
-In Experiment 2, we tried to improve our method for controlling linguistic information by phonetically manipulating improvised dialogic speech.
-Although this decision allowed us to follow prior work more closely \citep{de-ruiter2006, keitel2013}, it also limits our interpretation of children's behavior.
-Low-pass filtered and robotic-like speech are not typical within children's speech environment, and so it may have been unfair to compare the limited cue conditions in Experiment 2 to the normal speech condition, which not only had full linguistic cues, but is also more typically present in children's speech environments.
-One alternative to explore for the future would be scripted speech, cross-spliced speech, or speech that is carefully selected from a large corpus, given a set of specific linguistic constraints.
-
-Relatedly, since we used improvised conversation recordings, we did not control how often the speakers switched back and forth, or how often a switch occurred after a pause in speech.
-It is possible that, to some extent, children could have formed expectations about the turn structure on the basis of the frequent switching alone \citep[see, e.g., ][]{thorgrimsson2015}.
-If this statistical expectation were the only thing leading children's anticipatory gaze switches, we would have predicted more uniformity in looking behavior across the conditions and no major effects of linguistic signal, age, or transition type.
-But that is not what we saw.
-Nevertheless, another advantage of using scripted conversation would be the ability to control switch frequency and eliminate this issue.
-
-Finally, we carefully designed a scheme for identifying anticipatory turn switches in the data, building off of Keitel and colleagues' (2013) analyses.
-In both our study and theirs, anticipatory switches that happen very early in the turn are not counted.
-Instead, they would actually count as ``random'' switches in the baseline analyses.
-These same early switches are counted as non-random in other studies of children's gaze to conversational stimuli \citep{bakker2011, hofsten2009}.
-Since most studies, even those using a wider switching window \citep[e.g., ][]{hirvenkari2013} find that switches primarily happen close to or immediately after the inter-turn gap begins, the approach used here and in Keitel et al.
-(2013) probably captures most of participants' turn-relevant switching behavior.
-But if early looks do indeed increase as children get older and have more access to early syntactic cues, it would be helpful to develop a second measure of very early anticipatory looks.
-
-Despite extreme differences in linguistic information across conditions, independent effects of linguistic information (outside of interactions) were minimal in our experiments.
-But, using a button-press measure with similar stimuli, de Ruiter and colleagues (2006) found robust effects of linguistic condition.
-This disparity in findings could be partially explained by task differences.
-A button-press method explicitly encourages participants to make an early response, while observer gaze measures are purely passive.
-Participants' spontaneous anticipatory gaze switches may not elicit enough early responses to see large linguistically-driven differences in prediction.
-Indeed, most switches happened during the inter-turn gap (Keitel et al., 2013; Hirvenkari, 2013; Tice and Henetz, \citeyear{TiceHenetz11}).
-Observers' gaze patterns suggest that they aren't targeting potential turn-ends with their predictions.
-Rather, they are targeting switches in speakership.
-So although button press measures may be better equipped to tell us what participants \textit{can} do in predicting turn ends, they may also emphasize early responding (and consequently, lexicosyntactic cues) to a greater extent than we experience in everyday conversation.
-Within observer gaze experiments, questions may be the key to naturally eliciting early responses.
-We have seen in the current study that questions are interactionally sufficient to motivate early predictions.
+Future work could instead carefully script or cross-splice parts of turns to control for the presence or absence of linguistic cues for turn transition.
+
+The prediction measure we present is based on an observer's view of third-party conversation but, because participants' role in the interaction could affect their online predictions about turn taking, an ideal experimental measure would capture first-person behavior.
+First-person measures of spontaneous turn prediction will be key to revealing how participants distribute their attention over linguistic and non-verbal cues while taking part in everyday interaction, the implications of which relate to theories of online language processing for both language learning and everyday talk.
 
 \subsection{Conclusions}
-%
-%Conversation plays a central role in children's language learning.
+
+Conversation plays a central role in children's language learning.
 It is the driving force behind what children say and what they hear.
-And yet many conversational skills, like turn taking, rely heavily on language.
-Adults use language to accurately predict turn structure in conversation, which facilitates their online comprehension and allows them to respond relevantly and on time.
-But young children do not have enough linguistic knowledge to take turns the way that adults do.
-Instead, they first acquire a preverbal turn-taking system for interaction, and then somehow assimilate language into that system as they acquire it.
-The simultaneous acquisition and integration of interactive and linguistic skills is no easy task, as evidenced by children's protracted development of turn-taking skills.
-In the current study we have investigated children's language acquisition through the lens of turn taking, thereby bringing into focus children's application of linguistic knowledge in real-time interaction.
-We found that children acquire basic predictive skills early on, but that the integration of linguistic cues takes time and is primarily driven by sequences of action---in our case, by questions and their answers.
+Adults use linguistic information to accurately predict turn structure in conversation, which facilitates their online comprehension and allows them to respond relevantly and on time.
+In the current study we have investigated how children's predictions about turn structure changes as their linguistic skills develop in the first six years.
+We found that, although some basic knowledge about turn taking exists at age one, the integration of linguistic cues into children's predictions about turn structure takes time and is, like adults, primarily driven by sequences of action---in our case, by questions and their answers.
 
@@
 
-% PART UNFINISHED: APPENDIX
+\appendix
+\setcounter{figure}{0}  
+In all of the following plots, the gray dots represent the randomly permuted data's model estimates for the value listed (beta or standard error), the white dots represent the model estimates from the original data, and the triangles represent the 95th percentile for each distribution being shown.
+
+\begin{figure}
+  \centering
+  \subfloat[Children]{\includegraphics[width=0.49\textwidth]{figures/E1-chi-randrun-betas-absolute.png}\label{fig:E1-ChiBs}}
+  \hfill
+  \subfloat[Adults]{\includegraphics[width=0.49\textwidth]{figures/E1-adu-randrun-betas-absolute.png}\label{fig:E1-AduBs}}
+  \caption{Random-permutation and original $|$\textit{$\beta$}-values$|$ for predictors of children and adults' anticipatory gaze rates in Experiment 1.}
+\end{figure}
+
+\begin{figure}
+  \centering
+  \subfloat[Children]{\includegraphics[width=0.49\textwidth]{figures/E1-chi-randrun-SEs-absolute.png}\label{fig:E1-ChiSEs}}
+  \hfill
+  \subfloat[Adults]{\includegraphics[width=0.49\textwidth]{figures/E1-adu-randrun-SEs-absolute.png}\label{fig:E1-AduSEs}}
+  \caption{Random-permutation and original $|$\textit{SE}-values$|$ for predictors of children and adults' anticipatory gaze rates in Experiment 1.}
+\end{figure}
+
+\begin{figure}
+  \centering
+  \subfloat[Children]{\includegraphics[width=0.49\textwidth]{figures/E2-chi-randrun-betas-absolute.png}\label{fig:E2-ChiBs}}
+  \hfill
+  \subfloat[Adults]{\includegraphics[width=0.49\textwidth]{figures/E2-adu-randrun-betas-absolute.png}\label{fig:E2-AduBs}}
+  \caption{Random-permutation and original $|$\textit{$\beta$}-values$|$ for predictors of children and adults' anticipatory gaze rates in Experiment 2.}
+\end{figure}
+
+\begin{figure}
+  \centering
+  \subfloat[Children]{\includegraphics[width=0.49\textwidth]{figures/E2-chi-randrun-SEs-absolute.png}\label{fig:E2-ChiSEs}}
+  \hfill
+  \subfloat[Adults]{\includegraphics[width=0.49\textwidth]{figures/E2-adu-randrun-SEs-absolute.png}\label{fig:E2-AduSEs}}
+  \caption{Random-permutation and original $|$\textit{SE}-values$|$ for predictors of children and adults' anticipatory gaze rates in Experiment 2..}
+\end{figure}
 

### 2015-12-03 paper/Letters/ResponseLetter-Dec15.tex
draft of letter with responses, extra supplementary graphs

\documentclass[a4paper]{letter}
\pagenumbering{gobble}

%----------------------------------------------------------------------------------------
%	DOCUMENT MARGINS
%----------------------------------------------------------------------------------------

\textwidth 6.75in
\textheight 9.25in
\oddsidemargin -.25in
\evensidemargin -.25in
\topmargin -1in
\longindentation 0.50\textwidth
\parindent 0.4in

\signature{Marisa Casillas \\ marisa.casillas$@$mpi.nl}
\address{Wundtlaan 1 \\ 6525 XD Nijmegen \\ The Netherlands}

\begin{document}
\begin{letter}{}

\opening{Dear Profs. Dr. Pickering and Dr. Gerrig,}

We resubmit our manuscript on ``The development of children's ability to track and predict turn structure in conversation'' to the \textit{Journal of Memory and Language}. We apologize for our delay in getting the manuscript resubmitted; it took quite a long time to implement one of the suggestions from Reviewer 3, mostly because I had to learn quite a bit of new technical skill to carry his or her suggested analysis out on our computation-heavy data set. In addition to that, we have included a two other follow-up analyses in the Supplementary materials, both inspired by the other reviewers' comments. We have tried to make the text more succinct and the results and discussion sections more distinct and straightforward, following requests from all three reviewers. We hope that, with these changes, we have greatly improved the paper's clarity. Please find a description of specific comments addressed below. Thank you very much for your reconsideration. Please let us know if there is any further information that you require about the submission.

\textbf{Manuscript identifying details:}

\begin{tabular}{ll}
\textit{ID:} & JML-14-235 \\
\textit{Title:} & The development of children's ability to track and \\
& predict turn structure in conversation \\
\textit{Authors:} & Marisa Casillas\textsuperscript{a} and Michael C. Frank\textsuperscript{b} \\
\textit{Affiliations:} & a) Max Planck Institute for Psycholinguistics and \\
& b) Stanford University
\end{tabular}

\closing{Sincerely,}

We thank all three reviewers for their thorough and helpful comments on the previously submitted manuscript. We have tried to address all of the reviewers' comments, within the manuscript and/or here in the response letter.

\textbf{Reviewer 1}

\textit{1. 333ms may be too long for older children's saccadic planning time and may therefore inflate the anticipatory looking rates for older children.}

We were concerned about this, too, but we unfortunately did not collect saccadic planning norms for children ages 1--6. We did not feel comfortable (semi-)arbitrarily assigning saccadic planning times to different ages, so in the newly analyzed data we take the conservative approach of using adult-like planning times (200ms) for children at all ages in both experiments.

\smallskip

\textit{2. The description of the random baseline permutations was not clear.}

We have tried our best to clarify this description (Section XX) and have added specific pieces of information where requested.

\smallskip

\textit{3. Are anticipatory gaze shifts driven by boredom (and not by anticipation)? Do children generally just look away as the speaker goes on?}

This alternative hypothesis makes a valuable test of what we measured in the experiments. In the supplementary materials you can now find a graph comparing theoretically boredom-driven lookers and our real lookers. We modeled the proportion of participants looking at the current speaker for our real data and for (fake) boredom-driven lookers (who look away from the current speaker at a constant rate, starting 1 second after the onset of speech) across turns of different length. Even in short turns, where there is very little temporal room for children to behave differently from the boredom-driven lookers, there is a clear difference between the boredom-driven and the real data. (Figure XX). We take this to mean that children are unlikely to be looking away simply due to boredom.

\smallskip

\textit{4. The reported results form the models models are hard to follow.}

Thank you for pointing this out! We have added a little more information about how each variable was coded and have put summary tables of every model's output into the text.

\smallskip

\textit{5. ``No speech'' should be the reference level against which the linguistic conditions are compared in Experiment 2.}

This is a very useful comment, thanks! We agree and have changed our analyses to accommodate it. We had started with the ``normal speech'' condition as the reference level because that is what was done in related work on adult turn-end prediction. In the end, however, we too preferred the additive model, and have changed the text accordingly.

\newpage

\textit{6. The children appear to not use linguistic information at all.}

XX

\smallskip

\textit{7. It is too hard to link the conclusions in the discussion back to the statistical results and the effects themselves are not talked about in a clear enough way.}

We have tried our best to make this clearer, first by clarifying the statistical analyses and outputs, and second by trying to explicitly link each claim to an individual result (through the use of figure and table references). We hope that this will help readers better connect the results to our interpretations.

\smallskip

\textit{8. There are too many references to unpublished work.}

The single upside to taking so long in implementing these comments is that all of the unpublished work referred to previously is now published or in press!

\smallskip

\textit{Small miscellany has been addressed directly in the text.}

\bigskip

\textbf{Reviewer 2}

\textit{1. It is not clear which claims are backed up by which analyses. Analyses should be carried out for each group separately.}

Thanks! We have tried to improve the linkage between our claims and analyses. Please see responses 4 and 7 under Reviewer 1 for a little more detail. The main focus of our analyses was the effect that age (as a continuous variable) has on anticipatory gaze. However there are two cases in which our main analyses are not satisfying without further, age-group analyses: (1) age sometimes showed itself to interact with other factors (e.g., age and language condition), and (2) without further tests we can not say whether particular age groups themselves statistically differ from chance, across or within conditions. Following your suggestion, we have therefore done two things. First, we added follow-up two-tailed t-tests, making pairwise comparisons between age groups for significant interactions so that we could find out how the interacting factor (e.g., language condition) significantly changes across age groups (Sections XX). Second, we have added individual models of the youngest age groups (ages 1, 2, and 3), comparing their real looking behavior to our randomly permuted data across conditions (Section XX). These follow-up analyses, which are summarized in the text, are primarily described in the supplementary materials (Section XX).

\smallskip

\textbf{Reviewer 3}

\end{letter}
\end{document}

### 2015-12-03 paper/Letters/ResponseLetter-Dec2015.tex
whoops! here's the actual letter draft

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Professional Formal Letter
% LaTeX Template
% Version 1.0 (28/12/13)
%
% This template has been downloaded from:
% http://www.LaTeXTemplates.com
%
% Original author:
% Brian Moses (http://www.ms.uky.edu/~math/Resources/Templates/LaTeX/)
% with extensive modifications by Vel (vel@latextemplates.com)
%
% License:
% CC BY-NC-SA 3.0 (http://creativecommons.org/licenses/by-nc-sa/3.0/)
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%----------------------------------------------------------------------------------------
%	PACKAGES AND OTHER DOCUMENT CONFIGURATIONS
%----------------------------------------------------------------------------------------

\documentclass[11pt,a4paper]{letter} % Specify the font size (10pt, 11pt and 12pt) and paper size (letterpaper, a4paper, etc)

\usepackage{graphicx} % Required for including pictures
\usepackage{microtype} % Improves typography
\usepackage{gfsdidot} % Use the GFS Didot font: http://www.tug.dk/FontCatalogue/gfsdidot/
\usepackage[T1]{fontenc} % Required for accented characters

% Create a new command for the horizontal rule in the document which allows thickness specification
\makeatletter
\def\vhrulefill#1{\leavevmode\leaders\hrule\@height#1\hfill \kern\z@}
\makeatother

%----------------------------------------------------------------------------------------
%	DOCUMENT MARGINS
%----------------------------------------------------------------------------------------

\textwidth 6.75in
\textheight 9.25in
\oddsidemargin -.25in
\evensidemargin -.25in
\topmargin -1in
\longindentation 0.50\textwidth
\parindent 0.4in

%----------------------------------------------------------------------------------------
%	SENDER INFORMATION
%----------------------------------------------------------------------------------------

\def\Who{Marisa Casillas} % Your name
\def\What{} % Your title
\def\Where{MPI for Psycholingusitics} % Your department/institution
\def\Address{Wundtlaan 1} % Your address
\def\CityZip{6525XD Nijmegen, NL} % Your city, zip code, country, etc
\def\Email{E-mail: marisa.casillas@mpi.nl} % Your email address
\def\TEL{Phone: +31 (024) 3521266} % Your phone number
%\def\URL{URL: http://www.johnsmith.com} % Your URL

%----------------------------------------------------------------------------------------
%	HEADER AND FROM ADDRESS STRUCTURE
%----------------------------------------------------------------------------------------

\address{
\includegraphics[width=1in]{logo.jpg} % Include the logo of your institution
\hspace{5.1in} % Position of the institution logo, increase to move left, decrease to move right
\vskip -1.07in~\\ % Position of the text in relation to the institution logo, increase to move down, decrease to move up
\Large\hspace{1.5in}Max Planck Institute \hfill ~\\[0.05in] % First line of institution name, adjust hspace if your logo is wide
\hspace{1.5in}for Psycholinguistics \hfill \normalsize % Second line of institution name, adjust hspace if your logo is wide
%\makebox[0ex][r]{\bf \Who \What }\hspace{0.08in} % Print your name and title with a little whitespace to the right
~\\[-0.11in] % Reduce the whitespace above the horizontal rule
\hspace{1.5in}\vhrulefill{1pt} \\ % Horizontal rule, adjust hspace if your logo is wide and \vhrulefill for the thickness of the rule
\hspace{\fill}\parbox[t]{2.85in}{ % Create a box for your details underneath the horizontal rule on the right
\footnotesize % Use a smaller font size for the details
%\Who \\ \em % Your name, all text after this will be italicized
%\Where\\ % Your department
\Address\\ % Your address
\CityZip\\ % Your city and zip code
\TEL\\ % Your phone number
\Email\\ % Your email address
\\
\today\\
%\URL % Your URL
}
\hspace{-1in} % Horizontal position of this block, increase to move left, decrease to move right
\vspace{-1.2in} % Move the letter content up for a more compact look
}

%----------------------------------------------------------------------------------------
%	TO ADDRESS STRUCTURE
%----------------------------------------------------------------------------------------

\def\opening#1{\thispagestyle{empty}
{\centering\fromaddress \vspace{0.6in} \\ % Print the header and from address here, add whitespace to move date down
\hspace*{\longindentation}\hspace*{\fill}\par} % Print today's date, remove \today to not display it: \hspace*{\longindentation}\today\hspace*{\fill}\par}
{\raggedright \toname \\ \toaddress \par} % Print the to name and address
\vspace{0.4in} % White space after the to address
\noindent #1 % Print the opening line
% Uncomment the 4 lines below to print a footnote with custom text
%\def\thefootnote{}
%\def\footnoterule{\hrule}
%\footnotetext{\hspace*{\fill}{\footnotesize\em Footnote text}}
%\def\thefootnote{\arabic{footnote}}
}

%----------------------------------------------------------------------------------------
%	SIGNATURE STRUCTURE
%----------------------------------------------------------------------------------------

\signature{\Who \What} % The signature is a combination of your name and title

\long\def\closing#1{
\vspace{0.1in} % Some whitespace after the letter content and before the signature
\noindent % Stop paragraph indentation
\hspace*{\longindentation} % Move the signature right
\parbox{\indentedwidth}{\raggedright
#1 % Print the signature text
\vskip 0.65in % Whitespace between the signature text and your name
\fromsig}} % Print your name and title

%----------------------------------------------------------------------------------------

\begin{document}

%----------------------------------------------------------------------------------------
%	TO ADDRESS
%----------------------------------------------------------------------------------------

\begin{letter}{}
%{Prof. Jones\\
%Mathematics Search Committee\\
%Department of Mathematics\\
%University of California\\
%Berkeley, California 12345\\
%}

%----------------------------------------------------------------------------------------
%	LETTER CONTENT
%----------------------------------------------------------------------------------------

\opening{Dear Profs. Dr. Pickering and Dr. Gerrig,}

\smallskip

We resubmit our manuscript on ``The development of children's ability to track and predict turn structure in conversation'' to the \textit{Journal of Memory and Language}. We apologize for our delay in getting the manuscript resubmitted; it took quite a long time to implement one of the suggestions from Reviewer 3, mostly because I had to learn quite a bit of new technical skill to carry his or her suggested analysis out on our computation-heavy data set. In addition to that, we have included a two other follow-up analyses in the Supplementary materials, both inspired by the other reviewers' comments. We have tried to make the text more succinct and the results and discussion sections more distinct and straightforward, following requests from all three reviewers. We hope that, with these changes, we have greatly improved the paper's clarity. Please find a description of specific comments addressed below. Thank you very much for your reconsideration. Please let us know if there is any further information that you require about the submission.

\textbf{Manuscript identifying details:}

\begin{tabular}{ll}
\textit{ID:} & JML-14-235 \\
\textit{Title:} & The development of children's ability to track and \\
& predict turn structure in conversation \\
\textit{Authors:} & Marisa Casillas\textsuperscript{a} and Michael C. Frank\textsuperscript{b} \\
\textit{Affiliations:} & a) Max Planck Institute for Psycholinguistics and \\
& b) Stanford University
\end{tabular}

\closing{Sincerely,}

%----------------------------------------------------------------------------------------
\newpage
%----------------------------------------------------------------------------------------

\noindent We thank all three reviewers for their thorough and helpful comments on the previously submitted manuscript. We have tried to address all of the reviewers' comments, within the manuscript and/or here in the response letter.

\medskip

\noindent \textbf{Reviewer 1}

\noindent \textit{1. 333ms may be too long for older children's saccadic planning time and may therefore inflate the anticipatory looking rates for older children.}

We were concerned about this, too, but we unfortunately did not collect saccadic planning norms for children ages 1--6. We did not feel comfortable (semi-)arbitrarily assigning saccadic planning times to different ages, so in the newly analyzed data we take the conservative approach of using adult-like planning times (200ms) for children at all ages in both experiments.

\smallskip

\noindent \textit{2. The description of the random baseline permutations was not clear.}

\noindent We have tried our best to clarify this description (Section XX) and have added specific pieces of information where requested.

\smallskip

\noindent \textit{3. Are anticipatory gaze shifts driven by boredom (and not by anticipation)? Do children generally just look away as the speaker goes on?}

\noindent This alternative hypothesis makes a valuable test of what we measured in the experiments. In the supplementary materials you can now find a graph comparing theoretically boredom-driven lookers and our real lookers. We modeled the proportion of participants looking at the current speaker for our real data and for (fake) boredom-driven lookers (who look away from the current speaker at a constant rate, starting 1 second after the onset of speech) across turns of different length. Even in short turns, where there is very little temporal room for children to behave differently from the boredom-driven lookers, there is a clear difference between the boredom-driven and the real data. (Figure XX). We take this to mean that children are unlikely to be looking away simply due to boredom.

\smallskip

\noindent \textit{4. The reported results form the models models are hard to follow.}

\noindent Thank you for pointing this out! We have added a little more information about how each variable was coded and have put summary tables of every model's output into the text.

\smallskip

\noindent \textit{5. ``No speech'' should be the reference level against which the linguistic conditions are compared in Experiment 2.}

\noindent This is a very useful comment, thanks! We agree and have changed our analyses to accommodate it. We had started with the ``normal speech'' condition as the reference level because that is what was done in related work on adult turn-end prediction. In the end, however, we too preferred the additive model, and have changed the text accordingly.

\noindent \textit{6. Looking at the data, it seems possible that the children are not using any linguistic information.}

\noindent Thanks for asking for clarification here. We have tried to integrate aspects of our answer here into the manuscript. The lack of simple language condition effects in the children's data in Experiment 2 does not imply that they were not using linguistic information at all in their predictions (i.e., solely relying on visual information in the stimuli and the duration of inter-turn gap). An unexpected but quite important take away of our findings is that children and adults are far more likely to make an anticipatory gaze switch after hearing a question. Because speech act seems to be so effective in driving participants' anticipations, the extent to which they use linguistic cues is probably limited to their use of those cues to identify questions in the ongoing turn (which in turn allows them to expect an upcoming speaker switch with high certainty). The children's data showed a 2-way effect of age and transition type (question vs. non-question), but only in the ``normal'' speech condition, suggesting that it was the only condition in which they could extract the linguistic cues needed to accurately identify questions, perhaps because prosody and lexical cues to questionhood often work together in forming questions during everyday speech, but perhaps because the manipulated speech was simply too unfamiliar for them to apply their normal interactive prediction processing.

Either way, the existence of the question effect in the normal condition indicates that children do, in fact, use linguistic information (to identify questions) and the interaction of that effect with age suggests that they get better at doing so as they get older. These effects, for example, cannot be explained by duration alone (and its effects are already accounted for in the statistical analyses), and there is no visual cue advantage difference between the normal condition and the others. We are then left to conclude that they use linguistic information to support the question effect. A similar story can be told for the data in Experiment 1: non-verbal cues to questions vs. non-questions are available in both the English and non-English stimuli, but participants (adults and children) made more anticipatory switches when they had access to the lexical information. Again, this cannot be explained by gap duration or unequal visual advantages across the stimuli.

The fact that we have found most of our linguistic processing effects within an effect of speech act paints a very different picture from prior work (e.g., de Ruiter et al., 2006), which finds much clearer evidence of linguistic processing across multiple phonetically manipulated conditions, but without question-non-question comparisons. No turn prediction measure yet is capable of capturing predictions that occur in spontaneous first-person conversation, but we think that our results indicate that there are multiple types of prediction relevant to anticipating upcoming turn structure during interaction, and that any account of how listeners use linguistic cues while participating in conversation should be able to account for shifts in how listeners attend to cues to make more or less specific predictions (i.e., ``what words are upcoming'' vs. ``what types of action are upcoming'').

In sum, while the linguistic effects in our data are somewhat limited, they are present, consistent, developmentally changing, and important to our developing theories about turn prediction. We believe that this sets up future work quite well to dig into the specific linguistic cues (lexical and prosodic, though probably mostly lexical) that are being used in spontaneous question identification.

\smallskip

\noindent \textit{7. It is too hard to link the conclusions in the discussion back to the statistical results and the effects themselves are not talked about in a clear enough way.}

\noindent We have tried our best to make this clearer, first by clarifying the statistical analyses and outputs, and second by trying to explicitly link each claim to an individual result (through the use of figure and table references). We hope that this will help readers better connect the results to our interpretations.

\smallskip

\noindent \textit{8. There are too many references to unpublished work.}

\noindent The single upside to taking so long in implementing these comments is that all of the unpublished work referred to previously is now published or in press!

\smallskip

\noindent \textit{(Reviewer 1's minor points/corrections have all been addressed directly in the text.)}

\bigskip

\noindent \textbf{Reviewer 2}

\noindent \textit{1. It is not clear which claims are backed up by which analyses. Analyses should be carried out for each group separately.}

\noindent Thanks! We have tried to improve the linkage between our claims and analyses and now have presented the statistical results in tables, as requested. Please see responses 4 and 7 under Reviewer 1 for a little more detail. The main focus of our analyses was the effect that age (as a continuous variable) has on anticipatory gaze. However there are two cases in which our main analyses are not satisfying without further, age-group analyses: (a) age sometimes showed itself to interact with other factors (e.g., age and language condition), and (b) without further tests we can not say whether particular age groups themselves statistically differ from chance, across or within conditions. Following your suggestion, we have therefore done two things. First, we added follow-up two-tailed t-tests, making pairwise comparisons between age groups for significant interactions so that we could find out how the interacting factor (e.g., language condition) significantly changes across age groups (Sections XX). Second, we have added individual models of the youngest age groups (ages 1, 2, and 3), comparing their real looking behavior to our randomly permuted data across conditions (Section XX). These follow-up analyses, which are summarized in the text, are primarily described in the supplementary materials (Section XX).

\smallskip

\noindent \textit{2. The results and discussion for each study would be better if they were separated into different sections.}

\noindent Our original intention in combining results and discussion was to make the findings easier to understand. Since we were not able to achieve this in the prior submission, we have followed Reviewer 2's advice and now use a more traditional results-only then discussion-only structure. We think that this has improved the clarity of the paper and we give thanks for the suggestion.

\smallskip

\noindent \textit{3. The methods sections need more elaboration in a few places, and the baseline analysis section could be much clearer.}

\noindent We thank the reviewer for his or her careful attention to our methods section (and its missing details.) We have added in the experimental methods information requested and have re-checked each section for missing information someone might need to replicate our methods. We've also worked to clarify the description of the baseline analysis, including a new figure to accompany it (Section XX; Figure XX). We hope that it is now easier to understand.

\smallskip

\noindent \textit{4. The adult data deserve further discussion because they raise the question of whether this measure taps into linguistic processing at all.}

\noindent We actually think the adults' results are well in-line with prior work showing that (a) adults' predictions primarily rely on lexical cues (E1 \& E2) and (b) adults are better anticipators than young children (E1 \& E2). The complicating factor is that we report a new and important interacting factor with their linguistic processing: speech act. Like the children in our data set, adults made more anticipatory switches following questions, which means that most of our evidence about their use of linguistic cues comes from question transitions. This enriches the picture of linguistic processing for turn prediction, as painted by prior work: linguistic processing is brought into play more often in some contexts than others. From these points we conclude that our experiments do indeed tap into participants' use of linguistic cues to predict  upcoming turn structure. That isn't to say that the method wouldn't be dramatically improved with more natural linguistic controls and tighter controls on gap duration---it would! But we think our current results still make a valuable contribution to the current understanding of turn anticipation in children \textit{and} adults, and lays critical groundwork for discovering what specific types of cues participants spontaneously monitor in everyday interaction.

\smallskip

\noindent \textit{5. There is a lot of unpublished work referred to in the manuscript, with errors in the reference list.}

\noindent These papers are now all published or in press. The errors in the reference list were partly due to half-filled BibTex entries used to produce the previous manuscript. We have checked them again and hope that there are very few (or no) errors remaining.

\bigskip

\noindent \textbf{Reviewer 3}

\noindent \textit{1. The manuscript could be shortened substantially without loss of content.}

\noindent We agree that the paper was overly long. We have significantly shortened the introduction and tried to remove redundancies where possible in the rest of the text. We have also moved some of the secondary analyses to the Supplementary Materials section so that the main text can be read without getting into the nitty gritty of every single analysis. We hope that this has helped make the main text more straightforward and approachable length-wise.

\smallskip

\noindent \textit{2. There is a confound between linguistic condition and puppet pair in Experiment 2.}

\noindent We thank Reviewer 3 for this point. Our design does not fully cross puppet pair (e.g., robots, blue puppets) with linguistic condition (e.g., ``words only'' and ``no speech''). Even though each puppet pair is associated with different conversation clips across children (e.g. robots talking about kitties, birthday parties, or pancakes), robots were only associated with ``words only'' speech, merpeople were only associated with ``prosody only'' speech, and the puppets with fancy clothes were only associated with the ``no speech'' condition. We did this to increase the pragmatic felicity of the experiments for the older children (i.e., robots make robot sounds, merpeople's voices are muffled under water, the fancy-clothed puppets are in a room with main other voices). It is therefore fair to point out a possible confound between linguistic condition and puppet pair. Thankfully, we also ran a short follow-up study at the museum with 3--5-year-olds in which each child only saw one video---the normal speech conversation about birthday parties---with a randomly assigned puppet pair performing the conversation. Five children watched each puppet pair, for a total of 30 children across the six pairs (Figure XX). This experiment holds all things constant except for the appearance of the puppets. We then used a mixed effects logistic regression of children's anticipatory switches (yes or no at each transition), with puppet pair (robots/merpeople/fancy dress/other-3; Figure XX) as a fixed effect and participant and turn transition as random effects. In four versions of this model we systematically varied the reference level to check for differences between every puppet pair, finding no significant affects of puppet type on switching rate. We take this as evidence that, although we did not fully cross puppet pairs and linguistic conditions in Experiment 2, it was unlikely to have had strong effects on children's looking rates above and beyond the intended effects of linguistic condition. We have included details and graphs for this analysis in the Supplementary Materials.

\smallskip

\noindent \textit{3. Computer animation would have been a better choice than puppets.}

\noindent We completely agree with the reviewer on this point. We did not have the ability to create custom animated conversation videos at the time. But, since then, the first author has developed a method for doing exactly that. Unfortunately, for this past work, we are still stuck with the puppet stimuli.

\smallskip

\noindent \textit{4. The model design needs to be further justified and the way the effects are talked about needs clarification.}

\noindent We corrected small mistakes in the model description (thanks for spotting them!), for example, making it clear that we used a logistic regression (for our binary response variable of switch-no switch) and have added short bits of information justifying our use of separate models for adults and children and our use of random slopes. We have also cleaned up the way we talk about the significant effects in the model output because they were sometimes misleading. There was no concern about dummy coding since this is carried out automatically across factor predictors by the glmer package in R (as is now clearer with the presentation of results in tables XX and XX).

\smallskip

\noindent \textit{5. An ideal-observer analysis of the usefulness of the linguistic cues would give a better idea about what can affect anticipatory gaze.}

\noindent We really liked this suggestion but felt that it was too far outside the scope of the current paper, especially because our stimulus set is probably too small to build a generalizable model of cue usefulness. We will, however, seriously consider this type of analysis for future work, with adult and child turn-taking, in experiments and in spontaneous speech corpora.

\smallskip

\noindent \textit{6. The random baseline would be better if it were converted to a full permutation analysis so that the baseline rate of anticipatory gazing is not treated as an additive factor, which it probably is not.}

\noindent We thank Reviewer 3 very much for his/her suggestion and clear explanation of how to conduct the analysis. To be sure about the results we decided to run 5,000 random permutations on the data from each experiment, resulting in 20,000 glmer runs. We apologize for the delay of this resubmission, which is primarily due to the implementation of this random permutation analysis. To analyze our data on this scale, the first author had to learn some new R libraries, optimization techniques, and how to use the local computer cluster, dealing with many technical problems along the way! That being said, we think that the new permutation analysis is indeed simpler and more theoretically intuitive than our prior method (described in XX and XX, visualized in XX and XX). Please note that there were two problems with conducting the analyses as originally instructed in the reviewer's comments: (a) we used \textit{t} values instead of \textit{$\beta$} estimates because the standard error estimates for the randomly permuted data were nearly categorically higher, dramatically so in some cases, and (b) many of the models (especially in Experiment 2) resulted in convergence warnings, which we have dealt with by only reporting results for converging models.

\smallskip

\noindent \textit{7. The theoretical contribution of this paper is not clear.}

\noindent We have tried to improve the clarity of our findings' theoretical importance in the revised manuscript. I summarize here the points that we cover. One of the most important theoretical contributions of our work is in establishing and replicating speech act effects in children and adults' spontaneous predictions---while participants may always use linguistic information to predict upcoming speaker changes, our results to not support the idea that they always use linguistic information to predict upcoming turn ends, as is assumed in metalinguistic measures of turn-end prediction (e.g., pressing a button while listening to speech). Our results suggest that participants' spontaneous predictions are the result of question-monitoring, presumably achieved by recruiting linguistic cues to questionhood from the unfolding signal. The trends in our data suggest that participants are primarily recruiting lexical cues to do this, though establishing this pattern will require follow up work with more focused stimuli.

A second, but related theoretical takeaway is that lexical information alone is not equivalent to full linguistic information for children, as it has been shown to be (XX) and replicated here (XX) with adults. If this effect arose in our data because children do not respond well to phonetically manipulated speech in conversational contexts, this point can be overturned by work that controls linguistic information through other means. If not, it suggests that there is something specially informative about the combined prosodic-lexical marking on question turns, for which there is also prior support (XX) though our finding would be the first to show such effects developmentally.

A third theoretical contribution is that young children (e.g., at ages one and two) do make anticipatory gaze prediction more often than would be expected by chance alone, but not by much. Since older children and adults' gaze shifts are primarily driven by question turns, this may suggest that, although children \textit{can} make predictions about upcoming turn structure, it doesn't count for much until they have acquired the linguistic skill to pick out question turns. This finding, in turn, has potentially important implications for how participant role (first- instead of third-person) and cultural differences (high vs. low parent-infant interaction styles) might affect children's early predictive skill. It also bridges prior work demonstrating that children begin taking non-linguistic turns in infancy (XX), but still have trouble integrating linguistic aspects of responding at age 3--4 (XX).

The fourth contribution is in our interpretation of discrepancies between our methods and results compared to prior work (using more explicit, metalinguistic methods and showing more obvious online predictions using linguistic information). The analyses of these discrepancies led us to put forth a new idea: participants in first-person spontaneous conversation may use multiple strategies in making predictions about upcoming turn structure, using more passive prediction to detect upcoming speaker transitions (e.g., questions), switching into precise turn-end prediction mode when necessary. A flexible system like this allows listeners to continuously monitor ongoing conversation at a low cost while still managing to plan their responses and come in quickly when needed. As far as we know, we are the first to articulate this idea, integrating findings from multiple methods, age groups, and styles of linguistic control.

%----------------------------------------------------------------------------------------

\end{letter}
\end{document}

### 2015-12-03 paper/CasillasFrank-Anticip.tex
intro revisions

@@
 In both experiments, children and adults anticipated more often after hearing questions.
-Like adults, prosody alone did not improve childrenÕs predictive gaze shifts.
+Like adults, prosody alone did not improve children's predictive gaze shifts.
 But, unlike adults, lexical cues alone were also not sufficient to improve prediction---children's performance was best overall with access to lexicosyntax and prosody together.
@@
 Infant turn taking is heavily scaffolded by caregivers, has distinct timing in comparison to adult turn taking, and lacks semantic content \citep{hilbrink2015, jaffe2001}.
-However, children's early, turn-structured social interactions are presumably a critical precursor to their conversational turn taking.
-Along these lines, early non-verbal interactions establish the protocol by which children come to use language with others.
-And as they acquire language, they also come to integrate it into their preverbal turn-taking systems.
+However, children's early, turn-structured social interactions are presumably a critical precursor to their conversational turn taking: early non-verbal interactions likely establish the protocol by which children come to use language with others.
+How do children integrate linguistic knowledge with these preverbal turn-taking abilities?
+
+% And as children acquire language, they also come to integrate it into their preverbal turn-taking systems.
 
@@
 
-\subsection{Turn taking}
+\subsection{Adults' turn taking}
 
@@
 The groups can span turn-by-turn exchanges (e.g., simple question--response, ``How are you?''--``Fine.'') or sequence-by-sequence exchanges (e.g., reciprocals, ``How are you?''--``Fine, and you?''--``Great!'').
-Sequences of action drive the conversation forward into the next, relevant sequences of talk (e.g., ''And you?''--``Great!''--``Why's that?''; \citealp{schegloff2007}).
+
+% Sequences of action drive the conversation forward into the next, relevant sequences of talk (e.g., ''And you?''--``Great!''--``Why's that?''; \citealp{schegloff2007}).
 To take a turn, participants need to make predictions about what conversational content will be relevant next.
@@
 
-Despite this complexity, conversational turn taking is often precise in its timing.
+Despite this complexity, conversational turn taking is often precise in its timing, and this precision requires prediction.
 Across a diverse sample of conversations in 10 languages, one study found a consistent average turn transition time of 0--200 msec at points of speaker switch \citep{stivers2009}.
@@
 
-\subsection{Adults' turn prediction}
+% \subsection{}
 
@@
 The speech snippets were controlled for the amount of linguistic information present; some were normal, but others had flattened pitch, low-pass filtered speech, or further manipulations.
-De Ruiter and colleagues found that, with pitch-flattened speech, the timing of participants' button responses was comparable to their timing with the full linguistic signal.
+With pitch-flattened speech, the timing of participants' button responses was comparable to their timing with the full linguistic signal.
 But when no lexical information was available, participants' responses were significantly earlier.
@@
 
-The role of prosody for adult turn prediction is still a matter of debate.
+Despite this evidence, the role of prosody for adult turn prediction is still a matter of debate.
 De Ruiter and colleagues' (2006) experiment focused on the role of intonation, which is only a partial index of prosody.
@@
 These experimental findings corroborate other corpus and experimental work promoting a combination of cues (lexicosyntactic, prosodic, and pragmatic) as key for accurate turn-end prediction \citep{duncan1972, ford1996, hirvenkari2013}.
-
-In sum, adults accurately and spontaneously make predictions about upcoming turn structure.
+We next turn to evidence on children's developing turn-taking ability.
+
+\subsection{Children's turn prediction}
+
+% Adults accurately and spontaneously make predictions about upcoming turn structure.
 Their predictions rely on a sophisticated body of knowledge about linguistic structure, non-verbal signals, and social actions.
@@
 So a primary role for lexicosyntactic cues doesn't fit well with children's pre-verbal turn taking.
-
-\subsection{Children's turn prediction}
-
-\subsubsection{Observational studies}
+How does children's turn-taking
+
+% \subsubsection{Observational studies}
 
@@
 Children's first turn-like structures appear as early as two to three months in proto-conversation with their caregivers \citep{bruner1975, bruner1985}.
-During proto-conversations, caregivers interact with their infants as if they were capable of making meaningful contributions; they take every look, vocalization, arm flail, and burp as ``utterances'' in the joint discourse \citep{bateson1975, jaffe2001, snow1977}.
+During proto-conversations, caregivers interact with their infants as if they were capable of making meaningful contributions: they take every look, vocalization, arm flail, and burp as ``utterances'' in the joint discourse \citep{bateson1975, jaffe2001, snow1977}.
 Infants catch onto the structure of proto-conversations quickly.
 By three to four months they notice disturbances to the contingency of their caregivers' response and, in reaction, change the rate and quality of their vocalizations \citep{k-bloom1988, masataka1993}.
-Infants at this age also notice changes to social contingency outside of turn structure.
+
+% Infants at this age also notice changes to social contingency outside of turn structure.
 In the Still Face paradigm, caregivers interact with their infants and then suddenly halt, taking on a neutral expression with a sustained gaze.
@@
 
-The timing of children's responses to their caregivers' speech shows a non-linear pattern of fall-rise-fall from early infancy to middle childhood.
+The timing of children's responses to their caregivers' speech shows a non-linear pattern.
 A recent study by Hilbrink et al.
@@
 
-The protracted development of children's timing may be attributable to their linguistic development: Taking turns on time is easier when the response is a simple vocalization rather than a linguistic utterance.
+This puzzling pattern is likely due to their linguistic development: Taking turns on time is easier when the response is a simple vocalization rather than a linguistic utterance.
 Integrating language into the turn-taking system may be one major factor in children's delayed responses \citep{casillas2016}.
-If response planning (i.e., language production) is the primary hurdle in children's spontaneous turn taking, we should find evidence that children understand turn-taking behaviors before they are able to produce the behaviors themselves; this hypothesis has been recently explored in experimental settings.
-
-\subsubsection{Experimental studies}
-
-Children begin to develop specific expectations about conversational behavior before they begin to speak.
-Sometime between four and six months, children begin to attend differently to face-to-face and back-to-back conversation; six-month-olds follow conversational speakers more with their gaze when at least one speaker is looking at the other \citep{augusti2010}.
-At ten months, infants expect people to look and talk at other people, and not to objects \citep{beier2012}.
-At twelve months infants expect to see responses to verbal (but not non-speech) utterances in face-to-face contexts \citep{thorgrimsson2015}.
-
-There are mixed results regarding when children begin to anticipate turn structure in conversation.
+Consistent with this hypotheiss, during the first year, children's prosodic abilities are relatively sophisticated while their lexical knowledge is limited.
+Infatns can distinguish their native language's rhythm type from others soon after birth \citep{mehler1988, nazzi2003}; they
+% Beginning between four and five months, infants prefer pauses in speech to be inserted at prosodic boundaries, and by 6 months they can start using prosodic markers to pick out sub-clausal syntactic units \citep{jusczyk1995, soderstrom2003}.
+show preference for the typical stress patterns of their native language over others by 6--9 months (e.g., iambic vs.
+trochaic), and can use prosodic information to segment the speech stream into smaller chunks from 8 months onward \citep{johnson2001, morgan1995}.
+In comparison, children show at best a very limited lexical inventory before the first birthday \citep{bergelson2013, shi2010}.
+% Two-month-olds also notice changes in word order, but this ability appears to rely on prosodic cueing \citep{mandel1996}. 
+% Generally speaking then, our current knowledge about children's linguistic development points to a possible early advantage for prosody in children's turn-taking predictions.
+
+If response planning (i.e., language production) is the primary hurdle in young children's spontaneous turn taking, we should find evidence that children understand turn-taking behaviors before they are able to produce the behaviors themselves.
+This hypothesis has been recently explored in experimental settings, but results are mixed.
 One study found that 12-month-olds make more predictive gaze shifts to a responder while watching human verbal conversation compared to conversation-like interactions with objects \citep{bakker2011}, but another only found a similar effect at 36 months \citep{hofsten2009}.
@@
 
-Keitel and colleagues \citeyearpar{keitel2013} addressed the random baseline issue in their study of 6-, 12-, 24-, and 36-month-olds.
+% Children begin to develop specific expectations about conversational behavior before they begin to speak.
+Sometime between four and six months, children begin to attend differently to face-to-face and back-to-back conversation; six-month-olds follow conversational speakers more with their gaze when at least one speaker is looking at the other \citep{augusti2010}.
+At ten months, infants expect people to look and talk at other people, and not to objects \citep{beier2012}.
+At twelve months infants expect to see responses to verbal (but not non-speech) utterances in face-to-face contexts \citep{thorgrimsson2015}.
+
+Keitel and colleagues \citeyearpar{keitel2013} addressed the random baseline issue in a study of 6-, 12-, 24-, and 36-month-olds.
 They asked participants to watch short videos of conversation and tracked their eye movements at points of speaker change.
 They found that children's anticipatory gaze frequency was only greater than chance for 36-month-olds and adults.
-Their study was the first to focus on the role of linguistic processing in children's turn predictions.
+Their study was also the first to focus on the role of linguistic processing in children's turn predictions.
 They showed their participants two types of conversation videos: One normal and one with flattened pitch (i.e., with flattened intonation contours), finding that only 36-month-olds were affected by a lack of intonation contours.
@@
 First, their estimates of baseline gaze frequency (``random'' in their terminology) were not random.
-Instead, they used gaze switches during ongoing speech as a baseline, during which switching is least likely to occur \citep{hirvenkari2013} and thereby maximizing their chances of finding a difference between gaze frequency at turn transitions and their baseline rate.
+Instead, they used gaze switches during ongoing speech as a baseline.
+Ongoing speech is perhaps the period in which switching is least likely to occur \citep{hirvenkari2013}, thus maximizing chances of finding a difference between gaze frequency at turn transitions and their baseline rate.
 A more conservative baseline would be to compare participants' looking behavior at turn transitions to their looking behavior during randomly selected windows of time throughout the stimulus, including turn transitions.
@@
 
-Second, the conversation stimuli they used were somewhat unusual.
+Second, the conversation stimuli \citet{keitel2013} used were somewhat unusual.
 The average gap between turns was 900 msec, which is much longer than typical adult timing, where gaps average around 200 msec \citep{stivers2009}.
@@
 Additionally, in order to produce more naturalistic conversation, it would have been ideal to localize the sound sources for the two voices in the video (i.e., to have the voices come out of separate left and right speakers).
-But both voices were recorded and played back on the same audio channel, which may have made it more difficult to distinguish the two talkers.
-Again, we attempt to address these issues in our current study.
-
+But both voices were recorded and played back on the same audio channel, which may have made it more difficult to distinguish the two talkers (again, we attempt to address these issues in our current study).
 Despite these minor methodological issues, the Keitel et al.
-\citeyearpar{keitel2013} study still demonstrates intriguing age-based differences in children's ability to predict upcoming turn structure, and the results suggest that both semantic and intonational development \textit{do} play a role in children's looking patterns.
+\citeyearpar{keitel2013} study still demonstrates intriguing age-based differences in children's ability to predict upcoming turn structure.
 Our current work thus takes this paradigm as a starting point.\footnote{See also \citet{casillas2012, casillas2013}.}
 
-\subsubsection{Prosodic development}
-The roles of prosody and lexicosyntax in children's turn predictions are currently unknown, but children understand more about prosody than lexicosyntax early in life.
-Children begin to acquire prosody in the womb, and can distinguish their native language's rhythm type from others (e.g., syllable-timed vs.
-stress-timed) 2--5 days after birth \citep{mehler1988, moon1993, nazzi2003}.
-Beginning between four and five months, infants prefer pauses in speech to be inserted at prosodic boundaries, and by 6 months they can start using prosodic markers to pick out sub-clausal syntactic units \citep{jusczyk1995, soderstrom2003}.
-They show preference for the typical stress patterns of their native language over others by 6--9 months (e.g., iambic vs.
-trochaic), and can use prosodic information to segment the speech stream into smaller chunks from 8 months onward \citep{johnson2001, jusczyk1993, morgan1995}.
-In comparison, children show only a limited lexical inventory at six months, and begin to recognize function words just before their first birthdays, with syntactic categorization beginning around 14 months \citep{bergelson2013, shi2010}.
-Two-month-olds also notice changes in word order, but this ability appears to rely on prosodic cueing \citep{mandel1996}.
-Generally speaking then, our current knowledge about children's linguistic development points to a possible early advantage for prosody in children's turn-taking predictions.
-
-We report here on the role of linguistic processing in children's predictions about upcoming turn structure, in particular on how children use prosodic and lexicosyntactic information to make their predictions.
-Prior work has focused mainly on lexicosyntax and intonation, and not on prosody proper (\citealp{de-ruiter2006}; \citealp{keitel2013}, but see \citealp{torreira2015}), even though infants seem to acquire the basic rhythmic properties of the prosodic signal first \citep{mehler1988, moon1993, nazzi2003}.
+Our goal in the current study is to understand the extent to which FILL ME IN BY DESCRIBING WHAT IS UNKNOWN BEFORE OUR STUDY, AND SUMMARIZE WHAT OUR EXPERIMENTS DO FOR THAT GOAL.
+
+% We report here on the role of linguistic processing in children's predictions about upcoming turn structure, in particular on how children use prosodic and lexicosyntactic information to make their predictions.
+
+% Prior work has focused mainly on lexicosyntax and intonation, and not on prosody proper (\citealp{de-ruiter2006}; \citealp{keitel2013}, but see \citealp{torreira2015}), even though infants seem to acquire the basic rhythmic properties of the prosodic signal first \citep{mehler1988, moon1993, nazzi2003}.
 

### 2015-12-04 paper/CasillasFrank-Anticip.tex
updated text for submission; some small changes to go

@@
 \usepackage{longtable}
+\usepackage[section]{placeins}
 
@@
 Of course, infant turn taking is different from adult turn taking in several ways.
-Infant turn taking is heavily scaffolded by caregivers, has distinct timing in comparison to adult turn taking, and lacks semantic content \citep{hilbrink2015, jaffe2001}.
-However, children's early, turn-structured social interactions are presumably a critical precursor to their conversational turn taking: early non-verbal interactions likely establish the protocol by which children come to use language with others.
+Infant turn taking is heavily scaffolded by caregivers, has different timing from adult turn taking, and lacks semantic content \citep{hilbrink2015, jaffe2001}.
+However, children's early, turn-structured social interactions are presumably a critical precursor to their later conversational turn taking: Early non-verbal interactions likely establish the protocol by which children come to use language with others.
 How do children integrate linguistic knowledge with these preverbal turn-taking abilities?
-
-% And as children acquire language, they also come to integrate it into their preverbal turn-taking systems.
 
@@
 In the examples above (traffic, games, and monkeys) the set of sequence and action types is far more limited and predictable than what we find in everyday talk.
-For example, conversational turns come grouped into semantically-contingent sequences of action.
+Conversational turns come grouped into semantically-contingent sequences of action.
 The groups can span turn-by-turn exchanges (e.g., simple question--response, ``How are you?''--``Fine.'') or sequence-by-sequence exchanges (e.g., reciprocals, ``How are you?''--``Fine, and you?''--``Great!'').
@@
 Prediction thus plays a key role in timely turn taking.
-
-% \subsection{}
 
@@
 Children's first turn-like structures appear as early as two to three months in proto-conversation with their caregivers \citep{bruner1975, bruner1985}.
-During proto-conversations, caregivers interact with their infants as if they were capable of making meaningful contributions: they take every look, vocalization, arm flail, and burp as ``utterances'' in the joint discourse \citep{bateson1975, jaffe2001, snow1977}.
+During proto-conversations, caregivers interact with their infants as if they were capable of making meaningful contributions: They take every look, vocalization, arm flail, and burp as ``utterances'' in the joint discourse \citep{bateson1975, jaffe2001, snow1977}.
 Infants catch onto the structure of proto-conversations quickly.
@@
 The timing of children's responses to their caregivers' speech shows a non-linear pattern.
-A recent study by Hilbrink et al.
-(2015) finds that infants' turn timing at three months is often too early or too late: They start vocalizing in overlap on $~$40\% of their caregivers' turns, and their non-overlapped vocalizations come after an average inter-turn silent gap of 350--900 msec (adult average: 200 msec).
-Between four and nine months, children begin to reduce the number of turns happening in overlap while also improving on their average response latency.
-But then, later on, children's response latencies slow down again, peaking at average gaps of more than 1000 msec at nine months, with only very gradual improvement after that \citep{hilbrink2015}.
-While children's avoidance of overlap is nearly adult-like by nine months, the timing of their non-overlapped responses stays much longer than the 200 msec standard for the next few years \citep{casillas2016, garvey1984, ervin-tripp1979}.
-
+Infants' contingent vocalizations in the first few months of life show very fast timing (though with a lot of vocal overlap) that, by nine months, slows down considerably, only gradually speeding up again after 12 months \citep{hilbrink2015}.
+Taking turns with brief transitions between speakers is difficult for children; while their avoidance of overlap is nearly adult-like by nine months, the timing of their non-overlapped responses stays much longer than the 200 msec standard for the next few years \citep{casillas2016, garvey1984, ervin-tripp1979}.
 This puzzling pattern is likely due to their linguistic development: Taking turns on time is easier when the response is a simple vocalization rather than a linguistic utterance.
 Integrating language into the turn-taking system may be one major factor in children's delayed responses \citep{casillas2016}.
-Consistent with this hypotheiss, during the first year, children's prosodic abilities are relatively sophisticated while their lexical knowledge is limited.
-Infatns can distinguish their native language's rhythm type from others soon after birth \citep{mehler1988, nazzi2003}; they
-% Beginning between four and five months, infants prefer pauses in speech to be inserted at prosodic boundaries, and by 6 months they can start using prosodic markers to pick out sub-clausal syntactic units \citep{jusczyk1995, soderstrom2003}.
-show preference for the typical stress patterns of their native language over others by 6--9 months (e.g., iambic vs.
+
+While children, like adults, could use linguistic cues in the ongoing turn to make predictions about upcoming turn structure, studies of early linguistic development point to a possible early advantage for prosody over lexicosyntax in children's turn-taking predictions.
+Infants can distinguish their native language's rhythm type from others soon after birth \citep{mehler1988, nazzi2003}; they show preference for the typical stress patterns of their native language over others by 6--9 months (e.g., iambic vs.
 trochaic), and can use prosodic information to segment the speech stream into smaller chunks from 8 months onward \citep{johnson2001, morgan1995}.
-In comparison, children show at best a very limited lexical inventory before the first birthday \citep{bergelson2013, shi2010}.
-% Two-month-olds also notice changes in word order, but this ability appears to rely on prosodic cueing \citep{mandel1996}. 
-% Generally speaking then, our current knowledge about children's linguistic development points to a possible early advantage for prosody in children's turn-taking predictions.
-
-If response planning (i.e., language production) is the primary hurdle in young children's spontaneous turn taking, we should find evidence that children understand turn-taking behaviors before they are able to produce the behaviors themselves.
+Four- to five-month-olds also prefer pauses in speech to be inserted at prosodic boundaries, and by 6 months they can start using prosodic markers to pick out sub-clausal syntactic units, both of which are useful for extracting turn structure from ongoing speech \citep{jusczyk1995, soderstrom2003}.
+In comparison, children show at best a very limited lexical inventory before their first birthday \citep{bergelson2013, shi2010}.
+
+%If response planning (i.e., language production) is the primary hurdle in young children's spontaneous turn taking, we should find evidence that children understand turn-taking behaviors before they are able to produce the behaviors themselves.
 This hypothesis has been recently explored in experimental settings, but results are mixed.
@@
 
-Keitel and colleagues \citeyearpar{keitel2013} addressed the random baseline issue in a study of 6-, 12-, 24-, and 36-month-olds.
-They asked participants to watch short videos of conversation and tracked their eye movements at points of speaker change.
-They found that children's anticipatory gaze frequency was only greater than chance for 36-month-olds and adults.
-Their study was also the first to focus on the role of linguistic processing in children's turn predictions.
-They showed their participants two types of conversation videos: One normal and one with flattened pitch (i.e., with flattened intonation contours), finding that only 36-month-olds were affected by a lack of intonation contours.
-The adult control group made equal numbers of anticipatory looks in the videos, with and without intonation contours, consistent with prior adult findings \citep{de-ruiter2006}.
-Keitel and colleagues concluded that children's ability to predict upcoming turn structure relies on their ability to comprehend the stimuli (emerging around 36 months), especially with respect to semantic access.
-They also suggest that intonation takes a secondary role in turn prediction, but only \textit{after} children acquire more sophisticated, adult-like language comprehension systems (sometime after 36 months).
+% Prior work has focused mainly on lexicosyntax and intonation, and not on prosody proper (\citealp{de-ruiter2006}; \citealp{keitel2013}, but see \citealp{torreira2015}), even though infants seem to acquire the basic rhythmic properties of the prosodic signal first \citep{mehler1988, moon1993, nazzi2003}.
+
+Keitel and colleagues \citeyearpar{keitel2013} were one of the first to explore how children use linguistic cues to predict upcoming turn structure.
+They asked 6-, 12-, 24-, 36-month-old, and adult participants to watch short videos of conversation and tracked their eye movements at points of speaker change.
+They showed their participants two types of conversation videos---one normal and one with flattened pitch (i.e., with flattened intonation contours)---to test the role of intonation in participants' anticipatory predictions about upcoming speech.
+Comparing children's anticipatory gaze frequency to a random baseline, they found that only 36-month-olds and adults made anticipatory gaze switches more often than expected by chance.
+Among those, only 36-month-olds were affected by a lack of intonation contours, leading Keitel and colleagues to conclude that children's ability to predict upcoming turn structure relies on their ability to comprehend the stimuli lexicosemantically.
+They also suggest that intonation takes a secondary role in turn prediction, following their and other studies findings with adults (e.g., \citealp{de-ruiter2006}), but only \textit{after} children acquire more sophisticated, adult-like language comprehension systems (sometime after 36 months in their data).
 
 Although the Keitel et al.
-\citeyearpar{keitel2013} study constitutes a substantial advance over previous work, it has its own limitations.
+\citeyearpar{keitel2013} study constitutes a substantial advance over previous work in this domain, it has its own limitations.
 Because these limitations directly inform our own study design, we review them in some detail.
@@
 
-Our goal in the current study is to understand the extent to which FILL ME IN BY DESCRIBING WHAT IS UNKNOWN BEFORE OUR STUDY, AND SUMMARIZE WHAT OUR EXPERIMENTS DO FOR THAT GOAL.
-
-% We report here on the role of linguistic processing in children's predictions about upcoming turn structure, in particular on how children use prosodic and lexicosyntactic information to make their predictions.
-
-% Prior work has focused mainly on lexicosyntax and intonation, and not on prosody proper (\citealp{de-ruiter2006}; \citealp{keitel2013}, but see \citealp{torreira2015}), even though infants seem to acquire the basic rhythmic properties of the prosodic signal first \citep{mehler1988, moon1993, nazzi2003}.
+Our goal in the current study is to find out when children begin to make predictions about upcoming turn structure and to understand how their predictions are affected by linguistic cues across development.
+We present two experiments in which we measure children's anticipatory gaze to responders while watching conversation videos with natural (people using English vs.
+non-English; Experiment 1) and non-natural (puppets with phonetically manipulated speech; Experiment 2) control over the presence of lexical and prosodic cues.
+We tested children across a wide range of ages (Experiment 1: 3--5 years; Experiment 2: 1--6 years), with adult control participants in each experiment.
+
+Our findings suggest that children and adults use linguistic cues to make predictions about upcoming turn structure, but that they primarily do so to predict speaker transitions after a question.
+We also find that children make more predictions than  expected by chance, even at age XX, but that it takes several years before they recruit linguistic cues to anticipate responses more often after questions.
+We found no direct evidence of an early prosody advantage in children's anticipations and, further, no evidence that prosodic or lexical cues alone can substitute for their combination in the full linguistic signal, as is proposed for adults \citep{de-ruiter2006}.
 
@@
 non-question).
-The non-English stimuli were coded for speech act from a monolingual English-speaker's perspective, i.e., which turns ``sound like'' questions, and which don't: we asked five native American English speakers to listen to the audio signal for each turn and judge whether it sounded like a question.
+The non-English stimuli were coded for speech act from a monolingual English-speaker's perspective, i.e., which turns ``sound like'' questions, and which don't: We asked five native American English speakers to listen to the audio signal for each turn and judge whether it sounded like a question.
 We then coded turns with at least 80\% ``yes'' responses as questions.
@@
 Overall, participants' pattern of gaze to current speakers demonstrated that they performed basic turn tracking during the videos, regardless of language.
+Figure \ref{fig:E1-randvsreal} shows participants' anticipatory gaze rates across age, language condition, and transition type, in the real data and the random baseline (described below).
 
@@
 \end{table}
+
+\begin{figure}[h]
+\begin{center}
+\includegraphics[width=0.8\textwidth]{figures/E1-samples-by-lang-groups-trans-types.png}
+\end{center}
+\caption{Anticipatory gaze rates across language condition and transition type for the real (red and blue) and randomly permuted baseline (gray). Vertical bars represent the standard error.} 
+\label{fig:E1-randvsreal}
+\end{figure}
 
@@
 We would need multiple items from each of the languages to check for similarity effects of specific linguistic features.}, transition type (question vs.
-non-question), age (3, 4, 5), and duration of the inter-turn gap (seconds, e.g., 0.441) as predictors, with full interactions between condition, transition type, and age.
-We included the duration of the inter-turn gap as a predictor since longer gaps also provide more opportunities to make anticipatory switches (Figure \ref{fig:criterion}).
+non-question), age (3, 4, 5; numeric), and duration of the inter-turn gap (seconds, e.g., 0.441) as predictors, with full interactions between condition, transition type, and age.
+We included the duration of the inter-turn gap as a predictor since longer gaps provide more opportunities to make anticipatory switches (Figure \ref{fig:criterion}).
 We additionally included random effects of item (turn transition) and participant, with random slopes of condition, transition type, and their interaction for participants \citep{barr2013}.\footnote{The models we report are all qualitatively unchanged by the exclusion of their random slopes.
@@
 We estimated the probability that these patterns were the result of random looking by running the same regression models on participants' real eye-tracking data, only this time calculating their anticipatory gaze switches with respect to randomly permuted turn transition windows.
-This process involved: (1) randomizing the order and temporal placement of the analysis windows within each stimulus (Figure \ref{fig:shuffling}; ``analysis window'' is defined in Figure \ref{fig:criterion}), thereby randomly redistributing the analysis windows across the eye-tracking signal, (2) re-running each participant's eye tracking data through switch identification (described in \ref{sec:algorithm}), this time using the randomly permuted analysis windows, and (3) modeling the anticipatory gazes from the randomly permuted data with the same statistical models we used for the original data (Section \ref{sec:models1}; Table \ref{tab:E1-models}).
+This process involved: (1) Randomizing the order and temporal placement of the analysis windows within each stimulus (Figure \ref{fig:shuffling}; ``analysis window'' is defined in Figure \ref{fig:criterion}), thereby randomly redistributing the analysis windows across the eye-tracking signal, (2) re-running each participant's eye tracking data through switch identification (described in Section \ref{sec:algorithm}), this time using the randomly permuted analysis windows, and (3) modeling the anticipatory gazes from the randomly permuted data with the same statistical models we used for the original data (Section \ref{sec:models1}; Table \ref{tab:E1-models}).
 Importantly, although the onset time of each transition was shuffled within the eye-tracking signal, the other intrinsic properties of each turn transition (e.g., prior speaker identity, transition type, gap duration, language condition, etc.) stayed constant across each random permutation.
@@
 \end{center}
-\caption{Example of shuffling for five turn transition analysis windows. The windows were $\pm$300 msec around the inter-turn gap.} 
+\caption{Example of analysis window permutations for a stimulus with five turn transitions. The windows were $\pm$300 msec around the inter-turn gap.} 
 \label{fig:shuffling}
@@
 We completed this baseline procedure on 5,000 random permutations of the original turn transition analysis windows and compared the \textit{t}-values from each predictor in the original models (Table \ref{tab:E1-models}) to the distribution of \textit{t}-values for each predictor in the 5,000 models of the randomly permuted datasets.\footnote{We report \textit{t}-values rather than beta estimates because the standard errors in the randomly permuted data models were much higher than for the original data.
-For those interested, plots of the beta and standard error distributions are available in the Appendix.} We could then test whether significant effects from the original statistical models differed from the random baseline by calculating the proportion of random data \textit{t}-values exceeded by the original \textit{t}-value for each predictor, using the absolute value of all \textit{t}-values for a two-tailed test.
-For example, children's original ``language condition'' \textit{t}-value was $|$3.429$|$, which is greater than 99.9\% of all $|$\textit{t}-value$|$ estimates from the randomly-permuted data models (i.e., \textit{p} $=$ .001).
+For those interested, plots of the beta and standard error distributions are available in the Supplementary Materials.} We could then test whether significant effects from the original statistical models differed from the random baseline by calculating the proportion of random data \textit{t}-values exceeded by the original \textit{t}-value for each predictor, using the absolute value of all \textit{t}-values for a two-tailed test.
+For example, children's original ``language condition'' \textit{t}-value was $|$3.429$|$, which is greater than 99.9\% of all $|$\textit{t}-value$|$ estimates from the randomly-permuted data models (i.e., \textit{p}$=$.001).
 This leads us to conclude that that the effect of language condition in the original model was highly unlikely to be the result of random gaze shifting.
 
-\begin{figure}[h!]
-  \centering
-  \subfloat[Children]{\includegraphics[width=0.49\textwidth]{figures/E1-chi-randrun-t-vals-absolute.png}\label{fig:E1-ChiTs}}
-  \hfill
-  \subfloat[Adults]{\includegraphics[width=0.49\textwidth]{figures/E1-adu-randrun-t-vals-absolute.png}\label{fig:E1-AduTs}}
-  \caption{Random-permutation and original $|$\textit{t}-values$|$ for predictors of children and adults' anticipatory gaze rates.
-Gray dots $=$ random model estimates, White dots $=$ original model estimates, Triangles $=$ 95th percentile for each \textit{t}-value distribution.}
-\end{figure}
-
-Our baseline analyses revealed that none of the significant predictors from models of the original, turn-related data can be explained by random looking.
-The children's data showed strong evidence of differentiation from the randomly permuted data for all four significant effects in the original model (Table \ref{tab:E1-models}: Children):  the original \textit{t}-values for language condition, gap duration, the age-language condition interaction, and the language condition-transition type interaction were all greater than 95\% of \textit{t}-values for the randomly permuted data (99.9\%, 95.5\%, 99.4\%, and 96\%, respectively; Figure \ref{fig:E1-ChiTs}).
-Similarly, the adults' data showed significant differentiation from the randomly permuted data for two of the three originally significant predictors---transition type and the transition type-gap duration interaction (greater than 99.9\% and 99.7\% of random \textit{t}-values, respectively)---with marginal differentiation for the interaction of language condition and transition type (greater than 94.7\% of random \textit{t}-values; Figure \ref{fig:E1-AduTs}).
-The effects of language condition and transition type for the real and randomly permuted data can also be observed in Figure \ref{fig:E1-randvsreal}.
-We excluded the output of random-permutation models that did not ultimately converge to remove unreliable model results from our percentile calculations below (78\% and 76\% of models for children and adults, respectively).
-
-\begin{figure}[h]
-\begin{center}
-\includegraphics[width=0.8\textwidth]{figures/E1-samples-by-lang-groups-trans-types.png}
-\end{center}
-\caption{Anticipatory gaze rates across language condition and transition type for the real (red and blue) and randomly permuted (gray) data. Vertical bars represent the standard error.} 
-\label{fig:E1-randvsreal}
-\end{figure}
-
-\begin{figure}[h!]
-\begin{center}
-\includegraphics[width=0.8\textwidth]{figures/E1-child-randvsreal-ttest-agebylg.png}
-\end{center}
-\caption{Pairwise comparisons of the language condition effect across ages for the original data (black dots) and the 5,000 randomly permuted datasets (distribution).} 
-\label{fig:E1-lgageinteraction}
-\end{figure}
+We excluded the output of random-permutation models that gave convergence warnings in order to remove unreliable estimates from our analyses (non-converging models were 22.4\% and 24.4\% of all models for children and adults, respectively; see the Supplementary Materials for more information on model exclusion).
+
+Our baseline analyses revealed that none of the significant predictors from models of the original, turn-related data can be explained by random looking (see Figures \ref{fig:E1-ChiTs} and \ref{fig:E1-AduTs} for the full distributional data).
+For the children's data, the original \textit{t}-values for language condition, gap duration, the age-language condition interaction, and the language condition-transition type interaction were all greater than 95\% of \textit{t}-values for the randomly permuted data (99.9\%, 95.5\%, 99.4\%, and 96\%, respectively).
+Similarly, the adults' data showed significant differentiation from the randomly permuted data for two of the three originally significant predictors---transition type and the transition type-gap duration interaction (greater than 99.9\% and 99.7\% of random \textit{t}-values, respectively)---with marginal differentiation for the interaction of language condition and transition type (greater than 94.7\% of random \textit{t}-values).
 
 \paragraph{Developmental effects}
-The model of the children's data revealed a significant interaction of age and language condition (Table \ref{tab:E1-models}) that was highly unlikely to have derived from random looking (Figure \ref{fig:E1-randvsreal}).
-To further explore this effect, we compared the average effect of language condition across all ages: we extracted the average difference score for the two language conditions (English minus non-English) for each subject, computing an overall average for each random permutation of the data.
-For each random permutation, we then made pairwise comparisons of the average difference scores across ages.
-Figure \ref{fig:E1-lgageinteraction} plots the real-data difference scores against the random-data difference score distribution for each pairwise age comparison, showing that 3- and 4-year olds were affected equally by language condition, but that 5-year-olds affected less than both 3- and 4-year-olds (with 99.52\% and 99.96\% of difference scores greater than the randomly permuted data, respectively, i.e., differences of \textit{p} $<$ .01 and \textit{p} $<$ .001).
+The original statistical model of the children's real data revealed a significant interaction of age and language condition (Table \ref{tab:E1-models}) that was highly unlikely to have come from random looking (Figure \ref{fig:E1-randvsreal}).
+To further explore this effect, we compared the average effect of language condition for each age group: We extracted the average difference score for the two language conditions (English minus non-English) for each subject, computing an overall average for each random permutation of the data.
+For each random permutation, we then made pairwise comparisons of the average difference scores across pariticipant age groups, finding that, while 3- and 4-year olds showed similarly large effects of language condition, 5-year-olds showed a significantly smaller effect of language condition, compared to both younger age groups.
+In other words, the difference in the effect of language condition for 5-year-olds compared to younger children was larger than would be expected by chance in 99.52\% of the randomly permuted data sets for 3-year-olds and 99.96\% of the data sets for 4-year-olds---differences of \textit{p}$<$.01 and \textit{p}$<$.001, respectively (see Figure \ref{fig:E1-lgageinteraction} for difference score distributions).
+
+When does spontaneous turn prediction emerge developmentally during natural speech?
+To test whether the youngest age group (3-year-olds) already exceeded chance in their anticipatory gaze switches, we used two-tailed \textit{t}-tests to compare their real gaze rates to the random baseline in the most familiar speech condition for our participants: English.
+We found that three-year-olds made anticipatory gaze switches significantly above chance, both when all transitions were considered (\textit{t}(22.824)$=$-4.147, \textit{p}$<$.001) and for question transitions alone (\textit{t}(21.677)$=$-5.268, \textit{p}$<$.001).
 
@@
 First, children made more anticipatory switches overall in English videos, compared to non-English videos.
-This effect suggests that lexical access is important for children's ability to anticipate upcoming turn structure; children had no lexical access to the speech in the non-English videos, though they did have access to (non-native) prosodic cues and non-verbal behavior, consistent with prior work on turn-end prediction in adults (\citealp{de-ruiter2006}; \citealp{magyari2012}) and children \citep{keitel2013}.
+This effect suggests that lexical access is important for children's ability to anticipate upcoming turn structure; children had no lexical access to the speech in the non-English videos, though they did have access to (non-native) prosodic cues and non-verbal behavior.
+This finding is consistent with prior work on turn-end prediction in adults (\citealp{de-ruiter2006}; \citealp{magyari2012}) and children \citep{keitel2013}.
 Second, children systematically made more anticipatory switches after hearing a question compared to a non-question, but only in the English condition, suggesting that, when children have access to lexical cues, they are more likely to make an anticipatory gaze switch if they can expect an immediate response from the addressee.
-If so, then children's attention to lexical cues for turn taking may primarily be in monitoring the signal for cues to questionhood (e.g., subject-auxiliary inversion, \textit{wh}-words, etc.).
+If so, then children's (and adults') attention to lexical cues for turn taking may primarily be in monitoring the signal for cues to questionhood (e.g., subject-auxiliary inversion, \textit{wh}-words, etc.).
 
@@
 Finally, children showed an effect of gap duration (Table \ref{tab:E1-models}).
-This effect is straightforward: longer gaps resulted in longer analysis windows, yielding more time for children to make an anticipatory gaze.
+This effect is straightforward: Longer gaps resulted in longer analysis windows, yielding more time for children to make an anticipatory gaze.
 
@@
 Like children, adults made more anticipatory switches after hearing questions compared to non-questions, suggesting that anticipation mattered more to them when an immediate response was expected.
-Also like children, the advantage for questions was driven by lexical access such that adults must have relied on lexicosyntactic cues to questionhood in picking out turns that potentially require an immediate response, though this effect was only marginally divergent from the distribution of randomly permuted data (\textit{p} $=$ .053; Figure \ref{fig:E1-AduTs}).
-Finally, adults' anticipation rates were also affected by gap duration, but more so for questions than non-questions (Table \ref{tab:E1-models}).
-This interaction suggests that adults were less likely overall to make anticipatory switches at non-questions (as is evident for adults and children in Figure \ref{fig:E1-randvsreal}), and so did not benefit from extra time to do so compared to long gaps for questions.
+Also like children, the advantage for questions was driven by lexical access such that adults must have relied on lexicosyntactic cues to questionhood in picking out turns that potentially require an immediate response, though this effect was only marginally divergent from the distribution of randomly permuted data (\textit{p}$=$.053; Figure \ref{fig:E1-AduTs}).
+Finally, adults' anticipation rates were also affected by gap duration, but more so for questions than non-questions (Table \ref{tab:E1-models}), suggesting that adults were less likely overall to make switches at non-questions, and so did not benefit from extra time to do so.
 
@@
 Even in the English videos, when participants had full access to linguistic cues, their rates of anticipation were relatively low---in fact, comparable to the non-English videos---unless the turn was a question (Figure \ref{fig:E1-randvsreal}).
-Prior work using online, metalinguistic tasks has shows that participants can use linguistic cues to accurately predict upcoming turn ends.
-The current results suggest that, in their spontaneous predictions about third-party conversation, both children and adults monitor the linguistic structure of unfolding turns for cues to upcoming responses.
-
-Children and adults generally behaved relatively similarly in this first experiment and our language manipulation (English vs.
-non-English) was too coarse to comment on when children begin to use different types of native linguistic cues (e.g., prosody vs.
-lexicosyntax); we would instead need to directly compare lexicosyntactic and prosodic cues in the participants' native language, controlling for the presence of non-verbal cues.
+Prior work using online, metalinguistic tasks has shown that participants can use linguistic cues to accurately predict upcoming turn ends \citep{torreira2015, magyari12, de-ruiter2006}.
+The current results add a new dimension to our understanding of how listeners make predictions about turn ends: Both children and adults spontaneously monitor the linguistic structure of unfolding turns for cues to upcoming responses.
+
+Children and adults behaved relatively similarly in this first experiment, and our language manipulation (English vs.
+non-English) was too coarse to comment on when children begin to use specific linguistic cues (e.g., prosody vs.
+lexicosyntax).
+We would instead need to directly compare lexicosyntactic and prosodic cues in the participants' native language, controlling for the presence of non-verbal cues.
 To see the emergence of anticipatory gaze switching we would also need to include younger children, since participants already reliably made anticipatory gaze switches at age three.
-Experiment 1 thus lays the analytic groundwork for a method that allows for greater experimental control, which we introduce in Experiment 2.
+We follow up on both of these ideas in Experiment 2.
 
@@
 \textit{Audio Manipulation}.
-We created four versions of each audio clip: \textit{Normal}, \textit{words only}, \textit{prosody only}, and \textit{no speech}.
+We created four versions of each audio clip: \textit{normal}, \textit{words only}, \textit{prosody only}, and \textit{no speech}.
 That is, one version with a full linguistic signal (\textit{normal}), and three with incomplete linguistic information (hereafter ``limited cue'' conditions).
@@
 There were 85 turn transitions for analysis after excluding transitions longer than 550 msec and shorter than 90 msec.
-The remaining turn transitions had slightly more questions than non-question (N$=$50 and N$=$35, respectively), with transitions distributed somewhat evenly across conditions (keeping in mind that there were three \textit{normal} videos and only one limited cue video for each experiment version): \textit{Normal} (N$=$36), \textit{words only} (N$=$13), \textit{prosody only} (N$=$17), and {no speech} (N$=$19).
-Inter-turn gaps for questions (mean$=$365, median$=$427) were longer than those for non-questions (mean$=$302, median$=$323) on average, but gap duration was overall comparable across conditions: \textit{Normal} (mean$=$334, median$=$321), \textit{words only} (mean$=$347, median$=$369), \textit{prosody only} (mean$=$365, median$=$369), and \textit{no words} (mean$=$319, median$=$329).
+The remaining turn transitions had slightly more questions than non-question (N$=$50 and N$=$35, respectively), with transitions distributed somewhat evenly across conditions (keeping in mind that there were three \textit{normal} videos and only one limited cue video for each experiment version): \textit{normal} (N$=$36), \textit{words only} (N$=$13), \textit{prosody only} (N$=$17), and \textit{no speech} (N$=$19).
+Inter-turn gaps for questions (mean$=$365, median$=$427) were longer than those for non-questions (mean$=$302, median$=$323) on average, but gap duration was overall comparable across conditions: \textit{normal} (mean$=$334, median$=$321), \textit{words only} (mean$=$347, median$=$369), \textit{prosody only} (mean$=$365, median$=$369), and \textit{no words} (mean$=$319, median$=$329).
 The longer gaps for question transitions could give them an advantage because our anticipatory measure includes shifts initiated during the gap between turns (Figure \ref{fig:criterion}).
@@
 Participants' pattern of gaze indicated that they performed basic turn tracking across all ages and in all conditions.
-Participants looked at the screen most of the time during video playback (82\% and 86\% average for children and adults, respectively).
-Children and adults primarily kept their eyes on the person who was currently speaking: They gazed at the current speaker between 44\% and 69\% of the time, looking back at the addressee between 11\% and 14\% of the time (Table 2).
+Participants again looked at the screen most of the time during video playback (82\% and 86\% average for children and adults, respectively), primarily looking at the person who was currently speaking (Table 2).
 They tracked the current speaker in every condition---even one-year-olds looked more at the current speaker than at anything else in the three limited cue conditions (40\% for \textit{words only}, 43\% for \textit{prosody only}, and 39\% for \textit{no speech}).
@@
 Looks to the addressee also decreased with age, but the change was minimal.
+Figure \ref{fig:E2-randvsreal} shows participants' anticipatory gaze rates across age, the four language conditions, and transition type, in the real data and the random baseline.
 
@@
 
+\begin{figure}[h]
+\begin{center}
+\includegraphics[width=0.9\textwidth]{figures/E2-samples-by-lang-groups-trans-types.png}
+\end{center}
+\caption{Anticipatory gaze rates across language condition and transition type for the real (blue, dark green, light green, and red) and randomly permuted baseline (gray).
+Vertical bars represent the standard error.}
+\label{fig:E2-randvsreal}
+\end{figure}
+
 \subsubsection{Statistical models}
 \label{sec:models2}
-
-We identified anticipatory gaze switches for all 85 usable turn transitions, and analyzed them for effects of language condition, transition type, and age with two mixed-effects logistic regressions \citep{lme4, R}.
+We identified anticipatory gaze switches for all 85 usable turn transitions, and analyzed them for effects of language condition, transition type, and age with two mixed-effects logistic regressions.
 We again built separate models for children and adults because effects of age were only pertinent to the children's data.
 The child model included condition (normal/prosody only/words only/no speech; with no speech as the reference level), transition type (question vs.
-non-question), age (1, 2, 3, 4, 5, 6), and duration of the inter-turn gap (in seconds) as predictors, with full interactions between language condition, transition type, and age.
-We again included the duration of the inter-turn gap as a control predictor and added random effects of item (turn transition) and participant, with random slopes of transition type for participants \citep{barr2013}.
+non-question), age (1, 2, 3, 4, 5, 6; numeric), and duration of the inter-turn gap (in seconds) as predictors, with full interactions between language condition, transition type, and age.
+We again included the duration of the inter-turn gap as a control predictor and added random effects of item (turn transition) and participant, with random slopes of transition type for participants.
 The adult model included condition, transition type, their interactions, and duration as a control predictor, with participant and item included as random effects and random slopes of condition and transition type.
@@
 
-Children's anticipatory gaze switches showed an effect of gap duration (\textit{$\beta$}=4.18, \textit{SE}=0.624, \textit{t}=6.689, \textit{p}$<$.001), a two-way interaction of age and language condition (for prosody only speech compared to the no speech reference level; \textit{$\beta$}=0.393, \textit{SE}=0.189, \textit{t}=2.08, \textit{p}$<$.05), and a three-way interaction of age, transition type, and language condition (for normal speech compared to the no speech reference level; \textit{$\beta$}=-0.375, \textit{SE}=0.17, \textit{t}=-2.213, \textit{p}$<$.05).
-There were no significant effects of age or transition type alone (Table \ref{sec:models2}), with only a marginal effect of language condition (for prosody only compared to the no speech reference level; \textit{$\beta$}=-1.634, \textit{SE}=0.864, \textit{t}=-1.89, \textit{p}$=$.06)
-
-Adults' anticipatory gaze switches showed effects of gap duration (\textit{$\beta$}=4.75, \textit{SE}=1.248, \textit{t}=3.806, \textit{p}$<$.001) and language condition (for normal speech \textit{$\beta$}=1.256, \textit{SE}=0.563, \textit{t}=2.229, \textit{p}$<$05.
-and words only speech \textit{$\beta$}=1.594, \textit{SE}=0.721, \textit{t}=2.211, \textit{p}$<$.05  compared to the no speech reference level).
+Children's anticipatory gaze switches showed an effect of gap duration (\textit{$\beta$}=4.18, \textit{SE}=0.624, \textit{t}=6.689, \textit{p}$<$.001), a two-way interaction of age and language condition (for \textit{prosody only} speech compared to the \textit{no speech} reference level; \textit{$\beta$}=0.393, \textit{SE}=0.189, \textit{t}=2.08, \textit{p}$<$.05), and a three-way interaction of age, transition type, and language condition (for \textit{normal} speech compared to the \textit{no speech} reference level; \textit{$\beta$}=-0.375, \textit{SE}=0.17, \textit{t}=-2.213, \textit{p}$<$.05).
+There were no significant effects of age or transition type alone (Table \ref{sec:models2}), with only a marginal effect of language condition (for \textit{prosody only} compared to the \textit{no speech} reference level; \textit{$\beta$}=-1.634, \textit{SE}=0.864, \textit{t}=-1.89, \textit{p}$=$.06)
+
+Adults' anticipatory gaze switches showed effects of gap duration (\textit{$\beta$}=4.75, \textit{SE}=1.248, \textit{t}=3.806, \textit{p}$<$.001) and language condition (for \textit{normal} speech \textit{$\beta$}=1.256, \textit{SE}=0.563, \textit{t}=2.229, \textit{p}$<$.05 and \textit{words only} speech \textit{$\beta$}=1.594, \textit{SE}=0.721, \textit{t}=2.211, \textit{p}$<$.05  compared to the \textit{no speech} reference level).
 There were no effects of transition type (\textit{$\beta$}=0.429, \textit{SE}=0.609, \textit{t}=0.705, \textit{p}$=$.48).
-
-\begin{figure}[h!]
-  \centering
-  \subfloat[Children]{\includegraphics[width=0.49\textwidth]{figures/E2-chi-randrun-t-vals-absolute.png}\label{fig:E2-ChiTs}}
-  \hfill
-  \subfloat[Adults]{\includegraphics[width=0.49\textwidth]{figures/E2-adu-randrun-t-vals-absolute.png}\label{fig:E2-AduTs}}
-  \caption{Random-permutation and original $|$\textit{t}-values$|$ for predictors of children and adults' anticipatory gaze rates.
-Gray dots $=$ random model estimates, White dots $=$ original model estimates, Triangles $=$ 95th percentile for each \textit{t}-value distribution.}
-\end{figure}
 
@@
 
-Using the same technique described in experiment 1 (Section \ref{sec:randbaseline1}), we created and modeled 5,000 random permutations of participants' anticipatory gaze.
+Using the same technique described in Experiment 1 (Section \ref{sec:randbaseline1}), we created and modeled 5,000 random permutations of participants' anticipatory gaze (see Figures \ref{fig:E2-ChiTs} and \ref{fig:E2-AduTs} for the full distributional data).
 Our baseline analyses revealed that none of the significant predictors from models of the original, turn-related data (Table \ref{tab:E2-models}: Children) can be explained by random looking.
-In the children's data, the original model's \textit{t}-values for language condition (prosody only), gap duration, the two-way interaction of age and language condition (prosody only) and the three-way interaction of age, transition type, and language condition (normal speech) were all greater than 95\% of the randomly permuted \textit{t}-values (96.2\%, 100\%, 95.1\%, and 95.1\%, respectively; Figure \ref{fig:E2-ChiTs}).
-Similarly, the adults' data showed significant differentiation from the randomly permuted data for all originally significant predictors: gap duration and language condition for normal speech and words-only speech (greater than 100\%, 96.8\%, and 98.7\% of random \textit{t}-values, respectively; Figure \ref{fig:E2-AduTs}).
+In the children's data, the original model's \textit{t}-values for language condition (\textit{prosody only}), gap duration, the two-way interaction of age and language condition (\textit{prosody only}) and the three-way interaction of age, transition type, and language condition (\textit{normal} speech) were all greater than 95\% of the randomly permuted \textit{t}-values (96.2\%, 100\%, 95.1\%, and 95.1\%, respectively).
+Similarly, the adults' data showed significant differentiation from the randomly permuted data for all originally significant predictors: gap duration and language condition for \textit{normal} speech and words-only speech (greater than 100\%, 96.8\%, and 98.7\% of random \textit{t}-values, respectively).
 The effects of language condition and transition type for the real and randomly permuted data can also be observed in Figure \ref{fig:E1-randvsreal}.
-We excluded the output of random-permutation models that did not ultimately converge to remove unreliable model results from our percentile calculations below (69\% and 70\% of models for children and adults, respectively).
-
-\begin{figure}[h]
-\begin{center}
-\includegraphics[width=0.9\textwidth]{figures/E2-samples-by-lang-groups-trans-types.png}
-\end{center}
-\caption{Anticipatory gaze rates across language condition and transition type for the real (blue, dark green, light green, and red) and randomly permuted (gray) data.
-Vertical bars represent the standard error.}
-\label{fig:E2-randvsreal}
-\end{figure}
-
-\begin{figure}[h!]
-\begin{center}
-\includegraphics[width=0.9\textwidth]{figures/E2-child-randvsreal-ttest-muffledages.png}
-\end{center}
-\caption{Significant pairwise comparisons of the prosody only-no speech linguisitic condition effect, across ages, for the original data (black dots) and the 5,000 randomly permuted datasets (distribution).}
-\label{fig:E2-lgageinteraction}
-\end{figure}
-
-\begin{figure}[h!]
-\begin{center}
-\includegraphics[width=0.9\textwidth]{figures/E2-child-randvsreal-ttest-normaltypesages.png}
-\end{center}
-\caption{Significant pairwise comparisons of the normal speech-no speech language condition effect for questions status, across ages, for the original data (black dots) and the 5,000 randomly permuted datasets (distribution).}
-\label{fig:E2-lgagetypeinteraction}
-\end{figure}
+
+As before, we excluded the output of random-permutation models that resulted in convergence warnings in order to remove unreliable model estimates from our analyses (non-converging models made up 69\% and 70\% of models for children and adults, respectively; see the Supplementary Materials for more information on model exclusion).
 
@@
 The model of the children's data revealed two significant interactions with age, neither of which derived from random looking (Table \ref{tab:E2-models}; Figure \ref{fig:E2-randvsreal}).
-The first was a significant interaction of age and language condition (for prosody only compared to the no speech reference level), suggesting a different age effect between the two linguistic conditions.
-As in Experiment 1, we further explored the age effect by extracting the average difference score over subjects for this age-language condition interaction in each random permutation of the data, making pairwise comparisons between the six age groups.
-Figure \ref{fig:E2-lgageinteraction} shows the comparisons that demonstrate a significant difference in age for the no speech baseline vs.
-the prosody only condition, revealing that anticipatory gaze in the prosody only condition significantly improves with age, especially at ages 5 and 6 (all with difference scores greater than 95\% of the random data scores; \textit{p} $<$ .05).
-
-The second interaction with age was a three-way interaction of age, transition type, and language condition (for normal speech compared to the no speech baseline).
-We again created pairwise comparisons of the average difference scores for this age-transition type-language condition interaction in each random permutation of the data, with significant differences shown in Figure \ref{fig:E2-lgagetypeinteraction}.
-These pairwise comparisons showed that the effect of transition type in the normal speech condition became larger with age, with significant improvements by age 4 over ages 1 and 2 (99.9\% and 98.86\%, respectively), by age 5 over age 4 (97.54\%), and by age 6 over ages 1, 2, and 5 (99.5\%, 97.36\%, and 95.04\%), all significantly different from chance (\textit{p} $<$ .05).
+The first was a significant interaction of age and language condition (for \textit{prosody only} compared to the \textit{no speech} reference level), suggesting a different age effect between the two linguistic conditions.
+As in Experiment 1, we further explored each age interaction by extracting an average difference score over subjects for the effect of language condition (\textit{no speech} vs.
+\textit{prosody only}) within each random permutation of the data, making pairwise comparisons between the six age groups.
+These tests revealed that children's anticipation in the \textit{prosody only} condition significantly improved at ages five and six (difference scores greater than 95\% of the random data scores; \textit{p}$<$.05; see Figure \ref{fig:E2-lgageinteraction} for the full distributions).
+
+The second age-based interaction was a three-way interaction of age, transition type, and language condition (for \textit{normal} speech compared to the \textit{no speech} baseline).
+We again created pairwise comparisons of the average difference scores for the transition type-language condition interaction across age groups in each random permutation of the data, finding that the effect of transition type in the \textit{normal} speech condition became larger with age, with significant improvements by age 4 over ages 1 and 2 (99.9\% and 98.86\%, respectively), by age 5 over age 4 (97.54\%), and by age 6 over ages 1, 2, and 5 (99.5\%, 97.36\%, and 95.04\%), all significantly different from chance (\textit{p}$<$.05; see Figure \ref{fig:E2-lgagetypeinteraction} for the full distributions).
+
+Our main goal in extending the age range to 1- and 2-year-olds in Experiment 2 was to find the age of emergence for spontaneous turn structure predictions.
+As in Experiment 1, we used two-tailed \textit{t}-tests to compare children's real gaze rates to the random baseline in the \textit{normal} speech condition, in which the speech stimulus is most like what children hear every day.
+We tested real gaze rates against baseline for three age groups: ages one, two, and three.
+Two- and three-year-old children made anticipatory gaze switches significantly above chance both when all transitions were considered (2-year-olds: \textit{t}(26.193)$=$-4.137, \textit{p}$<$.001; 3-year-olds: \textit{t}(22.757)$=$-2.662, \textit{p}$<$.05) and for question transitions alone (2-year-olds: \textit{t}(25.345)$=$-4.269, \textit{p}$<$.001; 3-year-olds: \textit{t}(21.555)$=$-3.03, \textit{p}$<$.01).
+One-year-olds, however, made anticipatory gaze shifts that were only marginally above chance for turn transitions overall and for question turns alone (overall: \textit{t}(24.784)$=$-2.049, \textit{p}$=$.051; questions: \textit{t}(25.009)$=$-2.03, \textit{p}$=$.053).
 
@@
 \label{sec:discussion2}
-As in Experiment 1, children and adults spontaneously tracked the turn structure of the conversations, making anticipatory gaze switches at an above-chance rate across all ages but not in all conditions (Table \ref{tab:e2_look}; Figure \ref{fig:E2-randvsreal}).
-In the normal speech condition, however, when children had access to full linguistic information, they made more anticipatory gaze switches than expected by chance even at age one.
-Children's anticipatory gaze rates were affected by gap duration, plus interactions of language condition, transition type and age (Table \ref{tab:E2-models}), none of which could be explained by a baseline of random gaze switching (Figure \ref{fig:E2-ChiTs}).
-
-Two of the three linguistic conditions affected children's anticipatory gaze switches, compared to the no-speech reference level: prosody-only speech and normal speech.
-Prosody only speech resulted in fewer anticipatory gazes overall, compared to the no-speech baseline, though the effect was only marginal (\textit{p}$=$.06).
-However, prosody-only speech \textit{did} significantly differ from the no speech condition in its interaction with age: whereas age gains were negligible in the no speech condition, 5- and 6-year-olds in the prosody-only condition showed significantly more anticipatory gaze switches than younger children (Figure \ref{fig:E2-lgageinteraction}), going from below- and at-chance anticipatory gaze rates to being well above chance (Figure \ref{fig:E2-randvsreal}).
-
-The normal speech condition showed a significantly more gains in anticipatory gaze for questions with age compared to the no-speech condition.
+As in Experiment 1, children and adults spontaneously tracked the turn structure of the conversations, making anticipatory gaze switches at an above-chance rate across all ages, but not in all conditions (Table \ref{tab:look_e2}; Figure \ref{fig:E2-randvsreal}).
+However, when children had access to full linguistic information (in the \textit{normal} speech condition), they made more anticipatory gaze switches than expected by chance at age two (and even marginally at age one).
+Across conditions, children's anticipatory gaze rates were affected by gap duration, plus interactions of language condition, transition type and age (Table \ref{tab:E2-models}), none of which could be explained by a baseline of random gaze switching (Figure \ref{fig:E2-ChiTs}).
+
+Two of the three linguistic conditions affected children's anticipatory gaze switches, compared to the no-speech reference level: prosody-only speech and \textit{normal} speech.
+\textit{Prosody only} speech resulted in fewer anticipatory gazes overall, compared to the no-speech baseline, though the effect was only marginal (\textit{p}$=$.06).
+However, prosody-only speech \textit{did} significantly differ from the \textit{no speech} condition in its interaction with age: Whereas age gains were negligible in the \textit{no speech} condition, 5- and 6-year-olds in the prosody-only condition showed significantly more anticipatory gaze switches than younger children (Figure \ref{fig:E2-lgageinteraction}), going from at-chance anticipatory gaze rates to rates well above chance (Figure \ref{fig:E2-randvsreal}).
+
+Anticipatory gaze for questions increased significantly more with age in the \textit{normal} speech condition compared to the no-speech baseline.
 As in Experiment 1, children consistently made more anticipatory gazes after hearing questions when they had access to lexical material.
 But now, with a larger age span in Experiment 2, we can begin to see a developmental path for the question effect.
-At least for normal speech, greater anticipatory looking for questions is not present from the start: 4-year-olds are the first to make significant gains on 1- and 2-year-olds, but then there are further significant gains at age 5, and again age 6 (Figure \ref{fig:E2-lgagetypeinteraction}).
-While children at ages five and six showed adult-like differentiation of questions and non-questions in the normal speech condition, 1-year-olds had nearly identical switch rates for the two transition types.
+At least for \textit{normal} speech, greater anticipatory looking for questions is not present from the start: 4-year-olds are the first to make significant gains on 1- and 2-year-olds, but then there are further significant gains at age 5, and again age 6 (Figure \ref{fig:E2-lgagetypeinteraction}).
+While children at ages five and six showed adult-like differentiation of questions and non-questions in the \textit{normal} speech condition, 1-year-olds had nearly identical switch rates for the two transition types.
 This suggests that the participants' tendency to make more anticipatory switches for questions emerges after age one and continues developing, with rapid gains in ages three through age six.
@@
 Adults' anticipatory gaze rates were affected by gap duration and two of the language conditions (Table \ref{tab:E2-models}), none of which could be explained by a baseline of random gaze switching (Figure \ref{fig:E2-AduTs}).
-Adults made more anticipatory switches overall for the normal speech and words-only conditions compared to the no-speech condition, falling in-line with past work showing that adults primarily use lexical information in making predictions about upcoming turn structure \citep{de-ruiter2006}.
-Though adults did make more anticipatory switches for questions than non-questions on average in the two lexical conditions (Figure \ref{fig:E2-randvsreal}), the effect of transition type was not significant in either (Table \ref{tab:E2-models}), unlike Experiment 1.
+Adults made more anticipatory switches overall for the \textit{normal} speech and words-only conditions compared to the no-speech condition, falling in-line with past work showing that adults primarily use lexical information in making predictions about upcoming turn structure \citep{de-ruiter2006}.
+Though adults did make more anticipatory switches for questions than non-questions on average in the two lexical conditions (Figure \ref{fig:E2-randvsreal}), the effect of transition type was not significant in either one (Table \ref{tab:E2-models}).
 Like children, adults also showed a straightforward effect of gap duration.
@@
 
-In fact, children and adults did show more anticipatory gaze switching in the normal condition compared to the no-speech condition, but for children this effect only emerged in the form of anticipations following question turns, which increased with age faster than it did with no linguistic cues at all (i.e., in the no-speech condition).
-Adults also showed significantly higher anticipatory switch rates in the words only condition but no effects of transition type, alone or within linguistic conditions.
-Taken together, these results partly replicate the findings from Experiment 1: participants make more anticipatory switches when they have access to lexical information and, when they do, tend to make more anticipatory switches for questions compared to non-questions.
-
-We had anticipated significant gains in anticipatory switching with age, but children only showed significant developmental increases in the prosody only condition and the normal condition (for question transitions).
-Rather than showing an early advantage for prosody over lexical information, children did not show significant improvement (or even above-chance performance) until age five.
-In contrast, their anticipatory gaze switches were, on average, already above chance at age one in the normal condition (with both lexical and prosodic information) and at age three in the words only condition (with lexical information).
-These findings do not support an early role for prosody in children's spontaneous predictions about upcoming turn structure.
-On the contrary, their predictions were best when lexical information was present, especially when following a question.
-
-However, these results do not support the idea that lexical information is sufficient for children (as is proposed for adults; \citealp{de-ruiter2006}).
-On the contrary, children showed significant gains in the normal speech condition, but \textit{not} in the words only condition.
-Notably, the normal speech condition, in addition to having both lexical and prosodic information, is also the one most likely to occur in our participants' daily lives (compared to muffled or robotic speech) and therefore may have an extra advantage over the other conditions.
-
-Finally, on average, participants at all ages made anticipatory gaze switches more often than would be expected by chance in the \textit{no speech} condition.
-One interpretation of this finding is that participants were not using linguistic information to predict upcoming turns at all---the only cue to turn taking in the \textit{no speech} condition was the alternating mouth movements of the conversing puppets.
-But this interpretation is not compatible with the effects of linguistic information that do emerge in the adult and child data: An advantage for lexical cues shaped by transition type and age.
-
-The core aims of Experiment 2 were to gain better traction on the individual roles of prosody and lexicosyntax in children's turn predictions, and to expand our age range to capture more developmental change.
-We found that effects of linguistic processing and age \textit{were} present in the dataset, but were primarily related to the question effect, and primarily occur in the normal speech condition.
-The relation between prosody and lexicosyntax in young children's anticipations is still not clear: The results of the prosody only manipulation suggest that children do not reliably use low-pass filtered speech to anticipate upcoming turn structure until age five, but on the other hand, the words only condition did not show significant overall differences from the no-speech baseline, including its interactions with age or transition type.
-If anything, the results do not support the idea that there is a strong early advantage for prosody in turn predictions.
+In fact, children and adults did show more anticipatory gaze switching in the \textit{normal} speech condition, compared to the no-speech condition, but for children this effect only emerged in the form of anticipations following question turns, which increased with age significantly faster than it did when there were no linguistic cues present at all.
+Adults also showed significantly higher anticipatory switch rates in the \textit{words only} condition but no effects of transition type, alone or within linguistic conditions.
+Taken together, these results partly replicate the findings from Experiment 1: Participants make more anticipatory switches when they have access to lexical information and, when they do, tend to make more anticipatory switches for questions compared to non-questions.
+
+We had anticipated significant gains in anticipatory switching with age, but children only showed significant developmental increases in the \textit{prosody only} condition and the \textit{normal} condition (for question transitions).
+Rather than showing an early advantage for prosody over lexical information, children's anticipation were at chance with prosody-only speech and did not show significant improvement until age five.
+In contrast, their anticipatory gaze switches were already marginally above chance at age one and significantly so at age two for \textit{normal} speech.
+These findings therefore do not support an early role for prosody in children's spontaneous turn structure predictions.
+On the contrary, children's predictions were best when lexical information was present, especially when following a question.
+
+However, these results do not support the idea that lexical information is sufficient for children, as has been proposed for adults previously \citealp{de-ruiter2006} and replicated in the present study.
+On the contrary, children showed significant gains in the \textit{normal} speech condition, but \textit{not} in the \textit{words only} condition.
+Notably, the \textit{normal} speech condition, in addition to having both lexical and prosodic information, is also the one most likely to occur in our participants' daily lives (compared to muffled or robotic speech) and therefore may have an extra advantage of familiarity over the other conditions.
+
+Finally, though participants were often able to make anticipatory gaze switches more often than would be expected by chance with no linguistic information (in the \textit{no speech} condition), the advantage for lexical cues emerging in the children's data with age and in the adult's data overall suggests that participants were using linguistic information in making their predictions when it was available.
+
+The core aims of Experiment 2 were to gain better traction on the individual roles of prosody and lexicosyntax in children's turn predictions, and to find the age of emergence for spontaneous turn anticipation.
+Replicating many of our findings from Experiment 1, we found that children and adults both use linguistic cues, in children's case, primarily to anticipate responses after question turns and primarily when the full speech signal was available.
+Adults, on the other hand, showed broadly increased anticipation for lexical conditions, with no overall effects of transition type.
+
+Rather than finding an early advantage for prosodic cues, we found that children struggled to make above-chance anticipations in the \textit{prosody only} condition until age five.
+They didn't show significant overall effects for lexical information alone either, suggesting that children need the full linguistic signal to extract predictive benefits from ongoing conversational speech.
+This finding is surprising given children's earlier development of prosodic skills but unsurprising in the light of their overall question bias---compared to prosodic cues, lexicosyntactic cues to questionhood are categorical units, have clearer form-to-meaning mappings, and occur early in the utterance.
+This finding has implications for our conceptions about adults and children's use of linguistic cues for spontaneous turn prediction, which we expand on in the General Discussion.
 
@@
 In two experiments investigating children's anticipatory gaze to upcoming speakers, we found evidence that turn prediction develops early in childhood and that spontaneous predictions are primarily driven by participants' expectation of an immediate response in the next turn.
-In making predictions about upcoming turn structure, children used a combination lexical and prosodic cues; neither prosodic nor lexical cues alone were sufficient to support increased anticipatory gaze.
-We also found no early advantage for prosody over lexicosyntax, and instead found that children were unable to make above-chance anticipatory gazes in the prosody only condition until age five.
+In making predictions about upcoming turn structure, children used a combination of lexical and prosodic cues; neither lexical nor prosodic cues alone were sufficient to support increased anticipatory gaze.
+We also found no early advantage for prosody over lexicosyntax, and instead found that children were unable to make above-chance anticipatory gazes in the \textit{prosody only} condition until age five.
 We discuss these findings with respect to the role of linguistic cues in predictions about upcoming turn structure, the importance of questions in spontaneous predictions about conversation, and children's developing competence as conversationalists.
@@
 Prior work with adults has found a consistent and critical role for lexicosyntax in predicting upcoming turn structure \citep{de-ruiter2006, magyari2012}, with the role of prosody still under debate \citep{duncan1972, ford1996, torreira2015}.
-Knowing that children comprehend more about prosody than lexicosyntax early on (see introduction; also see \citealp{speer2009} for a review), we thought it possible that young children would instead show an advantage for prosody over lexicosyntax.
+Knowing that children comprehend more about prosody than lexicosyntax early on (see introduction; also see \citealp{speer2009} for a review), we thought it possible that young children would instead show an advantage for prosody in their predictions about turn structure in conversation.
 Our results suggest that, on the contrary, when presented with \textit{only} prosodic information, children's spontaneous predictions about upcoming turn structure are limited until age five.
-Do children instead rely on lexicosyntax, as adults are proposed to do?
-
-We found no evidence that, for children, lexicosyntax alone is ``sufficient'' (equal to full linguistic information) for spontaneous turn prediction \citep[pg.
-531]{de-ruiter2006}: In both experiments, children's performance was best in conditions when they had access to the full linguistic signal.
-Adults on the other hand, showed significant gains in anticipatory gaze switching in both conditions with lexical cues, compared to a no speech baseline.
-
-Participants appeared to use linguistic cues to make more anticipatory gaze switches in both experiments.
-Questions are often marked both prosodically and lexicosyntactically for their speech act status.
-The close link between prosodic and syntactic structure makes it difficult to tease apart how predictive processing in one domain is distinct from the other in turn prediction.
-That being said, compared to prosodic contours (e.g., final rising intonation), lexicosyntactic cues like \textit{wh}-words, \textit{do}-insertion, and subject-auxiliary inversion are frequent, categorical, and early-occurring in the utterance.
-Children may therefore have an easier time picking out and interpreting lexical cues to questionhood on the fly.
-Question turns started yielding significantly more anticipatory gazes by age 3--4 in the normal speech condition of Experiment 2, by which time children frequently hear and use a variety of polar \textit{wh}-questions \citep{clark2009}.
+
+Importantly, we also found no evidence that lexical information alone is equivalent to full linguistic information for children, as has been shown before \cite{magyari2012, de-ruiter2006} and replicated in the current study for adult participants: In both experiments, children's performance was best in conditions when they had access to the full linguistic signal.
+Adults on the other hand, showed significant gains in anticipatory gaze switching in both conditions with lexical cues.
+If this effect arose in our data because children do not make predictions based on phonetically manipulated speech in conversational contexts, the current findings can be overturned by follow-up work controlling linguistic information through other means.
+But if not, there may be something specially informative about the combined prosodic and lexical cues to questionhood that boosts children's anticipations before they can use these cues separately.
+Even in adults, Torreira and colleagues (\citeyear{torreira2015}) were able to show that the trade-off in informativity between lexical and prosodic cues is more subtle in semi-natural (spliced) speech.
+The present findings are the first to show evidence of a similar effect developmentally.
+
+\subsubsection{The question effect}
+
+In both experiments, anticipatory looking was primarily driven by question transitions, a pattern that had not been previously reported in other anticipatory gaze studies, on children or adults (Keitel et al., 2013; Hirvenkari, 2013; Tice and Henetz, \citeyear{TiceHenetz11}).
+Questions make an upcoming speaker switch immediately relevant, helping the listener to predict with high certainty what will happen next (i.e., an answer from the addressee), and are often easily identifiable by overt prosodic and lexicosyntactic cues.
+
+Compared to prosodic cues (e.g., final rising intonation), lexicosyntactic cues (e.g., \textit{wh}-words, \textit{do}-insertion, and subject-auxiliary inversion) were frequent, categorical, and early-occurring in the utterance.
+Children may have therefore had an easier time picking out and interpreting lexical cues to questionhood.
+The question effect showed its first significant gains between ages three and four in the \textit{normal} speech condition of Experiment 2, by which time children frequently hear and use a variety of polar \textit{wh}-questions \citep{clark2009}.
 Furthermore, while lexicosyntactic question cues were available on every instance of \textit{wh}- and \textit{yes/no} questions in our stimuli, prosodic question cues were only salient on \textit{yes/no} questions and, even then, the mapping of prosodic contour to speech act (e.g., high final rises for polar questions) is far from one-to-one.
 
-\subsubsection{The question effect}
-
-In both experiments, anticipatory looking was primarily driven by question transitions, a pattern that had not been previously reported in other observer gaze studies, on children or adults (Keitel et al., 2013; Hirvenkari, 2013; Tice and Henetz, \citeyear{TiceHenetz11}).
-Questions make a speaker switch immediately relevant, helping the listener to predict with high certainty what will happen next (i.e., an answer from the addressee).
-As mentioned, linguistic cues to questionhood also often occur early in the utterance, giving observers time to plan a gaze switch.
-Because the form of a question can constrain the type of response that will come next (e.g., a location after a \textit{where} question), questions can even help listeners predict specific upcoming content in the next turn.
-
-Our results suggest that question turns start significantly affecting children's predictions between ages three and four, but further testing with finer-grained age samples with stimuli focused on specific communicative acts and linguistic cues is needed to better sketch out the developmental trajectory of this effect.
+%Because the form of a question can constrain the type of response that will come next (e.g., a location after a \textit{where} question), questions can even help listeners predict specific upcoming content in the next turn.
+
+%Our results suggest that question turns start significantly affecting children's predictions between ages three and four, but further testing with finer-grained age samples with stimuli focused on specific communicative acts and linguistic cues is needed to better sketch out the developmental trajectory of this effect.
 For example, the effect size and age of emergence might differ by question type (e.g., \textit{wh-} vs.
@@
 
-All that being said, our current data do not tell us what it is about questions that makes children and adults more likely to anticipatorily switch their gaze to addressees.
+Two important questions for future work are then: (a) How does children's ability to monitor for questions in conversation relate to their prior experience with questions?
+and (b) what is it about questions that makes children and adults more likely to anticipatorily switch their gaze to addressees?
 Other request formats, such as imperatives, compliments, and complaints make a response from the addressee highly likely in the next turn \citep{schegloff2007}.
 Rhetorical and tag questions, on the other hand, take a similar form to prototypical polar questions, but often do not require an answer.
-So, though it is clear that participants anticipated responses more often for questions than non-questions, we do not yet know whether their predictive action is limited to turns formatted as questions or is generally applicable to turn structures that project an immediate response from the addressee.
-
-Much recent work on prediction during turn taking has focused on participants' use of linguistic cues to predict the end of the current turn \citep{torreira2015, magyari2012, de-ruiter2006, ford1996, duncan1972}, in some cases finding that lexical information was sufficient for prediction.
-Our current results suggest that, spontaneous predictions are instead driven by predictions about what is \textit{beyond} the end of the current turn---that questions are sufficient for prediction.
-
-To integrate these results and understand how listeners make predictions for turn taking, it is crucial to account for the participants' role in the interaction.
+So, though it is clear that adults and children anticipated responses more often for questions than non-questions, we do not yet know whether their predictive action is limited to turns formatted as questions or is generally applicable to turn structures that project an immediate response from the addressee.
+
+The question effect itself has implications for our current theories about turn prediction.
+While participants may always use linguistic information to predict upcoming speaker changes (as they have in the two studies presented here), they might not always use linguistic information to predict upcoming turn ends, as is assumed in metalinguistic measures of turn-end prediction (e.g., pressing a button while listening to speech: \citealp{torreira2015, magyari2012, de-ruiter2006}) and as has been the focus of other turn-end prediction studies \citealp{ford1996, duncan1972}.
+Our results rather suggest that participants' \textit{spontaneous} predictions, at least while viewing third-party conversation, are the results of question-monitoring, presumably achieved by recruiting linguistic cues to questionhood from the unfolding signal.
+In other words, our results suggest that predictions are driven by what is \textit{beyond} the end of the current turn---that questions, not lexical cues, are sufficient for prediction at the first level of conversation monitoring.
+
+There is at least one clear hypothesis that can bridge these apparently conflicting results: Listeners in spontaneous, first-person conversation may use multiple strategies in making predictions about upcoming turn structure, using more passive prediction to detect upcoming speaker transition (e.g., questions), and then switching into precise turn-end prediction mode (\`{a} la \citealp{de-ruiter2006}) when necessary.
+A flexible prediction system like this one allows listeners to continuously monitor ongoing conversation at a low cost while still managing to plan their responses and come in quickly when needed.
+
+To test this hypothesis, which integrates findings from turn-structure prediction with multiple measures, age groups, and styles of linguistic control, it will be crucial to look at prediction from a first-person perspective.
 The results we present here are based on predictions about third-party conversation, which enables participants to follow interactions with no chance of actually participating.
-Although recent work has shown that similar anticipatory eye gazes do occur in spontaneous conversation \citep{holler2015}, we do not yet know if the same question advantage occurs, or which linguistic cues seem to drive it.
-It is possible that participants track conversation for cues to upcoming opportunities/obligations to speak and, when one is found, focus more attention on predicting the precise timing of the current turn's end.
-To answer these questions we will need to innovate first-person prediction measures that can be used in real-time interaction.
+Although recent work has shown that similar anticipatory eye gazes do occur in spontaneous conversation \citep{holler2015}, more work is needed to determine if the same question advantage occurs, which linguistic cues seem to drive it, and whether participants switch into a ``precision'' mode when they detect imminent speaker change.
 
@@
 
-One of the core aims of our study was to test whether children show an early competence for turn taking, as is proposed by studies of spontaneous mother-infant interaction and theories about the mechanisms underlying human interaction \citep{hilbrink2015, levinson2006}.
-Although children and adults' gaze patterns were quite similar in Experiment 1, children in Experiment two showed change with age in their anticipatory looking, even in the most prototypical speech condition (``normal'' speech), and 6-year-olds still did not achieve adult-like levels of anticipatory looking on average.
-This may indicate that children rely more in non-verbal cues in anticipating turn transitions or, alternatively, that adults are better at flexibly adapting to the turn-relevant cues present at any moment.
-When they \textit{did} have full linguistic information, children on average still made above-chance anticipatory gazes to responders, suggesting that at least some turn-taking competence for third-party conversation is present in infancy, though it is a weak effect.
-
-Taken together, the data suggest that turn-taking skills do begin to emerge in infancy, but that adult-like competence is not achieved until much later, mirroring results from children's spontaneous interactions with their caregivers (infants: \citealp{hilbrink2015, jaffe2001, snow1977}; older children: \citealp{casillas2016, garvey1984, ervin-tripp1979}).
-It is possible, however, that first-person measures of children's predictions would show more frequent turn structure anticipations at younger ages.
+One of the core aims of our study was to test whether children show an early competence for turn taking, as is proposed by studies of spontaneous mother-infant proto-conversation and theories about the mechanisms underlying human interaction in general \citep{hilbrink2015, levinson2006}.
+We did find evidence that young children already make spontaneous predictions about upcoming turn structure, definitely at age two and even marginally at age one.
+However, ``above chance'' performance was far from adult-like predictive behavior, and children in our studies did not show adult-like competency in their predictions, even at age six.
+This may indicate that children rely more on non-verbal cues in anticipating turn transitions or, alternatively, that adults are better at flexibly adapting to the turn-relevant cues present at any moment.
+
+% We could add here that this flexibility might emerge around age 5: That's when we see some mini-adult-like effects in the children's data.
+In E1 that take the form of their (potentially) using either non-verbal or non-native prosodic cues in the non-English videos.
+In E2 this takes the form of 5 and 6 year olds showing significant improvement in the \textit{prosody only} condition.
+
+Taken together, the data suggest that turn-taking skills do begin to emerge in infancy, but that their predictions don't help them anticipate much until they have acquired the ability to pick out question turns.
+This finding leads us to wonder how participant role (first- instead of third-person) and cultural differences (e.g., high vs.
+low parent-infant interaction styles) might feed into this early predictive skill.
+It also bridges the prior work showing a predisposition for turn taking in infancy (e.g., \citealp{hilbrink2015}) but late acquisition of adult-like competence when it comes to integrating linguistic information into turn-taking behaviors \citep{casillas2016, garvey1984, ervin-tripp1979}.
 
@@
 
-Although Experiments 1 and 2 have offered new findings regarding the relation to speech act in online turn predictions and the emergence of those predictions in the first six years, there remain at least two major limitations to our work: speech naturalness and participant role.
-
-Following prior work \citep{de-ruiter2006, keitel2013}, we used phonetically manipulated speech in Experiment 2, resulting in speech sounds that children don't usually hear in their natural environment.
+There are at least two major limitations to our work: Speech naturalness and participant role.Following prior work \citep{de-ruiter2006, keitel2013}, we used phonetically manipulated speech in Experiment 2, resulting in speech sounds that children don't usually hear in their natural environment.
 Many prior studies have used phonetically-altered speech with infants and young children \citep[cf.][]{jusczyk2000}, but almost none of them have done so in a conversational context.
-Children could have had trouble processing the \textit{words only} and \textit{prosody only} conditions because they were unfamiliar, and not just because they had less linguistic information available.
 Future work could instead carefully script or cross-splice parts of turns to control for the presence or absence of linguistic cues for turn transition.
 
-The prediction measure we present is based on an observer's view of third-party conversation but, because participants' role in the interaction could affect their online predictions about turn taking, an ideal experimental measure would capture first-person behavior.
+The prediction measure used in these studies is based on an observer's view of third-party conversation but, because participants' role in the interaction could affect their online predictions about turn taking, an ideal experimental measure would capture first-person behavior.
 First-person measures of spontaneous turn prediction will be key to revealing how participants distribute their attention over linguistic and non-verbal cues while taking part in everyday interaction, the implications of which relate to theories of online language processing for both language learning and everyday talk.
@@
 Adults use linguistic information to accurately predict turn structure in conversation, which facilitates their online comprehension and allows them to respond relevantly and on time.
-In the current study we have investigated how children's predictions about turn structure changes as their linguistic skills develop in the first six years.
-We found that, although some basic knowledge about turn taking exists at age one, the integration of linguistic cues into children's predictions about turn structure takes time and is, like adults, primarily driven by sequences of action---in our case, by questions and their answers.
+The present study offers new findings regarding the role of speech acts and linguistic processing in online turn prediction, and has given evidence that turn prediction emerges by age two is not integrated with linguistic cues until much later.
+Using language to make predictions about upcoming interactive content takes time and, for both children and adults, is primarily driven by participants' orientation to what will happen beyond the end of the current turn.
 
@@
 Clark, and the First Language Acquisition group at Radboud University for their feedback on earlier versions of this work.
+The analysis code and raw data for this project can be found on GitHub at https://github.com/langcog/turn\_taking/.
 
@@
 \appendix
+
+\section{Real vs. randomly permuted model outcomes: \textit{t}, \textit{$\beta$}, and \textit{SE}}
 \setcounter{figure}{0}
+
 In all of the following plots, the gray dots represent the randomly permuted data's model estimates for the value listed (beta or standard error), the white dots represent the model estimates from the original data, and the triangles represent the 95th percentile for each distribution being shown.
 
-\begin{figure}
+\subsection{Experiment 1}
+
+\begin{figure}[!htb]
   \centering
-  \subfloat[Children]{\includegraphics[width=0.49\textwidth]{figures/E1-chi-randrun-betas-absolute.png}\label{fig:E1-ChiBs}}
+  \subfloat[Children]{\includegraphics[width=0.44\textwidth]{figures/E1-chi-randrun-t-vals-absolute.png}\label{fig:E1-ChiTs}}
   \hfill
-  \subfloat[Adults]{\includegraphics[width=0.49\textwidth]{figures/E1-adu-randrun-betas-absolute.png}\label{fig:E1-AduBs}}
-  \caption{Random-permutation and original $|$\textit{$\beta$}-values$|$ for predictors of children and adults' anticipatory gaze rates in Experiment 1.}
-\end{figure}
-
-\begin{figure}
+  \subfloat[Adults]{\includegraphics[width=0.44\textwidth]{figures/E1-adu-randrun-t-vals-absolute.png}\label{fig:E1-AduTs}}
+  \caption{Random-permutation and original $|$\textit{t}-values$|$ for predictors of anticipatory gaze rates in Experiment 1.}
+\end{figure}
+
+\begin{figure}[!htb]
   \centering
-  \subfloat[Children]{\includegraphics[width=0.49\textwidth]{figures/E1-chi-randrun-SEs-absolute.png}\label{fig:E1-ChiSEs}}
+  \subfloat[Children]{\includegraphics[width=0.44\textwidth]{figures/E1-chi-randrun-betas-absolute.png}\label{fig:E1-ChiBs}}
   \hfill
-  \subfloat[Adults]{\includegraphics[width=0.49\textwidth]{figures/E1-adu-randrun-SEs-absolute.png}\label{fig:E1-AduSEs}}
-  \caption{Random-permutation and original $|$\textit{SE}-values$|$ for predictors of children and adults' anticipatory gaze rates in Experiment 1.}
-\end{figure}
-
-\begin{figure}
+  \subfloat[Adults]{\includegraphics[width=0.44\textwidth]{figures/E1-adu-randrun-betas-absolute.png}\label{fig:E1-AduBs}}
+  \caption{Random-permutation and original $|$\textit{$\beta$}-values$|$ for predictors of gaze rates in Experiment 1.}
+\end{figure}
+
+\begin{figure}[!htb]
   \centering
-  \subfloat[Children]{\includegraphics[width=0.49\textwidth]{figures/E2-chi-randrun-betas-absolute.png}\label{fig:E2-ChiBs}}
+  \subfloat[Children]{\includegraphics[width=0.44\textwidth]{figures/E1-chi-randrun-SEs-absolute.png}\label{fig:E1-ChiSEs}}
   \hfill
-  \subfloat[Adults]{\includegraphics[width=0.49\textwidth]{figures/E2-adu-randrun-betas-absolute.png}\label{fig:E2-AduBs}}
-  \caption{Random-permutation and original $|$\textit{$\beta$}-values$|$ for predictors of children and adults' anticipatory gaze rates in Experiment 2.}
-\end{figure}
-
-\begin{figure}
+  \subfloat[Adults]{\includegraphics[width=0.44\textwidth]{figures/E1-adu-randrun-SEs-absolute.png}\label{fig:E1-AduSEs}}
+  \caption{Random-permutation and original $|$\textit{SE}-values$|$ for predictors of anticipatory gaze rates in Experiment 1.}
+\end{figure}
+
+\clearpage
+
+\subsection{Experiment 2}
+
+\begin{figure}[!htb]
   \centering
-  \subfloat[Children]{\includegraphics[width=0.49\textwidth]{figures/E2-chi-randrun-SEs-absolute.png}\label{fig:E2-ChiSEs}}
+  \subfloat[Children]{\includegraphics[width=0.44\textwidth]{figures/E2-chi-randrun-t-vals-absolute.png}\label{fig:E2-ChiTs}}
   \hfill
-  \subfloat[Adults]{\includegraphics[width=0.49\textwidth]{figures/E2-adu-randrun-SEs-absolute.png}\label{fig:E2-AduSEs}}
-  \caption{Random-permutation and original $|$\textit{SE}-values$|$ for predictors of children and adults' anticipatory gaze rates in Experiment 2..}
-\end{figure}
+  \subfloat[Adults]{\includegraphics[width=0.44\textwidth]{figures/E2-adu-randrun-t-vals-absolute.png}\label{fig:E2-AduTs}}
+  \caption{Random-permutation and original $|$\textit{t}-values$|$ for predictors of anticipatory gaze rates in Experiment 2.}
+\end{figure}
+
+\begin{figure}[!htb]
+  \centering
+  \subfloat[Children]{\includegraphics[width=0.44\textwidth]{figures/E2-chi-randrun-betas-absolute.png}\label{fig:E2-ChiBs}}
+  \hfill
+  \subfloat[Adults]{\includegraphics[width=0.44\textwidth]{figures/E2-adu-randrun-betas-absolute.png}\label{fig:E2-AduBs}}
+  \caption{Random-permutation and original $|$\textit{$\beta$}-values$|$ for predictors of anticipatory gaze rates in Experiment 2.}
+\end{figure}
+
+\begin{figure}[!htb]
+  \centering
+  \subfloat[Children]{\includegraphics[width=0.44\textwidth]{figures/E2-chi-randrun-SEs-absolute.png}\label{fig:E2-ChiSEs}}
+  \hfill
+  \subfloat[Adults]{\includegraphics[width=0.44\textwidth]{figures/E2-adu-randrun-SEs-absolute.png}\label{fig:E2-AduSEs}}
+  \caption{Random-permutation and original $|$\textit{SE}-values$|$ for predictors of anticipatory gaze rates in Experiment 2.}
+\end{figure}
+
+\section{Pairwise developmental tests: Real vs. randomly permuted effects}
+\setcounter{figure}{0}
+
+In each of the plots below, the dot represents the original data value for the effect and the 5,000 randomly permuted data effect sizes are shown in the distribution.
+The percentage shown is the percentage of random permutation values exceeded by the original data value (taking the absolute value of all data points for a two-tailed test.)
+
+\begin{figure}[!htb]
+\begin{center}
+\includegraphics[width=0.8\textwidth]{figures/E1-child-randvsreal-ttest-agebylg.png}
+\end{center}
+\caption{Pairwise comparisons of the language condition effect across ages in Experiment 1.} 
+\label{fig:E1-lgageinteraction}
+\end{figure}
+
+\begin{figure}[!htb]
+\begin{center}
+\includegraphics[width=0.9\textwidth]{figures/E2-child-randvsreal-ttest-muffledages.png}
+\end{center}
+\caption{Significant pairwise comparisons of the \textit{prosody only}-\textit{no speech} linguisitic condition effect, across ages in Experiment 2} 
+\label{fig:E2-lgageinteraction}
+\end{figure}
+
+\begin{figure}[!htb]
+\begin{center}
+\includegraphics[width=0.9\textwidth]{figures/E2-child-randvsreal-ttest-normaltypesages.png}
+\end{center}
+\caption{Significant pairwise comparisons of the \textit{normal speech}-\textit{no speech} language condition effect for questions status, across ages, in Experiment 2.} 
+\label{fig:E2-lgagetypeinteraction}
+\end{figure}
+
+\section{Non-convergent models}
+\setcounter{figure}{0}
 

### 2015-12-04 paper/Letters/ResponseLetter-Dec2015.tex
resubmitted

@@
 We apologize for our delay in getting the manuscript resubmitted; it took quite a long time to implement one of the suggestions from Reviewer 3, mostly because I had to learn quite a bit of new technical skill to carry his or her suggested analysis out on our computation-heavy data set.
-In addition to that, we have included a two other follow-up analyses in the Supplementary materials, both inspired by the other reviewers' comments.
+In addition to that, we have included a two other follow-up analyses in the Appendix, both inspired by the other reviewers' comments.
 We have tried to make the text more succinct and the results and discussion sections more distinct and straightforward, following requests from all three reviewers.
@@
 
-\noindent We have tried our best to clarify this description (Section XX) and have added specific pieces of information where requested.
+\noindent We have tried our best to clarify this description (Section 2.2.2) and have added specific pieces of information where requested.
 
@@
 \noindent This alternative hypothesis makes a valuable test of what we measured in the experiments.
-In the supplementary materials you can now find a graph comparing theoretically boredom-driven lookers and our real lookers.
+In the Appendix you can now find a graph comparing theoretically boredom-driven lookers and our real lookers.
 We modeled the proportion of participants looking at the current speaker for our real data and for (fake) boredom-driven lookers (who look away from the current speaker at a constant rate, starting 1 second after the onset of speech) across turns of different length.
 Even in short turns, where there is very little temporal room for children to behave differently from the boredom-driven lookers, there is a clear difference between the boredom-driven and the real data.
-(Figure XX).
+(Figure D.2).
 We take this to mean that children are unlikely to be looking away simply due to boredom.
@@
 
-\noindent Thank you for pointing this out! We have added a little more information about how each variable was coded and have put summary tables of every model's output into the text.
+\noindent We have added a little more information about how each variable was coded and have put summary tables of every model's output into the text.
 
@@
 Following your suggestion, we have therefore done two things.
-First, we added follow-up two-tailed t-tests, making pairwise comparisons between age groups for significant interactions so that we could find out how the interacting factor (e.g., language condition) significantly changes across age groups (Sections XX).
-Second, we have added individual models of the youngest age groups (ages 1, 2, and 3), comparing their real looking behavior to our randomly permuted data across conditions (Section XX).
-These follow-up analyses, which are summarized in the text, are primarily described in the supplementary materials (Section XX).
+First, we added follow-up two-tailed t-tests, making pairwise comparisons between age groups for significant interactions so that we could find out how the interacting factor (e.g., language condition) significantly changes across age groups (Sections B.1--3).
+Second, we have added individual models of the youngest age groups (ages 1, 2, and 3), comparing their real looking behavior to our randomly permuted data across conditions (at the ends of Sections 2.2.2 and 3.3.2).
+These follow-up analyses, which are summarized in the text, are primarily described in the Appendix (Section B).
 
@@
 \noindent We thank the reviewer for his or her careful attention to our methods section (and its missing details.) We have added in the experimental methods information requested and have re-checked each section for missing information someone might need to replicate our methods.
-We've also worked to clarify the description of the baseline analysis, including a new figure to accompany it (Section XX; Figure XX).
+We've also worked to clarify the description of the baseline analysis, including a new figure to accompany it (Section 2.2.2; Figure 4).
 We hope that it is now easier to understand.
@@
 We have significantly shortened the introduction and tried to remove redundancies where possible in the rest of the text.
-We have also moved some of the secondary analyses to the Supplementary Materials section so that the main text can be read without getting into the nitty gritty of every single analysis.
+We have also moved some of the secondary analyses to the Appendix so that the main text can be read without getting into the nitty gritty of every single analysis.
 We hope that this has helped make the main text more straightforward and approachable length-wise.
@@
 \noindent We thank Reviewer 3 for this point.
-Our design does not fully cross puppet pair (e.g., robots, blue puppets) with linguistic condition (e.g., ``words only'' and ``no speech'').
+We repeat here some of the information now available in the Appendix for easy summary: Our design does not fully cross puppet pair (e.g., robots, blue puppets) with linguistic condition (e.g., ``words only'' and ``no speech'').
 Even though each puppet pair is associated with different conversation clips across children (e.g.
@@
 Thankfully, we also ran a short follow-up study at the museum with 3--5-year-olds in which each child only saw one video---the normal speech conversation about birthday parties---with a randomly assigned puppet pair performing the conversation.
-Five children watched each puppet pair, for a total of 30 children across the six pairs (Figure XX).
+Five children watched each puppet pair, for a total of 30 children across the six pairs (shown Figure 3).
 This experiment holds all things constant except for the appearance of the puppets.
-We then used a mixed effects logistic regression of children's anticipatory switches (yes or no at each transition), with puppet pair (robots/merpeople/fancy dress/other-3; Figure XX) as a fixed effect and participant and turn transition as random effects.
+We then used a mixed effects logistic regression of children's anticipatory switches (yes or no at each transition), with puppet pair (robots/merpeople/fancy dress/normal-speech-puppets) as a fixed effect and participant and turn transition as random effects.
 In four versions of this model we systematically varied the reference level to check for differences between every puppet pair, finding no significant affects of puppet type on switching rate.
 We take this as evidence that, although we did not fully cross puppet pairs and linguistic conditions in Experiment 2, it was unlikely to have had strong effects on children's looking rates above and beyond the intended effects of linguistic condition.
-We have included details and graphs for this analysis in the Supplementary Materials.
+We have included details and graphs for this analysis in the Appendix.
 
@@
 We have also cleaned up the way we talk about the significant effects in the model output because they were sometimes misleading.
-There was no concern about dummy coding since this is carried out automatically across factor predictors by the glmer package in R (as is now clearer with the presentation of results in tables XX and XX).
+There was no concern about dummy coding since this is carried out automatically across factor predictors by the glmer package in R (as is now clearer with the presentation of results in Tables 2 and 5).
 
@@
 To analyze our data on this scale, the first author had to learn some new R libraries, optimization techniques, and how to use the local computer cluster, dealing with many technical problems along the way!
-That being said, we think that the new permutation analysis is indeed simpler and more theoretically intuitive than our prior method (described in XX and XX, visualized in XX and XX).
+That being said, we think that the new permutation analysis is indeed simpler and more theoretically intuitive than our prior method (described in 2.2.2 and 3.3.2, visualized in Section A).
 Please note that there were two problems with conducting the analyses as originally instructed in the reviewer's comments: (a) we used \textit{t} values instead of \textit{$\beta$} estimates because the standard error estimates for the randomly permuted data were nearly categorically higher, dramatically so in some cases, and (b) many of the models (especially in Experiment 2) resulted in convergence warnings, which we have dealt with by only reporting results for converging models.
@@
 
-A second, but related theoretical takeaway is that lexical information alone is not equivalent to full linguistic information for children, as it has been shown to be (XX) and replicated here (XX) with adults.
+A second, but related theoretical takeaway is that lexical information alone is not equivalent to full linguistic information for children, as it has been shown to be (e.g., de Ruiter et al., 2006) and replicated here (Tables 2 and 5) with adults.
 If this effect arose in our data because children do not respond well to phonetically manipulated speech in conversational contexts, this point can be overturned by work that controls linguistic information through other means.
-If not, it suggests that there is something specially informative about the combined prosodic-lexical marking on question turns, for which there is also prior support (XX) though our finding would be the first to show such effects developmentally.
+If not, it suggests that there is something specially informative about the combined prosodic-lexical marking on question turns, for which there is also prior support (Torreira et al., 2015) though our finding would be the first to show such effects developmentally.
 
@@
 low parent-infant interaction styles) might affect children's early predictive skill.
-It also bridges prior work demonstrating that children begin taking non-linguistic turns in infancy (XX), but still have trouble integrating linguistic aspects of responding at age 3--4 (XX).
+It also bridges prior work demonstrating that children begin taking non-linguistic turns in infancy (e.g., Hilbrink et al., 2015), but still have trouble integrating linguistic aspects of responding at age 3--4 (Casillas et al., in press; Garvey \& Berninger, 1984).
 

### 2015-12-04 paper/CasillasFrank-Anticip.tex
resubmitted

@@
 
+\linespread{2}
 \journal{}
@@
 Our findings suggest that children and adults use linguistic cues to make predictions about upcoming turn structure, but that they primarily do so to predict speaker transitions after a question.
-We also find that children make more predictions than  expected by chance, even at age XX, but that it takes several years before they recruit linguistic cues to anticipate responses more often after questions.
+We also find that children make more predictions than expected by chance at age two, but that it takes several years before they recruit linguistic cues to anticipate responses more often after questions.
 We found no direct evidence of an early prosody advantage in children's anticipations and, further, no evidence that prosodic or lexical cues alone can substitute for their combination in the full linguistic signal, as is proposed for adults \citep{de-ruiter2006}.
@@
 For the children's data, the original \textit{t}-values for language condition, gap duration, the age-language condition interaction, and the language condition-transition type interaction were all greater than 95\% of \textit{t}-values for the randomly permuted data (99.9\%, 95.5\%, 99.4\%, and 96\%, respectively).
-Similarly, the adults' data showed significant differentiation from the randomly permuted data for two of the three originally significant predictors---transition type and the transition type-gap duration interaction (greater than 99.9\% and 99.7\% of random \textit{t}-values, respectively)---with marginal differentiation for the interaction of language condition and transition type (greater than 94.7\% of random \textit{t}-values).
+Similarly, the adults' data showed significant differentiation from the randomly permuted data for two of the three originally significant predictors---transition type and the transition type-gap duration interaction (greater than 99.9\% and 99.7\% of random \textit{t}-values, respectively)---with marginal differentiation for the interaction of language condition and transition type (greater than 94.6\% of random \textit{t}-values).
 
@@
 	\hline
-    Condition$=$\textit{muffled}         &   -1.63429  &   0.86390 &  -1.892   & 0.0585 .      \\
-    Age*Condition$=$\textit{muffled}       &   0.39317 &     0.18907 &   2.080  &  0.0376 *      \\
-    Condition$=$\textit{muffled}*     &  1.77190  &   1.24864  &  1.419 &   0.1559        \\
+    Condition$=$\textit{prosody}         &   -1.63429  &   0.86390 &  -1.892   & 0.0585 .      \\
+    Age*Condition$=$\textit{prosody}       &   0.39317 &     0.18907 &   2.080  &  0.0376 *      \\
+    Condition$=$\textit{prosody}*     &  1.77190  &   1.24864  &  1.419 &   0.1559        \\
     \hspace*{5mm} Type$=$\textit{non-Question} &&&& \\
-    Age*Condition$=$\textit{muffled}* &  -0.47057 &    0.28703 &  -1.639 &   0.1011        \\
+    Age*Condition$=$\textit{prosody}* &  -0.47057 &    0.28703 &  -1.639 &   0.1011        \\
     \hspace*{5mm} Type$=$\textit{non-Question} &&&& \\
 	\hline
-    Condition$=$\textit{robot}            &  -0.26741 &    0.59071 &  -0.453  &  0.6508        \\
-    Age*Condition$=$\textit{robot}           &   0.13740  &   0.13568   & 1.013  &  0.3112        \\
-    Condition$=$\textit{robot}*     &   -1.02193  &   1.01227  & -1.010  &  0.3127        \\
+    Condition$=$\textit{words}            &  -0.26741 &    0.59071 &  -0.453  &  0.6508        \\
+    Age*Condition$=$\textit{words}           &   0.13740  &   0.13568   & 1.013  &  0.3112        \\
+    Condition$=$\textit{words}*     &   -1.02193  &   1.01227  & -1.010  &  0.3127        \\
     \hspace*{5mm} Type$=$\textit{non-Question} &&&& \\
-    Age*Condition$=$\textit{robot}*  &   0.08946  &   0.22349   & 0.400 &   0.6890        \\
+    Age*Condition$=$\textit{words}*  &   0.08946  &   0.22349   & 0.400 &   0.6890        \\
     \hspace*{5mm} Type$=$\textit{non-Question} &&&& \\
@@
 	\hline
-    Condition$=$\textit{muffled}     &     0.3349  &    0.8965  &  0.374 & 0.708692     \\ 
-    Condition$=$\textit{muffled}*  &  0.6627  &    1.2138 &   0.546 & 0.585108     \\ 
+    Condition$=$\textit{prosody}     &     0.3349  &    0.8965  &  0.374 & 0.708692     \\ 
+    Condition$=$\textit{prosody}*  &  0.6627  &    1.2138 &   0.546 & 0.585108     \\ 
     \hspace*{5mm} Type$=$\textit{non-Question} &&&& \\
 	\hline
-    Condition$=$\textit{robot}       &     1.5938  &    0.7208 &   2.211 & 0.027023 *   \\ 
-    Condition$=$\textit{robot}*   &  -1.1265  &    0.9109  & -1.237 & 0.216201     \\ 
+    Condition$=$\textit{words}       &     1.5938  &    0.7208 &   2.211 & 0.027023 *   \\ 
+    Condition$=$\textit{words}*   &  -1.1265  &    0.9109  & -1.237 & 0.216201     \\ 
     \hspace*{5mm} Type$=$\textit{non-Question} &&&& \\
@@
 \end{center}
-\caption{Significant pairwise comparisons of the \textit{normal speech}-\textit{no speech} language condition effect for questions status, across ages, in Experiment 2.} 
+\caption{Significant pairwise comparisons of the \textit{normal speech}-\textit{no speech} language condition effect for transition type, across ages, in Experiment 2.} 
 \label{fig:E2-lgagetypeinteraction}
@@
 \section{Non-convergent models}
+\setcounter{table}{0}
+
+Non-convergent models made up 22--24\% of the 5,000 models of randomly permuted data in Experiment 1 and 69--70\% of the 5,000 for Experiment 2.
+We excluded these non-convergent models because they displayed erratic $\beta$ and \textit{SE} estimates, as summarized below in the table of \textit{t}-values from convergent and non-convergent models in Experiment 1.
+The non-convergent models from Experiment 2 showed similar patterns.
+The high frequency of problematic models persisted even when we changed optimizers and we suspect the issue derives from data sparsity in some of the random runs.
+
+\begin{table}[h]
+  \begin{scriptsize}
+\centering
+  \begin{tabular}{lccccc}
+    Variable & Mean & Median & \textit{SD} & Minimum & Maximum\\ 
+    \hline
+	\textbf{\textit{Children}} &&&&& \\
+    \hline
+    (Intercept)						&	-2.52 (-458.42)	& -2.54 (-2.86)	& 0.87 (1319.22)		& -5.53 (-8185.36)	& 0.41 (0.97)			\\
+    Age								& -0.51 (-17.83)		& -0.49 (-0.53)	& 0.79 (83.78)		& -3.71 (-672.2)		& 2.3 (342.8)			\\
+    LgCond					& -0.53 (-109.91)		& -0.55 (-0.63)	& 0.93 (564.42)		& -3.93 (-4418.74)	& 3.23 (2296.19)		\\
+    Type							& -0.10 (-29.66)		& -0.09 (-0.1)		& 0.98 (515.12)		& -4.06 (-4383.92)	& 3.23 (2296.19)		\\
+    Duration						& 0.99 (345.53)		& 0.98 (1.15)		& 1.07 (1323.13)		& -2.44 (-5048.24)	& 3.36 (3416.68)		\\
+    Age*LgCond				& 0.19 (10.64)		& 0.2 (0.18)		& 0.9 (109.6)			& -3.31 (-581.61)		& 5.78 (9985.16)		\\
+    Age*Type						& 0.02 (-1.8)			& 0.001 (-0.05)	& 0.9 (98.27)			& -3.36 (-884.36)		& 3.59 (946.81)		\\
+    LgCond*Type			& 0.2 (45.32)			& 0.2 (0.27)		& 0.96 (691.3)		& -3.12 (-4160.06)	& 3.45 (640.43)		\\
+    Age*LgCond*Type		& -0.12 (-14.23)		& -0.12 (-0.15)	& 0.93 (156.72)		& -2.98 (-1318.26)	& 3.39 (5107.64)		\\
+    \hline
+	\textbf{\textit{Adults}} &&&&& \\
+    \hline
+    (Intercept)    					&    -1.63 (-126.14) 	& -1.71 (-1.73)	& 0.97 (713.39)		& -4.08 (-12111.22)	& 2.15 (649.55)		\\
+    LgCond  					&      -0.26 (-679.6) 	& -0.3 (-0.53)		& 1.02 (15894.33)	& -3.45 (-494979.7)	& 3.35 (88581.58)	\\
+    Type 							&       -0.11 (6.29) 	& -0.13 (-0.04)	& 1.11 (501.5)		& -3.85 (-6420.76)	& 3.28 (8177.88)		\\
+    Duration  						&       0.25 (84.09) 	& 0.27 (0.26)		& 1.1 (1152.94)		& -3.25 (-10864.51)	& 3.46 (18540.62)	\\
+    LgCond*Type  			&       0.12 (-242.27) 	& 0.1 (0.34)		& 1.07 (26836.7)		& -3.41 (-62264.27)	& 3.81 (509198.4)	\\
+    LgCond*Duration  		&       0.15 (780.03) 	& 0.16 (0.39)		& 1.04 (44105.02)	& -3.84 (-798498.6)	& 3.55 (1145951)		\\
+    Type*Duration 				&        0.05 (-6.56) 	& 0.05 (0.02)		& 1.13 (1389.9)		& -3.54 (-15979.22)	& 3.87 (16419.46)	\\
+    LgCond*Type*Duration  &      -0.06 (1083.63)	& -0.08 (-0.21)	& 1.1 (63116.54)		& -4.21 (-1201895)	& 4.02 (1284965)		\\
+    \hline
+  \end{tabular}
+  \caption{Estimated \textit{t}-values for each predictor in in the adult and child models from Experiment 1. Converging (`C') and non-converging (`non-C') values are shown as: `C (non-C)'. }
+\label{tab:nonconv_e1}
+ \end{scriptsize}
+\end{table}
+
+\clearpage
+
+\section{Miscellaneous}
 \setcounter{figure}{0}
 
+One alternative hypothesis for children's anticipatory gazes is that they simply grow bored and start looking away at a constant rate after a turn begins.
+This data plotted here show a hypothetical group of participants who begin to lose interest (at a linear rate) after one second of a turn (gray dots) compared to participants' real data from Experiment 2 (black dots).
+This pattern suggests that, though children do look away with time, their looks away are not simply driven by boredom.
+
+\begin{figure}[!htb]
+\begin{center}
+\includegraphics[width=0.9\textwidth]{figures/boredom-hypothesis.png}
+\end{center}
+\caption{Proportion participants looking at the current speaker: hypothetical boredom-driven data (gray dots) versus real data from Experiment 2 (black dots).} 
+\label{fig:boredomhypothesis}
+\end{figure}
+
+\bigskip
+
+The design for Experiment 2 does not fully cross puppet pair (e.g., robots, blue puppets) with linguistic condition (e.g., ``words only'' and ``no speech'').
+Even though each puppet pair is associated with different conversation clips across children (e.g.
+robots talking about kitties, birthday parties, or pancakes), robots were only associated with ``words only'' speech, merpeople were only associated with ``prosody only'' speech, and the puppets with fancy clothes were only associated with the ``no speech'' condition.
+We did this to increase the pragmatic felicity of the experiments for the older children (i.e., robots make robot sounds, merpeople's voices are muffled under water, the fancy-clothed puppets are in a room with main other voices).
+It is therefore fair to point out a possible confound between linguistic condition and puppet pair.
+Thankfully, we also ran a short follow-up study at the museum with 3--5-year-olds in which each child only saw one video---the normal speech conversation about birthday parties---with a randomly assigned puppet pair performing the conversation.
+Five children watched each puppet pair, for a total of 30 children across the six pairs.
+This experiment holds all things constant except for the appearance of the puppets.
+We then used a mixed effects logistic regression of children's anticipatory switches (yes or no at each transition), with puppet pair (robots/merpeople/fancy dress/other-3; Figure \ref{fig:speakers}) as a fixed effect and participant and turn transition as random effects.
+In four versions of this model we systematically varied the reference level to check for differences between every puppet pair, finding no significant affects of puppet type on switching rate.
+We take this as evidence that, although we did not fully cross puppet pairs and linguistic conditions in Experiment 2, it was unlikely to have had strong effects on children's looking rates above and beyond the intended effects of linguistic condition.
+
+\begin{figure}[!htb]
+\begin{center}
+\includegraphics[width=0.9\textwidth]{figures/all-puppetdyads-bars.png}
+\end{center}
+\caption{Proportion gaze switches across puppet pairs when linguistic condition and conversation are held constant.} 
+\label{fig:pairconfound}
+\end{figure}
+
 \end{document}

### 2015-12-11 paper/CasillasFrank-Anticip.tex
a few last paper changes for the resubmission

@@
 Even in the English videos, when participants had full access to linguistic cues, their rates of anticipation were relatively low---in fact, comparable to the non-English videos---unless the turn was a question (Figure \ref{fig:E1-randvsreal}).
-Prior work using online, metalinguistic tasks has shown that participants can use linguistic cues to accurately predict upcoming turn ends \citep{torreira2015, magyari12, de-ruiter2006}.
+Prior work using online, metalinguistic tasks has shown that participants can use linguistic cues to accurately predict upcoming turn ends \citep{torreira2015, magyari2012, de-ruiter2006}.
 The current results add a new dimension to our understanding of how listeners make predictions about turn ends: Both children and adults spontaneously monitor the linguistic structure of unfolding turns for cues to upcoming responses.
@@
 
-However, these results do not support the idea that lexical information is sufficient for children, as has been proposed for adults previously \citealp{de-ruiter2006} and replicated in the present study.
+However, these results do not support the idea that lexical information is sufficient for children, as has been proposed for adults previously (\citealp{de-ruiter2006}) and replicated in the present study.
 On the contrary, children showed significant gains in the \textit{normal} speech condition, but \textit{not} in the \textit{words only} condition.
@@
 
-Importantly, we also found no evidence that lexical information alone is equivalent to full linguistic information for children, as has been shown before \cite{magyari2012, de-ruiter2006} and replicated in the current study for adult participants: In both experiments, children's performance was best in conditions when they had access to the full linguistic signal.
+Importantly, we also found no evidence that lexical information alone is equivalent to full linguistic information for children, as has been shown before \citep{magyari2012, de-ruiter2006} and replicated in the current study for adult participants: In both experiments, children's performance was best in conditions when they had access to the full linguistic signal.
 Adults on the other hand, showed significant gains in anticipatory gaze switching in both conditions with lexical cues.
@@
 The question effect itself has implications for our current theories about turn prediction.
-While participants may always use linguistic information to predict upcoming speaker changes (as they have in the two studies presented here), they might not always use linguistic information to predict upcoming turn ends, as is assumed in metalinguistic measures of turn-end prediction (e.g., pressing a button while listening to speech: \citealp{torreira2015, magyari2012, de-ruiter2006}) and as has been the focus of other turn-end prediction studies \citealp{ford1996, duncan1972}.
+While participants may always use linguistic information to predict upcoming speaker changes (as they have in the two studies presented here), they might not always use linguistic information to predict upcoming turn ends, as is assumed in metalinguistic measures of turn-end prediction (e.g., pressing a button while listening to speech: \citealp{torreira2015, magyari2012, de-ruiter2006}) and as has been the focus of other turn-end prediction studies (\citealp{ford1996, duncan1972}).
 Our results rather suggest that participants' \textit{spontaneous} predictions, at least while viewing third-party conversation, are the results of question-monitoring, presumably achieved by recruiting linguistic cues to questionhood from the unfolding signal.

### 2015-12-15 paper/Letters/ResponseLetter-Dec2015.tex
pdf
text changes

@@
 \usepackage{microtype} % Improves typography
-\usepackage{gfsdidot} % Use the GFS Didot font: http://www.tug.dk/FontCatalogue/gfsdidot/
+% \usepackage{gfsdidot} % Use the GFS Didot font: http://www.tug.dk/FontCatalogue/gfsdidot/
 \usepackage[T1]{fontenc} % Required for accented characters
@@
 
-We resubmit our manuscript on ``The development of children's ability to track and predict turn structure in conversation'' to the \textit{Journal of Memory and Language}.
-We apologize for our delay in getting the manuscript resubmitted; it took quite a long time to implement one of the suggestions from Reviewer 3, mostly because I had to learn quite a bit of new technical skill to carry his or her suggested analysis out on our computation-heavy data set.
-In addition to that, we have included a two other follow-up analyses in the Appendix, both inspired by the other reviewers' comments.
-We have tried to make the text more succinct and the results and discussion sections more distinct and straightforward, following requests from all three reviewers.
-We hope that, with these changes, we have greatly improved the paper's clarity.
+\noindent Please accept our resubmission of our manuscript on ``The development of children's ability to track and predict turn structure in conversation'' to the \textit{Journal of Memory and Language}.
+We apologize for our delay in getting the manuscript resubmitted.
+It took quite a long time to implement one of the suggestions from Reviewer 3, mostly because I had to learn quite a bit of new technical skill to carry out a suggested analysis out on our computation-heavy data set.
+Nevertheless, I believe that this analysis has substantially strengthened the paper as a whole, so I appreciate your patience.
+
+\noindent In addition to the new analysis, we have also included two other follow-up analyses in the Appendix, both inspired by the other reviewers' comments.
+And in other revisions, we have tried hard to make the text more succinct and the results and discussion sections more distinct and more straightforward, following requests from all three reviewers.
+
+\noindent We hope that, with these changes, we have greatly improved the paper's clarity.
 Please find a description of specific comments addressed below.
@@
 Please let us know if there is any further information that you require about the submission.
+
+\closing{Sincerely,}
+
+\bigskip 
+\bigskip
 
@@
 
-\closing{Sincerely,}
-
 %----------------------------------------------------------------------------------------
@@
 \noindent We thank all three reviewers for their thorough and helpful comments on the previously submitted manuscript.
-We have tried to address all of the reviewers' comments, within the manuscript and/or here in the response letter.
+We have tried to address all of the reviewers' comments, within the manuscript as well as here in the response letter.
 
@@
 
-We were concerned about this, too, but we unfortunately did not collect saccadic planning norms for children ages 1--6.
-We did not feel comfortable (semi-)arbitrarily assigning saccadic planning times to different ages, so in the newly analyzed data we take the conservative approach of using adult-like planning times (200ms) for children at all ages in both experiments.
+\noindent Thank you for this comment; we were concerned about this issue as well, but we unfortunately did not collect independent saccadic planning norms for children ages 1--6 (and we do not know of any authoritative source for these norms).
+In the newly analyzed data, we take the conservative approach of using adult-like planning times (200ms) for children at all ages in both experiments.
 
@@
 
-\noindent We have tried our best to clarify this description (Section 2.2.2) and have added specific pieces of information where requested.
+\noindent In our revision, we have completely redone all random permutation analyses and have attempted to clarify their description (Section 2.2.2).
 
@@
 
-\noindent This alternative hypothesis makes a valuable test of what we measured in the experiments.
-In the Appendix you can now find a graph comparing theoretically boredom-driven lookers and our real lookers.
-We modeled the proportion of participants looking at the current speaker for our real data and for (fake) boredom-driven lookers (who look away from the current speaker at a constant rate, starting 1 second after the onset of speech) across turns of different length.
-Even in short turns, where there is very little temporal room for children to behave differently from the boredom-driven lookers, there is a clear difference between the boredom-driven and the real data.
-(Figure D.2).
-We take this to mean that children are unlikely to be looking away simply due to boredom.
+\noindent Thanks for pointing out an important alternative hypothesis.
+To address this comment, we modeled the proportion of participants looking at the current speaker for our real data and for (fake) boredom-driven lookers (who look away from the current speaker at a constant rate, starting 1 second after the onset of speech) across turns of different length.
+Even in short turns, where there is very little temporal room for children to behave differently from the boredom-driven lookers, we still see a clear difference between the boredom-driven simulations and the real data (Figure D.2).
+We take this result to be evidence against the alternative hypothesis; we report the details of the analysis in an appendix.
 
@@
 
-\noindent \textit{4. The reported results form the models models are hard to follow.}
+\noindent \textit{4. The reported results from the models models are hard to follow.}
 
-\noindent We have added a little more information about how each variable was coded and have put summary tables of every model's output into the text.
+\noindent We apologize---the data from this experiment are very complex.
+To aid comprehension, we have added more information about how each variable was coded and have put summary tables of every model's output into the text.
+In addition, we have reorganized our results sections in hopes of making them more straightforward.
 
@@
 
-\noindent This is a very useful comment, thanks!
-We agree and have changed our analyses to accommodate it.
-We had started with the ``normal speech'' condition as the reference level because that is what was done in related work on adult turn-end prediction.
-In the end, however, we too preferred the additive model, and have changed the text accordingly.
+\noindent This is another very useful comment, thank you!
+We agree and have changed our analyses to set ``no speech'' as the reference level.
+We had started with the ``normal speech'' condition as the reference level because this choice mirrors the coding in related work on adult turn-end prediction.
+But the additive model implied by the ``no speech'' baseline is easier to interpret.
 
@@
 
-\noindent Thanks for asking for clarification here.
-We have tried to integrate aspects of our answer here into the manuscript.
-The lack of simple language condition effects in the children's data in Experiment 2 does not imply that they were not using linguistic information at all in their predictions (i.e., solely relying on visual information in the stimuli and the duration of inter-turn gap).
-An unexpected but quite important take away of our findings is that children and adults are far more likely to make an anticipatory gaze switch after hearing a question.
-Because speech act seems to be so effective in driving participants' anticipations, the extent to which they use linguistic cues is probably limited to their use of those cues to identify questions in the ongoing turn (which in turn allows them to expect an upcoming speaker switch with high certainty).
-The children's data showed a 2-way effect of age and transition type (question vs.
-non-question), but only in the ``normal'' speech condition, suggesting that it was the only condition in which they could extract the linguistic cues needed to accurately identify questions, perhaps because prosody and lexical cues to questionhood often work together in forming questions during everyday speech, but perhaps because the manipulated speech was simply too unfamiliar for them to apply their normal interactive prediction processing.
+\noindent Thanks for raising this challenging and important point.
+To summarize: We think that the robust question (speech act) effects we observed in both experiments provide strong evidence for the use of linguistic information.
 
-Either way, the existence of the question effect in the normal condition indicates that children do, in fact, use linguistic information (to identify questions) and the interaction of that effect with age suggests that they get better at doing so as they get older.
-These effects, for example, cannot be explained by duration alone (and its effects are already accounted for in the statistical analyses), and there is no visual cue advantage difference between the normal condition and the others.
+In Experiment 2, children's data showed a 2-way interaction of age and transition type (question vs.
+non-question), but only in the ``normal'' speech condition, suggesting that it was the only condition in which they could extract the linguistic cues needed to accurately identify questions.
+We speculate that this is because prosody and lexical cues to questionhood often work together in forming questions during everyday speech, but it could also be that the manipulated speech was simply too unfamiliar for them to predict effectively.
+Either way, the existence of the question effect in the normal condition indicates that children do, in fact, use linguistic information (to identify questions) and the interaction with age suggests that they get better at doing so as they get older.
+These effects cannot be explained by duration alone (and duration is controlled in our analysis), and there is no visual cue advantage difference between the normal condition and the others.
 We are then left to conclude that they use linguistic information to support the question effect.
+
 A similar story can be told for the data in Experiment 1: non-verbal cues to questions vs.
@@
 Again, this cannot be explained by gap duration or unequal visual advantages across the stimuli.
+Thus, we believe that both experiments provide support for the use of linguistic cues, albeit for a subset of these cues.
 
-The fact that we have found most of our linguistic processing effects within an effect of speech act paints a very different picture from prior work (e.g., de Ruiter et al., 2006), which finds much clearer evidence of linguistic processing across multiple phonetically manipulated conditions, but without question-non-question comparisons.
-No turn prediction measure yet is capable of capturing predictions that occur in spontaneous first-person conversation, but we think that our results indicate that there are multiple types of prediction relevant to anticipating upcoming turn structure during interaction, and that any account of how listeners use linguistic cues while participating in conversation should be able to account for shifts in how listeners attend to cues to make more or less specific predictions (i.e., ``what words are upcoming'' vs.
+% The fact that we have found most of our linguistic processing effects within an effect of speech act differs from prior work (e.g., de Ruiter et al., 2006), however, where there was clearer evidence of linguistic processing across multiple phonetically manipulated conditions, but without question-non-question comparisons.
+
+% More work is needed to square these two patterns of data, but we do believe that the presence o 
+% No turn prediction measure yet is capable of capturing predictions that occur in spontaneous first-person conversation, but we think that our results indicate that there are multiple types of prediction relevant to anticipating upcoming turn structure during interaction, and that any account of how listeners use linguistic cues while participating in conversation should be able to account for shifts in how listeners attend to cues to make more or less specific predictions (i.e., ``what words are upcoming'' vs.
 ``what types of action are upcoming'').
-
-In sum, while the linguistic effects in our data are somewhat limited, they are present, consistent, developmentally changing, and important to our developing theories about turn prediction.
+% In sum, while the linguistic effects in our data are somewhat limited, they are present, consistent, developmentally changing, and important to our developing theories about turn prediction.
 We believe that this sets up future work quite well to dig into the specific linguistic cues (lexical and prosodic, though probably mostly lexical) that are being used in spontaneous question identification.
@@
 
-\noindent We have tried our best to make this clearer, first by clarifying the statistical analyses and outputs, and second by trying to explicitly link each claim to an individual result (through the use of figure and table references).
+\noindent We apologize.
+In our revision, we have tried our best to make these links clearer, first by clarifying the statistical analyses and outputs, and second by trying to explicitly link each claim to an individual result (through the use of figure and table references).
 We hope that this will help readers better connect the results to our interpretations.
@@
 
-\noindent The single upside to taking so long in implementing these comments is that all of the unpublished work referred to previously is now published or in press!
+\noindent The single upside to our delayed resubmission is that all of the unpublished work referred to previously is now published or in press!
 
@@
 
-\noindent Thanks!
-We have tried to improve the linkage between our claims and analyses and now have presented the statistical results in tables, as requested.
+\noindent As discussed above, we have tried to improve the linkage between our claims and analyses and now have presented the statistical results in tables, as requested.
 Please see responses 4 and 7 under Reviewer 1 for a little more detail.
+
 The main focus of our analyses was the effect that age (as a continuous variable) has on anticipatory gaze.
 However there are two cases in which our main analyses are not satisfying without further, age-group analyses: (a) age sometimes showed itself to interact with other factors (e.g., age and language condition), and (b) without further tests we can not say whether particular age groups themselves statistically differ from chance, across or within conditions.
-Following your suggestion, we have therefore done two things.
+
+Following your suggestion, we have thus made two additional changes.
 First, we added follow-up two-tailed t-tests, making pairwise comparisons between age groups for significant interactions so that we could find out how the interacting factor (e.g., language condition) significantly changes across age groups (Sections B.1--3).
@@
 \noindent Our original intention in combining results and discussion was to make the findings easier to understand.
-Since we were not able to achieve this in the prior submission, we have followed Reviewer 2's advice and now use a more traditional results-only then discussion-only structure.
-We think that this has improved the clarity of the paper and we give thanks for the suggestion.
+Since we were not able to achieve this in the prior submission, we have followed Reviewer 2's advice and now use a more traditional ``results-first'' structure.
+We hope this change has improved the clarity of the paper.
 
@@
 
-\noindent We thank the reviewer for his or her careful attention to our methods section (and its missing details.) We have added in the experimental methods information requested and have re-checked each section for missing information someone might need to replicate our methods.
-We've also worked to clarify the description of the baseline analysis, including a new figure to accompany it (Section 2.2.2; Figure 4).
-We hope that it is now easier to understand.
+\noindent We thank the reviewer for careful attention to our methods section (and its missing details).
+We have added in the experimental methods information requested and have re-checked each section for missing information that might be useful in replicating our methods.
+We've also worked to clarify the description of the baseline analysis (now reported in depth in the Appendix), including a new figure (Section 2.2.2; Figure 4).
 
@@
 From these points we conclude that our experiments do indeed tap into participants' use of linguistic cues to predict upcoming turn structure.
-That isn't to say that the method wouldn't be dramatically improved with more natural linguistic controls and tighter controls on gap duration---it would!
+ % That isn't to say that the method wouldn't be improved by more natural linguistic controls and tighter controls on gap duration---it would!
 But we think our current results still make a valuable contribution to the current understanding of turn anticipation in children \textit{and} adults, and lays critical groundwork for discovering what specific types of cues participants spontaneously monitor in everyday interaction.
@@
 
-\noindent These papers are now all published or in press.
-The errors in the reference list were partly due to half-filled BibTex entries used to produce the previous manuscript.
-We have checked them again and hope that there are very few (or no) errors remaining.
+\noindent These papers are now all published or in press. We apologize for errors in the reference list; we have checked them again and hope that there are no remaining errors.
 
@@
 
-\noindent We agree that the paper was overly long.
-We have significantly shortened the introduction and tried to remove redundancies where possible in the rest of the text.
+\noindent We have significantly shortened the introduction and tried to remove redundancies where possible in the rest of the text.
 We have also moved some of the secondary analyses to the Appendix so that the main text can be read without getting into the nitty gritty of every single analysis.
@@
 robots talking about kitties, birthday parties, or pancakes), robots were only associated with ``words only'' speech, merpeople were only associated with ``prosody only'' speech, and the puppets with fancy clothes were only associated with the ``no speech'' condition.
-We did this to increase the pragmatic felicity of the experiments for the older children (i.e., robots make robot sounds, merpeople's voices are muffled under water, the fancy-clothed puppets are in a room with main other voices).
+We made this choice to increase the pragmatic felicity of the experiments for the older children (i.e., robots make robot sounds, merpeople's voices are muffled under water, the fancy-clothed puppets are in a room with main other voices).
 It is therefore fair to point out a possible confound between linguistic condition and puppet pair.
-Thankfully, we also ran a short follow-up study at the museum with 3--5-year-olds in which each child only saw one video---the normal speech conversation about birthday parties---with a randomly assigned puppet pair performing the conversation.
+
+Thankfully, we also ran a short follow-up study at the museum with 3--5-year-olds that addresses this issue.
+In the follow-up study, each child only saw one video---the normal speech conversation about birthday parties---with a randomly assigned puppet pair performing the conversation.
 Five children watched each puppet pair, for a total of 30 children across the six pairs (shown Figure 3).
@@
 We take this as evidence that, although we did not fully cross puppet pairs and linguistic conditions in Experiment 2, it was unlikely to have had strong effects on children's looking rates above and beyond the intended effects of linguistic condition.
-We have included details and graphs for this analysis in the Appendix.
+We have included details and graphs for this norming experiment in the Appendix.
 
@@
 
-\noindent We completely agree with the reviewer on this point.
-We did not have the ability to create custom animated conversation videos at the time.
-But, since then, the first author has developed a method for doing exactly that.
-Unfortunately, for this past work, we are still stuck with the puppet stimuli.
+\noindent We agree with the reviewer that computer animation is one method for providing greater experimental control.
+We did not have the ability to create custom animated conversation videos at the time the research was conducted, though we do use it in some ongoing work.
+Nevertheless, we believe that the control experiment reported above provides some data that speaks against a ``puppet saliency'' hypothesis accounting for the results of Experiment 2.
 
@@
 
-\noindent We corrected small mistakes in the model description (thanks for spotting them!), for example, making it clear that we used a logistic regression (for our binary response variable of switch-no switch) and have added short bits of information justifying our use of separate models for adults and children and our use of random slopes.
-We have also cleaned up the way we talk about the significant effects in the model output because they were sometimes misleading.
-There was no concern about dummy coding since this is carried out automatically across factor predictors by the glmer package in R (as is now clearer with the presentation of results in Tables 2 and 5).
+\noindent In our revision, we correct some small issues in our model description, for example, making it clear that we used a logistic regression (for our binary response variable of switch-no switch).
+We have also added a justification of our use of separate models for adults and children and our decisions regarding random slopes.
+We have also cleaned up the way we talk about the significant effects in the model output.
+In particular, we believe that the concern regarding dummy coding was a misunderstanding; dummy coding was in fact used for categorical predictors (as is now clearer with the presentation of results in Tables 2 and 5).
 
@@
 
-\noindent We really liked this suggestion but felt that it was too far outside the scope of the current paper, especially because our stimulus set is probably too small to build a generalizable model of cue usefulness.
-We will, however, seriously consider this type of analysis for future work, with adult and child turn-taking, in experiments and in spontaneous speech corpora.
+\noindent We agree---this is a great suggestion, but it may venture too far outside the scope of the current paper, especially because our stimulus set is probably too small to support building a generalizable model of cue use.
+In addition, the interpretation of our experiments is already complex without the addition of a computational model.
+% We will, however, seriously consider this type of analysis for future work, with adult and child turn-taking, in experiments and in spontaneous speech corpora.
 
@@
 
-\noindent We thank Reviewer 3 very much for his/her suggestion and clear explanation of how to conduct the analysis.
-To be sure about the results we decided to run 5,000 random permutations on the data from each experiment, resulting in 20,000 glmer runs.
-We apologize for the delay of this resubmission, which is primarily due to the implementation of this random permutation analysis.
-To analyze our data on this scale, the first author had to learn some new R libraries, optimization techniques, and how to use the local computer cluster, dealing with many technical problems along the way!
-That being said, we think that the new permutation analysis is indeed simpler and more theoretically intuitive than our prior method (described in 2.2.2 and 3.3.2, visualized in Section A).
-Please note that there were two problems with conducting the analyses as originally instructed in the reviewer's comments: (a) we used \textit{t} values instead of \textit{$\beta$} estimates because the standard error estimates for the randomly permuted data were nearly categorically higher, dramatically so in some cases, and (b) many of the models (especially in Experiment 2) resulted in convergence warnings, which we have dealt with by only reporting results for converging models.
+\noindent We thank Reviewer 3 very much for the suggestion and clear explanation of how to conduct the analysis.
+Unfortunately, this was not a trivial analysis to implement, causing some delays in the manuscript resubmission.
+Running 5,000 random permutations and 20,000 mixed models required a substantial investment in terms of parallel computing infrastructure.
+
+That being said, we think that the new permutation analysis is indeed simpler and more theoretically intuitive than our prior method (described in 2.2.2 and 3.3.2, and visualized in Section A).
+Please note that there were two problems with conducting the analyses as originally instructed in the reviewer's comments.
+1) We used $t$ values instead of $\beta$ estimates because the standard error estimates for the randomly permuted data were categorically higher (due to data sparsity issues).
+The use of $t$ values is closer to the practice in neuroimaging data analysis, where statistical significance is computed based on the empirical distribution of test statistics.
+2) Many of the models resulted in convergence warnings.
+We made the decision to report results across all convergent models, but we are happy to revisit this decision if the Reviewer believes there is an alternative approach that we should follow.
 
@@
 
-\noindent We have tried to improve the clarity of our findings' theoretical importance in the revised manuscript.
-I summarize here the points that we cover.
-One of the most important theoretical contributions of our work is in establishing and replicating speech act effects in children and adults' spontaneous predictions---while participants may always use linguistic information to predict upcoming speaker changes, our results to not support the idea that they always use linguistic information to predict upcoming turn ends, as is assumed in metalinguistic measures of turn-end prediction (e.g., pressing a button while listening to speech).
-Our results suggest that participants' spontaneous predictions are the result of question-monitoring, presumably achieved by recruiting linguistic cues to questionhood from the unfolding signal.
+\noindent We have tried to improve the clarity of our findings' theoretical importance in the revised manuscript. I summarize here the points that we cover.
+
+First, one of the most important theoretical contributions of our work is in establishing and replicating the effect of speech act (questions) in children and adults' spontaneous predictions.
+While participants may always use linguistic information to predict upcoming speaker changes, our results to not support the idea that they always use linguistic information to predict upcoming \emph{turn ends}, as is assumed in metalinguistic measures of turn-end prediction (e.g., pressing a button while listening to speech).
+Instead, our results suggest that participants' spontaneous predictions are the result of question-monitoring, presumably achieved by recruiting linguistic cues to questionhood from the unfolding signal.
 The trends in our data suggest that participants are primarily recruiting lexical cues to do this, though establishing this pattern will require follow up work with more focused stimuli.
 
-A second, but related theoretical takeaway is that lexical information alone is not equivalent to full linguistic information for children, as it has been shown to be (e.g., de Ruiter et al., 2006) and replicated here (Tables 2 and 5) with adults.
-If this effect arose in our data because children do not respond well to phonetically manipulated speech in conversational contexts, this point can be overturned by work that controls linguistic information through other means.
+Second, lexical information alone is not equivalent to full linguistic information for children, as it has been shown to be with adults (e.g., de Ruiter et al., 2006; replicated in our work).
+While this result will require replication with different stimulus items, it still points to an important developmental effect.
+% If this effect arose in our data because children do not respond well to phonetically manipulated speech in conversational contexts, this point can be overturned by work that controls linguistic information through other means.
 If not, it suggests that there is something specially informative about the combined prosodic-lexical marking on question turns, for which there is also prior support (Torreira et al., 2015) though our finding would be the first to show such effects developmentally.
 
-A third theoretical contribution is that young children (e.g., at ages one and two) do make anticipatory gaze prediction more often than would be expected by chance alone, but not by much.
+Third, young children (e.g., at ages one and two) do make anticipatory gaze prediction more often than would be expected by chance alone (albeit at low rates).
 Since older children and adults' gaze shifts are primarily driven by question turns, this may suggest that, although children \textit{can} make predictions about upcoming turn structure, it doesn't count for much until they have acquired the linguistic skill to pick out question turns.
@@
 
-The fourth contribution is in our interpretation of discrepancies between our methods and results compared to prior work (using more explicit, metalinguistic methods and showing more obvious online predictions using linguistic information).
+% Fourth, our work helps understand we  our interpretation of discrepancies between our methods and results compared to prior work (using more explicit, metalinguistic methods and showing more obvious online predictions using linguistic information).
 The analyses of these discrepancies led us to put forth a new idea: participants in first-person spontaneous conversation may use multiple strategies in making predictions about upcoming turn structure, using more passive prediction to detect upcoming speaker transitions (e.g., questions), switching into precise turn-end prediction mode when necessary.

### 2015-12-15 paper/CasillasFrank-Anticip.tex
text changes

@@
 
-\linespread{2}
+% \linespread{2}
 \journal{}
@@
 In two eye-tracking experiments, we measured children's anticipatory gaze to upcoming responders while controlling linguistic cues to upcoming turn structure.
-In Experiment 1, we showed English and non-English conversations to English-speaking participants, finding minimal differences between the predictive looking behavior of preschoolers and adults.
-In Experiment 2, we phonetically controlled lexicosyntactic and prosodic cues in English-only speech, finding that children's predictive looking behavior improved from ages one to six, but that even one-year-olds made more anticipatory looks than would be expected by chance.
+In Experiment 1, we showed English and non-English conversations to English-speaking adults and children.
+%, finding minimal differences between the predictive looking behavior of preschoolers and adults. 
+In Experiment 2, we phonetically controlled lexicosyntactic and prosodic cues in English-only speech. 
+Children's predictive looking behavior improved from ages one to six, but even one-year-olds made more anticipatory looks than would be expected by chance.
 In both experiments, children and adults anticipated more often after hearing questions. 
@@
 But, unlike adults, lexical cues alone were also not sufficient to improve prediction---children's performance was best overall with access to lexicosyntax and prosody together.
-Our findings support an account in which turn prediction emerges in infancy, but takes several years before becoming fully integrated with linguistic processing.
+Our findings support an account in which turn prediction emerges in infancy, but becomes fully integrated with linguistic processing only gradually.
 \end{abstract}
@@
 Turn-taking behaviors are also cross-culturally consistent in their basic features and the details of their implementation \citep{de-vos2015, dingemanse2013, stivers2009}.
-How does this ability develop?
+% How does this ability develop?
 
 Children participate in sequential coordination with their caregivers starting at three months of age---before they can rely on any linguistic cues in taking turns \citep[see, among others, ][]{bateson1975, hilbrink2015, jaffe2001, snow1977}.
-Of course, infant turn taking is different from adult turn taking in several ways.
+Infant turn taking is different from adult turn taking in several ways, however.
 Infant turn taking is heavily scaffolded by caregivers, has different timing from adult turn taking, and lacks semantic content \citep{hilbrink2015, jaffe2001}.
-However, children's early, turn-structured social interactions are presumably a critical precursor to their later conversational turn taking: Early non-verbal interactions likely establish the protocol by which children come to use language with others.
-How do children integrate linguistic knowledge with these preverbal turn-taking abilities?
+But children's early, turn-structured social interactions are presumably a critical precursor to their later conversational turn taking: Early non-verbal interactions likely establish the protocol by which children come to use language with others.
+How do children integrate linguistic knowledge with these preverbal turn-taking abilities, and how does this integration change over the course of childhood?
 
 In this study, we investigate when children begin to make predictions about upcoming turn structure in conversation, and how they integrate language into their predictions as they grow older.
-In what follows, we first give a basic review of turn-taking research and the state of current knowledge about adult turn prediction.
+In the remainder of the introduction, we first give a basic review of turn-taking research and the state of current knowledge about adult turn prediction.
 We then discuss recent work on the development of turn-taking skills before turning to the details of our own study.
@@
 
-Despite this complexity, conversational turn taking is often precise in its timing, and this precision requires prediction.
+Despite this complexity, conversational turn taking is precise in its timing.
 Across a diverse sample of conversations in 10 languages, one study found a consistent average turn transition time of 0--200 msec at points of speaker switch \citep{stivers2009}.
@@
 
-Despite this evidence, the role of prosody for adult turn prediction is still a matter of debate.
+Despite this body of evidence, the role of prosody for adult turn prediction is still a matter of debate.
 De Ruiter and colleagues' (2006) experiment focused on the role of intonation, which is only a partial index of prosody.
-Prosodic structure is also tied closely to the syntax of an utterance, and so the two linguistic signals are difficult to control independently \citep{ford1996}.
+And in addition, prosody is tied closely to the syntax of an utterance, so the two linguistic signals are difficult to control independently \citep{ford1996}.
 \citet*{torreira2015} used a combination of button-press and verbal responses to investigate the relationship between lexicosyntactic and prosodic cues in turn-end prediction.
@@
 These experimental findings corroborate other corpus and experimental work promoting a combination of cues (lexicosyntactic, prosodic, and pragmatic) as key for accurate turn-end prediction \citep{duncan1972, ford1996, hirvenkari2013}.
-We next turn to evidence on children's developing turn-taking ability.
 
@@
 Among those, only 36-month-olds were affected by a lack of intonation contours, leading Keitel and colleagues to conclude that children's ability to predict upcoming turn structure relies on their ability to comprehend the stimuli lexicosemantically.
-They also suggest that intonation takes a secondary role in turn prediction, following their and other studies findings with adults (e.g., \citealp{de-ruiter2006}), but only \textit{after} children acquire more sophisticated, adult-like language comprehension systems (sometime after 36 months in their data).
+They also suggested that intonation might play a secondary role in turn prediction, but only after children acquire more sophisticated, adult-like language comprehension abilities.
 
 Although the Keitel et al.
-\citeyearpar{keitel2013} study constitutes a substantial advance over previous work in this domain, it has its own limitations.
+\citeyearpar{keitel2013} study constitutes a substantial advance over previous work in this domain, it has some limitations.
 Because these limitations directly inform our own study design, we review them in some detail.
@@
 Instead, they used gaze switches during ongoing speech as a baseline.
-Ongoing speech is perhaps the period in which switching is least likely to occur \citep{hirvenkari2013}, thus maximizing chances of finding a difference between gaze frequency at turn transitions and their baseline rate.
+But ongoing speech is perhaps the period in which switching is least likely to occur \citep{hirvenkari2013}--- thus, this particular baseline maximizes the chance of finding a difference between gaze frequency at turn transitions and baseline.
 A more conservative baseline would be to compare participants' looking behavior at turn transitions to their looking behavior during randomly selected windows of time throughout the stimulus, including turn transitions.
@@
 
+\subsection{The current study}
+
 Our goal in the current study is to find out when children begin to make predictions about upcoming turn structure and to understand how their predictions are affected by linguistic cues across development.
@@
 
-Our findings suggest that children and adults use linguistic cues to make predictions about upcoming turn structure, but that they primarily do so to predict speaker transitions after a question.
-We also find that children make more predictions than expected by chance at age two, but that it takes several years before they recruit linguistic cues to anticipate responses more often after questions.
-We found no direct evidence of an early prosody advantage in children's anticipations and, further, no evidence that prosodic or lexical cues alone can substitute for their combination in the full linguistic signal, as is proposed for adults \citep{de-ruiter2006}.
+Because the results of our experiments are complex, we highlight three primary findings.
+First, although children and adults use linguistic cues to make predictions about upcoming turn structure, they do so primarily in predict speaker transitions after questions (a speech act effect).
+This speech act effect, which we did not initially predict, is intriguing and suggests that previous work may have neglected an important dimension of linguistic cues.
+Second, we find that children make more predictions than expected by chance starting at age two, but that this effect is small with our stimuli.
+Third, we find no evidence of an early prosody advantage in children's anticipations and, further, no evidence that prosodic or lexical cues alone can substitute for their combination in the full linguistic signal, as is proposed for adults \citep{de-ruiter2006}; instead, anticipation is strongest for a stimulus with the full range of cues.
+In sum, our findings support an account in which turn prediction emerges in infancy, but becomes fully integrated with linguistic processing only gradually across development.
 
@@
 \begin{center}
-\includegraphics[width=0.8\textwidth]{figures/E1-samples-by-lang-groups-trans-types.png}
+\includegraphics[width=1\textwidth]{figures/E1-samples-by-lang-groups-trans-types.png}
 \end{center}
@@
 \begin{table}[h!]
-  \begin{scriptsize}
+\begin{small}
 \begin{center}
@@
 
-\begin{figure}[ht!]
+\begin{figure}[t!]
 \begin{center}
@@
 By running the real and randomly permuted data sets through identical statistical models, we can also estimate how likely it is that predictor effects in the original data (e.g., the effect of language condition; Table \ref{tab:E1-models}) arose from random looking.
-
-We completed this baseline procedure on 5,000 random permutations of the original turn transition analysis windows and compared the \textit{t}-values from each predictor in the original models (Table \ref{tab:E1-models}) to the distribution of \textit{t}-values for each predictor in the 5,000 models of the randomly permuted datasets.\footnote{We report \textit{t}-values rather than beta estimates because the standard errors in the randomly permuted data models were much higher than for the original data.
-For those interested, plots of the beta and standard error distributions are available in the Supplementary Materials.} We could then test whether significant effects from the original statistical models differed from the random baseline by calculating the proportion of random data \textit{t}-values exceeded by the original \textit{t}-value for each predictor, using the absolute value of all \textit{t}-values for a two-tailed test.
-For example, children's original ``language condition'' \textit{t}-value was $|$3.429$|$, which is greater than 99.9\% of all $|$\textit{t}-value$|$ estimates from the randomly-permuted data models (i.e., \textit{p}$=$.001).
-This leads us to conclude that that the effect of language condition in the original model was highly unlikely to be the result of random gaze shifting.
-
-We excluded the output of random-permutation models that gave convergence warnings in order to remove unreliable estimates from our analyses (non-converging models were 22.4\% and 24.4\% of all models for children and adults, respectively; see the Supplementary Materials for more information on model exclusion).
-
-Our baseline analyses revealed that none of the significant predictors from models of the original, turn-related data can be explained by random looking (see Figures \ref{fig:E1-ChiTs} and \ref{fig:E1-AduTs} for the full distributional data).
+Because these analyses are complex, we report their full details in Appendix A.
+
+Our baseline analyses revealed that none of the significant predictors from models of the original, turn-related data can be explained by random looking.
 For the children's data, the original \textit{t}-values for language condition, gap duration, the age-language condition interaction, and the language condition-transition type interaction were all greater than 95\% of \textit{t}-values for the randomly permuted data (99.9\%, 95.5\%, 99.4\%, and 96\%, respectively).
@@
 
-\paragraph{Developmental effects}
-The original statistical model of the children's real data revealed a significant interaction of age and language condition (Table \ref{tab:E1-models}) that was highly unlikely to have come from random looking (Figure \ref{fig:E1-randvsreal}).
-To further explore this effect, we compared the average effect of language condition for each age group: We extracted the average difference score for the two language conditions (English minus non-English) for each subject, computing an overall average for each random permutation of the data.
-For each random permutation, we then made pairwise comparisons of the average difference scores across pariticipant age groups, finding that, while 3- and 4-year olds showed similarly large effects of language condition, 5-year-olds showed a significantly smaller effect of language condition, compared to both younger age groups.
+\subsubsection{Developmental effects}
+
+The models reported above revealed a significant interaction of age and language condition (Table \ref{tab:E1-models}) that was unlikely be due to random looking (Figure \ref{fig:E1-randvsreal}).
+To further explore this effect, we compared the average effect of language condition for each age group: Using the permutation analyses above, we extracted the average difference score for the two language conditions (English minus non-English) for each subject, computing an overall average for each random permutation of the data.
+For each random permutation, we then made pairwise comparisons of the average difference scores across pariticipant age groups.
+Details are given in Appendix B.
+
+These analyses showed that, while 3- and 4-year olds showed similarly large effects of language condition, 5-year-olds showed a significantly smaller effect of language condition, compared to both younger age groups.
 In other words, the difference in the effect of language condition for 5-year-olds compared to younger children was larger than would be expected by chance in 99.52\% of the randomly permuted data sets for 3-year-olds and 99.96\% of the data sets for 4-year-olds---differences of \textit{p}$<$.01 and \textit{p}$<$.001, respectively (see Figure \ref{fig:E1-lgageinteraction} for difference score distributions).
@@
 When does spontaneous turn prediction emerge developmentally during natural speech?
-To test whether the youngest age group (3-year-olds) already exceeded chance in their anticipatory gaze switches, we used two-tailed \textit{t}-tests to compare their real gaze rates to the random baseline in the most familiar speech condition for our participants: English.
-We found that three-year-olds made anticipatory gaze switches significantly above chance, both when all transitions were considered (\textit{t}(22.824)$=$-4.147, \textit{p}$<$.001) and for question transitions alone (\textit{t}(21.677)$=$-5.268, \textit{p}$<$.001).
+To test whether the youngest age group (3-year-olds) already exceeded chance in their anticipatory gaze switches, we used two-tailed \textit{t}-tests to compare their real gaze rates to the random baseline in the English condition.
+Although the overall effect was small, we found that three-year-olds made anticipatory gaze switches significantly above chance, when all transitions were considered (\textit{t}(22.824)$=$-4.147, \textit{p}$<$.001) as well as for question transitions alone (\textit{t}(21.677)$=$-5.268, \textit{p}$<$.001).
 
@@
 
-Children and adults spontaneously tracked the turn structure of the conversations, making anticipatory gaze switches at an above-chance rate across all ages and conditions (Table \ref{tab:e1_look}; Figure \ref{fig:E1-randvsreal}).
+Children and adults spontaneously tracked the turn structure of the conversations, making anticipatory gaze switches at an above-chance rate across all ages and conditions.
 Children's anticipatory gaze rates were affected by language condition, transition type, age, and gap duration (Table \ref{tab:E1-models}), none of which could be explained by a baseline of random gaze switching (Figure \ref{fig:E1-ChiTs}).
-
-Language condition (English vs.
+These data show a number of important features that bear on our questions of interest.
+
+First, both adults' and children's anticipations were strongly affected by transition type.
+Both groups made more anticipatory switches after hearing questions, compared to non-questions.
+Even in the English videos, when participants had full access to linguistic cues, their rates of anticipation were relatively low---in fact, comparable to the non-English videos---unless the turn was a question.
+Prior work using online, metalinguistic tasks has shown that participants can use linguistic cues to accurately predict upcoming turn ends \citep{torreira2015, magyari2012, de-ruiter2006}.
+The current results add a new dimension to our understanding of how listeners make predictions about turn ends: Both children and adults spontaneously monitor the linguistic structure of unfolding turns for cues to upcoming responses.
+
+Second, we saw developmental effects such that older children anticipated more reliably, but only in the non-English videos.
+In the English videos all children anticipated, especially foro questions.
+But the language condition interaction suggests that the 5-year-olds were able to leverage anticipatory cues in the non-English videos in a way that 3- and 4-year-olds could not, possibly by shifting more attention to the non-native prosodic or non-verbal cues.
+Prior work on children's turn-structure anticipation proposed that children's turn-end predictions rely primarily on lexicosyntactic structure (and not, e.g., prosody) as they get older \citep{keitel2013}.
+The current results suggest more flexibility in children's predictions; when they do not have access to lexical information, older children and adults are likely to find alternative cues to turn taking behavior.
+
+In Experiment 2 we follow up on these findings, improving on two aspects of the design.
+Our language manipulation in this first experiment was too coarse to provide data regarding specific linguistic cues (e.g., prosody vs.
+lexicosyntax).
+In Experiment 2, to compare lexicosyntactic and prosodic cues directly, controlling for the presence of non-verbal cues, wed use artificial stimuli.
+In addition, we saw above-chance anticipation in the youngest children we tested in Experiment 1.
+Although there were developmental changes in prediction for non-English stimuli, all children anticipated successfully after questions.
+In Experiment 2 we explore a wider developmental range.
+
+% Language condition (English vs.
 non-English) affected children's anticipations in two ways (Table \ref{tab:E1-models};  Figure \ref{fig:E1-randvsreal}).
@@
 
-Children's anticipatory gaze switches were also affected by their age, but only in the non-English videos: 3- and 4-year-olds made many more anticipatory switches when watching videos in English compared to non-English, but this effect of language condition had attenuated significantly by age 5 (Table \ref{tab:E1-models};  Figure \ref{fig:E1-randvsreal}; Figure \ref{fig:E1-lgageinteraction}).
-This interaction suggests that the 5-year-olds were able to leverage anticipatory cues in the non-English videos in a way that 3- and 4-year-olds could not, possibly by shifting more attention to the non-native prosodic or non-verbal cues.
-Prior work on children's turn-structure anticipation proposed that children's turn-end predictions rely primarily on lexicosyntactic structure (and not, e.g., prosody) as they get older \citep{keitel2013}.
-The current results suggest more flexibility in children's predictions; when they do not have access to lexical information, older children are likely to find alternative cues to turn taking behavior.
-
-Finally, children showed an effect of gap duration (Table \ref{tab:E1-models}).
+% Finally, children showed an effect of gap duration (Table \ref{tab:E1-models}).
 This effect is straightforward: Longer gaps resulted in longer analysis windows, yielding more time for children to make an anticipatory gaze.
 
-Adults' anticipatory gaze rates were also affected by transition type, language condition, and gap duration (Table \ref{tab:E1-models}), none of which could be easily explained by a baseline of random gaze switching (Figure \ref{fig:E1-AduTs}).
+% Adults' anticipatory gaze rates were also affected by transition type, language condition, and gap duration (Table \ref{tab:E1-models}), none of which could be easily explained by a baseline of random gaze switching (Figure \ref{fig:E1-AduTs}).
 Like children, adults made more anticipatory switches after hearing questions compared to non-questions, suggesting that anticipation mattered more to them when an immediate response was expected.
-Also like children, the advantage for questions was driven by lexical access such that adults must have relied on lexicosyntactic cues to questionhood in picking out turns that potentially require an immediate response, though this effect was only marginally divergent from the distribution of randomly permuted data (\textit{p}$=$.053; Figure \ref{fig:E1-AduTs}).
+Also like children, the advantage for questions was driven % by lexical access such that adults must have relied on lexicosyntactic cues to questionhood in picking out turns that potentially require an immediate response, though this effect was only marginally divergent from the distribution of randomly permuted data (\textit{p}$=$.053; Figure \ref{fig:E1-AduTs}).
 Finally, adults' anticipation rates were also affected by gap duration, but more so for questions than non-questions (Table \ref{tab:E1-models}), suggesting that adults were less likely overall to make switches at non-questions, and so did not benefit from extra time to do so.
 
-\subsubsection{Summary}
-\label{sec:summary1}
-
-Children and adults' predictions alike were benefited by access to lexical information (English) and speech act status (questionhood), suggesting that linguistic cues, particularly lexical ones, facilitate their spontaneous predictions about upcoming turn structure through the identification of turns with immediate responses.
+% % \subsubsection{Summary}
+% % \label{sec:summary1}
+
+% Children and adults' predictions alike were benefited by access to lexical information (English) and speech act status (questionhood), suggesting that linguistic cues, particularly lexical ones, facilitate their spontaneous predictions about upcoming turn structure through the identification of turns with immediate responses.
 Children's anticipatory gaze rates for questions and non-questions in English was stable across ages and comparable to adult behavior (Figure \ref{fig:E1-randvsreal}), suggesting that they can identify questions in native stimuli with adult-like competence by age three.
@@
 
-Interestingly, adults and children both were strongly affected by transition type, in that they made more anticipatory switches after hearing questions, compared to non-questions.
-Even in the English videos, when participants had full access to linguistic cues, their rates of anticipation were relatively low---in fact, comparable to the non-English videos---unless the turn was a question (Figure \ref{fig:E1-randvsreal}).
-Prior work using online, metalinguistic tasks has shown that participants can use linguistic cues to accurately predict upcoming turn ends \citep{torreira2015, magyari2012, de-ruiter2006}.
-The current results add a new dimension to our understanding of how listeners make predictions about turn ends: Both children and adults spontaneously monitor the linguistic structure of unfolding turns for cues to upcoming responses.
-
-Children and adults behaved relatively similarly in this first experiment, and our language manipulation (English vs.
-non-English) was too coarse to comment on when children begin to use specific linguistic cues (e.g., prosody vs.
-lexicosyntax).
-We would instead need to directly compare lexicosyntactic and prosodic cues in the participants' native language, controlling for the presence of non-verbal cues.
-To see the emergence of anticipatory gaze switching we would also need to include younger children, since participants already reliably made anticipatory gaze switches at age three.
-We follow up on both of these ideas in Experiment 2.
+% Children and adults behaved relatively similarly in this first experiment, and
+
+%
 
@@
 
-We improved our design by using native-language stimuli, controlling for lexical and prosodic information, eliminating non-verbal cues, and testing children from a wider age range.
-All of the videos in Experiment 2 were in the participants' native language (American English).
+Experiment 2 used native-language stimuli, controlled for lexical and prosodic information, eliminating non-verbal cues, and tested children from a wider age range.
 To tease apart the role of lexical and prosodic information, we phonetically manipulated the speech signal for pitch, syllable duration, and lexical access.
@@
 Non-verbal cues in Experiment 1 (e.g., gaze and gesture) could have helped participants make predictions about upcoming turn structure  \citep{rossano2009, stivers2010}.
-Since our focus is on linguistic cues, we eliminated all gaze and gestural signals in Experiment 2 by replacing the videos of human actors with videos of puppets.
+Since our focus was on linguistic cues, we eliminated all gaze and gestural signals in Experiment 2 by replacing the videos of human actors with videos of puppets.
 Puppets are less realistic and expressive than human actors, but they create a natural context for having somewhat motionless talkers in the videos (thereby allowing us to eliminate gestural and gaze cues).
@@
 \begin{center}
-\includegraphics[width=0.9\textwidth]{figures/E2-samples-by-lang-groups-trans-types.png}
+\includegraphics[width=\textwidth]{figures/E2-samples-by-lang-groups-trans-types.png}
 \end{center}
@@
 
-\section{Real vs. randomly permuted model outcomes: \textit{t}, \textit{$\beta$}, and \textit{SE}}
+\section{Permutation Analyses}
 \setcounter{figure}{0}
+
+We completed this baseline procedure on 5,000 random permutations of the original turn transition analysis windows and compared the \textit{t}-values from each predictor in the original models (Table \ref{tab:E1-models}) to the distribution of \textit{t}-values for each predictor in the 5,000 models of the randomly permuted datasets.\footnote{We report \textit{t}-values rather than beta estimates because the standard errors in the randomly permuted data models were much higher than for the original data.
+For those interested, plots of the beta and standard error distributions are available in the Supplementary Materials.} We could then test whether significant effects from the original statistical models differed from the random baseline by calculating the proportion of random data \textit{t}-values exceeded by the original \textit{t}-value for each predictor, using the absolute value of all \textit{t}-values for a two-tailed test.
+For example, children's original ``language condition'' \textit{t}-value was $|$3.429$|$, which is greater than 99.9\% of all $|$\textit{t}-value$|$ estimates from the randomly-permuted data models (i.e., \textit{p}$=$.001).
+This leads us to conclude that that the effect of language condition in the original model was highly unlikely to be the result of random gaze shifting.
+
+We excluded the output of random-permutation models that gave convergence warnings in order to remove unreliable estimates from our analyses (non-converging models were 22.4\% and 24.4\% of all models for children and adults, respectively; see the Supplementary Materials for more information on model exclusion).
 

### 2015-12-17 paper/CasillasFrank-Anticip.tex
partial MCF pass

@@
 Overall, participants' pattern of gaze to current speakers demonstrated that they performed basic turn tracking during the videos, regardless of language.
-Figure \ref{fig:E1-randvsreal} shows participants' anticipatory gaze rates across age, language condition, and transition type, in the real data and the random baseline (described below).
+Figure \ref{fig:E1-randvsreal} shows participants' anticipatory gaze rates across age, language condition, and transition type.
 
@@
 \end{center}
-  \end{scriptsize}
+  \end{small}
   \caption{Model output for children and adults' anticipatory gaze switches.}
@@
 We recruited 27 undergraduate adults and 129 children between ages 1;0--6;11 to participate in our experiment.
-We recruited our child participants from the Children's Discovery Museum in San Jose, California, targeting approximately 20 children for each of the six 1-year age groups (range$=$20--23).
+We recruited our child participants from Children's Discovery Museum of San Jose, California, targeting approximately 20 children for each of the six one-year age groups (range: 20--23).
 All participants were native English speakers, though some parents (N$=$27) reported that their child heard a second (and sometimes third) language at home.
@@
 \begin{center}
-\includegraphics[width=0.9\textwidth]{figures/FIG-EN-stim.png}
+\includegraphics[width=\textwidth]{figures/FIG-EN-stim.png}
 \end{center}
@@
 Looks to the addressee also decreased with age, but the change was minimal.
-Figure \ref{fig:E2-randvsreal} shows participants' anticipatory gaze rates across age, the four language conditions, and transition type, in the real data and the random baseline.
-
-\begin{table}[h]
+Figure \ref{fig:E2-randvsreal} shows participants' anticipatory gaze rates across age, the four language conditions, and transition type.
+
+\begin{table}[t]
 \begin{center}
@@
 
-\begin{table}[h]
+\begin{table}[t]
 \begin{center}
@@
 \begin{center}
-\includegraphics[width=\textwidth]{figures/E2-samples-by-lang-groups-trans-types.png}
+\includegraphics[width=\textwidth]{figures/E2-samples-by-lang-groups-trans-types_mod.png}
 \end{center}
@@
 \label{sec:models2}
+
 We identified anticipatory gaze switches for all 85 usable turn transitions, and analyzed them for effects of language condition, transition type, and age with two mixed-effects logistic regressions.
@@
 
-\begin{scriptsize}
+\begin{footnotesize}
 \begin{longtable}{lcccl}
@@
 \end{longtable}
-\end{scriptsize}
+\end{footnotesize}
 
@@
 
-Using the same technique described in Experiment 1 (Section \ref{sec:randbaseline1}), we created and modeled 5,000 random permutations of participants' anticipatory gaze (see Figures \ref{fig:E2-ChiTs} and \ref{fig:E2-AduTs} for the full distributional data).
-Our baseline analyses revealed that none of the significant predictors from models of the original, turn-related data (Table \ref{tab:E2-models}: Children) can be explained by random looking.
+Using the same technique described in Experiment 1 (Section \ref{sec:randbaseline1}), we created and modeled random permutations of participants' anticipatory gaze.
+These analyses revealed that none of the significant predictors from models of the original, turn-related data could be explained by random looking.
 In the children's data, the original model's \textit{t}-values for language condition (\textit{prosody only}), gap duration, the two-way interaction of age and language condition (\textit{prosody only}) and the three-way interaction of age, transition type, and language condition (\textit{normal} speech) were all greater than 95\% of the randomly permuted \textit{t}-values (96.2\%, 100\%, 95.1\%, and 95.1\%, respectively).
 Similarly, the adults' data showed significant differentiation from the randomly permuted data for all originally significant predictors: gap duration and language condition for \textit{normal} speech and words-only speech (greater than 100\%, 96.8\%, and 98.7\% of random \textit{t}-values, respectively).
-The effects of language condition and transition type for the real and randomly permuted data can also be observed in Figure \ref{fig:E1-randvsreal}.
-
-As before, we excluded the output of random-permutation models that resulted in convergence warnings in order to remove unreliable model estimates from our analyses (non-converging models made up 69\% and 70\% of models for children and adults, respectively; see the Supplementary Materials for more information on model exclusion).
-
-\paragraph{Developmental effects}
-The model of the children's data revealed two significant interactions with age, neither of which derived from random looking (Table \ref{tab:E2-models}; Figure \ref{fig:E2-randvsreal}).
+% The effects of language condition and transition type for the real and randomly permuted data can also be observed in Figure \ref{fig:E1-randvsreal}.
+
+\subsubsection{Developmental effects}
+
+Our main goal in extending the age range to 1- and 2-year-olds in Experiment 2 was to find the age of emergence for spontaneous turn structure predictions in our paradigm.
+As in Experiment 1, we used two-tailed \textit{t}-tests to compare children's real gaze rates to the random baseline in the \textit{normal} speech condition, in which the speech stimulus is most like what children hear every day.
+We tested real gaze rates against baseline for three age groups: ages one, two, and three.
+Two- and three-year-old children made anticipatory gaze switches significantly above chance both when all transitions were considered (2-year-olds: \textit{t}(26.193)$=$-4.137, \textit{p}$<$.001; 3-year-olds: \textit{t}(22.757)$=$-2.662, \textit{p}$<$.05) and for question transitions alone (2-year-olds: \textit{t}(25.345)$=$-4.269, \textit{p}$<$.001; 3-year-olds: \textit{t}(21.555)$=$-3.03, \textit{p}$<$.01).
+One-year-olds, however, made anticipatory gaze shifts that were not statistically significant for turn transitions overall and for question turns alone (overall: \textit{t}(24.784)$=$-2.049, \textit{p}$=$.051; questions: \textit{t}(25.009)$=$-2.03, \textit{p}$=$.053).
+
+Regression models for the children's data also revealed two significant interactions with age.
 The first was a significant interaction of age and language condition (for \textit{prosody only} compared to the \textit{no speech} reference level), suggesting a different age effect between the two linguistic conditions.
@@
 \textit{prosody only}) within each random permutation of the data, making pairwise comparisons between the six age groups.
-These tests revealed that children's anticipation in the \textit{prosody only} condition significantly improved at ages five and six (difference scores greater than 95\% of the random data scores; \textit{p}$<$.05; see Figure \ref{fig:E2-lgageinteraction} for the full distributions).
+These tests revealed that children's anticipation in the \textit{prosody only} condition significantly improved at ages five and six (difference scores greater than 95\% of the random data scores (\textit{p}$<$.05).
 
 The second age-based interaction was a three-way interaction of age, transition type, and language condition (for \textit{normal} speech compared to the \textit{no speech} baseline).
-We again created pairwise comparisons of the average difference scores for the transition type-language condition interaction across age groups in each random permutation of the data, finding that the effect of transition type in the \textit{normal} speech condition became larger with age, with significant improvements by age 4 over ages 1 and 2 (99.9\% and 98.86\%, respectively), by age 5 over age 4 (97.54\%), and by age 6 over ages 1, 2, and 5 (99.5\%, 97.36\%, and 95.04\%), all significantly different from chance (\textit{p}$<$.05; see Figure \ref{fig:E2-lgagetypeinteraction} for the full distributions).
-
-Our main goal in extending the age range to 1- and 2-year-olds in Experiment 2 was to find the age of emergence for spontaneous turn structure predictions.
-As in Experiment 1, we used two-tailed \textit{t}-tests to compare children's real gaze rates to the random baseline in the \textit{normal} speech condition, in which the speech stimulus is most like what children hear every day.
-We tested real gaze rates against baseline for three age groups: ages one, two, and three.
-Two- and three-year-old children made anticipatory gaze switches significantly above chance both when all transitions were considered (2-year-olds: \textit{t}(26.193)$=$-4.137, \textit{p}$<$.001; 3-year-olds: \textit{t}(22.757)$=$-2.662, \textit{p}$<$.05) and for question transitions alone (2-year-olds: \textit{t}(25.345)$=$-4.269, \textit{p}$<$.001; 3-year-olds: \textit{t}(21.555)$=$-3.03, \textit{p}$<$.01).
-One-year-olds, however, made anticipatory gaze shifts that were only marginally above chance for turn transitions overall and for question turns alone (overall: \textit{t}(24.784)$=$-2.049, \textit{p}$=$.051; questions: \textit{t}(25.009)$=$-2.03, \textit{p}$=$.053).
+We again created pairwise comparisons of the average difference scores for the transition type-language condition interaction across age groups in each random permutation of the data, finding that the effect of transition type in the \textit{normal} speech condition became larger with age, with significant improvements by age 4 over ages 1 and 2 (99.9\% and 98.86\%, respectively), by age 5 over age 4 (97.54\%), and by age 6 over ages 1, 2, and 5 (99.5\%, 97.36\%, and 95.04\%), all significantly different from chance (\textit{p}$<$.05).
 
@@
 \label{sec:discussion2}
-As in Experiment 1, children and adults spontaneously tracked the turn structure of the conversations, making anticipatory gaze switches at an above-chance rate across all ages, but not in all conditions (Table \ref{tab:look_e2}; Figure \ref{fig:E2-randvsreal}).
-However, when children had access to full linguistic information (in the \textit{normal} speech condition), they made more anticipatory gaze switches than expected by chance at age two (and even marginally at age one).
-Across conditions, children's anticipatory gaze rates were affected by gap duration, plus interactions of language condition, transition type and age (Table \ref{tab:E2-models}), none of which could be explained by a baseline of random gaze switching (Figure \ref{fig:E2-ChiTs}).
-
-Two of the three linguistic conditions affected children's anticipatory gaze switches, compared to the no-speech reference level: prosody-only speech and \textit{normal} speech.
-\textit{Prosody only} speech resulted in fewer anticipatory gazes overall, compared to the no-speech baseline, though the effect was only marginal (\textit{p}$=$.06).
-However, prosody-only speech \textit{did} significantly differ from the \textit{no speech} condition in its interaction with age: Whereas age gains were negligible in the \textit{no speech} condition, 5- and 6-year-olds in the prosody-only condition showed significantly more anticipatory gaze switches than younger children (Figure \ref{fig:E2-lgageinteraction}), going from at-chance anticipatory gaze rates to rates well above chance (Figure \ref{fig:E2-randvsreal}).
-
-Anticipatory gaze for questions increased significantly more with age in the \textit{normal} speech condition compared to the no-speech baseline.
-As in Experiment 1, children consistently made more anticipatory gazes after hearing questions when they had access to lexical material.
-But now, with a larger age span in Experiment 2, we can begin to see a developmental path for the question effect.
-At least for \textit{normal} speech, greater anticipatory looking for questions is not present from the start: 4-year-olds are the first to make significant gains on 1- and 2-year-olds, but then there are further significant gains at age 5, and again age 6 (Figure \ref{fig:E2-lgagetypeinteraction}).
-While children at ages five and six showed adult-like differentiation of questions and non-questions in the \textit{normal} speech condition, 1-year-olds had nearly identical switch rates for the two transition types.
-This suggests that the participants' tendency to make more anticipatory switches for questions emerges after age one and continues developing, with rapid gains in ages three through age six.
-Finally, children showed a straightforward effect of gap duration (Table \ref{tab:E2-models}), as in Experiment 1.
-
-Adults' anticipatory gaze rates were affected by gap duration and two of the language conditions (Table \ref{tab:E2-models}), none of which could be explained by a baseline of random gaze switching (Figure \ref{fig:E2-AduTs}).
-Adults made more anticipatory switches overall for the \textit{normal} speech and words-only conditions compared to the no-speech condition, falling in-line with past work showing that adults primarily use lexical information in making predictions about upcoming turn structure \citep{de-ruiter2006}.
-Though adults did make more anticipatory switches for questions than non-questions on average in the two lexical conditions (Figure \ref{fig:E2-randvsreal}), the effect of transition type was not significant in either one (Table \ref{tab:E2-models}).
-Like children, adults also showed a straightforward effect of gap duration.
-
-\subsubsection{Summary}
-\label{sec:summary2}
-
-Children and adults both showed more anticipatory gaze switches with increased linguistic information, but only for a subset of the linguistic conditions and transition types.
-We had expected to see the most anticipatory switches in the \textit{normal} condition and the least anticipatory switches in the \textit{no speech} condition because they contained the most and least linguistic information, respectively.
-We had also expected to replicate our finding from Experiment 1 that questions result in more anticipatory switches than non-questions, with the added hypothesis that the question effect is driven by lexicosyntactic cues.
-We additionally anticipated an overall increase in anticipatory switches with age.
-Finally, since the development of prosodic skills partially precedes the development of lexicosyntax, we expected to see more switches in the \textit{prosody only} condition compared to the \textit{words only} condition in the youngest age groups.
-
-In fact, children and adults did show more anticipatory gaze switching in the \textit{normal} speech condition, compared to the no-speech condition, but for children this effect only emerged in the form of anticipations following question turns, which increased with age significantly faster than it did when there were no linguistic cues present at all.
-Adults also showed significantly higher anticipatory switch rates in the \textit{words only} condition but no effects of transition type, alone or within linguistic conditions.
-Taken together, these results partly replicate the findings from Experiment 1: Participants make more anticipatory switches when they have access to lexical information and, when they do, tend to make more anticipatory switches for questions compared to non-questions.
-
-We had anticipated significant gains in anticipatory switching with age, but children only showed significant developmental increases in the \textit{prosody only} condition and the \textit{normal} condition (for question transitions).
-Rather than showing an early advantage for prosody over lexical information, children's anticipation were at chance with prosody-only speech and did not show significant improvement until age five.
-In contrast, their anticipatory gaze switches were already marginally above chance at age one and significantly so at age two for \textit{normal} speech.
-These findings therefore do not support an early role for prosody in children's spontaneous turn structure predictions.
-On the contrary, children's predictions were best when lexical information was present, especially when following a question.
-
-However, these results do not support the idea that lexical information is sufficient for children, as has been proposed for adults previously (\citealp{de-ruiter2006}) and replicated in the present study.
-On the contrary, children showed significant gains in the \textit{normal} speech condition, but \textit{not} in the \textit{words only} condition.
-Notably, the \textit{normal} speech condition, in addition to having both lexical and prosodic information, is also the one most likely to occur in our participants' daily lives (compared to muffled or robotic speech) and therefore may have an extra advantage of familiarity over the other conditions.
-
-Finally, though participants were often able to make anticipatory gaze switches more often than would be expected by chance with no linguistic information (in the \textit{no speech} condition), the advantage for lexical cues emerging in the children's data with age and in the adult's data overall suggests that participants were using linguistic information in making their predictions when it was available.
 
 The core aims of Experiment 2 were to gain better traction on the individual roles of prosody and lexicosyntax in children's turn predictions, and to find the age of emergence for spontaneous turn anticipation.
-Replicating many of our findings from Experiment 1, we found that children and adults both use linguistic cues, in children's case, primarily to anticipate responses after question turns and primarily when the full speech signal was available.
-Adults, on the other hand, showed broadly increased anticipation for lexical conditions, with no overall effects of transition type.
-
-Rather than finding an early advantage for prosodic cues, we found that children struggled to make above-chance anticipations in the \textit{prosody only} condition until age five.
-They didn't show significant overall effects for lexical information alone either, suggesting that children need the full linguistic signal to extract predictive benefits from ongoing conversational speech.
-This finding is surprising given children's earlier development of prosodic skills but unsurprising in the light of their overall question bias---compared to prosodic cues, lexicosyntactic cues to questionhood are categorical units, have clearer form-to-meaning mappings, and occur early in the utterance.
-This finding has implications for our conceptions about adults and children's use of linguistic cues for spontaneous turn prediction, which we expand on in the General Discussion.
+Taken together, our results replicate the findings from Experiment 1: Participants make more anticipatory switches when they have access to lexical information and, when they do, tend to make more anticipatory switches for questions compared to non-questions.
+
+As in Experiment 1, with normal speech, children and adults spontaneously tracked the turn structure of the conversations, making anticipatory gaze switches at an above-chance rate across all ages.
+And in addition, they made far more anticipations for questions than for non-question turns---at least for two-year-olds and older.
+But these effects were different for the two comparison conditions, \textit{prosody-only} and \textit{words-only}.
+In the \textit{prosody-only} condition, performance was low for younger children and increased substantially for older children (especially for questions).
+In the \textit{words-only} condition, anticipation performance was relatively robust for questions for two-year-olds and older (much like in normal speech), but never rose above chance for the children for non-question turns.
+Thus, these findings do not support an early role for prosody in children's spontaneous turn structure predictions.
+On the contrary, children's predictions appeared to stem more from lexical information, probably primarily from question words.
 
@@
 
-We excluded the output of random-permutation models that gave convergence warnings in order to remove unreliable estimates from our analyses (non-converging models were 22.4\% and 24.4\% of all models for children and adults, respectively; see the Supplementary Materials for more information on model exclusion).
+E1: We excluded the output of random-permutation models that gave convergence warnings in order to remove unreliable estimates from our analyses (non-converging models were 22.4\% and 24.4\% of all models for children and adults, respectively; see the Supplementary Materials for more information on model exclusion).
+
+E2: As before, we excluded the output of random-permutation models that resulted in convergence warnings in order to remove unreliable model estimates from our analyses (non-converging models made up 69\% and 70\% of models for children and adults, respectively; see the Supplementary Materials for more information on model exclusion).
 

### 2015-12-17 paper/CasillasFrank-Anticip.tex
cuts in GD

@@
 Until recently, we have had very little data on how children weave language into their already-existing turn-taking behaviors.
-
-In two experiments investigating children's anticipatory gaze to upcoming speakers, we found evidence that turn prediction develops early in childhood and that spontaneous predictions are primarily driven by participants' expectation of an immediate response in the next turn.
+In two experiments investigating children's anticipatory gaze to upcoming speakers, we found evidence that turn prediction develops early in childhood and that spontaneous predictions are primarily driven by participants' expectation of an immediate response in the next turn (e.g., for questions).
 In making predictions about upcoming turn structure, children used a combination of lexical and prosodic cues; neither lexical nor prosodic cues alone were sufficient to support increased anticipatory gaze.
@@
 
-Importantly, we also found no evidence that lexical information alone is equivalent to full linguistic information for children, as has been shown before \citep{magyari2012, de-ruiter2006} and replicated in the current study for adult participants: In both experiments, children's performance was best in conditions when they had access to the full linguistic signal.
+Importantly, we also found no evidence that lexical information alone is equivalent to full linguistic information for children, as has been shown before \citep{magyari2012, de-ruiter2006} and replicated in the current study for adult participants.
+In both experiments, children's performance was best in conditions when they had access to the full linguistic signal.
 Adults on the other hand, showed significant gains in anticipatory gaze switching in both conditions with lexical cues.
@@
 
-\subsubsection{The question effect}
+\subsection{The question effect}
 
@@
 
-%Because the form of a question can constrain the type of response that will come next (e.g., a location after a \textit{where} question), questions can even help listeners predict specific upcoming content in the next turn.
-
-%Our results suggest that question turns start significantly affecting children's predictions between ages three and four, but further testing with finer-grained age samples with stimuli focused on specific communicative acts and linguistic cues is needed to better sketch out the developmental trajectory of this effect.
-For example, the effect size and age of emergence might differ by question type (e.g., \textit{wh-} vs.
-\textit{yes-no}) or the location of question-identifying cues within the unfolding utterance (early vs.
-late), but we would not be able to see it given the current design.
-
-Prior work on children's acquisition of questions indicates that they may already have some understanding about question-answer sequences by the time they begin to speak: Questions make up approximately one third of the utterances children hear, before and after the onset of speech, and even into their preschool years, even though the types and complexity of questions change throughout development \citep{casillas2016, fitneva2012, henning2005, shatz1979}.\footnote{There is substantial variation question frequency by individual and socioeconomic class \citep{hart1992}.} For the first few years, many of the questions directed to children are ``test' questions---questions that the caregiver already has the answer to (e.g., ``What does a cat say?''), but this changes as children get older.
+Prior work on children's acquisition of questions indicates that they may already have some understanding about question-answer sequences by the time they begin to speak: Questions make up approximately one third of the utterances children hear, before and after the onset of speech, and even into their preschool years, even though the types and complexity of questions change throughout development \citep{casillas2016, fitneva2012, henning2005, shatz1979}.\footnote{There is substantial variation question frequency by individual and socioeconomic class \citep{hart1992}.} For the first few years, many of the questions directed to children are ``test'' questions---questions that the caregiver already has the answer to (e.g., ``What does a cat say?''), but this changes as children get older.
 Questions help caregivers to get their young children's attention and to ensure that information is in common ground, even if the responses are non-verbal or infelicitous \citep{bruner1985, fitneva2012, snow1977}.
@@
 
-Two important questions for future work are then: (a) How does children's ability to monitor for questions in conversation relate to their prior experience with questions?
-and (b) what is it about questions that makes children and adults more likely to anticipatorily switch their gaze to addressees?
+Two important questions for future work are then: (1) How does children's ability to monitor for questions in conversation relate to their prior experience with questions?
+and (2) what is it about questions that makes children and adults more likely to anticipatorily switch their gaze to addressees?
 Other request formats, such as imperatives, compliments, and complaints make a response from the addressee highly likely in the next turn \citep{schegloff2007}.
@@
 
-The question effect itself has implications for our current theories about turn prediction.
+% The question effect itself has implications for our current theories about turn prediction.
 While participants may always use linguistic information to predict upcoming speaker changes (as they have in the two studies presented here), they might not always use linguistic information to predict upcoming turn ends, as is assumed in metalinguistic measures of turn-end prediction (e.g., pressing a button while listening to speech: \citealp{torreira2015, magyari2012, de-ruiter2006}) and as has been the focus of other turn-end prediction studies (\citealp{ford1996, duncan1972}).
@@
 
-There is at least one clear hypothesis that can bridge these apparently conflicting results: Listeners in spontaneous, first-person conversation may use multiple strategies in making predictions about upcoming turn structure, using more passive prediction to detect upcoming speaker transition (e.g., questions), and then switching into precise turn-end prediction mode (\`{a} la \citealp{de-ruiter2006}) when necessary.
+% There is at least one clear hypothesis that can bridge these apparently conflicting results: Listeners in spontaneous, first-person conversation may use multiple strategies in making predictions about upcoming turn structure, using more passive prediction to detect upcoming speaker transition (e.g., questions), and then switching into precise turn-end prediction mode (\`{a} la \citealp{de-ruiter2006}) when necessary.
 A flexible prediction system like this one allows listeners to continuously monitor ongoing conversation at a low cost while still managing to plan their responses and come in quickly when needed.
 
-To test this hypothesis, which integrates findings from turn-structure prediction with multiple measures, age groups, and styles of linguistic control, it will be crucial to look at prediction from a first-person perspective.
+% To test this hypothesis, which integrates findings from turn-structure prediction with multiple measures, age groups, and styles of linguistic control, it will be crucial to look at prediction from a first-person perspective.
 The results we present here are based on predictions about third-party conversation, which enables participants to follow interactions with no chance of actually participating.
@@
 
-\subsubsection{Early competence for turn taking?}
+\subsection{Early competence for turn taking?}
 
@@
 
-There are at least two major limitations to our work: Speech naturalness and participant role.Following prior work \citep{de-ruiter2006, keitel2013}, we used phonetically manipulated speech in Experiment 2, resulting in speech sounds that children don't usually hear in their natural environment.
+There are at least two major limitations to our work: Speech naturalness and participant role.
+Following prior work \citep{de-ruiter2006, keitel2013}, we used phonetically manipulated speech in Experiment 2, resulting in speech sounds that children don't usually hear in their natural environment.
 Many prior studies have used phonetically-altered speech with infants and young children \citep[cf.][]{jusczyk2000}, but almost none of them have done so in a conversational context.

### 2016-01-08 paper/CasillasFrank-Anticip.tex
Updated figures and tables

@@
 \usepackage[section]{placeins}
+\usepackage{rotating}
 
@@
 \end{center}
-\caption{Anticipatory gaze rates across language condition and transition type for the real (red and blue) and randomly permuted baseline (gray). Vertical bars represent the standard error.} 
+\caption{Anticipatory gaze rates across language condition and transition type for the real (red and blue) and randomly permuted baseline (gray). Vertical bars represent 95\% confidence intervals.} 
 \label{fig:E1-randvsreal}
@@
     \hline
-    (Intercept)   &      -0.96146 &   0.84901 & -1.132 & 0.257446 \\    
-    Age             &    -0.18268 &   0.17507 & -1.043 & 0.296725   \\  
-    LgCond$=$\textit{non-English} &          -3.29347 &   0.96045 & -3.429 & 0.000606 *** \\
-    Type$=$\textit{non-Question}          &     -1.10129 &   0.86494 & -1.273 & 0.202925     \\
-    Duration      &       3.40169 &   1.22826 &  2.770 & 0.005614  **\\
-    Age*LgCond$=$\textit{non-English} &       0.52065 &   0.21190 &  2.457 & 0.014008 ** \\  
-    Age*TypeS$=$\textit{non-Question}          & -0.01628 &   0.19437 & -0.084 & 0.933232     \\
-    LgCond$=$\textit{non-English}* &     2.68166 &   1.35016 &  1.986 & 0.047013 * \\
+    (Intercept)   						& -0.96145	& 0.84915		& -1.132	& 0.257531 	\\    
+    Age             						& -0.18268	& 0.17509		& -1.043	& 0.296764   	\\  
+    LgCond$=$\textit{non-English} 		& -3.29349	& 0.96055		& -3.429	& 0.000606 *** 	\\
+    Type$=$\textit{non-Question}          	& -1.10131	& 0.86520		& -1.273	& 0.203055     	\\
+    Duration      						& 3.40171		& 1.22878		&  2.768	& 0.005634  **	\\
+    Age*LgCond$=$\textit{non-English} 	& 0.52066		& 0.21192		&  2.457	& 0.014015 * 	\\  
+    Age*TypeS$=$\textit{non-Question}      & -0.01628	& 0.19442		& -0.084	& 0.933262     	\\
+    LgCond$=$\textit{non-English}*		& 2.68171		& 1.35045		& 1.986	& 0.047057 * 	\\
     \hspace*{5mm} Type$=$\textit{non-Question} &&&& \\
-    Age*LgCond$=$\textit{non-English}* & -0.45632 &   0.30163 & -1.513 & 0.130315  \\
+    Age*LgCond$=$\textit{non-English}*	& -0.45633	& 0.30168		& -1.513	& 0.130378  \\
     \hspace*{5mm} Type$=$\textit{non-Question} &&&& \\
@@
     \hline
-    (Intercept)       &        -0.1966 &    0.6942 & -0.283 & 0.776988     \\
-    LgCond$=$\textit{non-English}    &             -0.8812 &    0.9602 & -0.918 & 0.358754   \\  
-    Type$=$\textit{non-Question}             &        -4.4953 &    1.3139 & -3.421 & 0.000623 *** \\
-    Duration          &        -1.1227 &    1.9880 & -0.565 & 0.572238     \\
-    LgCond$=$\textit{non-English}* &           3.2972 &    1.6101 &  2.048 & 0.040581 * \\ 
-    Type$=$\textit{non-Question} &&&& \\
-    LgCond$=$\textit{non-English}* &        1.3626 &    3.0077 &  0.453 & 0.650527     \\
+    (Intercept)						& -0.1966		& 0.6945		& -0.283	& 0.777062     	\\
+    LgCond$=$\textit{non-English}		& -0.8812		& 0.9607		& -0.917	& 0.359028   	\\  
+    Type$=$\textit{non-Question}		& -4.4953		& 1.3147		& -3.419	& 0.000628 *** \\
+    Duration						& -1.1227		& 1.9889		& -0.565	& 0.572414     	\\
+    LgCond$=$\textit{non-English}*		& 3.2972		& 1.6115		& 2.046	& 0.040747 * 	\\ 
+    \hspace*{5mm} Type$=$\textit{non-Question} &&&& \\
+    LgCond$=$\textit{non-English}*		& 1.3625		& 3.0097		& 0.453	& 0.650749     	\\
     \hspace*{5mm} Duration &&&& \\
-    Type$=$\textit{non-Question}*          &  10.5107 &    3.3459 &  3.141 & 0.001682  ** \\
+    Type$=$\textit{non-Question}*		& 10.5107		& 3.3482		& 3.139	& 0.001694  ** 	\\
     \hspace*{5mm} Duration &&&& \\
-    LgCond$=$\textit{non-English}* & -6.3156 &    4.4926 & -1.406 & 0.159790 \\    
-    Type$=$\textit{non-Question}* &&&& \\
+    LgCond$=$\textit{non-English}*		& -6.3156		& 4.4969		& -1.404	& 0.160191 	\\    
+    \hspace*{5mm} Type$=$\textit{non-Question}* &&&& \\
     \hspace*{5mm} Duration &&&& \\
@@
 \begin{center}
-\includegraphics[width=\textwidth]{figures/E2-samples-by-lang-groups-trans-types_mod.png}
+\includegraphics[width=\textwidth]{figures/E2-samples-by-lang-groups-trans-types.png}
 \end{center}
 \caption{Anticipatory gaze rates across language condition and transition type for the real (blue, dark green, light green, and red) and randomly permuted baseline (gray).
-Vertical bars represent the standard error.}
+Vertical bars represent 95\% confidence intervals.}
 \label{fig:E2-randvsreal}
@@
     \hline
-    (Intercept)                & -3.57414  &   0.48576 &  -7.358 & 1.87e-13 ***    \\
-    Age                         & 0.02543 &    0.10260  &  0.248   & 0.8042        \\
-    Type$=$\textit{non-Question}                      & -0.81873  &   0.59985 &  -1.365  &  0.1723        \\
-    Duration                   &  4.17672  &   0.62446  &  6.689 & 2.25e-11 ***    \\
-    Age*Type$=$\textit{non-Question}                &    0.15116   &  0.13643  &  1.108  &  0.2679        \\
-	\hline
-    Condition$=$\textit{normal}            &  0.36710 &    0.43296 &   0.848  &  0.3965        \\
-    Age*Condition$=$\textit{normal}       &   0.12919 &    0.10227  &  1.263  &  0.2065        \\
-    Condition$=$\textit{normal}*     &   0.91059  &   0.72095  &  1.263 &   0.2066        \\
+    (Intercept)						& -3.49403	& 0.48454		& -7.211	& 5.55e-13 ***    	\\
+    Age							& 0.02436		& 0.10249		& 0.238	& 0.8121        		\\
+    Type$=$\textit{non-Question}		& -0.88900	& 0.61192		& -1.453	& 0.1463        		\\
+    Duration						& 3.94743		& 0.61668		& 6.401	& 1.54e-10 ***    	\\
+    Age*Type$=$\textit{non-Question}	& 0.15359		& 0.13996		& 1.097	& 0.2725        		\\
+	\hline
+    Condition$=$\textit{normal}			& 0.37337		& 0.43421		& 0.860	& 0.3899        		\\
+    Age*Condition$=$\textit{normal}		& 0.12950		& 0.10217		& 1.267	& 0.2050        		\\
+    Condition$=$\textit{normal}*			& 0.91074		& 0.72581		& 1.255	& 0.2096        		\\
     \hspace*{5mm} Type$=$\textit{non-Question} &&&& \\
-    Age*Condition$=$\textit{normal}* &   -0.37542  &   0.16963  & -2.213 &   0.0269 *      \\
+    Age*Condition$=$\textit{normal}*		& -0.37965	& 0.17031		& -2.229	& 0.0258 *      		\\
     \hspace*{5mm} Type$=$\textit{non-Question} &&&& \\
 	\hline
-    Condition$=$\textit{prosody}         &   -1.63429  &   0.86390 &  -1.892   & 0.0585 .      \\
-    Age*Condition$=$\textit{prosody}       &   0.39317 &     0.18907 &   2.080  &  0.0376 *      \\
-    Condition$=$\textit{prosody}*     &  1.77190  &   1.24864  &  1.419 &   0.1559        \\
+    Condition$=$\textit{prosody}			& -1.60734	& 0.86680		& -1.854	& 0.0637 .      		\\
+    Age*Condition$=$\textit{prosody}		& 0.39271		& 0.18905		& 2.077	& 0.0378 *      		\\
+    Condition$=$\textit{prosody}*		& 1.68552		& 1.05414		& 1.599	& 0.1098        		\\
     \hspace*{5mm} Type$=$\textit{non-Question} &&&& \\
-    Age*Condition$=$\textit{prosody}* &  -0.47057 &    0.28703 &  -1.639 &   0.1011        \\
+    Age*Condition$=$\textit{prosody}*	& -0.32360	& 0.23229		& -1.393	& 0.1636        		\\
     \hspace*{5mm} Type$=$\textit{non-Question} &&&& \\
 	\hline
-    Condition$=$\textit{words}            &  -0.26741 &    0.59071 &  -0.453  &  0.6508        \\
-    Age*Condition$=$\textit{words}           &   0.13740  &   0.13568   & 1.013  &  0.3112        \\
-    Condition$=$\textit{words}*     &   -1.02193  &   1.01227  & -1.010  &  0.3127        \\
+    Condition$=$\textit{words}			& -0.26996	& 0.59313		& -0.455	& 0.6490        		\\
+    Age*Condition$=$\textit{words}		& 0.14044		& 0.13565		& 1.035	& 0.3005        		\\
+    Condition$=$\textit{words}*			& -1.03066	& 1.01610		& -1.014	& 0.3104        		\\
     \hspace*{5mm} Type$=$\textit{non-Question} &&&& \\
-    Age*Condition$=$\textit{words}*  &   0.08946  &   0.22349   & 0.400 &   0.6890        \\
+    Age*Condition$=$\textit{words}*		& 0.08829		& 0.22387		& 0.394	& 0.6933        		\\
     \hspace*{5mm} Type$=$\textit{non-Question} &&&& \\
@@
     \hline
-    (Intercept)           &  -3.4557   &   0.7199 &  -4.800 & 1.58e-06 *** \\ 
-    Type$=$\textit{non-Question}                  &   0.4292  &    0.6089  &  0.705 & 0.480916     \\ 
-    Duration             &     4.7500   &   1.2480  &  3.806 & 0.000141 *** \\ 
-	\hline
-    Condition$=$\textit{normal}    &      1.2556   &   0.5633  &  2.229 & 0.025805 *   \\ 
-    Condition$=$\textit{normal}* &   -0.9452  &    0.7631 &  -1.239 & 0.215475     \\ 
-    \hspace*{5mm} Type$=$\textit{non-Question} &&&& \\
-	\hline
-    Condition$=$\textit{prosody}     &     0.3349  &    0.8965  &  0.374 & 0.708692     \\ 
-    Condition$=$\textit{prosody}*  &  0.6627  &    1.2138 &   0.546 & 0.585108     \\ 
-    \hspace*{5mm} Type$=$\textit{non-Question} &&&& \\
-	\hline
-    Condition$=$\textit{words}       &     1.5938  &    0.7208 &   2.211 & 0.027023 *   \\ 
-    Condition$=$\textit{words}*   &  -1.1265  &    0.9109  & -1.237 & 0.216201     \\ 
+    (Intercept)						& -3.3811		& 0.6884		& -4.912	& 9.03e-07 *** 		\\ 
+    Type$=$\textit{non-Question}		& 0.4375		& 0.5854		& 0.747	& 0.4549     		\\ 
+    Duration						& 4.6961		& 1.1804		& 3.978	& 6.94e-05 *** 		\\ 
+	\hline
+    Condition$=$\textit{normal}			& 1.2033		& 0.5359		& 2.245	& 0.0247 *   		\\ 
+    Condition$=$\textit{normal}*			& -0.9627		& 0.7358		& -1.308	& 0.1907     		\\ 
+    \hspace*{5mm} Type$=$\textit{non-Queston} &&&& \\
+	\hline
+    Condition$=$\textit{prosody}			& 0.2407		& 0.8011		& 0.301	& 0.7638     		\\ 
+    Condition$=$\textit{prosody}*		& 0.5525		& 0.9374		& 0.589	& 0.5556     		\\ 
+    \hspace*{5mm} Type$=$\textit{non-Queston} &&&& \\
+	\hline
+    Condition$=$\textit{words}			& 1.5613		& 0.7087		& 2.203	& 0.0276 *   		\\ 
+    Condition$=$\textit{words}*			& -1.1557		& 0.8854		& -1.305	& 0.1918     		\\ 
     \hspace*{5mm} Type$=$\textit{non-Question} &&&& \\
@@
 
-\begin{table}[h]
+\begin{sidewaystable}[!htbp]
   \begin{scriptsize}
 \centering
-  \begin{tabular}{lccccc}
-    Variable & Mean & Median & \textit{SD} & Minimum & Maximum\\ 
+  \begin{tabular}{lllllllllll}
+    Variable 				& Mean-C & Mean-NC & Median-C & Median-NC & \textit{SD}-C & \textit{SD}-NC & Min-C & Min-NC & Max-C & Max-NC \\ 
     \hline
@@
     \hline
-    (Intercept)						&	-2.52 (-458.42)	& -2.54 (-2.86)	& 0.87 (1319.22)		& -5.53 (-8185.36)	& 0.41 (0.97)			\\
-    Age								& -0.51 (-17.83)		& -0.49 (-0.53)	& 0.79 (83.78)		& -3.71 (-672.2)		& 2.3 (342.8)			\\
-    LgCond					& -0.53 (-109.91)		& -0.55 (-0.63)	& 0.93 (564.42)		& -3.93 (-4418.74)	& 3.23 (2296.19)		\\
-    Type							& -0.10 (-29.66)		& -0.09 (-0.1)		& 0.98 (515.12)		& -4.06 (-4383.92)	& 3.23 (2296.19)		\\
-    Duration						& 0.99 (345.53)		& 0.98 (1.15)		& 1.07 (1323.13)		& -2.44 (-5048.24)	& 3.36 (3416.68)		\\
-    Age*LgCond				& 0.19 (10.64)		& 0.2 (0.18)		& 0.9 (109.6)			& -3.31 (-581.61)		& 5.78 (9985.16)		\\
-    Age*Type						& 0.02 (-1.8)			& 0.001 (-0.05)	& 0.9 (98.27)			& -3.36 (-884.36)		& 3.59 (946.81)		\\
-    LgCond*Type			& 0.2 (45.32)			& 0.2 (0.27)		& 0.96 (691.3)		& -3.12 (-4160.06)	& 3.45 (640.43)		\\
-    Age*LgCond*Type		& -0.12 (-14.23)		& -0.12 (-0.15)	& 0.93 (156.72)		& -2.98 (-1318.26)	& 3.39 (5107.64)		\\
+    (Intercept)				& -2.52	& -458.42		& -2.54	& -2.86	& 0.87	& 1319.22		& -5.53	& -8185.36	& 0.41	& 0.97		\\
+    Age					& -0.51	& -17.83		& -0.49	& -0.53	& 0.79	& 83.78		& -3.71	& -672.2		& 2.3		& 342.8		\\
+    LgCond					& -0.53	& -109.91		& -0.55	& -0.63	& 0.93	& 564.42		& -3.93	& -4418.74	& 3.23	& 2296.19		\\
+    Type					& -0.10	& -29.66		& -0.09	& -0.1	& 0.98	& 515.12		& -4.06	& -4383.92	& 3.23	& 2296.19		\\
+    Duration				& 0.99	& 345.53		& 0.98	& 1.15	& 1.07	& 1323.13		& -2.44	& -5048.24	& 3.36	& 3416.68		\\
+    Age*LgCond				& 0.19	& 10.64		& 0.2		& 0.18	& 0.9		& 109.6		& -3.31	& -581.61		& 5.78	& 9985.16		\\
+    Age*Type				& 0.02	& -1.8		& 0.001	& -0.05	& 0.9		& 98.27		& -3.36	& -884.36		& 3.59	& 946.81		\\
+    LgCond*Type			& 0.2		& 45.32		& 0.2		& 0.27	& 0.96	& 691.3		& -3.12	& -4160.06	& 3.45	& 640.43		\\
+    Age*LgCond*Type		& -0.12	& -14.23		& -0.12	& -0.15	& 0.93	& 156.72		& -2.98	& -1318.26	& 3.39	& 5107.64		\\
     \hline
@@
     \hline
-    (Intercept)    					&    -1.63 (-126.14) 	& -1.71 (-1.73)	& 0.97 (713.39)		& -4.08 (-12111.22)	& 2.15 (649.55)		\\
-    LgCond  					&      -0.26 (-679.6) 	& -0.3 (-0.53)		& 1.02 (15894.33)	& -3.45 (-494979.7)	& 3.35 (88581.58)	\\
-    Type 							&       -0.11 (6.29) 	& -0.13 (-0.04)	& 1.11 (501.5)		& -3.85 (-6420.76)	& 3.28 (8177.88)		\\
-    Duration  						&       0.25 (84.09) 	& 0.27 (0.26)		& 1.1 (1152.94)		& -3.25 (-10864.51)	& 3.46 (18540.62)	\\
-    LgCond*Type  			&       0.12 (-242.27) 	& 0.1 (0.34)		& 1.07 (26836.7)		& -3.41 (-62264.27)	& 3.81 (509198.4)	\\
-    LgCond*Duration  		&       0.15 (780.03) 	& 0.16 (0.39)		& 1.04 (44105.02)	& -3.84 (-798498.6)	& 3.55 (1145951)		\\
-    Type*Duration 				&        0.05 (-6.56) 	& 0.05 (0.02)		& 1.13 (1389.9)		& -3.54 (-15979.22)	& 3.87 (16419.46)	\\
-    LgCond*Type*Duration  &      -0.06 (1083.63)	& -0.08 (-0.21)	& 1.1 (63116.54)		& -4.21 (-1201895)	& 4.02 (1284965)		\\
+    (Intercept)    				& -1.63	& -126.14		& -1.71	& -1.73	& 0.97	& 713.39		& -4.08	& -12111.22	& 2.15	& 649.55		\\
+    LgCond  				& -0.26	& -679.6		& -0.3	& -0.53	& 1.02	& 15894.33	& -3.45	& -494979.7	& 3.35	& 88581.58	\\
+    Type 					& -0.11	& 6.29		& -0.13	& -0.04	& 1.11	& 501.5		& -3.85	& -6420.76	& 3.28	& 8177.88		\\
+    Duration  				& 0.25	& 84.09		& 0.27	& 0.26	& 1.1		& 1152.94		& -3.25	& -10864.51	& 3.46	& 18540.62	\\
+    LgCond*Type  			& 0.12	& -242.27		& 0.1		& 0.34	& 1.07	& 26836.7		& -3.41	& -62264.27	& 3.81	& 509198.4	\\
+    LgCond*Duration  		& 0.15	& 780.03		& 0.16	& 0.39	& 1.04	& 44105.02	& -3.84	& -798498.6	& 3.55	& 1145951	\\
+    Type*Duration 			& 0.05	& -6.56		& 0.05	& 0.02	& 1.13	& 1389.9		& -3.54	& -15979.22	& 3.87	& 16419.46	\\
+    LgCond*Type*Duration  	& -0.06	& 1083.63		& -0.08	& -0.21	& 1.1		& 63116.54	& -4.21	& -1201895	& 4.02	& 1284965	\\
     \hline
   \end{tabular}
-  \caption{Estimated \textit{t}-values for each predictor in in the adult and child models from Experiment 1. Converging (`C') and non-converging (`non-C') values are shown as: `C (non-C)'. }
+  \caption{Estimated \textit{t}-values for each predictor in converging (`-C') and non-converging (`-NC') adult and child models from Experiment 1. }
 \label{tab:nonconv_e1}
  \end{scriptsize}
-\end{table}
+\end{sidewaystable}
 

### 2016-01-12 paper/CasillasFrank-Anticip.tex
updates to discussion and appendices

@@
 
-Children's anticipatory gaze switches showed effects of language condition (\textit{$\beta$}=-3.29, \textit{SE}=0.961, \textit{t}=-3.43, \textit{p}$<$.001) and gap duration (\textit{$\beta$}=3.4, \textit{SE}=1.229, \textit{t}=2.77, \textit{p}$<$.01) with additional effects of an age-by-language condition interaction (\textit{$\beta$}=0.52, \textit{SE}=0.212, \textit{t}=2.46, \textit{p}$<$.05) and a language condition-by-transition type interaction (\textit{$\beta$}=2.68, \textit{SE}=1.35, \textit{t}=1.99, \textit{p}$<$.05).
-There were no significant effects of age or transition type alone (\textit{$\beta$}=-0.18, \textit{SE}=0.175, \textit{t}=-1.04, \textit{p}$=$.3 and \textit{$\beta$}=-1.10, \textit{SE}=0.865, \textit{t}=-1.27, \textit{p}$=$.2, respectively).
-
-Adults' anticipatory gaze switches shows an effect of transition type (\textit{$\beta$}=-4.5, \textit{SE}=1.314, \textit{t}=-3.42, \textit{p}$<$.001) and significant interactions between language condition and transition type (\textit{$\beta$}=3.3, \textit{SE}=1.61, \textit{t}=2.05, \textit{p}$<$.05) and transition type and gap duration (\textit{$\beta$}=10.51, \textit{SE}=3.346, \textit{t}=3.141, \textit{p}$<$.01).
+Children's anticipatory gaze switches showed effects of language condition (\textit{$\beta$}=-3.29, \textit{SE}=0.961, \textit{z}=-3.43, \textit{p}$<$.001) and gap duration (\textit{$\beta$}=3.4, \textit{SE}=1.229, \textit{z}=2.77, \textit{p}$<$.01) with additional effects of an age-by-language condition interaction (\textit{$\beta$}=0.52, \textit{SE}=0.212, \textit{z}=2.46, \textit{p}$<$.05) and a language condition-by-transition type interaction (\textit{$\beta$}=2.68, \textit{SE}=1.35, \textit{z}=1.99, \textit{p}$<$.05).
+There were no significant effects of age or transition type alone (\textit{$\beta$}=-0.18, \textit{SE}=0.175, \textit{z}=-1.04, \textit{p}$=$.3 and \textit{$\beta$}=-1.10, \textit{SE}=0.865, \textit{z}=-1.27, \textit{p}$=$.2, respectively).
+
+Adults' anticipatory gaze switches shows an effect of transition type (\textit{$\beta$}=-4.5, \textit{SE}=1.315, \textit{z}=-3.42, \textit{p}$<$.001) and significant interactions between language condition and transition type (\textit{$\beta$}=3.3, \textit{SE}=1.61, \textit{z}=2.05, \textit{p}$<$.05) and transition type and gap duration (\textit{$\beta$}=10.51, \textit{SE}=3.348, \textit{z}=3.139, \textit{p}$<$.01).
 
@@
 By running the real and randomly permuted data sets through identical statistical models, we can also estimate how likely it is that predictor effects in the original data (e.g., the effect of language condition; Table \ref{tab:E1-models}) arose from random looking.
-Because these analyses are complex, we report their full details in Appendix A.
+Because these analyses are complex, we report their full details in \ref{sec:permutation}.
 
 Our baseline analyses revealed that none of the significant predictors from models of the original, turn-related data can be explained by random looking.
-For the children's data, the original \textit{t}-values for language condition, gap duration, the age-language condition interaction, and the language condition-transition type interaction were all greater than 95\% of \textit{t}-values for the randomly permuted data (99.9\%, 95.5\%, 99.4\%, and 96\%, respectively).
-Similarly, the adults' data showed significant differentiation from the randomly permuted data for two of the three originally significant predictors---transition type and the transition type-gap duration interaction (greater than 99.9\% and 99.7\% of random \textit{t}-values, respectively)---with marginal differentiation for the interaction of language condition and transition type (greater than 94.6\% of random \textit{t}-values).
+For the children's data, the original \textit{z}-values for language condition, gap duration, the age-language condition interaction, and the language condition-transition type interaction were all greater than 95\% of \textit{z}-values for the randomly permuted data (99.9\%, 95.5\%, 99.4\%, and 96\%, respectively).
+Similarly, the adults' data showed significant differentiation from the randomly permuted data for two of the three originally significant predictors---transition type and the transition type-gap duration interaction (greater than 99.9\% and 99.7\% of random \textit{z}-values, respectively)---with marginal differentiation for the interaction of language condition and transition type (greater than 94.6\% of random \textit{z}-values).
 
@@
 The models reported above revealed a significant interaction of age and language condition (Table \ref{tab:E1-models}) that was unlikely be due to random looking (Figure \ref{fig:E1-randvsreal}).
-To further explore this effect, we compared the average effect of language condition for each age group: Using the permutation analyses above, we extracted the average difference score for the two language conditions (English minus non-English) for each subject, computing an overall average for each random permutation of the data.
-For each random permutation, we then made pairwise comparisons of the average difference scores across pariticipant age groups.
-Details are given in Appendix B.
-
-These analyses showed that, while 3- and 4-year olds showed similarly large effects of language condition, 5-year-olds showed a significantly smaller effect of language condition, compared to both younger age groups.
-In other words, the difference in the effect of language condition for 5-year-olds compared to younger children was larger than would be expected by chance in 99.52\% of the randomly permuted data sets for 3-year-olds and 99.96\% of the data sets for 4-year-olds---differences of \textit{p}$<$.01 and \textit{p}$<$.001, respectively (see Figure \ref{fig:E1-lgageinteraction} for difference score distributions).
+To further explore this effect, we compared the average effect of language condition for each age group: Using the permutation analyses above, we extracted the average difference score for the two language conditions (English minus non-English) for each participant, computing an overall average for each random permutation of the data.
+Then, within each permutation, we made pairwise comparisons of the average difference scores across participant age groups.
+This process yielded a distribution of random permutation-based difference scores that we could then compare to the difference score in the actual data.
+Details are given in \ref{sec:pairwisedev}.
+
+These analyses revealed that, while 3- and 4-year olds showed similarly-sized effects of language condition, 5-year-olds showed a significantly smaller effect of language condition, compared to both younger age groups.
+The difference in language condition effect between 5-year-olds and 3-year-olds was greater than would be expected by chance in 99.52\% of the randomly permuted data sets (\textit{p}$<$.01).
+Similarly, the difference in language condition effect between 5-year-olds and 4-year-olds was greater than would be expected by chance in  99.96\% of the data sets (\textit{p}$<$.001; see Figure \ref{fig:E1-lgageinteraction} for difference score distributions).
 
 When does spontaneous turn prediction emerge developmentally during natural speech?
-To test whether the youngest age group (3-year-olds) already exceeded chance in their anticipatory gaze switches, we used two-tailed \textit{t}-tests to compare their real gaze rates to the random baseline in the English condition.
-Although the overall effect was small, we found that three-year-olds made anticipatory gaze switches significantly above chance, when all transitions were considered (\textit{t}(22.824)$=$-4.147, \textit{p}$<$.001) as well as for question transitions alone (\textit{t}(21.677)$=$-5.268, \textit{p}$<$.001).
+To test whether the youngest age group (3-year-olds) already exceeded chance in their anticipatory gaze switches, we  compared children's real gaze rates to the random baseline in the English condition with two-tailed \textit{t}-tests.
+We found that three-year-olds made anticipatory gaze switches significantly above chance, when all transitions were considered (\textit{t}(22.824)$=$-4.147, \textit{p}$<$.001) as well as for question transitions alone (\textit{t}(21.677)$=$-5.268, \textit{p}$<$.001).
 
@@
 Children and adults spontaneously tracked the turn structure of the conversations, making anticipatory gaze switches at an above-chance rate across all ages and conditions.
-Children's anticipatory gaze rates were affected by language condition, transition type, age, and gap duration (Table \ref{tab:E1-models}), none of which could be explained by a baseline of random gaze switching (Figure \ref{fig:E1-ChiTs}).
+Children's anticipatory gaze rates were affected by language condition, transition type, age, and gap duration (Table \ref{tab:E1-models}), none of which could be explained by a baseline of random gaze switching (\ref{sec:permutation}; Figure \ref{fig:E1-ChiTs}).
 These data show a number of important features that bear on our questions of interest.
@@
 Both groups made more anticipatory switches after hearing questions, compared to non-questions.
-Even in the English videos, when participants had full access to linguistic cues, their rates of anticipation were relatively low---in fact, comparable to the non-English videos---unless the turn was a question.
+Even in the English videos, when participants had full access to linguistic cues, their rates of anticipation were relatively low---comparable to the non-English videos---unless the turn was a question.
 Prior work using online, metalinguistic tasks has shown that participants can use linguistic cues to accurately predict upcoming turn ends \citep{torreira2015, magyari2012, de-ruiter2006}.
-The current results add a new dimension to our understanding of how listeners make predictions about turn ends: Both children and adults spontaneously monitor the linguistic structure of unfolding turns for cues to upcoming responses.
-
-Second, we saw developmental effects such that older children anticipated more reliably, but only in the non-English videos.
-In the English videos all children anticipated, especially foro questions.
-But the language condition interaction suggests that the 5-year-olds were able to leverage anticipatory cues in the non-English videos in a way that 3- and 4-year-olds could not, possibly by shifting more attention to the non-native prosodic or non-verbal cues.
+The current results add a new dimension to our understanding of how listeners make predictions about turn ends: Both children and adults spontaneously monitor the linguistic structure of unfolding turns for cues to imminent responses.
+
+Second, children made more anticipatory switches overall in English videos, compared to non-English videos.
+This effect suggests that lexical access is important for children's ability to anticipate upcoming turn structure, consistent with prior work on turn-end prediction in adults (\citealp{de-ruiter2006}; \citealp{magyari2012}) and children \citep{keitel2013}.
+
+Third, we saw developmental effects such that older children anticipated more reliably than younger children, but only in the non-English videos.
+In the English videos all children anticipated, especially after hearing questions.
+This interaction between age and language condition suggests that the 5-year-olds were able to leverage anticipatory cues in the non-English videos in a way that 3- and 4-year-olds could not, possibly by shifting more attention to the non-native prosodic or non-verbal cues.
 Prior work on children's turn-structure anticipation proposed that children's turn-end predictions rely primarily on lexicosyntactic structure (and not, e.g., prosody) as they get older \citep{keitel2013}.
@@
 
-In Experiment 2 we follow up on these findings, improving on two aspects of the design.
-Our language manipulation in this first experiment was too coarse to provide data regarding specific linguistic cues (e.g., prosody vs.
+In Experiment 2 we follow up on these findings, improving on two aspects of the design: First, our language manipulation in this first experiment was too coarse to provide data regarding specific linguistic channels (e.g., prosody vs.
 lexicosyntax).
-In Experiment 2, to compare lexicosyntactic and prosodic cues directly, controlling for the presence of non-verbal cues, wed use artificial stimuli.
-In addition, we saw above-chance anticipation in the youngest children we tested in Experiment 1.
-Although there were developmental changes in prediction for non-English stimuli, all children anticipated successfully after questions.
+In Experiment 2 we compared lexicosyntactic and prosodic cues with phonetically altered speech and used puppets to eliminate non-verbal cues to turn taking.
+Second, we were not able to pinpoint the emergence of anticipatory switching because the youngest age group in our first sample were already able to make anticipatory switches above chance rates.
 In Experiment 2 we explore a wider developmental range.
-
-% Language condition (English vs.
-non-English) affected children's anticipations in two ways (Table \ref{tab:E1-models};  Figure \ref{fig:E1-randvsreal}).
-First, children made more anticipatory switches overall in English videos, compared to non-English videos.
-This effect suggests that lexical access is important for children's ability to anticipate upcoming turn structure; children had no lexical access to the speech in the non-English videos, though they did have access to (non-native) prosodic cues and non-verbal behavior.
-This finding is consistent with prior work on turn-end prediction in adults (\citealp{de-ruiter2006}; \citealp{magyari2012}) and children \citep{keitel2013}.
-Second, children systematically made more anticipatory switches after hearing a question compared to a non-question, but only in the English condition, suggesting that, when children have access to lexical cues, they are more likely to make an anticipatory gaze switch if they can expect an immediate response from the addressee.
-If so, then children's (and adults') attention to lexical cues for turn taking may primarily be in monitoring the signal for cues to questionhood (e.g., subject-auxiliary inversion, \textit{wh}-words, etc.).
 
@@
 Finally, adults' anticipation rates were also affected by gap duration, but more so for questions than non-questions (Table \ref{tab:E1-models}), suggesting that adults were less likely overall to make switches at non-questions, and so did not benefit from extra time to do so.
-
-% % \subsubsection{Summary}
-% % \label{sec:summary1}
 
@@
 Relatedly, though lexical access clearly facilitated participants' anticipatory gaze rate, it was not necessary for participants---especially adults---in order to exceed chance switching rates (Figure \ref{fig:E1-randvsreal}), suggesting that participants use non-lexical cues (e.g., prosody, non-verbal behavior) to make anticipatory eye movements at least some of the time.
-
-% Children and adults behaved relatively similarly in this first experiment, and
-
-%
 
@@
     Condition$=$\textit{normal}*			& -0.9627		& 0.7358		& -1.308	& 0.1907     		\\ 
-    \hspace*{5mm} Type$=$\textit{non-Queston} &&&& \\
+    \hspace*{5mm} Type$=$\textit{non-Question} &&&& \\
 	\hline
@@
     Condition$=$\textit{prosody}*			& 0.5525			& 0.9374		& 0.589	& 0.5556     		\\ 
-    \hspace*{5mm} Type$=$\textit{non-Queston} &&&& \\
+    \hspace*{5mm} Type$=$\textit{non-Question} &&&& \\
 	\hline
@@
 
-Children's anticipatory gaze switches showed an effect of gap duration (\textit{$\beta$}=4.18, \textit{SE}=0.624, \textit{t}=6.689, \textit{p}$<$.001), a two-way interaction of age and language condition (for \textit{prosody only} speech compared to the \textit{no speech} reference level; \textit{$\beta$}=0.393, \textit{SE}=0.189, \textit{t}=2.08, \textit{p}$<$.05), and a three-way interaction of age, transition type, and language condition (for \textit{normal} speech compared to the \textit{no speech} reference level; \textit{$\beta$}=-0.375, \textit{SE}=0.17, \textit{t}=-2.213, \textit{p}$<$.05).
-There were no significant effects of age or transition type alone (Table \ref{sec:models2}), with only a marginal effect of language condition (for \textit{prosody only} compared to the \textit{no speech} reference level; \textit{$\beta$}=-1.634, \textit{SE}=0.864, \textit{t}=-1.89, \textit{p}$=$.06)
-
-Adults' anticipatory gaze switches showed effects of gap duration (\textit{$\beta$}=4.75, \textit{SE}=1.248, \textit{t}=3.806, \textit{p}$<$.001) and language condition (for \textit{normal} speech \textit{$\beta$}=1.256, \textit{SE}=0.563, \textit{t}=2.229, \textit{p}$<$.05 and \textit{words only} speech \textit{$\beta$}=1.594, \textit{SE}=0.721, \textit{t}=2.211, \textit{p}$<$.05  compared to the \textit{no speech} reference level).
-There were no effects of transition type (\textit{$\beta$}=0.429, \textit{SE}=0.609, \textit{t}=0.705, \textit{p}$=$.48).
+Children's anticipatory gaze switches showed an effect of gap duration (\textit{$\beta$}=3.95, \textit{SE}=0.617, \textit{z}=6.401, \textit{p}$<$.001), a two-way interaction of age and language condition (for \textit{prosody only} speech compared to the \textit{no speech} reference level; \textit{$\beta$}=0.393, \textit{SE}=0.189, \textit{z}=2.08, \textit{p}$<$.05), and a three-way interaction of age, transition type, and language condition (for \textit{normal} speech compared to the \textit{no speech} reference level; \textit{$\beta$}=-0.38, \textit{SE}=0.17, \textit{z}=-2.229, \textit{p}$<$.05).
+There were no significant effects of age or transition type alone (Table \ref{sec:models2}), with only a marginal effect of language condition (for \textit{prosody only} compared to the \textit{no speech} reference level; \textit{$\beta$}=-1.607, \textit{SE}=0.867, \textit{z}=-1.85, \textit{p}$=$.064)
+
+Adults' anticipatory gaze switches showed effects of gap duration (\textit{$\beta$}=4.7, \textit{SE}=1.18, \textit{z}=3.978, \textit{p}$<$.001) and language condition (for \textit{normal} speech \textit{$\beta$}=1.203, \textit{SE}=0.536, \textit{z}=2.245, \textit{p}$<$.05 and \textit{words only} speech \textit{$\beta$}=1.561, \textit{SE}=0.709, \textit{z}=2.203, \textit{p}$<$.05  compared to the \textit{no speech} reference level).
+There were no effects of transition type (\textit{$\beta$}=0.437, \textit{SE}=0.585, \textit{z}=0.747, \textit{p}$=$.45).
 
@@
 These analyses revealed that none of the significant predictors from models of the original, turn-related data could be explained by random looking.
-In the children's data, the original model's \textit{t}-values for language condition (\textit{prosody only}), gap duration, the two-way interaction of age and language condition (\textit{prosody only}) and the three-way interaction of age, transition type, and language condition (\textit{normal} speech) were all greater than 95\% of the randomly permuted \textit{t}-values (96.2\%, 100\%, 95.1\%, and 95.1\%, respectively).
-Similarly, the adults' data showed significant differentiation from the randomly permuted data for all originally significant predictors: gap duration and language condition for \textit{normal} speech and words-only speech (greater than 100\%, 96.8\%, and 98.7\% of random \textit{t}-values, respectively).
-% The effects of language condition and transition type for the real and randomly permuted data can also be observed in Figure \ref{fig:E1-randvsreal}.
+In the children's data, the original model's \textit{z}-values for language condition (\textit{prosody only}), gap duration, the two-way interaction of age and language condition (\textit{prosody only}) and the three-way interaction of age, transition type, and language condition (\textit{normal} speech) were all greater than 95\% of the randomly permuted \textit{z}-values (96.5\%, 100\%, 95.6\%, and 95.5\%, respectively).
+Similarly, the adults' data showed significant differentiation from the randomly permuted data for all originally significant predictors: gap duration and language condition for \textit{normal} speech and words-only speech (greater than 100\%, 97.8\%, and 99.1\% of random \textit{z}-values, respectively).
 
@@
 Two- and three-year-old children made anticipatory gaze switches significantly above chance both when all transitions were considered (2-year-olds: \textit{t}(26.193)$=$-4.137, \textit{p}$<$.001; 3-year-olds: \textit{t}(22.757)$=$-2.662, \textit{p}$<$.05) and for question transitions alone (2-year-olds: \textit{t}(25.345)$=$-4.269, \textit{p}$<$.001; 3-year-olds: \textit{t}(21.555)$=$-3.03, \textit{p}$<$.01).
-One-year-olds, however, made anticipatory gaze shifts that were not statistically significant for turn transitions overall and for question turns alone (overall: \textit{t}(24.784)$=$-2.049, \textit{p}$=$.051; questions: \textit{t}(25.009)$=$-2.03, \textit{p}$=$.053).
+One-year-olds, however, only made anticipatory gaze shifts marginally above chance for turn transitions overall and for question turns alone (overall: \textit{t}(24.784)$=$-2.049, \textit{p}$=$.051; questions: \textit{t}(25.009)$=$-2.03, \textit{p}$=$.053).
 
@@
 The first was a significant interaction of age and language condition (for \textit{prosody only} compared to the \textit{no speech} reference level), suggesting a different age effect between the two linguistic conditions.
-As in Experiment 1, we further explored each age interaction by extracting an average difference score over subjects for the effect of language condition (\textit{no speech} vs.
+As in Experiment 1, we explored each age interaction by extracting an average difference score over participants for the effect of language condition (\textit{no speech} vs.
 \textit{prosody only}) within each random permutation of the data, making pairwise comparisons between the six age groups.
@@
 As in Experiment 1, with normal speech, children and adults spontaneously tracked the turn structure of the conversations, making anticipatory gaze switches at an above-chance rate across all ages.
-And in addition, they made far more anticipations for questions than for non-question turns---at least for two-year-olds and older.
-But these effects were different for the two comparison conditions, \textit{prosody-only} and \textit{words-only}.
-In the \textit{prosody-only} condition, performance was low for younger children and increased substantially for older children (especially for questions).
-In the \textit{words-only} condition, anticipation performance was relatively robust for questions for two-year-olds and older (much like in normal speech), but never rose above chance for the children for non-question turns.
-Thus, these findings do not support an early role for prosody in children's spontaneous turn structure predictions.
-On the contrary, children's predictions appeared to stem more from lexical information, probably primarily from question words.
+They also made far more anticipations for questions than for non-question turns---at least for children two years olds and older.
+But these effects were different for the two conditions with partial linguistic information: \textit{prosody-only} and \textit{words-only}.
+In the \textit{prosody-only} condition, performance was low for younger children and increased significantly with age (especially for questions).
+In the \textit{words-only} condition, children age two and older showed robust anticipatory switching for questions (much like in normal speech), but never rose above chance for non-question turns.
+These findings do not therefore support an early role for prosody in children's spontaneous turn structure predictions.
+On the contrary, children's predictions appeared to stem more from lexical information, possibly from lexicosyntactic cues to questionhood.
 
@@
 
-Children begin to develop conversational turn-taking skills long before their first words \citep{bateson1975, hilbrink2015, jaffe2001, snow1977}.
+Children begin to develop conversational turn-taking skills long before their first words emerge\citep{bateson1975, hilbrink2015, jaffe2001, snow1977}.
 As they acquire language, they also acquire the information needed to make accurate predictions about upcoming turn structure.
 Until recently, we have had very little data on how children weave language into their already-existing turn-taking behaviors.
-In two experiments investigating children's anticipatory gaze to upcoming speakers, we found evidence that turn prediction develops early in childhood and that spontaneous predictions are primarily driven by participants' expectation of an immediate response in the next turn (e.g., for questions).
+In two experiments investigating children's anticipatory gaze to upcoming speakers, we found evidence that turn prediction develops early in childhood and that spontaneous predictions are primarily driven by participants' expectation of an immediate response in the next turn (e.g., after questions).
 In making predictions about upcoming turn structure, children used a combination of lexical and prosodic cues; neither lexical nor prosodic cues alone were sufficient to support increased anticipatory gaze.
@@
 
-Prior work with adults has found a consistent and critical role for lexicosyntax in predicting upcoming turn structure \citep{de-ruiter2006, magyari2012}, with the role of prosody still under debate \citep{duncan1972, ford1996, torreira2015}.
-Knowing that children comprehend more about prosody than lexicosyntax early on (see introduction; also see \citealp{speer2009} for a review), we thought it possible that young children would instead show an advantage for prosody in their predictions about turn structure in conversation.
-Our results suggest that, on the contrary, when presented with \textit{only} prosodic information, children's spontaneous predictions about upcoming turn structure are limited until age five.
-
-Importantly, we also found no evidence that lexical information alone is equivalent to full linguistic information for children, as has been shown before \citep{magyari2012, de-ruiter2006} and replicated in the current study for adult participants.
-In both experiments, children's performance was best in conditions when they had access to the full linguistic signal.
-Adults on the other hand, showed significant gains in anticipatory gaze switching in both conditions with lexical cues.
-If this effect arose in our data because children do not make predictions based on phonetically manipulated speech in conversational contexts, the current findings can be overturned by follow-up work controlling linguistic information through other means.
-But if not, there may be something specially informative about the combined prosodic and lexical cues to questionhood that boosts children's anticipations before they can use these cues separately.
-Even in adults, Torreira and colleagues (\citeyear{torreira2015}) were able to show that the trade-off in informativity between lexical and prosodic cues is more subtle in semi-natural (spliced) speech.
+Prior work with adults has found a consistent role for lexicosyntax in predicting upcoming turn structure \citep{de-ruiter2006, magyari2012}, whereas the role of prosody still under debate \citep{duncan1972, ford1996, torreira2015}.
+Knowing that children comprehend more about prosody than lexicosyntax early on (see \ref{sec:intro}; also see \citealp{speer2009} for a review), we thought it possible that young children would instead show an advantage for prosody in their predictions about turn structure in conversation.
+Our results suggest that, on the contrary, exclusively presenting prosodic information to children limits their spontaneous predictions about upcoming turn structure until age five.
+
+Perhaps surprisingly, we also found no evidence that lexical information alone is equivalent to the full linguistic signal in driving children's predictions, as has been shown previously for adults \citep{magyari2012, de-ruiter2006} and is replicated with adult participants in the current study.
+That said, our findings point more toward early lexical effects in children's turn anticipations than prosodic ones: children's performance in both experiments was consistently better when they had access to lexical information, especially after question turns.
+And even though the \textit{words only} condition in Experiment 2 was not itself significantly different from the baseline \textit{no speech} condition, children's anticipations trended toward above-chance rates.
+
+Above all, children and adults anticipated best with they had access to the full linguistic signal.
+There may be something informative about combined prosodic and lexical cues to questionhood that helps to boost children's anticipations before they can use these cues separately.
+Even in adults, Torreira and colleagues (\citeyear{torreira2015}) showed that the trade-off in informativity between lexical and prosodic cues is more subtle in semi-natural (spliced) speech.
 The present findings are the first to show evidence of a similar effect developmentally.
@@
 
-Compared to prosodic cues (e.g., final rising intonation), lexicosyntactic cues (e.g., \textit{wh}-words, \textit{do}-insertion, and subject-auxiliary inversion) were frequent, categorical, and early-occurring in the utterance.
+Compared to prosodic cues (e.g., final rising intonation), lexicosyntactic cues to questionhood (e.g., \textit{wh}-words, \textit{do}-insertion, and subject-auxiliary inversion) are categorical, and early-occurring in the utterance.
 Children may have therefore had an easier time picking out and interpreting lexical cues to questionhood.
-The question effect showed its first significant gains between ages three and four in the \textit{normal} speech condition of Experiment 2, by which time children frequently hear and use a variety of polar \textit{wh}-questions \citep{clark2009}.
-Furthermore, while lexicosyntactic question cues were available on every instance of \textit{wh}- and \textit{yes/no} questions in our stimuli, prosodic question cues were only salient on \textit{yes/no} questions and, even then, the mapping of prosodic contour to speech act (e.g., high final rises for polar questions) is far from one-to-one.
-
-Prior work on children's acquisition of questions indicates that they may already have some understanding about question-answer sequences by the time they begin to speak: Questions make up approximately one third of the utterances children hear, before and after the onset of speech, and even into their preschool years, even though the types and complexity of questions change throughout development \citep{casillas2016, fitneva2012, henning2005, shatz1979}.\footnote{There is substantial variation question frequency by individual and socioeconomic class \citep{hart1992}.} For the first few years, many of the questions directed to children are ``test'' questions---questions that the caregiver already has the answer to (e.g., ``What does a cat say?''), but this changes as children get older.
+The question effect showed its first significant gains between ages three and four in the \textit{normal} speech condition of Experiment 2, by which time children frequently hear and use a variety of polar and \textit{wh}-questions \citep{clark2009}.
+Furthermore, while lexicosyntactic question cues were available on every instance of \textit{wh}- and \textit{yes/no} questions in our stimuli, prosodic question cues were only salient on \textit{yes/no} questions.
+Even still, the mapping of prosodic contour to speech act (e.g., high final rises for polar questions) is far from one-to-one, leaving substantial room for uncertainty in prosodic contour interpretation.
+
+Prior work on children's acquisition of questions indicates that they may already have some knowledge of question-answer sequences by the time they begin to speak: Questions make up approximately one third of the utterances children hear, before and after the onset of speech, and even into their preschool years, though the types and complexity of questions change throughout development \citep{casillas2016, fitneva2012, henning2005, shatz1979}.\footnote{There is substantial variation question frequency by individual and socioeconomic class \citep{hart1992}.} For the first few years, many of the questions directed to children are ``test'' questions---questions that the caregiver already has the answer to (e.g., ``What does a cat say?''), but this changes as children get older.
 Questions help caregivers to get their young children's attention and to ensure that information is in common ground, even if the responses are non-verbal or infelicitous \citep{bruner1985, fitneva2012, snow1977}.
@@
 
-% The question effect itself has implications for our current theories about turn prediction.
-While participants may always use linguistic information to predict upcoming speaker changes (as they have in the two studies presented here), they might not always use linguistic information to predict upcoming turn ends, as is assumed in metalinguistic measures of turn-end prediction (e.g., pressing a button while listening to speech: \citealp{torreira2015, magyari2012, de-ruiter2006}) and as has been the focus of other turn-end prediction studies (\citealp{ford1996, duncan1972}).
-Our results rather suggest that participants' \textit{spontaneous} predictions, at least while viewing third-party conversation, are the results of question-monitoring, presumably achieved by recruiting linguistic cues to questionhood from the unfolding signal.
-In other words, our results suggest that predictions are driven by what is \textit{beyond} the end of the current turn---that questions, not lexical cues, are sufficient for prediction at the first level of conversation monitoring.
+More broadly, our results suggest that participants' \textit{spontaneous} predictions, at least while viewing third-party conversation, are driven by what lies \textit{beyond} the end of the current turn---not by the upcoming end of the turn itself, as has been often assumed \citep{torreira2015, keitel2013, magyari2012, de-ruiter2006}.
+In future work, it will be crucial to measure prediction from a first-person perspective to resolve this apparent contradiction (see also \citealp{holler2015}).
 
@@
 One of the core aims of our study was to test whether children show an early competence for turn taking, as is proposed by studies of spontaneous mother-infant proto-conversation and theories about the mechanisms underlying human interaction in general \citep{hilbrink2015, levinson2006}.
-We did find evidence that young children already make spontaneous predictions about upcoming turn structure, definitely at age two and even marginally at age one.
-However, ``above chance'' performance was far from adult-like predictive behavior, and children in our studies did not show adult-like competency in their predictions, even at age six.
-This may indicate that children rely more on non-verbal cues in anticipating turn transitions or, alternatively, that adults are better at flexibly adapting to the turn-relevant cues present at any moment.
-
-% We could add here that this flexibility might emerge around age 5: That's when we see some mini-adult-like effects in the children's data.
-In E1 that take the form of their (potentially) using either non-verbal or non-native prosodic cues in the non-English videos.
-In E2 this takes the form of 5 and 6 year olds showing significant improvement in the \textit{prosody only} condition.
-
-Taken together, the data suggest that turn-taking skills do begin to emerge in infancy, but that their predictions don't help them anticipate much until they have acquired the ability to pick out question turns.
+We found evidence that young children make spontaneous predictions about upcoming turn structure: definitely at age two and even marginally at age one.
+
+These results contrast with Keitel and colleagues' \citeyearpar{keitel2013} finding that children cannot anticipate upcoming turn structure at above-chance rates until age three.
+The current study used an appreciably more conservative random baseline than the one used in Keitel and colleagues' study.
+Therefore, this difference in age of emergence more likely stems from our use of a more engaging speech style, stereo speech playback, and more typical turn transition durations.
+
+To be clear, young children's ``above chance'' performance was often still far from adult-like predictive behavior---even at age six, children in our studies did not show adult-like competency in their predictions.
+This may indicate that young children rely more on non-verbal cues in anticipating turn transitions or, alternatively, that adults are better at flexibly adapting to the turn-relevant cues present at any moment.
+
+% This added flexibility might emerge around age 5, when we see the first adult-like effects in the children's data: In Experiment 1, 5-year-olds using non-verbal or non-native prosodic cues to make predictions in the non-English videos comparable to their predictions for English.
+In Experiment 2, 5- and 6-year-olds showing significant advances over younger children in their anticipations for the \textit{prosody only} condition.
+
+Taken together, our data suggest that turn-taking skills do begin to emerge in infancy, but that children cannot make effective predictions until they can pick out question turns.
 This finding leads us to wonder how participant role (first- instead of third-person) and cultural differences (e.g., high vs.
-low parent-infant interaction styles) might feed into this early predictive skill.
-It also bridges the prior work showing a predisposition for turn taking in infancy (e.g., \citealp{hilbrink2015}) but late acquisition of adult-like competence when it comes to integrating linguistic information into turn-taking behaviors \citep{casillas2016, garvey1984, ervin-tripp1979}.
+low parent-infant interaction styles) feed into this early predictive skill.
+This finding also bridges  prior work showing a predisposition for turn taking in infancy (e.g., \citealp{hilbrink2015}) but late acquisition of adult-like competence for turn-taking in children's conversations \citep{casillas2016, garvey1984, ervin-tripp1979}.
 
@@
 There are at least two major limitations to our work: Speech naturalness and participant role.
-Following prior work \citep{de-ruiter2006, keitel2013}, we used phonetically manipulated speech in Experiment 2, resulting in speech sounds that children don't usually hear in their natural environment.
+Following prior work \citep{de-ruiter2006, keitel2013}, we used phonetically manipulated speech in Experiment 2.
+This decision resulted in speech sounds that children don't usually hear in their natural environment.
 Many prior studies have used phonetically-altered speech with infants and young children \citep[cf.][]{jusczyk2000}, but almost none of them have done so in a conversational context.
-Future work could instead carefully script or cross-splice parts of turns to control for the presence or absence of linguistic cues for turn transition.
-
-The prediction measure used in these studies is based on an observer's view of third-party conversation but, because participants' role in the interaction could affect their online predictions about turn taking, an ideal experimental measure would capture first-person behavior.
+Future work could instead carefully script or cross-splice sub-parts of turns to control for the presence or absence of linguistic cues for turn transition (see, e.g., \citealp{torreira2015}).
+
+The prediction measure used in our studies is based on an observer's view of third-party conversation but, because participants' role in the interaction could affect their online predictions about turn taking, an better measure would instead capture first-person behavior.
 First-person measures of spontaneous turn prediction will be key to revealing how participants distribute their attention over linguistic and non-verbal cues while taking part in everyday interaction, the implications of which relate to theories of online language processing for both language learning and everyday talk.
@@
 \section{Permutation Analyses}
+\label{sec:permutation}
 \setcounter{figure}{0}
-
-We completed this baseline procedure on 5,000 random permutations of the original turn transition analysis windows and compared the \textit{t}-values from each predictor in the original models (Table \ref{tab:E1-models}) to the distribution of \textit{t}-values for each predictor in the 5,000 models of the randomly permuted datasets.\footnote{We report \textit{t}-values rather than beta estimates because the standard errors in the randomly permuted data models were much higher than for the original data.
-For those interested, plots of the beta and standard error distributions are available in the Supplementary Materials.} We could then test whether significant effects from the original statistical models differed from the random baseline by calculating the proportion of random data \textit{t}-values exceeded by the original \textit{t}-value for each predictor, using the absolute value of all \textit{t}-values for a two-tailed test.
-For example, children's original ``language condition'' \textit{t}-value was $|$3.429$|$, which is greater than 99.9\% of all $|$\textit{t}-value$|$ estimates from the randomly-permuted data models (i.e., \textit{p}$=$.001).
-This leads us to conclude that that the effect of language condition in the original model was highly unlikely to be the result of random gaze shifting.
-
-E1: We excluded the output of random-permutation models that gave convergence warnings in order to remove unreliable estimates from our analyses (non-converging models were 22.4\% and 24.4\% of all models for children and adults, respectively; see the Supplementary Materials for more information on model exclusion).
-
-E2: As before, we excluded the output of random-permutation models that resulted in convergence warnings in order to remove unreliable model estimates from our analyses (non-converging models made up 69\% and 70\% of models for children and adults, respectively; see the Supplementary Materials for more information on model exclusion).
-
-In all of the following plots, the gray dots represent the randomly permuted data's model estimates for the value listed (beta or standard error), the white dots represent the model estimates from the original data, and the triangles represent the 95th percentile for each distribution being shown.
-
-\subsection{Experiment 1}
-
-\begin{figure}[!htb]
+\setcounter{table}{0}  
+How can we be sure that our primary dependent measure (anticipatory gaze switching) actually relates to turn transitions?
+Even if children were gazing back and forth randomly during the experiment, we would have still captured some false hits---switches that ended up in the turn-transition windows by chance.
+
+We estimated the baseline probability of making an anticipatory switch by randomly permuting the placement of the transition windows within each stimulus (Figure \ref{fig:shuffling}).
+We then used the switch identification procedure from Experiments 1 and 2 (Section \ref{sec:algorithm}) to find out how often participants made ``anticipatory'' switches within these randomly permuted windows.
+This procedure de-links participants' gaze data from turn structure by randomly re-assigning the onset time of each turn-transition in each permutation.
+We created 5,000 of these permutations for each experiment to get an anticipatory switch baselines over all possible starting points.
+
+Importantly, the randomized windows were not allowed to overlap with each other, keeping true to the original stimuli.
+We also made sure that the properties of each turn transition stayed constant across permutations.
+So, while ``transition window A'' might start 2 seconds into Random Permutation 1 and 17 seconds into Random Permutation 2, it maintains the same prior speaker identity, transition type, gap duration, language condition, etc., across both permutations.
+
+We then re-ran the statistical models from the original data on each of the random permutations, e.g., using Experiment 1's original model to analyze the anticipatory switches from each random permutation of the Experiment 1 looking data.
+We could then calculate the proportion of random data \textit{z}-values exceeded by the original \textit{z}-value for each predictor.
+We used the absolute value of all \textit{z}-values to conduct a two-tailed test.
+If the original effect of a predictor exceeded 95\% of the random model effects for that same predictor, we deemed that predictor's effect to be significantly different from the random baseline.
+
+For example, children's ``language condition'' effect from Experiment 1 had a \textit{z}-value of $|$3.429$|$, which is greater than 99.9\% of all $|$\textit{z}-value$|$ estimates from Experiment 1's random permutation models (i.e., \textit{p}$=$.001).
+It is therefore highly unlikely that the effect of language condition in the original model derived from random gaze shifting.
+
+We used this procedure to derive the random-baseline comparison values in the main text (above).
+However, we ran into two issues along the way: First, we had to report \textit{z}-values rather than beta estimates.
+Second, we had to exclude a substantial portion of the models, especially in Experiment 2 because of convergence issues.
+We address each of these issues below.
+
+\subsection{Beta, standard error, and z estimates in the data}
+We reported \textit{z}-values in the main text rather than beta estimates because the standard errors in the randomly permuted data models were much higher than for the original data.
+The distributions of each predictor's beta estimate, standard error, and \textit{z}-value for adults and children in each experiment are shown in the graphs below (Figures \ref{fig:E1-ChiTs}--\ref{fig:E2-AduSEs}).
+In each plot, the gray dots represent randomly permuted model estimates for the value listed (beta, standard error, or \textit{z}), the white dots represent the model estimates from the original data, and the triangles represent the 95th percentile for each random distribution.
+
+\begin{sidewaysfigure}[!htb]
   \centering
-  \subfloat[Children]{\includegraphics[width=0.44\textwidth]{figures/E1-chi-randrun-t-vals-absolute.png}\label{fig:E1-ChiTs}}
+  \textbf{Experiment 1: \textit{z}-value estimates}\par\medskip
+  \subfloat[Children]{\includegraphics[width=0.44\textwidth]{figures/E1-chi-randrun-z-vals-absolute.png}\label{fig:E1-ChiTs}}
   \hfill
-  \subfloat[Adults]{\includegraphics[width=0.44\textwidth]{figures/E1-adu-randrun-t-vals-absolute.png}\label{fig:E1-AduTs}}
-  \caption{Random-permutation and original $|$\textit{t}-values$|$ for predictors of anticipatory gaze rates in Experiment 1.}
-\end{figure}
-
-\begin{figure}[!htb]
+  \subfloat[Adults]{\includegraphics[width=0.44\textwidth]{figures/E1-adu-randrun-z-vals-absolute.png}\label{fig:E1-AduTs}}
+  \caption{Random-permutation and original $|$\textit{z}-values$|$ for predictors of anticipatory gaze rates in Experiment 1.}
+\label{fig:E1-Ts}
+\end{sidewaysfigure}
+
+\begin{sidewaysfigure}[!htb]
   \centering
+  \textbf{Experiment 1: \textit{$\beta$} estimates}\par\medskip
   \subfloat[Children]{\includegraphics[width=0.44\textwidth]{figures/E1-chi-randrun-betas-absolute.png}\label{fig:E1-ChiBs}}
@@
   \caption{Random-permutation and original $|$\textit{$\beta$}-values$|$ for predictors of gaze rates in Experiment 1.}
-\end{figure}
-
-\begin{figure}[!htb]
+\label{fig:E1-Bs}
+\end{sidewaysfigure}
+
+\begin{sidewaysfigure}[!htb]
   \centering
+  \textbf{Experiment 1: \textit{SE} estimates}\par\medskip
   \subfloat[Children]{\includegraphics[width=0.44\textwidth]{figures/E1-chi-randrun-SEs-absolute.png}\label{fig:E1-ChiSEs}}
@@
   \caption{Random-permutation and original $|$\textit{SE}-values$|$ for predictors of anticipatory gaze rates in Experiment 1.}
-\end{figure}
-
-\clearpage
-
-\subsection{Experiment 2}
-
-\begin{figure}[!htb]
+\label{fig:E1-SEs}
+\end{sidewaysfigure}
+
+\begin{sidewaysfigure}[!htb]
   \centering
-  \subfloat[Children]{\includegraphics[width=0.44\textwidth]{figures/E2-chi-randrun-t-vals-absolute.png}\label{fig:E2-ChiTs}}
+  \textbf{Experiment 2: \textit{z} estimates}\par\medskip
+  \subfloat[Children]{\includegraphics[width=0.44\textwidth]{figures/E2-chi-randrun-z-vals-absolute.png}\label{fig:E2-ChiTs}}
   \hfill
-  \subfloat[Adults]{\includegraphics[width=0.44\textwidth]{figures/E2-adu-randrun-t-vals-absolute.png}\label{fig:E2-AduTs}}
-  \caption{Random-permutation and original $|$\textit{t}-values$|$ for predictors of anticipatory gaze rates in Experiment 2.}
-\end{figure}
-
-\begin{figure}[!htb]
+  \subfloat[Adults]{\includegraphics[width=0.44\textwidth]{figures/E2-adu-randrun-z-vals-absolute.png}\label{fig:E2-AduTs}}
+  \caption{Random-permutation and original $|$\textit{z}-values$|$ for predictors of anticipatory gaze rates in Experiment 2.}
+\label{fig:E2-Ts}
+\end{sidewaysfigure}
+
+\begin{sidewaysfigure}[!htb]
   \centering
+  \textbf{Experiment 2: \textit{$\beta$} estimates}\par\medskip
   \subfloat[Children]{\includegraphics[width=0.44\textwidth]{figures/E2-chi-randrun-betas-absolute.png}\label{fig:E2-ChiBs}}
@@
   \caption{Random-permutation and original $|$\textit{$\beta$}-values$|$ for predictors of anticipatory gaze rates in Experiment 2.}
-\end{figure}
-
-\begin{figure}[!htb]
+\label{fig:E2-Bs}
+\end{sidewaysfigure}
+
+\begin{sidewaysfigure}[!htb]
   \centering
+  \textbf{Experiment 2: \textit{SE} estimates}\par\medskip
   \subfloat[Children]{\includegraphics[width=0.44\textwidth]{figures/E2-chi-randrun-SEs-absolute.png}\label{fig:E2-ChiSEs}}
@@
   \caption{Random-permutation and original $|$\textit{SE}-values$|$ for predictors of anticipatory gaze rates in Experiment 2.}
-\end{figure}
-
-\section{Pairwise developmental tests: Real vs. randomly permuted effects}
-\setcounter{figure}{0}
-
-In each of the plots below, the dot represents the original data value for the effect and the 5,000 randomly permuted data effect sizes are shown in the distribution.
-The percentage shown is the percentage of random permutation values exceeded by the original data value (taking the absolute value of all data points for a two-tailed test.)
-
-\begin{figure}[!htb]
-\begin{center}
-\includegraphics[width=0.8\textwidth]{figures/E1-child-randvsreal-ttest-agebylg.png}
-\end{center}
-\caption{Pairwise comparisons of the language condition effect across ages in Experiment 1.} 
-\label{fig:E1-lgageinteraction}
-\end{figure}
-
-\begin{figure}[!htb]
-\begin{center}
-\includegraphics[width=0.9\textwidth]{figures/E2-child-randvsreal-ttest-muffledages.png}
-\end{center}
-\caption{Significant pairwise comparisons of the \textit{prosody only}-\textit{no speech} linguisitic condition effect, across ages in Experiment 2} 
-\label{fig:E2-lgageinteraction}
-\end{figure}
-
-\begin{figure}[!htb]
-\begin{center}
-\includegraphics[width=0.9\textwidth]{figures/E2-child-randvsreal-ttest-normaltypesages.png}
-\end{center}
-\caption{Significant pairwise comparisons of the \textit{normal speech}-\textit{no speech} language condition effect for transition type, across ages, in Experiment 2.} 
-\label{fig:E2-lgagetypeinteraction}
-\end{figure}
-
-\section{Non-convergent models}
-\setcounter{table}{0}
-
-Non-convergent models made up 22--24\% of the 5,000 models of randomly permuted data in Experiment 1 and 69--70\% of the 5,000 for Experiment 2.
-We excluded these non-convergent models because they displayed erratic $\beta$ and \textit{SE} estimates, as summarized below in the table of \textit{t}-values from convergent and non-convergent models in Experiment 1.
-The non-convergent models from Experiment 2 showed similar patterns.
-The high frequency of problematic models persisted even when we changed optimizers and we suspect the issue derives from data sparsity in some of the random runs.
+\label{fig:E2-SEs}
+\end{sidewaysfigure}
+
+\clearpage
+
+\subsection{Non-convergent models}
+In comparing the real and randomly permuted datasets, we excluded the output of random-permutation models that gave convergence warnings in order to remove erratic model estimates from our analyses.
+Non-convergent models made up 22.4--24.4\% of the random permutation models in Experiment 1 and 69--70\% of the random permutation models in Experiment 2.
+The \textit{z}-values for each predictor in the converging and non-converging models from Experiment 1 are shown in table \ref{tab:nonconv_e1}.
+
+Although many of the non-converging models show estimates within range of the converging models (e.g., with a mean difference of only 0.09 in median \textit{z}-value across predictors), they also show many radically outlying estimates (e.g., showing a mean difference of 237.3 in mean \textit{z}-value across predictors).
+Similar patterns were obtained in the non-converging models for Experiment 2 and persisted even when we changed optimizers.
+
+We suspect that the issue derives from data sparsity in some of the random runs.
+This problem is known to occur when there are limited numbers of binary observations in each of design matrix's many bins \citep{allison2004}.
+We could instead use zero-inflated poisson (ZIP) or negative binomial regression models to allow for overdispersion in our data \citep{allison2012}.
+However, these would give us baselines for the normal, convergent model, which is not the aim of this analysis.
 
@@
   \begin{tabular}{lllllllllll}
-    Variable 				& Mean-C & Mean-NC & Median-C & Median-NC & \textit{SD}-C & \textit{SD}-NC & Min-C & Min-NC & Max-C & Max-NC \\ 
+    Variable & Mean$_{C}$ & Mean$_{NC}$ & Median$_{C}$ & Median$_{NC}$ & \textit{SD}$_{C}$ & \textit{SD}$_{NC}$ & Min$_{C}$ & Min$_{NC}$ & Max$_{C}$ & Max$_{NC}$ \\ 
     \hline
@@
     LgCond							& -0.53	& -109.91	& -0.55	& -0.63	& 0.93	& 564.42	& -3.93	& -4418.74		& 3.23	& 2296.19 \\
-    Type					& -0.10	& -29.66		& -0.09	& -0.1	& 0.98	& 515.12		& -4.06	& -4383.92	& 3.23	& 2296.19		\\
-    Duration				& 0.99	& 345.53		& 0.98	& 1.15	& 1.07	& 1323.13		& -2.44	& -5048.24	& 3.36	& 3416.68		\\
-    Age*LgCond				& 0.19	& 10.64		& 0.2		& 0.18	& 0.9		& 109.6		& -3.31	& -581.61		& 5.78	& 9985.16		\\
-    Age*Type				& 0.02	& -1.8		& 0.001	& -0.05	& 0.9		& 98.27		& -3.36	& -884.36		& 3.59	& 946.81		\\
-    LgCond*Type			& 0.2		& 45.32		& 0.2		& 0.27	& 0.96	& 691.3		& -3.12	& -4160.06	& 3.45	& 640.43		\\
-    Age*LgCond*Type		& -0.12	& -14.23		& -0.12	& -0.15	& 0.93	& 156.72		& -2.98	& -1318.26	& 3.39	& 5107.64		\\
+    Type							& -0.1	& -29.66	& -0.09	& -0.1	& 0.98	& 515.12	& -4.06	& -4383.92		& 3.36	& 3416.68 \\
+    Duration						& 0.99	& 345.53	& 0.98	& 1.15	& 1.07	& 1323.13	& -2.44	& -5048.24		& 5.78	& 9985.16 \\
+    Age*LgCond					& 0.19	& 10.64		& 0.2		& 0.18	& 0.9		& 109.6		& -3.31	& -581.61		& 3.59	& 946.81 \\
+    Age*Type						& 0.02	& -1.8		& 0.001	& -0.04	& 0.9		& 98.27		& -3.36	& -884.36		& 3.45	& 640.43 \\
+    LgCond*Type					& 0.2		& 45.32		& 0.2		& 0.27	& 0.96	& 691.3		& -3.12	& -4160.06		& 3.39	& 5107.64 \\
+    Age*LgCond*Type			& -0.12	& -14.23	& -0.12	& -0.15	& 0.93	& 156.72	& -2.98	& -1318.26		& 2.90	& 927.69 \\
     \hline
@@
     LgCond  						& -0.26	& -679.6	& -0.3	& -0.53	& 1.02	& 15894.33	& -3.45	& -494979.7	& 3.35	& 88581.58	\\
-    Type 					& -0.11	& 6.29		& -0.13	& -0.04	& 1.11	& 501.5		& -3.85	& -6420.76	& 3.28	& 8177.88		\\
+    Type 							& -0.11	& 6.29		& -0.13	& -0.04	& 1.11	& 501.5		& -3.85	& -6420.75		& 3.28	& 8177.88		\\
     Duration  					& 0.25	& 84.09		& 0.27	& 0.26	& 1.1		& 1152.94	& -3.25	& -10864.51	& 3.46	& 18540.62	\\
-    LgCond*Type  			& 0.12	& -242.27		& 0.1		& 0.34	& 1.07	& 26836.7		& -3.41	& -62264.27	& 3.81	& 509198.4	\\
+    LgCond*Type  				& 0.12	& -242.27	& 0.1		& 0.34	& 1.07	& 26836.7	& -3.41	& -622642.7	& 3.81	& 509198.4	\\
     LgCond*Duration			& 0.15	& 780.03	& 0.16	& 0.39	& 1.04	& 44105.02	& -3.84	& -798498.6	& 3.55	& 1145951	\\
@@
   \end{tabular}
-  \caption{Estimated \textit{t}-values for each predictor in converging (`-C') and non-converging (`-NC') adult and child models from Experiment 1. }
+  \caption{Estimated \textit{z}-values for each predictor in converging (\textit{C}) and non-converging (\textit{NC}) adult and child models from Experiment 1. }
 \label{tab:nonconv_e1}
@@
 
-\clearpage
-
-\section{Miscellaneous}
+\section{Pairwise developmental tests}
+\label{sec:pairwisedev}
 \setcounter{figure}{0}
+Experiments 1 and 2 both showed effects of age in interaction with linguistic condition and transition type (e.g., English vs.
+non-English).
+To explore these effects in more depth, we recorded the average difference score for the predictor that interacted with age (e.g., linguistic condition) for each participant (e.g., English minus non-English anticipatory switches), using these values to compute an average difference score over participants in each age group (e.g., age 3, 4, and 5) within each random permutation.
+That averaging process produces 5,000 baseline-derived difference scores for each age group.
+
+We then completed pairwise age comparisons of these difference scores (e.g., the linguistic condition effect in 3-year-olds vs.
+4-year-olds), computing the percent of random-permutation difference scores exceeded by the real-data difference score.
+If the real-data difference score exceeded 95\% of the random-data age difference scores, we deemed it significantly different from chance---a significant difference in age between the two groups being compared (e.g., a significant difference between ages three and four in the effect of linguistic condition).
+
+This procedure is essentially a two-tailed \textit{t}-test, adapted for use with the randomly permuted baseline data.
+
+In each of the plots below, the dot represents the real data value for the effect being shown.
+The 5,000 randomly permuted data effect sizes are shown in the distribution.
+The percentage displayed is the percentage of random permutation values exceeded by the original data value (taking the absolute value of all data points for a two-tailed test).
+Comparisons marked with 95\% or higher are significant at the \textit{p}$<$0.05 level.
+
+\begin{sidewaysfigure}[!htb]
+\begin{center}
+\textbf{Experiment 1: Age and linguistic condition}\par\medskip
+\includegraphics[width=0.95\textwidth]{figures/E1-child-randvsreal-ttest-agebylg.png}
+\end{center}
+\caption{Pairwise comparisons of the language condition effect across ages in Experiment 1.} 
+\label{fig:E1-lgageinteraction}
+\end{sidewaysfigure}
+
+\begin{sidewaysfigure}[!htb]
+\begin{center}
+\textbf{Experiment 2: Age and the prosody only condition}\par\medskip
+\includegraphics[width=0.95\textwidth]{figures/E2-child-randvsreal-ttest-muffledages.png}
+\end{center}
+\caption{Significant pairwise comparisons of the \textit{prosody only}-\textit{no speech} linguisitic condition effect, across ages in Experiment 2} 
+\label{fig:E2-lgageinteraction}
+\end{sidewaysfigure}
+
+\begin{sidewaysfigure}[!htb]
+\begin{center}
+\textbf{Experiment 2: Age, transition type, and normal speech}\par\medskip
+\includegraphics[width=0.95\textwidth]{figures/E2-child-randvsreal-ttest-normaltypesages.png}
+\end{center}
+\caption{Significant pairwise comparisons of the \textit{normal speech}-\textit{no speech} language condition effect for transition type, across ages, in Experiment 2.} 
+\label{fig:E2-lgagetypeinteraction}
+\end{sidewaysfigure}
+
+\section{Boredom-driven anticipatory looking}
+\label{sec:boredlooks}
+\setcounter{figure}{0}
 
 One alternative hypothesis for children's anticipatory gazes is that they simply grow bored and start looking away at a constant rate after a turn begins.
-This data plotted here show a hypothetical group of participants who begin to lose interest (at a linear rate) after one second of a turn (gray dots) compared to participants' real data from Experiment 2 (black dots).
-This pattern suggests that, though children do look away with time, their looks away are not simply driven by boredom.
+This data plotted here show a hypothetical group of boredom-driven participants (gray dots) compared to participants from the actual data in Experiment 2 (black dots).
+The hypothetical boredom-driven participants look away from the current speaker at a linear rate, beginning one second after the start of a turn.
 
@@
 \begin{center}
-\includegraphics[width=0.9\textwidth]{figures/boredom-hypothesis.png}
-\end{center}
-\caption{Proportion participants looking at the current speaker: hypothetical boredom-driven data (gray dots) versus real data from Experiment 2 (black dots).} 
+\includegraphics[width=0.95\textwidth]{figures/boredom-hypothesis.png}
+\end{center}
+\caption{Proportion of participants looking at the current speaker: hypothetical boredom-driven data (gray dots) versus actual data from Experiment 2 (black dots).
+Vertical bars indicate standard error in the experiment data.}
 \label{fig:boredomhypothesis}
@@
 
-\bigskip
-
+If children's switches away from the current speaker were driven by boredom, they would switch away equally quickly on long and short turns.
+However, as turn duration increased, children in the experiment looked at the current speaker for a longer period, suggesting that they were not switching away at a constant rate.
+We can see this pattern most clearly by looking at the time when 50\% of the children had switched away from the current (indicated by the horizontal line in Figure \ref{fig:boredomhypothesis}).
+Children's gaze crossed this 50\% threshold at 2.0, 2.9, and 3.6 seconds after the start of speech for turns with durations of 1--2, 2--3, and 3--4 seconds, respectively.
+In contrast, the hypothetical boredom-driven children always hit the 50\% threshold at 2.5 seconds after the start of speech.
+This pattern suggests that, though children do look away with time, their looks away are not simply driven by boredom.
+
+\clearpage
+
+\section{Puppet pair and linguistic condition}
+\label{sec:puppetconfound}
+\setcounter{figure}{0}  
 The design for Experiment 2 does not fully cross puppet pair (e.g., robots, blue puppets) with linguistic condition (e.g., ``words only'' and ``no speech'').
 Even though each puppet pair is associated with different conversation clips across children (e.g.
-robots talking about kitties, birthday parties, or pancakes), robots were only associated with ``words only'' speech, merpeople were only associated with ``prosody only'' speech, and the puppets with fancy clothes were only associated with the ``no speech'' condition.
-We did this to increase the pragmatic felicity of the experiments for the older children (i.e., robots make robot sounds, merpeople's voices are muffled under water, the fancy-clothed puppets are in a room with main other voices).
-It is therefore fair to point out a possible confound between linguistic condition and puppet pair.
-Thankfully, we also ran a short follow-up study at the museum with 3--5-year-olds in which each child only saw one video---the normal speech conversation about birthday parties---with a randomly assigned puppet pair performing the conversation.
-Five children watched each puppet pair, for a total of 30 children across the six pairs.
-This experiment holds all things constant except for the appearance of the puppets.
-We then used a mixed effects logistic regression of children's anticipatory switches (yes or no at each transition), with puppet pair (robots/merpeople/fancy dress/other-3; Figure \ref{fig:speakers}) as a fixed effect and participant and turn transition as random effects.
-In four versions of this model we systematically varied the reference level to check for differences between every puppet pair, finding no significant affects of puppet type on switching rate.
-We take this as evidence that, although we did not fully cross puppet pairs and linguistic conditions in Experiment 2, it was unlikely to have had strong effects on children's looking rates above and beyond the intended effects of linguistic condition.
+robots talking about kitties, birthday parties, and pancakes), the robot puppets themselves were exclusively associated with the ``words only'' condition.
+Similarly, merpeople were exclusively associated with ``prosody only'' speech, and the puppets wearing dress clothes were exclusively associated with the ``no speech'' condition.
+We designed the experiment this way to increase its pragmatic felicity for older children (i.e., robots make robot sounds, merpeople's voices are muffled under the water, the fancy-clothed puppets are in a `party' room with many other voices).
+There is therefore a possible confound between linguistic condition and puppet pair; for example, children could have made fewer anticipatory switches in the \textit{prosody only} condition because the puppets were less interesting.
+To test whether puppet pair drove the condition-based differences found in Experiment 2, we ran a short follow-up study.
+
+\bigskip
+\noindent \textbf{Methods}
+\medskip
+
+\noindent We recruited 30 children between ages 3;0--5;11 from Children's Discovery Museum of San Jose, California to participate in our experiment.
+All participants were native English speakers.
+Children were randomly assigned to one of six videos (five children per video).
+
+\medskip
+\noindent \textit{Materials}.
+We created 6 short videos from the stimulus recordings made for Experiment 2.
+Each video featured a puppet pair (red/blue/yellow/robot/ merpeople/dressy; Figure \ref{fig:puppets}).
+Puppets in all six videos performed the exact same conversation recording (`birthday party'; Experiment 2) with normal, unmanipulated speech; this experiment therefore holds all things constant except for the appearance of the puppets.
+
+\medskip
+\noindent \textit{Procedure}.
+We used the same experimental apparatus and procedure as in Experiments 1 and 2.
+Each participant was randomly assigned to watch only one of the six puppet videos.
+Five children watched each video in total.
+As in Experiment 2, the experimenter immediately began each session with calibration and then stimulus presentation because no special instructions were required.
+The entire experiment took less than three minutes.
+
+\medskip
+\noindent \textit{Data preparation}. We identified anticipatory gaze switches to the upcoming speaker using the same method as in Experiments 1 and 2.
+
+\bigskip
+\noindent \textbf{Results and discussion}
+\medskip
+
+\noindent We modeled children's anticipatory switches (yes or no at each transition) with mixed effects logistic regression, including puppet pair (robots/mer- people/fancy dress/other-3) as a fixed effect and participant and turn transition as random effects.
+We grouped the red, blue, and yellow puppets together because they collectively represent the puppets used in the \textit{normal} speech condition and this follow-up experiment is meant to test whether the condition-based differences from Experiment 2 arose from the puppets used in each condition.
 
@@
 \begin{center}
-\includegraphics[width=0.9\textwidth]{figures/all-puppetdyads-bars.png}
+\includegraphics[width=0.95\textwidth]{figures/all-puppetdyads-bars.png}
 \end{center}
@@
 
+\clearpage
+
+In four versions of this model we systematically varied the reference level of the puppet pair to check for any cross-condition differences.
+We found no significant effects of puppet pair on switching rate (all \textit{p} $>$ 0.25; Table \ref{tab:control_exp}).
+
+\begin{table}
+\begin{center}
+  \begin{tabular}{llcccc}
+    \hline
+           &  Estimate & Std. Error & z value & Pr($>$$|$z$|$) \\    
+    \hline
+		 \footnotesize{\textit{Intercept: normal-condition puppets}} &&&& \\
+    \hline
+     (Intercept) 								& -0.14790	& 0.32796	& -0.451		& 0.652 \\ 
+     Puppets$=$\textit{mermaid} 	& -0.07581	& 0.65532	& -0.116		& 0.908 \\ 
+     Puppets$=$\textit{robot} 		& -0.07104	& 0.65321	& -0.109		& 0.913 \\ 
+     Puppets$=$\textit{dressy} 		& -0.78206	& 0.68699	& -1.138		& 0.255 \\ 
+    \hline
+		 \footnotesize{\textit{Intercept: mermaid puppets}} &&&& \\
+    \hline
+     (Intercept) 								& -0.22371	& 0.56832	& -0.394		& 0.694 \\ 
+     Puppets$=$\textit{robot} 		& 0.004763	& 0.80096	& 0.006		& 0.995 \\ 
+     Puppets$=$\textit{dressy} 		& -0.70626	& 0.82742	& -0.854		& 0.393 \\ 
+    \hline
+		 \footnotesize{\textit{Intercept: robot puppets}} &&&& \\
+    \hline
+     (Intercept) 								& -0.21895	& 0.56565	& -0.387		& 0.699 \\ 
+     Puppets$=$\textit{dressy} 		& -0.71102	& 0.82657	& -0.860		& 0.390 \\ 
+    \hline
+		 \footnotesize{\textit{Intercept: dressy puppets}} &&&& \\
+    \hline
+     (Intercept) 								& -0.9300	& 0.6067		& -1.533		& 0.125 \\ 
+  \end{tabular}
+\end{center}
+  \caption{Model output for children's anticipatory gaze switches with reference levels varied to show all possible pairwise differences between puppet pairs.}
+\label{tab:control_exp}
+\end{table}
+
+We take this finding as evidence that our decision to not fully cross puppet pairs and linguistic conditions in Experiment 2 was unlikely to have strongly affected children's anticipatory gaze rates above and beyond the intended effects of linguistic condition.
+
 \end{document}

### 2016-01-13 paper/CasillasFrank-Anticip.tex
small changes to text and formatting

@@
 Children begin developing turn-taking skills in infancy but take several years to assimilate their growing knowledge of language into their turn-taking behavior.
-In two eye-tracking experiments, we measured children's anticipatory gaze to upcoming responders while controlling linguistic cues to upcoming turn structure.
+In two eye-tracking experiments, we measured children's anticipatory gaze to upcoming responders while controlling linguistic cues to turn structure.
 In Experiment 1, we showed English and non-English conversations to English-speaking adults and children.
-%, finding minimal differences between the predictive looking behavior of preschoolers and adults. 
 In Experiment 2, we phonetically controlled lexicosyntactic and prosodic cues in English-only speech.
-Children's predictive looking behavior improved from ages one to six, but even one-year-olds made more anticipatory looks than would be expected by chance.
-In both experiments, children and adults anticipated more often after hearing questions. 
+Children spontaneously made anticipatory gaze switches by age two and continued improving through age six.
+In both experiments, children and adults made more anticipatory switches after hearing questions.
 Like adults, prosody alone did not improve children's predictive gaze shifts.
-But, unlike adults, lexical cues alone were also not sufficient to improve prediction---children's performance was best overall with access to lexicosyntax and prosody together.
+But, unlike adults, lexical cues alone were not sufficient to improve prediction---children's performance was best overall with lexicosyntax and prosody together.
 Our findings support an account in which turn prediction emerges in infancy, but becomes fully integrated with linguistic processing only gradually.
@@
 \begin{keyword}
-Turn taking \sep Conversation \sep Development \sep Prosody \sep Lexical \sep Questions \sep Eye-tracking \sep Anticipation
+Turn taking \sep Conversation \sep Development \sep Prosody \sep Questions \sep Eye-tracking \sep Anticipation
 %% MSC codes here, in the form: \MSC code \sep code
@@
 Like other types of human interaction, it is organized at its core by the roles and goals of its participants.
-But what sets conversation apart is its structure: Sequences of interconnected, communicative actions that take place across alternating turns at talk.
+But, what sets conversation apart is its structure: sequences of interconnected, communicative actions that take place across alternating turns at talk.
 Sequential, turn-based structures in conversation are strikingly uniform across language communities and linguistic modalities.
 Turn-taking behaviors are also cross-culturally consistent in their basic features and the details of their implementation \citep{de-vos2015, dingemanse2013, stivers2009}.
-% How does this ability develop?
-
-Children participate in sequential coordination with their caregivers starting at three months of age---before they can rely on any linguistic cues in taking turns \citep[see, among others, ][]{bateson1975, hilbrink2015, jaffe2001, snow1977}.
-Infant turn taking is different from adult turn taking in several ways, however.
-Infant turn taking is heavily scaffolded by caregivers, has different timing from adult turn taking, and lacks semantic content \citep{hilbrink2015, jaffe2001}.
-But children's early, turn-structured social interactions are presumably a critical precursor to their later conversational turn taking: Early non-verbal interactions likely establish the protocol by which children come to use language with others.
+
+Children participate in sequential coordination (proto-turn taking) with their caregivers starting at three months of age---before they can rely on any linguistic cues \citep[see, among others, ][]{bateson1975, hilbrink2015, jaffe2001, snow1977}.
+Infant turn taking is different from adult turn taking in several ways, however: it is heavily scaffolded by caregivers, has different timing from adult turn taking, and lacks semantic content \citep{hilbrink2015, jaffe2001}.
+But children's early, turn-structured social interactions are presumably a critical precursor to their later conversational turn taking: early non-verbal interactions likely establish the protocol by which children come to use language with others.
 How do children integrate linguistic knowledge with these preverbal turn-taking abilities, and how does this integration change over the course of childhood?
@@
 Children's early games (e.g., give-and-take, peek-a-boo) have built-in, predictable turn structure \citep{ratner1978, ross1987}.
-Even monkeys take turns: Non-human primates such as marmosets and Campbell's monkeys vocalize contingently with each other in both natural and lab-controlled environments \citep{lemasson2011, takahashi2013}.
-In all these cases, turn taking serves as a protocol for interaction, allowing the participants to coordinate with one another through sequences of contingent action.
-
-Conversation distinguishes itself from non-conversational turn-taking behaviors by the complexity of the turn sequencing involved.
-In the examples above (traffic, games, and monkeys) the set of sequence and action types is far more limited and predictable than what we find in everyday talk.
+Even monkeys take turns: non-human primates such as marmosets and Campbell's monkeys vocalize contingently with each other in both natural and lab-controlled environments \citep{lemasson2011, takahashi2013}.
+In all these cases, turn taking serves as a protocol for interaction, allowing the participants to coordinate with each other through sequences of contingent action.
+
+Conversational turn taking distinguishes itself from other turn-taking behaviors by the complexity of the sequencing involved.
 Conversational turns come grouped into semantically-contingent sequences of action.
 The groups can span turn-by-turn exchanges (e.g., simple question--response, ``How are you?''--``Fine.'') or sequence-by-sequence exchanges (e.g., reciprocals, ``How are you?''--``Fine, and you?''--``Great!'').
+Compared to other turn-taking behaviors, the possible sequence and action types in everyday talk are can be diverse and unpredictable.
 
@@
 Non-verbal cues (e.g., gaze, posture, and pointing) often appear at turn-boundaries and can sometimes act as late indicators of an upcoming speaker switch \citep{rossano2009, stivers2010}.
-Additionally, the sequential context of a turn can make it clear what will come next: Answers after questions, thanks or denial after compliments, et cetera \citep{schegloff2007}.
+Additionally, the sequential context of a turn can make it clear what will come next: answers after questions, thanks or denial after compliments, etc.
+\citep{schegloff2007}.
 
@@
 The authors concluded that lexicosyntactic information\footnote{The ``lexicosyntactic'' condition only included flattened pitch and so was not exclusively lexicosyntactic---the speech would still have residual prosodic structure, including syllable duration and intensity.} was necessary and possibly sufficient for turn-end projection, while intonation was neither necessary nor sufficient.
-Congruent evidence comes from studies varying the predictability of lexicosyntactic and pragmatic content: Adults anticipate turn ends better when they can more accurately predict the exact words that will come next (\citealp{magyari2012}; see also \citealp{magyari2014}).
+Congruent evidence comes from studies varying the predictability of lexicosyntactic and pragmatic content: adults anticipate turn ends better when they can more accurately predict the exact words that will come next (\citealp{magyari2012}; see also \citealp{magyari2014}).
 They can also identify speech acts within the first word of an utterance \citep{gisladottir2015}, allowing them to start planning their response at the first moment possible \citep{bogels2015}.
@@
 De Ruiter and colleagues' (2006) experiment focused on the role of intonation, which is only a partial index of prosody.
-And in addition, prosody is tied closely to the syntax of an utterance, so the two linguistic signals are difficult to control independently \citep{ford1996}.
+Prosody is tied closely to the syntax of an utterance, so the two linguistic signals are difficult to control independently \citep{ford1996}.
 \citet*{torreira2015} used a combination of button-press and verbal responses to investigate the relationship between lexicosyntactic and prosodic cues in turn-end prediction.
@@
 
-% \subsubsection{Observational studies}
-
 The majority of work on children's early turn taking has focused on observations of spontaneous interaction.
-Children's first turn-like structures appear as early as two to three months in proto-conversation with their caregivers \citep{bruner1975, bruner1985}.
-During proto-conversations, caregivers interact with their infants as if they were capable of making meaningful contributions: They take every look, vocalization, arm flail, and burp as ``utterances'' in the joint discourse \citep{bateson1975, jaffe2001, snow1977}.
+Children's first turn-like structures appear as early as two to three months after birth, in proto-conversation with their caregivers \citep{bruner1975, bruner1985}.
+During proto-conversations, caregivers treat their infants as capable of making meaningful contributions: they take every look, vocalization, arm flail, and burp as ``utterances'' in the joint discourse \citep{bateson1975, jaffe2001, snow1977}.
 Infants catch onto the structure of proto-conversations quickly.
@@
 The timing of children's responses to their caregivers' speech shows a non-linear pattern.
-Infants' contingent vocalizations in the first few months of life show very fast timing (though with a lot of vocal overlap) that, by nine months, slows down considerably, only gradually speeding up again after 12 months \citep{hilbrink2015}.
-Taking turns with brief transitions between speakers is difficult for children; while their avoidance of overlap is nearly adult-like by nine months, the timing of their non-overlapped responses stays much longer than the 200 msec standard for the next few years \citep{casillas2016, garvey1984, ervin-tripp1979}.
-This puzzling pattern is likely due to their linguistic development: Taking turns on time is easier when the response is a simple vocalization rather than a linguistic utterance.
+Infants' contingent vocalizations in the first few months of life show very fast timing (though with a lot of vocal overlap).
+But by nine months, their timing slows down considerably, only to gradually speed up again after 12 months \citep{hilbrink2015}.
+Taking turns with brief transitions between speakers is difficult for children; while their avoidance of overlap is nearly adult-like by nine months, the timing of their non-overlapped responses stays much longer than the adult 200 msec standard for the next few years \citep{casillas2016, garvey1984, garvey1981, ervin-tripp1979}.
+This puzzling pattern is likely due to their linguistic development: taking turns on time is easier when the response is a simple vocalization rather than a linguistic utterance.
 Integrating language into the turn-taking system may be one major factor in children's delayed responses \citep{casillas2016}.
 
-While children, like adults, could use linguistic cues in the ongoing turn to make predictions about upcoming turn structure, studies of early linguistic development point to a possible early advantage for prosody over lexicosyntax in children's turn-taking predictions.
+While children, like adults, might use linguistic cues in the ongoing turn to make predictions about upcoming turn structure, studies of early linguistic development point to a possible early advantage for prosody over lexicosyntax in children's turn-taking predictions.
 Infants can distinguish their native language's rhythm type from others soon after birth \citep{mehler1988, nazzi2003}; they show preference for the typical stress patterns of their native language over others by 6--9 months (e.g., iambic vs.
 trochaic), and can use prosodic information to segment the speech stream into smaller chunks from 8 months onward \citep{johnson2001, morgan1995}.
-Four- to five-month-olds also prefer pauses in speech to be inserted at prosodic boundaries, and by 6 months they can start using prosodic markers to pick out sub-clausal syntactic units, both of which are useful for extracting turn structure from ongoing speech \citep{jusczyk1995, soderstrom2003}.
+Four- to five-month-olds also prefer pauses in speech to be inserted at prosodic boundaries, and by 6 months infants can use prosodic markers to pick out sub-clausal syntactic units, both of which are useful for extracting turn structure from ongoing speech \citep{jusczyk1995, soderstrom2003}.
 In comparison, children show at best a very limited lexical inventory before their first birthday \citep{bergelson2013, shi2010}.
@@
 Keitel and colleagues \citeyearpar{keitel2013} were one of the first to explore how children use linguistic cues to predict upcoming turn structure.
-They asked 6-, 12-, 24-, 36-month-old, and adult participants to watch short videos of conversation and tracked their eye movements at points of speaker change.
+They asked 6-, 12-, 24-, and 36-month-old infants, and adult participants to watch short videos of conversation and tracked their eye movements at points of speaker change.
 They showed their participants two types of conversation videos---one normal and one with flattened pitch (i.e., with flattened intonation contours)---to test the role of intonation in participants' anticipatory predictions about upcoming speech.
@@
 Among those, only 36-month-olds were affected by a lack of intonation contours, leading Keitel and colleagues to conclude that children's ability to predict upcoming turn structure relies on their ability to comprehend the stimuli lexicosemantically.
-They also suggested that intonation might play a secondary role in turn prediction, but only after children acquire more sophisticated, adult-like language comprehension abilities.
+They also suggest that intonation might play a secondary role in turn prediction, but only after children acquire more sophisticated, adult-like language comprehension abilities (also see \citealp{keitel2015}).
 
@@
 Instead, they used gaze switches during ongoing speech as a baseline.
-But ongoing speech is perhaps the period in which switching is least likely to occur \citep{hirvenkari2013}--- thus, this particular baseline maximizes the chance of finding a difference between gaze frequency at turn transitions and baseline.
-A more conservative baseline would be to compare participants' looking behavior at turn transitions to their looking behavior during randomly selected windows of time throughout the stimulus, including turn transitions.
-We follow this conservative approach in our work.
+But ongoing speech is the period in which switching is least likely to occur \citep{hirvenkari2013}---their baseline thus maximizes the chance of finding a difference in gaze frequency at turn transitions compared to the baseline.
+A more conservative baseline would compare participants' looking behavior at turn transitions to their looking behavior during randomly selected windows of time throughout the stimulus, including turn transitions.
+We follow this conservative approach in the current study.
 
 Second, the conversation stimuli \citet{keitel2013} used were somewhat unusual.
-The average gap between turns was 900 msec, which is much longer than typical adult timing, where gaps average around 200 msec \citep{stivers2009}.
-The speakers in the videos were also asked to minimize their movements while performing a scripted and adult-directed conversation, which would have created a somewhat unnatural stimulus.
-Additionally, in order to produce more naturalistic conversation, it would have been ideal to localize the sound sources for the two voices in the video (i.e., to have the voices come out of separate left and right speakers).
-But both voices were recorded and played back on the same audio channel, which may have made it more difficult to distinguish the two talkers (again, we attempt to address these issues in our current study).
+The average gap between turns was 900 msec, which is much longer than typical adult timing, which averages around 200 msec \citep{stivers2009}.
+The speakers in the videos were also asked to minimize their movements while performing scripted, adult-directed conversation, which would have created a somewhat unnatural interaction.
+Additionally, to produce more naturalistic conversation, it would have been ideal to localize the sound sources for the two voices in the video (i.e., to have the voices come out of separate left and right speakers).
+But both voices were recorded and played back on the same audio channel, which may have made it difficult to distinguish the two talkers.
+Again, we attempt to address these issues in our current study.
 Despite these minor methodological issues, the Keitel et al.
 \citeyearpar{keitel2013} study still demonstrates intriguing age-based differences in children's ability to predict upcoming turn structure.
-Our current work thus takes this paradigm as a starting point.\footnote{See also \citet{casillas2012, casillas2013}.}
+Our current work thus takes this paradigm as a starting point.\footnote{But also see \citet{casillas2012, casillas2013}.}
 
@@
 Our goal in the current study is to find out when children begin to make predictions about upcoming turn structure and to understand how their predictions are affected by linguistic cues across development.
-We present two experiments in which we measure children's anticipatory gaze to responders while watching conversation videos with natural (people using English vs.
+We present two experiments in which we measured children's anticipatory gaze to responders while they watched conversation videos with natural (people speaking English vs.
 non-English; Experiment 1) and non-natural (puppets with phonetically manipulated speech; Experiment 2) control over the presence of lexical and prosodic cues.
@@
 
-Because the results of our experiments are complex, we highlight three primary findings.
-First, although children and adults use linguistic cues to make predictions about upcoming turn structure, they do so primarily in predict speaker transitions after questions (a speech act effect).
-This speech act effect, which we did not initially predict, is intriguing and suggests that previous work may have neglected an important dimension of linguistic cues.
-Second, we find that children make more predictions than expected by chance starting at age two, but that this effect is small with our stimuli.
-Third, we find no evidence of an early prosody advantage in children's anticipations and, further, no evidence that prosodic or lexical cues alone can substitute for their combination in the full linguistic signal, as is proposed for adults \citep{de-ruiter2006}; instead, anticipation is strongest for a stimulus with the full range of cues.
+We highlight three primary findings: first, although children and adults use linguistic cues to make predictions about upcoming turn structure, they do so primarily in predict speaker transitions after questions (a speech act effect).
+This intriguing effect, which has not been reported previously, suggests that participants track unfolding speech for cues to upcoming speaker change, which may affect how they use linguistic cues more generally for anticipatory processing in conversation.
+Second, we find that children make more predictions than expected by chance starting at age two, but that this effect is small at first, and continues to improve through age six.
+Third, we find no evidence of an early prosody advantage in children's anticipations and, further, no evidence that prosodic or lexical cues alone can substitute for their combination in the full linguistic signal (as is proposed for adults; \citealp{de-ruiter2006}); instead, anticipation is strongest for stimuli with the full range of cues.
 In sum, our findings support an account in which turn prediction emerges in infancy, but becomes fully integrated with linguistic processing only gradually across development.
@@
 The participants, all native English speakers, were only expected to understand the two videos in English.
-We showed participants non-English videos to limit their access to lexical information while maintaining their access to other cues to turn boundaries (e.g., (non-native) prosody, gaze, breath, phrase final lengthening).
+We showed participants non-English videos to limit their access to lexical information while maintaining their access to other cues to turn boundaries (e.g., (non-native) prosody, gaze, inbreaths, phrase final lengthening).
 Using this method, we compared children and adult's anticipatory looks from the current speaker to the upcoming speaker at points of turn transition in English and non-English videos.
@@
 \textit{Video recordings}.
-We recorded pairs of talkers while they conversed in a sound-attenuated booth (see sample frame in Figure \ref{fig:speakers}).
+We recorded pairs of talkers while they conversed in a sound-attenuated booth (see a sample frame in Figure \ref{fig:speakers}).
 Each talker was a native speaker of the language being recorded, and each talker pair was male-female.
@@
 
-After recording, we combined the audio and video files by hand, and cropped each recording to the 30-second interval with the most turn activity.
+After recording, we combined the audio and video recordings by hand, and cropped each recording to the 30-second interval with the most turn activity.
 Because we recorded the conversations in stereo, the male and female voices came out of separate speakers during video playback.
@@
 non-question).
-The non-English stimuli were coded for speech act from a monolingual English-speaker's perspective, i.e., which turns ``sound like'' questions, and which don't: We asked five native American English speakers to listen to the audio signal for each turn and judge whether it sounded like a question.
+The non-English stimuli were coded for speech act from a monolingual English-speaker's perspective, i.e., which turns ``sound like'' questions, and which do not: we asked five native American English speakers to listen to the audio recording for each turn and judge whether it sounded like a question.
 We then coded turns with at least 80\% ``yes'' responses as questions.
@@
 We measured the duration of each turn transition from the audio recording associated with each video.
-We excluded turn transitions longer than 550 msec and shorter than 90 msec, including overlapped transitions, from analysis.\footnote{Overlap occurs when a responder begins a new turn before the current turn is finished.
-When overlap occurs, observers cannot switch their gaze in anticipation of the response because the response began earlier than expected; participants expect conversations to proceed with ``one speaker at a time'' \citep{sacks1974}.
-As such, they would still be fixated on the prior speaker when the overlap started, and then would have to switch their gaze \textit{reactively} to the responder.} This left approximately equal numbers of turn transitions available for analysis in the English (N$=$20) and non-English (N$=$16) videos.
+We excluded turn transitions longer than 550 msec and shorter than 90 msec, also excluding overlapped transitions, from analysis.\footnote{Overlap occurs when a responder begins a new turn before the current turn is finished.
+When overlap occurs, observers cannot switch their gaze in anticipation of the response because the response began earlier than expected.
+Participants expect conversations to proceed with ``one speaker at a time'' \citep{sacks1974}.
+They would therefore still be fixated on the prior speaker when the overlap started, and would have to switch their gaze \textit{reactively} to the responder.} This left approximately equal numbers of turn transitions available for analysis in the English (N$=$20) and non-English (N$=$16) videos.
 On average, the inter-turn gaps for English videos (mean$=$318, median$=$302, stdev$=$112 msec) were slightly longer than for non-English videos (mean$=$286, median$=$251, stdev$=$122 msec).
-The longer gaps in the English videos could give them a slight advantage: Our definition of an ``anticipatory gaze shift'' includes shifts that are initiated during the gap between turns (Figure \ref{fig:criterion}), so participants had slightly more time to make anticipatory shifts in the English videos.
+The longer gaps in the English videos could give them a slight advantage: our definition of an ``anticipatory gaze shift'' includes shifts that are initiated during the gap between turns (Figure \ref{fig:criterion}), so participants had slightly more time to make anticipatory shifts in the English videos.
 
@@
 
-Children were less inclined to simply sit and watch videos of conversation in languages they didn't speak, so we used a different procedure to keep them engaged: The experimenter started each session by asking the child about what languages he or she could speak, and about what other languages he or she had heard of.
+Children were less inclined to simply sit and watch videos of conversation in languages they didn't speak, so we used a different procedure to keep them engaged: the experimenter started each session by asking the child about what languages he or she could speak, and about what other languages he or she had heard of.
 Then the experimenter expressed her own enthusiasm for learning about new languages, and invited the child to watch a video about ``new and different languages'' together.
@@
 
-We checked each participant's gaze at each turn transition for three characteristics (Figure \ref{fig:criterion}): (1) That the participant fixated on the prior speaker for at least 100 msec at the end of the prior turn, (2) that sometime thereafter the participant switched to fixate on the upcoming speaker for at least 100 ms, and (3) that the switch in gaze was initiated within the first 200 msec of the response turn, or earlier.
-These criteria guarantee that we only counted gaze shifts when: (1) Participants were tracking the previous speaker, (2) switched their gaze to track the upcoming speaker, and (3) did so before they could have simply reacted to the onset of speech in the response.
+We checked each participant's gaze at each turn transition for three characteristics (Figure \ref{fig:criterion}): (1) that the participant fixated on the prior speaker for at least 100 msec at the end of the prior turn, (2) that sometime thereafter the participant switched to fixate on the upcoming speaker for at least 100 ms, and (3) that the switch in gaze was initiated within the first 200 msec of the response turn, or earlier.
+These criteria guarantee that we only counted gaze shifts when: (1) participants were tracking the previous speaker, (2) switched their gaze to track the upcoming speaker, and (3) did so before they could have simply reacted to the onset of speech in the response.
 Under this assumption, a gaze shift that was initiated within the first 200 msec of the response (or earlier) was planned \textit{before} the child could react to the onset of speech itself.
@@
 Participants looked at the screen most of the time during video playback (81\% and 91\% on average for children and adults, respectively).
-They primarily kept their eyes on the person who was currently speaking in both English and non-English videos: They gazed at the current speaker between 38\% and 63\% of the time, looking back at the addressee between 15\% and 20\% of the time (Table \ref{tab:e1_look}).
+They primarily kept their eyes on the person who was currently speaking in both English and non-English videos: they gazed at the current speaker between 38\% and 63\% of the time, looking back at the addressee between 15\% and 20\% of the time (Table \ref{tab:e1_look}).
 Even three-year-olds looked more at the current speaker than anything else, whether the videos were in a language they could understand or not.
@@
     \hline
-                   & Estimate & Std. Error & z value & Pr($>$$|$z$|$)    \\
+			           &  Estimate & Std. Error & \textit{z} value & Pr($>$$|$\textit{z}$|$) \\    
     \hline
@@
     \hline
-                       &  Estimate & Std. Error & z value & Pr($>$$|$z$|$) \\    
+           &  Estimate & Std. Error & \textit{z} value & Pr($>$$|$\textit{z}$|$) \\    
     \hline
@@
 non-question), age (3, 4, 5; numeric), and duration of the inter-turn gap (seconds, e.g., 0.441) as predictors, with full interactions between condition, transition type, and age.
-We included the duration of the inter-turn gap as a predictor since longer gaps provide more opportunities to make anticipatory switches (Figure \ref{fig:criterion}).
+We included the duration of the inter-turn gap as a control predictor since longer gaps provide more opportunities to make anticipatory switches (Figure \ref{fig:criterion}).
 We additionally included random effects of item (turn transition) and participant, with random slopes of condition, transition type, and their interaction for participants \citep{barr2013}.\footnote{The models we report are all qualitatively unchanged by the exclusion of their random slopes.
@@
 
-Adults' anticipatory gaze switches shows an effect of transition type (\textit{$\beta$}=-4.5, \textit{SE}=1.315, \textit{z}=-3.42, \textit{p}$<$.001) and significant interactions between language condition and transition type (\textit{$\beta$}=3.3, \textit{SE}=1.61, \textit{z}=2.05, \textit{p}$<$.05) and transition type and gap duration (\textit{$\beta$}=10.51, \textit{SE}=3.348, \textit{z}=3.139, \textit{p}$<$.01).
+Adults' anticipatory gaze switches showed an effect of transition type (\textit{$\beta$}=-4.5, \textit{SE}=1.315, \textit{z}=-3.42, \textit{p}$<$.001) and significant interactions between language condition and transition type (\textit{$\beta$}=3.3, \textit{SE}=1.61, \textit{z}=2.05, \textit{p}$<$.05) and transition type and gap duration (\textit{$\beta$}=10.51, \textit{SE}=3.348, \textit{z}=3.139, \textit{p}$<$.01).
 
@@
 We estimated the probability that these patterns were the result of random looking by running the same regression models on participants' real eye-tracking data, only this time calculating their anticipatory gaze switches with respect to randomly permuted turn transition windows.
-This process involved: (1) Randomizing the order and temporal placement of the analysis windows within each stimulus (Figure \ref{fig:shuffling}; ``analysis window'' is defined in Figure \ref{fig:criterion}), thereby randomly redistributing the analysis windows across the eye-tracking signal, (2) re-running each participant's eye tracking data through switch identification (described in Section \ref{sec:algorithm}), this time using the randomly permuted analysis windows, and (3) modeling the anticipatory gazes from the randomly permuted data with the same statistical models we used for the original data (Section \ref{sec:models1}; Table \ref{tab:E1-models}).
+This process involved: (1) randomizing the order and temporal placement of the analysis windows within each stimulus (Figure \ref{fig:shuffling}; ``analysis window'' is defined in Figure \ref{fig:criterion}) to randomly redistribute the analysis windows across the eye-tracking signal, (2) re-running each participant's eye tracking data through switch identification (described in Section \ref{sec:algorithm}) on each of the randomly permuted analysis windows, and (3) modeling the anticipatory switches from the randomly permuted data with the same statistical models we used for the original data (Section \ref{sec:models1}; Table \ref{tab:E1-models}).
 Importantly, although the onset time of each transition was shuffled within the eye-tracking signal, the other intrinsic properties of each turn transition (e.g., prior speaker identity, transition type, gap duration, language condition, etc.) stayed constant across each random permutation.
@@
 
-This procedure effectively de-links participants' gaze data from the turn structure in the original stimulus, thereby allowing us to compare turn-related (original) and non-turn-related (randomly permuted) looking behavior using the same eye movements.
+This procedure effectively de-links participants' gaze data from the turn structure in the original stimulus, thereby allowing us to compare turn-related (original) and non-turn-related (randomly permuted) looking behavior using the same eye movement data.
 The resulting anticipatory gazes from the randomly permuted analysis windows represent an average anticipatory gaze rate over all possible starting points: a random baseline.
-By running the real and randomly permuted data sets through identical statistical models, we can also estimate how likely it is that predictor effects in the original data (e.g., the effect of language condition; Table \ref{tab:E1-models}) arose from random looking.
+
+By running the real and randomly permuted data sets through identical statistical models, we can estimate how likely it is that predictor effects in the original data (e.g., the effect of language condition; Table \ref{tab:E1-models}) arose from random looking.
 Because these analyses are complex, we report their full details in \ref{sec:permutation}.
@@
 Our baseline analyses revealed that none of the significant predictors from models of the original, turn-related data can be explained by random looking.
-For the children's data, the original \textit{z}-values for language condition, gap duration, the age-language condition interaction, and the language condition-transition type interaction were all greater than 95\% of \textit{z}-values for the randomly permuted data (99.9\%, 95.5\%, 99.4\%, and 96\%, respectively).
-Similarly, the adults' data showed significant differentiation from the randomly permuted data for two of the three originally significant predictors---transition type and the transition type-gap duration interaction (greater than 99.9\% and 99.7\% of random \textit{z}-values, respectively)---with marginal differentiation for the interaction of language condition and transition type (greater than 94.6\% of random \textit{z}-values).
+For the children's data, the original \textit{z}-values for language condition, gap duration, the age-language condition interaction, and the language condition-transition type interaction were all greater than 95\% of \textit{z}-values for the randomly permuted data (99.9\%, 95.5\%, 99.4\%, and 96\%, respectively, all \textit{p}$<$.05).
+Similarly, the adults' data showed significant differentiation from the randomly permuted data for two of the three originally significant predictors---transition type and the transition type-gap duration interaction (greater than 99.9\% and 99.7\% of random \textit{z}-values, respectively, all \textit{p}$<$.01)---with marginal differentiation for the interaction of language condition and transition type (greater than 94.6\% of random \textit{z}-values; \textit{p}$=$.054).
 
@@
 The models reported above revealed a significant interaction of age and language condition (Table \ref{tab:E1-models}) that was unlikely be due to random looking (Figure \ref{fig:E1-randvsreal}).
-To further explore this effect, we compared the average effect of language condition for each age group: Using the permutation analyses above, we extracted the average difference score for the two language conditions (English minus non-English) for each participant, computing an overall average for each random permutation of the data.
+To further explore this effect, we compared the effect of language condition across age groups: Using the permutation analyses above, we extracted the average difference score for the two language conditions (English minus non-English) for each participant, computing an overall average for each random permutation of the data.
 Then, within each permutation, we made pairwise comparisons of the average difference scores across participant age groups.
@@
 
-These analyses revealed that, while 3- and 4-year olds showed similarly-sized effects of language condition, 5-year-olds showed a significantly smaller effect of language condition, compared to both younger age groups.
-The difference in language condition effect between 5-year-olds and 3-year-olds was greater than would be expected by chance in 99.52\% of the randomly permuted data sets (\textit{p}$<$.01).
-Similarly, the difference in language condition effect between 5-year-olds and 4-year-olds was greater than would be expected by chance in  99.96\% of the data sets (\textit{p}$<$.001; see Figure \ref{fig:E1-lgageinteraction} for difference score distributions).
-
-When does spontaneous turn prediction emerge developmentally during natural speech?
+These analyses revealed that, while 3- and 4-year olds showed similarly-sized effects of language condition, 5-year-olds had a significantly smaller effect of language condition, compared to both younger age groups.
+The difference in the language condition effect between 5-year-olds and 3-year-olds was greater than would be expected by chance (99.52\% of the randomly permuted data sets; \textit{p}$<$.01).
+Similarly, the difference in the language condition effect between 5-year-olds and 4-year-olds was greater than would be expected by chance (99.96\% of the data sets; \textit{p}$<$.001).
+See Figure \ref{fig:E1-lgageinteraction} for each difference score distribution.
+
+When does spontaneous turn prediction emerge developmentally?
 To test whether the youngest age group (3-year-olds) already exceeded chance in their anticipatory gaze switches, we compared children's real gaze rates to the random baseline in the English condition with two-tailed \textit{t}-tests.
+We used the English condition because we are most interested in finding out when children begin to make spontaneous turn predictions for natural speech.
 We found that three-year-olds made anticipatory gaze switches significantly above chance, when all transitions were considered (\textit{t}(22.824)$=$-4.147, \textit{p}$<$.001) as well as for question transitions alone (\textit{t}(21.677)$=$-5.268, \textit{p}$<$.001).
@@
 Prior work using online, metalinguistic tasks has shown that participants can use linguistic cues to accurately predict upcoming turn ends \citep{torreira2015, magyari2012, de-ruiter2006}.
-The current results add a new dimension to our understanding of how listeners make predictions about turn ends: Both children and adults spontaneously monitor the linguistic structure of unfolding turns for cues to imminent responses.
+The current results add a new dimension to our understanding of how listeners make predictions about turn ends: both children and adults spontaneously monitor the linguistic structure of unfolding turns for cues to imminent responses.
 
@@
 
-Third, we saw developmental effects such that older children anticipated more reliably than younger children, but only in the non-English videos.
-In the English videos all children anticipated, especially after hearing questions.
+Third, we saw that older children made anticipatory switches more reliably than younger children, but only in the non-English videos.
+In the English videos, children anticipated well at all ages, especially after hearing questions.
 This interaction between age and language condition suggests that the 5-year-olds were able to leverage anticipatory cues in the non-English videos in a way that 3- and 4-year-olds could not, possibly by shifting more attention to the non-native prosodic or non-verbal cues.
-Prior work on children's turn-structure anticipation proposed that children's turn-end predictions rely primarily on lexicosyntactic structure (and not, e.g., prosody) as they get older \citep{keitel2013}.
+Prior work on children's turn-structure anticipation has proposed that children's turn-end predictions rely primarily on lexicosyntactic structure (and not, e.g., prosody) as they get older \citep{keitel2013}.
 The current results suggest more flexibility in children's predictions; when they do not have access to lexical information, older children and adults are likely to find alternative cues to turn taking behavior.
 
-In Experiment 2 we follow up on these findings, improving on two aspects of the design: First, our language manipulation in this first experiment was too coarse to provide data regarding specific linguistic channels (e.g., prosody vs.
+In Experiment 2 we follow up on these findings, improving on two aspects of the design: first, our language manipulation in this first experiment was too coarse to provide data regarding specific linguistic channels (e.g., prosody vs.
 lexicosyntax).
 In Experiment 2 we compared lexicosyntactic and prosodic cues with phonetically altered speech and used puppets to eliminate non-verbal cues to turn taking.
-Second, we were not able to pinpoint the emergence of anticipatory switching because the youngest age group in our first sample were already able to make anticipatory switches above chance rates.
+Second, we were not able to pinpoint the emergence of anticipatory switching because the youngest age group in our sample was already able to make anticipatory switches at above chance rates.
 In Experiment 2 we explore a wider developmental range.
@@
 % Finally, children showed an effect of gap duration (Table \ref{tab:E1-models}).
-This effect is straightforward: Longer gaps resulted in longer analysis windows, yielding more time for children to make an anticipatory gaze.
+This effect is straightforward: longer gaps resulted in longer analysis windows, yielding more time for children to make an anticipatory gaze.
 
@@
 To tease apart the role of lexical and prosodic information, we phonetically manipulated the speech signal for pitch, syllable duration, and lexical access.
-By testing one- to six-year-olds we hoped to find the developmental onset of turn-predictive gaze.
+By testing 1- to 6-year-olds we hoped to find the developmental onset of turn-predictive gaze.
 We also hoped to measure changes in the relative roles of prosody and lexicosyntax across development.
@@
 We recruited 27 undergraduate adults and 129 children between ages 1;0--6;11 to participate in our experiment.
-We recruited our child participants from Children's Discovery Museum of San Jose, California, targeting approximately 20 children for each of the six one-year age groups (range: 20--23).
+We recruited our child participants from the Children's Discovery Museum of San Jose, California\footnote{We ran Experiment 2 at a local children's museum because it gave us access to children with a more diverse range of ages.}, targeting approximately 20 children for each of the six one-year age groups (range: 20--23).
 All participants were native English speakers, though some parents (N$=$27) reported that their child heard a second (and sometimes third) language at home.
 None of the adult participants reported fluency in a second language.
-We ran Experiment 2 at a local children's museum because it gave us access to children with a more diverse range of ages.
 
@@
 We created four versions of each audio clip: \textit{normal}, \textit{words only}, \textit{prosody only}, and \textit{no speech}.
-That is, one version with a full linguistic signal (\textit{normal}), and three with incomplete linguistic information (hereafter ``limited cue'' conditions).
+That is, one version with a full linguistic signal (\textit{normal}), and three with incomplete linguistic information (hereafter ``partial cue'' conditions).
 The \textit{normal} clips were the unmanipulated, original audio clips.
 
-The \textit{words only} clips were manipulated to have robot-like speech: We flattened the intonation contours to each talker's average pitch (F0) and we reset the duration of every nucleus and coda to each talker's average nucleus and coda duration.\footnote{We excluded hyper-lengthened words like [w\textipa{aU:}] `woooow!'.
+The \textit{words only} clips were manipulated to have robot-like speech: we flattened the intonation contours to each talker's average pitch (F$_{0}$) and we reset the duration of every nucleus and coda to each talker's average nucleus and coda duration.\footnote{We excluded hyper-lengthened words like [w\textipa{aU:}] `woooow!'.
 These were rare in the clips.} We made duration and pitch manipulations using PSOLA resynthesis in Praat \citep{Praat}.
@@
 The \textit{no speech} condition served as a non-linguistic baseline.
-For this condition, we replaced the original clip with multi-talker babble: We overlaid different child-oriented conversations (not including the original one), and then cropped the result to the duration of the original video.
+For this condition, we replaced the original clip with multi-talker babble: we overlaid different child-oriented conversations (not including the original one), and then cropped the result to the duration of the original video.
 Thus, the \textit{no speech} audio clips lacked any linguistic information to upcoming turn boundaries---the only cue to turn taking was the opening and closing of the puppets' mouths.
@@
 We created puppet video recordings to match the manipulated 20-second audio clips.
-The puppets were minimally expressive; the experimenter could only control the opening and closing of their mouths; their head, eyes, arms, and body stayed still.
+The puppets were minimally expressive; so that the puppeteer could only control the opening and closing of their mouths; the puppets' heads, eyes, arms, and bodies stayed still.
 Puppets were positioned looking forward to eliminate shared gaze as a cue to turn structure \citep{thorgrimsson2015}.
@@
 
-We used three pairs of puppets used for the \textit{normal} condition---`red', `blue' and `yellow'---and one pair of puppets for each limited cue condition: ``robots'', ``merpeople'', and ``party-goers'' (Figure 8).
-We randomly assigned half of the conversation topics (`birthday cake', `pets', and `breakfast') to the \textit{normal} condition, and half to the limited cue conditions (`riding bikes', `rainy days', and `the library').
+We used three pairs of puppets used for the \textit{normal} condition---`red', `blue' and `yellow'---and one pair of puppets for each partial cue condition: ``robots'', ``merpeople'', and ``party-goers'' (Figure 8).
+We randomly assigned half of the conversation topics (`birthday cake', `pets', and `breakfast') to the \textit{normal} condition, and half to the partial cue conditions (`riding bikes', `rainy days', and `the library').
 We then created three versions of the experiment, so that each of the six puppet pairs was associated with three different conversation topics across the different versions of the experiment (18 videos in total).
@@
 The duration of turn transitions and the number of speaker changes across videos was variable because the conversations were recorded semi-spontaneously.
-We measured turn transitions from the audio recording of the \textit{normal}, \textit{words only}, and \textit{prosody only} conditions.
-There was no audio from the original conversation in the \textit{no speech} condition videos, so we measured turn transitions from the video recording, using ELAN video editing software \citep{ELAN}.
+We measured turn transitions from the audio signal of the \textit{normal}, \textit{words only}, and \textit{prosody only} conditions.
+There was no audio from the original conversation in the \textit{no speech} condition videos, so we measured turn transitions from puppets' mouth movements in the video signal, using ELAN video annotation software \citep{ELAN}.
 
 There were 85 turn transitions for analysis after excluding transitions longer than 550 msec and shorter than 90 msec.
-The remaining turn transitions had slightly more questions than non-question (N$=$50 and N$=$35, respectively), with transitions distributed somewhat evenly across conditions (keeping in mind that there were three \textit{normal} videos and only one limited cue video for each experiment version): \textit{normal} (N$=$36), \textit{words only} (N$=$13), \textit{prosody only} (N$=$17), and \textit{no speech} (N$=$19).
+The remaining turn transitions had slightly more questions than non-questions (N$=$50 and N$=$35, respectively), with transitions distributed somewhat evenly across conditions (keeping in mind that there were three \textit{normal} videos and only one partial cue video for each experiment version): \textit{normal} (N$=$36), \textit{words only} (N$=$13), \textit{prosody only} (N$=$17), and \textit{no speech} (N$=$19).
 Inter-turn gaps for questions (mean$=$365, median$=$427) were longer than those for non-questions (mean$=$302, median$=$323) on average, but gap duration was overall comparable across conditions: \textit{normal} (mean$=$334, median$=$321), \textit{words only} (mean$=$347, median$=$369), \textit{prosody only} (mean$=$365, median$=$369), and \textit{no words} (mean$=$319, median$=$329).
@@
 Three of the puppet videos had \textit{normal} audio while the other three had \textit{words only}, \textit{prosody only}, and \textit{no speech} audio.
-This experiment required no special instructions so the experimenter immediately began each session with calibration (same as before) and then stimulus presentation.
+This experiment required no special instructions so, as before, the experimenter immediately began each session with calibration and then stimulus presentation.
 The entire experiment took less than five minutes.
-
-\subsubsection{Data preparation and coding}
-We coded each turn transition for its linguistic condition (\textit{normal}, \textit{words only}, \textit{prosody only}, and \textit{no speech}) and transition type (question/non-question)\footnote{We coded \textit{wh-}questions as ``non-questions'' for the \textit{prosody only} videos.
-Polar questions had a final rising prosodic contour, but \textit{wh-}questions did not  \citep{hedberg2010}.} and identified anticipatory gaze switches to the upcoming speaker using the methods from Experiment 1.
-
-\subsection{Results}
-\label{sec:results2}
-
-Participants' pattern of gaze indicated that they performed basic turn tracking across all ages and in all conditions.
-Participants again looked at the screen most of the time during video playback (82\% and 86\% average for children and adults, respectively), primarily looking at the person who was currently speaking (Table 2).
-They tracked the current speaker in every condition---even one-year-olds looked more at the current speaker than at anything else in the three limited cue conditions (40\% for \textit{words only}, 43\% for \textit{prosody only}, and 39\% for \textit{no speech}).
-There was a steady overall increase in looks to the current speaker with age and added linguistic information (Tables \ref{tab:look_e2} and \ref{tab:look_e2b}).
-Looks to the addressee also decreased with age, but the change was minimal.
-Figure \ref{fig:E2-randvsreal} shows participants' anticipatory gaze rates across age, the four language conditions, and transition type.
 
@@
 
+\subsubsection{Data preparation and coding}
+We coded each turn transition for its linguistic condition (\textit{normal}, \textit{words only}, \textit{prosody only}, and \textit{no speech}) and transition type (question/non-question)\footnote{We coded \textit{wh-}questions as ``non-questions'' for the \textit{prosody only} videos.
+Polar questions often have a final rising intonational contour, but \textit{wh-}questions often do not  \citep{hedberg2010}.}, and identified anticipatory gaze switches to the upcoming speaker using the methods from Experiment 1.
+
+\subsection{Results}
+\label{sec:results2}
+
+Participants' pattern of gaze indicated that they performed basic turn tracking across all ages and in all conditions.
+Participants looked at the screen most of the time during video playback (82\% and 86\% average for children and adults, respectively), primarily looking at the person who was currently speaking (Table 2).
+They tracked the current speaker in every condition---even one-year-olds looked more at the current speaker than at anything else in the three partial cue conditions (40\% for \textit{words only}, 43\% for \textit{prosody only}, and 39\% for \textit{no speech}).
+There was a steady overall increase in looks to the current speaker with age and added linguistic information (Tables \ref{tab:look_e2} and \ref{tab:look_e2b}).
+Looks to the addressee also decreased with age, but the change was minimal.
+Figure \ref{fig:E2-randvsreal} shows participants' anticipatory gaze rates across age, the four language conditions, and transition type.
+
 \begin{table}[t]
@@
 We again built separate models for children and adults because effects of age were only pertinent to the children's data.
-The child model included condition (normal/prosody only/words only/no speech; with no speech as the reference level), transition type (question vs.
+The child model included condition (\textit{normal}/\textit{prosody only}/\textit{words only}/\textit{no speech}; with \textit{no speech} as the reference level), transition type (question vs.
 non-question), age (1, 2, 3, 4, 5, 6; numeric), and duration of the inter-turn gap (in seconds) as predictors, with full interactions between language condition, transition type, and age.
@@
 
+Children's anticipatory gaze switches showed an effect of gap duration (\textit{$\beta$}=3.95, \textit{SE}=0.617, \textit{z}=6.401, \textit{p}$<$.001), a two-way interaction of age and language condition (for \textit{prosody only} speech compared to the \textit{no speech} reference level; \textit{$\beta$}=0.393, \textit{SE}=0.189, \textit{z}=2.08, \textit{p}$<$.05), and a three-way interaction of age, transition type, and language condition (for \textit{normal} speech compared to the \textit{no speech} reference level; \textit{$\beta$}=-0.38, \textit{SE}=0.17, \textit{z}=-2.229, \textit{p}$<$.05).
+There were no significant effects of age or transition type alone (Table \ref{sec:models2}), with only a marginal effect of language condition (for \textit{prosody only} compared to the \textit{no speech} reference level; \textit{$\beta$}=-1.607, \textit{SE}=0.867, \textit{z}=-1.85, \textit{p}$=$.064)
+
+Adults' anticipatory gaze switches showed effects of gap duration (\textit{$\beta$}=4.7, \textit{SE}=1.18, \textit{z}=3.978, \textit{p}$<$.001) and language condition (for \textit{normal} speech \textit{$\beta$}=1.203, \textit{SE}=0.536, \textit{z}=2.245, \textit{p}$<$.05 and \textit{words only} speech \textit{$\beta$}=1.561, \textit{SE}=0.709, \textit{z}=2.203, \textit{p}$<$.05  compared to the \textit{no speech} reference level).
+There were no effects of transition type (\textit{$\beta$}=0.437, \textit{SE}=0.585, \textit{z}=0.747, \textit{p}$=$.45).
+
+\subsubsection{Random baseline comparison}
+\label{sec:randbaseline2}
+
+Using the same technique described in Experiment 1 (Section \ref{sec:randbaseline1}), we created and modeled random permutations of participants' anticipatory gaze.
+These analyses revealed that none of the significant predictors from models of the original, turn-related data could be explained by random looking.
+In the children's data, the original model's \textit{z}-values for language condition (\textit{prosody only}), gap duration, the two-way interaction of age and language condition (\textit{prosody only}) and the three-way interaction of age, transition type, and language condition (\textit{normal} speech) were all greater than 95\% of the randomly permuted \textit{z}-values (96.5\%, 100\%, 95.6\%, and 95.5\%, respectively, all \textit{p}$<$.05).
+Similarly, the adults' data showed significant differentiation from the randomly permuted data for all originally significant predictors: gap duration and language condition for \textit{normal} speech and \textit{words only} speech (greater than 100\%, 97.8\%, and 99.1\% of random \textit{z}-values, respectively, all \textit{p}$<$.05).
+
+\begin{minipage}[t]{0.95\linewidth}
 \begin{footnotesize}
@@
     \hline
-                       							& Estimate & Std. Error & z value & Pr($>$$|$z$|$)    \\
+           &  Estimate & Std. Error & \textit{z} value & Pr($>$$|$\textit{z}$|$) \\    
     \hline
@@
     \hspace*{5mm} Type$=$\textit{non-Question} &&&& \\
-
     \hline
@@
     \hline
-                       							&  Estimate & Std. Error & z value & Pr($>$$|$z$|$) \\    
+           &  Estimate & Std. Error & \textit{z} value & Pr($>$$|$\textit{z}$|$) \\    
     \hline
@@
 \end{footnotesize}
-
-Children's anticipatory gaze switches showed an effect of gap duration (\textit{$\beta$}=3.95, \textit{SE}=0.617, \textit{z}=6.401, \textit{p}$<$.001), a two-way interaction of age and language condition (for \textit{prosody only} speech compared to the \textit{no speech} reference level; \textit{$\beta$}=0.393, \textit{SE}=0.189, \textit{z}=2.08, \textit{p}$<$.05), and a three-way interaction of age, transition type, and language condition (for \textit{normal} speech compared to the \textit{no speech} reference level; \textit{$\beta$}=-0.38, \textit{SE}=0.17, \textit{z}=-2.229, \textit{p}$<$.05).
-There were no significant effects of age or transition type alone (Table \ref{sec:models2}), with only a marginal effect of language condition (for \textit{prosody only} compared to the \textit{no speech} reference level; \textit{$\beta$}=-1.607, \textit{SE}=0.867, \textit{z}=-1.85, \textit{p}$=$.064)
-
-Adults' anticipatory gaze switches showed effects of gap duration (\textit{$\beta$}=4.7, \textit{SE}=1.18, \textit{z}=3.978, \textit{p}$<$.001) and language condition (for \textit{normal} speech \textit{$\beta$}=1.203, \textit{SE}=0.536, \textit{z}=2.245, \textit{p}$<$.05 and \textit{words only} speech \textit{$\beta$}=1.561, \textit{SE}=0.709, \textit{z}=2.203, \textit{p}$<$.05  compared to the \textit{no speech} reference level).
-There were no effects of transition type (\textit{$\beta$}=0.437, \textit{SE}=0.585, \textit{z}=0.747, \textit{p}$=$.45).
-
-\subsubsection{Random baseline comparison}
-\label{sec:randbaseline2}
-
-Using the same technique described in Experiment 1 (Section \ref{sec:randbaseline1}), we created and modeled random permutations of participants' anticipatory gaze.
-These analyses revealed that none of the significant predictors from models of the original, turn-related data could be explained by random looking.
-In the children's data, the original model's \textit{z}-values for language condition (\textit{prosody only}), gap duration, the two-way interaction of age and language condition (\textit{prosody only}) and the three-way interaction of age, transition type, and language condition (\textit{normal} speech) were all greater than 95\% of the randomly permuted \textit{z}-values (96.5\%, 100\%, 95.6\%, and 95.5\%, respectively).
-Similarly, the adults' data showed significant differentiation from the randomly permuted data for all originally significant predictors: gap duration and language condition for \textit{normal} speech and words-only speech (greater than 100\%, 97.8\%, and 99.1\% of random \textit{z}-values, respectively).
+\end{minipage}
 
@@
 
-Our main goal in extending the age range to 1- and 2-year-olds in Experiment 2 was to find the age of emergence for spontaneous turn structure predictions in our paradigm.
-As in Experiment 1, we used two-tailed \textit{t}-tests to compare children's real gaze rates to the random baseline in the \textit{normal} speech condition, in which the speech stimulus is most like what children hear every day.
+Our main goal in extending the age range to 1- and 2-year-olds in Experiment 2 was to find the age of emergence for spontaneous predictions about upcoming turn structure.
+As in Experiment 1, we used two-tailed \textit{t}-tests to compare children's real gaze rates to the random baseline rates in the \textit{normal} speech condition, in which the speech stimulus is most like what children hear every day.
 We tested real gaze rates against baseline for three age groups: ages one, two, and three.
@@
 
-Regression models for the children's data also revealed two significant interactions with age.
+The regression models for the children's data also revealed two significant interactions with age.
 The first was a significant interaction of age and language condition (for \textit{prosody only} compared to the \textit{no speech} reference level), suggesting a different age effect between the two linguistic conditions.
@@
 \textit{prosody only}) within each random permutation of the data, making pairwise comparisons between the six age groups.
-These tests revealed that children's anticipation in the \textit{prosody only} condition significantly improved at ages five and six (difference scores greater than 95\% of the random data scores (\textit{p}$<$.05).
+These tests revealed that children's anticipation in the \textit{prosody only} condition significantly improved at ages five and six (with difference scores greater than 95\% of the random data scores; \textit{p}$<$.05).
+See Figure \ref{fig:E2-lgageinteraction} for these \textit{prosody only} difference score distributions.
 
@@
 We again created pairwise comparisons of the average difference scores for the transition type-language condition interaction across age groups in each random permutation of the data, finding that the effect of transition type in the \textit{normal} speech condition became larger with age, with significant improvements by age 4 over ages 1 and 2 (99.9\% and 98.86\%, respectively), by age 5 over age 4 (97.54\%), and by age 6 over ages 1, 2, and 5 (99.5\%, 97.36\%, and 95.04\%), all significantly different from chance (\textit{p}$<$.05).
+See Figure \ref{fig:E2-lgagetypeinteraction} for these \textit{normal} speech difference score distributions.
 
@@
 The core aims of Experiment 2 were to gain better traction on the individual roles of prosody and lexicosyntax in children's turn predictions, and to find the age of emergence for spontaneous turn anticipation.
-Taken together, our results replicate the findings from Experiment 1: Participants make more anticipatory switches when they have access to lexical information and, when they do, tend to make more anticipatory switches for questions compared to non-questions.
-
-As in Experiment 1, with normal speech, children and adults spontaneously tracked the turn structure of the conversations, making anticipatory gaze switches at an above-chance rate across all ages.
+Many of our results replicate the findings from Experiment 1: participants made more anticipatory switches when they had access to lexical information and, when they did, tended to make more anticipatory switches for questions compared to non-questions.
+
+As in Experiment 1, children and adults spontaneously tracked the turn structure of the conversations, making anticipatory gaze switches at above-chance rates across all ages when listening to natural speech.
 They also made far more anticipations for questions than for non-question turns---at least for children two years olds and older.
-But these effects were different for the two conditions with partial linguistic information: \textit{prosody-only} and \textit{words-only}.
-In the \textit{prosody-only} condition, performance was low for younger children and increased significantly with age (especially for questions).
-In the \textit{words-only} condition, children age two and older showed robust anticipatory switching for questions (much like in normal speech), but never rose above chance for non-question turns.
+But these effects were different for the two conditions with partial linguistic information: \textit{prosody only} and \textit{words only}.
+In the \textit{prosody only} condition, performance was low for younger children and increased significantly with age (especially for questions).
+In the \textit{words only} condition, children age two and older showed robust anticipatory switching for questions (much like in \textit{normal} speech), but never rose above chance for non-question turns.
 These findings do not therefore support an early role for prosody in children's spontaneous turn structure predictions.
-On the contrary, children's predictions appeared to stem more from lexical information, possibly from lexicosyntactic cues to questionhood.
+There is also no evidence that lexical information is sufficient on its own to support children's anticipatory switching.
 
@@
 
-Children begin to develop conversational turn-taking skills long before their first words emerge\citep{bateson1975, hilbrink2015, jaffe2001, snow1977}.
+Children begin to develop conversational turn-taking skills long before their first words emerge \citep{bateson1975, hilbrink2015, jaffe2001, snow1977}.
 As they acquire language, they also acquire the information needed to make accurate predictions about upcoming turn structure.
@@
 Prior work with adults has found a consistent role for lexicosyntax in predicting upcoming turn structure \citep{de-ruiter2006, magyari2012}, whereas the role of prosody still under debate \citep{duncan1972, ford1996, torreira2015}.
-Knowing that children comprehend more about prosody than lexicosyntax early on (see \ref{sec:intro}; also see \citealp{speer2009} for a review), we thought it possible that young children would instead show an advantage for prosody in their predictions about turn structure in conversation.
+Knowing that children comprehend more about prosody than lexicosyntax early on (Section \ref{sec:intro}; also see \citealp{speer2009} for a review), we thought it possible that young children would instead show an advantage for prosody in their predictions about turn structure in conversation.
 Our results suggest that, on the contrary, exclusively presenting prosodic information to children limits their spontaneous predictions about upcoming turn structure until age five.
 
-Perhaps surprisingly, we also found no evidence that lexical information alone is equivalent to the full linguistic signal in driving children's predictions, as has been shown previously for adults \citep{magyari2012, de-ruiter2006} and is replicated with adult participants in the current study.
+Perhaps surprisingly, we also found no evidence that lexical information alone is equivalent to the full linguistic signal in driving children's predictions, as has been shown previously for adults \citep{magyari2012, de-ruiter2006} and as is replicated with adult participants in the current study.
 That said, our findings point more toward early lexical effects in children's turn anticipations than prosodic ones: children's performance in both experiments was consistently better when they had access to lexical information, especially after question turns.
-And even though the \textit{words only} condition in Experiment 2 was not itself significantly different from the baseline \textit{no speech} condition, children's anticipations trended toward above-chance rates.
+And although the \textit{words only} condition in Experiment 2 was not significantly different from the baseline \textit{no speech} condition, children's anticipations trended toward above-chance rates.
 
@@
 There may be something informative about combined prosodic and lexical cues to questionhood that helps to boost children's anticipations before they can use these cues separately.
-Even in adults, Torreira and colleagues (\citeyear{torreira2015}) showed that the trade-off in informativity between lexical and prosodic cues is more subtle in semi-natural (spliced) speech.
+Even in adults, Torreira and colleagues (\citeyear{torreira2015}) showed that the trade-off in informativity between lexical and prosodic cues is more subtle in semi-natural speech.
 The present findings are the first to show evidence of a similar effect developmentally.
@@
 
-In both experiments, anticipatory looking was primarily driven by question transitions, a pattern that had not been previously reported in other anticipatory gaze studies, on children or adults (Keitel et al., 2013; Hirvenkari, 2013; Tice and Henetz, \citeyear{TiceHenetz11}).
+In both experiments, anticipatory looking was primarily driven by question transitions, a pattern that has not been previously reported in other anticipatory gaze studies, on children or adults (Keitel et al., 2013; Hirvenkari, 2013; Tice and Henetz, \citeyear{TiceHenetz11}).
 Questions make an upcoming speaker switch immediately relevant, helping the listener to predict with high certainty what will happen next (i.e., an answer from the addressee), and are often easily identifiable by overt prosodic and lexicosyntactic cues.
@@
 Children may have therefore had an easier time picking out and interpreting lexical cues to questionhood.
-The question effect showed its first significant gains between ages three and four in the \textit{normal} speech condition of Experiment 2, by which time children frequently hear and use a variety of polar and \textit{wh}-questions \citep{clark2009}.
+The question effect showed its first significant gains between ages three and four in the \textit{normal} speech condition of Experiment 2 (Figure \ref{fig:E2-lgagetypeinteraction}), by which time children frequently hear and use a variety of polar and \textit{wh}-questions \citep{clark2009}.
 Furthermore, while lexicosyntactic question cues were available on every instance of \textit{wh}- and \textit{yes/no} questions in our stimuli, prosodic question cues were only salient on \textit{yes/no} questions.
-Even still, the mapping of prosodic contour to speech act (e.g., high final rises for polar questions) is far from one-to-one, leaving substantial room for uncertainty in prosodic contour interpretation.
-
-Prior work on children's acquisition of questions indicates that they may already have some knowledge of question-answer sequences by the time they begin to speak: Questions make up approximately one third of the utterances children hear, before and after the onset of speech, and even into their preschool years, though the types and complexity of questions change throughout development \citep{casillas2016, fitneva2012, henning2005, shatz1979}.\footnote{There is substantial variation question frequency by individual and socioeconomic class \citep{hart1992}.} For the first few years, many of the questions directed to children are ``test'' questions---questions that the caregiver already has the answer to (e.g., ``What does a cat say?''), but this changes as children get older.
+Finally, the mapping of prosodic contour to speech act (e.g., high final rises for polar questions) is far from one-to-one, leaving substantial room for uncertainty in prosodic contour interpretation in general.
+
+Prior work on children's acquisition of questions indicates that they may already have some knowledge of question-answer sequences by the time they begin to speak: questions make up approximately one third of the utterances children hear, before and after the onset of speech, and even into their preschool years, though the types and complexity of questions change throughout development \citep{casillas2016, fitneva2012, henning2005, shatz1979}.\footnote{There is substantial variation question frequency by individual and socioeconomic class \citep{hart1992}.} For the first few years, many of the questions directed to children are ``test'' questions---questions that the caregiver already has the answer to (e.g., ``What does a cat say?''), but this changes as children get older.
 Questions help caregivers to get their young children's attention and to ensure that information is in common ground, even if the responses are non-verbal or infelicitous \citep{bruner1985, fitneva2012, snow1977}.
@@
 
-Two important questions for future work are then: (1) How does children's ability to monitor for questions in conversation relate to their prior experience with questions?
+Two important questions for future work are then: (1) how does children's ability to monitor for questions in conversation relate to their prior experience with questions?
 and (2) what is it about questions that makes children and adults more likely to anticipatorily switch their gaze to addressees?
@@
 
-More broadly, our results suggest that participants' \textit{spontaneous} predictions, at least while viewing third-party conversation, are driven by what lies \textit{beyond} the end of the current turn---not by the upcoming end of the turn itself, as has been often assumed \citep{torreira2015, keitel2013, magyari2012, de-ruiter2006}.
+More broadly, our results suggest that participants' spontaneous predictions, at least while viewing third-party conversation, are driven by what lies \textit{beyond} the end of the current turn---not by the upcoming end of the turn itself, as has been focused on in prior work \citep{torreira2015, keitel2013, magyari2012, de-ruiter2006}.
 In future work, it will be crucial to measure prediction from a first-person perspective to resolve this apparent contradiction (see also \citealp{holler2015}).
 
-% There is at least one clear hypothesis that can bridge these apparently conflicting results: Listeners in spontaneous, first-person conversation may use multiple strategies in making predictions about upcoming turn structure, using more passive prediction to detect upcoming speaker transition (e.g., questions), and then switching into precise turn-end prediction mode (\`{a} la \citealp{de-ruiter2006}) when necessary.
+% There is at least one clear hypothesis that can bridge these apparently conflicting results: listeners in spontaneous, first-person conversation may use multiple strategies in making predictions about upcoming turn structure, using more passive prediction to detect upcoming speaker transition (e.g., questions), and then switching into precise turn-end prediction mode (\`{a} la \citealp{de-ruiter2006}) when necessary.
 A flexible prediction system like this one allows listeners to continuously monitor ongoing conversation at a low cost while still managing to plan their responses and come in quickly when needed.
@@
 One of the core aims of our study was to test whether children show an early competence for turn taking, as is proposed by studies of spontaneous mother-infant proto-conversation and theories about the mechanisms underlying human interaction in general \citep{hilbrink2015, levinson2006}.
-We found evidence that young children make spontaneous predictions about upcoming turn structure: definitely at age two and even marginally at age one.
+We found evidence that young children make spontaneous predictions about upcoming turn structure: definitely at age two and marginally at age one.
 
@@
 
-To be clear, young children's ``above chance'' performance was often still far from adult-like predictive behavior---even at age six, children in our studies did not show adult-like competency in their predictions.
+To be clear, young children's ``above chance'' performance was often still far from adult-like predictive behavior---children at ages one and two were still very close to chance in their anticipations and, even at age six, children were not fully adult-like in their predictions.
 This may indicate that young children rely more on non-verbal cues in anticipating turn transitions or, alternatively, that adults are better at flexibly adapting to the turn-relevant cues present at any moment.
 
-% This added flexibility might emerge around age 5, when we see the first adult-like effects in the children's data: In Experiment 1, 5-year-olds using non-verbal or non-native prosodic cues to make predictions in the non-English videos comparable to their predictions for English.
+% This added flexibility might emerge around age 5, when we see the first adult-like effects in the children's data: in Experiment 1, 5-year-olds using non-verbal or non-native prosodic cues to make predictions in the non-English videos comparable to their predictions for English.
 In Experiment 2, 5- and 6-year-olds showing significant advances over younger children in their anticipations for the \textit{prosody only} condition.
@@
 low parent-infant interaction styles) feed into this early predictive skill.
-This finding also bridges  prior work showing a predisposition for turn taking in infancy (e.g., \citealp{hilbrink2015}) but late acquisition of adult-like competence for turn-taking in children's conversations \citep{casillas2016, garvey1984, ervin-tripp1979}.
+It also bridges  prior work showing a predisposition for turn taking in infancy (e.g., \citealp{bateson1975, hilbrink2015, jaffe2001, snow1977}) but late acquisition of adult-like competence for turn-taking in children's conversations \citep{casillas2016, garvey1984, garvey1981, ervin-tripp1979}.
 
@@
 
-There are at least two major limitations to our work: Speech naturalness and participant role.
+There are at least two major limitations to our work: speech naturalness and participant role.
 Following prior work \citep{de-ruiter2006, keitel2013}, we used phonetically manipulated speech in Experiment 2.
@@
 Many prior studies have used phonetically-altered speech with infants and young children \citep[cf.][]{jusczyk2000}, but almost none of them have done so in a conversational context.
-Future work could instead carefully script or cross-splice sub-parts of turns to control for the presence or absence of linguistic cues for turn transition (see, e.g., \citealp{torreira2015}).
+Future work could instead carefully script or cross-splice sub-parts of turns to control for the presence of linguistic cues for turn transition (see, e.g., \citealp{torreira2015}).
 
 The prediction measure used in our studies is based on an observer's view of third-party conversation but, because participants' role in the interaction could affect their online predictions about turn taking, an better measure would instead capture first-person behavior.
-First-person measures of spontaneous turn prediction will be key to revealing how participants distribute their attention over linguistic and non-verbal cues while taking part in everyday interaction, the implications of which relate to theories of online language processing for both language learning and everyday talk.
+First-person measures of spontaneous turn prediction will be key to revealing how participants distribute their attention over linguistic and non-linguistic cues while taking part in everyday interaction, the implications of which relate to theories of online language processing for both language learning and everyday talk.
 
@@
 Adults use linguistic information to accurately predict turn structure in conversation, which facilitates their online comprehension and allows them to respond relevantly and on time.
-The present study offers new findings regarding the role of speech acts and linguistic processing in online turn prediction, and has given evidence that turn prediction emerges by age two is not integrated with linguistic cues until much later.
+The present study offers new findings regarding the role of speech acts and linguistic processing in online turn prediction, and has given evidence that turn prediction emerges by age two, but is not integrated with linguistic cues until much later.
 Using language to make predictions about upcoming interactive content takes time and, for both children and adults, is primarily driven by participants' orientation to what will happen beyond the end of the current turn.
@@
 This work was supported by an ERC Advanced Grant to Stephen C.
-Levinson (269484-INTERACT), NSF graduate research and dissertation improvement fellowships to the first author, and a Merck Foundation fellowship to the second author.
+Levinson (269484-INTERACT), an NSF graduate research and dissertation improvement fellowship to the first author, and a Merck Foundation fellowship to the second author.
 Earlier versions of these data and analyses were presented to conference audiences \citep{casillas2012, casillas2013}.
@@
 Clark, and the First Language Acquisition group at Radboud University for their feedback on earlier versions of this work.
-The analysis code and raw data for this project can be found on GitHub at https://github.com/langcog/turn\_taking/.
+The analysis code for this project can be found on GitHub at https://github.com/langcog/turn\_taking/.
 
@@
 \bibliography{anticip}
+\clearpage
 
@@
 We also made sure that the properties of each turn transition stayed constant across permutations.
-So, while ``transition window A'' might start 2 seconds into Random Permutation 1 and 17 seconds into Random Permutation 2, it maintains the same prior speaker identity, transition type, gap duration, language condition, etc., across both permutations.
+So, while ``transition window A'' might start 2 seconds into Random Permutation 1 and 17 seconds into Random Permutation 2, it maintained the same prior speaker identity, transition type, gap duration, language condition, etc., across both permutations.
 
@@
 We used the absolute value of all \textit{z}-values to conduct a two-tailed test.
-If the original effect of a predictor exceeded 95\% of the random model effects for that same predictor, we deemed that predictor's effect to be significantly different from the random baseline.
+If the original effect of a predictor exceeded 95\% of the random model effects for that same predictor, we deemed that predictor's effect to be significantly different from the random baseline (i.e., \textit{p}$<$.05).
 
@@
 We used this procedure to derive the random-baseline comparison values in the main text (above).
-However, we ran into two issues along the way: First, we had to report \textit{z}-values rather than beta estimates.
-Second, we had to exclude a substantial portion of the models, especially in Experiment 2 because of convergence issues.
+However, we ran into two issues along the way: first, we had to report \textit{z}-values rather than beta estimates.
+Second, we had to exclude a substantial portion of the models, especially in Experiment 2 because of model non-convergence.
 We address each of these issues below.
 
-\subsection{Beta, standard error, and z estimates in the data}
+\subsection{Beta, standard error, and z estimates}
 We reported \textit{z}-values in the main text rather than beta estimates because the standard errors in the randomly permuted data models were much higher than for the original data.
@@
 \subsection{Non-convergent models}
-In comparing the real and randomly permuted datasets, we excluded the output of random-permutation models that gave convergence warnings in order to remove erratic model estimates from our analyses.
+In comparing the real and randomly permuted datasets, we excluded the output of random-permutation models that gave convergence warnings to remove erratic model estimates from our analyses.
 Non-convergent models made up 22.4--24.4\% of the random permutation models in Experiment 1 and 69--70\% of the random permutation models in Experiment 2.
@@
 Although many of the non-converging models show estimates within range of the converging models (e.g., with a mean difference of only 0.09 in median \textit{z}-value across predictors), they also show many radically outlying estimates (e.g., showing a mean difference of 237.3 in mean \textit{z}-value across predictors).
-Similar patterns were obtained in the non-converging models for Experiment 2 and persisted even when we changed optimizers.
-
-We suspect that the issue derives from data sparsity in some of the random runs.
-This problem is known to occur when there are limited numbers of binary observations in each of design matrix's many bins \citep{allison2004}.
-We could instead use zero-inflated poisson (ZIP) or negative binomial regression models to allow for overdispersion in our data \citep{allison2012}.
+Similar patterns were obtained in the non-converging models for Experiment 2 and persisted even when we tried other optimizers.
+
+We suspect that the issue derives from data sparsity in some of the random permutations.
+This problem is known to occur when there are limited numbers of binary observations in each of a design matrix's bins \citep{allison2004}.
+We could instead use zero-inflated poisson or negative binomial regression models to allow for overdispersion in our data \citep{allison2012}.
 However, these would give us baselines for the normal, convergent model, which is not the aim of this analysis.
@@
   \begin{tabular}{lllllllllll}
-    Variable & Mean$_{C}$ & Mean$_{NC}$ & Median$_{C}$ & Median$_{NC}$ & \textit{SD}$_{C}$ & \textit{SD}$_{NC}$ & Min$_{C}$ & Min$_{NC}$ & Max$_{C}$ & Max$_{NC}$ \\ 
+    & Mean$_{C}$ & Mean$_{NC}$ & Median$_{C}$ & Median$_{NC}$ & \textit{SD}$_{C}$ & \textit{SD}$_{NC}$ & Min$_{C}$ & Min$_{NC}$ & Max$_{C}$ & Max$_{NC}$ \\ 
     \hline
@@
   \end{tabular}
-  \caption{Estimated \textit{z}-values for each predictor in converging (\textit{C}) and non-converging (\textit{NC}) adult and child models from Experiment 1. }
+  \caption{Estimated \textit{z}-values for each predictor in converging (\textit{C}) and non-converging (\textit{NC}) child and adult models from Experiment 1. }
 \label{tab:nonconv_e1}
@@
 non-English).
-To explore these effects in more depth, we recorded the average difference score for the predictor that interacted with age (e.g., linguistic condition) for each participant (e.g., English minus non-English anticipatory switches), using these values to compute an average difference score over participants in each age group (e.g., age 3, 4, and 5) within each random permutation.
+To explore these effects in more depth, we recorded the average difference score for the predictor that interacted with age for each participant (e.g., English minus non-English anticipatory switches), using these values to compute an average difference score over participants in each age group (e.g., age 3, 4, and 5) within each random permutation.
 That averaging process produces 5,000 baseline-derived difference scores for each age group.
 
-We then completed pairwise age comparisons of these difference scores (e.g., the linguistic condition effect in 3-year-olds vs.
+We then made pairwise age comparisons of these difference scores (e.g., the linguistic condition effect in 3-year-olds vs.
 4-year-olds), computing the percent of random-permutation difference scores exceeded by the real-data difference score.
-If the real-data difference score exceeded 95\% of the random-data age difference scores, we deemed it significantly different from chance---a significant difference in age between the two groups being compared (e.g., a significant difference between ages three and four in the effect of linguistic condition).
-
+If the real-data difference score exceeded 95\% of the random-data age difference scores, we deemed it to be an age effect significantly different from chance---e.g., a significant difference between ages three and four in the effect of linguistic condition.
 This procedure is essentially a two-tailed \textit{t}-test, adapted for use with the randomly permuted baseline data.
@@
 In each of the plots below, the dot represents the real data value for the effect being shown.
-The 5,000 randomly permuted data effect sizes are shown in the distribution.
+The effect sizes from the 5,000 randomly permuted data sets are shown in the distribution.
 The percentage displayed is the percentage of random permutation values exceeded by the original data value (taking the absolute value of all data points for a two-tailed test).
@@
 \begin{center}
-\textbf{Experiment 2: Age and the prosody only condition}\par\medskip
+\textbf{Experiment 2: Age and the \textit{prosody only} condition}\par\medskip
 \includegraphics[width=0.95\textwidth]{figures/E2-child-randvsreal-ttest-muffledages.png}
@@
 \begin{center}
-\textbf{Experiment 2: Age, transition type, and normal speech}\par\medskip
+\textbf{Experiment 2: Age, transition type, and \textit{normal} speech}\par\medskip
 \includegraphics[width=0.95\textwidth]{figures/E2-child-randvsreal-ttest-normaltypesages.png}
@@
 \end{center}
-\caption{Proportion of participants looking at the current speaker: hypothetical boredom-driven data (gray dots) versus actual data from Experiment 2 (black dots).
-Vertical bars indicate standard error in the experiment data.}
+\caption{Proportion of participants (hypothetical boredom-driven$=$gray; actual Experiment 2$=$black) looking at the current speaker, split by turn duration.
+Vertical bars indicate standard error in the experimental data.}
 \label{fig:boredomhypothesis}
@@
 If children's switches away from the current speaker were driven by boredom, they would switch away equally quickly on long and short turns.
-However, as turn duration increased, children in the experiment looked at the current speaker for a longer period, suggesting that they were not switching away at a constant rate.
-We can see this pattern most clearly by looking at the time when 50\% of the children had switched away from the current (indicated by the horizontal line in Figure \ref{fig:boredomhypothesis}).
-Children's gaze crossed this 50\% threshold at 2.0, 2.9, and 3.6 seconds after the start of speech for turns with durations of 1--2, 2--3, and 3--4 seconds, respectively.
+However, as turn duration increased, children in the experiment looked at the current speaker for a longer period before looking away, suggesting that they were not switching away at a constant rate.
+We can see this pattern most clearly in the time at which 50\% of the children had switched away from the current (indicated by the horizontal line in Figure \ref{fig:boredomhypothesis}): children's gaze crossed this 50\% threshold at 2.0, 2.9, and 3.6 seconds after the start of speech for turns with durations of 1--2, 2--3, and 3--4 seconds, respectively.
 In contrast, the hypothetical boredom-driven children always hit the 50\% threshold at 2.5 seconds after the start of speech.
@@
 \setcounter{figure}{0}  
-The design for Experiment 2 does not fully cross puppet pair (e.g., robots, blue puppets) with linguistic condition (e.g., ``words only'' and ``no speech'').
-Even though each puppet pair is associated with different conversation clips across children (e.g.
-robots talking about kitties, birthday parties, and pancakes), the robot puppets themselves were exclusively associated with the ``words only'' condition.
-Similarly, merpeople were exclusively associated with ``prosody only'' speech, and the puppets wearing dress clothes were exclusively associated with the ``no speech'' condition.
-We designed the experiment this way to increase its pragmatic felicity for older children (i.e., robots make robot sounds, merpeople's voices are muffled under the water, the fancy-clothed puppets are in a `party' room with many other voices).
-There is therefore a possible confound between linguistic condition and puppet pair; for example, children could have made fewer anticipatory switches in the \textit{prosody only} condition because the puppets were less interesting.
+The design for Experiment 2 does not fully cross puppet pair (e.g., robots, blue puppets) with linguistic condition (e.g., \textit{words only} and \textit{no speech}).
+Even though each puppet pair is associated with different conversation clips across children (e.g., robots talking about kitties, birthday parties, and pancakes), the robot puppets themselves were exclusively associated with the \textit{words only} condition.
+Similarly, merpeople were exclusively associated with \textit{prosody only} speech, and the puppets wearing dress clothes were exclusively associated with the \textit{no speech} condition.
+We designed the experiment this way to increase its pragmatic felicity for older children (i.e., robots make robot sounds, merpeople's voices are muffled under the water, the party-going puppets are in a `party' room with many other voices).
+There is therefore a confound between linguistic condition and puppet pair; for example, children could have made fewer anticipatory switches in the \textit{prosody only} condition because the puppets were less interesting.
 To test whether puppet pair drove the condition-based differences found in Experiment 2, we ran a short follow-up study.
@@
 
-\noindent We recruited 30 children between ages 3;0--5;11 from Children's Discovery Museum of San Jose, California to participate in our experiment.
+\noindent We recruited 30 children between ages 3;0 and 5;11 from the Children's Discovery Museum of San Jose, California to participate in our experiment.
 All participants were native English speakers.
@@
 We created 6 short videos from the stimulus recordings made for Experiment 2.
-Each video featured a puppet pair (red/blue/yellow/robot/ merpeople/dressy; Figure \ref{fig:puppets}).
-Puppets in all six videos performed the exact same conversation recording (`birthday party'; Experiment 2) with normal, unmanipulated speech; this experiment therefore holds all things constant except for the appearance of the puppets.
+Each video featured a puppet pair (red/blue/yellow/robot/ merpeople/party-goer; Figure \ref{fig:puppets}).
+Puppets in all six videos performed the exact same conversation recording (`birthday party'; Experiment 2) with normal, unmanipulated speech; this experiment therefore holds all things constant across stimuli except for the appearance of the puppets.
 
@@
 Each participant was randomly assigned to watch only one of the six puppet videos.
-Five children watched each video in total.
+Five children watched each video.
 As in Experiment 2, the experimenter immediately began each session with calibration and then stimulus presentation because no special instructions were required.
@@
 
-\noindent We modeled children's anticipatory switches (yes or no at each transition) with mixed effects logistic regression, including puppet pair (robots/mer- people/fancy dress/other-3) as a fixed effect and participant and turn transition as random effects.
-We grouped the red, blue, and yellow puppets together because they collectively represent the puppets used in the \textit{normal} speech condition and this follow-up experiment is meant to test whether the condition-based differences from Experiment 2 arose from the puppets used in each condition.
+\noindent We modeled children's anticipatory switches (yes or no at each transition) with mixed effects logistic regression, including puppet pair (robots/mer- people/party-goers/other-3) as a fixed effect and participant and turn transition as random effects.
+We grouped the red, blue, and yellow puppets together because they collectively represented the puppets used in the \textit{normal} speech condition---this follow-up experiment is meant to test whether the condition-based differences from Experiment 2 arose from the puppets used in each condition.
 
@@
 
-In four versions of this model we systematically varied the reference level of the puppet pair to check for any cross-condition differences.
-We found no significant effects of puppet pair on switching rate (all \textit{p} $>$ 0.25; Table \ref{tab:control_exp}).
-
 \begin{table}
@@
   \begin{tabular}{llcccc}
-    \hline
-           &  Estimate & Std. Error & z value & Pr($>$$|$z$|$) \\    
-    \hline
-		 \footnotesize{\textit{Intercept: normal-condition puppets}} &&&& \\
+%    \hline
+           &  Estimate & Std. Error & \textit{z} value & Pr($>$$|$\textit{z}$|$) \\    
+    \hline
+		 \footnotesize{\textit{Reference level: normal-condition puppets}} &&&& \\
     \hline
@@
      Puppets$=$\textit{robot} 		& -0.07104	& 0.65321	& -0.109		& 0.913 \\ 
-     Puppets$=$\textit{dressy} 		& -0.78206	& 0.68699	& -1.138		& 0.255 \\ 
-    \hline
-		 \footnotesize{\textit{Intercept: mermaid puppets}} &&&& \\
+     Puppets$=$\textit{party} 		& -0.78206	& 0.68699	& -1.138		& 0.255 \\ 
+    \hline
+		 \footnotesize{\textit{Reference level: mermaid puppets}} &&&& \\
     \hline
@@
      Puppets$=$\textit{robot} 		& 0.004763	& 0.80096	& 0.006		& 0.995 \\ 
-     Puppets$=$\textit{dressy} 		& -0.70626	& 0.82742	& -0.854		& 0.393 \\ 
-    \hline
-		 \footnotesize{\textit{Intercept: robot puppets}} &&&& \\
+     Puppets$=$\textit{party} 		& -0.70626	& 0.82742	& -0.854		& 0.393 \\ 
+    \hline
+		 \footnotesize{\textit{Reference level: robot puppets}} &&&& \\
     \hline
      (Intercept) 								& -0.21895	& 0.56565	& -0.387		& 0.699 \\ 
-     Puppets$=$\textit{dressy} 		& -0.71102	& 0.82657	& -0.860		& 0.390 \\ 
-    \hline
-		 \footnotesize{\textit{Intercept: dressy puppets}} &&&& \\
+     Puppets$=$\textit{party} 		& -0.71102	& 0.82657	& -0.860		& 0.390 \\ 
+    \hline
+		 \footnotesize{\textit{Reference level: party-goer puppets}} &&&& \\
     \hline
@@
 
+In four versions of this model, we systematically varied the reference level of the puppet pair to check for any cross-condition differences.
+We found no significant effects of puppet pair on switching rate (all \textit{p}$>$0.25; Table \ref{tab:control_exp}).
+
 We take this finding as evidence that our decision to not fully cross puppet pairs and linguistic conditions in Experiment 2 was unlikely to have strongly affected children's anticipatory gaze rates above and beyond the intended effects of linguistic condition.

### 2016-01-29 paper/Letters/ResponseLetter-Dec2015.tex
Updated paper changes and code used in submission Jan 2016.

@@
 \def\Address{Wundtlaan 1} % Your address
-\def\CityZip{6525XD Nijmegen, NL} % Your city, zip code, country, etc
+\def\CityZip{6525 XD Nijmegen, NL} % Your city, zip code, country, etc
 \def\Email{E-mail: marisa.casillas@mpi.nl} % Your email address
@@
 
-\opening{Dear Profs. Dr. Pickering and Dr. Gerrig,}
+\opening{Dear Profs.  Dr. Gerrig and Dr. Pickering,}
 
@@
 
-\noindent Please accept our resubmission of our manuscript on ``The development of children's ability to track and predict turn structure in conversation'' to the \textit{Journal of Memory and Language}.
-We apologize for our delay in getting the manuscript resubmitted.
-It took quite a long time to implement one of the suggestions from Reviewer 3, mostly because I had to learn quite a bit of new technical skill to carry out a suggested analysis out on our computation-heavy data set.
-Nevertheless, I believe that this analysis has substantially strengthened the paper as a whole, so I appreciate your patience.
+\noindent I hereby resubmit a manuscript on ``The development of children's ability to track and predict turn structure in conversation'' to the \textit{Journal of Memory and Language}.
+I apologize for the delay in getting the manuscript resubmitted.
+It took quite a long time to implement one of the suggestions from Reviewer 3, mostly because I had to acquire some new technical skills in order to carry out his/her suggested analysis on our computation-heavy data set.
+Nevertheless, I believe that the added analysis has substantially strengthened the paper as a whole, so I appreciate your patience very much.
 
-\noindent In addition to the new analysis, we have also included two other follow-up analyses in the Appendix, both inspired by the other reviewers' comments.
-And in other revisions, we have tried hard to make the text more succinct and the results and discussion sections more distinct and more straightforward, following requests from all three reviewers.
+\noindent Please note that, because of a miscommunication, I prematurely resubmitted the manuscript a few weeks ago (without my co-author's final sign-off).
+I immediately withdrew the manuscript when our miscommunication came to light and we have since then made minor improvements to the paper.
+With my co-author's final approval, I am ready to return the paper to JML.
+However, now my only option is to resubmit the manuscript as ``new''.
+If it is at all possible, we would request that the paper go back to the original reviewers, if only because we did quite a lot of work to address their specific (and substantial) requests.
+In case this is possible, I have included our responses to the reviewers below.
 
-\noindent We hope that, with these changes, we have greatly improved the paper's clarity.
+\noindent For example, in addition to the new random permutation analysis suggested by Reviewer 3, we have also included two further follow-up analyses in the Appendix, each inspired by the other reviewers' comments.
+We have also tried to make the text much more succinct, with distinct and straightforward results and discussion sections, following requests from all three reviewers.
+
+\noindent We hope that, with these changes, we have greatly improved the paper's quality.
 Please find a description of specific comments addressed below.
@@
 \begin{tabular}{ll}
-\textit{ID:} & JML-14-235 \\
+\textit{ID:} & JML-14-235 (from the previous submission) \\
 \textit{Title:} & The development of children's ability to track and \\
@@
 \noindent We thank all three reviewers for their thorough and helpful comments on the previously submitted manuscript.
-We have tried to address all of the reviewers' comments, within the manuscript as well as here in the response letter.
+We have tried to address all of their comments, in the manuscript and here in the response letter.
 
@@
 \noindent Thanks for pointing out an important alternative hypothesis.
-To address this comment, we modeled the proportion of participants looking at the current speaker for our real data and for (fake) boredom-driven lookers (who look away from the current speaker at a constant rate, starting 1 second after the onset of speech) across turns of different length.
+To address this comment, we modeled the proportion of participants looking at the current speaker for our real data and for (hypothetical) boredom-driven lookers (who look away from the current speaker at a constant rate, starting 1 second after the onset of speech) across turns of different length.
 Even in short turns, where there is very little temporal room for children to behave differently from the boredom-driven lookers, we still see a clear difference between the boredom-driven simulations and the real data (Figure D.2).
-We take this result to be evidence against the alternative hypothesis; we report the details of the analysis in an appendix.
+We take this result to be evidence against the alternative hypothesis; we report the details of the analysis in the Appendix (Section C).
 
@@
 We agree and have changed our analyses to set ``no speech'' as the reference level.
-We had started with the ``normal speech'' condition as the reference level because this choice mirrors the coding in related work on adult turn-end prediction.
+We had started with the ``normal speech'' condition as the reference level because this choice mirrors the coding in related work on adult turn-end prediction (de Ruiter et al., 2006).
 But the additive model implied by the ``no speech'' baseline is easier to interpret.
@@
 \noindent Thanks for raising this challenging and important point.
-To summarize: We think that the robust question (speech act) effects we observed in both experiments provide strong evidence for the use of linguistic information.
+To summarize: We think that the robust question (speech act) effects we observed in both experiments provides strong evidence for the use of linguistic information.
 
@@
 Following your suggestion, we have thus made two additional changes.
-First, we added follow-up two-tailed t-tests, making pairwise comparisons between age groups for significant interactions so that we could find out how the interacting factor (e.g., language condition) significantly changes across age groups (Sections B.1--3).
+First, we added follow-up two-tailed \textit{t}-tests, making pairwise comparisons between age groups for significant interactions so that we could find out how the interacting factor (e.g., language condition) significantly changes across age groups (Sections B.1--3).
 Second, we have added individual models of the youngest age groups (ages 1, 2, and 3), comparing their real looking behavior to our randomly permuted data across conditions (at the ends of Sections 2.2.2 and 3.3.2).
@@
 \noindent We thank Reviewer 3 for this point.
-We repeat here some of the information now available in the Appendix for easy summary: Our design does not fully cross puppet pair (e.g., robots, blue puppets) with linguistic condition (e.g., ``words only'' and ``no speech'').
+We repeat here some of the information now available in the Appendix (Section D) for easy summary: Our design does not fully cross puppet pair (e.g., robots, blue puppets) with linguistic condition (e.g., ``words only'' and ``no speech'').
 Even though each puppet pair is associated with different conversation clips across children (e.g.
@@
 In the follow-up study, each child only saw one video---the normal speech conversation about birthday parties---with a randomly assigned puppet pair performing the conversation.
-Five children watched each puppet pair, for a total of 30 children across the six pairs (shown Figure 3).
+Five children watched each puppet pair, for a total of 30 children across the six pairs (shown in Figure 3).
 This experiment holds all things constant except for the appearance of the puppets.
@@
 We take this as evidence that, although we did not fully cross puppet pairs and linguistic conditions in Experiment 2, it was unlikely to have had strong effects on children's looking rates above and beyond the intended effects of linguistic condition.
-We have included details and graphs for this norming experiment in the Appendix.
+We have included details and graphs for this norming experiment in the Appendix (Section D).
 
@@
 Unfortunately, this was not a trivial analysis to implement, causing some delays in the manuscript resubmission.
-Running 5,000 random permutations and 20,000 mixed models required a substantial investment in terms of parallel computing infrastructure.
+Running 20,000 random permutations with mixed models required a substantial investment in terms of parallel computing infrastructure.
 
@@
 Please note that there were two problems with conducting the analyses as originally instructed in the reviewer's comments.
-1) We used $t$ values instead of $\beta$ estimates because the standard error estimates for the randomly permuted data were categorically higher (due to data sparsity issues).
-The use of $t$ values is closer to the practice in neuroimaging data analysis, where statistical significance is computed based on the empirical distribution of test statistics.
+1) We used $z$ values instead of $\beta$ estimates because the standard error estimates for the randomly permuted data were categorically higher (due to data sparsity issues).
+The use of $z$ values is closer to the practice in neuroimaging data analysis, where statistical significance is computed based on the empirical distribution of test statistics.
 2) Many of the models resulted in convergence warnings.
-We made the decision to report results across all convergent models, but we are happy to revisit this decision if the Reviewer believes there is an alternative approach that we should follow.
+We found that models with convergence warnings were more likely to have extreme variance in their estimates (see Appendix Table A.1).
+We therefore made the decision to report results across all convergent models, but we are happy to revisit this decision if the Reviewer believes there is an alternative approach that we should follow.
 
@@
 First, one of the most important theoretical contributions of our work is in establishing and replicating the effect of speech act (questions) in children and adults' spontaneous predictions.
-While participants may always use linguistic information to predict upcoming speaker changes, our results to not support the idea that they always use linguistic information to predict upcoming \emph{turn ends}, as is assumed in metalinguistic measures of turn-end prediction (e.g., pressing a button while listening to speech).
+While participants may always use linguistic information to predict upcoming speaker changes, our results do not support the idea that they always use linguistic information to predict upcoming \emph{turn ends}, as is assumed in metalinguistic measures of turn-end prediction (e.g., pressing a button while listening to speech).
 Instead, our results suggest that participants' spontaneous predictions are the result of question-monitoring, presumably achieved by recruiting linguistic cues to questionhood from the unfolding signal.
@@
 low parent-infant interaction styles) might affect children's early predictive skill.
-It also bridges prior work demonstrating that children begin taking non-linguistic turns in infancy (e.g., Hilbrink et al., 2015), but still have trouble integrating linguistic aspects of responding at age 3--4 (Casillas et al., in press; Garvey \& Berninger, 1984).
+It also bridges prior work demonstrating that children begin taking non-linguistic turns in infancy (e.g., Hilbrink et al., 2015) with work that shows that children still have trouble integrating linguistic aspects of responding at age 3--4 (Casillas et al., in press; Garvey \& Berninger, 1984).
 

### 2016-01-29 paper/CasillasFrank-Anticip.tex
Updated paper changes and code used in submission Jan 2016.

@@
 
-% \linespread{2}
+\linespread{2}
 \journal{}
@@
 But, unlike adults, lexical cues alone were not sufficient to improve prediction---children's performance was best overall with lexicosyntax and prosody together.
-Our findings support an account in which turn prediction emerges in infancy, but becomes fully integrated with linguistic processing only gradually.
+Our findings support an account in which turn prediction emerges in infancy, but then only gradually becomes fully integrated with linguistic processing.
 \end{abstract}
@@
 \begin{keyword}
-Turn taking \sep Conversation \sep Development \sep Prosody \sep Questions \sep Eye-tracking \sep Anticipation
+Turn taking \sep Conversation \sep Development \sep Prosody \sep Questions \sep Eye-tracking
 %% MSC codes here, in the form: \MSC code \sep code
@@
 Children participate in sequential coordination (proto-turn taking) with their caregivers starting at three months of age---before they can rely on any linguistic cues \citep[see, among others, ][]{bateson1975, hilbrink2015, jaffe2001, snow1977}.
-Infant turn taking is different from adult turn taking in several ways, however: it is heavily scaffolded by caregivers, has different timing from adult turn taking, and lacks semantic content \citep{hilbrink2015, jaffe2001}.
-But children's early, turn-structured social interactions are presumably a critical precursor to their later conversational turn taking: early non-verbal interactions likely establish the protocol by which children come to use language with others.
+However, infant turn taking is different from adult turn taking in several ways: it is heavily scaffolded by caregivers, has different timing from adult turn taking, and lacks semantic content \citep{hilbrink2015, jaffe2001}.
+But children's early, turn-structured social interactions are presumably a critical precursor to their later conversational turn taking.
+Early non-verbal interactions likely establish the protocol by which children come to use language with others.
 How do children integrate linguistic knowledge with these preverbal turn-taking abilities, and how does this integration change over the course of childhood?
@@
 
-\subsection{Adults' turn taking}
+\subsection{Adult turn taking}
 
@@
 Experimental results and current models of speech production suggest that it takes approximately 600 msec to produce a content word, and even longer to produce a simple utterance \citep{griffin2000, levelt1989}.
-So in order to achieve 200 msec turn transitions, speakers must begin formulating their response before the prior turn has ended \citep{levinson2013}.
+So in order to achieve 200 msec turn transitions, speakers must begin formulating their response before the prior turn has ended \citep{levinson2013, levinson2016}.
 Moreover, to formulate their response early on, speakers must track and anticipate what types of response might become relevant next.
@@
 
-Prior work suggests that adult listeners primarily use lexicosyntactic information to accurately predict upcoming turn structure \citep{de-ruiter2006}.
+Prior work suggests that adult listeners primarily use lexicosyntactic information to accurately predict upcoming turn structure.
 De Ruiter and colleagues \citeyearpar{de-ruiter2006} asked participants to listen to snippets of spontaneous conversation and to press a button whenever they anticipated that the current speaker was about to finish his or her turn.
@@
 
-\subsection{Children's turn prediction}
+\subsection{Turn taking in development}
 
@@
 So a primary role for lexicosyntactic cues doesn't fit well with children's pre-verbal turn taking.
-How does children's turn-taking
 
@@
 Infants catch onto the structure of proto-conversations quickly.
-By three to four months they notice disturbances to the contingency of their caregivers' response and, in reaction, change the rate and quality of their vocalizations \citep{k-bloom1988, masataka1993}.
+By three to four months they notice disturbances to the contingency of their caregivers' response and, in reaction, change the rate and quality of their vocalizations \citep{k-bloom1988, masataka1993, toda1993}.
 
@@
 But by nine months, their timing slows down considerably, only to gradually speed up again after 12 months \citep{hilbrink2015}.
-Taking turns with brief transitions between speakers is difficult for children; while their avoidance of overlap is nearly adult-like by nine months, the timing of their non-overlapped responses stays much longer than the adult 200 msec standard for the next few years \citep{casillas2016, garvey1984, garvey1981, ervin-tripp1979}.
+For children, taking turns with brief transitions between speakers is more difficult than avoiding speaker overlap; children's incidence of overlap is nearly adult-like by nine months, but the timing of their non-overlapped responses stays much longer than the adult 200 msec standard for the next few years \citep{casillas2016, garvey1984, garvey1981, ervin-tripp1979}.
 This puzzling pattern is likely due to their linguistic development: taking turns on time is easier when the response is a simple vocalization rather than a linguistic utterance.
-Integrating language into the turn-taking system may be one major factor in children's delayed responses \citep{casillas2016}.
-
-While children, like adults, might use linguistic cues in the ongoing turn to make predictions about upcoming turn structure, studies of early linguistic development point to a possible early advantage for prosody over lexicosyntax in children's turn-taking predictions.
+Integrating language into the turn-taking system may therefore be one major factor in children's delayed responses \citep{casillas2016}.
+
+Children, like adults, might use linguistic cues in the ongoing turn to make predictions about upcoming turn structure.
+Studies of early linguistic development point to a possible early advantage for prosody over lexicosyntax in children's turn-taking predictions.
 Infants can distinguish their native language's rhythm type from others soon after birth \citep{mehler1988, nazzi2003}; they show preference for the typical stress patterns of their native language over others by 6--9 months (e.g., iambic vs.
@@
 Second, the conversation stimuli \citet{keitel2013} used were somewhat unusual.
-The average gap between turns was 900 msec, which is much longer than typical adult timing, which averages around 200 msec \citep{stivers2009}.
+The average gap between turns was 900 msec, a duration much longer than typical adult timing, which averages around 200 msec \citep{stivers2009}.
 The speakers in the videos were also asked to minimize their movements while performing scripted, adult-directed conversation, which would have created a somewhat unnatural interaction.
@@
 \citeyearpar{keitel2013} study still demonstrates intriguing age-based differences in children's ability to predict upcoming turn structure.
-Our current work thus takes this paradigm as a starting point.\footnote{But also see \citet{casillas2012, casillas2013}.}
+Our current work takes this paradigm as a starting point.\footnote{But also see \citet{casillas2012, casillas2013}.}
 
@@
 Second, we find that children make more predictions than expected by chance starting at age two, but that this effect is small at first, and continues to improve through age six.
-Third, we find no evidence of an early prosody advantage in children's anticipations and, further, no evidence that prosodic or lexical cues alone can substitute for their combination in the full linguistic signal (as is proposed for adults; \citealp{de-ruiter2006}); instead, anticipation is strongest for stimuli with the full range of cues.
+Third, we find no evidence of an early prosody advantage in children's anticipations and, further, no evidence that prosodic or lexical cues alone can substitute for their combination in the full linguistic signal (as is proposed for adults; \citealp{de-ruiter2006}).
+Instead, anticipation is strongest for stimuli with the full range of cues.
 In sum, our findings support an account in which turn prediction emerges in infancy, but becomes fully integrated with linguistic processing only gradually across development.
@@
 We recorded participants' eye movements as they watched six short videos of two-person (dyadic) conversation interspersed with attention-getting filler videos.
-Each conversation video featured an improvised discourse in one of five languages (English, German, Hebrew, Japanese, and Korean); participants saw two videos in English and one in every other language.
+Each conversation video featured an improvised discourse in one of five languages (English, German, Hebrew, Japanese, and Korean).
+Participants saw two videos in English and one in every other language.
 The participants, all native English speakers, were only expected to understand the two videos in English.
@@
 
-After recording, we combined the audio and video recordings by hand, and cropped each recording to the 30-second interval with the most turn activity.
+After recording, we combined the audio and video recordings by hand, and cropped each recording to the (approximate) 30-second interval with the most turn activity.
 Because we recorded the conversations in stereo, the male and female voices came out of separate speakers during video playback.
@@
 non-question).
-The non-English stimuli were coded for speech act from a monolingual English-speaker's perspective, i.e., which turns ``sound like'' questions, and which do not: we asked five native American English speakers to listen to the audio recording for each turn and judge whether it sounded like a question.
-We then coded turns with at least 80\% ``yes'' responses as questions.
+The non-English stimuli were coded for speech act from a monolingual English-speaker's perspective, i.e., which turns ``sound like'' questions, and which do not: we asked five native American English speakers to listen to the audio recording for each non-English turn and judge whether it sounded like a question.
+We marked non-English turns as questions when at least 4 of the 5 listeners (80\%) said that the turn ``sounded like a question''.
 
@@
 We therefore allowed switches to occur as early as 200 msec before the end of the prior turn.
-For very early and very late switches, our requirement for 100 msec of fixation on each speaker would sometimes extend outside of the transition window boundaries (200 msec before and after the inter-turn gap).
-The maximally available fixation window was 100 msec before and after the earliest and latest possible switch point (300 msec before and after the inter-turn gap).
+Again, because it takes 200 msec to plan an eye movement, we counted anticipatory switches, at the latest, 200 msec after the onset of speech.
+Therefore, for very early and very late switches, our requirement of 100 msec of fixation on each speaker would sometimes extend outside of the transition window boundaries (200 msec before and after the inter-turn gap).
+The maximally available fixation window was therefore 100 msec before and after the earliest and latest possible switch point (300 msec before and after the inter-turn gap).
 We did not count switches made during the fixation window as anticipatory.
@@
 
+\linespread{1}
 \begin{table}[t]
@@
 \end{table}
+\linespread{2}
 
@@
 \end{center}
-\caption{Anticipatory gaze rates across language condition and transition type for the real (red and blue) and randomly permuted baseline (gray). Vertical bars represent 95\% confidence intervals.} 
+\caption{Anticipatory gaze rates across language condition and transition type for the real data (red and blue) and the randomly permuted baselines (gray).
+Vertical bars represent 95\% confidence intervals.}
 \label{fig:E1-randvsreal}
@@
 
+\linespread{1}
 \begin{table}[h!]
@@
     Type$=$\textit{non-Question}      	& -1.10131	& 0.86520		& -1.273	& 0.203055     	\\
-    Duration      										& 3.40171	& 1.22878		&  2.768	& 0.005634  **	\\
+    GapDuration      										& 3.40171	& 1.22878		&  2.768	& 0.005634  **	\\
     Age*LgCond$=$\textit{non-English} 	& 0.52066	& 0.21192		&  2.457	& 0.014015 * 	\\  
-    Age*TypeS$=$\textit{non-Question} 	& -0.01628	& 0.19442		& -0.084	& 0.933262     	\\
+    Age*Type$=$\textit{non-Question} 	& -0.01628	& 0.19442		& -0.084	& 0.933262     	\\
     LgCond$=$\textit{non-English}*			& 2.68171	& 1.35045		& 1.986	& 0.047057 * 	\\
@@
     Type$=$\textit{non-Question}				& -4.4953		& 1.3147		& -3.419	& 0.000628 *** \\
-    Duration													& -1.1227		& 1.9889		& -0.565	& 0.572414     	\\
+    GapDuration													& -1.1227		& 1.9889		& -0.565	& 0.572414     	\\
     LgCond$=$\textit{non-English}*			& 3.2972			& 1.6115		& 2.046	& 0.040747 * 	\\ 
@@
     LgCond$=$\textit{non-English}*			& 1.3625			& 3.0097		& 0.453	& 0.650749     	\\
-    \hspace*{5mm} Duration &&&& \\
+    \hspace*{5mm} GapDuration &&&& \\
     Type$=$\textit{non-Question}*				& 10.5107		& 3.3482		& 3.139	& 0.001694 ** 	\\
-    \hspace*{5mm} Duration &&&& \\
+    \hspace*{5mm} GapDuration &&&& \\
     LgCond$=$\textit{non-English}*			& -6.3156		& 4.4969		& -1.404	& 0.160191 	\\    
     \hspace*{5mm} Type$=$\textit{non-Question}* &&&& \\
-    \hspace*{5mm} Duration &&&& \\
+    \hspace*{5mm} GapDuration &&&& \\
     \hline
@@
   \end{small}
-  \caption{Model output for children and adults' anticipatory gaze switches.}
+  \caption{Model output for children and adults' anticipatory gaze switches in Experiment 1.}
 \label{tab:E1-models}
 \end{table}
+\linespread{2}
 
@@
 We estimated the probability that these patterns were the result of random looking by running the same regression models on participants' real eye-tracking data, only this time calculating their anticipatory gaze switches with respect to randomly permuted turn transition windows.
-This process involved: (1) randomizing the order and temporal placement of the analysis windows within each stimulus (Figure \ref{fig:shuffling}; ``analysis window'' is defined in Figure \ref{fig:criterion}) to randomly redistribute the analysis windows across the eye-tracking signal, (2) re-running each participant's eye tracking data through switch identification (described in Section \ref{sec:algorithm}) on each of the randomly permuted analysis windows, and (3) modeling the anticipatory switches from the randomly permuted data with the same statistical models we used for the original data (Section \ref{sec:models1}; Table \ref{tab:E1-models}).
+This process involved: (1) randomizing the order and temporal placement of the analysis windows within each stimulus (see Figure \ref{fig:shuffling}; ``analysis window'' is defined in Figure \ref{fig:criterion}) to randomly redistribute the analysis windows across the eye-tracking signal, (2) re-running each participant's eye tracking data through switch identification (described in Section \ref{sec:algorithm}) on each of the randomly permuted analysis windows, and (3) modeling the anticipatory switches from the randomly permuted data with the same statistical models we used for the original data (Section \ref{sec:models1}; Table \ref{tab:E1-models}).
 Importantly, although the onset time of each transition was shuffled within the eye-tracking signal, the other intrinsic properties of each turn transition (e.g., prior speaker identity, transition type, gap duration, language condition, etc.) stayed constant across each random permutation.
@@
 Similarly, the adults' data showed significant differentiation from the randomly permuted data for two of the three originally significant predictors---transition type and the transition type-gap duration interaction (greater than 99.9\% and 99.7\% of random \textit{z}-values, respectively, all \textit{p}$<$.01)---with marginal differentiation for the interaction of language condition and transition type (greater than 94.6\% of random \textit{z}-values; \textit{p}$=$.054).
+See Section \ref{sec:permutation} for more information on each predictor's random permutation distribution.
 
@@
 The models reported above revealed a significant interaction of age and language condition (Table \ref{tab:E1-models}) that was unlikely be due to random looking (Figure \ref{fig:E1-randvsreal}).
-To further explore this effect, we compared the effect of language condition across age groups: Using the permutation analyses above, we extracted the average difference score for the two language conditions (English minus non-English) for each participant, computing an overall average for each random permutation of the data.
+To further explore this effect, we compared the effect of language condition across age groups: using the permuted datasets described above, we extracted the average difference score for the two language conditions (English minus non-English) for each participant, computing an overall average for each random permutation of the data.
 Then, within each permutation, we made pairwise comparisons of the average difference scores across participant age groups.
@@
 Both groups made more anticipatory switches after hearing questions, compared to non-questions.
-Even in the English videos, when participants had full access to linguistic cues, their rates of anticipation were relatively low---comparable to the non-English videos---unless the turn was a question.
+Even in the English videos, when participants had full access to linguistic cues, their rates of anticipation were relatively low (even comparable to the non-English videos) unless the turn was a question.
 Prior work using online, metalinguistic tasks has shown that participants can use linguistic cues to accurately predict upcoming turn ends \citep{torreira2015, magyari2012, de-ruiter2006}.
@@
 
-In Experiment 2 we follow up on these findings, improving on two aspects of the design: first, our language manipulation in this first experiment was too coarse to provide data regarding specific linguistic channels (e.g., prosody vs.
+In Experiment 2, we follow up on these findings, improving on two aspects of the design: first, our language manipulation in this first experiment was too coarse to provide data regarding specific linguistic channels (e.g., prosody vs.
 lexicosyntax).
-In Experiment 2 we compared lexicosyntactic and prosodic cues with phonetically altered speech and used puppets to eliminate non-verbal cues to turn taking.
+In Experiment 2, we compared lexicosyntactic and prosodic cues with phonetically altered speech and used puppets to eliminate non-verbal cues to turn taking.
 Second, we were not able to pinpoint the emergence of anticipatory switching because the youngest age group in our sample was already able to make anticipatory switches at above chance rates.
-In Experiment 2 we explore a wider developmental range.
+In Experiment 2, we explore a wider developmental range.
 
@@
 As before, we asked talkers to speak ``as if they were on a children's television show'' and to ask at least a few questions during the improvisation.
-We cut each audio recording down to the 20-second interval with the most turn activity.
+We cut each audio recording down to the (approximate) 20-second interval with the most turn activity.
 The 20-second clips were then phonetically manipulated and used in the final video stimuli.
@@
 The \textit{no speech} condition served as a non-linguistic baseline.
-For this condition, we replaced the original clip with multi-talker babble: we overlaid different child-oriented conversations (not including the original one), and then cropped the result to the duration of the original video.
+For this condition, we replaced the original clip with multi-talker babble: we overlaid multiple child-oriented conversations (not including the original one), and then cropped the result to the duration of the original video.
 Thus, the \textit{no speech} audio clips lacked any linguistic information to upcoming turn boundaries---the only cue to turn taking was the opening and closing of the puppets' mouths.
@@
 There were 85 turn transitions for analysis after excluding transitions longer than 550 msec and shorter than 90 msec.
-The remaining turn transitions had slightly more questions than non-questions (N$=$50 and N$=$35, respectively), with transitions distributed somewhat evenly across conditions (keeping in mind that there were three \textit{normal} videos and only one partial cue video for each experiment version): \textit{normal} (N$=$36), \textit{words only} (N$=$13), \textit{prosody only} (N$=$17), and \textit{no speech} (N$=$19).
-Inter-turn gaps for questions (mean$=$365, median$=$427) were longer than those for non-questions (mean$=$302, median$=$323) on average, but gap duration was overall comparable across conditions: \textit{normal} (mean$=$334, median$=$321), \textit{words only} (mean$=$347, median$=$369), \textit{prosody only} (mean$=$365, median$=$369), and \textit{no words} (mean$=$319, median$=$329).
+The remaining turn transitions had more questions than non-questions (N$=$47 and N$=$38, respectively), with transitions distributed somewhat evenly across conditions (keeping in mind that there were three \textit{normal} videos and only one partial cue video for each experiment version): \textit{normal} (N$=$36), \textit{words only} (N$=$13), \textit{prosody only} (N$=$17), and \textit{no speech} (N$=$19).
+Inter-turn gaps for questions (mean$=$366, median$=$438, stdev$=$138 msec) were longer than those for non-questions (mean$=$305, median$=$325, stdev$=$94 msec) on average, but gap duration was overall comparable across conditions: \textit{normal} (mean$=$334, median$=$321, stdev$=$130 msec), \textit{words only} (mean$=$347, median$=$369, stdev$=$ 115 msec), \textit{prosody only} (mean$=$365, median$=$369, stdev$=$104 msec), and \textit{no words} (mean$=$319, median$=$329, stdev$=$136 msec).
 The longer gaps for question transitions could give them an advantage because our anticipatory measure includes shifts initiated during the gap between turns (Figure \ref{fig:criterion}).
@@
 \end{table}
+\linespread{2}
+
+\linespread{1}
+\begin{table}
+\begin{center}
+  \begin{tabular}{llcccc}
+    \hline
+    Condition & Speaker & Addressee & Other onscreen & Offscreen\\ 
+    \hline
+    Normal 				& 0.58 & 0.12 & 0.17 & 0.13 \\ 
+    Words only 		& 0.54 & 0.11 & 0.24 & 0.10 \\ 
+    Prosody only 	& 0.48 & 0.12 & 0.26 & 0.15 \\ 
+    No speech 		& 0.44 & 0.13 & 0.26 & 0.18 \\
+    \hline
+  \end{tabular}
+\end{center}
+  \caption{Average proportion of gaze to the current speaker and addressee during periods of talk across conditions.}
+\label{tab:look_e2b}
+\end{table}
+\linespread{2}
+
+% Average prop gaze to the current speaker across conditions for one-year-olds
+%                           speaker addressee other-screen offscreen
+%Normal   		  0.4737582 0.1366432    0.2003952 0.1892035
+%Words only    0.3977261 0.1283417    0.3112005 0.1627317
+%Prosody only 0.4290202 0.1611453    0.2558657 0.1539688
+%No speech    0.3786925 0.1228615    0.2356423 0.2628037
 
@@
 Polar questions often have a final rising intonational contour, but \textit{wh-}questions often do not  \citep{hedberg2010}.}, and identified anticipatory gaze switches to the upcoming speaker using the methods from Experiment 1.
+
+\begin{figure}[!ht]
+\begin{center}
+\includegraphics[width=\textwidth]{figures/E2-samples-by-lang-groups-trans-types.png}
+\end{center}
+\caption{Anticipatory gaze rates across language condition and transition type for the real data (blue, dark green, light green, and red) and the randomly permuted baselines (gray).
+Vertical bars represent 95\% confidence intervals.}
+\label{fig:E2-randvsreal}
+\end{figure}
 
@@
 
-\begin{table}[t]
-\begin{center}
-  \begin{tabular}{llcccc}
-    \hline
-    Condition & Speaker & Addressee & Other onscreen & Offscreen\\ 
-    \hline
-    Normal 				& 0.58 & 0.12 & 0.17 & 0.13 \\ 
-    Words only 		& 0.54 & 0.11 & 0.24 & 0.10 \\ 
-    Prosody only 	& 0.48 & 0.12 & 0.26 & 0.15 \\ 
-    No speech 		& 0.44 & 0.13 & 0.26 & 0.18 \\
-    \hline
-  \end{tabular}
-\end{center}
-  \caption{Average proportion of gaze to the current speaker and addressee during periods of talk across conditions.}
-\label{tab:look_e2b}
-\end{table}
-
-% Average prop gaze to the current speaker across conditions for one-year-olds
-%                           speaker addressee other-screen offscreen
-%Normal   		  0.4737582 0.1366432    0.2003952 0.1892035
-%Words only    0.3977261 0.1283417    0.3112005 0.1627317
-%Prosody only 0.4290202 0.1611453    0.2558657 0.1539688
-%No speech    0.3786925 0.1228615    0.2356423 0.2628037
-
-\begin{figure}[h]
-\begin{center}
-\includegraphics[width=\textwidth]{figures/E2-samples-by-lang-groups-trans-types.png}
-\end{center}
-\caption{Anticipatory gaze rates across language condition and transition type for the real (blue, dark green, light green, and red) and randomly permuted baseline (gray).
-Vertical bars represent 95\% confidence intervals.}
-\label{fig:E2-randvsreal}
-\end{figure}
-
 \subsubsection{Statistical models}
@@
 Similarly, the adults' data showed significant differentiation from the randomly permuted data for all originally significant predictors: gap duration and language condition for \textit{normal} speech and \textit{words only} speech (greater than 100\%, 97.8\%, and 99.1\% of random \textit{z}-values, respectively, all \textit{p}$<$.05).
-
+See Section \ref{sec:permutation} for more information on each predictor's random permutation distribution.
+
+\linespread{1}
 \begin{minipage}[t]{0.95\linewidth}
@@
     Type$=$\textit{non-Question}			& -0.88900		& 0.61192		& -1.453	& 0.1463        		\\
-    Duration												& 3.94743		& 0.61668		& 6.401	& 1.54e-10 ***    	\\
+    GapDuration												& 3.94743		& 0.61668		& 6.401	& 1.54e-10 ***    	\\
     Age*Type$=$\textit{non-Question}	& 0.15359		& 0.13996		& 1.097	& 0.2725        		\\
@@
     Type$=$\textit{non-Question}			& 0.4375			& 0.5854		& 0.747	& 0.4549     		\\ 
-    Duration												& 4.6961			& 1.1804		& 3.978	& 6.94e-05 *** 		\\ 
+    GapDuration												& 4.6961			& 1.1804		& 3.978	& 6.94e-05 *** 		\\ 
 	\hline
@@
     &&&& \\
-  \caption{Model output for children and adults' anticipatory gaze switches.}
+  \caption{Model output for children and adults' anticipatory gaze switches in Experiment 2.}
 \label{tab:E2-models}
@@
 \end{minipage}
+\linespread{2}
 
@@
 The core aims of Experiment 2 were to gain better traction on the individual roles of prosody and lexicosyntax in children's turn predictions, and to find the age of emergence for spontaneous turn anticipation.
-Many of our results replicate the findings from Experiment 1: participants made more anticipatory switches when they had access to lexical information and, when they did, tended to make more anticipatory switches for questions compared to non-questions.
+Many of our results replicate the findings from Experiment 1: participants often made more anticipatory switches when they had access to lexical information and, when they did, tended to make more anticipatory switches for questions compared to non-questions.
 
 As in Experiment 1, children and adults spontaneously tracked the turn structure of the conversations, making anticipatory gaze switches at above-chance rates across all ages when listening to natural speech.
-They also made far more anticipations for questions than for non-question turns---at least for children two years olds and older.
+They also made far more anticipations for questions than for non-question turns---at least for children two years old and older.
 But these effects were different for the two conditions with partial linguistic information: \textit{prosody only} and \textit{words only}.
@@
 Children may have therefore had an easier time picking out and interpreting lexical cues to questionhood.
-The question effect showed its first significant gains between ages three and four in the \textit{normal} speech condition of Experiment 2 (Figure \ref{fig:E2-lgagetypeinteraction}), by which time children frequently hear and use a variety of polar and \textit{wh}-questions \citep{clark2009}.
+Developmentally, the question effect showed its first significant gains between ages three and four in the \textit{normal} speech condition of Experiment 2 (Figure \ref{fig:E2-lgagetypeinteraction}), by which time children frequently hear and use a variety of polar and \textit{wh}-questions \citep{clark2009}.
 Furthermore, while lexicosyntactic question cues were available on every instance of \textit{wh}- and \textit{yes/no} questions in our stimuli, prosodic question cues were only salient on \textit{yes/no} questions.
@@
 
-Prior work on children's acquisition of questions indicates that they may already have some knowledge of question-answer sequences by the time they begin to speak: questions make up approximately one third of the utterances children hear, before and after the onset of speech, and even into their preschool years, though the types and complexity of questions change throughout development \citep{casillas2016, fitneva2012, henning2005, shatz1979}.\footnote{There is substantial variation question frequency by individual and socioeconomic class \citep{hart1992}.} For the first few years, many of the questions directed to children are ``test'' questions---questions that the caregiver already has the answer to (e.g., ``What does a cat say?''), but this changes as children get older.
+Prior work on children's acquisition of questions indicates that they may already have some knowledge of question-answer sequences by the time they begin to speak: questions make up approximately one third of the utterances children hear, before and after the onset of speech, and even into their preschool years, though the type and complexity of questions changes throughout development \citep{casillas2016, fitneva2012, henning2005, shatz1979}.\footnote{There is substantial variation in question frequency by individual and socioeconomic class \citep{hart1992, weisleder2012}.} For the first few years, many of the questions directed to children are ``test'' questions---questions that the caregiver already has the answer to (e.g., ``What does a cat say?''), but this changes as children get older.
 Questions help caregivers to get their young children's attention and to ensure that information is in common ground, even if the responses are non-verbal or infelicitous \citep{bruner1985, fitneva2012, snow1977}.
@@
 and (2) what is it about questions that makes children and adults more likely to anticipatorily switch their gaze to addressees?
-Other request formats, such as imperatives, compliments, and complaints make a response from the addressee highly likely in the next turn \citep{schegloff2007}.
+Other turn types, such as imperatives, compliments, and complaints make a response from the addressee highly likely in the next turn \citep{schegloff2007}.
 Rhetorical and tag questions, on the other hand, take a similar form to prototypical polar questions, but often do not require an answer.
@@
 
-More broadly, our results suggest that participants' spontaneous predictions, at least while viewing third-party conversation, are driven by what lies \textit{beyond} the end of the current turn---not by the upcoming end of the turn itself, as has been focused on in prior work \citep{torreira2015, keitel2013, magyari2012, de-ruiter2006}.
+More broadly, our results suggest that participants' spontaneous predictions, at least while viewing third-party conversation, are driven by what lies \textit{beyond} the end of the current turn---not just by the upcoming end of the turn itself, as has been focused on in prior work \citep{torreira2015, keitel2013, magyari2012, de-ruiter2006}.
 In future work, it will be crucial to measure prediction from a first-person perspective to resolve this apparent contradiction (see also \citealp{holler2015}).
@@
 low parent-infant interaction styles) feed into this early predictive skill.
-It also bridges  prior work showing a predisposition for turn taking in infancy (e.g., \citealp{bateson1975, hilbrink2015, jaffe2001, snow1977}) but late acquisition of adult-like competence for turn-taking in children's conversations \citep{casillas2016, garvey1984, garvey1981, ervin-tripp1979}.
+It also bridges  prior work showing a predisposition for turn taking in infancy (e.g., \citealp{bateson1975, hilbrink2015, jaffe2001, snow1977}) with children's apparently \textit{late} acquisition of adult-like competence for turn taking in actual conversation \citep{casillas2016, garvey1984, garvey1981, ervin-tripp1979}.
 
@@
 The present study offers new findings regarding the role of speech acts and linguistic processing in online turn prediction, and has given evidence that turn prediction emerges by age two, but is not integrated with linguistic cues until much later.
-Using language to make predictions about upcoming interactive content takes time and, for both children and adults, is primarily driven by participants' orientation to what will happen beyond the end of the current turn.
+Using language to make predictions about upcoming interactive content takes time and, for both children and adults, is primarily driven by participants' orientation to what will happen next---beyond the end of the current turn.
 
@@
 This work was supported by an ERC Advanced Grant to Stephen C.
-Levinson (269484-INTERACT), an NSF graduate research and dissertation improvement fellowship to the first author, and a Merck Foundation fellowship to the second author.
+Levinson (269484-INTERACT), an NSF Graduate Research Fellowship and NSF Dissertation Improvement Grant to MC, and a Merck Foundation fellowship to MCF.
 Earlier versions of these data and analyses were presented to conference audiences \citep{casillas2012, casillas2013}.
@@
 We reported \textit{z}-values in the main text rather than beta estimates because the standard errors in the randomly permuted data models were much higher than for the original data.
-The distributions of each predictor's beta estimate, standard error, and \textit{z}-value for adults and children in each experiment are shown in the graphs below (Figures \ref{fig:E1-ChiTs}--\ref{fig:E2-AduSEs}).
-In each plot, the gray dots represent randomly permuted model estimates for the value listed (beta, standard error, or \textit{z}), the white dots represent the model estimates from the original data, and the triangles represent the 95th percentile for each random distribution.
+The distributions for each predictor's beta estimate, standard error, and \textit{z}-value for adults and children in each experiment are shown in the graphs below (Figures \ref{fig:E1-ChiTs}--\ref{fig:E2-AduSEs}).
+In each plot, the gray dots represent the absolute value of the 5,000 randomly permuted model estimates for the estimate type plotted (beta, standard error, or \textit{z}), the white circles represent the model estimates from the original data, and the black triangles represent the 95th percentile for each random distribution.
 
@@
   \subfloat[Adults]{\includegraphics[width=0.44\textwidth]{figures/E1-adu-randrun-SEs-absolute.png}\label{fig:E1-AduSEs}}
-  \caption{Random-permutation and original $|$\textit{SE}-values$|$ for predictors of anticipatory gaze rates in Experiment 1.}
+  \caption{Random-permutation and original \textit{SE}-values for predictors of anticipatory gaze rates in Experiment 1.}
 \label{fig:E1-SEs}
@@
   \subfloat[Adults]{\includegraphics[width=0.44\textwidth]{figures/E2-adu-randrun-SEs-absolute.png}\label{fig:E2-AduSEs}}
-  \caption{Random-permutation and original $|$\textit{SE}-values$|$ for predictors of anticipatory gaze rates in Experiment 2.}
+  \caption{Random-permutation and original \textit{SE}-values for predictors of anticipatory gaze rates in Experiment 2.}
 \label{fig:E2-SEs}
@@
 Non-convergent models made up 22.4--24.4\% of the random permutation models in Experiment 1 and 69--70\% of the random permutation models in Experiment 2.
-The \textit{z}-values for each predictor in the converging and non-converging models from Experiment 1 are shown in table \ref{tab:nonconv_e1}.
+The \textit{z}-values for each predictor in the converging and non-converging models from Experiment 1 are shown in Table \ref{tab:nonconv_e1}.
 
@@
 
+\linespread{1}
 \begin{sidewaystable}[!htbp]
@@
 \end{sidewaystable}
+\linespread{2}
 
@@
 
-In each of the plots below, the dot represents the real data value for the effect being shown.
+In each of the plots below, the black dot represents the real data value for the effect being shown.
 The effect sizes from the 5,000 randomly permuted data sets are shown in the distribution.
@@
 Even though each puppet pair is associated with different conversation clips across children (e.g., robots talking about kitties, birthday parties, and pancakes), the robot puppets themselves were exclusively associated with the \textit{words only} condition.
-Similarly, merpeople were exclusively associated with \textit{prosody only} speech, and the puppets wearing dress clothes were exclusively associated with the \textit{no speech} condition.
+Similarly, merpeople were exclusively associated with \textit{prosody only} speech, and the puppets wearing dressy clothes were exclusively associated with the \textit{no speech} condition.
 We designed the experiment this way to increase its pragmatic felicity for older children (i.e., robots make robot sounds, merpeople's voices are muffled under the water, the party-going puppets are in a `party' room with many other voices).
@@
 
+\linespread{1}
 \begin{table}
@@
     \hline
-		 \footnotesize{\textit{Reference level: mermaid puppets}} &&&& \\
+		 \footnotesize{\textit{Reference level: mer-puppets}} &&&& \\
     \hline

### 2016-01-29 analysis/E1/0-helper.R

@@ -14,7 +14,7 @@ library(car)
 # Misc helper functions
 ################################################################################
 # Stat summary functions
-tstat <- function(m) fixef(m)/sqrt(diag(vcov(m)))
+zstat <- function(m) fixef(m)/sqrt(diag(vcov(m)))
 SEstat <- function(m) sqrt(diag(vcov(m)))
 
 # Add some style elements for ggplot2
@@ -167,7 +167,7 @@ smoothLD <- function(dt) {
 ################################################################################
 # Random run plot functions
 ################################################################################
-combine.runs <- function(betas, ses, ts, fits,
+combine.runs <- function(betas, ses, zs, fits,
 	modelfiles, infofiles, realm, reali) {
 	# Random run info
 	for (file in modelfiles) {
@@ -185,7 +185,7 @@ combine.runs <- function(betas, ses, ts, fits,
 		# ts
 		insert <- data.frame(t(data$t))
 		colnames(insert) <- data$Predictor
-		ts[Run == run, (data$Predictor) := insert]
+		zs[Run == run, (data$Predictor) := insert]
 	}
 	for (file in infofiles) {
 		data <- fread(paste(model.path, file, sep=""))
@@ -210,33 +210,16 @@ combine.runs <- function(betas, ses, ts, fits,
 	# ts
 	insert <- data.frame(t(realmodel$t))
 	colnames(insert) <- realmodel$Predictor
-	ts[Run == 0, (realmodel$Predictor) := insert]
+	zs[Run == 0, (realmodel$Predictor) := insert]
 	# fits
 	fits[Run == 0, LogLik := realinfo$LogLik]
 	fits[Run == 0, AIC := realinfo$AIC]
 	fits[Run == 0, Error := realinfo$Error]
 
 	# Return the filled in DTs as a list
-	setDTs <- list(B=betas, SE=ses, T=ts, F=fits)
+	setDTs <- list(B=betas, SE=ses, Z=zs, F=fits)
 	return(setDTs)
 }
-
-# combine.runs <- function(DT, filelist, name) {
-	# curr.row <- 1
-	# for (file in filelist) {
-		# # Read in a file
-		# data <- fread(paste(processed.data.path, file, sep=""))
-		# # Add the data to the big data table
-		# DT[curr.row, names(DT) := data]
-		# curr.row <- curr.row + 1
-	# }
-	# setkeyv(DT, "Run")
-	# # Write it out
[…]

### 2016-01-29 analysis/E1/3-plots.stats.analysis.R

@@ -23,7 +23,7 @@ subject.info <- fread(paste(info.path,"SubjectInfo.csv",sep=""))
 
 # Color and line palettes
 stoplightPalette <- c("dodgerblue3", "firebrick2")
-stoplightPalette2 <- c("gray26", "dodgerblue3", "firebrick2")
+stoplightPalette2 <- c("dodgerblue3", "firebrick2", "gray26")
 stoplightPalette3 <- c("dodgerblue3", "firebrick2")
 lines <- c("solid", "dashed")
 lines2 <- c("longdash", "dotted")
@@ -82,25 +82,40 @@ switch.all <- switch.all %>%
 write.csv(switch.all, paste(processed.data.path,
 	"SwitchAll.csv", sep=""), row.names=FALSE)
 
+# Needed for plotting below
+conf.interval <- 0.95
+switch.real <- filter(switch.all, Sample == "Transition") %>%
+               select(-Sample)
+
 ################################################################################
 # Create plots
 ################################################################################
 ###### DIFF BETWEEN RANDOM AND TRANSITION SAMPLE SETS ##########################
-mean.agg <- aggregate(Switch ~ LgGroup + Sample + Age, switch.all, mean)
+# Mean + vars
+mean.agg <- group_by(switch.all, LgGroup, Sample, Age) %>%
+	summarise(mean = mean(Switch))
 langs <- mean.agg$LgGroup
 ages <- mean.agg$Age
 samples <- mean.agg$Sample
-means <- mean.agg$Switch
-errs <- aggregate(Switch ~ LgGroup + Sample + Age, switch.all, sem)$Switch
-linecolors <- c(
-	"black", "black",
-	"green2", "red",
-	"black", "black",
-	"green2", "red",
-	"black", "black",
-	"green2", "red",
-	"black", "black",
-	"green2", "red")
+means <- mean.agg$mean
+# SE
+err.agg <- group_by(switch.all, LgGroup, Sample, Age) %>%
+	summarise(se = sem(Switch))
+errs <- err.agg$se
+# SD
+sd.agg <- group_by(switch.all, LgGroup, Sample, Age) %>%
+	summarise(sd = sd(Switch))
+sds <- sd.agg$sd
+# N
+sampszs <- switch.all %>%
+    group_by(LgGroup, Sample, Age) %>%
+    summarise(Switch = n())
[…]

### 2016-01-29 analysis/E1/4-model-functions.R

@@ -13,8 +13,6 @@ library(data.table)
 processed.data.path <- "processed_data/" #"../../data/E1/processed_data/"
 model.path <- "processed_data/models/"
 switch.all <- fread(paste(processed.data.path,"SwitchAll.csv", sep=""))
-chi.coefs <- fread(paste(processed.data.path,"chi.coefs-0.csv", sep=""))
-adu.coefs <- fread(paste(processed.data.path,"adu.coefs-0.csv", sep=""))
 
 run.models <- function(ns) {
     # Split the data into adults and children for separate models
@@ -43,7 +41,7 @@ run.models <- function(ns) {
 			chi.max.r.m <- data.frame(
 				B = fixef(chi.max.r),
 				SE = SEstat(chi.max.r),
-				t = tstat(chi.max.r))
+				z = zstat(chi.max.r))
 			chi.max.r.m <- cbind(Predictor = rownames(chi.max.r.m), chi.max.r.m)
 			rownames(chi.max.r.m) <- NULL
 				
@@ -92,7 +90,7 @@ run.models <- function(ns) {
 			adu.max.r.m <- data.frame(
 				B = fixef(adu.max.r),
 				SE = SEstat(adu.max.r),
-				t = tstat(adu.max.r))
+				z = zstat(adu.max.r))
 			adu.max.r.m <- cbind(Predictor = rownames(adu.max.r.m), adu.max.r.m)
 			rownames(adu.max.r.m) <- NULL
 				

### 2016-01-29 analysis/E1/5-compare.R

@@ -34,9 +34,9 @@ aduSE <- data.table(Run = aruns)
 for (var in aduvars) {
 	aduSE[,c(as.character(var)) := rep(0, nrow(aduSE))]	
 }
-aduT <- data.table(Run = aruns)
+aduZ <- data.table(Run = aruns)
 for (var in aduvars) {
-	aduT[,c(as.character(var)) := rep(0, nrow(aduT))]	
+	aduZ[,c(as.character(var)) := rep(0, nrow(aduZ))]	
 }
 chiB <- data.table(Run = cruns)
 for (var in chivars) {
@@ -46,9 +46,9 @@ chiSE <- data.table(Run = cruns)
 for (var in chivars) {
 	chiSE[,c(as.character(var)) := rep(0, nrow(chiSE))]	
 }
-chiT <- data.table(Run = cruns)
+chiZ <- data.table(Run = cruns)
 for (var in chivars) {
-	chiT[,c(as.character(var)) := rep(0, nrow(chiT))]	
+	chiZ[,c(as.character(var)) := rep(0, nrow(chiZ))]	
 }
 # Model fit table
 aduFit <- data.table(Run = aruns,
@@ -60,10 +60,10 @@ chiFit <- data.table(Run = cruns,
 	AIC = rep(0, length(cruns)),
 	Error = rep("", length(cruns)))
 
-aduruns <- combine.runs(aduB, aduSE, aduT, aduFit,
+aduruns <- combine.runs(aduB, aduSE, aduZ, aduFit,
 	adu.r.m.files, adu.r.i.files, "adu.max.m-0.csv", "adu.max.i-0.csv")
 aduruns[[4]]$ErrorBin <- ifelse(aduruns[[4]]$Error == "", 0, 1)
-chiruns <- combine.runs(chiB, chiSE, chiT, chiFit,
+chiruns <- combine.runs(chiB, chiSE, chiZ, chiFit,
 	chi.r.m.files, chi.r.i.files, "chi.max.m-0.csv", "chi.max.i-0.csv")
 chiruns[[4]]$ErrorBin <- ifelse(chiruns[[4]]$Error == "", 0, 1)
 
@@ -77,25 +77,46 @@ chiruns <- lapply(chiruns, merge, chinoerrors, by="Run")
 ### Adults ###
 aduBetas <- melt(aduruns[[1]], names(aduruns[[1]])[1])
[…]

### 2016-01-29 analysis/E1/_RUNME.R

@@ -5,7 +5,7 @@ source("2-find.switches.R")
 get.switches(1, FALSE)
 
 # Find switches in randomly-shuffled gap versions
-rand.runs <- 301:5000
+rand.runs <- 1:5000
 make.rand.gaps(rand.runs)
 get.switches(rand.runs, TRUE)
 

### 2016-01-29 analysis/E2/0-helper.R

@@ -13,7 +13,7 @@ library(car)
 # Misc helper functions (from Mike)
 ################################################################################
 # Stat summary functions
-tstat <- function(m) fixef(m)/sqrt(diag(vcov(m)))
+zstat <- function(m) fixef(m)/sqrt(diag(vcov(m)))
 SEstat <- function(m) sqrt(diag(vcov(m)))
 
 # Add some style elements for ggplot2
@@ -66,6 +66,12 @@ getstars <- function(x) {
     if (x < .001) {return("***")}
     if (x < .01) {return("**")}
     if (x < .05) {return("*")}
+}
+
+factorizeDF <- function(df) {
+	idxs <- sapply(df, is.character)
+	df[idxs] <- lapply(df[idxs], as.factor)
+	return(df)
 }
 
 ################################################################################
@@ -175,7 +181,7 @@ smoothLD <- function(dt) {
 ################################################################################
 # Random run plot functions
 ################################################################################
-combine.runs <- function(betas, ses, ts, fits,
+combine.runs <- function(betas, ses, zs, fits,
 	modelfiles, infofiles, realm, reali) {
 	# Random run info
 	for (file in modelfiles) {
@@ -193,7 +199,7 @@ combine.runs <- function(betas, ses, ts, fits,
 		# ts
 		insert <- data.frame(t(data$t))
 		colnames(insert) <- data$Predictor
-		ts[Run == run, (data$Predictor) := insert]
+		zs[Run == run, (data$Predictor) := insert]
 	}
 	for (file in infofiles) {
 		data <- fread(paste(model.path, file, sep=""))
@@ -218,33 +224,16 @@ combine.runs <- function(betas, ses, ts, fits,
 	# ts
 	insert <- data.frame(t(realmodel$t))
 	colnames(insert) <- realmodel$Predictor
-	ts[Run == 0, (realmodel$Predictor) := insert]
+	zs[Run == 0, (realmodel$Predictor) := insert]
 	# fits
 	fits[Run == 0, LogLik := realinfo$LogLik]
 	fits[Run == 0, AIC := realinfo$AIC]
 	fits[Run == 0, Error := realinfo$Error]
 
 	# Return the filled in DTs as a list
-	setDTs <- list(B=betas, SE=ses, T=ts, F=fits)
+	setDTs <- list(B=betas, SE=ses, Z=zs, F=fits)
 	return(setDTs)
 }
-
[…]

### 2016-01-29 analysis/E2/3-plots.stats.analysis.R

@@ -25,8 +25,8 @@ vid.info <- fread(paste(info.path,"VideoSegmentInfo.csv",sep=""))
 # Color and line palettes
 stoplightPalette <- c("dodgerblue3", "forestgreen",
 	"green3", "firebrick2")
-stoplightPalette2 <- c("gray26", "firebrick2",
-     "forestgreen", "dodgerblue3", "green3")
+stoplightPalette2 <- c("firebrick2", "forestgreen",
+	"dodgerblue3", "gray26", "green3")
 stoplightPalette3 <- c("dodgerblue3", "forestgreen",
 	"green3", "firebrick2")
 lines <- c("solid", "dashed")
@@ -74,9 +74,8 @@ switch.all <- rbind(switch.final, switch.final.r)
 gap.info <- gap.info %>%
             filter(Exclude == 0) %>%
             select(Gap, Duration, Type, QType, Segment) %>%
-            left_join(select(vid.info, Segment, Condition), by = "Segment") %>%
-            mutate(Type = ifelse(QType == "wh" &
-            	Condition == "muffled", "S", Type))
+            left_join(select(vid.info, Segment, Condition), by = "Segment")
+gap.info[which(Condition == "muffled" & QType == "wh"),]$Type <- "S"
 
 sub.info <- select(subject.info, Subject, Age) %>%
             filter(Age != "NA")
@@ -90,32 +89,41 @@ write.csv(switch.all, paste(processed.data.path,
 	"SwitchAll.csv", sep=""), row.names=FALSE)
 
 
+# Needed for plotting below
+conf.interval <- 0.95
+switch.real <- filter(switch.all, Sample == "Transition") %>%
+               select(-Sample)
+
 ################################################################################
 # Create plots
 ################################################################################
 ###### DIFF BETWEEN RANDOM AND TRANSITION SAMPLE SETS ##########################
-mean.agg <- aggregate(Switch ~ Condition + Sample + Age, switch.all, mean)
+# Mean + vars
+mean.agg <- group_by(switch.all, Condition, Sample, Age) %>%
+	summarise(mean = mean(Switch))
 conds <- mean.agg$Condition
 ages <- mean.agg$Age
 ages[ages == "21"] <- "A"
 samples <- mean.agg$Sample
-means <- mean.agg$Switch
[…]

### 2016-01-29 analysis/E2/4-model-functions.R

@@ -13,8 +13,6 @@ library(data.table)
 processed.data.path <- "processed_data/" #"../../data/E1/processed_data/"
 model.path <- "processed_data/models/"
 switch.all <- fread(paste(processed.data.path,"SwitchAll.csv",sep=""))
-chi.coef <- fread(paste(processed.data.path,"chi.coefs-0.csv",sep=""))
-adu.coef <- fread(paste(processed.data.path,"adu.coefs-0.csv",sep=""))
 
 run.models <- function(ns) {
     # Split the data into adults and children for separate models
@@ -45,7 +43,7 @@ run.models <- function(ns) {
 			chi.max.r.m <- data.frame(
 				B = fixef(chi.max.r),
 				SE = SEstat(chi.max.r),
-				t = tstat(chi.max.r))
+				z = zstat(chi.max.r))
 			chi.max.r.m <- cbind(Predictor = rownames(chi.max.r.m), chi.max.r.m)
 			rownames(chi.max.r.m) <- NULL
 				
@@ -94,7 +92,7 @@ run.models <- function(ns) {
 			adu.max.r.m <- data.frame(
 				B = fixef(adu.max.r),
 				SE = SEstat(adu.max.r),
-				t = tstat(adu.max.r))
+				z = zstat(adu.max.r))
 			adu.max.r.m <- cbind(Predictor = rownames(adu.max.r.m), adu.max.r.m)
 			rownames(adu.max.r.m) <- NULL
 				

### 2016-01-29 analysis/E2/5-compare.R

@@ -34,9 +34,9 @@ aduSE <- data.table(Run = aruns)
 for (var in aduvars) {
 	aduSE[,c(as.character(var)) := rep(0, nrow(aduSE))]	
 }
-aduT <- data.table(Run = aruns)
+aduZ <- data.table(Run = aruns)
 for (var in aduvars) {
-	aduT[,c(as.character(var)) := rep(0, nrow(aduT))]	
+	aduZ[,c(as.character(var)) := rep(0, nrow(aduZ))]	
 }
 chiB <- data.table(Run = cruns)
 for (var in chivars) {
@@ -46,9 +46,9 @@ chiSE <- data.table(Run = cruns)
 for (var in chivars) {
 	chiSE[,c(as.character(var)) := rep(0, nrow(chiSE))]	
 }
-chiT <- data.table(Run = cruns)
+chiZ <- data.table(Run = cruns)
 for (var in chivars) {
-	chiT[,c(as.character(var)) := rep(0, nrow(chiT))]	
+	chiZ[,c(as.character(var)) := rep(0, nrow(chiZ))]	
 }
 # Model fit table
 aduFit <- data.table(Run = aruns,
@@ -60,10 +60,10 @@ chiFit <- data.table(Run = cruns,
 	AIC = rep(0, length(cruns)),
 	Error = rep("", length(cruns)))
 
-aduruns <- combine.runs(aduB, aduSE, aduT, aduFit,
+aduruns <- combine.runs(aduB, aduSE, aduZ, aduFit,
 	adu.r.m.files, adu.r.i.files, "adu.max.m-0.csv", "adu.max.i-0.csv")
 aduruns[[4]]$ErrorBin <- ifelse(aduruns[[4]]$Error == "", 0, 1)
-chiruns <- combine.runs(chiB, chiSE, chiT, chiFit,
+chiruns <- combine.runs(chiB, chiSE, chiZ, chiFit,
 	chi.r.m.files, chi.r.i.files, "chi.max.m-0.csv", "chi.max.i-0.csv")
 chiruns[[4]]$ErrorBin <- ifelse(chiruns[[4]]$Error == "", 0, 1)
 
@@ -77,25 +77,59 @@ chiruns <- lapply(chiruns, merge, chinoerrors, by="Run")
 ### Adults ###
 aduBetas <- melt(aduruns[[1]], names(aduruns[[1]])[1])
[…]

### 2016-01-29 analysis/E2/6-boredom.hypoth.R

@@ -1,0 +1,318 @@
+# library(grid)
+# library(ggplot2)
+# library(bootstrap)
+# library(lme4)
+# library(stringr)
+# library(plotrix)
+# library(reshape)
+# library(plyr)
+# library(car)
+# library(gridExtra)
+# library(scales)  
+# library(Hmisc)
+# library(moments)
+rm(list = ls())
+source("0-helper.R")
+library(data.table)
+library(bit64)
+
+################################################################################
+################################################################################
+info.path <- "info/"
+processed.data.path <- "processed_data_boredom/"
+plot.path <- "plots/"
+
+colorPalette <- c("black", "gray")
+
+# Read in supplemental data and prep
+vid.info <- fread(paste(info.path, "VideoSegmentInfo.csv", sep=""))
+#vid.info$VideoDuration <- NULL
+turn.info <- fread(paste(info.path, "TurnInfo.csv", sep=""))
+subject.info <- fread(paste(info.path, "SubjectInfo.csv", sep=""))
+
+# Round the time values for turns
+turn.info = within(turn.info, {
+  Onset = round(Onset, 5)
+  Offset = round(Offset, 5)
+})
+
+# Read in main datafile
+all.data <- fread("processed_data/imported.data.csv")
+
+# Merge in the stimulus (video) information
+setkey(all.data, "Segment")
+setkey(vid.info, "Segment")
+all.data <- vid.info[all.data]
+# Subset the data to only include target stimuli
+all.data <- all.data[StimulusGroup == "target"]
+
+# Collapse male-looks and female-looks into a single look-direction column
+# F and M are each TRUE or FALSE (coerced to 1 or 0)
+all.data[, Look.Dir := ifelse(
+	F.looks > M.looks, "F", ifelse(
+	M.looks > F.looks, "M", "0"))]
+
+# Write out normal data up until now
+write.csv(all.data,
+    paste(processed.data.path, "all.data.before.analysis.bh.csv", sep=""),
+    row.names = FALSE)
+
+# Re-order the data table
+setkeyv(all.data, c("Subject", "Segment", "TimeSec"))
+
+all.data <- smoothLD(all.data)
+
+# Write out the prepped data for input into the transition analysis
+write.csv(all.data,
+    paste(processed.data.path, "prepped.data.bh.csv", sep=""),
+    row.names=FALSE)
+
[…]

### 2016-01-29 analysis/E2/Extras.R

@@ -1,98 +1,0 @@
-switch.rand <- filter(switch.all, Sample != "Transition")
-switch.real <- filter(switch.all, Sample == "Transition") %>%
-               select(-Sample)
-switch.real2 <- filter(switch.all, Sample == "Transition")
-randswitchrate <- aggregate(Switch ~ Run, switch.rand, mean)
-realswitchrate <- aggregate(Switch ~ Run, switch.real2, mean)
-switch.rand.C <- filter(switch.rand, Age != "A") %>%
-                 mutate(Age = as.numeric(Age))
-switch.rand.A <- filter(switch.rand, Age == "A")
-
-
-ggplot(randswitchrate, aes(x=Switch)) + geom_density() +
-	geom_point(data=realswitchrate, aes(x=Switch, y=0), size=4, color="black") +
-	geom_point(data=realswitchrate, aes(x=Switch, y=0), size=3, color="red") +
-	ggtitle("Anticipatory switches overall\n")
-
-# Set paths
-processed.data.path <- "processed_data/" #"../../data/E1/processed_data/"
-plot.path <- "plots/" #"plots/"
-
-# Read in all the random runs and combine them
-adu.r.files <- dir(processed.data.path, pattern="adu.coefs.r-[0-9]+.csv")
-chi.r.files <- dir(processed.data.path, pattern="chi.coefs.r-[0-9]+.csv")
-
-adu.r.model <- read.csv(paste(processed.data.path, adu.r.files[1], sep=""))
-chi.r.model <- read.csv(paste(processed.data.path, chi.r.files[1], sep=""))
-adu.model <- data.table(read.csv(paste(processed.data.path, pattern="adu.coefs-0.csv", sep="")))
-chi.model <- data.table(read.csv(paste(processed.data.path, pattern="chi.coefs-0.csv", sep="")))
-
-# Initialize an empty data table
-aduN <- length(adu.r.files)+1
-chiN <- length(chi.r.files)+1
-
-aduDT <- data.table(
-	matrix(0, aduN, length(adu.r.model),
-	dimnames=list(c(), colnames(adu.r.model))))
-aduDT$Sample <- rep("", aduN)
-aduDT$Error <- rep("", aduN)
-
[…]

### 2016-01-29 analysis/E2/puppetpair_control_study/0-find.transitions.R

@@ -1,0 +1,196 @@
+# Returns a data frame of transition info (whether or not switches were made
+# and, if so, when they started) for all subjects and transitions
+grab.transition.info <- function(obs, fixation.window) {
+    # Preset the total number of rows needed
+    subs <- unique(obs$Subject)
+    gaps <- sort(as.integer(unique(obs$GapNum)))
+    N <- length(subs) * length(gaps)
+    # Make the empty data frame
+    proportion.data <- data.frame(Subject=rep(NA, N), Gap=rep(NA, N),
+        Switch=rep(NA, N), Start.Time=rep(NA, N))
+    # Iterate through data frame rows, filling in transition information for
+    # each subject and gap
+    df.row = 1
+    # Grab a subject's data
+    for (i in 1:length(subs)) {
+        sub.data <- subset(obs, Subject == subs[i])
+        # Grab data for a particular gap
+        for (j in 1:length(gaps)) {
+            # Find the transition information (is there a switch? where?)
+            gap.transition <- find.transition(subset(sub.data,
+                GapNum == gaps[j]), fixation.window)
+            # Add the result to the data frame
+            proportion.data[df.row,] <- c(as.character(subs[i]),
+                gaps[j], gap.transition[1], gap.transition[2])
+            # Move onto the next row in the ouput data frame
+            df.row = df.row + 1
+        }
+    }
+    return(proportion.data)
+}
+
+
+# Parent function, linking the results of fixation.check and
+# find.transition.start together
+find.transition <- function(obs, fixation.window) {
+    if(nrow(obs) < 1) {
+        return(list(NA,NA))
+    }
+    if(is.numeric(max(obs$TimeSec)) == FALSE) {
+        return(list(NA,NA))
+    }
+    # Dummy minstart
+    minstart <- "NA"
+    # Search for a 100 msec fixation on the origin speaker during
+    # the ANCHOR period
+    anchor <- fixation.check(subset(obs, Anchor.Window == "ANCHOR"), "ANCHOR",
+        fixation.window, minstart)
+    minstart <- anchor[2]
+    # Search for a 100 msec fixation on the response speaker during
+    # the TRANSITION period
+    transition <- fixation.check(subset(obs, Transition.Window == "TRANSITION"),
+        "TRANSITION", fixation.window, minstart)
+    # Return the results of the search
+    if(anchor[1] == "Y" & transition[1] == "Y") {
+    	# If you found fixations on the ANCHOR and TRANSITION, find the starting
[…]

### 2016-01-29 analysis/E2/puppetpair_control_study/0-helper.R

@@ -1,0 +1,101 @@
+# Reads in the raw tracker data files and converts some basic raw values for us.
+read.data <- function(path,file.name) {
+    # E.g., A-ADU-P01.txt (for condition A, ADULT, and participant 01)
+    name.parts <- unlist(strsplit(file.name, "[-_.]"))
+    data <- read.csv(paste(path,file.name,sep=""))
+    # Within the tracker data for this participant...
+    data <- within(data, {
+        L.POR.X..px. <- to.n(L.POR.X..px.)
+        R.POR.X..px. <- to.n(R.POR.X..px.)
+        L.POR.Y..px. <- to.n(L.POR.Y..px.)
+        R.POR.Y..px. <- to.n(R.POR.Y..px.)
+        # Average x and y positions
+        x.pos <- rowMeans(data[, c("L.POR.X..px.", "R.POR.X..px.")])
+        y.pos <- rowMeans(data[, c("L.POR.Y..px.", "R.POR.Y..px.")])
+        # Object.Hits are when the eyes are looking in an area of interest
+        F.looks <- L.AOI.Hit == "FEMALE" | R.AOI.Hit == "FEMALE"
+        M.looks <- L.AOI.Hit == "MALE" | R.AOI.Hit == "MALE"
+        # Events are what the eyes are doing (e.g., fixating, saccading, etc.)
+        Events <- L.Event.Info != "Blink" & L.Event.Info != "-" &
+            R.Event.Info != "Blink" & R.Event.Info != "-"
+        Saccades <- L.Event.Info == "Saccade" | R.Event.Info == "Saccade"
+        Fixations <- L.Event.Info == "Fixation" | R.Event.Info == "Fixation"
+        # Add a version/agegroup/subject column based on the file name
+        Condition <- rep(name.parts[1], nrow(data))
+        Subject <- rep(paste(name.parts[1], name.parts[2], sep = "-"), nrow(data))
+        # Make the time into seconds
+        TimeSec <- round((Time - Time[1])/(1000000), 3)
+        MS.Increment <- c(0, diff(TimeSec))
+        ### Add a column of times for each video segment
+        TimeSecSeg <- rep(NA, nrow(data))
+        videos = unique(data$Stimulus)
+        for (video in videos[videos != "-"]) {
+          curr.time = 0
+          sub.indx <- which(data$Stimulus == video)
+          MS.Increment[sub.indx[1]] = 0
+          for (indx in sub.indx) {
+              curr.time = curr.time + MS.Increment[indx]
[…]

### 2016-01-29 analysis/E2/puppetpair_control_study/1-read.dataset.R

@@ -1,0 +1,31 @@
+rm(list = ls())
+source("0-useful.R")
+source("0-helper.R")
+
+raw.data.path <- "tracker_data/"
+info.path <- "info/"
+processed.data.path <- "processed_data/"
+
+files <- dir(raw.data.path, pattern="*.txt")
+all.data <- data.frame()
+
+# Loop through the files, grabbing the data we want
+for (file.name in files) {
+    # Print an update
+	print(file.name)
+	data <- read.data(raw.data.path, file.name)
+    # Take only the columns we want for further analysis
+	data <- subset(data, select = c("Time", "x.pos", "y.pos", "F.looks",
+	    "M.looks", "Events", "Condition", "Subject", "TimeSec","TimeSecSeg",
+	    "MS.Increment", "Stimulus", "Saccades", "Fixations"))
+    # Bind it together
+	all.data <- rbind(all.data, data)
+}
+
+# Rename one column
+colnames(all.data)[12] = "Segment"
+
+all.data <- subset(all.data, Segment != "")
+
+# Save the imported data
+write.csv(all.data, paste(processed.data.path,"imported.data.csv",sep=""), row.names=FALSE)

### 2016-01-29 analysis/E2/puppetpair_control_study/2-prep.analysis.R

@@ -1,0 +1,166 @@
+rm(list = ls())
+source("0-useful.R")
+source("0-helper.R")
+source("0-find.transitions.R")
+
+raw.data.path <- "tracker_data/"
+info.path <- "info/"
+processed.data.path <- "processed_data/"
+
+# Read in supplemental data and prep
+vid.info <- read.csv(paste(info.path,"VideoSegmentInfo.csv",sep=""))
+turn.info <- read.csv(paste(info.path,"TurnInfo.csv",sep=""))
+subject.info <- read.csv(paste(info.path,"SubjectInfo.csv",sep=""))
+gap.info <- read.csv(paste(info.path,"GapInfo.csv",sep=""))
+
+# Round the time values for turns and gaps
+turn.info = within(turn.info, {
+  Onset = round(Onset, 3)
+  Offset = round(Offset, 3)
+})
+gap.info = within(gap.info, {
+  Onset = round(Onset, 3)
+  Offset = round(Offset, 3)
+})
+
+
+################################################################################
+# First we'll merge the gaze data with the information about the stimuli
+
+# Read in main datafile
+all.data <- read.csv(paste(processed.data.path,"imported.data.csv",sep=""))
+
+# Add in the video.info data, keep only target videos, and re-order the data
+all.data <- merge(all.data, vid.info, by = "Segment")
+all.data <- all.data[order(all.data$Subject, all.data$TimeSecSeg),]
+
+# Merge in turn sites, using onsets
+all.data$Onset <- rep(NA, nrow(all.data))
+for (vid in unique(all.data$Segment)) {
+    # For each TimeSecSeg measurement, find the onset of the segment
+    # (from turn.info) within which it lies
+    onsets <- turn.info$Onset[turn.info$Segment == vid]
+    # REMINDER: example use of findInterval:
+    # findInterval(c(2,5,7,8), c(4,5,6,8,9,10)) => [1] 0 2 3 4
+    # In this case it returns a list as long as all.data$TimeSecSeg, with
+    # the index of the video onset that the each time measurement falls within
+    # (e.g., t = 5, vid[2] and t=7, vid[3] in the above example)
+    vid.onset.indices <- findInterval(
+        all.data$TimeSecSeg[all.data$Segment == vid], onsets)
+    # Maps these values onto the subset of times it was derived from
+    all.data$Onset[all.data$Segment == vid] <- onsets[vid.onset.indices]
+}
+# Add in turn.info data and re-order the main data frame
[…]

### 2016-01-29 analysis/E2/puppetpair_control_study/3-plots.stats.analysis.R

@@ -1,0 +1,255 @@
+rm(list = ls())
+library(ggplot2)
+library(lme4)
+library(gridExtra)
+library(scales)  
+library(Hmisc)
+source("0-useful.R")
+
+raw.data.path <- "tracker_data/"
+info.path <- "info/"
+processed.data.path <- "processed_data/"
+plot.path <- "plots/"
+
+# Color and line palettes
+puppetPalette <- c("gray20", "gray40", "gray80", "white")
+puppetPaletteline <- c("black", "black", "black", "black")
+
+# Default setting for all.data in case you don't need the table values and
+# aren't actually loading all.data
+all.data <- 0
+
+# Takes forever to load, only uncomment if you want the table values for looks 
+# to current speaker (at the top)
+all.data <- read.csv(paste(processed.data.path, "all.data.before.analysis.csv",
+    sep=""))
+
+# Read in gaze switch proportions
+# Actual data
+switch.final <- read.csv(paste(processed.data.path, "switch.final.csv", sep=""))
+switch.all <- switch.final
+
+# Read in cumulative gaze switch data
+cumulative.data <- read.csv(paste(processed.data.path,
+    "cumulative.conditions.age.c.csv", sep=""))
+
+
+################################################################################
+###### TABLE VALUES: LOOKS TO CURRENT SPEAKER ##################################
+if (length(all.data) > 1) {
+    sp.f <- subset(all.data, True.Sp.F == "1")
+    sp.m <- subset(all.data, True.Sp.M == "1")
+    # Find looks...
+    # To current speaker
+    spk <- nrow(subset(sp.f, F.looks == TRUE)) +
+        nrow(subset(sp.m, M.looks == TRUE))
+    # To current addressee
+    add <- nrow(subset(sp.f, M.looks == TRUE)) +
+        nrow(subset(sp.m, F.looks == TRUE))
+    # To elsewhere on screen
+    elsew <- nrow(subset(sp.f, M.looks == FALSE & F.looks == FALSE &
+        Events == TRUE)) + nrow(subset(sp.m, M.looks == FALSE &
+        F.looks == FALSE & Events == TRUE))
+    # Off screen
+    lost <- nrow(subset(sp.f, M.looks == FALSE & F.looks == FALSE &
[…]

### 2016-01-29 analysis/get.random.runs.R

@@ -1,5 +1,0 @@
-source("2-find.switches.R")
-
-ns <- as.integer(commandArgs(trailingOnly = TRUE))
-make.rand.gaps(ns)
-get.switches(ns, random = TRUE)

### 2016-01-29 analysis/run.models.R

@@ -1,4 +1,0 @@
-source("3-plots.stats.analysis.R")
-
-ns <- as.integer(commandArgs(trailingOnly = TRUE))
-run.models(ns)

### 2016-02-01 paper/Letters/ResponseLetter-Jan2016.tex
Minor changes to cover letter, manuscript title page, and highlights.

--- 2016-01-29 paper/Letters/ResponseLetter-Dec2015.tex
@@
 \noindent I hereby resubmit a manuscript on ``The development of children's ability to track and predict turn structure in conversation'' to the \textit{Journal of Memory and Language}.
-I apologize for the delay in getting the manuscript resubmitted.
+Please note that Prof.
+Dr.
+Pickering has volunteered to continue handling this revised manuscript.
+
+\noindent I apologize for our delay in getting the manuscript resubmitted.
 It took quite a long time to implement one of the suggestions from Reviewer 3, mostly because I had to acquire some new technical skills in order to carry out his/her suggested analysis on our computation-heavy data set.
@@
 
-\noindent Please note that, because of a miscommunication, I prematurely resubmitted the manuscript a few weeks ago (without my co-author's final sign-off).
-I immediately withdrew the manuscript when our miscommunication came to light and we have since then made minor improvements to the paper.
-With my co-author's final approval, I am ready to return the paper to JML.
-However, now my only option is to resubmit the manuscript as ``new''.
-If it is at all possible, we would request that the paper go back to the original reviewers, if only because we did quite a lot of work to address their specific (and substantial) requests.
-In case this is possible, I have included our responses to the reviewers below.
-
-\noindent For example, in addition to the new random permutation analysis suggested by Reviewer 3, we have also included two further follow-up analyses in the Appendix, each inspired by the other reviewers' comments.
-We have also tried to make the text much more succinct, with distinct and straightforward results and discussion sections, following requests from all three reviewers.
+\noindent Since our last submission, we have done quite a lot of work to improve our paper by addressing the reviewers' comments.
+For example, in addition to the new random permutation analysis suggested by Reviewer 3 (mentioned above), we have also included two further follow-up analyses in the Appendix (including additional experimental data), each inspired by the other reviewers' comments.
+We have also tried to make the text much more succinct, with separate and straightforward sections for the results and discussions, following requests from all three reviewers.
 
 \noindent We hope that, with these changes, we have greatly improved the paper's quality.
-Please find a description of specific comments addressed below.
+Please find a description of the specific comments addressed below.
 Thank you very much for your reconsideration.
@@
 \begin{tabular}{ll}
-\textit{ID:} & JML-14-235 (from the previous submission) \\
+\textit{ID:} & JML-14-235R2 \\
 \textit{Title:} & The development of children's ability to track and \\
@@
 \noindent We thank all three reviewers for their thorough and helpful comments on the previously submitted manuscript.
-We have tried to address all of their comments, in the manuscript and here in the response letter.
+We have tried to address all of their comments, both in the manuscript and here in the response letter.
 

### 2016-02-01 paper/CasillasFrank-Anticip.tex
Minor changes to cover letter, manuscript title page, and highlights.

@@
 %\address[StanfordLX]{Department of Linguistics, Stanford University}
-\cortext[cor1]{Corresponding author}
+\cortext[cor1]{Corresponding author. \\ Address: Wundtlaan 1, 6525 XD, Nijmegen, The Netherlands\\ Email: marisa.casillas@mpi.nl \\Telephone: +31 024 3521 566; Fax: +31 024 3521 213}
 
@@
 \begin{keyword}
-Turn taking \sep Conversation \sep Development \sep Prosody \sep Questions \sep Eye-tracking
+Turn taking \sep Conversation \sep Development \sep Questions \sep Eye-tracking \sep Anticipation
 %% MSC codes here, in the form: \MSC code \sep code
@@
 
-\bigskip
+\newpage
 \noindent \textbf{Methods}

### 2016-06-05 paper/CasillasFrank-Anticip.tex
most changes for resubmission

@@
 \usepackage{rotating}
-
-\linespread{2}
+%\usepackage{apacite}
+
+%\linespread{2}
 \journal{}
@@
 In both experiments, children and adults made more anticipatory switches after hearing questions.
-Like adults, prosody alone did not improve children's predictive gaze shifts.
-But, unlike adults, lexical cues alone were not sufficient to improve prediction---children's performance was best overall with lexicosyntax and prosody together.
-Our findings support an account in which turn prediction emerges in infancy, but then only gradually becomes fully integrated with linguistic processing.
+Consistent with prior findings on adult turn prediction, prosodic information alone did not increase children's anticipatory gaze shifts.
+But, unlike prior work with adults, lexical information alone was not sufficient either---children's performance was best overall with lexicosyntax and prosody together.
+Our findings support an account in which turn tracking and prediction emerges in infancy, and then only gradually becomes integrated with linguistic processing.
 \end{abstract}
@@
 
-\section{Introduction}
+\section*{Introduction}
 \label{sec:intro}
@@
 
-\subsection{Adult turn taking}
+\subsection*{Adult turn taking}
 
@@
 With pitch-flattened speech, the timing of participants' button responses was comparable to their timing with the full linguistic signal.
-But when no lexical information was available, participants' responses were significantly earlier.
+But when no lexical information was available, participants' responded significantly earlier within the turn.
 The authors concluded that lexicosyntactic information\footnote{The ``lexicosyntactic'' condition only included flattened pitch and so was not exclusively lexicosyntactic---the speech would still have residual prosodic structure, including syllable duration and intensity.} was necessary and possibly sufficient for turn-end projection, while intonation was neither necessary nor sufficient.
@@
 
-\subsection{Turn taking in development}
+\subsection*{Turn taking in development}
 
@@
 But by nine months, their timing slows down considerably, only to gradually speed up again after 12 months \citep{hilbrink2015}.
-For children, taking turns with brief transitions between speakers is more difficult than avoiding speaker overlap; children's incidence of overlap is nearly adult-like by nine months, but the timing of their non-overlapped responses stays much longer than the adult 200 msec standard for the next few years \citep{casillas2016, garvey1984, garvey1981, ervin-tripp1979}.
-This puzzling pattern is likely due to their linguistic development: taking turns on time is easier when the response is a simple vocalization rather than a linguistic utterance.
-Integrating language into the turn-taking system may therefore be one major factor in children's delayed responses \citep{casillas2016}.
-
-Children, like adults, might use linguistic cues in the ongoing turn to make predictions about upcoming turn structure.
+For children, taking turns with brief transitions between speakers is more difficult than avoiding speaker overlap; children's incidence of overlap is nearly adult-like by nine months, but the timing of their non-overlapped (i.e., gapped) responses remains longer than the adult 200 msec standard for the next few years \citep{casillas2016, garvey1984, garvey1981, ervin-tripp1979}.
+This puzzling pattern is likely due to children's linguistic development: taking turns on time is easier when their response is a simple vocalization rather than a linguistic utterance.
+Integrating language into the turn-taking system may therefore be a major factor in children's delayed responses \citep{casillas2016}.
+
+Before children manage to integrate linguistic cues into their turn-taking behaviors (for both turn prediction and production), they can rely on non-verbal interactional cues, including silence (e.g., between turns), eye gaze, body orientation, and gesture, to identify the boundaries of social actions.
+For example, with little to no linguistic knowledge, children are often able to infer and anticipate desired responses to offers and requests by taking account of their interlocutor's non-verbal communicative behavior, the structure of routine events, and the affordances of the current interactional context \citep{reddy13, nomikou11, shatz78}.
+With respect to turn taking in particular, children's spontaneous vocalizations during interaction demonstrate a sensitivity to short inter-speaker gaps from infancy \citep{hilbrink2015}, even when they are not yet using language to participate.
+Thus, before they can anticipate turn structure form linguistic cues, children might instead react to silence as a cue to upcoming speaker change.
+Interactional silence itself may then serve as an one of children's first cues to turn structure, giving them information about when to respond before they can rely on language.
+
+As children's language competence increases, they can use linguistic cues to make predictions about upcoming turn structure.
 Studies of early linguistic development point to a possible early advantage for prosody over lexicosyntax in children's turn-taking predictions.
@@
 They asked 6-, 12-, 24-, and 36-month-old infants, and adult participants to watch short videos of conversation and tracked their eye movements at points of speaker change.
-They showed their participants two types of conversation videos---one normal and one with flattened pitch (i.e., with flattened intonation contours)---to test the role of intonation in participants' anticipatory predictions about upcoming speech.
-Comparing children's anticipatory gaze frequency to a random baseline, they found that only 36-month-olds and adults made anticipatory gaze switches more often than expected by chance.
-Among those, only 36-month-olds were affected by a lack of intonation contours, leading Keitel and colleagues to conclude that children's ability to predict upcoming turn structure relies on their ability to comprehend the stimuli lexicosemantically.
+They showed their participants two types of conversation videos---one normal and one with flattened pitch---to test the role of intonation in participants' anticipatory predictions about upcoming speech.
+Comparing children's anticipatory gaze frequency to a random baseline, they found that only 36-month-olds and adults made anticipatory gaze switches more often than expected by chance, and only 36-month-olds were affected by a lack of intonation contours.
+This finding led Keitel and colleagues to conclude that children's ability to predict upcoming turn structure relies on their ability to comprehend the stimuli lexicosemantically.
 They also suggest that intonation might play a secondary role in turn prediction, but only after children acquire more sophisticated, adult-like language comprehension abilities (also see \citealp{keitel2015}).
@@
 \citeyearpar{keitel2013} study still demonstrates intriguing age-based differences in children's ability to predict upcoming turn structure.
-Our current work takes this paradigm as a starting point.\footnote{But also see \citet{casillas2012, casillas2013}.}
-
-\subsection{The current study}
+Our current work takes these findings as a starting point.\footnote{But also see \citet{casillas2012, casillas2013}.}
+
+\subsection*{The current study}
 
@@
 We tested children across a wide range of ages (Experiment 1: 3--5 years; Experiment 2: 1--6 years), with adult control participants in each experiment.
-
-We highlight three primary findings: first, although children and adults use linguistic cues to make predictions about upcoming turn structure, they do so primarily in predict speaker transitions after questions (a speech act effect).
+We additionally tested for the use of one non-verbal cue: inter-turn silence.
+
+We highlight four primary findings: first, although children and adults use linguistic cues to make predictions about upcoming turn structure, they do so primarily to predict speaker transitions after questions (a speech act effect).
 This intriguing effect, which has not been reported previously, suggests that participants track unfolding speech for cues to upcoming speaker change, which may affect how they use linguistic cues more generally for anticipatory processing in conversation.
-Second, we find that children make more predictions than expected by chance starting at age two, but that this effect is small at first, and continues to improve through age six.
-Third, we find no evidence of an early prosody advantage in children's anticipations and, further, no evidence that prosodic or lexical cues alone can substitute for their combination in the full linguistic signal (as is proposed for adults; \citealp{de-ruiter2006}).
-Instead, anticipation is strongest for stimuli with the full range of cues.
-In sum, our findings support an account in which turn prediction emerges in infancy, but becomes fully integrated with linguistic processing only gradually across development.
-
-\section{Experiment 1}
+Second, we find that children make more predictions than expected by chance starting at age two, but that this effect is small at first, and continues to improve through age six, along with children's use of linguistic cues to anticipate answers after question turns.
+Third, children and adults at all ages often used inter-turn silence (a non-verbal cue to turn structure) to make more predictive gaze switches to the responder, suggesting that non-verbal cues are useful for predicting turn structure early on and continue to be important in adulthood.
+Finally, we find no evidence for an early prosodic advantage in children's anticipations and, further, no evidence that lexical cues alone are comparable to the full linguistic signal in aiding participants' predictions (as is proposed for adults; \citealp{de-ruiter2006}).
+Anticipation is strongest for stimuli with the full range of linguistic cues.
+Our findings support an account in which turn prediction emerges in infancy, but becomes fully integrated with linguistic processing only gradually across development.
+
+\section*{Experiment 1}
 \label{sec:exp1}
@@
 The participants, all native English speakers, were only expected to understand the two videos in English.
-We showed participants non-English videos to limit their access to lexical information while maintaining their access to other cues to turn boundaries (e.g., (non-native) prosody, gaze, inbreaths, phrase final lengthening).
+We showed participants non-English videos to limit their access to lexical information while maintaining their access to other cues to turn boundaries (e.g., non-English prosody, gaze, in-breaths, phrase final lengthening).
 Using this method, we compared children and adult's anticipatory looks from the current speaker to the upcoming speaker at points of turn transition in English and non-English videos.
 
-\subsection{Methods}
+\subsection*{Methods}
 \label{sec:methods1}
 
-\subsubsection{Participants}
-
-We recruited 74 children between ages 3;0--5;11 and 11 undergraduate adults to participate in the experiment.
+\subsubsection*{Participants}
+
+We recruited 74 children ages 3;0--5;11 and 11 undergraduate adults to participate in the experiment.
+We recruited adult participants through the Stanford University Psychology participant database.
+Adult participants were either paid or received course credit for their time.
 Our child sample included 19 three-year-olds, 32 four-year-olds, and 23 five-year-olds, all enrolled in a local nursery school.
-All participants were native English speakers.
-Approximately one-third (N=25) of the children's parents and teachers reported that their child regularly heard a second (and sometimes third or further) language, but only one child frequently heard a language that was used in our non-English video stimuli, and we excluded his data from analyses.
-None of the adult participants reported fluency in a second language.
+All participants were native English speakers and volunteered their time.
+Approximately one-third (N$=$25) of the children's parents and teachers reported that their child regularly heard a second (and sometimes third or further) language, but only one child frequently heard a language that was used in our non-English video stimuli, and we excluded his data from the analyses.\footnote{Multilingual children may make predictions about upcoming turn structure differently from their monolingual peers due to their more varied experiences with language-relevant turn taking cues, but we are unable to test this hypothesis here due to the variability in multilingual language input and languages being learned in our sample.
+The same note applies to Experiment 2 below.} None of the adult participants reported fluency in a second language.
 
@@
 
-\subsubsection{Materials}
+\subsubsection*{Materials}
 
@@
 We marked non-English turns as questions when at least 4 of the 5 listeners (80\%) said that the turn ``sounded like a question''.
+This procedure means that the cues for recognizing ``questions'' in the non-English condition only resembled native English question cues, and therefore were likely harder to identify than cues to questionhood in the English condition.
+However, since participants did not speak the non-English languages and would only therefore ever treat ``question-sounding'' turns as questions, we proceeded with these analyses to see how pervasive question effects were---could they show up even without lexical access in a foreign language?
+If participants primarily rely on prosodic cues to question turns, it's possible that even non-English prosody can elicit anticipatory gaze switches for question-like turns.
 
@@
 We measured the duration of each turn transition from the audio recording associated with each video.
-We excluded turn transitions longer than 550 msec and shorter than 90 msec, also excluding overlapped transitions, from analysis.\footnote{Overlap occurs when a responder begins a new turn before the current turn is finished.
+We excluded turn transitions longer than 550 msec and shorter than 90 msec from analysis, additionally excluding overlapped transitions.\footnote{Overlap occurs when a responder begins a new turn before the current turn is finished.
 When overlap occurs, observers cannot switch their gaze in anticipation of the response because the response began earlier than expected.
@@
 On average, the inter-turn gaps for English videos (mean$=$318, median$=$302, stdev$=$112 msec) were slightly longer than for non-English videos (mean$=$286, median$=$251, stdev$=$122 msec).
-The longer gaps in the English videos could give them a slight advantage: our definition of an ``anticipatory gaze shift'' includes shifts that are initiated during the gap between turns (Figure \ref{fig:criterion}), so participants had slightly more time to make anticipatory shifts in the English videos.
 
@@
 
-\subsubsection{Procedure} 
+\subsubsection*{Procedure} 
 Participants sat in front of an SMI 120Hz corneal reflection eye-tracker mounted beneath a large flatscreen display.
@@
 
-\subsubsection{Data preparation and coding}
+\subsubsection*{Data preparation and coding}
 \label{sec:algorithm}
@@
 
-We checked each participant's gaze at each turn transition for three characteristics (Figure \ref{fig:criterion}): (1) that the participant fixated on the prior speaker for at least 100 msec at the end of the prior turn, (2) that sometime thereafter the participant switched to fixate on the upcoming speaker for at least 100 ms, and (3) that the switch in gaze was initiated within the first 200 msec of the response turn, or earlier.
+We checked each participant's gaze at each turn transition for three characteristics (Figure \ref{fig:criterion}): (1) that the participant fixated on the prior speaker for at least 100 msec at the end of the prior turn, (2) that immediately thereafter the participant switched to fixate on the upcoming speaker for at least 100 ms, and (3) that the switch in gaze was initiated within the first 200 msec of the response turn, or earlier.
 These criteria guarantee that we only counted gaze shifts when: (1) participants were tracking the previous speaker, (2) switched their gaze to track the upcoming speaker, and (3) did so before they could have simply reacted to the onset of speech in the response.
@@
 Early switches are consistent with the distribution of responses in explicit turn-boundary prediction tasks.
-For example, in a button press task, adult participants anticipate turn ends approximately 200 msec in advance of the turn's end, and anticipatory responses to pitch-flattened stimuli come even earlier \citep{de-ruiter2006}.
+For example, in a button press task, adult participants anticipated turn ends approximately 200 msec in advance of the turn's end, and anticipatory responses to pitch-flattened stimuli came even earlier \citep{de-ruiter2006}.
 We therefore allowed switches to occur as early as 200 msec before the end of the prior turn.
@@
 Our third prediction was that anticipatory looks would increase with development, along with children's increased linguistic competence.
-
-\subsection{Results}
+Finally, we predicted that transitions with longer inter-turn gaps would show greater anticipation because longer gaps provide (a) more time to make a gaze switch and (b) are themselves a cue to possible upcoming speaker switch.
+
+\subsection*{Results}
 \label{sec:results1}
@@
 \end{table}
-\linespread{2}
+%\linespread{2}
 
@@
 \end{center}
-\caption{Anticipatory gaze rates across language condition and transition type for the real data (red and blue) and the randomly permuted baselines (gray).
-Vertical bars represent 95\% confidence intervals.}
+\caption{Anticipatory gaze rates across language condition and transition type for the real and randomly permuted datasets. Vertical bars represent 95\% confidence intervals.} 
 \label{fig:E1-randvsreal}
@@
     \hline
-    (Intercept)   										& -0.96145	& 0.84915		& -1.132	& 0.257531 	\\    
-    Age             										& -0.18268	& 0.17509		& -1.043	& 0.296764   	\\  
-    LgCond$=$\textit{non-English} 			& -3.29349	& 0.96055		& -3.429	& 0.000606 *** 	\\
-    Type$=$\textit{non-Question}      	& -1.10131	& 0.86520		& -1.273	& 0.203055     	\\
-    GapDuration      										& 3.40171	& 1.22878		&  2.768	& 0.005634  **	\\
-    Age*LgCond$=$\textit{non-English} 	& 0.52066	& 0.21192		&  2.457	& 0.014015 * 	\\  
-    Age*Type$=$\textit{non-Question} 	& -0.01628	& 0.19442		& -0.084	& 0.933262     	\\
-    LgCond$=$\textit{non-English}*			& 2.68171	& 1.35045		& 1.986	& 0.047057 * 	\\
+    (Intercept)   													& -0.604	& 1.242	& -0.486  & 0.627		\\
+    Age             												& -0.002	& 0.261	& -0.009  & 0.993		\\
+    LgCond$=$\textit{non-English} 						& -3.65		& 1.16	& -3.146  & 0.002 **	\\
+    Type$=$\textit{non-Question}      					& -2.95		& 1.13	& -2.61	  & 0.009 **	\\
+    GapDuration      											&  2.247	& 3.194	&  0.704  & 0.482		\\
+    Age*LgCond$=$\textit{non-English} 					&  0.5		& 0.212	&  2.353  & 0.019 *	\\
+    Age*Type$=$\textit{non-Question} 					&  0.009	& 0.196	&  0.044  & 0.965		\\
+    LgCond$=$\textit{non-English}*						&  2.692	& 1.347	&  1.999  & 0.046 *	\\
     \hspace*{5mm} Type$=$\textit{non-Question} &&&& \\
-    Age*LgCond$=$\textit{non-English}*	& -0.45633	& 0.30168		& -1.513	& 0.130378  \\
+    Age*GapDuration												& -0.577	& 0.627	& -0.921  & 0.357		\\
+    LgCond$=$\textit{non-English}*GapDuration		&  1.143	& 2.287	&  0.5	  & 0.617		\\
+    Type$=$\textit{non-Question}*GapDuration			&  5.519	& 2.282	&  2.418  & 0.016 *	\\
+    Age*LgCond$=$\textit{non-English}*					& -0.433	& 0.304	& -1.426  & 0.154		\\
     \hspace*{5mm} Type$=$\textit{non-Question} &&&& \\
@@
     \hline
-    (Intercept)												& -0.1966		& 0.6945		& -0.283	& 0.777062     	\\
-    LgCond$=$\textit{non-English}				& -0.8812		& 0.9607		& -0.917	& 0.359028   	\\  
-    Type$=$\textit{non-Question}				& -4.4953		& 1.3147		& -3.419	& 0.000628 *** \\
-    GapDuration													& -1.1227		& 1.9889		& -0.565	& 0.572414     	\\
-    LgCond$=$\textit{non-English}*			& 3.2972			& 1.6115		& 2.046	& 0.040747 * 	\\ 
+    (Intercept)														& -0.584   & 0.64  & -0.913 & 0.361			\\
+    LgCond$=$\textit{non-English}							& -0.059   & 0.751 & -0.079 & 0.937			\\
+    Type$=$\textit{non-Question}							& -3.298   & 0.933 & -3.536 & 0.0004 ***	\\
+    GapDuration														&  0.132   & 1.766 &  0.075 & 0.941			\\ 
+    LgCond$=$\textit{non-English}*						&  1.234   & 0.629 &  1.961 & 0.0498 *		\\
     \hspace*{5mm} Type$=$\textit{non-Question} &&&& \\
-    LgCond$=$\textit{non-English}*			& 1.3625			& 3.0097		& 0.453	& 0.650749     	\\
-    \hspace*{5mm} GapDuration &&&& \\
-    Type$=$\textit{non-Question}*				& 10.5107		& 3.3482		& 3.139	& 0.001694 ** 	\\
-    \hspace*{5mm} GapDuration &&&& \\
-    LgCond$=$\textit{non-English}*			& -6.3156		& 4.4969		& -1.404	& 0.160191 	\\    
-    \hspace*{5mm} Type$=$\textit{non-Question}* &&&& \\
-    \hspace*{5mm} GapDuration &&&& \\
+    LgCond$=$\textit{non-English}*GapDuration		& -1.519   & 2.192 & -0.693 & 0.488			\\
+    Type$=$\textit{non-Question}*GapDuration			&  7.116   & 2.195 &  3.241 & 0.001 **		\\
     \hline
@@
 \end{table}
-\linespread{2}
-
-\subsubsection{Statistical models}
+%\linespread{2}
+
+\subsubsection*{Statistical models}
 \label{sec:models1}
@@
 We would need multiple items from each of the languages to check for similarity effects of specific linguistic features.}, transition type (question vs.
-non-question), age (3, 4, 5; numeric), and duration of the inter-turn gap (seconds, e.g., 0.441) as predictors, with full interactions between condition, transition type, and age.
-We included the duration of the inter-turn gap as a control predictor since longer gaps provide more opportunities to make anticipatory switches (Figure \ref{fig:criterion}).
-We additionally included random effects of item (turn transition) and participant, with random slopes of condition, transition type, and their interaction for participants \citep{barr2013}.\footnote{The models we report are all qualitatively unchanged by the exclusion of their random slopes.
-We have left the random slopes in because of minor participant-level variation in the predictors modeled.} The adult model included condition, transition type, duration, and their interactions as predictors with participant and item included as random effects and random slopes of condition, transition type, and their interaction for participant.
-
-Children's anticipatory gaze switches showed effects of language condition (\textit{$\beta$}=-3.29, \textit{SE}=0.961, \textit{z}=-3.43, \textit{p}$<$.001) and gap duration (\textit{$\beta$}=3.4, \textit{SE}=1.229, \textit{z}=2.77, \textit{p}$<$.01) with additional effects of an age-by-language condition interaction (\textit{$\beta$}=0.52, \textit{SE}=0.212, \textit{z}=2.46, \textit{p}$<$.05) and a language condition-by-transition type interaction (\textit{$\beta$}=2.68, \textit{SE}=1.35, \textit{z}=1.99, \textit{p}$<$.05).
-There were no significant effects of age or transition type alone (\textit{$\beta$}=-0.18, \textit{SE}=0.175, \textit{z}=-1.04, \textit{p}$=$.3 and \textit{$\beta$}=-1.10, \textit{SE}=0.865, \textit{z}=-1.27, \textit{p}$=$.2, respectively).
-
-Adults' anticipatory gaze switches showed an effect of transition type (\textit{$\beta$}=-4.5, \textit{SE}=1.315, \textit{z}=-3.42, \textit{p}$<$.001) and significant interactions between language condition and transition type (\textit{$\beta$}=3.3, \textit{SE}=1.61, \textit{z}=2.05, \textit{p}$<$.05) and transition type and gap duration (\textit{$\beta$}=10.51, \textit{SE}=3.348, \textit{z}=3.139, \textit{p}$<$.01).
-
-\subsubsection{Random baseline comparison}
+non-question), age (3, 4, 5; numeric; intercept as age=0), and duration of the inter-turn gap (seconds, e.g., 0.441) as predictors, with two-way interactions between gap duration and the other simple predictors (language condition, transition type, and age), also adding a three-way interaction between language condition, transition type, and age.
+We included the two-way interactions between gap duration and the other predictors in case the effect of inter-turn silence changes with age or linguistic cueing (e.g., if children older children rely less on silence as a cue)\footnote{We test these two-way interactions with gap duration in all of the models reported in this paper.
+Higher-order interactions with gap duration usually resulted in model non-convergence due to distributional sparsity when three or more predictor values were considered.}.
+We also included random effects of item (turn transition) and participant, with maximal random slopes of condition, transition type, and their interaction for participants \citep{barr2013}.\footnote{The models we report in this paper are all qualitatively unchanged by the exclusion of their random slopes.
+We have left the random slopes in because of minor participant-level variation in the predictors modeled.} The adult model included fixed effects of condition, transition type, and their interaction, plus two-way interactions between gap duration and language condition and transition type (as in the child model).
+The adult model also included random effects of item and participant with maximal random slopes of condition, transition type, and their interaction for participant.
+
+Children's anticipatory gaze switches showed effects of language condition (\textit{$\beta$}=-3.65, \textit{SE}=1.16, \textit{z}=-3.15, \textit{p}$<$.01) and transition type (\textit{$\beta$}=-2.95, \textit{SE}=1.13, \textit{z}=-2.61, \textit{p}$<$.01) with additional effects of an age-by-language condition interaction (\textit{$\beta$}=0.5, \textit{SE}=0.212, \textit{z}=2.35, \textit{p}$<$.05), a language condition-by-transition type interaction (\textit{$\beta$}=2.69, \textit{SE}=1.35, \textit{z}=1.99, \textit{p}$<$.05), and an interaction between transition type and gap duration (\textit{$\beta$}=5.52, \textit{SE}=2.28, \textit{z}=2.42, \textit{p}$<$.05).
+There were no significant effects of age or gap duration alone (\textit{$\beta$}=-0.002, \textit{SE}=0.26, \textit{z}=-0.009, \textit{p}$=$.99 and \textit{$\beta$}=2.25, \textit{SE}=3.19, \textit{z}=0.7, \textit{p}$=$.48, respectively).
+
+Adults' anticipatory gaze switches showed an effect of transition type (\textit{$\beta$}=-3.3, \textit{SE}=0.93, \textit{z}=-3.54, \textit{p}$<$.001) and significant interactions between language condition and transition type (\textit{$\beta$}=1.23, \textit{SE}=0.63, \textit{z}=1.96, \textit{p}$<$.05) and transition type and gap duration (\textit{$\beta$}=7.12, \textit{SE}=2.2, \textit{z}=3.24, \textit{p}$<$.01).
+There were no significant effects of language condition or gap duration alone (\textit{$\beta$}=-0.06, \textit{SE}=0.75, \textit{z}=-0.08, \textit{p}$=$.94 and \textit{$\beta$}=0.13, \textit{SE}=1.77, \textit{z}=0.08, \textit{p}$=$.94, respectively).
+
+\subsubsection*{Random baseline comparison}
 \label{sec:randbaseline1}
 
-We estimated the probability that these patterns were the result of random looking by running the same regression models on participants' real eye-tracking data, only this time calculating their anticipatory gaze switches with respect to randomly permuted turn transition windows.
-This process involved: (1) randomizing the order and temporal placement of the analysis windows within each stimulus (see Figure \ref{fig:shuffling}; ``analysis window'' is defined in Figure \ref{fig:criterion}) to randomly redistribute the analysis windows across the eye-tracking signal, (2) re-running each participant's eye tracking data through switch identification (described in Section \ref{sec:algorithm}) on each of the randomly permuted analysis windows, and (3) modeling the anticipatory switches from the randomly permuted data with the same statistical models we used for the original data (Section \ref{sec:models1}; Table \ref{tab:E1-models}).
-Importantly, although the onset time of each transition was shuffled within the eye-tracking signal, the other intrinsic properties of each turn transition (e.g., prior speaker identity, transition type, gap duration, language condition, etc.) stayed constant across each random permutation.
+Our primary analysis (above) makes the assumption that participants' eye movements generally follow the turn structure of the stimulus, i.e., that participants track the current speaker and switch their gaze to the upcoming speaker near turn transitions.
+Based on this assumption, we used linear mixed effects regressions to see how anticipatory looking is affected by aspects of participant group (age) and stimulus (e.g., transition type, language condition).
+But what if the assumption that participants generally track turn structure were wrong?
+Could these results have emerged if participants' eye movements were \textit{not} linked to turn structure?
+For example, if participants were randomly looking back and forth between the two speakers, we might still find some anticipatory switching by chance---we would see anticipatory switching no matter where the real turn transitions occurred.
+To test whether our primary results (the regression output above) could have arisen from random switching, not linked to turn structure, we conducted a secondary analysis comparing participants' anticipatory gaze at \textit{real} and \textit{randomly shuffled} points of turn transition.
+
+We conducted this analysis by running the same regression models on participants' eye-tracking data, only this time calculating their anticipatory gaze switches with respect to randomly permuted turn transition windows.
+This process involved: (1) randomizing the order and temporal placement of the analysis windows within each stimulus (see Figure \ref{fig:shuffling}; ``analysis window'' is defined in Figure \ref{fig:criterion}) to randomly redistribute the analysis windows across the eye-tracking signal, (2) re-running each participant's eye tracking data through switch identification (described in Section \ref{sec:algorithm}) on each of the randomly permuted analysis windows, and (3) modeling the anticipatory switches from the randomly permuted data (our random baseline dataset) with the same statistical models we used for the original dataset (Section \ref{sec:models1}; Table \ref{tab:E1-models}).
+Importantly, although the onset time of each transition was shuffled within the eye-tracking signal, the other intrinsic properties of each turn transition (e.g., prior speaker identity, transition type, gap duration, language condition, etc.) stayed constant across each permutation.
 
@@
 \end{center}
-\caption{Example of analysis window permutations for a stimulus with five turn transitions. The windows were $\pm$300 msec around the inter-turn gap.} 
+\caption{Example of analysis window permutations for a stimulus with five turn transitions. The windows included $\pm$300 msec around the inter-turn gap.} 
 \label{fig:shuffling}
@@
 
-This procedure effectively de-links participants' gaze data from the turn structure in the original stimulus, thereby allowing us to compare turn-related (original) and non-turn-related (randomly permuted) looking behavior using the same eye movement data.
-The resulting anticipatory gazes from the randomly permuted analysis windows represent an average anticipatory gaze rate over all possible starting points: a random baseline.
-
-By running the real and randomly permuted data sets through identical statistical models, we can estimate how likely it is that predictor effects in the original data (e.g., the effect of language condition; Table \ref{tab:E1-models}) arose from random looking.
-Because these analyses are complex, we report their full details in \ref{sec:permutation}.
+The random shuffling procedure de-links participants' gaze data from the turn structure in the original stimulus, thereby allowing us to compare turn-related (original) and non-turn-related (randomly permuted) looking behavior using the same eye movement data.
+We created 5,000 permutations of the original turn transitions (step 1 above), thereby creating 5,000 anticipatory gaze datasets with randomly de-linked gaze data (step 2 above).
+Because the randomly shuffled turn transitions could occur anywhere in the stimulus (so long as they didn't overlap each other), the resulting turn-transition windows collectively covered the entire stimulus---during speech and silence, during speaker change and speaker continuation, and during all turn transitions in the stimulus, even those excluded in the original analyses (e.g., because they were overlapped).\footnote{This technique crucially differs from that used by Keitel and colleagues (2013, 2015), which tests anticipatory gaze at turn transitions against anticipatory gaze during speech, thereby maximizing the possibility of finding a difference from the baseline measure.} Pooled together, the anticipatory gaze datasets yield an average anticipatory switch rate for each participant over all possible starting points in the stimuli: a random baseline.
+
+Using this technique we can compare participants' anticipatory switches at turn transition windows to their anticipatory switches over the stimulus as a whole.
+If participants were looking randomly back and forth between the speakers, we should similar patterns in both cases.
+Rather than simply comparing participants' overall anticipatory switch rates with real and random transition windows, we estimated the likelihood that each of the predictor effects in the original data (e.g., the effect of language condition; Table \ref{tab:E1-models}) could have arisen with random gaze switching: we ran identical statistical models on the real and randomly permuted data sets.
+This tells us not only whether participants' switches were above chance, but whether the specific underlying effects of their anticipatory gaze patterns (e.g., the effect of linguistic condition) were above that expected by chance.
+But, because these analyses are complex and secondary to the main results, we report their full details in \ref{sec:permutation}.
 
 Our baseline analyses revealed that none of the significant predictors from models of the original, turn-related data can be explained by random looking.
-For the children's data, the original \textit{z}-values for language condition, gap duration, the age-language condition interaction, and the language condition-transition type interaction were all greater than 95\% of \textit{z}-values for the randomly permuted data (99.9\%, 95.5\%, 99.4\%, and 96\%, respectively, all \textit{p}$<$.05).
-Similarly, the adults' data showed significant differentiation from the randomly permuted data for two of the three originally significant predictors---transition type and the transition type-gap duration interaction (greater than 99.9\% and 99.7\% of random \textit{z}-values, respectively, all \textit{p}$<$.01)---with marginal differentiation for the interaction of language condition and transition type (greater than 94.6\% of random \textit{z}-values; \textit{p}$=$.054).
-See Section \ref{sec:permutation} for more information on each predictor's random permutation distribution.
-
-\subsubsection{Developmental effects}
-
-The models reported above revealed a significant interaction of age and language condition (Table \ref{tab:E1-models}) that was unlikely be due to random looking (Figure \ref{fig:E1-randvsreal}).
+For the children's data, the original \textit{z}-values for language condition, transition type, the age-language condition interaction, the transition type-gap duration interaction, and the language condition-transition type interaction were all greater than 95\% of \textit{z}-values from models of the randomly permuted data (99.3\%, 99.1\%, 98.9\%, 97\%, and 96\%, respectively, all \textit{p}$<$.05).
+Similarly, the adults' data showed significant differentiation from the randomly permuted data for all three significant predictors from the real transition dataset---transition type, the interaction between transition type and gap duration, and the interaction between language condition and transition type (greater than 100\%, 99.8\%, and 95\% of random \textit{z}-values, respectively, all \textit{p}$\leq$.05).
+See Section \ref{sec:permutation} for more information on each predictor's random permutation distribution.\footnote{This baseline analysis tests ``random looking'' against ''turn-driven looking'', but it does not test subtypes of turn-driven looking.
+For example, children might switch their gaze from the current speaker to the addressee out of boredom with the ongoing speech rather than active anticipation of an upcoming response.
+We address this hypothesis about ``turn-transition'' gaze switchers vs.
+``boredom'' gaze switchers in \ref{sec:boredlooks}}
+
+\subsubsection*{Developmental effects}
+
+The models reported above revealed a significant interaction of age and language condition (Table \ref{tab:E1-models}) that was unlikely be due to random gaze switching (Figure \ref{fig:E1-randvsreal}).
 To further explore this effect, we compared the effect of language condition across age groups: using the permuted datasets described above, we extracted the average difference score for the two language conditions (English minus non-English) for each participant, computing an overall average for each random permutation of the data.
@@
 
-When does spontaneous turn prediction emerge developmentally?
-To test whether the youngest age group (3-year-olds) already exceeded chance in their anticipatory gaze switches, we compared children's real gaze rates to the random baseline in the English condition with two-tailed \textit{t}-tests.
+Given these findings, when does spontaneous turn prediction emerge developmentally?
+We tested whether the youngest age group (3-year-olds) already exceeded chance in their anticipatory gaze switches by comparing children's real gaze rates to the random baseline in the English condition with two-tailed \textit{t}-tests.
 We used the English condition because we are most interested in finding out when children begin to make spontaneous turn predictions for natural speech.
@@
 
-\subsection{Discussion}
+\subsection*{Discussion}
 \label{sec:discussion1}
@@
 First, both adults' and children's anticipations were strongly affected by transition type.
-Both groups made more anticipatory switches after hearing questions, compared to non-questions.
-Even in the English videos, when participants had full access to linguistic cues, their rates of anticipation were relatively low (even comparable to the non-English videos) unless the turn was a question.
+Both groups made more anticipatory switches after hearing questions, compared to non-questions, especially for the English stimuli compared to the non-English stimuli.
+Overall, participants made few anticipatory switches after non-questions, even in the English videos when they had full linguistic access.
 Prior work using online, metalinguistic tasks has shown that participants can use linguistic cues to accurately predict upcoming turn ends \citep{torreira2015, magyari2012, de-ruiter2006}.
@@
 Second, children made more anticipatory switches overall in English videos, compared to non-English videos.
-This effect suggests that lexical access is important for children's ability to anticipate upcoming turn structure, consistent with prior work on turn-end prediction in adults (\citealp{de-ruiter2006}; \citealp{magyari2012}) and children \citep{keitel2013}.
+This effect suggests that linguistic access is important for children's ability to anticipate upcoming turn structure, consistent with prior work on turn-end prediction in adults (\citealp{de-ruiter2006}; \citealp{magyari2012}) and children \citep{keitel2013}.
 
@@
 In the English videos, children anticipated well at all ages, especially after hearing questions.
-This interaction between age and language condition suggests that the 5-year-olds were able to leverage anticipatory cues in the non-English videos in a way that 3- and 4-year-olds could not, possibly by shifting more attention to the non-native prosodic or non-verbal cues.
+This interaction between age and language condition suggests that the 5-year-olds were able to leverage anticipatory cues in the non-English videos in a way that 3- and 4-year-olds could not, possibly by shifting more attention to the non-English prosodic or non-verbal cues.
 Prior work on children's turn-structure anticipation has proposed that children's turn-end predictions rely primarily on lexicosyntactic structure (and not, e.g., prosody) as they get older \citep{keitel2013}.
@@
 
-In Experiment 2, we follow up on these findings, improving on two aspects of the design: first, our language manipulation in this first experiment was too coarse to provide data regarding specific linguistic channels (e.g., prosody vs.
-lexicosyntax).
+Finally, children and adults made more anticipatory switches in transitions with longer inter-turn gaps, though this effect was limited to non-question turns (Table \ref{tab:E1-models}).
+This finding suggests that gap duration indeed serves as a cue to upcoming turn structure; while short gaps may be perceived as within-turn pauses \citep{mannel2009}, long gaps could instead be indicative of between-turn pauses (where speaker transition occurs).
+Participants might use these long silences to retroactively assign turn boundaries and anticipate speaker switches that were otherwise not anticipated (in this case, because the preceding turn was not a question).
+An alternative explanation for gap duration effects is that longer inter-turn gaps result in longer analysis windows, which yields more time for participants to make an anticipatory gaze.
+However, if participants are generally more likely to make a switch at question transitions, as our results suggest, we would expect that longer gaps would benefit questions more than non-questions---the opposite pattern from what the data show here.
+We take this as evidence that inter-turn silence may be most useful when participants have limited ability to make predictions about upcoming speaker transitions.
+
+In Experiment 2, we followed up on these findings, improving on two aspects of the design: first, our language manipulation in this first experiment was too coarse to provide data regarding specific linguistic information channels (e.g., the effect of prosodic information alone).
 In Experiment 2, we compared lexicosyntactic and prosodic cues with phonetically altered speech and used puppets to eliminate non-verbal cues to turn taking.
 Second, we were not able to pinpoint the emergence of anticipatory switching because the youngest age group in our sample was already able to make anticipatory switches at above chance rates.
-In Experiment 2, we explore a wider developmental range.
-
-% Finally, children showed an effect of gap duration (Table \ref{tab:E1-models}).
-This effect is straightforward: longer gaps resulted in longer analysis windows, yielding more time for children to make an anticipatory gaze.
+In Experiment 2, we explored a wider developmental range.
 
@@
 
-\section{Experiment 2}
+\section*{Experiment 2}
 \label{sec:exp2}
 
-Experiment 2 used native-language stimuli, controlled for lexical and prosodic information, eliminating non-verbal cues, and tested children from a wider age range.
+Experiment 2 used English-only stimuli, controlled for lexical and prosodic information, eliminated non-verbal cues, and tested children from a wider age range.
 To tease apart the role of lexical and prosodic information, we phonetically manipulated the speech signal for pitch, syllable duration, and lexical access.
@@
 
-Non-verbal cues in Experiment 1 (e.g., gaze and gesture) could have helped participants make predictions about upcoming turn structure  \citep{rossano2009, stivers2010}.
-Since our focus was on linguistic cues, we eliminated all gaze and gestural signals in Experiment 2 by replacing the videos of human actors with videos of puppets.
-Puppets are less realistic and expressive than human actors, but they create a natural context for having somewhat motionless talkers in the videos (thereby allowing us to eliminate gestural and gaze cues).
-Additionally, the prosody-controlled condition included small but global changes to syllable duration that would have required complex video manipulation or precise re-enactment with human talkers, neither of which was feasible.
-For these reasons, we decided to substitute puppet videos for human videos in the final stimuli.
-
+Non-verbal gestural cues in Experiment 1 could have helped participants make predictions about upcoming turn structure \citep{rossano2009, stivers2010}.
+Since our focus here is on linguistic cues, we eliminated all gaze and gestural signals in Experiment 2 by replacing the videos of human actors with videos of puppets.
+Puppets are less realistic and expressive than human actors, but they create a natural context for having somewhat motionless talkers in the videos.
+Additionally, the prosody-controlled condition (described below) included small but global changes to syllable duration that would have required complex video manipulation or precise re-enactment with human talkers, neither of which was feasible.
+For these reasons, we decided to use puppet videos rather than human videos in the final stimuli.
 As in the first experiment, we recorded participants' eye movements as they watched six short videos of dyadic conversation, and then analyzed their anticipatory glances from the current speaker to the upcoming speaker at points of turn transition.
 
-\subsection{Methods}
+\subsection*{Methods}
 \label{sec:methods2}
 
-\subsubsection{Participants}
-We recruited 27 undergraduate adults and 129 children between ages 1;0--6;11 to participate in our experiment.
-We recruited our child participants from the Children's Discovery Museum of San Jose, California\footnote{We ran Experiment 2 at a local children's museum because it gave us access to children with a more diverse range of ages.}, targeting approximately 20 children for each of the six one-year age groups (range: 20--23).
+\subsubsection*{Participants}
+We recruited 27 undergraduate adults and 129 children ages 1;0--6;11 to participate in our experiment.
+We recruited adult participants via the Stanford University Psychology participant database.
+Adult participants were either paid or received course credit for their time.
+We recruited our child participants from the Children's Discovery Museum in San Jose, California\footnote{We ran Experiment 2 at a local children's museum because it gave us access to children with a wider range of ages.
+Participants were volunteers.}, targeting approximately 20 children for each of the six one-year age groups (range: 20--23).
 All participants were native English speakers, though some parents (N$=$27) reported that their child heard a second (and sometimes third) language at home.
@@
 
-\subsubsection{Materials}
+\subsubsection*{Materials}
 We created 18 short videos of improvised, child-friendly conversation (Figure \ref{fig:puppets}).
@@
 \textit{Audio Manipulation}.
-We created four versions of each audio clip: \textit{normal}, \textit{words only}, \textit{prosody only}, and \textit{no speech}.
+We created four versions of each audio conversation: \textit{normal}, \textit{words only}, \textit{prosody only}, and \textit{no speech}.
 That is, one version with a full linguistic signal (\textit{normal}), and three with incomplete linguistic information (hereafter ``partial cue'' conditions).
-The \textit{normal} clips were the unmanipulated, original audio clips.
-
-The \textit{words only} clips were manipulated to have robot-like speech: we flattened the intonation contours to each talker's average pitch (F$_{0}$) and we reset the duration of every nucleus and coda to each talker's average nucleus and coda duration.\footnote{We excluded hyper-lengthened words like [w\textipa{aU:}] `woooow!'.
-These were rare in the clips.} We made duration and pitch manipulations using PSOLA resynthesis in Praat \citep{Praat}.
-Thus, the \textit{words only} versions of the audio clips had no pitch or durational cues to upcoming turn boundaries, but did have intact lexicosyntactic cues (and residual phonetic correlates of prosody, e.g., intensity).
-
-We created the \textit{prosody only} clips by low-pass filtering the original recording at 500 Hz with a 50 Hz Hanning window (following de Ruiter et al., 2006).
+The \textit{normal} conversations were the unmanipulated, original audio clips.
+
+The \textit{words only} conversations were manipulated to have robot-like speech: we flattened the intonation contours to each talker's average pitch (F$_{0}$) and we reset the duration of every nucleus and coda to each talker's average nucleus and coda duration.\footnote{We excluded hyper-lengthened words like [w\textipa{aU:}] `woooow!', but these were rare to begin with.} We made duration and pitch manipulations using PSOLA resynthesis in Praat \citep{Praat}.
+Thus, the \textit{words only} versions of the conversations had no pitch or durational cues to upcoming turn boundaries, but did have intact lexicosyntactic cues (and some residual phonetic correlates of prosody, e.g., intensity).
+
+We created the \textit{prosody only} conversations by low-pass filtering the original recording at 500 Hz with a 50 Hz Hanning window (following de Ruiter et al., 2006).
 This manipulation creates a ``muffled speech'' effect because low-pass filtering removes most of the phonetic information used to distinguish between phonemes.
-The \textit{prosody only} versions of the audio clips lacked lexical information, but retained their intonational and rhythmic cues to upcoming turn boundaries.
+The \textit{prosody only} versions of the conversations lacked lexical information, but retained their intonational and rhythmic cues to upcoming turn boundaries.
 
 The \textit{no speech} condition served as a non-linguistic baseline.
-For this condition, we replaced the original clip with multi-talker babble: we overlaid multiple child-oriented conversations (not including the original one), and then cropped the result to the duration of the original video.
-Thus, the \textit{no speech} audio clips lacked any linguistic information to upcoming turn boundaries---the only cue to turn taking was the opening and closing of the puppets' mouths.
-
-Finally, because low-pass filtering removes significant acoustic energy, the \textit{prosody only} clips were much quieter than the other three conditions.
+For this condition, we replaced the original audio clip for the conversation with multi-talker babble: we overlaid multiple child-oriented conversations (excluding the original one), and then cropped the result to the duration of the original conversation clip.
+Thus, the \textit{no speech} conversation lacked any linguistic information to upcoming turn boundaries---the only cue to turn taking was the opening and closing of the puppets' mouths.
+
+Finally, because low-pass filtering removes significant acoustic energy, the \textit{prosody only} conversations were much quieter than the other three conditions.
 Our last step was to downscale the intensity of the audio tracks in the three other conditions to match the volume of the \textit{prosody only} clips.
@@
 We created puppet video recordings to match the manipulated 20-second audio clips.
-The puppets were minimally expressive; so that the puppeteer could only control the opening and closing of their mouths; the puppets' heads, eyes, arms, and bodies stayed still.
-Puppets were positioned looking forward to eliminate shared gaze as a cue to turn structure \citep{thorgrimsson2015}.
-We took care to match the puppets' mouth movements to the syllable onsets as closely as possible, specifically avoiding any mouth movement before the onset of a turn.
-We then added the manipulated audio clips to the puppet video recordings by hand.
-
-We used three pairs of puppets used for the \textit{normal} condition---`red', `blue' and `yellow'---and one pair of puppets for each partial cue condition: ``robots'', ``merpeople'', and ``party-goers'' (Figure 8).
+The puppets were minimally expressive; the puppeteer could only control the opening and closing of their mouths.
+The puppets' heads, eyes, arms, and bodies stayed still.
+Puppets were positioned side-by-side, looking in the same direction to eliminate shared gaze as a cue to turn structure \citep{thorgrimsson2015}.
+We took care to match the puppets' mouth movements to the syllable onsets as closely as possible, specifically avoiding mouth movement before the onset of a turn.
+We then added the manipulated audio clips to the puppet video recordings by hand with video editing software.
+
+We used three pairs of puppets used for the \textit{normal} condition---`red', `blue' and `yellow'---and one pair of puppets for each partial cue condition: `robots', `merpeople', and `party-goers' (Figure 8).
 We randomly assigned half of the conversation topics (`birthday cake', `pets', and `breakfast') to the \textit{normal} condition, and half to the partial cue conditions (`riding bikes', `rainy days', and `the library').
-We then created three versions of the experiment, so that each of the six puppet pairs was associated with three different conversation topics across the different versions of the experiment (18 videos in total).
+We then created three versions of the experiment, so that each of the six puppet pairs was associated with three different conversation topics across the different versions of the experiment (18 videos in total; 6 videos per experiment version).
 We ensured that the position of the talkers (left and right) was counterbalanced in each version by flipping the video and audio channels as needed.
 
-The duration of turn transitions and the number of speaker changes across videos was variable because the conversations were recorded semi-spontaneously.
+As before, the duration of turn transitions and the number of speaker changes across videos was variable because the conversations were recorded semi-spontaneously.
 We measured turn transitions from the audio signal of the \textit{normal}, \textit{words only}, and \textit{prosody only} conditions.
@@
 Inter-turn gaps for questions (mean$=$366, median$=$438, stdev$=$138 msec) were longer than those for non-questions (mean$=$305, median$=$325, stdev$=$94 msec) on average, but gap duration was overall comparable across conditions: \textit{normal} (mean$=$334, median$=$321, stdev$=$130 msec), \textit{words only} (mean$=$347, median$=$369, stdev$=$ 115 msec), \textit{prosody only} (mean$=$365, median$=$369, stdev$=$104 msec), and \textit{no words} (mean$=$319, median$=$329, stdev$=$136 msec).
-The longer gaps for question transitions could give them an advantage because our anticipatory measure includes shifts initiated during the gap between turns (Figure \ref{fig:criterion}).
-
-\subsection{Procedure}
+
+\subsection*{Procedure}
 We used the same experimental apparatus and procedure as in the first experiment.
@@
 Three of the puppet videos had \textit{normal} audio while the other three had \textit{words only}, \textit{prosody only}, and \textit{no speech} audio.
-This experiment required no special instructions so, as before, the experimenter immediately began each session with calibration and then stimulus presentation.
+As before, the experimenter immediately began each session with calibration and then stimulus presentation.
+Participants were given no instruction about how to watch the videos or what their purpose was, they were simply encouraged to watch the ``fun/nice puppet videos''.
 The entire experiment took less than five minutes.
@@
 
-\subsubsection{Data preparation and coding}
+\subsubsection*{Data preparation and coding}
 We coded each turn transition for its linguistic condition (\textit{normal}, \textit{words only}, \textit{prosody only}, and \textit{no speech}) and transition type (question/non-question)\footnote{We coded \textit{wh-}questions as ``non-questions'' for the \textit{prosody only} videos.
@@
 \end{center}
-\caption{Anticipatory gaze rates across language condition and transition type for the real data (blue, dark green, light green, and red) and the randomly permuted baselines (gray).
-Vertical bars represent 95\% confidence intervals.}
+\caption{Anticipatory gaze rates across language condition and transition type for the real and randomly permuted datasets. Vertical bars represent 95\% confidence intervals.} 
 \label{fig:E2-randvsreal}
@@
 
-\subsection{Results}
+\subsection*{Results}
 \label{sec:results2}
@@
 
-\subsubsection{Statistical models}
+\subsubsection*{Statistical models}
 \label{sec:models2}
@@
 The child model included condition (\textit{normal}/\textit{prosody only}/\textit{words only}/\textit{no speech}; with \textit{no speech} as the reference level), transition type (question vs.
-non-question), age (1, 2, 3, 4, 5, 6; numeric), and duration of the inter-turn gap (in seconds) as predictors, with full interactions between language condition, transition type, and age.
-We again included the duration of the inter-turn gap as a control predictor and added random effects of item (turn transition) and participant, with random slopes of transition type for participants.
-The adult model included condition, transition type, their interactions, and duration as a control predictor, with participant and item included as random effects and random slopes of condition and transition type.
-
-Children's anticipatory gaze switches showed an effect of gap duration (\textit{$\beta$}=3.95, \textit{SE}=0.617, \textit{z}=6.401, \textit{p}$<$.001), a two-way interaction of age and language condition (for \textit{prosody only} speech compared to the \textit{no speech} reference level; \textit{$\beta$}=0.393, \textit{SE}=0.189, \textit{z}=2.08, \textit{p}$<$.05), and a three-way interaction of age, transition type, and language condition (for \textit{normal} speech compared to the \textit{no speech} reference level; \textit{$\beta$}=-0.38, \textit{SE}=0.17, \textit{z}=-2.229, \textit{p}$<$.05).
-There were no significant effects of age or transition type alone (Table \ref{sec:models2}), with only a marginal effect of language condition (for \textit{prosody only} compared to the \textit{no speech} reference level; \textit{$\beta$}=-1.607, \textit{SE}=0.867, \textit{z}=-1.85, \textit{p}$=$.064)
-
-Adults' anticipatory gaze switches showed effects of gap duration (\textit{$\beta$}=4.7, \textit{SE}=1.18, \textit{z}=3.978, \textit{p}$<$.001) and language condition (for \textit{normal} speech \textit{$\beta$}=1.203, \textit{SE}=0.536, \textit{z}=2.245, \textit{p}$<$.05 and \textit{words only} speech \textit{$\beta$}=1.561, \textit{SE}=0.709, \textit{z}=2.203, \textit{p}$<$.05  compared to the \textit{no speech} reference level).
-There were no effects of transition type (\textit{$\beta$}=0.437, \textit{SE}=0.585, \textit{z}=0.747, \textit{p}$=$.45).
-
-\subsubsection{Random baseline comparison}
+non-question), age (1, 2, 3, 4, 5, 6; numeric, intercept as age=0), and duration of the inter-turn gap (in seconds) as predictors, with full interactions between language condition, transition type, and age and two-way interactions between gap duration and the other basic predictors (age, linguistic condition, and transition type).
+We also included random effects of participant and item (turn transition), with maximal random slopes of transition type for participant.
+The adult model included condition, transition type, their interactions, gap duration, and two-way interactions between gap duration and condition and transition type, with participant and item as random effects and maximal random slopes of condition and transition type for participant.
+
+Children's anticipatory gaze switches showed an effect of gap duration (\textit{$\beta$}=3.85, \textit{SE}=1.73, \textit{z}=2.22, \textit{p}$<$.05), a two-way interaction of age and language condition (for \textit{prosody only} speech compared to the \textit{no speech} reference level; \textit{$\beta$}=0.38, \textit{SE}=0.19, \textit{z}=1.97, \textit{p}$<$.05), a marginal two-way interaction of language condition and gap duration (for \textit{prosody only} speech compared to the \textit{no speech} reference level; \textit{$\beta$}=-4.77, \textit{SE}=2.63, \textit{z}=-1.82, \textit{p}$=$.07), and a three-way interaction of age, transition type, and language condition (for \textit{normal} speech compared to the \textit{no speech} reference level; \textit{$\beta$}=-0.35, \textit{SE}=0.17, \textit{z}=-2.05, \textit{p}$<$.05).
+There were no significant effects of age or transition type alone (Table \ref{sec:models2}) (\textit{$\beta$}=-0.05, \textit{SE}=0.14, \textit{z}=-0.38, \textit{p}$=$.7 and \textit{$\beta$}=-1.22, \textit{SE}=0.96, \textit{z}=-1.27, \textit{p}$=$.2, respectively)
+
+Adults' anticipatory gaze switches showed a significant effect of language condition (for \textit{words only} speech compared to the \textit{no speech} reference level; \textit{$\beta$}=3.79, \textit{SE}=1.62, \textit{z}=2.34, \textit{p}$<$.05) and a marginal two-way interaction between language condition and transition type (for \textit{words only} speech compared to the \textit{no speech} reference level; \textit{$\beta$}=-1.68, \textit{SE}=0.89, \textit{z}=-1.89, \textit{p}$=$.06).
+There was no significant effect of transition type alone (\textit{$\beta$}=-0.02, \textit{SE}=1.44, \textit{z}=-0.02, \textit{p}$=$.99).
+
+\subsubsection*{Random baseline comparison}
 \label{sec:randbaseline2}
 
-Using the same technique described in Experiment 1 (Section \ref{sec:randbaseline1}), we created and modeled random permutations of participants' anticipatory gaze.
-These analyses revealed that none of the significant predictors from models of the original, turn-related data could be explained by random looking.
-In the children's data, the original model's \textit{z}-values for language condition (\textit{prosody only}), gap duration, the two-way interaction of age and language condition (\textit{prosody only}) and the three-way interaction of age, transition type, and language condition (\textit{normal} speech) were all greater than 95\% of the randomly permuted \textit{z}-values (96.5\%, 100\%, 95.6\%, and 95.5\%, respectively, all \textit{p}$<$.05).
-Similarly, the adults' data showed significant differentiation from the randomly permuted data for all originally significant predictors: gap duration and language condition for \textit{normal} speech and \textit{words only} speech (greater than 100\%, 97.8\%, and 99.1\% of random \textit{z}-values, respectively, all \textit{p}$<$.05).
+Using the same technique described in Experiment 1 (Section \ref{sec:randbaseline1}), we created and modeled random permutations of participants' anticipatory gaze switches.
+These analyses revealed that the significant predictors from models of the original, turn-related data were unlikely to be explained by random looking.
+In the children's data, the original model's \textit{z}-values for gap duration, the two-way interaction of age and language condition (\textit{prosody only}) and the three-way interaction of age, transition type, and language condition (\textit{normal} speech) were all greater than 93\% of the randomly permuted \textit{z}-values (95.6\%, 94\%, and 93.3\%, respectively, \textit{p}$=$.04, .06, and .07).
+Similarly, the adults' data showed significant differentiation from the randomly permuted data for the effect of language condition (\textit{words only} speech; greater than 98.3\% of random \textit{z}-values, \textit{p}$<$.02).
 See Section \ref{sec:permutation} for more information on each predictor's random permutation distribution.
@@
     \hline
-    (Intercept)											& -3.49403		& 0.48454		& -7.211	& 5.55e-13 ***    	\\
-    Age														& 0.02436		& 0.10249		& 0.238	& 0.8121        		\\
-    Type$=$\textit{non-Question}			& -0.88900		& 0.61192		& -1.453	& 0.1463        		\\
-    GapDuration												& 3.94743		& 0.61668		& 6.401	& 1.54e-10 ***    	\\
-    Age*Type$=$\textit{non-Question}	& 0.15359		& 0.13996		& 1.097	& 0.2725        		\\
-	\hline
-    Condition$=$\textit{normal}				& 0.37337		& 0.43421		& 0.860	& 0.3899        		\\
-    Age*Condition$=$\textit{normal}		& 0.12950		& 0.10217		& 1.267	& 0.2050        		\\
-    Condition$=$\textit{normal}*			& 0.91074		& 0.72581		& 1.255	& 0.2096        		\\
+    (Intercept)											& -3.452		& 0.76	& -4.543	& 5.55e-06 ***	\\
+    Age														& -0.054		& 0.143	& -0.379	& 0.705				\\
+    Type$=$\textit{non-Question}				& -1.217		& 0.958	& -1.27		& 0.204				\\
+    GapDuration											&  3.852		& 1.735	&  2.221	& 0.026 *			\\
+    Age*Type$=$\textit{non-Question}			&  0.152		& 0.141	&  1.081	& 0.28				\\
+    Age*GapDuration									&  0.214		& 0.266	&  0.805	& 0.421				\\
+    Type$=$\textit{non-Question}*				&  0.995		& 2.134	&  0.466	& 0.641				\\
+    \hspace*{5mm} GapDuration &&&& \\
+	\hline
+    Condition$=$\textit{normal}					&  0.54		& 0.742	&  0.728	& 0.467					\\
+    Age*Condition$=$\textit{normal}			&  0.125		& 0.103	&  1.221	& 0.222				\\
+    Condition$=$\textit{normal}*				&  0.908		& 0.748	&  1.215	& 0.224				\\
     \hspace*{5mm} Type$=$\textit{non-Question} &&&& \\
-    Age*Condition$=$\textit{normal}*	& -0.37965		& 0.17031		& -2.229	& 0.0258 *      		\\
+    Age*Condition$=$\textit{normal}*			& -0.355		& 0.173	& -2.051	& 0.04 *			\\
     \hspace*{5mm} Type$=$\textit{non-Question} &&&& \\
-	\hline
-    Condition$=$\textit{prosody}			& -1.60734		& 0.86680		& -1.854	& 0.0637 .      		\\
-    Age*Condition$=$\textit{prosody}	& 0.39271		& 0.18905		& 2.077	& 0.0378 *      		\\
-    Condition$=$\textit{prosody}*			& 1.68552		& 1.05414		& 1.599	& 0.1098        		\\
+    Condition$=$\textit{normal}*				& -0.431		& 1.67	& -0.258	& 0.797				\\
+    \hspace*{5mm} GapDuration &&&& \\
+	\hline
+    Condition$=$\textit{prosody}				&  0.549		& 1.452	&  0.378	& 0.705				\\
+    Age*Condition$=$\textit{prosody}			&  0.375		& 0.191	&  1.967	& 0.049 *			\\
+    Condition$=$\textit{prosody}*				&  1.076		& 1.105	&  0.974	& 0.33				\\
     \hspace*{5mm} Type$=$\textit{non-Question} &&&& \\
-    Age*Condition$=$\textit{prosody}*	& -0.32360		& 0.23229		& -1.393	& 0.1636        		\\
+    Age*Condition$=$\textit{prosody}*		& -0.296		& 0.235	& -1.257	& 0.209				\\
     \hspace*{5mm} Type$=$\textit{non-Question} &&&& \\
-	\hline
-    Condition$=$\textit{words}				& -0.26996		& 0.59313		& -0.455	& 0.6490        		\\
-    Age*Condition$=$\textit{words}		& 0.14044		& 0.13565		& 1.035	& 0.3005        		\\
-    Condition$=$\textit{words}*				& -1.03066		& 1.01610		& -1.014	& 0.3104        		\\
+    Condition$=$\textit{prosody}*				& -4.767		& 2.625	& -1.816	& 0.069 .			\\
+    \hspace*{5mm} GapDuration &&&& \\
+	\hline
+    Condition$=$\textit{words}				&  0.684		& 1.06	&  0.645	& 0.519					\\
+    Age*Condition$=$\textit{words}		&  0.127		& 0.136	&  0.934	& 0.350					\\
+    Condition$=$\textit{words}*				& -1.244		& 1.031	& -1.207	& 0.228					\\
     \hspace*{5mm} Type$=$\textit{non-Question} &&&& \\
-    Age*Condition$=$\textit{words}*		& 0.08829		& 0.22387		& 0.394	& 0.6933        		\\
+    Age*Condition$=$\textit{words}*		&  0.111		& 0.225	&  0.495	& 0.621					\\
     \hspace*{5mm} Type$=$\textit{non-Question} &&&& \\
+    Condition$=$\textit{words}*				& -2.285		& 2.232	& -1.024	& 0.306					\\
+    \hspace*{5mm} GapDuration &&&& \\
     \hline
     &&&& \\
-  \textbf{\textit{Adults}} &&&& \\
-    \hline
-           &  Estimate & Std. Error & \textit{z} value & Pr($>$$|$\textit{z}$|$) \\    
-    \hline
-    (Intercept)											& -3.3811		& 0.6884		& -4.912	& 9.03e-07 *** 		\\ 
-    Type$=$\textit{non-Question}			& 0.4375			& 0.5854		& 0.747	& 0.4549     		\\ 
-    GapDuration												& 4.6961			& 1.1804		& 3.978	& 6.94e-05 *** 		\\ 
-	\hline
-    Condition$=$\textit{normal}				& 1.2033			& 0.5359		& 2.245	& 0.0247 *   		\\ 
-    Condition$=$\textit{normal}*			& -0.9627		& 0.7358		& -1.308	& 0.1907     		\\ 
-    \hspace*{5mm} Type$=$\textit{non-Question} &&&& \\
-	\hline
-    Condition$=$\textit{prosody}			& 0.2407			& 0.8011		& 0.301	& 0.7638     		\\ 
-    Condition$=$\textit{prosody}*			& 0.5525			& 0.9374		& 0.589	& 0.5556     		\\ 
-    \hspace*{5mm} Type$=$\textit{non-Question} &&&& \\
-	\hline
-    Condition$=$\textit{words}				& 1.5613			& 0.7087		& 2.203	& 0.0276 *   		\\ 
-    Condition$=$\textit{words}*				& -1.1557		& 0.8854		& -1.305	& 0.1918     		\\ 
-    \hspace*{5mm} Type$=$\textit{non-Question} &&&& \\
-    \hline
-    &&&& \\
-  \caption{Model output for children and adults' anticipatory gaze switches in Experiment 2.}
-\label{tab:E2-models}
+  \caption{Model output for children's anticipatory gaze switches in Experiment 2.}
+\label{tab:E2-chimodels}
 \end{longtable}
@@
 \end{minipage}
-\linespread{2}
-
-\subsubsection{Developmental effects}
+%\linespread{2}
+
+\linespread{1}
+\begin{minipage}[t]{0.95\linewidth}
+\begin{footnotesize}
+\begin{longtable}{lcccl}
+  \textbf{\textit{Adults}} &&&& \\
+    \hline
+           &  Estimate & Std. Error & \textit{z} value & Pr($>$$|$\textit{z}$|$) \\    
+    \hline
+    (Intercept)										& -3.117		& 1.176	& -2.649	& 0.008 **				\\ 
+    Type$=$\textit{non-Question}			& -0.022		& 1.44	& -0.015	& 0.988					\\ 
+    GapDuration										&  4.073		& 2.947	&  1.382	& 0.167					\\ 
+    Type$=$\textit{non-Question}*			&  1.304		& 3.859	&  0.338	& 0.735					\\
+    \hspace*{5mm} GapDuration &&&& \\
+	\hline
+    Condition$=$\textit{normal}				&  0.39		& 1.316	&  0.296	& 0.767						\\ 
+    Condition$=$\textit{normal}*			& -0.709		& 0.754	& -0.94		& 0.347					\\ 
+    \hspace*{5mm} Type$=$\textit{non-Question} &&&& \\
+    Condition$=$\textit{normal}*			&  2.1			& 3.336	&  0.629	& 0.529					\\ 
+    \hspace*{5mm} GapDuration &&&& \\
+	\hline
+    Condition$=$\textit{prosody}			&  0.757		& 2.193	&  0.345	& 0.73					\\ 
+    Condition$=$\textit{prosody}*			&  0.386		& 1.065	&  0.362	& 0.717					\\ 
+    \hspace*{5mm} Type$=$\textit{non-Question} &&&& \\
+    Condition$=$\textit{prosody}*			& -1.118		& 4.543	& -0.246	& 0.805					\\ 
+    \hspace*{5mm} GapDuration &&&& \\
+	\hline
+    Condition$=$\textit{words}				&  3.792		& 1.621	&  2.338	& 0.019 *				\\ 
+    Condition$=$\textit{words}*				& -1.678		& 0.889	& -1.888	& 0.059 .				\\ 
+    \hspace*{5mm} Type$=$\textit{non-Question} &&&& \\
+    Condition$=$\textit{words}*				& -5.653		& 3.861	& -1.464	& 0.143					\\ 
+    \hspace*{5mm} GapDuration &&&& \\
+    \hline
+    &&&& \\
+  \caption{Model output for adults' anticipatory gaze switches in Experiment 2.}
+\label{tab:E2-adumodels}
+\end{longtable}
+\end{footnotesize}
+\end{minipage}
+%\linespread{2}
+
+\subsubsection*{Developmental effects}
 
@@
 As in Experiment 1, we used two-tailed \textit{t}-tests to compare children's real gaze rates to the random baseline rates in the \textit{normal} speech condition, in which the speech stimulus is most like what children hear every day.
-We tested real gaze rates against baseline for three age groups: ages one, two, and three.
+We tested real gaze rates against baseline rates for three age groups: one-, two-, and three-year-olds.
 Two- and three-year-old children made anticipatory gaze switches significantly above chance both when all transitions were considered (2-year-olds: \textit{t}(26.193)$=$-4.137, \textit{p}$<$.001; 3-year-olds: \textit{t}(22.757)$=$-2.662, \textit{p}$<$.05) and for question transitions alone (2-year-olds: \textit{t}(25.345)$=$-4.269, \textit{p}$<$.001; 3-year-olds: \textit{t}(21.555)$=$-3.03, \textit{p}$<$.01).
 One-year-olds, however, only made anticipatory gaze shifts marginally above chance for turn transitions overall and for question turns alone (overall: \textit{t}(24.784)$=$-2.049, \textit{p}$=$.051; questions: \textit{t}(25.009)$=$-2.03, \textit{p}$=$.053).
+
+We also tested the two baseline linguistic conditions against each other---\textit{no speech} and \textit{normal speech}---to find out when linguistic information made a difference in children's anticipations.
+Because, as we have seen, children primarily show linguistic effects in question-answer turn transitions, we investigated the use of linguistic cues across age by testing anticipation separately for question and non-question turns.
+Compared to the \textit{no speech} condition, children made significantly more anticipatory switches in the \textit{normal} speech condition for questions at ages 6, 4, and 3, and also marginally at age 2 (6-year-olds: \textit{t}(36.919)$=$3.8019, \textit{p}$<$.001; 4-year-olds: \textit{t}(41.449)$=$2.9777, \textit{p}$<$.01; 3-year-olds: \textit{t}(35.724)$=$2.4286, \textit{p}$<$.05; 2-year-olds: \textit{t}(41.078)$=$1.8018, \textit{p}$=$.079).
+Children's anticipatory switches for questions did not differ in the \textit{no speech} and \textit{normal} speech conditions at ages 5 or 1 (5-year-olds: \textit{t}(29.406)$=$1.2783, \textit{p}$=$.211; 1-year-olds: \textit{t}(35.907)$=$0.4961, \textit{p}$=$.623).
+In contrast, children's anticipatory switch rates for non-question turns were not significantly different between the \textit{no speech} and \textit{normal} speech conditions at any age (all \textit{p}$>$0.09).
+Consistent with the regression output, children were more likely to show an effect of linguistic content as they got older, but only for question transitions.
 
@@
 
-\subsection{Discussion}
+\subsection*{Discussion}
 \label{sec:discussion2}
@@
 The core aims of Experiment 2 were to gain better traction on the individual roles of prosody and lexicosyntax in children's turn predictions, and to find the age of emergence for spontaneous turn anticipation.
-Many of our results replicate the findings from Experiment 1: participants often made more anticipatory switches when they had access to lexical information and, when they did, tended to make more anticipatory switches for questions compared to non-questions.
-
-As in Experiment 1, children and adults spontaneously tracked the turn structure of the conversations, making anticipatory gaze switches at above-chance rates across all ages when listening to natural speech.
-They also made far more anticipations for questions than for non-question turns---at least for children two years old and older.
-But these effects were different for the two conditions with partial linguistic information: \textit{prosody only} and \textit{words only}.
-In the \textit{prosody only} condition, performance was low for younger children and increased significantly with age (especially for questions).
-In the \textit{words only} condition, children age two and older showed robust anticipatory switching for questions (much like in \textit{normal} speech), but never rose above chance for non-question turns.
-These findings do not therefore support an early role for prosody in children's spontaneous turn structure predictions.
-There is also no evidence that lexical information is sufficient on its own to support children's anticipatory switching.
-
-\section{General Discussion}
+Many of our results replicate the findings from Experiment 1: participants often made more anticipatory switches when they had access to linguistic information and, when they did, tended to make more anticipatory switches for questions compared to non-questions.
+
+As in Experiment 1, children and adults spontaneously tracked the turn structure of the conversations.
+Participants made anticipatory gaze switches at above-chance rates starting at age one for question transitions and age two for both questions and non-questions.
+Longer gaps had a broader impact on participants' anticipations in this experiment; we saw that, overall, longer inter-turn gaps resulted in more anticipatory switches, with the \textit{no speech} condition showing equal or significantly stronger effects of gap duration than all other conditions.
+
+As before, participants made far more anticipations for questions than for non-question turns---at least for those two years old and older.
+But these effects were different for the conditions with partial linguistic information: \textit{prosody only} and \textit{words only}.
+In the \textit{prosody only} condition, performance was initially low for young children and increased significantly with age.
+In the \textit{words only} condition, children age two and older showed robust switching for questions (much like in \textit{normal} speech), but never rose above chance for non-question turns (Figure \ref{fig:E2-randvsreal}), with no significant differences from the \textit{no speech} baseline.
+These findings do not support an early role for prosody or lexical information alone in children's spontaneous predictions about turn structure.
+They also give no evidence for the idea that lexical information is sufficient on its own to support children's anticipatory switching.
+They do underscore the developing relationship between the use of linguistic cues and speech act (transition type) in spontaneous predictions about upcoming turn structure.
+
+\section*{General Discussion}
 \label{sec:gendisc}
@@
 Until recently, we have had very little data on how children weave language into their already-existing turn-taking behaviors.
-In two experiments investigating children's anticipatory gaze to upcoming speakers, we found evidence that turn prediction develops early in childhood and that spontaneous predictions are primarily driven by participants' expectation of an immediate response in the next turn (e.g., after questions).
+In two experiments investigating children's anticipatory gaze to upcoming speakers, we found evidence that turn prediction develops early in childhood and that, when spontaneous predictions begin, they are primarily driven by participants' expectation of an immediate response in the next turn (e.g., after questions).
 In making predictions about upcoming turn structure, children used a combination of lexical and prosodic cues; neither lexical nor prosodic cues alone were sufficient to support increased anticipatory gaze.
-We also found no early advantage for prosody over lexicosyntax, and instead found that children were unable to make above-chance anticipatory gazes in the \textit{prosody only} condition until age five.
-We discuss these findings with respect to the role of linguistic cues in predictions about upcoming turn structure, the importance of questions in spontaneous predictions about conversation, and children's developing competence as conversationalists.
-
-\subsection{Predicting turn structure with linguistic cues}
+We also found no early advantage for prosody over lexicosyntax; children's anticipatory switch rates in the \textit{prosody only} condition were initially low, but showed significant gains by age five.
+We discuss these findings with respect to: the role of linguistic cues and inter-turn silence for predicting upcoming turn structure, the importance of questions in predictions about conversation, and children's developing competence as conversationalists.
+
+\subsection*{Predicting upcoming turn structure}
 
@@
 
-Perhaps surprisingly, we also found no evidence that lexical information alone is equivalent to the full linguistic signal in driving children's predictions, as has been shown previously for adults \citep{magyari2012, de-ruiter2006} and as is replicated with adult participants in the current study.
-That said, our findings point more toward early lexical effects in children's turn anticipations than prosodic ones: children's performance in both experiments was consistently better when they had access to lexical information, especially after question turns.
-And although the \textit{words only} condition in Experiment 2 was not significantly different from the baseline \textit{no speech} condition, children's anticipations trended toward above-chance rates.
-
-Above all, children and adults anticipated best with they had access to the full linguistic signal.
-There may be something informative about combined prosodic and lexical cues to questionhood that helps to boost children's anticipations before they can use these cues separately.
+Using prosody alone to accurately predict turn boundaries in conversation appears to be difficult for adults and children.
+Prosodic information is continuous, multidimensional, and indexically complex---it encodes syntactic structure, speech act, and extralinguistic orientation, sometimes simultaneously, though without clear one-to-one mappings between form and meaning \citep{cutler1997, shriberg1998, lammertink2015}.
+For these reasons, prosodic information alone may not be enough to both (a) make precise temporal predictions about turn structure, and (b) identify question turns, which otherwise appear to drive anticipatory gaze switching.
+Therefore, although children show early facility with prosodic discrimination \citep{nazzi2003, soderstrom2003, johnson2001, jusczyk1995, morgan1995, mehler1988}, using prosodic knowledge for turn prediction may generally be too difficult without additional information from lexical or syntactic cues.
+
+Our findings suggest that there is one prosodic cue that serves as an exception to this rule: inter-turn silence.
+Generally speaking, participants showed a greater anticipatory switches for longer inter-turn gaps, but the effect of inter-turn gap duration is strongest in our data when upcoming responses are less predictable, whether due to the question effect (Experiment 1) or the lack of non-verbal cues and any linguistic information (Experiment 2).
+Notably, there were no significant interactions of gap duration with participant age.
+This pattern of results suggests that, when predictive information about upcoming responses is absent, long silences may be used to retroactively assign turn-end boundaries, thereby increasing the participant's expectation for a speaker change and promoting more anticipatory gaze switches.
+The lack of interactions between age and gap duration suggests that the use of inter-turn silence remains important for older speakers under ambiguous turn structure conditions and the interactions between transition type and gap duration (Experiment 1) and condition and gap duration (Experiment 2; marginal), suggest that this effect is not straightforwardly the result of having more time to make a gaze switch.
+
+Because inter-turn silence works as a backwards-looking cue, it's predictive utility for speaker change is diminished compared to forward-looking linguistic cues (e.g., syntactic constituent completion).
+However, it is a strategy that can be usefully employed to predict upcoming responses some of the time, even without access to language.
+Pauses are detected and related to phrasal structure from early on; 5-month-old infants use pauses to parse intonational phrases \citep{mannel2009}.
+Our findings thus suggest that silence is an early and lasting cue for identifying turn structure online when other predictive information is not adequate.
+
+Perhaps surprisingly, we found no evidence that lexical information alone is equivalent to the full linguistic signal in driving children's predictions, as has been shown previously for adults \citep{magyari2012, de-ruiter2006} and as is replicated with adult participants in the current study.
+Unlike prosodic cues, lexicosyntactic cues are discreet and have much clearer form-to-meaning mappings, with clear lexicosyntactic cues to questionhood that occur early within turns (e.g., \textit{wh}-words, \textit{do}-insertion, and subject-auxiliary inversion).
+That said, although children's lexical and syntactic knowledge begins to develop early, it is limited for quite some time \citep{bergelson2013, shi2010, tomasello1999}.
+Our stimuli were made in a child-friendly style, they are still other-directed and fairly complex, with 20--30 seconds of continuous conversational speech.
+
+It is perhaps for this reason that children's performance was always best with the full signal, where lexicosyntactic information was supported by prosodic information and vice versa.
+There may be something extra informative about the combination of prosodic and lexical cues to questionhood that helps to boost children's anticipations before they can use these cues separately.
 Even in adults, Torreira and colleagues (\citeyear{torreira2015}) showed that the trade-off in informativity between lexical and prosodic cues is more subtle in semi-natural speech.
@@
 
-\subsection{The question effect}
+Finally, because most anticipatory switches were made following question turns, one could predict that the only lexical and prosodic cues that matter for spontaneous turn prediction are those that cue questionhood.
+Our linguistic manipulations here were focused on lexical and prosodic information globally, which may have affected lexical and prosodic cues to questionhood asymmetrically; while lexicosyntactic question cues were available on every instance of \textit{wh}- and \textit{yes/no} questions in our stimuli, prosodic question cues were only salient on \textit{yes/no} questions.
+Relatedly, many non-verbal cues also encode information about transition type, including gaze and gesture.
+We did not systematically test those cues here but, like inter-turn silence, they may play a critical role in parsing and making predictions about turn structure when other linguistic information is not sufficient to make accurate predictions.
+
+\subsection*{The question effect}
 
@@
 
-Compared to prosodic cues (e.g., final rising intonation), lexicosyntactic cues to questionhood (e.g., \textit{wh}-words, \textit{do}-insertion, and subject-auxiliary inversion) are categorical, and early-occurring in the utterance.
-Children may have therefore had an easier time picking out and interpreting lexical cues to questionhood.
-Developmentally, the question effect showed its first significant gains between ages three and four in the \textit{normal} speech condition of Experiment 2 (Figure \ref{fig:E2-lgagetypeinteraction}), by which time children frequently hear and use a variety of polar and \textit{wh}-questions \citep{clark2009}.
-Furthermore, while lexicosyntactic question cues were available on every instance of \textit{wh}- and \textit{yes/no} questions in our stimuli, prosodic question cues were only salient on \textit{yes/no} questions.
-Finally, the mapping of prosodic contour to speech act (e.g., high final rises for polar questions) is far from one-to-one, leaving substantial room for uncertainty in prosodic contour interpretation in general.
-
 Prior work on children's acquisition of questions indicates that they may already have some knowledge of question-answer sequences by the time they begin to speak: questions make up approximately one third of the utterances children hear, before and after the onset of speech, and even into their preschool years, though the type and complexity of questions changes throughout development \citep{casillas2016, fitneva2012, henning2005, shatz1979}.\footnote{There is substantial variation in question frequency by individual and socioeconomic class \citep{hart1992, weisleder2012}.} For the first few years, many of the questions directed to children are ``test'' questions---questions that the caregiver already has the answer to (e.g., ``What does a cat say?''), but this changes as children get older.
 Questions help caregivers to get their young children's attention and to ensure that information is in common ground, even if the responses are non-verbal or infelicitous \citep{bruner1985, fitneva2012, snow1977}.
-So, in addition to having a special interactive status (for adults and children alike), questions are a core characteristic of many caregiver-child interactions, motivating a general benefit for questions in turn structure anticipation.
+Moreover, because of their high frequency and relatively limited number of formats, questions, especially \textit{wh}-questions, may be more identifiable and predictable compared to other types of speech acts.
+So, in addition to having a special interactive status (for adults and children alike), questions are a frequent, predictable and core characteristic of many caregiver-child interactions, motivating a general benefit for questions in turn structure anticipation.
 
@@
 and (2) what is it about questions that makes children and adults more likely to anticipatorily switch their gaze to addressees?
-Other turn types, such as imperatives, compliments, and complaints make a response from the addressee highly likely in the next turn \citep{schegloff2007}.
-Rhetorical and tag questions, on the other hand, take a similar form to prototypical polar questions, but often do not require an answer.
-So, though it is clear that adults and children anticipated responses more often for questions than non-questions, we do not yet know whether their predictive action is limited to turns formatted as questions or is generally applicable to turn structures that project an immediate response from the addressee.
-
-More broadly, our results suggest that participants' spontaneous predictions, at least while viewing third-party conversation, are driven by what lies \textit{beyond} the end of the current turn---not just by the upcoming end of the turn itself, as has been focused on in prior work \citep{torreira2015, keitel2013, magyari2012, de-ruiter2006}.
-In future work, it will be crucial to measure prediction from a first-person perspective to resolve this apparent contradiction (see also \citealp{holler2015}).
-
-% There is at least one clear hypothesis that can bridge these apparently conflicting results: listeners in spontaneous, first-person conversation may use multiple strategies in making predictions about upcoming turn structure, using more passive prediction to detect upcoming speaker transition (e.g., questions), and then switching into precise turn-end prediction mode (\`{a} la \citealp{de-ruiter2006}) when necessary.
-A flexible prediction system like this one allows listeners to continuously monitor ongoing conversation at a low cost while still managing to plan their responses and come in quickly when needed.
-
-% To test this hypothesis, which integrates findings from turn-structure prediction with multiple measures, age groups, and styles of linguistic control, it will be crucial to look at prediction from a first-person perspective.
-The results we present here are based on predictions about third-party conversation, which enables participants to follow interactions with no chance of actually participating.
-Although recent work has shown that similar anticipatory eye gazes do occur in spontaneous conversation \citep{holler2015}, more work is needed to determine if the same question advantage occurs, which linguistic cues seem to drive it, and whether participants switch into a ``precision'' mode when they detect imminent speaker change.
-
-\subsection{Early competence for turn taking?}
+If this effect is particular to turns that require an immediate response (``adjacency pairs''; \citealp{schegloff2007}), other turn types, such as imperatives, compliments, and complaints should show similar patterns to questions.
+If the effect is instead about overall predictability of the syntactic frame, children would instead show similar patterns for other frequent frames in child-directed speech (e.g., ``Look at the X.''; \citealp{mintz2003}).
+Recognizability and predictability are likely to play a role as children become more sophisticated language users, even if the effect is truly about adjacency pairs; for example, rhetorical and tag questions take a very similar form to prototypical polar questions, but usually do not require an answer.
+So, though it is clear that adults and children anticipated responses more often for questions than non-questions, we do not yet know whether their predictive action is limited to turns formatted as questions, turns with high recognizability and predictability, or turns that project an immediate response from the addressee.
+
+The question effect suggests that participants' spontaneous predictions, at least while viewing third-party conversation, may be driven by what lies \textit{beyond} the end of the current turn---not just by the upcoming end of the turn itself, as has been focused on in prior work \citep{torreira2015, keitel2013, magyari2012, de-ruiter2006}.
+In future work, it will be crucial to measure prediction from a first-person perspective to find out what kinds of predictions are most relevant to addressees in conversation (see also \citealp{holler2015}).
+
+One possible scenario is that listeners in spontaneous, first-person conversation use multiple strategies to make predictions about upcoming turn structure: they could use semi-passively attend to incoming speech for cues to upcoming speaker transition (e.g., questions and other adjacency pairs) and, when possible upcoming transition is detected, switch into a more precise turn-end prediction mode (\`{a} la \citealp{de-ruiter2006}).
+A flexible prediction system like this one allows listeners to continuously monitor ongoing conversation for turn-related cues at a low cost while still managing to plan their responses and come in quickly when needed.
+
+To test this hypothesis, we would need to look at prediction from a first-person perspective, which very little work so far has accomplished (present work included).
+Although third-party measures enable us to measure participants' predictions without any interference from language production, they also limit our knowledge about how the need to give a response might itself play an important role in addressees' prediction strategies.
+Recent work has shown that shifts in addressee gaze similar to those measured here indeed occur in spontaneous conversation \citep{holler2015}, but much more work is needed to determine how participants make predictions about turn structure in first-person contexts and whether those mechanisms shift at points of imminent speaker change.
+
+\subsection*{Early competence for turn taking?}
 
@@
 Therefore, this difference in age of emergence more likely stems from our use of a more engaging speech style, stereo speech playback, and more typical turn transition durations.
+The child-friendly style of speech in particular may have helped in two ways: keeping children more engaged with the stimuli in general and using less syntactically complex and more prosodically exaggerated cues \citep{fernald1989, werker1989, snow1977} compared to what they would get with adult-adult speech.
 
 To be clear, young children's ``above chance'' performance was often still far from adult-like predictive behavior---children at ages one and two were still very close to chance in their anticipations and, even at age six, children were not fully adult-like in their predictions.
-This may indicate that young children rely more on non-verbal cues in anticipating turn transitions or, alternatively, that adults are better at flexibly adapting to the turn-relevant cues present at any moment.
-
-% This added flexibility might emerge around age 5, when we see the first adult-like effects in the children's data: in Experiment 1, 5-year-olds using non-verbal or non-native prosodic cues to make predictions in the non-English videos comparable to their predictions for English.
-In Experiment 2, 5- and 6-year-olds showing significant advances over younger children in their anticipations for the \textit{prosody only} condition.
-
-Taken together, our data suggest that turn-taking skills do begin to emerge in infancy, but that children cannot make effective predictions until they can pick out question turns.
-This finding leads us to wonder how participant role (first- instead of third-person) and cultural differences (e.g., high vs.
-low parent-infant interaction styles) feed into this early predictive skill.
-It also bridges  prior work showing a predisposition for turn taking in infancy (e.g., \citealp{bateson1975, hilbrink2015, jaffe2001, snow1977}) with children's apparently \textit{late} acquisition of adult-like competence for turn taking in actual conversation \citep{casillas2016, garvey1984, garvey1981, ervin-tripp1979}.
-
-\subsection{Limitations and future work}
+This may indicate that young children rely primarily on non-verbal cues (like inter-turn silence) in anticipating turn transitions while adults use both verbal and non-verbal cues to make predictions.
+Relatedly, adults may be more expert in flexibly adapting to the turn-relevant cues present at any moment, e.g., responding to non-English prosodic cues in Experiment 1.
+
+Taken together, our data suggest that turn-taking skills do begin to emerge in infancy, but that children cannot consistently make effective predictions until they can identify and react to question turns.
+This finding leads us to wonder how participant role (first- instead of third-person) and differences in early interactional experience (e.g., frequent vs.
+infrequent question-asking from caregivers) feed into this early predictive skill.
+It also bridges prior work showing a predisposition for turn taking in infancy (e.g., \citealp{bateson1975, hilbrink2015, jaffe2001, snow1977}) with children's apparently \textit{late} acquisition of adult-like competence for turn taking in actual conversation \citep{casillas2016, garvey1984, garvey1981, ervin-tripp1979} and reinforces the idea that it takes children several years to fully integrate linguistic information into their turn-taking systems \citep{casillas2016,garvey1981}.
+
+\subsection*{Limitations and future work}
 
@@
 This decision resulted in speech sounds that children don't usually hear in their natural environment.
-Many prior studies have used phonetically-altered speech with infants and young children \citep[cf.][]{jusczyk2000}, but almost none of them have done so in a conversational context.
-Future work could instead carefully script or cross-splice sub-parts of turns to control for the presence of linguistic cues for turn transition (see, e.g., \citealp{torreira2015}).
-
-The prediction measure used in our studies is based on an observer's view of third-party conversation but, because participants' role in the interaction could affect their online predictions about turn taking, an better measure would instead capture first-person behavior.
-First-person measures of spontaneous turn prediction will be key to revealing how participants distribute their attention over linguistic and non-linguistic cues while taking part in everyday interaction, the implications of which relate to theories of online language processing for both language learning and everyday talk.
-
-\subsection{Conclusions}
+Many prior studies have used phonetically-altered speech with infants and young children \citep[cf.][]{jusczyk2000}, but few of them have done so in a conversational context.
+Future work could instead carefully script speech or cross-splice sub-parts of turns to control for the presence of linguistic cues for turn transition (see, e.g., \citealp{torreira2015}).
+
+The prediction measure used in our studies is based on an observer's view of third-party conversation but, because participants' role in the interaction could affect their online predictions about turn taking, an ideal measure would instead capture first-person predictions.
+If conversational participants' predictions are partly shaped by their need to respond, first-person measures of spontaneous turn prediction will be key to revealing how participants distribute their attention over verbal and non-verbal cues while taking part in everyday interaction, the implications of which relate to theories of online language processing for both language learning and everyday talk.
+
+That said, the third-person paradigm used in the present study still has much to tell us about turn prediction.
+The task is natural and intuitive in that no instruction is required, which means that it captures spontaneous predictive behavior and can be used with participants of all ages.
+Frequencies of anticipatory gaze switching appear to be stable across language communities where similar tasks have been tested \citep{keitel2013, keitel2015, holler2015, hirvenkari2013}---even from a first-person perspective---so the task is one that measures robust predictive behavior relevant to conversational processing in many languages.
+It also lends itself to many possibilities for controlling the presence of individual verbal and non-verbal cues and has a clear method for assessing random switching baselines across the entire stimulus (because the participant's task is constant).
+Also, in the case that response preparation interferes with our ability to see prediction at the ends of incoming turns, third-person paradigms are one of the only ways to ensure that we are measuring prediction processes in isolation.
+
+The current findings also make important predictions about what we will see in first-person prediction paradigms.
+For example, the focus on upcoming points of speaker transition is only more important when participants are embedded in a live interaction; we would thus expect question-like effects to occur in first-person paradigms too, and perhaps even be amplified.
+If so, participants' use of linguistic information would still subserve this goal, with prediction at a premium.
+So we would predict that linguistic cues to upcoming speaker change would be most critical in a first-person measure.
+Regarding development, the same facts about the complexity of prediction with prosodic information and children's initial limited lexical inventories would still hold, as would the use of silence and non-verbal cues to assess and predict turn structure in the absence of clear predictive linguistic information.
+We therefore think that the paradigm presented here has important contributions to make in finding out how we attend to and make predictions about conversational interaction.
+
+\subsection*{Conclusions}
 
@@
 Adults use linguistic information to accurately predict turn structure in conversation, which facilitates their online comprehension and allows them to respond relevantly and on time.
-The present study offers new findings regarding the role of speech acts and linguistic processing in online turn prediction, and has given evidence that turn prediction emerges by age two, but is not integrated with linguistic cues until much later.
-Using language to make predictions about upcoming interactive content takes time and, for both children and adults, is primarily driven by participants' orientation to what will happen next---beyond the end of the current turn.
+The present study offers new findings regarding the role of speech acts and linguistic processing in online turn prediction, and has given evidence that turn prediction emerges by age two, increases with age, and is driven by question turns.
+However, turn prediction is not fully integrated with linguistic cues until much later and, in the absence of predictive linguistic cues, children and adults alike rely on retroactive cues such as inter-turn silence to predict upcoming speaker change.
+Using language to make predictions about upcoming interactive content takes time to develop and, for participants of all ages appears to be primarily driven by participants' orientation to what will happen next---beyond the end of the current turn.
 
@@
 We also thank Tania Henetz, Francisco Torreira, Stephen C.
-Levinson, Eve V.
-Clark, and the First Language Acquisition group at Radboud University for their feedback on earlier versions of this work.
+Levinson, and Eve V.
+Clark for their feedback on earlier versions of this work.
 The analysis code for this project can be found on GitHub at https://github.com/langcog/turn\_taking/.
 
-\bibliographystyle{elsarticle-harv}
+\bibliographystyle{model5-names}
+\biboptions{authoryear}
 \bibliography{anticip}
@@
 \end{sidewaystable}
-\linespread{2}
+%\linespread{2}
 
@@
 
-One alternative hypothesis for children's anticipatory gazes is that they simply grow bored and start looking away at a constant rate after a turn begins.
-This data plotted here show a hypothetical group of boredom-driven participants (gray dots) compared to participants from the actual data in Experiment 2 (black dots).
+One alternative hypothesis for children's anticipatory gazes is that they are looking at the current speaker at the start of each turn, but then grow bored and start looking away at a constant rate.
+Even though this alternative hypothesis does not predict the primary effects in our data (e.g., the difference between questions and non-questions), we cannot rule out the possibility that a portion of participants' saccades come from boredom.
+
+The data plotted here show a hypothetical group of boredom-driven participants (gray dots) compared to participants from the actual data in Experiment 2 (black dots).
 The hypothetical boredom-driven participants look away from the current speaker at a linear rate, beginning one second after the start of a turn.
@@
 
-If children's switches away from the current speaker were driven by boredom, they would switch away equally quickly on long and short turns.
-However, as turn duration increased, children in the experiment looked at the current speaker for a longer period before looking away, suggesting that they were not switching away at a constant rate.
-We can see this pattern most clearly in the time at which 50\% of the children had switched away from the current (indicated by the horizontal line in Figure \ref{fig:boredomhypothesis}): children's gaze crossed this 50\% threshold at 2.0, 2.9, and 3.6 seconds after the start of speech for turns with durations of 1--2, 2--3, and 3--4 seconds, respectively.
-In contrast, the hypothetical boredom-driven children always hit the 50\% threshold at 2.5 seconds after the start of speech.
-This pattern suggests that, though children do look away with time, their looks away are not simply driven by boredom.
+If children's switches away from the current speaker were purely driven by boredom, they would switch away equally quickly on long and short turns.
+Therefore, their crossover point---the point in time at which 50\% of the children have switched away from the current speaker---would be the same for all turns, no matter the length of the turn.
+This pattern is demonstrated in the hypothetical boredom-driven crossover points, which always occur 2.5 seconds after the start of speech (gray vertical line; Figure \ref{fig:boredomhypothesis}).
+
+In children's \textit{actual} looking data we see that crossover points increase with turn duration: 2.0, 2.9, and 3.6 seconds after the start of speech for turns with durations of 1--2, 2--3, and 3--4 seconds, respectively (black vertical line; Figure \ref{fig:boredomhypothesis}).
+This pattern suggests that, though children do look away as the turn is unfolding, their looks away are not simply driven by boredom.
+
+Are the looks away in Figure \ref{fig:boredomhypothesis} still too early to count as ``turn-transition'' anticipation?
+It is true that children start looking away after one second has passed (but then only gradually).
+It is possible that some of these early looks away are boredom-driven, but it is equally plausible that some of them are turn-driven.
+Early predictive behavior is common in turn-taking studies with adults, in both constrained turn-taking tasks (e.g., \citealp{de-ruiter2006, gisladottir2015, bogels2015}) and in spontaneous conversation (e.g., \citealp{holler2015, torreira2015}).
+Although this same pattern has yet to be established for children's turn predictions, the looking behavior here is at least consistent with adult response patterns in previous work.
+Additionally, because our analysis windows in the main study only overlapped with the pre-gap utterance by 300 msec (Figure \ref{fig:criterion}), our primary results are unlikely to capture any very early or early boredom-driven gaze switches.
+
+We conclude that the boredom-driven effects in our data are unlikely to change our primary results, though we acknowledge that characterizing different gaze switching strategies in this kind of data is an important avenue for future work.
 

### 2016-06-06 paper/Letters/ResponseLetter-Jun2016.tex
minor changes for R3 and template for the response letter (content to come soon)

--- 2015-12-15 paper/Letters/ResponseLetter-Dec2015.tex
@@
 
-\opening{Dear Profs. Dr. Pickering and Dr. Gerrig,}
+\opening{Dear Dr. Gerrig,}
 
@@
 
-\noindent Please accept our resubmission of our manuscript on ``The development of children's ability to track and predict turn structure in conversation'' to the \textit{Journal of Memory and Language}.
-We apologize for our delay in getting the manuscript resubmitted.
-It took quite a long time to implement one of the suggestions from Reviewer 3, mostly because I had to learn quite a bit of new technical skill to carry out a suggested analysis out on our computation-heavy data set.
-Nevertheless, I believe that this analysis has substantially strengthened the paper as a whole, so I appreciate your patience.
+\noindent I hereby resubmit a revised version of my manuscript ``The development of children's ability to track and predict turn structure in conversation'' to the \textit{Journal of Memory and Language}.
 
-\noindent In addition to the new analysis, we have also included two other follow-up analyses in the Appendix, both inspired by the other reviewers' comments.
-And in other revisions, we have tried hard to make the text more succinct and the results and discussion sections more distinct and more straightforward, following requests from all three reviewers.
+%\noindent I apologize for our delay in getting the manuscript resubmitted.
+It took quite a long time to implement one of the suggestions from Reviewer 3, mostly because I had to acquire some new technical skills in order to carry out his/her suggested analysis on our computation-heavy data set.
+Nevertheless, I believe that the added analysis has substantially strengthened the paper as a whole, so I appreciate your patience very much.
 
-\noindent We hope that, with these changes, we have greatly improved the paper's clarity.
-Please find a description of specific comments addressed below.
+%\noindent Since our last submission, we have done quite a lot of work to improve our paper by addressing the reviewers' comments.
+For example, in addition to the new random permutation analysis suggested by Reviewer 3 (mentioned above), we have also included two further follow-up analyses in the Appendix (including additional experimental data), each inspired by the other reviewers' comments.
+We have also tried to make the text much more succinct, with separate and straightforward sections for the results and discussions, following requests from all three reviewers.
+
+%\noindent We hope that, with these changes, we have greatly improved the paper's quality.
+Please find a description of the specific comments addressed below.
 Thank you very much for your reconsideration.
@@
 \begin{tabular}{ll}
-\textit{ID:} & JML-14-235 \\
+\textit{ID:} & JML-14-235R3 \\
 \textit{Title:} & The development of children's ability to track and \\
@@
 
-\noindent We thank all three reviewers for their thorough and helpful comments on the previously submitted manuscript.
-We have tried to address all of the reviewers' comments, within the manuscript as well as here in the response letter.
+%\noindent We thank all three reviewers for their thorough and helpful comments on the previously submitted manuscript.
+We have tried to address all of their comments, both in the manuscript and here in the response letter.
 
@@
 
-\noindent \textit{1. 333ms may be too long for older children's saccadic planning time and may therefore inflate the anticipatory looking rates for older children.}
-
-\noindent Thank you for this comment; we were concerned about this issue as well, but we unfortunately did not collect independent saccadic planning norms for children ages 1--6 (and we do not know of any authoritative source for these norms).
-In the newly analyzed data, we take the conservative approach of using adult-like planning times (200ms) for children at all ages in both experiments.
-
-\smallskip
-
-\noindent \textit{2. The description of the random baseline permutations was not clear.}
-
-\noindent In our revision, we have completely redone all random permutation analyses and have attempted to clarify their description (Section 2.2.2).
-
-\smallskip
-
-\noindent \textit{3. Are anticipatory gaze shifts driven by boredom (and not by anticipation)? Do children generally just look away as the speaker goes on?}
-
-\noindent Thanks for pointing out an important alternative hypothesis.
-To address this comment, we modeled the proportion of participants looking at the current speaker for our real data and for (fake) boredom-driven lookers (who look away from the current speaker at a constant rate, starting 1 second after the onset of speech) across turns of different length.
-Even in short turns, where there is very little temporal room for children to behave differently from the boredom-driven lookers, we still see a clear difference between the boredom-driven simulations and the real data (Figure D.2).
-We take this result to be evidence against the alternative hypothesis; we report the details of the analysis in an appendix.
-
-\smallskip
-
-\noindent \textit{4. The reported results from the models models are hard to follow.}
-
-\noindent We apologize---the data from this experiment are very complex.
-To aid comprehension, we have added more information about how each variable was coded and have put summary tables of every model's output into the text.
-In addition, we have reorganized our results sections in hopes of making them more straightforward.
-
-\smallskip
-
-\noindent \textit{5. ``No speech'' should be the reference level against which the linguistic conditions are compared in Experiment 2.}
-
-\noindent This is another very useful comment, thank you!
-We agree and have changed our analyses to set ``no speech'' as the reference level.
-We had started with the ``normal speech'' condition as the reference level because this choice mirrors the coding in related work on adult turn-end prediction.
-But the additive model implied by the ``no speech'' baseline is easier to interpret.
-
-\noindent \textit{6. Looking at the data, it seems possible that the children are not using any linguistic information.}
-
-\noindent Thanks for raising this challenging and important point.
-To summarize: We think that the robust question (speech act) effects we observed in both experiments provide strong evidence for the use of linguistic information.
-
-In Experiment 2, children's data showed a 2-way interaction of age and transition type (question vs.
-non-question), but only in the ``normal'' speech condition, suggesting that it was the only condition in which they could extract the linguistic cues needed to accurately identify questions.
-We speculate that this is because prosody and lexical cues to questionhood often work together in forming questions during everyday speech, but it could also be that the manipulated speech was simply too unfamiliar for them to predict effectively.
-Either way, the existence of the question effect in the normal condition indicates that children do, in fact, use linguistic information (to identify questions) and the interaction with age suggests that they get better at doing so as they get older.
-These effects cannot be explained by duration alone (and duration is controlled in our analysis), and there is no visual cue advantage difference between the normal condition and the others.
-We are then left to conclude that they use linguistic information to support the question effect.
-
-A similar story can be told for the data in Experiment 1: non-verbal cues to questions vs.
-non-questions are available in both the English and non-English stimuli, but participants (adults and children) made more anticipatory switches when they had access to the lexical information.
-Again, this cannot be explained by gap duration or unequal visual advantages across the stimuli.
-Thus, we believe that both experiments provide support for the use of linguistic cues, albeit for a subset of these cues.
-
-% The fact that we have found most of our linguistic processing effects within an effect of speech act differs from prior work (e.g., de Ruiter et al., 2006), however, where there was clearer evidence of linguistic processing across multiple phonetically manipulated conditions, but without question-non-question comparisons.
-
-% More work is needed to square these two patterns of data, but we do believe that the presence o 
-% No turn prediction measure yet is capable of capturing predictions that occur in spontaneous first-person conversation, but we think that our results indicate that there are multiple types of prediction relevant to anticipating upcoming turn structure during interaction, and that any account of how listeners use linguistic cues while participating in conversation should be able to account for shifts in how listeners attend to cues to make more or less specific predictions (i.e., ``what words are upcoming'' vs.
-``what types of action are upcoming'').
-% In sum, while the linguistic effects in our data are somewhat limited, they are present, consistent, developmentally changing, and important to our developing theories about turn prediction.
-We believe that this sets up future work quite well to dig into the specific linguistic cues (lexical and prosodic, though probably mostly lexical) that are being used in spontaneous question identification.
-
-\smallskip
-
-\noindent \textit{7. It is too hard to link the conclusions in the discussion back to the statistical results and the effects themselves are not talked about in a clear enough way.}
-
-\noindent We apologize.
-In our revision, we have tried our best to make these links clearer, first by clarifying the statistical analyses and outputs, and second by trying to explicitly link each claim to an individual result (through the use of figure and table references).
-We hope that this will help readers better connect the results to our interpretations.
-
-\smallskip
-
-\noindent \textit{8. There are too many references to unpublished work.}
-
-\noindent The single upside to our delayed resubmission is that all of the unpublished work referred to previously is now published or in press!
-
-\smallskip
-
-\noindent \textit{(Reviewer 1's minor points/corrections have all been addressed directly in the text.)}
-
-\bigskip
-
 \noindent \textbf{Reviewer 2}
-
-\noindent \textit{1. It is not clear which claims are backed up by which analyses. Analyses should be carried out for each group separately.}
-
-\noindent As discussed above, we have tried to improve the linkage between our claims and analyses and now have presented the statistical results in tables, as requested.
-Please see responses 4 and 7 under Reviewer 1 for a little more detail.
-
-The main focus of our analyses was the effect that age (as a continuous variable) has on anticipatory gaze.
-However there are two cases in which our main analyses are not satisfying without further, age-group analyses: (a) age sometimes showed itself to interact with other factors (e.g., age and language condition), and (b) without further tests we can not say whether particular age groups themselves statistically differ from chance, across or within conditions.
-
-Following your suggestion, we have thus made two additional changes.
-First, we added follow-up two-tailed t-tests, making pairwise comparisons between age groups for significant interactions so that we could find out how the interacting factor (e.g., language condition) significantly changes across age groups (Sections B.1--3).
-Second, we have added individual models of the youngest age groups (ages 1, 2, and 3), comparing their real looking behavior to our randomly permuted data across conditions (at the ends of Sections 2.2.2 and 3.3.2).
-These follow-up analyses, which are summarized in the text, are primarily described in the Appendix (Section B).
-
-\smallskip
-
-\noindent \textit{2. The results and discussion for each study would be better if they were separated into different sections.}
-
-\noindent Our original intention in combining results and discussion was to make the findings easier to understand.
-Since we were not able to achieve this in the prior submission, we have followed Reviewer 2's advice and now use a more traditional ``results-first'' structure.
-We hope this change has improved the clarity of the paper.
-
-\smallskip
-
-\noindent \textit{3. The methods sections need more elaboration in a few places, and the baseline analysis section could be much clearer.}
-
-\noindent We thank the reviewer for careful attention to our methods section (and its missing details).
-We have added in the experimental methods information requested and have re-checked each section for missing information that might be useful in replicating our methods.
-We've also worked to clarify the description of the baseline analysis (now reported in depth in the Appendix), including a new figure (Section 2.2.2; Figure 4).
-
-\smallskip
-
-\noindent \textit{4. The adult data deserve further discussion because they raise the question of whether this measure taps into linguistic processing at all.}
-
-\noindent We actually think the adults' results are well in-line with prior work showing that (a) adults' predictions primarily rely on lexical cues (E1 \& E2) and (b) adults are better anticipators than young children (E1 \& E2).
-The complicating factor is that we report a new and important interacting factor with their linguistic processing: speech act.
-Like the children in our data set, adults made more anticipatory switches following questions, which means that most of our evidence about their use of linguistic cues comes from question transitions.
-This enriches the picture of linguistic processing for turn prediction, as painted by prior work: linguistic processing is brought into play more often in some contexts than others.
-From these points we conclude that our experiments do indeed tap into participants' use of linguistic cues to predict upcoming turn structure.
- % That isn't to say that the method wouldn't be improved by more natural linguistic controls and tighter controls on gap duration---it would!
-But we think our current results still make a valuable contribution to the current understanding of turn anticipation in children \textit{and} adults, and lays critical groundwork for discovering what specific types of cues participants spontaneously monitor in everyday interaction.
-
-\smallskip
-
-\noindent \textit{5. There is a lot of unpublished work referred to in the manuscript, with errors in the reference list.}
-
-\noindent These papers are now all published or in press. We apologize for errors in the reference list; we have checked them again and hope that there are no remaining errors.
-
-\bigskip
-
-\noindent \textbf{Reviewer 3}
-
-\noindent \textit{1. The manuscript could be shortened substantially without loss of content.}
-
-\noindent We have significantly shortened the introduction and tried to remove redundancies where possible in the rest of the text.
-We have also moved some of the secondary analyses to the Appendix so that the main text can be read without getting into the nitty gritty of every single analysis.
-We hope that this has helped make the main text more straightforward and approachable length-wise.
-
-\smallskip
-
-\noindent \textit{2. There is a confound between linguistic condition and puppet pair in Experiment 2.}
-
-\noindent We thank Reviewer 3 for this point.
-We repeat here some of the information now available in the Appendix for easy summary: Our design does not fully cross puppet pair (e.g., robots, blue puppets) with linguistic condition (e.g., ``words only'' and ``no speech'').
-Even though each puppet pair is associated with different conversation clips across children (e.g.
-robots talking about kitties, birthday parties, or pancakes), robots were only associated with ``words only'' speech, merpeople were only associated with ``prosody only'' speech, and the puppets with fancy clothes were only associated with the ``no speech'' condition.
-We made this choice to increase the pragmatic felicity of the experiments for the older children (i.e., robots make robot sounds, merpeople's voices are muffled under water, the fancy-clothed puppets are in a room with main other voices).
-It is therefore fair to point out a possible confound between linguistic condition and puppet pair.
-
-Thankfully, we also ran a short follow-up study at the museum with 3--5-year-olds that addresses this issue.
-In the follow-up study, each child only saw one video---the normal speech conversation about birthday parties---with a randomly assigned puppet pair performing the conversation.
-Five children watched each puppet pair, for a total of 30 children across the six pairs (shown Figure 3).
-This experiment holds all things constant except for the appearance of the puppets.
-We then used a mixed effects logistic regression of children's anticipatory switches (yes or no at each transition), with puppet pair (robots/merpeople/fancy dress/normal-speech-puppets) as a fixed effect and participant and turn transition as random effects.
-In four versions of this model we systematically varied the reference level to check for differences between every puppet pair, finding no significant affects of puppet type on switching rate.
-We take this as evidence that, although we did not fully cross puppet pairs and linguistic conditions in Experiment 2, it was unlikely to have had strong effects on children's looking rates above and beyond the intended effects of linguistic condition.
-We have included details and graphs for this norming experiment in the Appendix.
-
-\smallskip
-
-\noindent \textit{3. Computer animation would have been a better choice than puppets.}
-
-\noindent We agree with the reviewer that computer animation is one method for providing greater experimental control.
-We did not have the ability to create custom animated conversation videos at the time the research was conducted, though we do use it in some ongoing work.
-Nevertheless, we believe that the control experiment reported above provides some data that speaks against a ``puppet saliency'' hypothesis accounting for the results of Experiment 2.
-
-\smallskip
-
-\noindent \textit{4. The model design needs to be further justified and the way the effects are talked about needs clarification.}
-
-\noindent In our revision, we correct some small issues in our model description, for example, making it clear that we used a logistic regression (for our binary response variable of switch-no switch).
-We have also added a justification of our use of separate models for adults and children and our decisions regarding random slopes.
-We have also cleaned up the way we talk about the significant effects in the model output.
-In particular, we believe that the concern regarding dummy coding was a misunderstanding; dummy coding was in fact used for categorical predictors (as is now clearer with the presentation of results in Tables 2 and 5).
-
-\smallskip
-
-\noindent \textit{5. An ideal-observer analysis of the usefulness of the linguistic cues would give a better idea about what can affect anticipatory gaze.}
-
-\noindent We agree---this is a great suggestion, but it may venture too far outside the scope of the current paper, especially because our stimulus set is probably too small to support building a generalizable model of cue use.
-In addition, the interpretation of our experiments is already complex without the addition of a computational model.
-% We will, however, seriously consider this type of analysis for future work, with adult and child turn-taking, in experiments and in spontaneous speech corpora.
-
-\smallskip
-
-\noindent \textit{6.
-The random baseline would be better if it were converted to a full permutation analysis so that the baseline rate of anticipatory gazing is not treated as an additive factor, which it probably is not.}
-
-\noindent We thank Reviewer 3 very much for the suggestion and clear explanation of how to conduct the analysis.
-Unfortunately, this was not a trivial analysis to implement, causing some delays in the manuscript resubmission.
-Running 5,000 random permutations and 20,000 mixed models required a substantial investment in terms of parallel computing infrastructure.
-
-That being said, we think that the new permutation analysis is indeed simpler and more theoretically intuitive than our prior method (described in 2.2.2 and 3.3.2, and visualized in Section A).
-Please note that there were two problems with conducting the analyses as originally instructed in the reviewer's comments.
-1) We used $t$ values instead of $\beta$ estimates because the standard error estimates for the randomly permuted data were categorically higher (due to data sparsity issues).
-The use of $t$ values is closer to the practice in neuroimaging data analysis, where statistical significance is computed based on the empirical distribution of test statistics.
-2) Many of the models resulted in convergence warnings.
-We made the decision to report results across all convergent models, but we are happy to revisit this decision if the Reviewer believes there is an alternative approach that we should follow.
-
-\smallskip
-
-\noindent \textit{7. The theoretical contribution of this paper is not clear.}
-
-\noindent We have tried to improve the clarity of our findings' theoretical importance in the revised manuscript. I summarize here the points that we cover.
-
-First, one of the most important theoretical contributions of our work is in establishing and replicating the effect of speech act (questions) in children and adults' spontaneous predictions.
-While participants may always use linguistic information to predict upcoming speaker changes, our results to not support the idea that they always use linguistic information to predict upcoming \emph{turn ends}, as is assumed in metalinguistic measures of turn-end prediction (e.g., pressing a button while listening to speech).
-Instead, our results suggest that participants' spontaneous predictions are the result of question-monitoring, presumably achieved by recruiting linguistic cues to questionhood from the unfolding signal.
-The trends in our data suggest that participants are primarily recruiting lexical cues to do this, though establishing this pattern will require follow up work with more focused stimuli.
-
-Second, lexical information alone is not equivalent to full linguistic information for children, as it has been shown to be with adults (e.g., de Ruiter et al., 2006; replicated in our work).
-While this result will require replication with different stimulus items, it still points to an important developmental effect.
-% If this effect arose in our data because children do not respond well to phonetically manipulated speech in conversational contexts, this point can be overturned by work that controls linguistic information through other means.
-If not, it suggests that there is something specially informative about the combined prosodic-lexical marking on question turns, for which there is also prior support (Torreira et al., 2015) though our finding would be the first to show such effects developmentally.
-
-Third, young children (e.g., at ages one and two) do make anticipatory gaze prediction more often than would be expected by chance alone (albeit at low rates).
-Since older children and adults' gaze shifts are primarily driven by question turns, this may suggest that, although children \textit{can} make predictions about upcoming turn structure, it doesn't count for much until they have acquired the linguistic skill to pick out question turns.
-This finding, in turn, has potentially important implications for how participant role (first- instead of third-person) and cultural differences (high vs.
-low parent-infant interaction styles) might affect children's early predictive skill.
-It also bridges prior work demonstrating that children begin taking non-linguistic turns in infancy (e.g., Hilbrink et al., 2015), but still have trouble integrating linguistic aspects of responding at age 3--4 (Casillas et al., in press; Garvey \& Berninger, 1984).
-
-% Fourth, our work helps understand we  our interpretation of discrepancies between our methods and results compared to prior work (using more explicit, metalinguistic methods and showing more obvious online predictions using linguistic information).
-The analyses of these discrepancies led us to put forth a new idea: participants in first-person spontaneous conversation may use multiple strategies in making predictions about upcoming turn structure, using more passive prediction to detect upcoming speaker transitions (e.g., questions), switching into precise turn-end prediction mode when necessary.
-A flexible system like this allows listeners to continuously monitor ongoing conversation at a low cost while still managing to plan their responses and come in quickly when needed.
-As far as we know, we are the first to articulate this idea, integrating findings from multiple methods, age groups, and styles of linguistic control.
 

### 2016-06-06 paper/CasillasFrank-Anticip.tex
minor changes for R3 and template for the response letter (content to come soon)

@@
 Like other types of human interaction, it is organized at its core by the roles and goals of its participants.
-But, what sets conversation apart is its structure: sequences of interconnected, communicative actions that take place across alternating turns at talk.
+But what sets conversation apart is its structure: sequences of interconnected, communicative actions that take place across alternating turns at talk.
 Sequential, turn-based structures in conversation are strikingly uniform across language communities and linguistic modalities.
@@
 But children's early, turn-structured social interactions are presumably a critical precursor to their later conversational turn taking.
-Early non-verbal interactions likely establish the protocol by which children come to use language with others.
-How do children integrate linguistic knowledge with these preverbal turn-taking abilities, and how does this integration change over the course of childhood?
-
-In this study, we investigate when children begin to make predictions about upcoming turn structure in conversation, and how they integrate language into their predictions as they grow older.
-In the remainder of the introduction, we first give a basic review of turn-taking research and the state of current knowledge about adult turn prediction.
-We then discuss recent work on the development of turn-taking skills before turning to the details of our own study.
+These early non-verbal interactions likely establish the protocol by which children come to use language with others.
+How then do children integrate linguistic knowledge with these preverbal turn-taking abilities?
+
+In this study, we investigate when children begin to make predictions about upcoming turn structure in conversation and how they integrate language into their predictions as they grow older.
+We first give a basic review of turn-taking research and the state of current knowledge about adult turn prediction.
+We then discuss recent work on the development of turn-taking skills before getting into the details of the present study.
 
@@
 The groups can span turn-by-turn exchanges (e.g., simple question--response, ``How are you?''--``Fine.'') or sequence-by-sequence exchanges (e.g., reciprocals, ``How are you?''--``Fine, and you?''--``Great!'').
-Compared to other turn-taking behaviors, the possible sequence and action types in everyday talk are can be diverse and unpredictable.
+Compared to other turn-taking behaviors, the possible sequence and action types in everyday talk are diverse and unpredictable.
 
@@
 Despite this complexity, conversational turn taking is precise in its timing.
-Across a diverse sample of conversations in 10 languages, one study found a consistent average turn transition time of 0--200 msec at points of speaker switch \citep{stivers2009}.
+Across a diverse sample of conversations in 10 languages, one study found a consistent average inter-turn silence of 0--200 msec at points of speaker switch \citep{stivers2009}.
 Experimental results and current models of speech production suggest that it takes approximately 600 msec to produce a content word, and even longer to produce a simple utterance \citep{griffin2000, levelt1989}.
-So in order to achieve 200 msec turn transitions, speakers must begin formulating their response before the prior turn has ended \citep{levinson2013, levinson2016}.
+In order to achieve 200 msec turn transitions, speakers must begin formulating their response before the prior turn has ended \citep{levinson2013, levinson2016}.
 Moreover, to formulate their response early on, speakers must track and anticipate what types of response might become relevant next.
@@
 
-Adults have a lot of information at their disposal to help make accurate predictions about upcoming turn content.
+Adults have a lot of information at their disposal to help make accurate predictions.
 Lexical, syntactic, and prosodic information (e.g., \textit{wh}- words, subject-auxiliary inversion, and list intonation) can all inform addressees about upcoming linguistic structure \citep{de-ruiter2006, duncan1972, ford1996, torreira2015}.
 Non-verbal cues (e.g., gaze, posture, and pointing) often appear at turn-boundaries and can sometimes act as late indicators of an upcoming speaker switch \citep{rossano2009, stivers2010}.
-Additionally, the sequential context of a turn can make it clear what will come next: answers after questions, thanks or denial after compliments, etc.
+Additionally, the sequential context of a turn can make the next action obvious: answers after questions, thanks or denial after compliments, etc.
 \citep{schegloff2007}.
@@
 Participants never verbally responded or pressed the ``turn-end'' button when hearing a syntactically-complete phrase without an intonational phrase boundary.
-And when intonational phrase boundaries were embedded in multi-utterance turns, participants were tricked into pressing the ``turn-end'' button 29\% of the time.
-Their results suggest that listeners actually do rely on prosodic cues to execute a response (see also de \citet{de-ruiter2006}:525).
+And when intonational phrase boundaries were embedded within multi-utterance turns, participants were tricked into pressing the ``turn-end'' button 29\% of the time.
+These findings suggest that listeners actually do rely on prosodic cues to execute a response in interaction with their predictions about the unfolding syntactic structure (see also de \citet{de-ruiter2006}:525).
 These experimental findings corroborate other corpus and experimental work promoting a combination of cues (lexicosyntactic, prosodic, and pragmatic) as key for accurate turn-end prediction \citep{duncan1972, ford1996, hirvenkari2013}.
@@
 The majority of work on children's early turn taking has focused on observations of spontaneous interaction.
-Children's first turn-like structures appear as early as two to three months after birth, in proto-conversation with their caregivers \citep{bruner1975, bruner1985}.
+Children's first turn-like structures appear as early as two to three months after birth, in proto-conversation with their caregivers \citep{bruner1975, bruner1985, snow1977}.
 During proto-conversations, caregivers treat their infants as capable of making meaningful contributions: they take every look, vocalization, arm flail, and burp as ``utterances'' in the joint discourse \citep{bateson1975, jaffe2001, snow1977}.
@@
 
-Before children manage to integrate linguistic cues into their turn-taking behaviors (for both turn prediction and production), they can rely on non-verbal interactional cues, including silence (e.g., between turns), eye gaze, body orientation, and gesture, to identify the boundaries of social actions.
-For example, with little to no linguistic knowledge, children are often able to infer and anticipate desired responses to offers and requests by taking account of their interlocutor's non-verbal communicative behavior, the structure of routine events, and the affordances of the current interactional context \citep{reddy13, nomikou11, shatz78}.
-With respect to turn taking in particular, children's spontaneous vocalizations during interaction demonstrate a sensitivity to short inter-speaker gaps from infancy \citep{hilbrink2015}, even when they are not yet using language to participate.
-Thus, before they can anticipate turn structure form linguistic cues, children might instead react to silence as a cue to upcoming speaker change.
-Interactional silence itself may then serve as an one of children's first cues to turn structure, giving them information about when to respond before they can rely on language.
+Before children manage to integrate linguistic cues into their turn-taking behaviors (for both turn prediction and production), they can rely on non-verbal interactional cues, including silence, eye gaze, body orientation, and gesture, to identify the boundaries of social actions.
+For example, with little to no linguistic knowledge, children are often able to infer desired responses to offers and requests by taking account of their interlocutor's non-verbal communicative behavior, the structure of routine events, and the affordances of the current interactional context \citep{reddy13, nomikou11, shatz78}.
+With respect to turn taking in particular, children's spontaneous vocalizations during interaction demonstrate a sensitivity to short inter-speaker gaps from infancy \citep{hilbrink2015}.
+Thus, before they can anticipate turn structure form linguistic cues, children might react to silence as a cue to upcoming speaker change.
+Interactional silence itself may then serve as one of children's first cues to turn structure, giving them information about when to respond before they can rely on language.
 
@@
 They asked 6-, 12-, 24-, and 36-month-old infants, and adult participants to watch short videos of conversation and tracked their eye movements at points of speaker change.
-They showed their participants two types of conversation videos---one normal and one with flattened pitch---to test the role of intonation in participants' anticipatory predictions about upcoming speech.
-Comparing children's anticipatory gaze frequency to a random baseline, they found that only 36-month-olds and adults made anticipatory gaze switches more often than expected by chance, and only 36-month-olds were affected by a lack of intonation contours.
+They showed their participants two types of videos---one normal and one with flattened pitch---to test the role of intonation in participants' anticipatory predictions about upcoming speech.
+Comparing children's anticipatory gaze frequency to a random baseline, they found that only 36-month-olds and adults made anticipatory gaze switches more often than expected by chance, and only 36-month-olds were affected by flattened intonation contours.
 This finding led Keitel and colleagues to conclude that children's ability to predict upcoming turn structure relies on their ability to comprehend the stimuli lexicosemantically.
@@
 Again, we attempt to address these issues in our current study.
-Despite these minor methodological issues, the Keitel et al.
-\citeyearpar{keitel2013} study still demonstrates intriguing age-based differences in children's ability to predict upcoming turn structure.
+Despite these minor methodological drawbacks, the Keitel et al.
+\citeyearpar{keitel2013} study still demonstrates interesting age-based differences in children's ability to predict upcoming turn structure.
 Our current work takes these findings as a starting point.\footnote{But also see \citet{casillas2012, casillas2013}.}
@@
 
-We highlight four primary findings: first, although children and adults use linguistic cues to make predictions about upcoming turn structure, they do so primarily to predict speaker transitions after questions (a speech act effect).
+We highlight four primary findings: first, although children and adults use linguistic cues to make predictions about upcoming turn structure, they do so primarily to predict speaker transitions after questions (a ``speech act'' effect).
 This intriguing effect, which has not been reported previously, suggests that participants track unfolding speech for cues to upcoming speaker change, which may affect how they use linguistic cues more generally for anticipatory processing in conversation.
 Second, we find that children make more predictions than expected by chance starting at age two, but that this effect is small at first, and continues to improve through age six, along with children's use of linguistic cues to anticipate answers after question turns.
-Third, children and adults at all ages often used inter-turn silence (a non-verbal cue to turn structure) to make more predictive gaze switches to the responder, suggesting that non-verbal cues are useful for predicting turn structure early on and continue to be important in adulthood.
-Finally, we find no evidence for an early prosodic advantage in children's anticipations and, further, no evidence that lexical cues alone are comparable to the full linguistic signal in aiding participants' predictions (as is proposed for adults; \citealp{de-ruiter2006}).
+Third, children and adults often used inter-turn silence (a non-verbal cue to turn structure) to make more predictive gaze switches to the responder, suggesting that non-verbal cues are useful for predicting turn structure early on and continue to be important in adulthood.
+Finally, we find no evidence for an early prosodic advantage in children's anticipations and, further, no evidence that lexical cues alone are comparable to the full linguistic signal in aiding children's predictions (as is proposed for adults; \citealp{de-ruiter2006}).
 Anticipation is strongest for stimuli with the full range of linguistic cues.
-Our findings support an account in which turn prediction emerges in infancy, but becomes fully integrated with linguistic processing only gradually across development.
+Our findings support an account in which turn prediction emerges in infancy, but becomes fully integrated with linguistic processing only gradually.
 
@@
 We showed participants non-English videos to limit their access to lexical information while maintaining their access to other cues to turn boundaries (e.g., non-English prosody, gaze, in-breaths, phrase final lengthening).
-Using this method, we compared children and adult's anticipatory looks from the current speaker to the upcoming speaker at points of turn transition in English and non-English videos.
+Using this method, we analyzed children and adult's anticipatory looks from the current speaker to the upcoming speaker at points of turn transition in English and non-English videos.
 
@@
 Adult participants were either paid or received course credit for their time.
-Our child sample included 19 three-year-olds, 32 four-year-olds, and 23 five-year-olds, all enrolled in a local nursery school.
-All participants were native English speakers and volunteered their time.
-Approximately one-third (N$=$25) of the children's parents and teachers reported that their child regularly heard a second (and sometimes third or further) language, but only one child frequently heard a language that was used in our non-English video stimuli, and we excluded his data from the analyses.\footnote{Multilingual children may make predictions about upcoming turn structure differently from their monolingual peers due to their more varied experiences with language-relevant turn taking cues, but we are unable to test this hypothesis here due to the variability in multilingual language input and languages being learned in our sample.
-The same note applies to Experiment 2 below.} None of the adult participants reported fluency in a second language.
+Our child sample included 19 three-year-olds, 32 four-year-olds, and 23 five-year-olds, all enrolled in a local nursery school and all of whom volunteered their time.
+All participants were native English speakers.
+Approximately one-third (N$=$25) of the children's parents and teachers reported that their child regularly heard a second (and sometimes third or further) language, but only one child frequently heard a language that was used in our non-English video stimuli, and we excluded his data from the analyses.\footnote{Multilingual children may make predictions about upcoming turn structure differently from their monolingual peers due to their more varied experiences with linguistic cues to turn taking.
+However, we are unable to test this hypothesis here due to the variability in multilingual language input and the diverse set of languages being learned in our sample.
+The same applies to Experiment 2.} None of the adult participants reported fluency in a second language.
 
@@
 \textit{Video recordings}.
-We recorded pairs of talkers while they conversed in a sound-attenuated booth (see a sample frame in Figure \ref{fig:speakers}).
+We recorded pairs of talkers while they conversed in a sound-attenuated booth (Figure \ref{fig:speakers}).
 Each talker was a native speaker of the language being recorded, and each talker pair was male-female.
@@
 Then we asked talkers to choose a new topic---one relevant to young children (e.g., riding a bike, eating breakfast)---and to improvise a dialogue on that topic.
-We asked them to speak as if they were on a children's television show in order to elicit child-directed speech toward each other.
-We recorded until the talkers achieved at least 30 seconds of uninterrupted discourse with enthusiastic, child-directed speech.
+We asked them to speak as if they were on a children's television show in order to elicit child-friendly speech toward each other.
+We recorded until the talkers achieved at least 30 seconds of uninterrupted discourse with enthusiastic, child-friendly speech.
 Most talker pairs took less than five minutes to complete the task, usually by agreeing on a rough script at the start.
@@
 
-After recording, we combined the audio and video recordings by hand, and cropped each recording to the (approximate) 30-second interval with the most turn activity.
+After recording, we combined the audio and video recordings by hand, and cropped each one to the (approximate) 30-second interval with the most turn activity.
 Because we recorded the conversations in stereo, the male and female voices came out of separate speakers during video playback.
@@
 We coded each turn transition in the videos for language condition (English vs.
-non-English), inter-turn gap duration (in milliseconds), and speech act (question vs.
+non-English), inter-turn gap duration (in milliseconds), and transition type (question vs.
 non-question).
-The non-English stimuli were coded for speech act from a monolingual English-speaker's perspective, i.e., which turns ``sound like'' questions, and which do not: we asked five native American English speakers to listen to the audio recording for each non-English turn and judge whether it sounded like a question.
+Each non-English turn was coded as a question or non-question from a monolingual English-speaker's perspective, i.e., turns that ``sound like'' questions and turns that do not.
+We asked five native American English speakers to listen to the audio recording for each non-English turn and judge whether it sounded like a question.
 We marked non-English turns as questions when at least 4 of the 5 listeners (80\%) said that the turn ``sounded like a question''.
-This procedure means that the cues for recognizing ``questions'' in the non-English condition only resembled native English question cues, and therefore were likely harder to identify than cues to questionhood in the English condition.
-However, since participants did not speak the non-English languages and would only therefore ever treat ``question-sounding'' turns as questions, we proceeded with these analyses to see how pervasive question effects were---could they show up even without lexical access in a foreign language?
+Thus, ``question'' cues in the non-English condition only resembled native English question cues, and therefore were likely harder to identify than cues to questionhood in the English condition.
+However, since participants did not speak the non-English languages and would only ever treat ``question-sounding'' turns as questions, we proceeded with these analyses to see how pervasive question effects were---could they show up even without lexical access in a foreign language?
 If participants primarily rely on prosodic cues to question turns, it's possible that even non-English prosody can elicit anticipatory gaze switches for question-like turns.
@@
 These criteria guarantee that we only counted gaze shifts when: (1) participants were tracking the previous speaker, (2) switched their gaze to track the upcoming speaker, and (3) did so before they could have simply reacted to the onset of speech in the response.
-Under this assumption, a gaze shift that was initiated within the first 200 msec of the response (or earlier) was planned \textit{before} the child could react to the onset of speech itself.
+Under the assumption that it takes at least 200 msec to plan an eye movement, gaze shifts initiated within the first 200 msec of the response (or earlier) were planned \textit{before} participants could react to the onset of speech itself.
 
@@
 Again, because it takes 200 msec to plan an eye movement, we counted anticipatory switches, at the latest, 200 msec after the onset of speech.
-Therefore, for very early and very late switches, our requirement of 100 msec of fixation on each speaker would sometimes extend outside of the transition window boundaries (200 msec before and after the inter-turn gap).
+Therefore, for very early and very late switches, our requirement of 100 msec of fixation on each speaker would sometimes extend outside of the gaze launch window boundaries (200 msec before and after the inter-turn gap; Figure \ref{fig:criterion}).
 The maximally available fixation window was therefore 100 msec before and after the earliest and latest possible switch point (300 msec before and after the inter-turn gap).
@@
 \end{center}
-\caption{Schematic summary of criteria for anticipatory gaze shifts from speaker A to speaker B during a turn transition.} 
+\caption{Schematic summary of the criteria for anticipatory gaze shifts from speaker A to speaker B during a turn transition.} 
 \label{fig:criterion}
@@
 They primarily kept their eyes on the person who was currently speaking in both English and non-English videos: they gazed at the current speaker between 38\% and 63\% of the time, looking back at the addressee between 15\% and 20\% of the time (Table \ref{tab:e1_look}).
-Even three-year-olds looked more at the current speaker than anything else, whether the videos were in a language they could understand or not.
+Even three-year-olds looked more at the current speaker than anything else, whether or not the videos were in a language they could understand.
 Children looked at the current speaker less than adults did during the non-English videos.
@@
 \end{center}
-  \caption{Average proportion of gaze to the current speaker and addressee during periods of talk.}
+  \caption{Average proportion of gaze to the current speaker and addressee during periods of talk across ages in Experiment 1.}
 \label{tab:e1_look}
@@
     LgCond$=$\textit{non-English} 						& -3.65		& 1.16	& -3.146  & 0.002 **	\\
-    Type$=$\textit{non-Question}      					& -2.95		& 1.13	& -2.61	  & 0.009 **	\\
+    TType$=$\textit{non-Question}      					& -2.95		& 1.13	& -2.61	  & 0.009 **	\\
     GapDuration      											&  2.247	& 3.194	&  0.704  & 0.482		\\
     Age*LgCond$=$\textit{non-English} 					&  0.5		& 0.212	&  2.353  & 0.019 *	\\
-    Age*Type$=$\textit{non-Question} 					&  0.009	& 0.196	&  0.044  & 0.965		\\
+    Age*TType$=$\textit{non-Question} 					&  0.009	& 0.196	&  0.044  & 0.965		\\
     LgCond$=$\textit{non-English}*						&  2.692	& 1.347	&  1.999  & 0.046 *	\\
-    \hspace*{5mm} Type$=$\textit{non-Question} &&&& \\
+    \hspace*{5mm} TType$=$\textit{non-Question} &&&& \\
     Age*GapDuration												& -0.577	& 0.627	& -0.921  & 0.357		\\
     LgCond$=$\textit{non-English}*GapDuration		&  1.143	& 2.287	&  0.5	  & 0.617		\\
-    Type$=$\textit{non-Question}*GapDuration			&  5.519	& 2.282	&  2.418  & 0.016 *	\\
+    TType$=$\textit{non-Question}*GapDuration			&  5.519	& 2.282	&  2.418  & 0.016 *	\\
     Age*LgCond$=$\textit{non-English}*					& -0.433	& 0.304	& -1.426  & 0.154		\\
-    \hspace*{5mm} Type$=$\textit{non-Question} &&&& \\
+    \hspace*{5mm} TType$=$\textit{non-Question} &&&& \\
     \hline
@@
     LgCond$=$\textit{non-English}							& -0.059   & 0.751 & -0.079 & 0.937			\\
-    Type$=$\textit{non-Question}							& -3.298   & 0.933 & -3.536 & 0.0004 ***	\\
+    TType$=$\textit{non-Question}							& -3.298   & 0.933 & -3.536 & 0.0004 ***	\\
     GapDuration														&  0.132   & 1.766 &  0.075 & 0.941			\\ 
     LgCond$=$\textit{non-English}*						&  1.234   & 0.629 &  1.961 & 0.0498 *		\\
-    \hspace*{5mm} Type$=$\textit{non-Question} &&&& \\
+    \hspace*{5mm} TType$=$\textit{non-Question} &&&& \\
     LgCond$=$\textit{non-English}*GapDuration		& -1.519   & 2.192 & -0.693 & 0.488			\\
-    Type$=$\textit{non-Question}*GapDuration			&  7.116   & 2.195 &  3.241 & 0.001 **		\\
+    TType$=$\textit{non-Question}*GapDuration			&  7.116   & 2.195 &  3.241 & 0.001 **		\\
     \hline
@@
   \end{small}
-  \caption{Model output for children and adults' anticipatory gaze switches in Experiment 1.}
+  \caption{Model output for participants' anticipatory gaze switches in Experiment 1.}
 \label{tab:E1-models}
@@
 
-We identified anticipatory gaze switches for all 36 usable turn transitions, based on the criteria outlined in Section \ref{sec:algorithm}, and analyzed them for effects of language, transition type, and age with two mixed-effects logistic regressions \citep{lme4, R}.
+We identified anticipatory gaze switches for all 36 usable turn transitions, based on the criteria outlined above, and analyzed them for effects of language, transition type, and age with two mixed-effects logistic regressions \citep{lme4, R}.
 We built one model each for children and adults.
 We modeled children and adults separately because effects of age are only pertinent to the children's data.
+
 The child model included condition (English vs.
@@
 We would need multiple items from each of the languages to check for similarity effects of specific linguistic features.}, transition type (question vs.
-non-question), age (3, 4, 5; numeric; intercept as age=0), and duration of the inter-turn gap (seconds, e.g., 0.441) as predictors, with two-way interactions between gap duration and the other simple predictors (language condition, transition type, and age), also adding a three-way interaction between language condition, transition type, and age.
-We included the two-way interactions between gap duration and the other predictors in case the effect of inter-turn silence changes with age or linguistic cueing (e.g., if children older children rely less on silence as a cue)\footnote{We test these two-way interactions with gap duration in all of the models reported in this paper.
+non-question), age (3, 4, 5; numeric; intercept as age=0), and duration of the inter-turn gap (seconds, e.g., 0.441) as predictors, with two-way interactions between gap duration and the other simple fixed effects (language condition, transition type, and age) and a three-way interaction between language condition, transition type, and age.
+We included the two-way interactions with gap duration in case the effect of inter-turn silence changes with age or linguistic cueing (e.g., if children older children rely less on silence as a cue)\footnote{We test these two-way interactions with gap duration in all of the models reported in this paper.
 Higher-order interactions with gap duration usually resulted in model non-convergence due to distributional sparsity when three or more predictor values were considered.}.
 We also included random effects of item (turn transition) and participant, with maximal random slopes of condition, transition type, and their interaction for participants \citep{barr2013}.\footnote{The models we report in this paper are all qualitatively unchanged by the exclusion of their random slopes.
-We have left the random slopes in because of minor participant-level variation in the predictors modeled.} The adult model included fixed effects of condition, transition type, and their interaction, plus two-way interactions between gap duration and language condition and transition type (as in the child model).
+We have left the random slopes in because of minor participant-level variation in the predictors modeled.}
+
+The adult model included fixed effects of condition, transition type, and their interaction, plus two-way interactions between gap duration and the other simple fixed effects (language condition and transition type, as in the child model).
 The adult model also included random effects of item and participant with maximal random slopes of condition, transition type, and their interaction for participant.
 
-Children's anticipatory gaze switches showed effects of language condition (\textit{$\beta$}=-3.65, \textit{SE}=1.16, \textit{z}=-3.15, \textit{p}$<$.01) and transition type (\textit{$\beta$}=-2.95, \textit{SE}=1.13, \textit{z}=-2.61, \textit{p}$<$.01) with additional effects of an age-by-language condition interaction (\textit{$\beta$}=0.5, \textit{SE}=0.212, \textit{z}=2.35, \textit{p}$<$.05), a language condition-by-transition type interaction (\textit{$\beta$}=2.69, \textit{SE}=1.35, \textit{z}=1.99, \textit{p}$<$.05), and an interaction between transition type and gap duration (\textit{$\beta$}=5.52, \textit{SE}=2.28, \textit{z}=2.42, \textit{p}$<$.05).
+Children's anticipatory gaze switches showed effects of language condition (\textit{$\beta$}=-3.65, \textit{SE}=1.16, \textit{z}=-3.15, \textit{p}$<$.01) and transition type (\textit{$\beta$}=-2.95, \textit{SE}=1.13, \textit{z}=-2.61, \textit{p}$<$.01) with additional effects of an age-by-language condition interaction (\textit{$\beta$}=0.5, \textit{SE}=0.212, \textit{z}=2.35, \textit{p}$<$.05), a language condition-by-transition type interaction (\textit{$\beta$}=2.69, \textit{SE}=1.35, \textit{z}=1.99, \textit{p}$<$.05), and a transition type-gap duration interaction (\textit{$\beta$}=5.52, \textit{SE}=2.28, \textit{z}=2.42, \textit{p}$<$.05).
 There were no significant effects of age or gap duration alone (\textit{$\beta$}=-0.002, \textit{SE}=0.26, \textit{z}=-0.009, \textit{p}$=$.99 and \textit{$\beta$}=2.25, \textit{SE}=3.19, \textit{z}=0.7, \textit{p}$=$.48, respectively).
@@
 Our primary analysis (above) makes the assumption that participants' eye movements generally follow the turn structure of the stimulus, i.e., that participants track the current speaker and switch their gaze to the upcoming speaker near turn transitions.
-Based on this assumption, we used linear mixed effects regressions to see how anticipatory looking is affected by aspects of participant group (age) and stimulus (e.g., transition type, language condition).
+As just described, based on this assumption, we used linear mixed effects regressions to see how anticipatory looking is affected by aspects of participant group (age) and stimulus (e.g., transition type, language condition).
 But what if the assumption that participants generally track turn structure were wrong?
 Could these results have emerged if participants' eye movements were \textit{not} linked to turn structure?
-For example, if participants were randomly looking back and forth between the two speakers, we might still find some anticipatory switching by chance---we would see anticipatory switching no matter where the real turn transitions occurred.
-To test whether our primary results (the regression output above) could have arisen from random switching, not linked to turn structure, we conducted a secondary analysis comparing participants' anticipatory gaze at \textit{real} and \textit{randomly shuffled} points of turn transition.
+For example, if participants were randomly looking back and forth between the two speakers, we might still find some anticipatory switching by chance.
+To test whether our primary results (the regression output above) could have arisen from random switching, not linked to turn structure, we conducted a secondary analysis comparing participants' anticipatory gaze at real and randomly shuffled points of turn transition.
 
 We conducted this analysis by running the same regression models on participants' eye-tracking data, only this time calculating their anticipatory gaze switches with respect to randomly permuted turn transition windows.
-This process involved: (1) randomizing the order and temporal placement of the analysis windows within each stimulus (see Figure \ref{fig:shuffling}; ``analysis window'' is defined in Figure \ref{fig:criterion}) to randomly redistribute the analysis windows across the eye-tracking signal, (2) re-running each participant's eye tracking data through switch identification (described in Section \ref{sec:algorithm}) on each of the randomly permuted analysis windows, and (3) modeling the anticipatory switches from the randomly permuted data (our random baseline dataset) with the same statistical models we used for the original dataset (Section \ref{sec:models1}; Table \ref{tab:E1-models}).
+This process involved: (1) randomizing the order and temporal placement of the analysis windows within each stimulus (see Figure \ref{fig:shuffling}; ``analysis window'' as shown in Figure \ref{fig:criterion}) to randomly redistribute the analysis windows across the eye-tracking signal, (2) re-running each participant's eye tracking data through switch identification (described above) on each of the randomly permuted analysis windows, and (3) modeling the anticipatory switches from the randomly permuted data (our random baseline dataset) with the same statistical models we used for the original dataset (Table \ref{tab:E1-models}).
 Importantly, although the onset time of each transition was shuffled within the eye-tracking signal, the other intrinsic properties of each turn transition (e.g., prior speaker identity, transition type, gap duration, language condition, etc.) stayed constant across each permutation.
@@
 We created 5,000 permutations of the original turn transitions (step 1 above), thereby creating 5,000 anticipatory gaze datasets with randomly de-linked gaze data (step 2 above).
-Because the randomly shuffled turn transitions could occur anywhere in the stimulus (so long as they didn't overlap each other), the resulting turn-transition windows collectively covered the entire stimulus---during speech and silence, during speaker change and speaker continuation, and during all turn transitions in the stimulus, even those excluded in the original analyses (e.g., because they were overlapped).\footnote{This technique crucially differs from that used by Keitel and colleagues (2013, 2015), which tests anticipatory gaze at turn transitions against anticipatory gaze during speech, thereby maximizing the possibility of finding a difference from the baseline measure.} Pooled together, the anticipatory gaze datasets yield an average anticipatory switch rate for each participant over all possible starting points in the stimuli: a random baseline.
-
-Using this technique we can compare participants' anticipatory switches at turn transition windows to their anticipatory switches over the stimulus as a whole.
-If participants were looking randomly back and forth between the speakers, we should similar patterns in both cases.
+Because the randomly shuffled turn transitions could occur anywhere in the stimulus (so long as they didn't overlap each other within a single iteration), the resulting turn-transition windows collectively covered the entire stimulus---during speech and silence, during speaker change and speaker continuation, and during all turn transitions in the stimulus, even those excluded in the original analyses (e.g., because they were overlapped).\footnote{This technique crucially differs from that used by Keitel and colleagues (2013, 2015), which tests anticipatory gaze at turn transitions against anticipatory gaze during speech, which maximized the possibility of finding a difference from the baseline measure.} Pooled together, the anticipatory gaze datasets yielded an average anticipatory switch rate for each participant over all possible starting points in the stimuli: a random baseline.
+
+Using this technique we compared participants' anticipatory switches at turn transition windows to their anticipatory switches over the stimulus as a whole.
+If participants looked randomly back and forth between the speakers, we would have seen similar patterns in both cases.
 Rather than simply comparing participants' overall anticipatory switch rates with real and random transition windows, we estimated the likelihood that each of the predictor effects in the original data (e.g., the effect of language condition; Table \ref{tab:E1-models}) could have arisen with random gaze switching: we ran identical statistical models on the real and randomly permuted data sets.
-This tells us not only whether participants' switches were above chance, but whether the specific underlying effects of their anticipatory gaze patterns (e.g., the effect of linguistic condition) were above that expected by chance.
-But, because these analyses are complex and secondary to the main results, we report their full details in \ref{sec:permutation}.
+This tells us not only whether participants' switches were above chance, but whether the specific underlying effects of their anticipatory gaze patterns (e.g., the effect of language condition) were above that expected by chance.
+Because these analyses are complex and secondary to the main results, we report their full details in \ref{sec:permutation}.
 
@@
 For the children's data, the original \textit{z}-values for language condition, transition type, the age-language condition interaction, the transition type-gap duration interaction, and the language condition-transition type interaction were all greater than 95\% of \textit{z}-values from models of the randomly permuted data (99.3\%, 99.1\%, 98.9\%, 97\%, and 96\%, respectively, all \textit{p}$<$.05).
-Similarly, the adults' data showed significant differentiation from the randomly permuted data for all three significant predictors from the real transition dataset---transition type, the interaction between transition type and gap duration, and the interaction between language condition and transition type (greater than 100\%, 99.8\%, and 95\% of random \textit{z}-values, respectively, all \textit{p}$\leq$.05).
+Similarly, the adults' data showed significant differentiation from the randomly permuted data for all three significant predictors from the real transition dataset.
+Transition type, the interaction between transition type and gap duration, and the interaction between language condition and transition type showed \textit{z}-values that exceeded 100\%, 99.8\%, and 95\% of random \textit{z}-values, respectively (all \textit{p}$\leq$.05).
 See Section \ref{sec:permutation} for more information on each predictor's random permutation distribution.\footnote{This baseline analysis tests ``random looking'' against ''turn-driven looking'', but it does not test subtypes of turn-driven looking.
 For example, children might switch their gaze from the current speaker to the addressee out of boredom with the ongoing speech rather than active anticipation of an upcoming response.
-We address this hypothesis about ``turn-transition'' gaze switchers vs.
-``boredom'' gaze switchers in \ref{sec:boredlooks}}
+We address this hypothesis about ``turn-transition'' gaze switches vs.
+``boredom'' gaze switches in \ref{sec:boredlooks}}
 
@@
 
-Given these findings, when does spontaneous turn prediction emerge developmentally?
+When does spontaneous turn prediction emerge developmentally?
 We tested whether the youngest age group (3-year-olds) already exceeded chance in their anticipatory gaze switches by comparing children's real gaze rates to the random baseline in the English condition with two-tailed \textit{t}-tests.
@@
 This interaction between age and language condition suggests that the 5-year-olds were able to leverage anticipatory cues in the non-English videos in a way that 3- and 4-year-olds could not, possibly by shifting more attention to the non-English prosodic or non-verbal cues.
-Prior work on children's turn-structure anticipation has proposed that children's turn-end predictions rely primarily on lexicosyntactic structure (and not, e.g., prosody) as they get older \citep{keitel2013}.
-The current results suggest more flexibility in children's predictions; when they do not have access to lexical information, older children and adults are likely to find alternative cues to turn taking behavior.
+Prior work on children's turn-structure anticipation has proposed that children's turn-end predictions rely primarily on lexicosemantic structure (and not, e.g., prosody) as they get older \citep{keitel2013}.
+The current results suggest more flexibility in children's predictions; when they do not have access to lexical information, older children and adults find alternative cues to turn taking behavior.
 
@@
 This finding suggests that gap duration indeed serves as a cue to upcoming turn structure; while short gaps may be perceived as within-turn pauses \citep{mannel2009}, long gaps could instead be indicative of between-turn pauses (where speaker transition occurs).
-Participants might use these long silences to retroactively assign turn boundaries and anticipate speaker switches that were otherwise not anticipated (in this case, because the preceding turn was not a question).
+Participants might use long silences to retroactively assign turn boundaries and anticipate speaker switches that were otherwise not anticipated (in this case, because the preceding turn was not a question).
 An alternative explanation for gap duration effects is that longer inter-turn gaps result in longer analysis windows, which yields more time for participants to make an anticipatory gaze.
@@
 In Experiment 2, we compared lexicosyntactic and prosodic cues with phonetically altered speech and used puppets to eliminate non-verbal cues to turn taking.
-Second, we were not able to pinpoint the emergence of anticipatory switching because the youngest age group in our sample was already able to make anticipatory switches at above chance rates.
+Second, we were not able to pinpoint the emergence of anticipatory switching because the youngest age group in our sample was already able to make anticipatory switches at above-chance rates.
 In Experiment 2, we explored a wider developmental range.
@@
 We recruited 27 undergraduate adults and 129 children ages 1;0--6;11 to participate in our experiment.
-We recruited adult participants via the Stanford University Psychology participant database.
-Adult participants were either paid or received course credit for their time.
+Adult participants were recruited again via the Stanford University Psychology participant database and were either paid or received course credit for their time.
 We recruited our child participants from the Children's Discovery Museum in San Jose, California\footnote{We ran Experiment 2 at a local children's museum because it gave us access to children with a wider range of ages.
@@
 
-The \textit{words only} conversations were manipulated to have robot-like speech: we flattened the intonation contours to each talker's average pitch (F$_{0}$) and we reset the duration of every nucleus and coda to each talker's average nucleus and coda duration.\footnote{We excluded hyper-lengthened words like [w\textipa{aU:}] `woooow!', but these were rare to begin with.} We made duration and pitch manipulations using PSOLA resynthesis in Praat \citep{Praat}.
+The \textit{words only} conversations were manipulated to have robot-like speech: we flattened the intonation contours to each talker's average pitch (F$_{0}$) and we reset the duration of every nucleus and coda to each talker's average nucleus and coda duration.\footnote{We excluded hyper-lengthened words like [w\textipa{aU:}] `woooow!'.} We made duration and pitch manipulations using PSOLA resynthesis in Praat \citep{Praat}.
 Thus, the \textit{words only} versions of the conversations had no pitch or durational cues to upcoming turn boundaries, but did have intact lexicosyntactic cues (and some residual phonetic correlates of prosody, e.g., intensity).
@@
 We created puppet video recordings to match the manipulated 20-second audio clips.
-The puppets were minimally expressive; the puppeteer could only control the opening and closing of their mouths.
-The puppets' heads, eyes, arms, and bodies stayed still.
+The puppets were minimally expressive; the puppeteer could only control the opening and closing of their mouths, and the puppets' heads, eyes, arms, and bodies stayed still.
 Puppets were positioned side-by-side, looking in the same direction to eliminate shared gaze as a cue to turn structure \citep{thorgrimsson2015}.
@@
 
-We used three pairs of puppets used for the \textit{normal} condition---`red', `blue' and `yellow'---and one pair of puppets for each partial cue condition: `robots', `merpeople', and `party-goers' (Figure 8).
+We used three pairs of puppets for the \textit{normal} condition---`red', `blue' and `yellow'---and one pair of puppets for each partial cue condition: `robots', `merpeople', and `party-goers' (Figure 8).
 We randomly assigned half of the conversation topics (`birthday cake', `pets', and `breakfast') to the \textit{normal} condition, and half to the partial cue conditions (`riding bikes', `rainy days', and `the library').
@@
 There were 85 turn transitions for analysis after excluding transitions longer than 550 msec and shorter than 90 msec.
-The remaining turn transitions had more questions than non-questions (N$=$47 and N$=$38, respectively), with transitions distributed somewhat evenly across conditions (keeping in mind that there were three \textit{normal} videos and only one partial cue video for each experiment version): \textit{normal} (N$=$36), \textit{words only} (N$=$13), \textit{prosody only} (N$=$17), and \textit{no speech} (N$=$19).
+The remaining turn transitions had more questions than non-questions (N$=$47 and N$=$38, respectively), with transitions distributed somewhat evenly across conditions (keeping in mind that there were three \textit{normal} videos and only one video for each partial cue condition in each experiment version): \textit{normal} (N$=$36), \textit{words only} (N$=$13), \textit{prosody only} (N$=$17), and \textit{no speech} (N$=$19).
 Inter-turn gaps for questions (mean$=$366, median$=$438, stdev$=$138 msec) were longer than those for non-questions (mean$=$305, median$=$325, stdev$=$94 msec) on average, but gap duration was overall comparable across conditions: \textit{normal} (mean$=$334, median$=$321, stdev$=$130 msec), \textit{words only} (mean$=$347, median$=$369, stdev$=$ 115 msec), \textit{prosody only} (mean$=$365, median$=$369, stdev$=$104 msec), and \textit{no words} (mean$=$319, median$=$329, stdev$=$136 msec).
@@
 We used the same experimental apparatus and procedure as in the first experiment.
-Each participant watched six puppet videos in random order, with five 15--30 second filler videos placed in-between (e.g., running puppies, moving balls, flying bugs).
+Each participant watched six puppet videos in random order, with 15--30 second filler videos placed in-between (e.g., running puppies, moving balls, flying bugs).
 Three of the puppet videos had \textit{normal} audio while the other three had \textit{words only}, \textit{prosody only}, and \textit{no speech} audio.
 As before, the experimenter immediately began each session with calibration and then stimulus presentation.
-Participants were given no instruction about how to watch the videos or what their purpose was, they were simply encouraged to watch the ``fun/nice puppet videos''.
+Participants were given no instruction about how to watch the videos or what their purpose was, they were simply encouraged to watch the ``(fun/nice) puppet videos''.
 The entire experiment took less than five minutes.
@@
 \end{center}
-  \caption{Average proportion of gaze to the current speaker and addressee during periods of talk across ages.}
+  \caption{Average proportion of gaze to the current speaker and addressee during periods of talk across ages in Experiment 2.}
 \label{tab:look_e2}
@@
 \end{center}
-  \caption{Average proportion of gaze to the current speaker and addressee during periods of talk across conditions.}
+  \caption{Average proportion of gaze to the current speaker and addressee during periods of talk across conditions in Experiment 2.}
 \label{tab:look_e2b}
@@
 We coded each turn transition for its linguistic condition (\textit{normal}, \textit{words only}, \textit{prosody only}, and \textit{no speech}) and transition type (question/non-question)\footnote{We coded \textit{wh-}questions as ``non-questions'' for the \textit{prosody only} videos.
-Polar questions often have a final rising intonational contour, but \textit{wh-}questions often do not  \citep{hedberg2010}.}, and identified anticipatory gaze switches to the upcoming speaker using the methods from Experiment 1.
+Polar questions often have a final rising intonational contour, but \textit{wh-}questions do not  \citep{hedberg2010}.}, and identified anticipatory gaze switches to the upcoming speaker using the methods from Experiment 1.
 
@@
 Participants' pattern of gaze indicated that they performed basic turn tracking across all ages and in all conditions.
-Participants looked at the screen most of the time during video playback (82\% and 86\% average for children and adults, respectively), primarily looking at the person who was currently speaking (Table 2).
+Participants looked at the screen most of the time during video playback (82\% and 86\% average for children and adults, respectively), primarily looking at the person who was currently speaking (Tables \ref{tab:look_e2} and \ref{tab:look_e2b}).
 They tracked the current speaker in every condition---even one-year-olds looked more at the current speaker than at anything else in the three partial cue conditions (40\% for \textit{words only}, 43\% for \textit{prosody only}, and 39\% for \textit{no speech}).
@@
 The child model included condition (\textit{normal}/\textit{prosody only}/\textit{words only}/\textit{no speech}; with \textit{no speech} as the reference level), transition type (question vs.
-non-question), age (1, 2, 3, 4, 5, 6; numeric, intercept as age=0), and duration of the inter-turn gap (in seconds) as predictors, with full interactions between language condition, transition type, and age and two-way interactions between gap duration and the other basic predictors (age, linguistic condition, and transition type).
+non-question), age (1, 2, 3, 4, 5, 6; numeric, intercept as age=0), and duration of the inter-turn gap (in seconds) as predictors, with full interactions between language condition, transition type, and age and two-way interactions between gap duration and the other basic fixed effects (age, linguistic condition, and transition type).
 We also included random effects of participant and item (turn transition), with maximal random slopes of transition type for participant.
@@
 Children's anticipatory gaze switches showed an effect of gap duration (\textit{$\beta$}=3.85, \textit{SE}=1.73, \textit{z}=2.22, \textit{p}$<$.05), a two-way interaction of age and language condition (for \textit{prosody only} speech compared to the \textit{no speech} reference level; \textit{$\beta$}=0.38, \textit{SE}=0.19, \textit{z}=1.97, \textit{p}$<$.05), a marginal two-way interaction of language condition and gap duration (for \textit{prosody only} speech compared to the \textit{no speech} reference level; \textit{$\beta$}=-4.77, \textit{SE}=2.63, \textit{z}=-1.82, \textit{p}$=$.07), and a three-way interaction of age, transition type, and language condition (for \textit{normal} speech compared to the \textit{no speech} reference level; \textit{$\beta$}=-0.35, \textit{SE}=0.17, \textit{z}=-2.05, \textit{p}$<$.05).
-There were no significant effects of age or transition type alone (Table \ref{sec:models2}) (\textit{$\beta$}=-0.05, \textit{SE}=0.14, \textit{z}=-0.38, \textit{p}$=$.7 and \textit{$\beta$}=-1.22, \textit{SE}=0.96, \textit{z}=-1.27, \textit{p}$=$.2, respectively)
+There were no significant effects of age or transition type alone (Table \ref{tab:E2-chimodels}; \textit{$\beta$}=-0.05, \textit{SE}=0.14, \textit{z}=-0.38, \textit{p}$=$.7 and \textit{$\beta$}=-1.22, \textit{SE}=0.96, \textit{z}=-1.27, \textit{p}$=$.2, respectively)
 
 Adults' anticipatory gaze switches showed a significant effect of language condition (for \textit{words only} speech compared to the \textit{no speech} reference level; \textit{$\beta$}=3.79, \textit{SE}=1.62, \textit{z}=2.34, \textit{p}$<$.05) and a marginal two-way interaction between language condition and transition type (for \textit{words only} speech compared to the \textit{no speech} reference level; \textit{$\beta$}=-1.68, \textit{SE}=0.89, \textit{z}=-1.89, \textit{p}$=$.06).
-There was no significant effect of transition type alone (\textit{$\beta$}=-0.02, \textit{SE}=1.44, \textit{z}=-0.02, \textit{p}$=$.99).
+There was no significant effect of transition type alone (Table \ref{tab:E2-adumodels}; \textit{$\beta$}=-0.02, \textit{SE}=1.44, \textit{z}=-0.02, \textit{p}$=$.99).
 
@@
 
-Using the same technique described in Experiment 1 (Section \ref{sec:randbaseline1}), we created and modeled random permutations of participants' anticipatory gaze switches.
+Using the same technique described in Experiment 1, we created and modeled random permutations of participants' anticipatory gaze switches.
 These analyses revealed that the significant predictors from models of the original, turn-related data were unlikely to be explained by random looking.
@@
     Age														& -0.054		& 0.143	& -0.379	& 0.705				\\
-    Type$=$\textit{non-Question}				& -1.217		& 0.958	& -1.27		& 0.204				\\
+    TType$=$\textit{non-Question}				& -1.217		& 0.958	& -1.27		& 0.204				\\
     GapDuration											&  3.852		& 1.735	&  2.221	& 0.026 *			\\
-    Age*Type$=$\textit{non-Question}			&  0.152		& 0.141	&  1.081	& 0.28				\\
+    Age*TType$=$\textit{non-Question}		&  0.152		& 0.141	&  1.081	& 0.28				\\
     Age*GapDuration									&  0.214		& 0.266	&  0.805	& 0.421				\\
-    Type$=$\textit{non-Question}*				&  0.995		& 2.134	&  0.466	& 0.641				\\
+    TType$=$\textit{non-Question}*			&  0.995		& 2.134	&  0.466	& 0.641				\\
     \hspace*{5mm} GapDuration &&&& \\
@@
     Condition$=$\textit{normal}*				&  0.908		& 0.748	&  1.215	& 0.224				\\
-    \hspace*{5mm} Type$=$\textit{non-Question} &&&& \\
+    \hspace*{5mm} TType$=$\textit{non-Question} &&&& \\
     Age*Condition$=$\textit{normal}*			& -0.355		& 0.173	& -2.051	& 0.04 *			\\
-    \hspace*{5mm} Type$=$\textit{non-Question} &&&& \\
+    \hspace*{5mm} TType$=$\textit{non-Question} &&&& \\
     Condition$=$\textit{normal}*				& -0.431		& 1.67	& -0.258	& 0.797				\\
@@
     Condition$=$\textit{prosody}*				&  1.076		& 1.105	&  0.974	& 0.33				\\
-    \hspace*{5mm} Type$=$\textit{non-Question} &&&& \\
+    \hspace*{5mm} TType$=$\textit{non-Question} &&&& \\
     Age*Condition$=$\textit{prosody}*		& -0.296		& 0.235	& -1.257	& 0.209				\\
-    \hspace*{5mm} Type$=$\textit{non-Question} &&&& \\
-    Condition$=$\textit{prosody}*				& -4.767		& 2.625	& -1.816	& 0.069 .			\\
+    \hspace*{5mm} TType$=$\textit{non-Question} &&&& \\
+    Condition$=$\textit{prosody}*				& -4.767		& 2.625	& -1.816	& 0.069 (.)			\\
     \hspace*{5mm} GapDuration &&&& \\
@@
     Condition$=$\textit{words}*				& -1.244		& 1.031	& -1.207	& 0.228					\\
-    \hspace*{5mm} Type$=$\textit{non-Question} &&&& \\
+    \hspace*{5mm} TType$=$\textit{non-Question} &&&& \\
     Age*Condition$=$\textit{words}*		&  0.111		& 0.225	&  0.495	& 0.621					\\
-    \hspace*{5mm} Type$=$\textit{non-Question} &&&& \\
+    \hspace*{5mm} TType$=$\textit{non-Question} &&&& \\
     Condition$=$\textit{words}*				& -2.285		& 2.232	& -1.024	& 0.306					\\
@@
     (Intercept)										& -3.117		& 1.176	& -2.649	& 0.008 **				\\ 
-    Type$=$\textit{non-Question}			& -0.022		& 1.44	& -0.015	& 0.988					\\ 
+    TType$=$\textit{non-Question}			& -0.022		& 1.44	& -0.015	& 0.988					\\ 
     GapDuration										&  4.073		& 2.947	&  1.382	& 0.167					\\ 
-    Type$=$\textit{non-Question}*			&  1.304		& 3.859	&  0.338	& 0.735					\\
+    TType$=$\textit{non-Question}*		&  1.304		& 3.859	&  0.338	& 0.735					\\
     \hspace*{5mm} GapDuration &&&& \\
@@
     Condition$=$\textit{normal}*			& -0.709		& 0.754	& -0.94		& 0.347					\\ 
-    \hspace*{5mm} Type$=$\textit{non-Question} &&&& \\
+    \hspace*{5mm} TType$=$\textit{non-Question} &&&& \\
     Condition$=$\textit{normal}*			&  2.1			& 3.336	&  0.629	& 0.529					\\ 
@@
     Condition$=$\textit{prosody}*			&  0.386		& 1.065	&  0.362	& 0.717					\\ 
-    \hspace*{5mm} Type$=$\textit{non-Question} &&&& \\
+    \hspace*{5mm} TType$=$\textit{non-Question} &&&& \\
     Condition$=$\textit{prosody}*			& -1.118		& 4.543	& -0.246	& 0.805					\\ 
@@
     Condition$=$\textit{words}				&  3.792		& 1.621	&  2.338	& 0.019 *				\\ 
-    Condition$=$\textit{words}*				& -1.678		& 0.889	& -1.888	& 0.059 .				\\ 
-    \hspace*{5mm} Type$=$\textit{non-Question} &&&& \\
+    Condition$=$\textit{words}*				& -1.678		& 0.889	& -1.888	& 0.059 (.)				\\ 
+    \hspace*{5mm} TType$=$\textit{non-Question} &&&& \\
     Condition$=$\textit{words}*				& -5.653		& 3.861	& -1.464	& 0.143					\\ 
@@
 Our main goal in extending the age range to 1- and 2-year-olds in Experiment 2 was to find the age of emergence for spontaneous predictions about upcoming turn structure.
-As in Experiment 1, we used two-tailed \textit{t}-tests to compare children's real gaze rates to the random baseline rates in the \textit{normal} speech condition, in which the speech stimulus is most like what children hear every day.
+As in Experiment 1, we used two-tailed \textit{t}-tests to compare children's real gaze rates to the random baseline rates in the \textit{normal} speech condition (in which the speech stimulus is most like what children hear every day).
 We tested real gaze rates against baseline rates for three age groups: one-, two-, and three-year-olds.
@@
 In contrast, children's anticipatory switch rates for non-question turns were not significantly different between the \textit{no speech} and \textit{normal} speech conditions at any age (all \textit{p}$>$0.09).
-Consistent with the regression output, children were more likely to show an effect of linguistic content as they got older, but only for question transitions.
+Consistent with the regression results, children were more likely to show an effect of linguistic content as they got older, but only for question transitions.
 
@@
 \textit{prosody only}) within each random permutation of the data, making pairwise comparisons between the six age groups.
-These tests revealed that children's anticipation in the \textit{prosody only} condition significantly improved at ages five and six (with difference scores greater than 95\% of the random data scores; \textit{p}$<$.05).
+These tests revealed that children's anticipation in the \textit{prosody only} condition significantly improved at ages five and six compared to the \textit{no speech} baseline (with difference scores greater than 95\% of the random data scores; \textit{p}$<$.05).
 See Figure \ref{fig:E2-lgageinteraction} for these \textit{prosody only} difference score distributions.
@@
 As in Experiment 1, children and adults spontaneously tracked the turn structure of the conversations.
-Participants made anticipatory gaze switches at above-chance rates starting at age one for question transitions and age two for both questions and non-questions.
-Longer gaps had a broader impact on participants' anticipations in this experiment; we saw that, overall, longer inter-turn gaps resulted in more anticipatory switches, with the \textit{no speech} condition showing equal or significantly stronger effects of gap duration than all other conditions.
+Participants made anticipatory gaze switches at above-chance rates starting at age two for both questions and non-questions.
+Longer gaps had a broader impact on participants' anticipations in this experiment; we saw that, overall, longer inter-turn gaps resulted in more anticipatory switches, with the \textit{no speech} condition showing equal or stronger effects of gap duration than all other conditions.
 
@@
 These findings do not support an early role for prosody or lexical information alone in children's spontaneous predictions about turn structure.
-They also give no evidence for the idea that lexical information is sufficient on its own to support children's anticipatory switching.
-They do underscore the developing relationship between the use of linguistic cues and speech act (transition type) in spontaneous predictions about upcoming turn structure.
+They also give no support for the idea that lexical information is sufficient on its own to support children's anticipatory switching.
+They do underscore the developing relationship between the use of linguistic cues, inter-turn silence, and speech act (transition type) in spontaneous predictions about upcoming turn structure.
 
@@
 In two experiments investigating children's anticipatory gaze to upcoming speakers, we found evidence that turn prediction develops early in childhood and that, when spontaneous predictions begin, they are primarily driven by participants' expectation of an immediate response in the next turn (e.g., after questions).
-In making predictions about upcoming turn structure, children used a combination of lexical and prosodic cues; neither lexical nor prosodic cues alone were sufficient to support increased anticipatory gaze.
+In making predictions about upcoming turn structure, children used a combination of lexical and prosodic cues; neither signal alone was sufficient to support increased anticipatory gaze.
 We also found no early advantage for prosody over lexicosyntax; children's anticipatory switch rates in the \textit{prosody only} condition were initially low, but showed significant gains by age five.
-We discuss these findings with respect to: the role of linguistic cues and inter-turn silence for predicting upcoming turn structure, the importance of questions in predictions about conversation, and children's developing competence as conversationalists.
+We discuss these findings with respect to the role of linguistic cues and inter-turn silence for predicting upcoming turn structure, the importance of questions in predictions about conversation, and children's developing competence as conversationalists.
 
@@
 
-Prior work with adults has found a consistent role for lexicosyntax in predicting upcoming turn structure \citep{de-ruiter2006, magyari2012}, whereas the role of prosody still under debate \citep{duncan1972, ford1996, torreira2015}.
-Knowing that children comprehend more about prosody than lexicosyntax early on (Section \ref{sec:intro}; also see \citealp{speer2009} for a review), we thought it possible that young children would instead show an advantage for prosody in their predictions about turn structure in conversation.
+Prior work with adults has found a consistent role for lexicosyntax in predicting upcoming turn structure \citep{de-ruiter2006, magyari2012}, whereas the role of prosody is still under debate \citep{duncan1972, ford1996, torreira2015}.
+Knowing that children comprehend more about prosody than lexicosyntax early on (see \citealp{speer2009} for a review), we thought it possible that young children would instead show an advantage for prosody in their predictions about turn structure in conversation.
 Our results suggest that, on the contrary, exclusively presenting prosodic information to children limits their spontaneous predictions about upcoming turn structure until age five.
 
-Using prosody alone to accurately predict turn boundaries in conversation appears to be difficult for adults and children.
-Prosodic information is continuous, multidimensional, and indexically complex---it encodes syntactic structure, speech act, and extralinguistic orientation, sometimes simultaneously, though without clear one-to-one mappings between form and meaning \citep{cutler1997, shriberg1998, lammertink2015}.
+Thus, using prosody alone to accurately predict turn boundaries in conversation appears to be difficult for adults and children.
+Prosodic information is continuous, multidimensional, and indexically complex---it encodes syntactic structure, speech act, and extralinguistic information, sometimes simultaneously, without clear one-to-one mappings between form and meaning \citep{cutler1997, shriberg1998, lammertink2015}.
 For these reasons, prosodic information alone may not be enough to both (a) make precise temporal predictions about turn structure, and (b) identify question turns, which otherwise appear to drive anticipatory gaze switching.
@@
 
-Our findings suggest that there is one prosodic cue that serves as an exception to this rule: inter-turn silence.
-Generally speaking, participants showed a greater anticipatory switches for longer inter-turn gaps, but the effect of inter-turn gap duration is strongest in our data when upcoming responses are less predictable, whether due to the question effect (Experiment 1) or the lack of non-verbal cues and any linguistic information (Experiment 2).
+Our findings suggest that there is one prosodic cue that is an exception to this rule: inter-turn silence.
+Generally speaking, participants showed a greater anticipatory switches for longer inter-turn gaps, but the effect of inter-turn gap duration is strongest in our data when upcoming responses are less predictable, whether due to the asymmetrical response expectations for questions vs.
+non-questions (Experiment 1) or the lack of non-verbal cues and any linguistic information (Experiment 2).
 Notably, there were no significant interactions of gap duration with participant age.
-This pattern of results suggests that, when predictive information about upcoming responses is absent, long silences may be used to retroactively assign turn-end boundaries, thereby increasing the participant's expectation for a speaker change and promoting more anticipatory gaze switches.
-The lack of interactions between age and gap duration suggests that the use of inter-turn silence remains important for older speakers under ambiguous turn structure conditions and the interactions between transition type and gap duration (Experiment 1) and condition and gap duration (Experiment 2; marginal), suggest that this effect is not straightforwardly the result of having more time to make a gaze switch.
-
-Because inter-turn silence works as a backwards-looking cue, it's predictive utility for speaker change is diminished compared to forward-looking linguistic cues (e.g., syntactic constituent completion).
+This pattern of results suggests that, when predictive information about upcoming responses is absent, long silences may increase participants' expectation for a speaker change and promote more anticipatory gaze switches.
+Pauses are detected and related to phrasal structure from early on; 5-month-old infants use pauses to parse intonational phrases \citep{mannel2009}.
+The lack of interactions between age and gap duration suggests that the use of inter-turn silence remains important for older speakers and the interactions between transition type and gap duration (Experiment 1) and condition and gap duration (Experiment 2; marginal), suggest that this effect is simply the result of having more time to make a gaze switch.
+These findings thus suggest that silence is an early and lasting cue for identifying turn structure online when other predictive information is not adequate.
+
+Notably, many other non-linguistic cues encode information about transition type, including gaze and gesture.
+We did not systematically test those cues here but, like inter-turn silence, they may play a critical role in parsing and making predictions about turn structure when other linguistic information is not sufficient to make accurate predictions.
+
+%Because inter-turn silence works as a backwards-looking cue, it's predictive utility for speaker change is diminished compared to forward-looking linguistic cues (e.g., syntactic constituent completion).
 However, it is a strategy that can be usefully employed to predict upcoming responses some of the time, even without access to language.
-Pauses are detected and related to phrasal structure from early on; 5-month-old infants use pauses to parse intonational phrases \citep{mannel2009}.
-Our findings thus suggest that silence is an early and lasting cue for identifying turn structure online when other predictive information is not adequate.
 
@@
 Unlike prosodic cues, lexicosyntactic cues are discreet and have much clearer form-to-meaning mappings, with clear lexicosyntactic cues to questionhood that occur early within turns (e.g., \textit{wh}-words, \textit{do}-insertion, and subject-auxiliary inversion).
-That said, although children's lexical and syntactic knowledge begins to develop early, it is limited for quite some time \citep{bergelson2013, shi2010, tomasello1999}.
+That said, children's lexical and syntactic knowledge is limited for quite some time (\citealp{tomasello1999}, but see also \citealp{bergelson2013, shi2010}).
 Our stimuli were made in a child-friendly style, they are still other-directed and fairly complex, with 20--30 seconds of continuous conversational speech.
@@
 It is perhaps for this reason that children's performance was always best with the full signal, where lexicosyntactic information was supported by prosodic information and vice versa.
-There may be something extra informative about the combination of prosodic and lexical cues to questionhood that helps to boost children's anticipations before they can use these cues separately.
 Even in adults, Torreira and colleagues (\citeyear{torreira2015}) showed that the trade-off in informativity between lexical and prosodic cues is more subtle in semi-natural speech.
@@
 
-Finally, because most anticipatory switches were made following question turns, one could predict that the only lexical and prosodic cues that matter for spontaneous turn prediction are those that cue questionhood.
+%Finally, because most anticipatory switches were made following question turns, one could predict that the only lexical and prosodic cues that matter for spontaneous turn prediction are those that cue questionhood.
 Our linguistic manipulations here were focused on lexical and prosodic information globally, which may have affected lexical and prosodic cues to questionhood asymmetrically; while lexicosyntactic question cues were available on every instance of \textit{wh}- and \textit{yes/no} questions in our stimuli, prosodic question cues were only salient on \textit{yes/no} questions.
-Relatedly, many non-verbal cues also encode information about transition type, including gaze and gesture.
-We did not systematically test those cues here but, like inter-turn silence, they may play a critical role in parsing and making predictions about turn structure when other linguistic information is not sufficient to make accurate predictions.
 
@@
 Moreover, because of their high frequency and relatively limited number of formats, questions, especially \textit{wh}-questions, may be more identifiable and predictable compared to other types of speech acts.
-So, in addition to having a special interactive status (for adults and children alike), questions are a frequent, predictable and core characteristic of many caregiver-child interactions, motivating a general benefit for questions in turn structure anticipation.
+So, in addition to having a special interactive status, questions are a frequent, predictable and core characteristic of many caregiver-child interactions, motivating a general benefit for questions in turn structure anticipation.
 
@@
 and (2) what is it about questions that makes children and adults more likely to anticipatorily switch their gaze to addressees?
-If this effect is particular to turns that require an immediate response (``adjacency pairs''; \citealp{schegloff2007}), other turn types, such as imperatives, compliments, and complaints should show similar patterns to questions.
-If the effect is instead about overall predictability of the syntactic frame, children would instead show similar patterns for other frequent frames in child-directed speech (e.g., ``Look at the X.''; \citealp{mintz2003}).
-Recognizability and predictability are likely to play a role as children become more sophisticated language users, even if the effect is truly about adjacency pairs; for example, rhetorical and tag questions take a very similar form to prototypical polar questions, but usually do not require an answer.
-So, though it is clear that adults and children anticipated responses more often for questions than non-questions, we do not yet know whether their predictive action is limited to turns formatted as questions, turns with high recognizability and predictability, or turns that project an immediate response from the addressee.
-
-The question effect suggests that participants' spontaneous predictions, at least while viewing third-party conversation, may be driven by what lies \textit{beyond} the end of the current turn---not just by the upcoming end of the turn itself, as has been focused on in prior work \citep{torreira2015, keitel2013, magyari2012, de-ruiter2006}.
+If this ``question'' effect exists for all turns that require an immediate response (``adjacency pairs''; \citealp{schegloff2007}), other turn types, such as imperatives, compliments, and complaints should show similar patterns.
+If the effect is instead about overall predictability of the syntactic frame, children would instead show similar patterns for other frequent frames from child-directed speech (e.g., ``Look at the X''; \citealp{mintz2003}).
+The recognizability and predictability of syntactic frames is likely to play a role in turn prediction as children become more sophisticated language users, even if the effect is truly about adjacency pairs; for example, rhetorical and tag questions take a very similar form to prototypical polar questions, but usually do not require an answer.
+So, though it is clear that adults and children anticipate responses more often for questions than non-questions, we do not yet know whether their predictive action is limited to turns formatted as questions, turns with high recognizability and predictability, or turns that project an immediate response from the addressee.
+
+The question effect suggests that participants' spontaneous predictions may be driven by what lies \textit{beyond} the end of the current turn---not just by the upcoming end of the turn itself, as has been focused on in prior work \citep{torreira2015, keitel2013, magyari2012, de-ruiter2006}.
 In future work, it will be crucial to measure prediction from a first-person perspective to find out what kinds of predictions are most relevant to addressees in conversation (see also \citealp{holler2015}).
 
-One possible scenario is that listeners in spontaneous, first-person conversation use multiple strategies to make predictions about upcoming turn structure: they could use semi-passively attend to incoming speech for cues to upcoming speaker transition (e.g., questions and other adjacency pairs) and, when possible upcoming transition is detected, switch into a more precise turn-end prediction mode (\`{a} la \citealp{de-ruiter2006}).
+One possible scenario is that listeners in spontaneous, first-person conversation use multiple strategies to make predictions about upcoming turn structure: they could semi-passively attend to incoming speech for cues to upcoming speaker transition (e.g., questions and other adjacency pairs) and, when possible upcoming transition is detected, switch into a more precise turn-end prediction mode (\`{a} la \citealp{de-ruiter2006}).
 A flexible prediction system like this one allows listeners to continuously monitor ongoing conversation for turn-related cues at a low cost while still managing to plan their responses and come in quickly when needed.
@@
 Therefore, this difference in age of emergence more likely stems from our use of a more engaging speech style, stereo speech playback, and more typical turn transition durations.
-The child-friendly style of speech in particular may have helped in two ways: keeping children more engaged with the stimuli in general and using less syntactically complex and more prosodically exaggerated cues \citep{fernald1989, werker1989, snow1977} compared to what they would get with adult-adult speech.
+The child-friendly style of speech in particular may have helped in two ways: keeping children more engaged with the stimuli and using less syntactically complex and more prosodically exaggerated speech \citep{fernald1989, werker1989, snow1977} compared to what they would get with adult-adult conversation.
 
@@
 
-The prediction measure used in our studies is based on an observer's view of third-party conversation but, because participants' role in the interaction could affect their online predictions about turn taking, an ideal measure would instead capture first-person predictions.
+The prediction measure used in our studies is based on an observer's view of conversation but, because participants' role in the interaction could affect their online predictions about turn taking, an ideal measure would instead capture first-person predictions.
 If conversational participants' predictions are partly shaped by their need to respond, first-person measures of spontaneous turn prediction will be key to revealing how participants distribute their attention over verbal and non-verbal cues while taking part in everyday interaction, the implications of which relate to theories of online language processing for both language learning and everyday talk.
@@
 The task is natural and intuitive in that no instruction is required, which means that it captures spontaneous predictive behavior and can be used with participants of all ages.
-Frequencies of anticipatory gaze switching appear to be stable across language communities where similar tasks have been tested \citep{keitel2013, keitel2015, holler2015, hirvenkari2013}---even from a first-person perspective---so the task is one that measures robust predictive behavior relevant to conversational processing in many languages.
-It also lends itself to many possibilities for controlling the presence of individual verbal and non-verbal cues and has a clear method for assessing random switching baselines across the entire stimulus (because the participant's task is constant).
-Also, in the case that response preparation interferes with our ability to see prediction at the ends of incoming turns, third-person paradigms are one of the only ways to ensure that we are measuring prediction processes in isolation.
-
-The current findings also make important predictions about what we will see in first-person prediction paradigms.
-For example, the focus on upcoming points of speaker transition is only more important when participants are embedded in a live interaction; we would thus expect question-like effects to occur in first-person paradigms too, and perhaps even be amplified.
+Frequencies of anticipatory gaze switching appear to be stable across language communities where similar tasks have been tested \citep{keitel2013, keitel2015, holler2015, hirvenkari2013}---even from a first-person perspective---so the task is one that measures robust predictive behavior relevant to conversational processing across languages.
+It also lends itself to many possibilities for controlling the presence of individual verbal and non-verbal cues and has a clear method for assessing random switching baselines across the entire stimulus.
+Also, if it is the case that response preparation interferes with our ability to see prediction at the ends of incoming turns, third-person paradigms are one of the only ways to measure prediction processes in isolation.
+
+The current findings also make predictions about what we would see in first-person paradigms.
+For example, a focus on possible upcoming speaker transitions is even more important when the participants themselves may need to respond; we would thus expect question-like effects to occur in first-person paradigms, and perhaps even be amplified compared to third-person paradigms.
 If so, participants' use of linguistic information would still subserve this goal, with prediction at a premium.
-So we would predict that linguistic cues to upcoming speaker change would be most critical in a first-person measure.
-Regarding development, the same facts about the complexity of prediction with prosodic information and children's initial limited lexical inventories would still hold, as would the use of silence and non-verbal cues to assess and predict turn structure in the absence of clear predictive linguistic information.
-We therefore think that the paradigm presented here has important contributions to make in finding out how we attend to and make predictions about conversational interaction.
+Regarding development, the same facts about the complexity of prosody-based prediction and children's initial limited lexical inventories would still hold, as would the use of silence and non-verbal cues to assess and predict turn structure in the absence of clear predictive linguistic information.
+The paradigm presented here thus has important contributions to make in our understanding of how participants attend to and make predictions about conversational interaction.
 
@@
 However, turn prediction is not fully integrated with linguistic cues until much later and, in the absence of predictive linguistic cues, children and adults alike rely on retroactive cues such as inter-turn silence to predict upcoming speaker change.
-Using language to make predictions about upcoming interactive content takes time to develop and, for participants of all ages appears to be primarily driven by participants' orientation to what will happen next---beyond the end of the current turn.
+Using language to make predictions about upcoming interactive content takes time to develop and, for participants of all ages appears to be primarily driven by participants' expectations about what will happen next, beyond the end of the current turn.
 
@@
 We estimated the baseline probability of making an anticipatory switch by randomly permuting the placement of the transition windows within each stimulus (Figure \ref{fig:shuffling}).
-We then used the switch identification procedure from Experiments 1 and 2 (Section \ref{sec:algorithm}) to find out how often participants made ``anticipatory'' switches within these randomly permuted windows.
+We then used the switch identification procedure from Experiments 1 and 2 to find out how often participants made ``anticipatory'' switches within these randomly permuted windows.
 This procedure de-links participants' gaze data from turn structure by randomly re-assigning the onset time of each turn-transition in each permutation.
@@
 
-For example, children's ``language condition'' effect from Experiment 1 had a \textit{z}-value of $|$3.429$|$, which is greater than 99.9\% of all $|$\textit{z}-value$|$ estimates from Experiment 1's random permutation models (i.e., \textit{p}$=$.001).
-It is therefore highly unlikely that the effect of language condition in the original model derived from random gaze shifting.
+For example, children's ``language condition'' effect from Experiment 1 had a \textit{z}-value of $|$3.65$|$, which is greater than 99.3\% of all $|$\textit{z}-value$|$ estimates from Experiment 1's random permutation models (i.e., \textit{p}$=$.007).
+It is therefore highly unlikely that the effect of language condition in the original model came from random gaze shifting.
 
 We used this procedure to derive the random-baseline comparison values in the main text (above).
-However, we ran into two issues along the way: first, we had to report \textit{z}-values rather than beta estimates.
+However, we ran into two issues along the way: first, we had to report \textit{z}-values rather than beta estimates of each effect.
 Second, we had to exclude a substantial portion of the models, especially in Experiment 2 because of model non-convergence.
@@
 Although many of the non-converging models show estimates within range of the converging models (e.g., with a mean difference of only 0.09 in median \textit{z}-value across predictors), they also show many radically outlying estimates (e.g., showing a mean difference of 237.3 in mean \textit{z}-value across predictors).
-Similar patterns were obtained in the non-converging models for Experiment 2 and persisted even when we tried other optimizers.
+Similar patterns were obtained in the non-converging models for Experiment 2 and persisted across multiple attempts with different optimizers.
 
@@
 \setcounter{figure}{0}
-Experiments 1 and 2 both showed effects of age in interaction with linguistic condition and transition type (e.g., English vs.
-non-English).
-To explore these effects in more depth, we recorded the average difference score for the predictor that interacted with age for each participant (e.g., English minus non-English anticipatory switches), using these values to compute an average difference score over participants in each age group (e.g., age 3, 4, and 5) within each random permutation.
-That averaging process produces 5,000 baseline-derived difference scores for each age group.
-
-We then made pairwise age comparisons of these difference scores (e.g., the linguistic condition effect in 3-year-olds vs.
+Experiments 1 and 2 both showed effects of age in interaction with linguistic condition and transition type.
+To explore these effects in more depth, in each permutation we recorded the average difference score for each participant, for each predictor that interacted with age (e.g., English minus non-English anticipatory switches for each participant).
+We then used these values to compute an average difference score over the participants in each age group (e.g., age 3, 4, and 5) within each random permutation.
+This averaging process produces 5,000 baseline-derived difference scores for each age group.
+
+We then made pairwise age comparisons of the difference scores (e.g., the linguistic condition effect in 3-year-olds vs.
 4-year-olds), computing the percent of random-permutation difference scores exceeded by the real-data difference score.
-If the real-data difference score exceeded 95\% of the random-data age difference scores, we deemed it to be an age effect significantly different from chance---e.g., a significant difference between ages three and four in the effect of linguistic condition.
+If the real-data difference score exceeded 95\% of the random-data age difference scores, we deemed it to be an age effect significantly different from chance, e.g., a significant difference between ages three and four in the effect of linguistic condition.
 This procedure is essentially a two-tailed \textit{t}-test, adapted for use with the randomly permuted baseline data.
@@
 In each of the plots below, the black dot represents the real data value for the effect being shown.
-The effect sizes from the 5,000 randomly permuted data sets are shown in the distribution.
-The percentage displayed is the percentage of random permutation values exceeded by the original data value (taking the absolute value of all data points for a two-tailed test).
+The effect sizes from the 5,000 randomly permuted data sets are shown as a distribution.
+The percentage displayed is the percentage of random permutation difference scores exceeded by the original data differences score (taking the absolute value of all data points for a two-tailed test).
 Comparisons marked with 95\% or higher are significant at the \textit{p}$<$0.05 level.
@@
 \end{center}
-\caption{Significant pairwise comparisons of the \textit{prosody only}-\textit{no speech} linguisitic condition effect, across ages in Experiment 2} 
+\caption{Significant pairwise comparisons of the \textit{prosody only}-\textit{no speech} linguisitic condition effect, across ages in Experiment 2. Non-significant comparisons are not shown here.} 
 \label{fig:E2-lgageinteraction}
@@
 \end{center}
-\caption{Significant pairwise comparisons of the \textit{normal speech}-\textit{no speech} language condition effect for transition type, across ages, in Experiment 2.} 
+\caption{Significant pairwise comparisons of the \textit{normal speech}-\textit{no speech} language condition effect for transition type, across ages, in Experiment 2.
+Non-significant comparisons are not shown here.}
 \label{fig:E2-lgagetypeinteraction}
@@
 
-One alternative hypothesis for children's anticipatory gazes is that they are looking at the current speaker at the start of each turn, but then grow bored and start looking away at a constant rate.
+One alternative hypothesis for children's anticipatory gazes is that they look at the current speaker at the start of each turn, but then grow bored and start looking away at a constant rate.
 Even though this alternative hypothesis does not predict the primary effects in our data (e.g., the difference between questions and non-questions), we cannot rule out the possibility that a portion of participants' saccades come from boredom.
 
-The data plotted here show a hypothetical group of boredom-driven participants (gray dots) compared to participants from the actual data in Experiment 2 (black dots).
+The data plotted here show a hypothetical group of boredom-driven participants (gray dots) and participants from the actual data in Experiment 2 (black dots).
 The hypothetical boredom-driven participants look away from the current speaker at a linear rate, beginning one second after the start of a turn.
@@
 Therefore, their crossover point---the point in time at which 50\% of the children have switched away from the current speaker---would be the same for all turns, no matter the length of the turn.
-This pattern is demonstrated in the hypothetical boredom-driven crossover points, which always occur 2.5 seconds after the start of speech (gray vertical line; Figure \ref{fig:boredomhypothesis}).
-
-In children's \textit{actual} looking data we see that crossover points increase with turn duration: 2.0, 2.9, and 3.6 seconds after the start of speech for turns with durations of 1--2, 2--3, and 3--4 seconds, respectively (black vertical line; Figure \ref{fig:boredomhypothesis}).
+This pattern is demonstrated in the hypothetical boredom-driven crossover points, which always occur 2.5 seconds after the start of speech (gray vertical lines; Figure \ref{fig:boredomhypothesis}).
+
+In children's \textit{actual} looking data we see that crossover points increase with turn duration: 2.0, 2.9, and 3.6 seconds after the start of speech for turns with durations of 1--2, 2--3, and 3--4 seconds, respectively (black vertical lines; Figure \ref{fig:boredomhypothesis}).
 This pattern suggests that, though children do look away as the turn is unfolding, their looks away are not simply driven by boredom.
@@
 Are the looks away in Figure \ref{fig:boredomhypothesis} still too early to count as ``turn-transition'' anticipation?
-It is true that children start looking away after one second has passed (but then only gradually).
-It is possible that some of these early looks away are boredom-driven, but it is equally plausible that some of them are turn-driven.
-Early predictive behavior is common in turn-taking studies with adults, in both constrained turn-taking tasks (e.g., \citealp{de-ruiter2006, gisladottir2015, bogels2015}) and in spontaneous conversation (e.g., \citealp{holler2015, torreira2015}).
+It is true that children start looking away after one second has passed, but then only gradually.
+Some of these early looks away may be boredom-driven, but it is equally plausible that some of them are turn-driven.
+Early predictive behavior is common in turn-taking studies with adults, in both constrained turn-taking tasks \citep{de-ruiter2006, gisladottir2015, bogels2015} and in spontaneous conversation \citep{holler2015, torreira2015}.
 Although this same pattern has yet to be established for children's turn predictions, the looking behavior here is at least consistent with adult response patterns in previous work.
-Additionally, because our analysis windows in the main study only overlapped with the pre-gap utterance by 300 msec (Figure \ref{fig:criterion}), our primary results are unlikely to capture any very early or early boredom-driven gaze switches.
-
-We conclude that the boredom-driven effects in our data are unlikely to change our primary results, though we acknowledge that characterizing different gaze switching strategies in this kind of data is an important avenue for future work.
+Additionally, because our analysis windows in the main study only overlapped with the pre-gap utterance by 300 msec (Figure \ref{fig:criterion}), our primary results are unlikely to capture any of these very early or early boredom-driven gaze switches, which makes them unproblematic either way in the current analysis.
+
+We therefore conclude that the boredom-driven effects in our data are unlikely to change our primary results, though we acknowledge that characterizing different gaze switching strategies in this kind of data is an important avenue for future work.
 
@@
 Each video featured a puppet pair (red/blue/yellow/robot/ merpeople/party-goer; Figure \ref{fig:puppets}).
-Puppets in all six videos performed the exact same conversation recording (`birthday party'; Experiment 2) with normal, unmanipulated speech; this experiment therefore holds all things constant across stimuli except for the appearance of the puppets.
+Puppets in all six videos performed the exact same conversation recording (`birthday party'; Experiment 2) with normal, unmanipulated speech.
+This experiment therefore holds all things constant across stimuli except for the appearance of the puppets.
 
@@
     \hline
-     (Intercept) 								& -0.14790	& 0.32796	& -0.451		& 0.652 \\ 
-     Puppets$=$\textit{mermaid} 	& -0.07581	& 0.65532	& -0.116		& 0.908 \\ 
-     Puppets$=$\textit{robot} 		& -0.07104	& 0.65321	& -0.109		& 0.913 \\ 
-     Puppets$=$\textit{party} 		& -0.78206	& 0.68699	& -1.138		& 0.255 \\ 
+     (Intercept) 							& -0.148	& 0.328		& -0.451		& 0.652 \\ 
+     Puppets$=$\textit{mermaid} 	& -0.076	& 0.655		& -0.116		& 0.908 \\ 
+     Puppets$=$\textit{robot} 		& -0.071	& 0.653		& -0.109		& 0.913 \\ 
+     Puppets$=$\textit{party} 		& -0.782	& 0.687		& -1.138		& 0.255 \\ 
     \hline
@@
     \hline
-     (Intercept) 								& -0.22371	& 0.56832	& -0.394		& 0.694 \\ 
-     Puppets$=$\textit{robot} 		& 0.004763	& 0.80096	& 0.006		& 0.995 \\ 
-     Puppets$=$\textit{party} 		& -0.70626	& 0.82742	& -0.854		& 0.393 \\ 
+     (Intercept) 							& -0.224	& 0.568		& -0.394		& 0.694 \\ 
+     Puppets$=$\textit{robot} 		& 0.0048	& 0.801		& 0.006			& 0.995 \\ 
+     Puppets$=$\textit{party} 		& -0.706	& 0.827		& -0.854		& 0.393 \\ 
     \hline
@@
     \hline
-     (Intercept) 								& -0.21895	& 0.56565	& -0.387		& 0.699 \\ 
-     Puppets$=$\textit{party} 		& -0.71102	& 0.82657	& -0.860		& 0.390 \\ 
+     (Intercept) 							& -0.219	& 0.566		& -0.387		& 0.699 \\ 
+     Puppets$=$\textit{party} 		& -0.711	& 0.827		& -0.860		& 0.390 \\ 
     \hline
@@
     \hline
-     (Intercept) 								& -0.9300	& 0.6067		& -1.533		& 0.125 \\ 
+     (Intercept) 							& -0.93		& 0.607		& -1.533		& 0.125 \\ 
   \end{tabular}
@@
 
-We take this finding as evidence that our decision to not fully cross puppet pairs and linguistic conditions in Experiment 2 was unlikely to have strongly affected children's anticipatory gaze rates above and beyond the intended effects of linguistic condition.
+We take this finding as evidence that our decision to not fully cross puppet pairs and linguistic conditions in Experiment 2 was unlikely to have affected children's anticipatory gaze rates above and beyond the intended effects of linguistic condition.
 

### 2016-06-06 paper/Letters/ResponseLetter-Jun2016.tex
draft of response letter; still missing page refs to manuscript

@@
 
-\noindent I hereby resubmit a revised version of my manuscript ``The development of children's ability to track and predict turn structure in conversation'' to the \textit{Journal of Memory and Language}.
+\noindent I hereby submit a revised version of the manuscript ``The development of children's ability to track and predict turn structure in conversation'' to the \textit{Journal of Memory and Language}.
 
-%\noindent I apologize for our delay in getting the manuscript resubmitted.
-It took quite a long time to implement one of the suggestions from Reviewer 3, mostly because I had to acquire some new technical skills in order to carry out his/her suggested analysis on our computation-heavy data set.
-Nevertheless, I believe that the added analysis has substantially strengthened the paper as a whole, so I appreciate your patience very much.
+\noindent Your and the reviewers' comments have helped us make substantial improvements in the quality and clarity of our developmental story and the theoretical discussion regarding turn prediction in third- and first-person measures.
+Since our last submission, we have updated all of the statistical models---including the 20k models of the random baseline data---to systematically test for effects of the inter-turn gap.
+We have integrated these new gap-related results into our discussion, which is now focused on the differing roles of (predictive) linguistic cues and inter-turn silence, a (reactive) cue in itself.
 
-%\noindent Since our last submission, we have done quite a lot of work to improve our paper by addressing the reviewers' comments.
-For example, in addition to the new random permutation analysis suggested by Reviewer 3 (mentioned above), we have also included two further follow-up analyses in the Appendix (including additional experimental data), each inspired by the other reviewers' comments.
-We have also tried to make the text much more succinct, with separate and straightforward sections for the results and discussions, following requests from all three reviewers.
+\noindent We have edited the text in response to each of the specific questions and comments from the most recent reviews (see below).
+In particular, we added more detail and theoretical heft to our discussion of inter-turn gap effects, the question effect and the implications of measuring of turn prediction in a third-person paradigm, making specific predictions about what we might find in a first-person measure.
+In response to some of the reviewers' comments, we've also worked to further clarify our descriptions of the random baseline analyses, the extent of the linguistic effects, and our test of boredom-driven saccades in the Appendix.
+Despite these additions, we have tried to keep the text as short as possible.
 
-%\noindent We hope that, with these changes, we have greatly improved the paper's quality.
+\noindent We hope that you will agree that these changes have greatly improved the paper's quality.
 Please find a description of the specific comments addressed below.
-Thank you very much for your reconsideration.
+Thank you very much for your reconsideration!
 Please let us know if there is any further information that you require about the submission.
@@
 
-%\noindent We thank all three reviewers for their thorough and helpful comments on the previously submitted manuscript.
-We have tried to address all of their comments, both in the manuscript and here in the response letter.
+\noindent We thank the editor and our two reviewers for their very helpful comments on the most recently submitted manuscript. We have tried to address their comments here and in the manuscript.
 
@@
 
-\noindent \textbf{Reviewer 1}
+\noindent \textit{1. Are the permutation analyses anti-conservative?}
 
-\noindent \textbf{Reviewer 2}
+\noindent R2 found our new description of the permutation analysis satisfactory, but R1 still expressed some doubts that we believe may have arisen from a miscommunication.
+We therefore focused the description of the permutation analyses more on the purpose of the analyses (PXX).
+R1 expresses concern that the permutation analyses are anti-conservative because we assume that ``baseline saccades are equiprobable across the turn''.
+While we appreciate R1's point, this characterization is not quite correct and we believe that our permutation analyses are doing exactly what they should: test the likelihood that we could get significant effects (of any predictor) if the saccades turned out to be randomly distributed---a true random baseline.
+Crucially, our permutations assume that baseline saccades are equiprobable across the entire stimulus, and not just across turns.
+R1 may imagine that we are comparing (a) anticipation rates at transition windows against (b) rates during turns, thereby maximizing our chances of finding a difference from the baseline.
+This is not what we are doing and is specifically what we criticize about the Keitel et al.
+(2013, 2015) studies.
+Instead, we compare (a) anticipation rates at transition windows to (b) transition rates over the whole stimulus.
+In other words, our baseline is agnostic to the placement of the real transition windows.
+
+\noindent R1 might argue in response that our baseline estimate is still not fair because, presumably, most saccades are at transition windows and so, by comparing them to saccades over the whole stimulus, we are merely comparing them to at a watered down saccadic rate.
+We have two responses to this: First, the primary aim of the permutation analysis is actually to test R1's assumption (i.e., anticipation at transition windows is greater than elsewhere), but has the substantial added benefit of testing whether the assumption holds for each predictor effect in the lmers.
+Second, the non-``transition window'' parts of the stimulus contain many behaviors that likely elicit anticipatory gaze behavior.
+For example, the transition windows in our real-turn analyses don't include: very early and very delayed anticipatory looks, looks to a responder in error (before the participant realizes the same speaker has actually continued speaking), and turn transitions that weren't included in our final analyses (e.g., because they had overlapped speech or very long inter-turn gaps).
+All three of these cases are common in our data and elicit similar looking behavior to the target ``transition window'' anticipatory gazes.
+But, because they are not part of our final ``real transition window'' dataset, they only count against us in the analyses---they \textit{only} appear in the baseline estimates, and never in the ``real'' transition estimates.
+
+\noindent We therefore conclude that our permutation analyses are (a) exactly fit to the question we pose, (b) present a more useful and conservative baseline than used in prior literature, and (c) are conservative in the sense that there are other turn-relevant behaviors hidden in the non-``transition window'' parts of the stimulus.
+By editing the motivation behind the analyses on PXX, we hope we have made this argument clearer.
+R1 may still be dissatisfied with the permutation analyses in that they do not capture anything more than ``random-looker'' vs.
+``turn-driven looker'' (in his/her words, that they are ``superfluous''), but this is exactly what they are intended to do.
+We do note, however, that we also test a more subtle hypothesis about anticipatory looking rates (i.e., ``boredom-driven looker'' vs.
+``turn-driven looker'') in a separate set of analyses (Boredom Hypothesis; Appendix C, also discussed below).
+
+\noindent \textit{2. Are some of the saccades still driven by boredom?}
+
+\noindent R1 notes that, although we show significant differences between our real lookers and hypothetical ``boredom-driven'' lookers in Appendix C, our additional analyses do not rule out the fact that boredom might play a role.
+R1 points to the Figure C.1 to highlight that children still sometimes look away from the speaker long before the turn ends.
+We agree that our analyses can't rule out the fact that some proportion of the saccades might be due to boredom (or to chance).
+Surely some of them are.
+That said, we do not think that this is a major issue for our conclusions.
+First, the primary effects in our data (which revolve around questions and linguistic access) are not predicted by a boredom account---a bored looker should be constant in their boredom for questions and for non-questions.
+Second, our data show dramatic difference in the timing of look-away rates across turns of different length, which is also not predicted by a boredom-based account: boredom-driven look-aways would occur at a constant rate, independent of the ongoing turn's ultimate length.
+Finally, early (within-turn) predictive behavior is a normal finding for turn-taking studies.
+The bulk of previous work with adults, in both constrained turn-taking tasks (e.g., de Ruiter et al., 2006, G\'{i}slad\'{o}ttir et al., 2015; B\"{o}gels et al., 2015) and spontaneous conversation (e.g., Holler \& Kendrick, 2015; Torreira et al., 2015) have found evidence for early speech act detection, early response planning, and early anticipatory looking, all occurring well before the end of the ongoing turn.
+We therefore believe that whatever boredom-driven effects exist in our data are likely small and do not interfere with the interpretation of our primary variables of interest.
+We have added discussion on this point to the Appendix section dedicated to the boredom hypothesis (PXX).
+We also added a footnote earlier in the manuscript, pointing readers to this extra analysis in case they are interested (PXX).
+
+\noindent \textit{3. Are the permutation-based t-tests of age effects trustworthy?}
+
+\noindent R1 is not satisfied with the age-based \textit{t}-tests in Section 3.3.3 (Expt 2) and would rather see individual comparisons between the Normal speech and No speech (``party'') conditions for each age.
+We have added R2's requested \textit{t}-tests to our analyses and have integrated them into the results for Expt 2.
+
+\noindent \textit{4. Shouldn't the effect of gap duration be part of the developmental story?}
+
+\noindent R1 points out that, we should treat gap duration as more than just a control variable; that its effect is both strong and theoretically interesting.
+Following R1's suggestion, we have added discussion about the role of inter-turn gaps for retroactive turn boundary assignment throughout the manuscript, from the Introduction to the General Discussion.
+We have worked to make it clear why gaps themselves can be a good cue to turn transition.
+We have also taken up R1's suggestion for reframing in the General Discussion; we now discuss the developmental trajectory with respect to both the condition-based linguistic effects and inter-turn silence, which more naturally integrates the effect of gap duration into our story (PXX).
+By doing this, we move ``away from the idea that linguistic prediction is the only explanatory factor'' as R1 suggests.
+R1 also noticed that we didn't include an age-by-gapduration or condition-by-gapduration interaction across all of our lmers.
+This is because of we were unable to add full interactions between gap duration and the other fixed effects in all models; in many models, higher-order interactions with gap duration create non-convergence issues due to data sparsity with some combinations of three or more predictor levels.
+
+\noindent We came up with a balanced response to R1's concerns: to test the effect of gap duration but also remain consistent across all of our models, we added a fixed effect of gap duration and two-way interactions between gap duration and all other simple fixed effects (i.e., gap duration and age, linguistic condition, and transition type) to every model.
+This way all of our models test both the overall effect of gap duration (as before) and any interaction between gap duration and age, linguistic condition, or transition type.
+This tests R1's hypothesis that the gap duration effect might change with age (which, incidentally, we do not see in our data) as well as other potential effects of gap duration.
+
+\noindent \textit{5. Is there strong evidence that children are using linguistic information?}
+
+\noindent R1 mentions that the linguistic effects in our study are relatively small, tucked within interactions.
+We take R1's point, but also expected from the start that linguistic effects would strongly interact with age (the reason we used a broad age range) and that anticipatory gazes would interact with speech act (the reason we included question coding).
+We did not, however, anticipate that the effect of speech act would dominate our results.
+The strong effects of speech act effectively funnels all the linguistic effects (and their age interactions) into the ``question'' subset of the data.
+Because the speech act effect is both new and theoretically motivated, and because there are still effects of age and linguistic condition associated with it, we believe that these findings are groundbreaking in what they tell us about the conditions under which children and adults engage in spontaneous linguistic prediction (i.e., linguistic effects) about turn taking in conversation.
+
+\noindent To better make this argument in the manuscript, we have added discussion on the origin of the question effect and why it is important (also addressing a request from R2 below).
+R1's comment may have partly arisen from the way we discussed the linguistic effects.
+We reviewed the text and found that, indeed, we occasionally overgeneralized linguistic effects to ``children'' (as a whole) when they really should be limited to ``older children'' or ``adults''.
+We have gone back through the text and made sure that, when discussing linguistic effects, it is clear which subset of participants we are referring to.
+
+\noindent R2 points out some conflicting statements about the use of lexical cues in Experiment 2.
+Thanks!
+These are now addressed in the text (PXX).
+R2 also urges us to consider our findings with respect to other cues that might drive anticipation (e.g., non-verbal cues).
+We have added a short point to the discussion that makes clear that the linguistic effects we study here are only a few of the possible sources for anticipatory information, and that the systematic study of other cues, especially non-verbal ones is still needed (PXX).
+
+\noindent \textit{6. Is the developmental story too focused on the condition-based linguistic cues?}
+
+\noindent R1 argues that it is an oversimplification for us to argue that ``turn prediction develops in early childhood'' and that a better story would highlight two components: (a) reactive cues (e.g., inter-turn gap duration) that children may use early on while they (b) gradually learn the association between certain linguistic structures/speech acts and speaker transitions (for which they must use linguistic cues).
+We thank R1 for this suggestion.
+As mentioned above, we have integrated discussions of gap duration as a cue to turn taking throughout the text, outlining this more nuanced developmental trajectory in multiple places (e.g., PXX).
+
+\noindent \textit{7. What is the source of the question effect?}
+
+\noindent R2 asks us to speculate about the causal explanation for the question effect, particularly regarding whether it is really about speech acts or about other effects that correlate with questionhood (e.g., predictability).
+He or she also raises the possibility that the question effect might have developmental roots, e.g., from the high frequency of questions in children's input.
+We thank R2 for the comment and have added some discussion concerning these points to PXX.
+
+\noindent \textit{8. Should we be concerned about these low rates of spontaneous anticipatory looks?}
+
+\noindent R2 suggests that the low rates of anticipatory gaze switching in our data may cast doubt on the idea that adults and children routinely anticipate turn ends, as is assumed in much theoretical and experimental work about the psycholinguistics of turn taking.
+If so, R2 would like us to comment on how, then, speakers still successfully participate in conversation.
+We have added a brief note on our overall anticipation rates to the text of the General Discussion (PXX).
+Because this is a somewhat niche question that most readers may not be interested in, we only respond more elaborately here:
+
+\noindent In our study, ``spontaneous prediction'' is a simplified indicator that participants successfully anticipated a transition from the prior speaker to the upcoming speaker with their eye movements.
+The absence of an anticipatory gaze, does not signal that no anticipation was made.
+Behavioral evidence from experiments with adult listeners (e.g., B\"{o}gels et al., 2015; Sjerps \& Meyer, 2015) also suggest that participants can flexibly calibrate the onset of their response planning, such that they may pay more or less attention to the turn-final content of incoming turns.
+There is no reason to think that children would, unlike adults, always use the same predictive strategies.
+Therefore, in some sense, we do see our rates of anticipation as challenging for the idea that participants always or even usually make precise prediction about upcoming turn ends, but our findings are not the first to be consistent with this idea.
+But what does this mean for first-person prediction?
+
+\noindent Other studies using similar methods in third- and first-person measures find very similar anticipation rates, so our findings are at least not anomalous (e.g., Keitel et al., 2013; 2015; Holler \& Kendrick, 2015; Hirvenkari, 2013).
+Our data also primarily tap into an effect that should generalize to a first-party situation: the question effect.
+The question effect might be amplified when participants themselves need to respond.
+We briefly address these points in the text (PXX).
+
+\noindent In sum, we think there is still good reason to believe that participants routinely anticipate upcoming turn structure (e.g., the ends of ongoing turns and the likelihood of upcoming turn transition), but our study runs counter to the idea that they do this spontaneously at all possible points of turn transition.
+Our findings are in-line with an account that emphasizes flexibility in how and when participants make predictions about turn taking.
+
+\noindent \textit{9. Do the multilingual children behave differently?}
+
+\noindent R2 suggests that the subgroup of participants who have exposure to a second or even third language might behave differently because of their more varied experiences with language and turn taking (particularly with respect to turn-relevant cues such as prosody).
+We agree that this is a fascinating question, but we will not analyze this subgroup separately for a few reasons: first, as the reviewer guesses, we would not have enough power.
+Second, and perhaps more importantly, we did not conduct a full language history of each participant and so we do not know whether the second language comes from parent(s), grandparents, a nanny, friends, school, etc., and therefore can't confidently group children into a uniform ``multilingual'' group.
+We have added a footnote on this point (PXX).
+
+\noindent \textit{10. Does speech style affect children's performance?}
+
+\noindent R2 wonders whether our results imply that children's performance in our study is better than in previous, similar work (i.e., Keitel et al., 2013, 2015) because we use child-friendly speech in our stimuli and, if so, why that might be.
+He or she suggests that the mechanism could be linguistic, attentional, or both.
+We agree that it is likely both linguistic and attentional, and have added a note about this in the manuscript (PXX).
+
+\noindent \textit{11. Are the non-English ``question cues'' comparable to English question cues?}
+
+\noindent R2 points out that the status of the prosodic cues in the non-English recordings from Experiment 1 is not clear because, though they might be recognizable by English speakers as ``similar to'' English question cues, they are still phonologically and phonetically not identical to actual English question cues.
+We agree, but we also think that this doesn't bear too much on our findings.
+However, to make sure that this fact doesn't pass readers by, we now acknowledge the difference the non-English prosodic cues more emphatically (PXX).
+
+\noindent \textit{12. Can you better emphasize the strengths of this method?}
+
+\noindent R2 would like us to more explicitly mention the benefits of our method for measuring anticipation.
+We thank R2 for his or her encouragement and have added a this to the General Discussion (PXX).
+
+\noindent \textit{13. Is it important that this is a third-party predictive measure?}
+
+\noindent The editor would like us to comment on how participant role might affect our results.
+We link this request to R1's comments about the use of gap duration for predictive and reactive turn tracking, and to R2's comments about the generalization of our anticipation rates to conversation more generally.
+As discussed above, we have taken some space in the General Discussion to speculate about the possible effects of measuring predictions about turn taking from a first- or third-person perspective (e.g., PXX).
+
+\noindent \textit{14. Miscellaneous issues with the graphs}
+
+\noindent Both reviewers made small requests regarding the graphs.
+In line with their requests, we have removed the legend for condition (color) in Figures 3 \& 6, matched the labels for ``question'' and ``non-question'' in all cases (i.e., not ``Q'' and ``S'' as they sometimes still appeared).
+We also made Figure 2 clearer with examples of the earliest and latest possible anticipatory gaze switches, plus a prototypical anticipatory gaze switch.
+
+\noindent \textit{15. Other minor issues}
+
+\noindent Finally, we addressed a number of minor issues, including removing internal numbering, using APA-formatted references, adding a brief note about the non-centered age variable for each experiment, and some smaller comments on confusing wording and typos.
+
+\medskip
+
+\noindent In making these changes, we feel that we have substantially improved the clarity of our developmental story and our theoretical and methodological contributions.
+We thank the editor and reviewers again for their helpful comments and hope that they also find the manuscript to be much improved.
 

### 2016-06-21 paper/CasillasFrank-Anticip.tex
minor tweaks

@@
 We first give a basic review of turn-taking research and the state of current knowledge about adult turn prediction.
-We then discuss recent work on the development of turn-taking skills before getting into the details of the present study.
+We then discuss recent work on the development of turn-taking skills before presenting the details of the present study.
 
@@
 Children's early games (e.g., give-and-take, peek-a-boo) have built-in, predictable turn structure \citep{ratner1978, ross1987}.
-Even monkeys take turns: non-human primates such as marmosets and Campbell's monkeys vocalize contingently with each other in both natural and lab-controlled environments \citep{lemasson2011, takahashi2013}.
+Even monkeys take turns: Non-human primates such as marmosets and Campbell's monkeys vocalize contingently with each other in both natural and lab-controlled environments \citep{lemasson2011, takahashi2013}.
 In all these cases, turn taking serves as a protocol for interaction, allowing the participants to coordinate with each other through sequences of contingent action.

### 2016-06-21 paper/Letters/ResponseLetter-Jun2016.tex
response

@@
 We therefore focused the description of the permutation analyses more on the purpose of the analyses (PXX).
-R1 expresses concern that the permutation analyses are anti-conservative because we assume that ``baseline saccades are equiprobable across the turn''.
-While we appreciate R1's point, this characterization is not quite correct and we believe that our permutation analyses are doing exactly what they should: test the likelihood that we could get significant effects (of any predictor) if the saccades turned out to be randomly distributed---a true random baseline.
+In particular, R1 expressed concern that the permutation analyses are anti-conservative because we assume that ``baseline saccades are equiprobable across the turn.'' While we appreciate R1's point, this characterization is not quite correct and we believe that our permutation analyses are doing exactly what they should: namely, testing the likelihood that we could get significant effects (of any predictor) if the saccades turned out to be randomly distributed---a true random baseline.
 Crucially, our permutations assume that baseline saccades are equiprobable across the entire stimulus, and not just across turns.
@@
 Instead, we compare (a) anticipation rates at transition windows to (b) transition rates over the whole stimulus.
-In other words, our baseline is agnostic to the placement of the real transition windows.
+In other words, our baseline is agnostic to the placement of the true transition windows.
 
-\noindent R1 might argue in response that our baseline estimate is still not fair because, presumably, most saccades are at transition windows and so, by comparing them to saccades over the whole stimulus, we are merely comparing them to at a watered down saccadic rate.
-We have two responses to this: First, the primary aim of the permutation analysis is actually to test R1's assumption (i.e., anticipation at transition windows is greater than elsewhere), but has the substantial added benefit of testing whether the assumption holds for each predictor effect in the lmers.
+\noindent One possible response would be that presumably, most saccades are at transition windows and so, by comparing them to saccades over the whole stimulus, we are merely comparing them to at a watered down saccadic rate.
+We have two responses to this objection, however: First, the primary aim of the permutation analysis is actually to test R1's assumption (i.e., anticipation at transition windows is greater than elsewhere), but has the substantial added benefit of testing whether the assumption holds for each predictor effect in the mixed effect models.
 Second, the non-``transition window'' parts of the stimulus contain many behaviors that likely elicit anticipatory gaze behavior.
@@
 
-\noindent We therefore conclude that our permutation analyses are (a) exactly fit to the question we pose, (b) present a more useful and conservative baseline than used in prior literature, and (c) are conservative in the sense that there are other turn-relevant behaviors hidden in the non-``transition window'' parts of the stimulus.
+\noindent We therefore conclude that our permutation analyses are (a)  fit to the question we pose, (b) present a more useful and conservative baseline than used in prior literature, and (c) are conservative in the sense that there are other turn-relevant behaviors hidden in the non-``transition window'' parts of the stimulus.
 By editing the motivation behind the analyses on PXX, we hope we have made this argument clearer.
@@
 
-\noindent R1 is not satisfied with the age-based \textit{t}-tests in Section 3.3.3 (Expt 2) and would rather see individual comparisons between the Normal speech and No speech (``party'') conditions for each age.
-We have added R2's requested \textit{t}-tests to our analyses and have integrated them into the results for Expt 2.
+\noindent To supplement these analyses, we have added standard (non-permutation based) comparisons between the Normal speech and No speech (``party'') conditions for each age and have integrated them into the results for Expt 2.
 
@@
 \noindent R1 points out that, we should treat gap duration as more than just a control variable; that its effect is both strong and theoretically interesting.
-Following R1's suggestion, we have added discussion about the role of inter-turn gaps for retroactive turn boundary assignment throughout the manuscript, from the Introduction to the General Discussion.
+Following R1's helpful suggestion, we have added discussion about the role of inter-turn gaps for retroactive turn boundary assignment throughout the manuscript, from the Introduction to the General Discussion.
 We have worked to make it clear why gaps themselves can be a good cue to turn transition.
@@
 By doing this, we move ``away from the idea that linguistic prediction is the only explanatory factor'' as R1 suggests.
-R1 also noticed that we didn't include an age-by-gapduration or condition-by-gapduration interaction across all of our lmers.
+R1 also noticed that we didn't include an age-by-duration or condition-by-duration interaction in our models.
 This is because of we were unable to add full interactions between gap duration and the other fixed effects in all models; in many models, higher-order interactions with gap duration create non-convergence issues due to data sparsity with some combinations of three or more predictor levels.
 
-\noindent We came up with a balanced response to R1's concerns: to test the effect of gap duration but also remain consistent across all of our models, we added a fixed effect of gap duration and two-way interactions between gap duration and all other simple fixed effects (i.e., gap duration and age, linguistic condition, and transition type) to every model.
+\noindent To address this issue, in our revision, we attempt to test the effect of gap duration but also remain consistent across all of our models.
+We added a fixed effect of gap duration and two-way interactions between gap duration and all other simple fixed effects (i.e., gap duration and age, linguistic condition, and transition type) to every model.
 This way all of our models test both the overall effect of gap duration (as before) and any interaction between gap duration and age, linguistic condition, or transition type.
-This tests R1's hypothesis that the gap duration effect might change with age (which, incidentally, we do not see in our data) as well as other potential effects of gap duration.
+This tests the hypothesis that the gap duration effect might change with age (which, incidentally, we do not see in our data) as well as other potential effects of gap duration.
 
@@
 \noindent R1 mentions that the linguistic effects in our study are relatively small, tucked within interactions.
-We take R1's point, but also expected from the start that linguistic effects would strongly interact with age (the reason we used a broad age range) and that anticipatory gazes would interact with speech act (the reason we included question coding).
+We take R1's point, but also expected from the start that linguistic effects would strongly interact with age (the reason we used a broad age range) and that anticipatory gaze would interact with speech act (the reason we included question coding).
 We did not, however, anticipate that the effect of speech act would dominate our results.
 The strong effects of speech act effectively funnels all the linguistic effects (and their age interactions) into the ``question'' subset of the data.
-Because the speech act effect is both new and theoretically motivated, and because there are still effects of age and linguistic condition associated with it, we believe that these findings are groundbreaking in what they tell us about the conditions under which children and adults engage in spontaneous linguistic prediction (i.e., linguistic effects) about turn taking in conversation.
+Because the speech act effect is both new and theoretically motivated, and because there are still effects of age and linguistic condition associated with it, we believe that these findings are informative: they tell us about the conditions under which children and adults engage in spontaneous linguistic prediction about turn taking in conversation.
 

### 2016-06-23 paper/Letters/ResponseLetter-Jun2016.tex
resubmitted manuscript

@@
 
-\noindent Your and the reviewers' comments have helped us make substantial improvements in the quality and clarity of our developmental story and the theoretical discussion regarding turn prediction in third- and first-person measures.
+\noindent Your and the reviewers' insightful comments have once again helped us make substantial improvements in the quality and clarity of our developmental story and also in the theoretical discussion regarding turn prediction with third- and first-person measures.
 Since our last submission, we have updated all of the statistical models---including the 20k models of the random baseline data---to systematically test for effects of the inter-turn gap.
@@
 Please find a description of the specific comments addressed below.
-Thank you very much for your reconsideration!
+Thank you very much again for your helpful comments and for your reconsideration!
 Please let us know if there is any further information that you require about the submission.
@@
 
-\noindent We thank the editor and our two reviewers for their very helpful comments on the most recently submitted manuscript. We have tried to address their comments here and in the manuscript.
+\noindent We greatly thank the editor and our two reviewers for their very helpful comments on the most recently submitted manuscript.
+We have done our best to address their comments here and in the manuscript.
 
@@
 
-\noindent R2 found our new description of the permutation analysis satisfactory, but R1 still expressed some doubts that we believe may have arisen from a miscommunication.
-We therefore focused the description of the permutation analyses more on the purpose of the analyses (PXX).
-In particular, R1 expressed concern that the permutation analyses are anti-conservative because we assume that ``baseline saccades are equiprobable across the turn.'' While we appreciate R1's point, this characterization is not quite correct and we believe that our permutation analyses are doing exactly what they should: namely, testing the likelihood that we could get significant effects (of any predictor) if the saccades turned out to be randomly distributed---a true random baseline.
+\noindent R2 found our new description of the permutation analysis satisfactory, but R1 still expressed some doubts that we believe may have arisen from a miscommunication on our part.
+We have therefore focused the description of the permutation analyses more on the purpose of the analyses (pp.
+32--33).
+In particular we wanted to address R1's concern that the permutation analyses are anti-conservative.
+R1 expresses worry that we assume that ``baseline saccades are equiprobable across the turn.'' We have clarified the text to show that this is luckily not the case.
 Crucially, our permutations assume that baseline saccades are equiprobable across the entire stimulus, and not just across turns.
+They test the likelihood that we could get significant effects (of any predictor) if the saccades turned out to be randomly distributed---a true random baseline.
 R1 may imagine that we are comparing (a) anticipation rates at transition windows against (b) rates during turns, thereby maximizing our chances of finding a difference from the baseline.
@@
 (2013, 2015) studies.
-Instead, we compare (a) anticipation rates at transition windows to (b) transition rates over the whole stimulus.
+Instead, we compare anticipation rates at transition windows to transition rates over the whole stimulus.
 In other words, our baseline is agnostic to the placement of the true transition windows.
 
-\noindent One possible response would be that presumably, most saccades are at transition windows and so, by comparing them to saccades over the whole stimulus, we are merely comparing them to at a watered down saccadic rate.
-We have two responses to this objection, however: First, the primary aim of the permutation analysis is actually to test R1's assumption (i.e., anticipation at transition windows is greater than elsewhere), but has the substantial added benefit of testing whether the assumption holds for each predictor effect in the mixed effect models.
+\noindent One possible response R1 might have to this is that, presumably, most saccades are at transition windows and so, by comparing them to saccades over the whole stimulus, we are merely comparing our time windows of interest to at a watered down saccadic rate.
+We have two responses to this objection: First, the primary aim of the permutation analysis is to test the assumption inherent in that idea (i.e., that anticipation at transition windows is greater than elsewhere), but has the substantial added benefit of testing whether the assumption holds for each individual predictor effect in the mixed effect models.
 Second, the non-``transition window'' parts of the stimulus contain many behaviors that likely elicit anticipatory gaze behavior.
@@
 \noindent We therefore conclude that our permutation analyses are (a)  fit to the question we pose, (b) present a more useful and conservative baseline than used in prior literature, and (c) are conservative in the sense that there are other turn-relevant behaviors hidden in the non-``transition window'' parts of the stimulus.
-By editing the motivation behind the analyses on PXX, we hope we have made this argument clearer.
-R1 may still be dissatisfied with the permutation analyses in that they do not capture anything more than ``random-looker'' vs.
+By editing the motivation behind the analyses, we hope we have made this argument clearer for readers.
+We hope too that this satisfies R1's concerns.
+He or she may still be somewhat unimpressed with the permutation analyses becauuse they do not capture anything more than ``random-looker'' vs.
 ``turn-driven looker'' (in his/her words, that they are ``superfluous''), but this is exactly what they are intended to do.
-We do note, however, that we also test a more subtle hypothesis about anticipatory looking rates (i.e., ``boredom-driven looker'' vs.
+We do think that R1's point about boredom-driven looking is interesting, however, which is why we test this more subtle hypothesis about anticipatory looking rates (i.e., ``boredom-driven looker'' vs.
 ``turn-driven looker'') in a separate set of analyses (Boredom Hypothesis; Appendix C, also discussed below).
+
+\newpage
 
@@
 R1 points to the Figure C.1 to highlight that children still sometimes look away from the speaker long before the turn ends.
-We agree that our analyses can't rule out the fact that some proportion of the saccades might be due to boredom (or to chance).
-Surely some of them are.
+We completely agree that our analyses can't rule out the fact that \textit{some} proportion of the saccades might be due to boredom (or to chance).
+Surely R1 is right that some of them are!
 That said, we do not think that this is a major issue for our conclusions.
-First, the primary effects in our data (which revolve around questions and linguistic access) are not predicted by a boredom account---a bored looker should be constant in their boredom for questions and for non-questions.
+First, the primary effects in our data (which revolve around questions and access to linguistic cues) are not predicted by a boredom account---a bored looker should be constant in their boredom for questions and for non-questions.
 Second, our data show dramatic difference in the timing of look-away rates across turns of different length, which is also not predicted by a boredom-based account: boredom-driven look-aways would occur at a constant rate, independent of the ongoing turn's ultimate length.
@@
 We therefore believe that whatever boredom-driven effects exist in our data are likely small and do not interfere with the interpretation of our primary variables of interest.
-We have added discussion on this point to the Appendix section dedicated to the boredom hypothesis (PXX).
-We also added a footnote earlier in the manuscript, pointing readers to this extra analysis in case they are interested (PXX).
+We have added discussion on this point to the Appendix section dedicated to the boredom hypothesis (pp.
+108--109).
+We also added a footnote earlier in the manuscript, pointing readers to this extra analysis in case they are also interested in this follow-up (p.
+37).
 
@@
 
-\noindent To supplement these analyses, we have added standard (non-permutation based) comparisons between the Normal speech and No speech (``party'') conditions for each age and have integrated them into the results for Expt 2.
+\noindent We appreciate that our use of the random baseline data to compute likelihood is an unusual approach.
+We chose it to fit the particular analysis needs of these studies.
+That said, R1's comments encouraged us to include more traditional tests to supplement these analyses (and perhaps thereby make the findings more comprehensible overall).
+We have therefore added standard (non-permutation based) comparisons between the Normal speech and No speech (``party'') conditions for each age and have integrated them into the results for Expt 2 (pp.
+57--58).
 
@@
 
-\noindent R1 points out that, we should treat gap duration as more than just a control variable; that its effect is both strong and theoretically interesting.
-Following R1's helpful suggestion, we have added discussion about the role of inter-turn gaps for retroactive turn boundary assignment throughout the manuscript, from the Introduction to the General Discussion.
+\noindent R1 suggests we should treat gap duration as more than just a control variable; that its effect is both strong and theoretically interesting.
+We think this is a really great point.
+Following R1's suggestion, we have added discussion about the role of inter-turn gaps for retroactive turn boundary assignment throughout the manuscript, from the Introduction to the General Discussion.
 We have worked to make it clear why gaps themselves can be a good cue to turn transition.
-We have also taken up R1's suggestion for reframing in the General Discussion; we now discuss the developmental trajectory with respect to both the condition-based linguistic effects and inter-turn silence, which more naturally integrates the effect of gap duration into our story (PXX).
-By doing this, we move ``away from the idea that linguistic prediction is the only explanatory factor'' as R1 suggests.
+We have also taken up R1's suggestion for reframing in the General Discussion; we now discuss the developmental trajectory with respect to both the condition-based linguistic effects and inter-turn silence, which more naturally integrates the effect of gap duration into our story.
+By doing this, we move ``away from the idea that linguistic prediction is the only explanatory factor'' as R1 proposes.
 R1 also noticed that we didn't include an age-by-duration or condition-by-duration interaction in our models.
-This is because of we were unable to add full interactions between gap duration and the other fixed effects in all models; in many models, higher-order interactions with gap duration create non-convergence issues due to data sparsity with some combinations of three or more predictor levels.
+That was because we were unable to add full interactions between gap duration and the other fixed effects in all models; in many models, higher-order interactions with gap duration create non-convergence issues due to data sparsity with some combinations of three or more predictor levels.
+We carefully considered how we could satisfy R1's comment while also retaining consistency and validity in our models.
 
-\noindent To address this issue, in our revision, we attempt to test the effect of gap duration but also remain consistent across all of our models.
-We added a fixed effect of gap duration and two-way interactions between gap duration and all other simple fixed effects (i.e., gap duration and age, linguistic condition, and transition type) to every model.
+\noindent Our analyses now test the effect of gap duration but also remain consistent across all of our models.
+We achieved this by adding a fixed effect of gap duration and two-way interactions between gap duration and all other simple fixed effects (i.e., gap duration and age, linguistic condition, and transition type) to every model.
 This way all of our models test both the overall effect of gap duration (as before) and any interaction between gap duration and age, linguistic condition, or transition type.
-This tests the hypothesis that the gap duration effect might change with age (which, incidentally, we do not see in our data) as well as other potential effects of gap duration.
+This tests the hypothesis that the gap duration effect might change with age that R1 put forward (which, incidentally, we do not see in our data) as well as other potential interactions between gap duration and other predictors.
 
@@
 
-\noindent R1 mentions that the linguistic effects in our study are relatively small, tucked within interactions.
-We take R1's point, but also expected from the start that linguistic effects would strongly interact with age (the reason we used a broad age range) and that anticipatory gaze would interact with speech act (the reason we included question coding).
-We did not, however, anticipate that the effect of speech act would dominate our results.
-The strong effects of speech act effectively funnels all the linguistic effects (and their age interactions) into the ``question'' subset of the data.
+\noindent R1 mentions that the linguistic effects in our study are relatively small, all tucked within interactions.
+R1's comment is correct, but we would also like to point out that we expected from the start that linguistic effects would strongly interact with age (the reason we used a broad age range) and that anticipatory gaze would interact with speech act (the reason we included question coding).
+We did not, however, anticipate that the effect of speech act would so strongly dominate our results.
+The strong effect of speech act effectively funnels all the linguistic effects (and their age interactions) into the ``question'' subset of the data.
 Because the speech act effect is both new and theoretically motivated, and because there are still effects of age and linguistic condition associated with it, we believe that these findings are informative: they tell us about the conditions under which children and adults engage in spontaneous linguistic prediction about turn taking in conversation.
@@
 Thanks!
-These are now addressed in the text (PXX).
+These are now addressed in the text.
 R2 also urges us to consider our findings with respect to other cues that might drive anticipation (e.g., non-verbal cues).
-We have added a short point to the discussion that makes clear that the linguistic effects we study here are only a few of the possible sources for anticipatory information, and that the systematic study of other cues, especially non-verbal ones is still needed (PXX).
+We have added a short point to the discussion that makes clear that the linguistic effects we study here are only a few of the possible sources for anticipatory information, and that the systematic study of other cues, especially non-verbal ones is still needed (p.
+64).
 
@@
 
-\noindent R1 argues that it is an oversimplification for us to argue that ``turn prediction develops in early childhood'' and that a better story would highlight two components: (a) reactive cues (e.g., inter-turn gap duration) that children may use early on while they (b) gradually learn the association between certain linguistic structures/speech acts and speaker transitions (for which they must use linguistic cues).
-We thank R1 for this suggestion.
-As mentioned above, we have integrated discussions of gap duration as a cue to turn taking throughout the text, outlining this more nuanced developmental trajectory in multiple places (e.g., PXX).
+\noindent R1 proposes that it is an oversimplification for us to argue that ``turn prediction develops in early childhood'', and that a better story would highlight two components: (a) reactive cues (e.g., inter-turn gap duration) that children may use early on while they (b) gradually learn the association between certain linguistic structures/speech acts and speaker transitions (for which they must use linguistic cues).
+We thank R1 for this suggestion!
+As mentioned above, we have integrated discussions of gap duration as a cue to turn taking throughout the text, outlining this more nuanced developmental trajectory in multiple places.
 
@@
 He or she also raises the possibility that the question effect might have developmental roots, e.g., from the high frequency of questions in children's input.
-We thank R2 for the comment and have added some discussion concerning these points to PXX.
+We thank R2 for the great comment and have added some discussion concerning these points (p.
+67).
+
+\newpage
 
@@
 If so, R2 would like us to comment on how, then, speakers still successfully participate in conversation.
-We have added a brief note on our overall anticipation rates to the text of the General Discussion (PXX).
-Because this is a somewhat niche question that most readers may not be interested in, we only respond more elaborately here:
+We have added a brief note on our overall anticipation rates to the text of the General Discussion (p.
+68).
+Because this is a somewhat niche question that most readers may not be interested in, we only respond more elaborately here.
+But we thank R2 for encouraging us to think more about how our findings fit in with others on turn prediction:
 
@@
 There is no reason to think that children would, unlike adults, always use the same predictive strategies.
-Therefore, in some sense, we do see our rates of anticipation as challenging for the idea that participants always or even usually make precise prediction about upcoming turn ends, but our findings are not the first to be consistent with this idea.
-But what does this mean for first-person prediction?
+Therefore, in some sense, we do see our rates of anticipation as challenging for the idea that participants always or even usually make precise prediction about upcoming turn ends, but our findings are not the first to support that idea.
+What does this mean for first-person prediction?
 
@@
 The question effect might be amplified when participants themselves need to respond.
-We briefly address these points in the text (PXX).
+We briefly address these points in the text (p.
+68).
 
-\noindent In sum, we think there is still good reason to believe that participants routinely anticipate upcoming turn structure (e.g., the ends of ongoing turns and the likelihood of upcoming turn transition), but our study runs counter to the idea that they do this spontaneously at all possible points of turn transition.
+\noindent In sum, we think there is still good reason to believe that participants routinely anticipate upcoming turn structure (e.g., the ends of ongoing turns and the likelihood of upcoming turn transition), but our study does run counter to the idea that they do this spontaneously at all possible points of turn transition.
 Our findings are in-line with an account that emphasizes flexibility in how and when participants make predictions about turn taking.
@@
 
-\noindent R2 suggests that the subgroup of participants who have exposure to a second or even third language might behave differently because of their more varied experiences with language and turn taking (particularly with respect to turn-relevant cues such as prosody).
-We agree that this is a fascinating question, but we will not analyze this subgroup separately for a few reasons: first, as the reviewer guesses, we would not have enough power.
+\noindent R2 makes the intriguing suggestion that the subgroup of participants who have exposure to a second or even third language might behave differently because of their more varied experiences with language and turn taking (particularly with respect to turn-relevant cues such as prosody).
+We agree that this is a fascinating question, but we will not analyze this subgroup separately for a few reasons: first, as the reviewer guesses, we would not have enough statistical power.
 Second, and perhaps more importantly, we did not conduct a full language history of each participant and so we do not know whether the second language comes from parent(s), grandparents, a nanny, friends, school, etc., and therefore can't confidently group children into a uniform ``multilingual'' group.
-We have added a footnote on this point (PXX).
+We do think this is an important point, though, so we've added a footnote on this point (p.
+17).
 
@@
 He or she suggests that the mechanism could be linguistic, attentional, or both.
-We agree that it is likely both linguistic and attentional, and have added a note about this in the manuscript (PXX).
+We agree that it is likely both linguistic and attentional, and have added a note about this in the manuscript (p.
+70).
 
@@
 
-\noindent R2 points out that the status of the prosodic cues in the non-English recordings from Experiment 1 is not clear because, though they might be recognizable by English speakers as ``similar to'' English question cues, they are still phonologically and phonetically not identical to actual English question cues.
-We agree, but we also think that this doesn't bear too much on our findings.
-However, to make sure that this fact doesn't pass readers by, we now acknowledge the difference the non-English prosodic cues more emphatically (PXX).
+\noindent R2 correctly points out that the status of the prosodic cues in the non-English recordings from Experiment 1 is not clear because, though they might be recognizable by English speakers as ``similar to'' English question cues, they are still phonologically and phonetically not identical to actual English question cues.
+We agree, but we also think that this doesn't bear too heavily on our findings.
+However, to do justice to R2's comment and to make sure that this fact doesn't pass readers by, we now acknowledge the difference the non-English prosodic cues more emphatically (p.
+20).
 
@@
 \noindent R2 would like us to more explicitly mention the benefits of our method for measuring anticipation.
-We thank R2 for his or her encouragement and have added a this to the General Discussion (PXX).
+We thank R2 for his or her encouragement and have added a this to the General Discussion (pp.
+73--74).
 
@@
 We link this request to R1's comments about the use of gap duration for predictive and reactive turn tracking, and to R2's comments about the generalization of our anticipation rates to conversation more generally.
-As discussed above, we have taken some space in the General Discussion to speculate about the possible effects of measuring predictions about turn taking from a first- or third-person perspective (e.g., PXX).
+As discussed above, we have taken some space in the General Discussion to speculate about the possible effects of measuring predictions about turn taking from a first- or third-person perspective (e.g., pp.
+68--69 and pp.
+72--74).
 
@@
 
-\noindent Finally, we addressed a number of minor issues, including removing internal numbering, using APA-formatted references, adding a brief note about the non-centered age variable for each experiment, and some smaller comments on confusing wording and typos.
+\noindent We addressed a number of minor issues, including removing internal numbering, using APA-formatted references, adding a brief note about the non-centered age variable for each experiment, and some smaller comments on confusing wording and typos.
 
@@
 \noindent In making these changes, we feel that we have substantially improved the clarity of our developmental story and our theoretical and methodological contributions.
-We thank the editor and reviewers again for their helpful comments and hope that they also find the manuscript to be much improved.
+We give big thanks to the editor and reviewers again for their insightful and helpful comments and hope that they also find the manuscript to be much improved!
 

### 2016-06-23 paper/June2016-Submission/CasillasFrank-Anticip.tex
resubmitted manuscript

--- 2016-06-21 paper/CasillasFrank-Anticip.tex
@@
 \begin{abstract}
-Children begin developing turn-taking skills in infancy but take several years to assimilate their growing knowledge of language into their turn-taking behavior.
+Children begin developing turn-taking skills in infancy but take several years to fluidly integrate their growing knowledge of language into their turn-taking behavior.
 In two eye-tracking experiments, we measured children's anticipatory gaze to upcoming responders while controlling linguistic cues to turn structure.
@@
 But, unlike prior work with adults, lexical information alone was not sufficient either---children's performance was best overall with lexicosyntax and prosody together.
-Our findings support an account in which turn tracking and prediction emerges in infancy, and then only gradually becomes integrated with linguistic processing.
+Our findings support an account in which turn tracking and turn prediction emerge in infancy and then gradually become integrated with children's online linguistic processing.
 \end{abstract}
@@
 Children participate in sequential coordination (proto-turn taking) with their caregivers starting at three months of age---before they can rely on any linguistic cues \citep[see, among others, ][]{bateson1975, hilbrink2015, jaffe2001, snow1977}.
-However, infant turn taking is different from adult turn taking in several ways: it is heavily scaffolded by caregivers, has different timing from adult turn taking, and lacks semantic content \citep{hilbrink2015, jaffe2001}.
-But children's early, turn-structured social interactions are presumably a critical precursor to their later conversational turn taking.
-These early non-verbal interactions likely establish the protocol by which children come to use language with others.
+However, infant turn taking is different from adult turn taking in several ways: it is heavily scaffolded by caregivers, has different inter-turn timing, and lacks semantic content \citep{hilbrink2015, jaffe2001}.
+But children's early, turn-structured social interactions are presumably a critical precursor to their later conversational turn taking, establishing the protocol by which children come to use language with others.
 How then do children integrate linguistic knowledge with these preverbal turn-taking abilities?
 
-In this study, we investigate when children begin to make predictions about upcoming turn structure in conversation and how they integrate language into their predictions as they grow older.
+In this study, we investigate when children begin to make predictions about upcoming turn structure in conversation and how online linguistic processing becomes integrated into their predictions as they grow older.
 We first give a basic review of turn-taking research and the state of current knowledge about adult turn prediction.
@@
 Adults have a lot of information at their disposal to help make accurate predictions.
-Lexical, syntactic, and prosodic information (e.g., \textit{wh}- words, subject-auxiliary inversion, and list intonation) can all inform addressees about upcoming linguistic structure \citep{de-ruiter2006, duncan1972, ford1996, torreira2015}.
+Lexical, syntactic, and prosodic information (e.g., \textit{wh}-words, subject-auxiliary inversion, and list intonation) can all inform addressees about upcoming linguistic structure \citep{de-ruiter2006, duncan1972, ford1996, bogelstorreira2015}.
 Non-verbal cues (e.g., gaze, posture, and pointing) often appear at turn-boundaries and can sometimes act as late indicators of an upcoming speaker switch \citep{rossano2009, stivers2010}.
@@
 With pitch-flattened speech, the timing of participants' button responses was comparable to their timing with the full linguistic signal.
-But when no lexical information was available, participants' responded significantly earlier within the turn.
+But when no lexical information was available, participants responded significantly earlier within the turn.
 The authors concluded that lexicosyntactic information\footnote{The ``lexicosyntactic'' condition only included flattened pitch and so was not exclusively lexicosyntactic---the speech would still have residual prosodic structure, including syllable duration and intensity.} was necessary and possibly sufficient for turn-end projection, while intonation was neither necessary nor sufficient.
@@
 Prosody is tied closely to the syntax of an utterance, so the two linguistic signals are difficult to control independently \citep{ford1996}.
-\citet*{torreira2015} used a combination of button-press and verbal responses to investigate the relationship between lexicosyntactic and prosodic cues in turn-end prediction.
+\citet*{bogelstorreira2015} used a combination of button-press and verbal responses to investigate the relationship between lexicosyntactic and prosodic cues in turn-end prediction.
 Critically, their stimuli were cross-spliced so that each item had full prosodic cues to accompany the lexicosyntax.
@@
 And when intonational phrase boundaries were embedded within multi-utterance turns, participants were tricked into pressing the ``turn-end'' button 29\% of the time.
-These findings suggest that listeners actually do rely on prosodic cues to execute a response in interaction with their predictions about the unfolding syntactic structure (see also de \citet{de-ruiter2006}:525).
+These findings suggest that listeners actually do rely on prosodic cues to execute a response, and that their use of prosodic cues interacts with their predictions about the unfolding syntactic structure (see also de \citet{de-ruiter2006}:525).
 These experimental findings corroborate other corpus and experimental work promoting a combination of cues (lexicosyntactic, prosodic, and pragmatic) as key for accurate turn-end prediction \citep{duncan1972, ford1996, hirvenkari2013}.
@@
 
-Before children manage to integrate linguistic cues into their turn-taking behaviors (for both turn prediction and production), they can rely on non-verbal interactional cues, including silence, eye gaze, body orientation, and gesture, to identify the boundaries of social actions.
+Before children manage to fully integrate linguistic processing into their turn-taking behaviors (for both turn prediction and production), they can rely on non-verbal interactional cues, including silence, eye gaze, body orientation, and gesture, to identify the boundaries of social actions.
 For example, with little to no linguistic knowledge, children are often able to infer desired responses to offers and requests by taking account of their interlocutor's non-verbal communicative behavior, the structure of routine events, and the affordances of the current interactional context \citep{reddy13, nomikou11, shatz78}.
 With respect to turn taking in particular, children's spontaneous vocalizations during interaction demonstrate a sensitivity to short inter-speaker gaps from infancy \citep{hilbrink2015}.
-Thus, before they can anticipate turn structure form linguistic cues, children might react to silence as a cue to upcoming speaker change.
+Thus, before children can anticipate turn structure by integrating linguistic cues from unfolding speech, they might react to silence as a cue to upcoming speaker change.
 Interactional silence itself may then serve as one of children's first cues to turn structure, giving them information about when to respond before they can rely on language.
 
-As children's language competence increases, they can use linguistic cues to make predictions about upcoming turn structure.
+As children's language competence and speed of processing increases \citep{kail1991}, they become better equipped to use linguistic cues in making predictions about upcoming turn structure.
 Studies of early linguistic development point to a possible early advantage for prosody over lexicosyntax in children's turn-taking predictions.
-Infants can distinguish their native language's rhythm type from others soon after birth \citep{mehler1988, nazzi2003}; they show preference for the typical stress patterns of their native language over others by 6--9 months (e.g., iambic vs.
+Infants can distinguish their native language's rhythm type from others soon after birth \citep{mehler1988, nazzi2003}.
+They also show preference for the typical stress patterns of their native language over others by 6--9 months (e.g., iambic vs.
 trochaic), and can use prosodic information to segment the speech stream into smaller chunks from 8 months onward \citep{johnson2001, morgan1995}.
@@
 
-% Prior work has focused mainly on lexicosyntax and intonation, and not on prosody proper (\citealp{de-ruiter2006}; \citealp{keitel2013}, but see \citealp{torreira2015}), even though infants seem to acquire the basic rhythmic properties of the prosodic signal first \citep{mehler1988, moon1993, nazzi2003}.
+% Prior work has focused mainly on lexicosyntax and intonation, and not on prosody proper (\citealp{de-ruiter2006}; \citealp{keitel2013}, but see \citealp{bogelstorreira2015}), even though infants seem to acquire the basic rhythmic properties of the prosodic signal first \citep{mehler1988, moon1993, nazzi2003}.
 
@@
 They showed their participants two types of videos---one normal and one with flattened pitch---to test the role of intonation in participants' anticipatory predictions about upcoming speech.
-Comparing children's anticipatory gaze frequency to a random baseline, they found that only 36-month-olds and adults made anticipatory gaze switches more often than expected by chance, and only 36-month-olds were affected by flattened intonation contours.
+Comparing children's anticipatory gaze frequency to a random baseline, they found that only 36-month-olds and adults made anticipatory gaze switches more often than expected by chance, and that only 36-month-olds were affected by flattened intonation contours.
 This finding led Keitel and colleagues to conclude that children's ability to predict upcoming turn structure relies on their ability to comprehend the stimuli lexicosemantically.
-They also suggest that intonation might play a secondary role in turn prediction, but only after children acquire more sophisticated, adult-like language comprehension abilities (also see \citealp{keitel2015}).
+They also suggest that intonation might play a secondary role in turn prediction, but only after children acquire more sophisticated, adult-like language comprehension skills (also see \citealp{keitel2015}).
 
@@
 Despite these minor methodological drawbacks, the Keitel et al.
-\citeyearpar{keitel2013} study still demonstrates interesting age-based differences in children's ability to predict upcoming turn structure.
+\citeyearpar{keitel2013} study still demonstrates interesting age-based differences in children's predictions about upcoming turn structure.
 Our current work takes these findings as a starting point.\footnote{But also see \citet{casillas2012, casillas2013}.}
@@
 
-Our goal in the current study is to find out when children begin to make predictions about upcoming turn structure and to understand how their predictions are affected by linguistic cues across development.
+Our goal in the current study is to find out when children begin to make predictions about upcoming turn structure and to understand how their predictions are affected by linguistic cues to turn taking across development.
 We present two experiments in which we measured children's anticipatory gaze to responders while they watched conversation videos with natural (people speaking English vs.
@@
 Anticipation is strongest for stimuli with the full range of linguistic cues.
-Our findings support an account in which turn prediction emerges in infancy, but becomes fully integrated with linguistic processing only gradually.
+Our findings support an account in which turn prediction emerges in infancy and becomes integrated with online linguistic processing gradually, possibly because of children's increased linguistic knowledge and speed of processing with development.
 
@@
 
-We recorded participants' eye movements as they watched six short videos of two-person (dyadic) conversation interspersed with attention-getting filler videos.
+We recorded participants' eye movements as they watched six short videos of two-person (dyadic) conversation that were interspersed with attention-getting filler videos.
 Each conversation video featured an improvised discourse in one of five languages (English, German, Hebrew, Japanese, and Korean).
@@
 Approximately one-third (N$=$25) of the children's parents and teachers reported that their child regularly heard a second (and sometimes third or further) language, but only one child frequently heard a language that was used in our non-English video stimuli, and we excluded his data from the analyses.\footnote{Multilingual children may make predictions about upcoming turn structure differently from their monolingual peers due to their more varied experiences with linguistic cues to turn taking.
-However, we are unable to test this hypothesis here due to the variability in multilingual language input and the diverse set of languages being learned in our sample.
+We are unable to test this hypothesis here due to the variability in multilingual language input and the diverse set of languages being learned in our sample.
 The same applies to Experiment 2.} None of the adult participants reported fluency in a second language.
@@
 We marked non-English turns as questions when at least 4 of the 5 listeners (80\%) said that the turn ``sounded like a question''.
-Thus, ``question'' cues in the non-English condition only resembled native English question cues, and therefore were likely harder to identify than cues to questionhood in the English condition.
-However, since participants did not speak the non-English languages and would only ever treat ``question-sounding'' turns as questions, we proceeded with these analyses to see how pervasive question effects were---could they show up even without lexical access in a foreign language?
+Thus, ``question'' cues in the non-English condition only \textit{resembled} native English question cues, and were therefore likely harder to identify than cues to questionhood in the English condition.
+However, since participants did not speak the non-English languages and would only ever treat ``question-sounding'' turns as questions, we proceeded with these analyses to see how pervasive question effects were---could they show up even without lexical access?
 If participants primarily rely on prosodic cues to question turns, it's possible that even non-English prosody can elicit anticipatory gaze switches for question-like turns.
@@
 Participants sat in front of an SMI 120Hz corneal reflection eye-tracker mounted beneath a large flatscreen display.
-The display and eye-tracker were secured to a table with an ergonomic arm that allowed the experimenter to position the whole apparatus at a comfortable height, approximately 60 cm from the viewer.
+The display and eye-tracker were secured to a table with an ergonomic arm that allowed the experimenter to position the whole apparatus at a comfortable height and approximately 60 cm from the viewer.
 We placed stereo speakers on the table, to the left and right of the display.
@@
 Again, because it takes 200 msec to plan an eye movement, we counted anticipatory switches, at the latest, 200 msec after the onset of speech.
-Therefore, for very early and very late switches, our requirement of 100 msec of fixation on each speaker would sometimes extend outside of the gaze launch window boundaries (200 msec before and after the inter-turn gap; Figure \ref{fig:criterion}).
+Therefore, for very early and very late switches, our requirement of 100 msec of fixation on each speaker would sometimes extend outside of the gaze launch window boundaries (200 msec before and after the inter-turn gap; dark gray boxes Figure \ref{fig:criterion}).
 The maximally available fixation window was therefore 100 msec before and after the earliest and latest possible switch point (300 msec before and after the inter-turn gap).
@@
 \end{center}
-\caption{Schematic summary of the criteria for anticipatory gaze shifts from speaker A to speaker B during a turn transition.}
+\caption{Schematic summary of the criteria for anticipatory gaze shifts from speaker A to speaker B during a turn transition.
+FIX = hypothetical fixation on speaker A or speaker B; dashed lines = hypothetical saccadic time.}
 \label{fig:criterion}
@@
 We expected participants to show greater anticipation in the English videos than in the non-English videos because of their increased access to linguistic information in English.
-We also predicted that anticipation would be greater following questions compared to non-questions; questions have early cues to upcoming turn transition (e.g., \textit{wh-} words, subject-auxiliary inversion), and also make a next response immediately relevant.
-Our third prediction was that anticipatory looks would increase with development, along with children's increased linguistic competence.
+We also predicted that anticipation would be greater following questions compared to non-questions; questions have early cues to upcoming turn transition (e.g., \textit{wh-}words, subject-auxiliary inversion) and also make a next response immediately relevant.
+Our third prediction was that anticipatory looks would increase with development, along with children's increased linguistic competence and speed of processing.
 Finally, we predicted that transitions with longer inter-turn gaps would show greater anticipation because longer gaps provide (a) more time to make a gaze switch and (b) are themselves a cue to possible upcoming speaker switch.
@@
 non-question), age (3, 4, 5; numeric; intercept as age=0), and duration of the inter-turn gap (seconds, e.g., 0.441) as predictors, with two-way interactions between gap duration and the other simple fixed effects (language condition, transition type, and age) and a three-way interaction between language condition, transition type, and age.
-We included the two-way interactions with gap duration in case the effect of inter-turn silence changes with age or linguistic cueing (e.g., if children older children rely less on silence as a cue)\footnote{We test these two-way interactions with gap duration in all of the models reported in this paper.
-Higher-order interactions with gap duration usually resulted in model non-convergence due to distributional sparsity when three or more predictor values were considered.}.
-We also included random effects of item (turn transition) and participant, with maximal random slopes of condition, transition type, and their interaction for participants \citep{barr2013}.\footnote{The models we report in this paper are all qualitatively unchanged by the exclusion of their random slopes.
+We included the two-way interactions with gap duration in case the effect of inter-turn silence changes with age or linguistic cueing (e.g., if children older children rely less on silence as a cue).\footnote{We test these two-way interactions with gap duration in all of the models reported in this paper.
+Higher-order interactions with gap duration usually resulted in model non-convergence due to distributional sparsity when three or more predictor values were considered, so we did not include them.} We also included random effects of item (turn transition) and participant, with maximal random slopes of condition, transition type, and their interaction for participants \citep{barr2013}.\footnote{The models we report in this paper are all qualitatively unchanged by the exclusion of their random slopes.
 We have left the random slopes in because of minor participant-level variation in the predictors modeled.}
@@
 Our primary analysis (above) makes the assumption that participants' eye movements generally follow the turn structure of the stimulus, i.e., that participants track the current speaker and switch their gaze to the upcoming speaker near turn transitions.
-As just described, based on this assumption, we used linear mixed effects regressions to see how anticipatory looking is affected by aspects of participant group (age) and stimulus (e.g., transition type, language condition).
+As just described, based on this assumption, we used linear mixed effects regressions to see how anticipatory looking is affected by aspects of participant group (e.g., age) and stimulus (e.g., transition type, language condition).
 But what if the assumption that participants generally track turn structure were wrong?
@@
 For example, if participants were randomly looking back and forth between the two speakers, we might still find some anticipatory switching by chance.
-To test whether our primary results (the regression output above) could have arisen from random switching, not linked to turn structure, we conducted a secondary analysis comparing participants' anticipatory gaze at real and randomly shuffled points of turn transition.
+To test whether our primary results (the regression output above) could have arisen from random switching we conducted a secondary analysis comparing participants' anticipatory gaze at real and randomly shuffled points of turn transition.
 
 We conducted this analysis by running the same regression models on participants' eye-tracking data, only this time calculating their anticipatory gaze switches with respect to randomly permuted turn transition windows.
-This process involved: (1) randomizing the order and temporal placement of the analysis windows within each stimulus (see Figure \ref{fig:shuffling}; ``analysis window'' as shown in Figure \ref{fig:criterion}) to randomly redistribute the analysis windows across the eye-tracking signal, (2) re-running each participant's eye tracking data through switch identification (described above) on each of the randomly permuted analysis windows, and (3) modeling the anticipatory switches from the randomly permuted data (our random baseline dataset) with the same statistical models we used for the original dataset (Table \ref{tab:E1-models}).
+This process involved: (1) randomizing the order and temporal placement of the analysis windows within each stimulus (Figure \ref{fig:shuffling}; ``analysis window'' is as shown in Figure \ref{fig:criterion}) to randomly redistribute the analysis windows across the eye-tracking signal, (2) re-running each participant's eye tracking data through switch identification (described above) on each of the randomly permuted analysis windows, and (3) modeling the anticipatory switches from the randomly permuted data (our random baseline dataset) with the same statistical models we used for the original dataset (Table \ref{tab:E1-models}).
 Importantly, although the onset time of each transition was shuffled within the eye-tracking signal, the other intrinsic properties of each turn transition (e.g., prior speaker identity, transition type, gap duration, language condition, etc.) stayed constant across each permutation.
@@
 The random shuffling procedure de-links participants' gaze data from the turn structure in the original stimulus, thereby allowing us to compare turn-related (original) and non-turn-related (randomly permuted) looking behavior using the same eye movement data.
-We created 5,000 permutations of the original turn transitions (step 1 above), thereby creating 5,000 anticipatory gaze datasets with randomly de-linked gaze data (step 2 above).
-Because the randomly shuffled turn transitions could occur anywhere in the stimulus (so long as they didn't overlap each other within a single iteration), the resulting turn-transition windows collectively covered the entire stimulus---during speech and silence, during speaker change and speaker continuation, and during all turn transitions in the stimulus, even those excluded in the original analyses (e.g., because they were overlapped).\footnote{This technique crucially differs from that used by Keitel and colleagues (2013, 2015), which tests anticipatory gaze at turn transitions against anticipatory gaze during speech, which maximized the possibility of finding a difference from the baseline measure.} Pooled together, the anticipatory gaze datasets yielded an average anticipatory switch rate for each participant over all possible starting points in the stimuli: a random baseline.
-
+We created 5,000 permutations of the original turn transitions, thereby creating 5,000 anticipatory gaze datasets with randomly de-linked gaze data.
+Because the randomly shuffled turn transitions could occur anywhere in the stimulus (so long as they didn't overlap each other within a single iteration), the resulting turn-transition windows collectively covered the entire stimulus---during speech and silence, during speaker change and speaker continuation, and during all turn transitions in the stimulus, even those excluded in the original analyses (e.g., because they were overlapped).
+This technique crucially differs from that used by Keitel and colleagues (2013, 2015), which tests anticipatory gaze at turn transitions against anticipatory gaze during speech.
+Pooled together, our 5,000 anticipatory gaze datasets yielded an average anticipatory switch rate for each participant over all possible starting points in the stimuli: a random baseline.
 Using this technique we compared participants' anticipatory switches at turn transition windows to their anticipatory switches over the stimulus as a whole.
 If participants looked randomly back and forth between the speakers, we would have seen similar patterns in both cases.
+
 Rather than simply comparing participants' overall anticipatory switch rates with real and random transition windows, we estimated the likelihood that each of the predictor effects in the original data (e.g., the effect of language condition; Table \ref{tab:E1-models}) could have arisen with random gaze switching: we ran identical statistical models on the real and randomly permuted data sets.
@@
 Transition type, the interaction between transition type and gap duration, and the interaction between language condition and transition type showed \textit{z}-values that exceeded 100\%, 99.8\%, and 95\% of random \textit{z}-values, respectively (all \textit{p}$\leq$.05).
-See Section \ref{sec:permutation} for more information on each predictor's random permutation distribution.\footnote{This baseline analysis tests ``random looking'' against ''turn-driven looking'', but it does not test subtypes of turn-driven looking.
-For example, children might switch their gaze from the current speaker to the addressee out of boredom with the ongoing speech rather than active anticipation of an upcoming response.
-We address this hypothesis about ``turn-transition'' gaze switches vs.
-``boredom'' gaze switches in \ref{sec:boredlooks}}
+See \ref{sec:permutation} for more information on each predictor's random permutation distribution.\footnote{This baseline analysis tests ``random looking'' against ''turn-driven looking'', but it does not test subtypes of turn-driven looking.
+For example, children might switch their gaze from the current speaker to the addressee out of boredom with the ongoing speech rather than from active anticipation of an upcoming response.
+We address this hypothesis about ``boredom'' gaze switches vs.
+``turn-transition'' gaze switches in \ref{sec:boredlooks}}
 
@@
 Overall, participants made few anticipatory switches after non-questions, even in the English videos when they had full linguistic access.
-Prior work using online, metalinguistic tasks has shown that participants can use linguistic cues to accurately predict upcoming turn ends \citep{torreira2015, magyari2012, de-ruiter2006}.
+Prior work using online, metalinguistic tasks has shown that participants can use linguistic cues to accurately predict upcoming turn ends \citep{bogelstorreira2015, magyari2012, de-ruiter2006}.
 The current results add a new dimension to our understanding of how listeners make predictions about turn ends: both children and adults spontaneously monitor the linguistic structure of unfolding turns for cues to imminent responses.
@@
 Participants might use long silences to retroactively assign turn boundaries and anticipate speaker switches that were otherwise not anticipated (in this case, because the preceding turn was not a question).
-An alternative explanation for gap duration effects is that longer inter-turn gaps result in longer analysis windows, which yields more time for participants to make an anticipatory gaze.
-However, if participants are generally more likely to make a switch at question transitions, as our results suggest, we would expect that longer gaps would benefit questions more than non-questions---the opposite pattern from what the data show here.
+An alternative explanation for effects of gap duration is that longer inter-turn gaps result in longer analysis windows, which gives participants more time to make an anticipatory gaze.
+However, if participants are generally more likely to make a switch at question transitions (as our results suggest), and if question-driven switches aren't already at ceiling when gaps are short, we would expect that longer gaps would benefit questions more than non-questions---the opposite pattern from what the data show here.
 We take this as evidence that inter-turn silence may be most useful when participants have limited ability to make predictions about upcoming speaker transitions.
@@
 
-We used three pairs of puppets for the \textit{normal} condition---`red', `blue' and `yellow'---and one pair of puppets for each partial cue condition: `robots', `merpeople', and `party-goers' (Figure 8).
+We used three pairs of puppets for the \textit{normal} condition---`red', `blue' and `yellow'---and one pair of puppets for each partial cue condition: `robots', `merpeople', and `party-goers' (Figure \ref{fig:puppets}).
 We randomly assigned half of the conversation topics (`birthday cake', `pets', and `breakfast') to the \textit{normal} condition, and half to the partial cue conditions (`riding bikes', `rainy days', and `the library').
@@
 There were 85 turn transitions for analysis after excluding transitions longer than 550 msec and shorter than 90 msec.
-The remaining turn transitions had more questions than non-questions (N$=$47 and N$=$38, respectively), with transitions distributed somewhat evenly across conditions (keeping in mind that there were three \textit{normal} videos and only one video for each partial cue condition in each experiment version): \textit{normal} (N$=$36), \textit{words only} (N$=$13), \textit{prosody only} (N$=$17), and \textit{no speech} (N$=$19).
+The remaining turn transitions had more questions than non-questions (N$=$47 and N$=$38, respectively), with transitions distributed somewhat evenly across conditions, keeping in mind that there were three \textit{normal} videos and only one video for each partial cue condition in each experiment version: \textit{normal} (N$=$36), \textit{words only} (N$=$13), \textit{prosody only} (N$=$17), and \textit{no speech} (N$=$19).
 Inter-turn gaps for questions (mean$=$366, median$=$438, stdev$=$138 msec) were longer than those for non-questions (mean$=$305, median$=$325, stdev$=$94 msec) on average, but gap duration was overall comparable across conditions: \textit{normal} (mean$=$334, median$=$321, stdev$=$130 msec), \textit{words only} (mean$=$347, median$=$369, stdev$=$ 115 msec), \textit{prosody only} (mean$=$365, median$=$369, stdev$=$104 msec), and \textit{no words} (mean$=$319, median$=$329, stdev$=$136 msec).
@@
 Similarly, the adults' data showed significant differentiation from the randomly permuted data for the effect of language condition (\textit{words only} speech; greater than 98.3\% of random \textit{z}-values, \textit{p}$<$.02).
-See Section \ref{sec:permutation} for more information on each predictor's random permutation distribution.
+See \ref{sec:permutation} for more information on each predictor's random permutation distribution.
 
@@
 Compared to the \textit{no speech} condition, children made significantly more anticipatory switches in the \textit{normal} speech condition for questions at ages 6, 4, and 3, and also marginally at age 2 (6-year-olds: \textit{t}(36.919)$=$3.8019, \textit{p}$<$.001; 4-year-olds: \textit{t}(41.449)$=$2.9777, \textit{p}$<$.01; 3-year-olds: \textit{t}(35.724)$=$2.4286, \textit{p}$<$.05; 2-year-olds: \textit{t}(41.078)$=$1.8018, \textit{p}$=$.079).
-Children's anticipatory switches for questions did not differ in the \textit{no speech} and \textit{normal} speech conditions at ages 5 or 1 (5-year-olds: \textit{t}(29.406)$=$1.2783, \textit{p}$=$.211; 1-year-olds: \textit{t}(35.907)$=$0.4961, \textit{p}$=$.623).
+Children's anticipatory switches for questions did not significantly differ in the \textit{no speech} and \textit{normal} speech conditions at ages 5 or 1 (5-year-olds: \textit{t}(29.406)$=$1.2783, \textit{p}$=$.211; 1-year-olds: \textit{t}(35.907)$=$0.4961, \textit{p}$=$.623).
 In contrast, children's anticipatory switch rates for non-question turns were not significantly different between the \textit{no speech} and \textit{normal} speech conditions at any age (all \textit{p}$>$0.09).
-Consistent with the regression results, children were more likely to show an effect of linguistic content as they got older, but only for question transitions.
+Thus, consistent with the regression results, children were more likely to show an effect of linguistic content as they got older, but only for question transitions.
 
@@
 Participants made anticipatory gaze switches at above-chance rates starting at age two for both questions and non-questions.
-Longer gaps had a broader impact on participants' anticipations in this experiment; we saw that, overall, longer inter-turn gaps resulted in more anticipatory switches, with the \textit{no speech} condition showing equal or stronger effects of gap duration than all other conditions.
+Longer gaps had a broader impact on participants' anticipations in this second experiment; we saw that, overall, longer inter-turn gaps resulted in more anticipatory switches, with the \textit{no speech} condition showing equal or stronger effects of gap duration than all other conditions.
 
@@
 They also give no support for the idea that lexical information is sufficient on its own to support children's anticipatory switching.
-They do underscore the developing relationship between the use of linguistic cues, inter-turn silence, and speech act (transition type) in spontaneous predictions about upcoming turn structure.
+They do underscore the developing relationship between the online use of linguistic cues, inter-turn silence, and speech act in spontaneous predictions about upcoming turn structure.
 
@@
 Children begin to develop conversational turn-taking skills long before their first words emerge \citep{bateson1975, hilbrink2015, jaffe2001, snow1977}.
-As they acquire language, they also acquire the information needed to make accurate predictions about upcoming turn structure.
+As they become fast and knowledgeable language users, they also become able to make accurate predictions about upcoming turn structure.
 Until recently, we have had very little data on how children weave language into their already-existing turn-taking behaviors.
@@
 We also found no early advantage for prosody over lexicosyntax; children's anticipatory switch rates in the \textit{prosody only} condition were initially low, but showed significant gains by age five.
-We discuss these findings with respect to the role of linguistic cues and inter-turn silence for predicting upcoming turn structure, the importance of questions in predictions about conversation, and children's developing competence as conversationalists.
+We discuss these findings with respect to the role of linguistic processing and inter-turn silence for predicting upcoming turn structure, the importance of questions in predictions about conversation, and children's developing competence as conversationalists.
 
@@
 
-Prior work with adults has found a consistent role for lexicosyntax in predicting upcoming turn structure \citep{de-ruiter2006, magyari2012}, whereas the role of prosody is still under debate \citep{duncan1972, ford1996, torreira2015}.
+Prior work with adults has found a consistent role for lexicosyntax in predicting upcoming turn structure \citep{de-ruiter2006, magyari2012}, whereas the role of prosody is still under debate \citep{duncan1972, ford1996, bogelstorreira2015}.
 Knowing that children comprehend more about prosody than lexicosyntax early on (see \citealp{speer2009} for a review), we thought it possible that young children would instead show an advantage for prosody in their predictions about turn structure in conversation.
@@
 Thus, using prosody alone to accurately predict turn boundaries in conversation appears to be difficult for adults and children.
-Prosodic information is continuous, multidimensional, and indexically complex---it encodes syntactic structure, speech act, and extralinguistic information, sometimes simultaneously, without clear one-to-one mappings between form and meaning \citep{cutler1997, shriberg1998, lammertink2015}.
-For these reasons, prosodic information alone may not be enough to both (a) make precise temporal predictions about turn structure, and (b) identify question turns, which otherwise appear to drive anticipatory gaze switching.
-Therefore, although children show early facility with prosodic discrimination \citep{nazzi2003, soderstrom2003, johnson2001, jusczyk1995, morgan1995, mehler1988}, using prosodic knowledge for turn prediction may generally be too difficult without additional information from lexical or syntactic cues.
+Prosodic information is continuous, multidimensional, and can index multiple meanings at once---it encodes syntactic structure, speech act, and extralinguistic information without clear one-to-one mappings between form and meaning \citep{cutler1997, shriberg1998, lammertink2015}.
+For these reasons, prosodic information alone may not be enough for young children to easily make precise temporal predictions about turn structure, and identify question turns in unfolding speech.
+Therefore, although children show early facility with prosodic discrimination \citep{nazzi2003, soderstrom2003, johnson2001, jusczyk1995, morgan1995, mehler1988}, using prosodic knowledge for turn prediction may be difficult without additional information from lexical or syntactic cues.
 
@@
 Pauses are detected and related to phrasal structure from early on; 5-month-old infants use pauses to parse intonational phrases \citep{mannel2009}.
-The lack of interactions between age and gap duration suggests that the use of inter-turn silence remains important for older speakers and the interactions between transition type and gap duration (Experiment 1) and condition and gap duration (Experiment 2; marginal), suggest that this effect is simply the result of having more time to make a gaze switch.
+The lack of interactions between age and gap duration suggests that the use of inter-turn silence remains important for older speakers and the interactions between transition type and gap duration (Experiment 1) and condition and gap duration (Experiment 2; marginal), suggest that this effect is not simply the result of having more time to make a gaze switch.
 These findings thus suggest that silence is an early and lasting cue for identifying turn structure online when other predictive information is not adequate.
@@
 That said, children's lexical and syntactic knowledge is limited for quite some time (\citealp{tomasello1999}, but see also \citealp{bergelson2013, shi2010}).
-Our stimuli were made in a child-friendly style, they are still other-directed and fairly complex, with 20--30 seconds of continuous conversational speech.
+Although our stimuli were made in a child-friendly style, they are still other-directed and fairly complex, with 20--30 seconds of continuous conversational speech.
 
 It is perhaps for this reason that children's performance was always best with the full signal, where lexicosyntactic information was supported by prosodic information and vice versa.
-Even in adults, Torreira and colleagues (\citeyear{torreira2015}) showed that the trade-off in informativity between lexical and prosodic cues is more subtle in semi-natural speech.
+Even in adults, B\"{o}gels and Torreira (\citeyear{bogelstorreira2015}) showed that the trade-off in informativity between lexical and prosodic cues is more subtle in semi-natural speech.
 The present findings are the first to show evidence of a similar effect developmentally.
@@
 Moreover, because of their high frequency and relatively limited number of formats, questions, especially \textit{wh}-questions, may be more identifiable and predictable compared to other types of speech acts.
-So, in addition to having a special interactive status, questions are a frequent, predictable and core characteristic of many caregiver-child interactions, motivating a general benefit for questions in turn structure anticipation.
-
-Two important questions for future work are then: (1) how does children's ability to monitor for questions in conversation relate to their prior experience with questions?
+So, in addition to having a special interactive status, questions are a frequent, predictable, and core characteristic of many caregiver-child interactions, motivating a general benefit for questions in turn structure anticipation.
+
+Two important routes for future work are then: (1) how does children's ability to monitor for questions in conversation relate to their prior experience with questions?
 and (2) what is it about questions that makes children and adults more likely to anticipatorily switch their gaze to addressees?
@@
 
-The question effect suggests that participants' spontaneous predictions may be driven by what lies \textit{beyond} the end of the current turn---not just by the upcoming end of the turn itself, as has been focused on in prior work \citep{torreira2015, keitel2013, magyari2012, de-ruiter2006}.
-In future work, it will be crucial to measure prediction from a first-person perspective to find out what kinds of predictions are most relevant to addressees in conversation (see also \citealp{holler2015}).
+A question effect suggests that participants' spontaneous predictions may be driven by what lies \textit{beyond} the end of the current turn---not just by the upcoming end of the turn itself, as has been focused on in prior work \citep{bogelstorreira2015, keitel2013, magyari2012, de-ruiter2006}.
+In future work, it will be crucial to measure prediction from a first-person perspective to find out what kinds of predictions are most relevant to addressees in conversation.
 
@@
 One of the core aims of our study was to test whether children show an early competence for turn taking, as is proposed by studies of spontaneous mother-infant proto-conversation and theories about the mechanisms underlying human interaction in general \citep{hilbrink2015, levinson2006}.
-We found evidence that young children make spontaneous predictions about upcoming turn structure: definitely at age two and marginally at age one.
+We found evidence that young children make spontaneous predictions about upcoming turn structure: definitely by age two and marginally by age one.
 
@@
 
-To be clear, young children's ``above chance'' performance was often still far from adult-like predictive behavior---children at ages one and two were still very close to chance in their anticipations and, even at age six, children were not fully adult-like in their predictions.
-This may indicate that young children rely primarily on non-verbal cues (like inter-turn silence) in anticipating turn transitions while adults use both verbal and non-verbal cues to make predictions.
-Relatedly, adults may be more expert in flexibly adapting to the turn-relevant cues present at any moment, e.g., responding to non-English prosodic cues in Experiment 1.
-
-Taken together, our data suggest that turn-taking skills do begin to emerge in infancy, but that children cannot consistently make effective predictions until they can identify and react to question turns.
+To be clear, young children's ``above chance'' performance was often still far from adult-like predictive behavior---turn prediction (and the concurrent use of linguistic cues from unfolding speech) increased only gradually with age.
+Children at ages one and two were still very close to chance in their anticipations and, even at age six, children were not fully adult-like in their predictions.
+This indicates that young children may at first rely primarily on non-verbal cues, like inter-turn silence, to anticipate turn transitions but that, by adulthood, listeners use both verbal and non-verbal cues to make predictions.
+Relatedly, adult listeners may be more expert in flexibly adapting to the turn-relevant cues present at any moment, e.g., responding to non-English prosodic cues in Experiment 1.
+
+Taken together, our data suggest that turn-taking skills do begin to emerge in infancy, but that children cannot consistently make effective predictions until they can identify question turns in unfolding speech and react to them quickly.
 This finding leads us to wonder how participant role (first- instead of third-person) and differences in early interactional experience (e.g., frequent vs.
 infrequent question-asking from caregivers) feed into this early predictive skill.
-It also bridges prior work showing a predisposition for turn taking in infancy (e.g., \citealp{bateson1975, hilbrink2015, jaffe2001, snow1977}) with children's apparently \textit{late} acquisition of adult-like competence for turn taking in actual conversation \citep{casillas2016, garvey1984, garvey1981, ervin-tripp1979} and reinforces the idea that it takes children several years to fully integrate linguistic information into their turn-taking systems \citep{casillas2016,garvey1981}.
+It also bridges prior work showing a predisposition for turn taking in infancy (e.g., \citealp{bateson1975, hilbrink2015, jaffe2001, snow1977}) with children's apparently \textit{late} acquisition of adult-like competence for turn taking in spontaneous conversation \citep{casillas2016, garvey1984, garvey1981, ervin-tripp1979}.
+It also reinforces the idea that it takes children several years to fully integrate linguistic information into their turn-taking systems \citep{casillas2016,garvey1981}.
+
+What makes the integration of linguistic information so gradual?
+We suspect that two slow-developing processes---children's linguistic knowledge (e.g., \textit{wh}-words, subject-auxiliary inversion) and their speed of processing for linguistic information (e.g., parsing and retrieval)---both contribute to their ability to make predictions about turn structure in unfolding speech.
+Children may be able to integrate predictive cues for turn taking from the start, but their knowledge of these cues and their speed in parsing and recognizing them may be too slow at first for use in online prediction.
+This account falls in line with the early and continued use of non-verbal cues  found in the current study, but more work is needed to tease these developmental threads apart.
 
@@
 Many prior studies have used phonetically-altered speech with infants and young children \citep[cf.][]{jusczyk2000}, but few of them have done so in a conversational context.
-Future work could instead carefully script speech or cross-splice sub-parts of turns to control for the presence of linguistic cues for turn transition (see, e.g., \citealp{torreira2015}).
+Future work could instead carefully script speech or cross-splice sub-parts of turns to control for the presence of linguistic cues for turn transition (see, e.g., \citealp{bogelstorreira2015}).
 
@@
 It also lends itself to many possibilities for controlling the presence of individual verbal and non-verbal cues and has a clear method for assessing random switching baselines across the entire stimulus.
-Also, if it is the case that response preparation interferes with our ability to see prediction at the ends of incoming turns, third-person paradigms are one of the only ways to measure prediction processes in isolation.
+Also, if it is the case that response preparation interferes with our ability to see prediction at the ends of incoming turns \citep{levinson2016}, third-person paradigms are one of the only ways to measure prediction processes in isolation.
 
@@
 Adults use linguistic information to accurately predict turn structure in conversation, which facilitates their online comprehension and allows them to respond relevantly and on time.
-The present study offers new findings regarding the role of speech acts and linguistic processing in online turn prediction, and has given evidence that turn prediction emerges by age two, increases with age, and is driven by question turns.
-However, turn prediction is not fully integrated with linguistic cues until much later and, in the absence of predictive linguistic cues, children and adults alike rely on retroactive cues such as inter-turn silence to predict upcoming speaker change.
+The present study offers new findings regarding the role of speech acts and linguistic processing in online turn prediction, and has given evidence that turn prediction emerges by age two, increases with age, and is driven by the ability to identify and react to question turns in unfolding speech.
+However, children's successful integration of online linguistic processing and online predictions about upcoming turn structure develops gradually.
+When participants can't use predictive linguistic cues (because they are absent, unfamiliar, or are processed too late), children and adults alike rely on retroactive cues such as inter-turn silence to predict upcoming speaker change.
 Using language to make predictions about upcoming interactive content takes time to develop and, for participants of all ages appears to be primarily driven by participants' expectations about what will happen next, beyond the end of the current turn.
@@
 
-We then re-ran the statistical models from the original data on each of the random permutations, e.g., using Experiment 1's original model to analyze the anticipatory switches from each random permutation of the Experiment 1 looking data.
+We then re-ran the statistical models from the original data on each of the random permutations, e.g., using Experiment 1's original model structure to analyze the anticipatory switches from each random permutation of the Experiment 1 looking data.
 We could then calculate the proportion of random data \textit{z}-values exceeded by the original \textit{z}-value for each predictor.
@@
 In comparing the real and randomly permuted datasets, we excluded the output of random-permutation models that gave convergence warnings to remove erratic model estimates from our analyses.
-Non-convergent models made up 22.4--24.4\% of the random permutation models in Experiment 1 and 69--70\% of the random permutation models in Experiment 2.
+Non-convergent models made up 13--14\% of the random permutation models in Experiment 1 and 46--49\% of the random permutation models in Experiment 2.
 The \textit{z}-values for each predictor in the converging and non-converging models from Experiment 1 are shown in Table \ref{tab:nonconv_e1}.
 
-Although many of the non-converging models show estimates within range of the converging models (e.g., with a mean difference of only 0.09 in median \textit{z}-value across predictors), they also show many radically outlying estimates (e.g., showing a mean difference of 237.3 in mean \textit{z}-value across predictors).
+Although many of the non-converging models show estimates within range of the converging models (e.g., with a mean difference of only 0.096 in median \textit{z}-value across predictors), they also show many radically outlying estimates (e.g., showing a mean difference of 146.7 in mean \textit{z}-value across predictors).
 Similar patterns were obtained in the non-converging models for Experiment 2 and persisted across multiple attempts with different optimizers.
@@
     \hline
-    (Intercept)					& -2.52	& -458.42	& -2.54	& -2.86	& 0.87	& 1319.22	& -5.53	& -8185.36		& 0.41	& 0.97 \\
-    Age								& -0.51	& -17.83	& -0.49	& -0.53	& 0.79	& 83.78		& -3.71	& -672.2		& 2.3		& 342.8 \\
-    LgCond							& -0.53	& -109.91	& -0.55	& -0.63	& 0.93	& 564.42	& -3.93	& -4418.74		& 3.23	& 2296.19 \\
-    Type							& -0.1	& -29.66	& -0.09	& -0.1	& 0.98	& 515.12	& -4.06	& -4383.92		& 3.36	& 3416.68 \\
-    Duration						& 0.99	& 345.53	& 0.98	& 1.15	& 1.07	& 1323.13	& -2.44	& -5048.24		& 5.78	& 9985.16 \\
-    Age*LgCond					& 0.19	& 10.64		& 0.2		& 0.18	& 0.9		& 109.6		& -3.31	& -581.61		& 3.59	& 946.81 \\
-    Age*Type						& 0.02	& -1.8		& 0.001	& -0.04	& 0.9		& 98.27		& -3.36	& -884.36		& 3.45	& 640.43 \\
-    LgCond*Type					& 0.2		& 45.32		& 0.2		& 0.27	& 0.96	& 691.3		& -3.12	& -4160.06		& 3.39	& 5107.64 \\
-    Age*LgCond*Type			& -0.12	& -14.23	& -0.12	& -0.15	& 0.93	& 156.72	& -2.98	& -1318.26		& 2.90	& 927.69 \\
+		(Intercept)			&	-1.56	&	-901.93	&	-1.59	&	-1.94	&	0.98	&	1945.68	&	-4.89	&	-11942.58	&	2.16	&	1840.2 \\
+		Age					&	-0.2	&	-26.08	&	-0.21	&	-0.28	&	0.92	&	193.6	&	-4.06	&	-1151.44	&	3.57	&	751.7 \\
+		LgCond				&	-0.54	&	-313.61	&	-0.56	&	-0.77	&	1.04	&	1281.84	&	-4.55	&	-7781.18	&	3.51	&	4341.3 \\
+		TType				&	-0.03	&	-22.27	&	-0.03	&	0.01	&	1.05	&	1099.5	&	-3.42	&	-7137.95	&	3.56	&	5034.84 \\
+		GapDur				&	0.42	&	511.04	&	0.45	&	0.61	&	1.09	&	3555.2	&	-3.86	&	-15899.54	&	3.88	&	21151.4 \\
+		Age*LgCond			&	0.18	&	6.46	&	0.18	&	0.23	&	0.91	&	160.59	&	-3.35	&	-791.57		&	3.69	&	950.17 \\
+		Age*TType			&	0.02	&	-7.08	&	-0.01	&	-0.05	&	0.9		&	152.2	&	-3.45	&	-815.06		&	3.43	&	741.38 \\
+		LgCond*TType		&	0.18	&	-5.76	&	0.2		&	0.21	&	0.97	&	1129.35	&	-3.26	&	-6230.78	&	3.4		&	5997.59 \\
+		Age*GapDur			&	-0.11	&	-24.39	&	-0.08	&	-0.12	&	0.99	&	536.89	&	-4.08	&	-2897.34	&	2.87	&	2602.11 \\
+		LgCond*GapDur		&	0.22	&	475.09	&	0.2		&	0.4		&	1.12	&	2988.88	&	-3.83	&	-14231.85	&	4.02	&	17307.34 \\
+		Ttype*GapDur		&	-0.02	&	-37.07	&	-0.03	&	-0.12	&	1.13	&	2824.93	&	-4.51	&	-16493.61	&	4.73	&	14994.45 \\
+		Age*LgCond*TType	&	-0.1	&	-2.92	&	-0.11	&	-0.21	&	0.93	&	241.44	&	-3.34	&	-1434.96	&	3.02	&	1333.34 \\
     \hline
@@
     \hline
-    (Intercept)    			& -1.63	& -126.14	& -1.71	& -1.73	& 0.97	& 713.39	& -4.08	& -12111.22	& 2.15	& 649.55		\\
-    LgCond  						& -0.26	& -679.6	& -0.3	& -0.53	& 1.02	& 15894.33	& -3.45	& -494979.7	& 3.35	& 88581.58	\\
-    Type 							& -0.11	& 6.29		& -0.13	& -0.04	& 1.11	& 501.5		& -3.85	& -6420.75		& 3.28	& 8177.88		\\
-    Duration  					& 0.25	& 84.09		& 0.27	& 0.26	& 1.1		& 1152.94	& -3.25	& -10864.51	& 3.46	& 18540.62	\\
-    LgCond*Type  				& 0.12	& -242.27	& 0.1		& 0.34	& 1.07	& 26836.7	& -3.41	& -622642.7	& 3.81	& 509198.4	\\
-    LgCond*Duration			& 0.15	& 780.03	& 0.16	& 0.39	& 1.04	& 44105.02	& -3.84	& -798498.6	& 3.55	& 1145951	\\
-    Type*Duration 				& 0.05	& -6.56		& 0.05	& 0.02	& 1.13	& 1389.9	& -3.54	& -15979.22	& 3.87	& 16419.46	\\
-    LgCond*Type*Duration	& -0.06	& 1083.63	& -0.08	& -0.21	& 1.1		& 63116.54	& -4.21	& -1201895		& 4.02	& 1284965	\\
+		(Intercept)		&	-1.85	&	-135.7	&	-1.9	&	-1.96	&	0.96	&	707.63	&	-4.48		&	-8056.34	&	1.61	&	654.56 \\
+		LgCond			&	-0.35	&	-57.44	&	-0.37	&	-0.5	&	1.09	&	625.12	&	-3.8		&	-6033.9		&	3.68	&	5343.37 \\
+		TType			&	-0.06	&	9.59	&	-0.06	&	0		&	1.09	&	403.93	&	-3.54		&	-4131.97	&	3.34	&	3793.07 \\
+		GapDur			&	0.31	&	97.73	&	0.32	&	0.38	&	1.12	&	1159.99	&	-3.11		&	-7149.74	&	3.89	&	10669.09 \\
+		LgCond*TType	&	0.18	&	31.6	&	0.18	&	0.22	&	1.03	&	560.99	&	-2.87		&	-7722.35	&	3.9		&	4377.92 \\
+		LgCond*GapDur	&	0.19	&	77.34	&	0.21	&	0.18	&	1.12	&	1047.37	&	-4.18		&	-7713.96	&	3.71	&	7764.19 \\
+		Ttype*GapDur	&	0		&	-50.12	&	0.01	&	-0.07	&	1.11	&	1065.37	&	-3.42		&	-10640.42	&	3.64	&	7868.74 \\
     \hline
@@
 %\linespread{2}
+
+%\linespread{1}
+%\begin{sidewaystable}[!htbp]
+%  \begin{scriptsize}
+%\centering
+%  \begin{tabular}{lllllllllll}
+%    & Mean$_{C}$ & Mean$_{NC}$ & Median$_{C}$ & Median$_{NC}$ & \textit{SD}$_{C}$ & \textit{SD}$_{NC}$ & Min$_{C}$ & Min$_{NC}$ & Max$_{C}$ & Max$_{NC}$ \\
+%    \hline
+%	\textbf{\textit{Children}} &&&&& \\
+%    \hline
+%		(Intercept)		&	-3.05	&	-2800.55	&	-3.07	&	-1496.47	&	0.84	&	3130.77		&	-5.75	&	-15569.16	&	0.26	&	-0.34 \\
+%		Age			&	0.75	&	81.43		&	0.76	&	1.44		&	1.11	&	224.08		&	-3.33	&	-793.12		&	4.27	&	1094.44 \\
+%		LgCond(Pros)		&	0.25	&	391.42		&	0.23	&	0.59		&	1.08	&	4135.69		&	-3.21	&	-135300.8	&	3.81	&	11272.92 \\
+%		LgCond(Norm)		&	0.12	&	165.95		&	0.1	&	0.2		&	1.05	&	1544.35		&	-3.59	&	-6994.53	&	3.67	&	12125.92 \\
+%		LgCond(Lex)		&	0.01	&	-320.01		&	-0.04	&	-0.33		&	1.19	&	2842.73		&	-3.37	&	-22569.54	&	4.18	&	10000.98 \\
+%		TType			&	-0.37	&	-389.38		&	-0.38	&	-0.75		&	1.09	&	1832.78		&	-3.78	&	-9806.93	&	3.04	&	8548.36 \\
+%		GapDur			&	0.36	&	843.66		&	0.38	&	0.74		&	1.05	&	3702.7		&	-3.52	&	-24538.48	&	3.77	&	24284.02 \\
+%		Age*LgCond(Pros)	&	-0.04	&	6.36		&	-0.04	&	-0.08		&	1.06	&	551.96		&	-3.04	&	-1483.99	&	3.11	&	21619.1 \\
+%		Age*LgCond(Norm)	&	-0.04	&	-15.15		&	-0.06	&	-0.2		&	1.11	&	173.55		&	-3.44	&	-816.8	&	&	3.48	&	1008.44 \\
+%		Age*LgCond(Lex)		&	0.04	&	1.65		&	0.01	&	0.11		&	1.18	&	244.02		&	-3.76	&	-1406.31	&	3.86	&	1167.23 \\
+%		Age*TType		&	0.04	&	2.05		&	0.03	&	0.02		&	1.09	&	204.49		&	-4.19	&	-958.44		&	3.73	&	1194.13 \\
+%		LgCond(Pros)*TType	&	-0.07	&	57.09		&	-0.07	&	-0.19		&	0.96	&	3402.56		&	-3.46	&	-9987.18	&	2.89	&	132110.78 \\
+%		LgCond(Norm)*TType	&	0.05	&	-0.15		&	0.06	&	0.03		&	1.01	&	1161.34		&	-4	&	-6401.71	&	3.52	&	5301.89 \\
+%		LgCond(Lex)*TType	&	-0.13	&	-64.62		&	-0.12	&	-0.18		&	1.08	&	1705.68		&	-3.8	&	-9255.27	&	3.18	&	10020.13 \\
+%		Age*GapDur		&	-0.21	&	-56.54		&	-0.24	&	-0.4		&	1.1	&	405.89		&	-3.73	&	-1958.19	&	3.27	&	1777.06 \\
+%		LgCond(Pros)*GapDur	&	-0.14	&	-635		&	-0.12	&	-0.27		&	1.1	&	5604.19		&	-3.9	&	-26632.24	&	3.28	&	24770.67 \\
+%		LgCond(Norm)*GapDur	&	0.27	&	529.45		&	0.29	&	0.48		&	1.08	&	3739.49		&	-3.37	&	-26880.55	&	3.99	&	24169.51 \\
+%		LgCond(Lex)*GapDur	&	0.14	&	-378.79		&	0.19	&	0.63		&	1.19	&	68742.57	&	-4.53	&	-3282549.89	&	3.45	&	53088.85 \\
+%		TType*GapDur		&	0.5	&	1361.03		&	0.5	&	0.94		&	1.13	&	4678.69		&	-2.85	&	-19023.27	&	4.3	&	28029.73 \\
+%		Age*LgCond(Pros)*TType	&	-0.24	&	-68.05		&	-0.25	&	-0.38		&	1.05	&	593.7		&	-3.37	&	-21989.44	&	2.72	&	1431.64 \\
+%		Age*LgCond(Norm)*TType	&	-0.11	&	-17.34		&	-0.12	&	-0.22		&	1.11	&	248.73		&	-3.78	&	-1136.43	&	3.75	&	1206.99 \\
+%		Age*LgCond(Lex)*TType	&	-0.08	&	-11.56		&	-0.08	&	0		&	1.15	&	349.91		&	-3.75	&	-1565.3		&	3.85	&	1702.72 \\
+%    \hline
+%	\textbf{\textit{Adults}} &&&&& \\
+%    \hline
+%		(Intercept)		&	-1.61	&	-457.88		&	-1.72	&	-1.76	&	0.79	&	1989.16		&	-3.36	&	-43709.32	&	1.1	&	907.26 \\
+%		LgCond(Pros)		&	0	&	944.53		&	-0.06	&	-0.05	&	1.01	&	43362.61	&	-2.51	&	-336814.65	&	3.45	&	1425038.89 \\
+%		LgCond(Norm)		&	-0.26	&	-6.07		&	-0.23	&	-0.23	&	1	&	1160.02		&	-3.4	&	-11796.29	&	2.39	&	18948.21 \\
+%		LgCond(Lex)		&	0.09	&	-4007.34	&	-0.01	&	-0.09	&	1.04	&	47289.25	&	-2.55	&	-1255648.45	&	3.3	&	186209.69 \\
+%		TType			&	-0.06	&	-15.67		&	-0.06	&	-0.1	&	1.03	&	1341.86		&	-3.07	&	-17433.35	&	3.02	&	16583.07 \\
+%		GapDur			&	0.37	&	216.75		&	0.43	&	0.43	&	0.97	&	2083.86		&	-2.39	&	-18593.61	&	3.06	&	27075.55 \\
+%		LgCond(Pros)*TType	&	0.03	&	-705.9		&	0.11	&	0.05	&	0.96	&	26567.05	&	-3.91	&	-873190.15	&	3.03	&	75193.36 \\
+%		LgCond(Norm)*TType	&	0.02	&	-70.87		&	-0.04	&	-0.03	&	1.04	&	1034.96		&	-2.66	&	-20885.3	&	3.05	&	6980.78 \\
+%		LgCond(Lex)*TType	&	0.04	&	916.68		&	0.04	&	0	&	1.06	&	9636.18		&	-3.22	&	-31426.17	&	3.61	&	194241.2 \\
+%		LgCond(Pros)*GapDur	&	-0.1	&	-3700.6		&	-0.12	&	-0.14	&	1.04	&	107644.13	&	-3.34	&	-3422355.79	&	2.95	&	582760.84 \\
+%		LgCond(Norm)*GapDur	&	0.13	&	29469.75	&	0.11	&	0.1	&	1	&	1461329.4	&	-2.71	&	-27108.45	&	3.15	&	72670835.07 \\
+%		LgCond(Lex)*GapDur	&	-0.16	&	5777.75		&	-0.12	&	-0.01	&	1.07	&	96011.19	&	-3.17	&	-1140956.82	&	2.52	&	2450843.03 \\
+%		TType*GapDur		&	0.03	&	110.05		&	0.02	&	0.09	&	1.05	&	3223.11		&	-3.84	&	-44017.73	&	2.93	&	43175.22 \\
+%    \hline
+%  \end{tabular}
+%  \caption{Estimated \textit{z}-values for each predictor in converging (\textit{C}) and non-converging (\textit{NC}) child and adult models from Experiment 2. }
+%\label{tab:nonconv_e2}
+% \end{scriptsize}
+%\end{sidewaystable}
+%%\linespread{2}
 
@@
 
-In each of the plots below, the black dot represents the real data value for the effect being shown.
+In each of the plots below, the black dot represents the real data value for the effect being shown (the difference score).
 The effect sizes from the 5,000 randomly permuted data sets are shown as a distribution.
@@
 \end{center}
-\caption{Significant pairwise comparisons of the \textit{prosody only}-\textit{no speech} linguisitic condition effect, across ages in Experiment 2. Non-significant comparisons are not shown here.}
+\caption{Significant pairwise comparisons of the \textit{prosody only}-\textit{no speech} linguisitic condition effect, across ages in Experiment 2. Non-significant comparisons are not shown.}
 \label{fig:E2-lgageinteraction}
@@
 \caption{Significant pairwise comparisons of the \textit{normal speech}-\textit{no speech} language condition effect for transition type, across ages, in Experiment 2.
-Non-significant comparisons are not shown here.}
+Non-significant comparisons are not shown.}
 \label{fig:E2-lgagetypeinteraction}
@@
 Some of these early looks away may be boredom-driven, but it is equally plausible that some of them are turn-driven.
-Early predictive behavior is common in turn-taking studies with adults, in both constrained turn-taking tasks \citep{de-ruiter2006, gisladottir2015, bogels2015} and in spontaneous conversation \citep{holler2015, torreira2015}.
+Early predictive behavior is common in turn-taking studies with adults, in both constrained turn-taking tasks \citep{de-ruiter2006, gisladottir2015, bogelstorreira2015} and in spontaneous conversation \citep{holler2015, bogels2015}.
 Although this same pattern has yet to be established for children's turn predictions, the looking behavior here is at least consistent with adult response patterns in previous work.

### 2016-06-27 paper/CasillasFrank-Anticip-SuppMaterials.tex
separated supplementary materials per editor's request

--- 2016-06-23 paper/June2016-Submission/CasillasFrank-Anticip.tex
@@
 \usepackage{rotating}
-%\usepackage{apacite}
 
@@
 \begin{document}
-
-\begin{frontmatter}
-
-\title{The development of children's ability to track and predict turn structure in conversation}
-
-\author[MPI]{Marisa Casillas\corref{cor1}}
-\address[MPI]{Max Planck Institute for Psycholinguistics, Nijmegen}
-%\address[StanfordLX]{Department of Linguistics, Stanford University}
-\cortext[cor1]{Corresponding author. \\ Address: Wundtlaan 1, 6525 XD, Nijmegen, The Netherlands\\ Email: marisa.casillas@mpi.nl \\Telephone: +31 024 3521 566; Fax: +31 024 3521 213}
-
-\author[StanfordPSY]{Michael C. Frank}
-
-\address[StanfordPSY]{Department of Psychology, Stanford University}
-
-\begin{abstract}
-Children begin developing turn-taking skills in infancy but take several years to fluidly integrate their growing knowledge of language into their turn-taking behavior.
-In two eye-tracking experiments, we measured children's anticipatory gaze to upcoming responders while controlling linguistic cues to turn structure.
-In Experiment 1, we showed English and non-English conversations to English-speaking adults and children.
-In Experiment 2, we phonetically controlled lexicosyntactic and prosodic cues in English-only speech.
-Children spontaneously made anticipatory gaze switches by age two and continued improving through age six.
-In both experiments, children and adults made more anticipatory switches after hearing questions.
-Consistent with prior findings on adult turn prediction, prosodic information alone did not increase children's anticipatory gaze shifts.
-But, unlike prior work with adults, lexical information alone was not sufficient either---children's performance was best overall with lexicosyntax and prosody together.
-Our findings support an account in which turn tracking and turn prediction emerge in infancy and then gradually become integrated with children's online linguistic processing.
-\end{abstract}
-
-\begin{keyword}
-Turn taking \sep Conversation \sep Development \sep Questions \sep Eye-tracking \sep Anticipation
-%% MSC codes here, in the form: \MSC code \sep code
-%% or \MSC[2008] code \sep code (2000 is the default)
-
-\end{keyword}
-
-\end{frontmatter}
-
-\linenumbers
-
-\section*{Introduction}
-\label{sec:intro}
-
-Spontaneous conversation is a universal context for using and learning language.
-Like other types of human interaction, it is organized at its core by the roles and goals of its participants.
-But what sets conversation apart is its structure: sequences of interconnected, communicative actions that take place across alternating turns at talk.
-Sequential, turn-based structures in conversation are strikingly uniform across language communities and linguistic modalities.
-Turn-taking behaviors are also cross-culturally consistent in their basic features and the details of their implementation \citep{de-vos2015, dingemanse2013, stivers2009}.
-
-Children participate in sequential coordination (proto-turn taking) with their caregivers starting at three months of age---before they can rely on any linguistic cues \citep[see, among others, ][]{bateson1975, hilbrink2015, jaffe2001, snow1977}.
-However, infant turn taking is different from adult turn taking in several ways: it is heavily scaffolded by caregivers, has different inter-turn timing, and lacks semantic content \citep{hilbrink2015, jaffe2001}.
-But children's early, turn-structured social interactions are presumably a critical precursor to their later conversational turn taking, establishing the protocol by which children come to use language with others.
-How then do children integrate linguistic knowledge with these preverbal turn-taking abilities?
-
-In this study, we investigate when children begin to make predictions about upcoming turn structure in conversation and how online linguistic processing becomes integrated into their predictions as they grow older.
-We first give a basic review of turn-taking research and the state of current knowledge about adult turn prediction.
-We then discuss recent work on the development of turn-taking skills before presenting the details of the present study.
-
-\subsection*{Adult turn taking}
-
-Turn taking itself is not unique to conversation.
-Many other human activities are organized around sequential turns at action.
-Traffic intersections and computer network communication both use turn-taking systems.
-Children's early games (e.g., give-and-take, peek-a-boo) have built-in, predictable turn structure \citep{ratner1978, ross1987}.
-Even monkeys take turns: Non-human primates such as marmosets and Campbell's monkeys vocalize contingently with each other in both natural and lab-controlled environments \citep{lemasson2011, takahashi2013}.
-In all these cases, turn taking serves as a protocol for interaction, allowing the participants to coordinate with each other through sequences of contingent action.
-
-Conversational turn taking distinguishes itself from other turn-taking behaviors by the complexity of the sequencing involved.
-Conversational turns come grouped into semantically-contingent sequences of action.
-The groups can span turn-by-turn exchanges (e.g., simple question--response, ``How are you?''--``Fine.'') or sequence-by-sequence exchanges (e.g., reciprocals, ``How are you?''--``Fine, and you?''--``Great!'').
-Compared to other turn-taking behaviors, the possible sequence and action types in everyday talk are diverse and unpredictable.
-
-% Sequences of action drive the conversation forward into the next, relevant sequences of talk (e.g., ''And you?''--``Great!''--``Why's that?''; \citealp{schegloff2007}).
-To take a turn, participants need to make predictions about what conversational content will be relevant next.
-In some cases, relevant next turns are somewhat obvious (e.g., question--response) while, in other cases, there are multiple relevant next actions to choose from or no obvious next action at all (e.g., after a closing).
-
-Despite this complexity, conversational turn taking is precise in its timing.
-Across a diverse sample of conversations in 10 languages, one study found a consistent average inter-turn silence of 0--200 msec at points of speaker switch \citep{stivers2009}.
-Experimental results and current models of speech production suggest that it takes approximately 600 msec to produce a content word, and even longer to produce a simple utterance \citep{griffin2000, levelt1989}.
-In order to achieve 200 msec turn transitions, speakers must begin formulating their response before the prior turn has ended \citep{levinson2013, levinson2016}.
-Moreover, to formulate their response early on, speakers must track and anticipate what types of response might become relevant next.
-They also need to predict the content and form of upcoming speech so that they can launch their articulation at exactly the right moment.
-Prediction thus plays a key role in timely turn taking.
-
-Adults have a lot of information at their disposal to help make accurate predictions.
-Lexical, syntactic, and prosodic information (e.g., \textit{wh}-words, subject-auxiliary inversion, and list intonation) can all inform addressees about upcoming linguistic structure \citep{de-ruiter2006, duncan1972, ford1996, bogelstorreira2015}.
-Non-verbal cues (e.g., gaze, posture, and pointing) often appear at turn-boundaries and can sometimes act as late indicators of an upcoming speaker switch \citep{rossano2009, stivers2010}.
-Additionally, the sequential context of a turn can make the next action obvious: answers after questions, thanks or denial after compliments, etc.
-\citep{schegloff2007}.
-
-Prior work suggests that adult listeners primarily use lexicosyntactic information to accurately predict upcoming turn structure.
-De Ruiter and colleagues \citeyearpar{de-ruiter2006} asked participants to listen to snippets of spontaneous conversation and to press a button whenever they anticipated that the current speaker was about to finish his or her turn.
-The speech snippets were controlled for the amount of linguistic information present; some were normal, but others had flattened pitch, low-pass filtered speech, or further manipulations.
-With pitch-flattened speech, the timing of participants' button responses was comparable to their timing with the full linguistic signal.
-But when no lexical information was available, participants responded significantly earlier within the turn.
-The authors concluded that lexicosyntactic information\footnote{The ``lexicosyntactic'' condition only included flattened pitch and so was not exclusively lexicosyntactic---the speech would still have residual prosodic structure, including syllable duration and intensity.} was necessary and possibly sufficient for turn-end projection, while intonation was neither necessary nor sufficient.
-Congruent evidence comes from studies varying the predictability of lexicosyntactic and pragmatic content: adults anticipate turn ends better when they can more accurately predict the exact words that will come next (\citealp{magyari2012}; see also \citealp{magyari2014}).
-They can also identify speech acts within the first word of an utterance \citep{gisladottir2015}, allowing them to start planning their response at the first moment possible \citep{bogels2015}.
-
-Despite this body of evidence, the role of prosody for adult turn prediction is still a matter of debate.
-De Ruiter and colleagues' (2006) experiment focused on the role of intonation, which is only a partial index of prosody.
-Prosody is tied closely to the syntax of an utterance, so the two linguistic signals are difficult to control independently \citep{ford1996}.
-\citet*{bogelstorreira2015} used a combination of button-press and verbal responses to investigate the relationship between lexicosyntactic and prosodic cues in turn-end prediction.
-Critically, their stimuli were cross-spliced so that each item had full prosodic cues to accompany the lexicosyntax.
-Because of the splicing, they were able to create items that had syntactically-complete units with no intonational phrase boundary at the end.
-Participants never verbally responded or pressed the ``turn-end'' button when hearing a syntactically-complete phrase without an intonational phrase boundary.
-And when intonational phrase boundaries were embedded within multi-utterance turns, participants were tricked into pressing the ``turn-end'' button 29\% of the time.
-These findings suggest that listeners actually do rely on prosodic cues to execute a response, and that their use of prosodic cues interacts with their predictions about the unfolding syntactic structure (see also de \citet{de-ruiter2006}:525).
-These experimental findings corroborate other corpus and experimental work promoting a combination of cues (lexicosyntactic, prosodic, and pragmatic) as key for accurate turn-end prediction \citep{duncan1972, ford1996, hirvenkari2013}.
-
-\subsection*{Turn taking in development}
-
-% Adults accurately and spontaneously make predictions about upcoming turn structure.
-Their predictions rely on a sophisticated body of knowledge about linguistic structure, non-verbal signals, and social actions.
-Knowing this, we could expect that children's acquisition of turn-taking skills is closely tied to their knowledge about language, gaze, gesture, and social cues.
-But children's turn taking starts early in infancy, long before their first words or gestures emerge.
-So a primary role for lexicosyntactic cues doesn't fit well with children's pre-verbal turn taking.
-
-The majority of work on children's early turn taking has focused on observations of spontaneous interaction.
-Children's first turn-like structures appear as early as two to three months after birth, in proto-conversation with their caregivers \citep{bruner1975, bruner1985, snow1977}.
-During proto-conversations, caregivers treat their infants as capable of making meaningful contributions: they take every look, vocalization, arm flail, and burp as ``utterances'' in the joint discourse \citep{bateson1975, jaffe2001, snow1977}.
-Infants catch onto the structure of proto-conversations quickly.
-By three to four months they notice disturbances to the contingency of their caregivers' response and, in reaction, change the rate and quality of their vocalizations \citep{k-bloom1988, masataka1993, toda1993}.
-
-% Infants at this age also notice changes to social contingency outside of turn structure.
-In the Still Face paradigm, caregivers interact with their infants and then suddenly halt, taking on a neutral expression with a sustained gaze.
-When faced with this sudden disappearance of social contingency, infants three months and older try a range of methods to reinitiate the interaction, such as vocalization, reaching, and smiling before looking away or getting upset \citep{rochat1998, toda1993}.
-
-The timing of children's responses to their caregivers' speech shows a non-linear pattern.
-Infants' contingent vocalizations in the first few months of life show very fast timing (though with a lot of vocal overlap).
-But by nine months, their timing slows down considerably, only to gradually speed up again after 12 months \citep{hilbrink2015}.
-For children, taking turns with brief transitions between speakers is more difficult than avoiding speaker overlap; children's incidence of overlap is nearly adult-like by nine months, but the timing of their non-overlapped (i.e., gapped) responses remains longer than the adult 200 msec standard for the next few years \citep{casillas2016, garvey1984, garvey1981, ervin-tripp1979}.
-This puzzling pattern is likely due to children's linguistic development: taking turns on time is easier when their response is a simple vocalization rather than a linguistic utterance.
-Integrating language into the turn-taking system may therefore be a major factor in children's delayed responses \citep{casillas2016}.
-
-Before children manage to fully integrate linguistic processing into their turn-taking behaviors (for both turn prediction and production), they can rely on non-verbal interactional cues, including silence, eye gaze, body orientation, and gesture, to identify the boundaries of social actions.
-For example, with little to no linguistic knowledge, children are often able to infer desired responses to offers and requests by taking account of their interlocutor's non-verbal communicative behavior, the structure of routine events, and the affordances of the current interactional context \citep{reddy13, nomikou11, shatz78}.
-With respect to turn taking in particular, children's spontaneous vocalizations during interaction demonstrate a sensitivity to short inter-speaker gaps from infancy \citep{hilbrink2015}.
-Thus, before children can anticipate turn structure by integrating linguistic cues from unfolding speech, they might react to silence as a cue to upcoming speaker change.
-Interactional silence itself may then serve as one of children's first cues to turn structure, giving them information about when to respond before they can rely on language.
-
-As children's language competence and speed of processing increases \citep{kail1991}, they become better equipped to use linguistic cues in making predictions about upcoming turn structure.
-Studies of early linguistic development point to a possible early advantage for prosody over lexicosyntax in children's turn-taking predictions.
-Infants can distinguish their native language's rhythm type from others soon after birth \citep{mehler1988, nazzi2003}.
-They also show preference for the typical stress patterns of their native language over others by 6--9 months (e.g., iambic vs.
-trochaic), and can use prosodic information to segment the speech stream into smaller chunks from 8 months onward \citep{johnson2001, morgan1995}.
-Four- to five-month-olds also prefer pauses in speech to be inserted at prosodic boundaries, and by 6 months infants can use prosodic markers to pick out sub-clausal syntactic units, both of which are useful for extracting turn structure from ongoing speech \citep{jusczyk1995, soderstrom2003}.
-In comparison, children show at best a very limited lexical inventory before their first birthday \citep{bergelson2013, shi2010}.
-
-%If response planning (i.e., language production) is the primary hurdle in young children's spontaneous turn taking, we should find evidence that children understand turn-taking behaviors before they are able to produce the behaviors themselves.
-This hypothesis has been recently explored in experimental settings, but results are mixed.
-One study found that 12-month-olds make more predictive gaze shifts to a responder while watching human verbal conversation compared to conversation-like interactions with objects \citep{bakker2011}, but another only found a similar effect at 36 months \citep{hofsten2009}.
-However, neither of these two studies had baselines to which the turn-relevant looking behavior could be compared.
-A baseline measurement is critical because there may be developmental differences in gaze shifting between conversational participants, even if the shifting is not related to turn structure.
-Such developmental differences could produce artifactual changes in measures of turn-contingent shifting.
-
-% Children begin to develop specific expectations about conversational behavior before they begin to speak.
-Sometime between four and six months, children begin to attend differently to face-to-face and back-to-back conversation; six-month-olds follow conversational speakers more with their gaze when at least one speaker is looking at the other \citep{augusti2010}.
-At ten months, infants expect people to look and talk at other people, and not to objects \citep{beier2012}.
-At twelve months infants expect to see responses to verbal (but not non-speech) utterances in face-to-face contexts \citep{thorgrimsson2015}.
-
-% Prior work has focused mainly on lexicosyntax and intonation, and not on prosody proper (\citealp{de-ruiter2006}; \citealp{keitel2013}, but see \citealp{bogelstorreira2015}), even though infants seem to acquire the basic rhythmic properties of the prosodic signal first \citep{mehler1988, moon1993, nazzi2003}.
-
-Keitel and colleagues \citeyearpar{keitel2013} were one of the first to explore how children use linguistic cues to predict upcoming turn structure.
-They asked 6-, 12-, 24-, and 36-month-old infants, and adult participants to watch short videos of conversation and tracked their eye movements at points of speaker change.
-They showed their participants two types of videos---one normal and one with flattened pitch---to test the role of intonation in participants' anticipatory predictions about upcoming speech.
-Comparing children's anticipatory gaze frequency to a random baseline, they found that only 36-month-olds and adults made anticipatory gaze switches more often than expected by chance, and that only 36-month-olds were affected by flattened intonation contours.
-This finding led Keitel and colleagues to conclude that children's ability to predict upcoming turn structure relies on their ability to comprehend the stimuli lexicosemantically.
-They also suggest that intonation might play a secondary role in turn prediction, but only after children acquire more sophisticated, adult-like language comprehension skills (also see \citealp{keitel2015}).
-
-Although the Keitel et al.
-\citeyearpar{keitel2013} study constitutes a substantial advance over previous work in this domain, it has some limitations.
-Because these limitations directly inform our own study design, we review them in some detail.
-First, their estimates of baseline gaze frequency (``random'' in their terminology) were not random.
-Instead, they used gaze switches during ongoing speech as a baseline.
-But ongoing speech is the period in which switching is least likely to occur \citep{hirvenkari2013}---their baseline thus maximizes the chance of finding a difference in gaze frequency at turn transitions compared to the baseline.
-A more conservative baseline would compare participants' looking behavior at turn transitions to their looking behavior during randomly selected windows of time throughout the stimulus, including turn transitions.
-We follow this conservative approach in the current study.
-
-Second, the conversation stimuli \citet{keitel2013} used were somewhat unusual.
-The average gap between turns was 900 msec, a duration much longer than typical adult timing, which averages around 200 msec \citep{stivers2009}.
-The speakers in the videos were also asked to minimize their movements while performing scripted, adult-directed conversation, which would have created a somewhat unnatural interaction.
-Additionally, to produce more naturalistic conversation, it would have been ideal to localize the sound sources for the two voices in the video (i.e., to have the voices come out of separate left and right speakers).
-But both voices were recorded and played back on the same audio channel, which may have made it difficult to distinguish the two talkers.
-Again, we attempt to address these issues in our current study.
-Despite these minor methodological drawbacks, the Keitel et al.
-\citeyearpar{keitel2013} study still demonstrates interesting age-based differences in children's predictions about upcoming turn structure.
-Our current work takes these findings as a starting point.\footnote{But also see \citet{casillas2012, casillas2013}.}
-
-\subsection*{The current study}
-
-Our goal in the current study is to find out when children begin to make predictions about upcoming turn structure and to understand how their predictions are affected by linguistic cues to turn taking across development.
-We present two experiments in which we measured children's anticipatory gaze to responders while they watched conversation videos with natural (people speaking English vs.
-non-English; Experiment 1) and non-natural (puppets with phonetically manipulated speech; Experiment 2) control over the presence of lexical and prosodic cues.
-We tested children across a wide range of ages (Experiment 1: 3--5 years; Experiment 2: 1--6 years), with adult control participants in each experiment.
-We additionally tested for the use of one non-verbal cue: inter-turn silence.
-
-We highlight four primary findings: first, although children and adults use linguistic cues to make predictions about upcoming turn structure, they do so primarily to predict speaker transitions after questions (a ``speech act'' effect).
-This intriguing effect, which has not been reported previously, suggests that participants track unfolding speech for cues to upcoming speaker change, which may affect how they use linguistic cues more generally for anticipatory processing in conversation.
-Second, we find that children make more predictions than expected by chance starting at age two, but that this effect is small at first, and continues to improve through age six, along with children's use of linguistic cues to anticipate answers after question turns.
-Third, children and adults often used inter-turn silence (a non-verbal cue to turn structure) to make more predictive gaze switches to the responder, suggesting that non-verbal cues are useful for predicting turn structure early on and continue to be important in adulthood.
-Finally, we find no evidence for an early prosodic advantage in children's anticipations and, further, no evidence that lexical cues alone are comparable to the full linguistic signal in aiding children's predictions (as is proposed for adults; \citealp{de-ruiter2006}).
-Anticipation is strongest for stimuli with the full range of linguistic cues.
-Our findings support an account in which turn prediction emerges in infancy and becomes integrated with online linguistic processing gradually, possibly because of children's increased linguistic knowledge and speed of processing with development.
-
-\section*{Experiment 1}
-\label{sec:exp1}
-
-We recorded participants' eye movements as they watched six short videos of two-person (dyadic) conversation that were interspersed with attention-getting filler videos.
-Each conversation video featured an improvised discourse in one of five languages (English, German, Hebrew, Japanese, and Korean).
-Participants saw two videos in English and one in every other language.
-The participants, all native English speakers, were only expected to understand the two videos in English.
-We showed participants non-English videos to limit their access to lexical information while maintaining their access to other cues to turn boundaries (e.g., non-English prosody, gaze, in-breaths, phrase final lengthening).
-Using this method, we analyzed children and adult's anticipatory looks from the current speaker to the upcoming speaker at points of turn transition in English and non-English videos.
-
-\subsection*{Methods}
-\label{sec:methods1}
-
-\subsubsection*{Participants}
-
-We recruited 74 children ages 3;0--5;11 and 11 undergraduate adults to participate in the experiment.
-We recruited adult participants through the Stanford University Psychology participant database.
-Adult participants were either paid or received course credit for their time.
-Our child sample included 19 three-year-olds, 32 four-year-olds, and 23 five-year-olds, all enrolled in a local nursery school and all of whom volunteered their time.
-All participants were native English speakers.
-Approximately one-third (N$=$25) of the children's parents and teachers reported that their child regularly heard a second (and sometimes third or further) language, but only one child frequently heard a language that was used in our non-English video stimuli, and we excluded his data from the analyses.\footnote{Multilingual children may make predictions about upcoming turn structure differently from their monolingual peers due to their more varied experiences with linguistic cues to turn taking.
-We are unable to test this hypothesis here due to the variability in multilingual language input and the diverse set of languages being learned in our sample.
-The same applies to Experiment 2.} None of the adult participants reported fluency in a second language.
-
-\begin{figure}[t]
 \begin{center}
-\includegraphics[width=0.6\textwidth]{figures/FIG-FL-stim.png}
+\Large
+\textbf{Supplementary Materials}
 \end{center}
-\caption{Example frame from a conversation video used in Experiment 1.}
-\label{fig:speakers}
-\end{figure}
-
-\subsubsection*{Materials}
-
-\textit{Video recordings}.
-We recorded pairs of talkers while they conversed in a sound-attenuated booth (Figure \ref{fig:speakers}).
-Each talker was a native speaker of the language being recorded, and each talker pair was male-female.
-Using a Marantz PMD 660 solid state field recorder, we captured audio from two lapel microphones, one attached to each participant, while simultaneously recording video from the built-in camera of a MacBook laptop computer.
-The talkers were volunteers and were acquainted with their recording partner ahead of time.
-
-Each recording session began with a 20-minute warm-up period of spontaneous conversation during which the pair talked for five minutes on four topics (favorite foods, entertainment, hometown layout, and pets).
-Then we asked talkers to choose a new topic---one relevant to young children (e.g., riding a bike, eating breakfast)---and to improvise a dialogue on that topic.
-We asked them to speak as if they were on a children's television show in order to elicit child-friendly speech toward each other.
-We recorded until the talkers achieved at least 30 seconds of uninterrupted discourse with enthusiastic, child-friendly speech.
-Most talker pairs took less than five minutes to complete the task, usually by agreeing on a rough script at the start.
-We encouraged talkers to ask at least a few questions to each other during the improvisation.
-The resulting conversations were therefore not entirely spontaneous, but were as close as possible while still remaining child-oriented in topic, prosodic pattern, and lexicosyntactic construction.\footnote{All of the non-English talkers were fluent in English as a second language, and some fluently spoke three or more languages.
-We chose male-female pairs as a natural way of creating contrast between the two talker voices.}
-
-After recording, we combined the audio and video recordings by hand, and cropped each one to the (approximate) 30-second interval with the most turn activity.
-Because we recorded the conversations in stereo, the male and female voices came out of separate speakers during video playback.
-This gave each voice in the videos a localized source (from the left or right loudspeaker).
-We coded each turn transition in the videos for language condition (English vs.
-non-English), inter-turn gap duration (in milliseconds), and transition type (question vs.
-non-question).
-Each non-English turn was coded as a question or non-question from a monolingual English-speaker's perspective, i.e., turns that ``sound like'' questions and turns that do not.
-We asked five native American English speakers to listen to the audio recording for each non-English turn and judge whether it sounded like a question.
-We marked non-English turns as questions when at least 4 of the 5 listeners (80\%) said that the turn ``sounded like a question''.
-Thus, ``question'' cues in the non-English condition only \textit{resembled} native English question cues, and were therefore likely harder to identify than cues to questionhood in the English condition.
-However, since participants did not speak the non-English languages and would only ever treat ``question-sounding'' turns as questions, we proceeded with these analyses to see how pervasive question effects were---could they show up even without lexical access?
-If participants primarily rely on prosodic cues to question turns, it's possible that even non-English prosody can elicit anticipatory gaze switches for question-like turns.
-
-Because the conversational stimuli were recorded semi-spontaneously, the duration of turn transitions and the number of speaker transitions in each video was variable.
-We measured the duration of each turn transition from the audio recording associated with each video.
-We excluded turn transitions longer than 550 msec and shorter than 90 msec from analysis, additionally excluding overlapped transitions.\footnote{Overlap occurs when a responder begins a new turn before the current turn is finished.
-When overlap occurs, observers cannot switch their gaze in anticipation of the response because the response began earlier than expected.
-Participants expect conversations to proceed with ``one speaker at a time'' \citep{sacks1974}.
-They would therefore still be fixated on the prior speaker when the overlap started, and would have to switch their gaze \textit{reactively} to the responder.} This left approximately equal numbers of turn transitions available for analysis in the English (N$=$20) and non-English (N$=$16) videos.
-On average, the inter-turn gaps for English videos (mean$=$318, median$=$302, stdev$=$112 msec) were slightly longer than for non-English videos (mean$=$286, median$=$251, stdev$=$122 msec).
-
-Questions made up exactly half of the turn transitions in the English (N$=$10) and non-English (N$=$8) videos.
-In the English videos, inter-turn gaps were slightly shorter for questions (mean$=$310, median$=$293, stdev$=$112 msec) than non-questions (mean$=$325, median$=$315, stdev$=$118 msec).
-Non-English videos did not show a large difference in transition time for questions (mean$=$270, median$=$257, stdev$=$116 msec) and non-questions (mean$=$302, median$=$252, stdev$=$134 msec).
-
-\subsubsection*{Procedure}
-Participants sat in front of an SMI 120Hz corneal reflection eye-tracker mounted beneath a large flatscreen display.
-The display and eye-tracker were secured to a table with an ergonomic arm that allowed the experimenter to position the whole apparatus at a comfortable height and approximately 60 cm from the viewer.
-We placed stereo speakers on the table, to the left and right of the display.
-
-Before the experiment started, we warned adult participants that they would see videos in several languages and that, though they weren't expected to understand the content of non-English videos, we \textit{would} ask them to answer general, non-language-based questions about the conversations.
-Then after each video we asked participants one of the following randomly-assigned questions: ``Which speaker talked more?'', ``Which speaker asked the most questions?'', ``Which speaker seemed more friendly?'', and ``Did the speakers' level of enthusiasm shift during the conversation?'' We also asked if the participants could understand any of what was said after each video.
-The participants responded verbally while an experimenter noted their responses.
-
-Children were less inclined to simply sit and watch videos of conversation in languages they didn't speak, so we used a different procedure to keep them engaged: the experimenter started each session by asking the child about what languages he or she could speak, and about what other languages he or she had heard of.
-Then the experimenter expressed her own enthusiasm for learning about new languages, and invited the child to watch a video about ``new and different languages'' together.
-If the child agreed to watch, the experimenter and the child sat together in front of the display, with the child centered in front of the tracker and the experimenter off to the side.
-Each conversation video was preceded and followed by a 15--30 second attention-getting filler video (e.g., running puppies, singing muppets, flying bugs).
-If the child began to look bored, the experimenter would talk during the fillers, either commenting on the previous conversation (``That was a neat language!'') or giving the language name for the next conversation (``This next one is called Hebrew.
-Let's see what it's like.'') The experimenter's comments reinforced the video-watching as a joint task.
-
-All participants (child and adult) completed a five-point calibration routine before the first video started.
-We used a dancing Elmo for the children's calibration image.
-During the experiment, participants watched all six 30-second conversation videos.
-The first and last conversations were in American English and the intervening conversations were Hebrew, Japanese, German, and Korean.
-The presentation order of the non-English videos was shuffled into four lists, which participants were assigned to randomly.
-The entire experiment, including instructions, took 10--15 minutes.
-
-\subsubsection*{Data preparation and coding}
-\label{sec:algorithm}
-
-To determine whether participants predicted upcoming turn transitions, we needed to define a set of criteria for what counted as an anticipatory gaze shift.
-Prior work using similar experimental procedures has found that adults and children make anticipatory gaze shifts to upcoming talkers within a wide time frame; the earliest shifts occur before the end of the prior turn, and the latest occur after the onset of the response turn, with most shifts occurring in the inter-turn gap (Keitel et al., 2013; Hirvenkari, 2013; Tice and Henetz, \citeyear{TiceHenetz11}).
-Following prior work, we measured how often our participants shifted their gaze from the prior to the upcoming speaker \textit{before} the shift in gaze could have been initiated in reaction to the onset of the speaker's response.
-In doing so, we assumed that it takes participants 200 msec to plan an eye movement, following standards from adult anticipatory processing studies \citep[e.g., ][]{kamide2003}.
-
-We checked each participant's gaze at each turn transition for three characteristics (Figure \ref{fig:criterion}): (1) that the participant fixated on the prior speaker for at least 100 msec at the end of the prior turn, (2) that immediately thereafter the participant switched to fixate on the upcoming speaker for at least 100 ms, and (3) that the switch in gaze was initiated within the first 200 msec of the response turn, or earlier.
-These criteria guarantee that we only counted gaze shifts when: (1) participants were tracking the previous speaker, (2) switched their gaze to track the upcoming speaker, and (3) did so before they could have simply reacted to the onset of speech in the response.
-Under the assumption that it takes at least 200 msec to plan an eye movement, gaze shifts initiated within the first 200 msec of the response (or earlier) were planned \textit{before} participants could react to the onset of speech itself.
-
-As mentioned, most anticipatory switches happen in the inter-turn gap, but we also allowed anticipatory gaze switches that occurred in the final syllables of the prior turn.
-Early switches are consistent with the distribution of responses in explicit turn-boundary prediction tasks.
-For example, in a button press task, adult participants anticipated turn ends approximately 200 msec in advance of the turn's end, and anticipatory responses to pitch-flattened stimuli came even earlier \citep{de-ruiter2006}.
-We therefore allowed switches to occur as early as 200 msec before the end of the prior turn.
-Again, because it takes 200 msec to plan an eye movement, we counted anticipatory switches, at the latest, 200 msec after the onset of speech.
-Therefore, for very early and very late switches, our requirement of 100 msec of fixation on each speaker would sometimes extend outside of the gaze launch window boundaries (200 msec before and after the inter-turn gap; dark gray boxes Figure \ref{fig:criterion}).
-The maximally available fixation window was therefore 100 msec before and after the earliest and latest possible switch point (300 msec before and after the inter-turn gap).
-We did not count switches made during the fixation window as anticipatory.
-We \textit{did} count switches made during the inter-turn gap.
-The period of time from the beginning of the possible fixation window on the prior speaker to the end of the possible fixation window on the responder was our total analysis window (300 msec $+$ the inter-turn gap $+$ 300 msec).
-
-\begin{figure}[t]
-\begin{center}
-\includegraphics[width=0.7\textwidth]{figures/FIG-AnticipCriteria.png}
-\end{center}
-\caption{Schematic summary of the criteria for anticipatory gaze shifts from speaker A to speaker B during a turn transition.
-FIX = hypothetical fixation on speaker A or speaker B; dashed lines = hypothetical saccadic time.}
-\label{fig:criterion}
-\end{figure}
-
-\paragraph{Predictions}
-We expected participants to show greater anticipation in the English videos than in the non-English videos because of their increased access to linguistic information in English.
-We also predicted that anticipation would be greater following questions compared to non-questions; questions have early cues to upcoming turn transition (e.g., \textit{wh-}words, subject-auxiliary inversion) and also make a next response immediately relevant.
-Our third prediction was that anticipatory looks would increase with development, along with children's increased linguistic competence and speed of processing.
-Finally, we predicted that transitions with longer inter-turn gaps would show greater anticipation because longer gaps provide (a) more time to make a gaze switch and (b) are themselves a cue to possible upcoming speaker switch.
-
-\subsection*{Results}
-\label{sec:results1}
-
-Participants looked at the screen most of the time during video playback (81\% and 91\% on average for children and adults, respectively).
-They primarily kept their eyes on the person who was currently speaking in both English and non-English videos: they gazed at the current speaker between 38\% and 63\% of the time, looking back at the addressee between 15\% and 20\% of the time (Table \ref{tab:e1_look}).
-Even three-year-olds looked more at the current speaker than anything else, whether or not the videos were in a language they could understand.
-Children looked at the current speaker less than adults did during the non-English videos.
-Despite this, their looks to the addressee did not increase substantially in the non-English videos, indicating that their looks away were probably related to boredom rather than confusion about ongoing turn structure.
-Overall, participants' pattern of gaze to current speakers demonstrated that they performed basic turn tracking during the videos, regardless of language.
-Figure \ref{fig:E1-randvsreal} shows participants' anticipatory gaze rates across age, language condition, and transition type.
-
-\linespread{1}
-\begin{table}[t]
-\begin{center}
-  \begin{tabular}{llcccc}
-    \hline
-    Age group & Condition & Speaker & Addressee & Other onscreen & Offscreen\\
-    \hline
-    3 & English & 0.61 & 0.16 & 0.14 & 0.08 \\
-    4 & English & 0.60 & 0.15 & 0.11 & 0.13 \\
-    5 & English & 0.57 & 0.15 & 0.16 & 0.12 \\
-    Adult & English & 0.63 & 0.16 & 0.16 & 0.05 \\
-    3 & Non-English & 0.38 & 0.17 & 0.20 & 0.25 \\
-    4 & Non-English & 0.43 & 0.19 & 0.21 & 0.18 \\
-    5 & Non-English & 0.40 & 0.16 & 0.26 & 0.18 \\
-    Adult & Non-English & 0.58 & 0.20 & 0.16 & 0.07 \\
-%   Overall & & 0.49 & 0.17 & 0.18 & 0.16 \\
-    \hline
-  \end{tabular}
-\end{center}
-  \caption{Average proportion of gaze to the current speaker and addressee during periods of talk across ages in Experiment 1.}
-\label{tab:e1_look}
-\end{table}
-%\linespread{2}
-
-\begin{figure}[h]
-\begin{center}
-\includegraphics[width=1\textwidth]{figures/E1-samples-by-lang-groups-trans-types.png}
-\end{center}
-\caption{Anticipatory gaze rates across language condition and transition type for the real and randomly permuted datasets. Vertical bars represent 95\% confidence intervals.}
-\label{fig:E1-randvsreal}
-\end{figure}
-
-\linespread{1}
-\begin{table}[h!]
-\begin{small}
-\begin{center}
-  \begin{tabular}{lcccl}
-  \textbf{\textit{Children}} &&&& \\
-    \hline
-			           &  Estimate & Std. Error & \textit{z} value & Pr($>$$|$\textit{z}$|$) \\
-    \hline
-    (Intercept)   													& -0.604	& 1.242	& -0.486  & 0.627		\\
-    Age             												& -0.002	& 0.261	& -0.009  & 0.993		\\
-    LgCond$=$\textit{non-English} 						& -3.65		& 1.16	& -3.146  & 0.002 **	\\
-    TType$=$\textit{non-Question}      					& -2.95		& 1.13	& -2.61	  & 0.009 **	\\
-    GapDuration      											&  2.247	& 3.194	&  0.704  & 0.482		\\
-    Age*LgCond$=$\textit{non-English} 					&  0.5		& 0.212	&  2.353  & 0.019 *	\\
-    Age*TType$=$\textit{non-Question} 					&  0.009	& 0.196	&  0.044  & 0.965		\\
-    LgCond$=$\textit{non-English}*						&  2.692	& 1.347	&  1.999  & 0.046 *	\\
-    \hspace*{5mm} TType$=$\textit{non-Question} &&&& \\
-    Age*GapDuration												& -0.577	& 0.627	& -0.921  & 0.357		\\
-    LgCond$=$\textit{non-English}*GapDuration		&  1.143	& 2.287	&  0.5	  & 0.617		\\
-    TType$=$\textit{non-Question}*GapDuration			&  5.519	& 2.282	&  2.418  & 0.016 *	\\
-    Age*LgCond$=$\textit{non-English}*					& -0.433	& 0.304	& -1.426  & 0.154		\\
-    \hspace*{5mm} TType$=$\textit{non-Question} &&&& \\
-    \hline
-  &&&& \\
-  \textbf{\textit{Adults}} &&&& \\
-    \hline
-           &  Estimate & Std. Error & \textit{z} value & Pr($>$$|$\textit{z}$|$) \\
-    \hline
-    (Intercept)														& -0.584   & 0.64  & -0.913 & 0.361			\\
-    LgCond$=$\textit{non-English}							& -0.059   & 0.751 & -0.079 & 0.937			\\
-    TType$=$\textit{non-Question}							& -3.298   & 0.933 & -3.536 & 0.0004 ***	\\
-    GapDuration														&  0.132   & 1.766 &  0.075 & 0.941			\\
-    LgCond$=$\textit{non-English}*						&  1.234   & 0.629 &  1.961 & 0.0498 *		\\
-    \hspace*{5mm} TType$=$\textit{non-Question} &&&& \\
-    LgCond$=$\textit{non-English}*GapDuration		& -1.519   & 2.192 & -0.693 & 0.488			\\
-    TType$=$\textit{non-Question}*GapDuration			&  7.116   & 2.195 &  3.241 & 0.001 **		\\
-    \hline
-  \end{tabular}
-\end{center}
-  \end{small}
-  \caption{Model output for participants' anticipatory gaze switches in Experiment 1.}
-\label{tab:E1-models}
-\end{table}
-%\linespread{2}
-
-\subsubsection*{Statistical models}
-\label{sec:models1}
-
-We identified anticipatory gaze switches for all 36 usable turn transitions, based on the criteria outlined above, and analyzed them for effects of language, transition type, and age with two mixed-effects logistic regressions \citep{lme4, R}.
-We built one model each for children and adults.
-We modeled children and adults separately because effects of age are only pertinent to the children's data.
-
-The child model included condition (English vs.
-non-English)\footnote{Because each non-English language was represented by a single stimulus, we cannot treat individual languages as factors.
-Gaze behavior might be best for non-native languages that have the most structural overlap with participants' native language: English speakers can make predictions about the strength of upcoming Swedish prosodic boundaries nearly as well as Swedish speakers do, but Chinese speakers are at a disadvantage in the same task \citep{carlson2005}.
-We would need multiple items from each of the languages to check for similarity effects of specific linguistic features.}, transition type (question vs.
-non-question), age (3, 4, 5; numeric; intercept as age=0), and duration of the inter-turn gap (seconds, e.g., 0.441) as predictors, with two-way interactions between gap duration and the other simple fixed effects (language condition, transition type, and age) and a three-way interaction between language condition, transition type, and age.
-We included the two-way interactions with gap duration in case the effect of inter-turn silence changes with age or linguistic cueing (e.g., if children older children rely less on silence as a cue).\footnote{We test these two-way interactions with gap duration in all of the models reported in this paper.
-Higher-order interactions with gap duration usually resulted in model non-convergence due to distributional sparsity when three or more predictor values were considered, so we did not include them.} We also included random effects of item (turn transition) and participant, with maximal random slopes of condition, transition type, and their interaction for participants \citep{barr2013}.\footnote{The models we report in this paper are all qualitatively unchanged by the exclusion of their random slopes.
-We have left the random slopes in because of minor participant-level variation in the predictors modeled.}
-
-The adult model included fixed effects of condition, transition type, and their interaction, plus two-way interactions between gap duration and the other simple fixed effects (language condition and transition type, as in the child model).
-The adult model also included random effects of item and participant with maximal random slopes of condition, transition type, and their interaction for participant.
-
-Children's anticipatory gaze switches showed effects of language condition (\textit{$\beta$}=-3.65, \textit{SE}=1.16, \textit{z}=-3.15, \textit{p}$<$.01) and transition type (\textit{$\beta$}=-2.95, \textit{SE}=1.13, \textit{z}=-2.61, \textit{p}$<$.01) with additional effects of an age-by-language condition interaction (\textit{$\beta$}=0.5, \textit{SE}=0.212, \textit{z}=2.35, \textit{p}$<$.05), a language condition-by-transition type interaction (\textit{$\beta$}=2.69, \textit{SE}=1.35, \textit{z}=1.99, \textit{p}$<$.05), and a transition type-gap duration interaction (\textit{$\beta$}=5.52, \textit{SE}=2.28, \textit{z}=2.42, \textit{p}$<$.05).
-There were no significant effects of age or gap duration alone (\textit{$\beta$}=-0.002, \textit{SE}=0.26, \textit{z}=-0.009, \textit{p}$=$.99 and \textit{$\beta$}=2.25, \textit{SE}=3.19, \textit{z}=0.7, \textit{p}$=$.48, respectively).
-
-Adults' anticipatory gaze switches showed an effect of transition type (\textit{$\beta$}=-3.3, \textit{SE}=0.93, \textit{z}=-3.54, \textit{p}$<$.001) and significant interactions between language condition and transition type (\textit{$\beta$}=1.23, \textit{SE}=0.63, \textit{z}=1.96, \textit{p}$<$.05) and transition type and gap duration (\textit{$\beta$}=7.12, \textit{SE}=2.2, \textit{z}=3.24, \textit{p}$<$.01).
-There were no significant effects of language condition or gap duration alone (\textit{$\beta$}=-0.06, \textit{SE}=0.75, \textit{z}=-0.08, \textit{p}$=$.94 and \textit{$\beta$}=0.13, \textit{SE}=1.77, \textit{z}=0.08, \textit{p}$=$.94, respectively).
-
-\subsubsection*{Random baseline comparison}
-\label{sec:randbaseline1}
-
-Our primary analysis (above) makes the assumption that participants' eye movements generally follow the turn structure of the stimulus, i.e., that participants track the current speaker and switch their gaze to the upcoming speaker near turn transitions.
-As just described, based on this assumption, we used linear mixed effects regressions to see how anticipatory looking is affected by aspects of participant group (e.g., age) and stimulus (e.g., transition type, language condition).
-But what if the assumption that participants generally track turn structure were wrong?
-Could these results have emerged if participants' eye movements were \textit{not} linked to turn structure?
-For example, if participants were randomly looking back and forth between the two speakers, we might still find some anticipatory switching by chance.
-To test whether our primary results (the regression output above) could have arisen from random switching we conducted a secondary analysis comparing participants' anticipatory gaze at real and randomly shuffled points of turn transition.
-
-We conducted this analysis by running the same regression models on participants' eye-tracking data, only this time calculating their anticipatory gaze switches with respect to randomly permuted turn transition windows.
-This process involved: (1) randomizing the order and temporal placement of the analysis windows within each stimulus (Figure \ref{fig:shuffling}; ``analysis window'' is as shown in Figure \ref{fig:criterion}) to randomly redistribute the analysis windows across the eye-tracking signal, (2) re-running each participant's eye tracking data through switch identification (described above) on each of the randomly permuted analysis windows, and (3) modeling the anticipatory switches from the randomly permuted data (our random baseline dataset) with the same statistical models we used for the original dataset (Table \ref{tab:E1-models}).
-Importantly, although the onset time of each transition was shuffled within the eye-tracking signal, the other intrinsic properties of each turn transition (e.g., prior speaker identity, transition type, gap duration, language condition, etc.) stayed constant across each permutation.
-
-\begin{figure}[t!]
-\begin{center}
-\includegraphics[width=0.8\textwidth]{figures/FIG-ShuffledWindows.png}
-\end{center}
-\caption{Example of analysis window permutations for a stimulus with five turn transitions. The windows included $\pm$300 msec around the inter-turn gap.}
-\label{fig:shuffling}
-\end{figure}
-
-The random shuffling procedure de-links participants' gaze data from the turn structure in the original stimulus, thereby allowing us to compare turn-related (original) and non-turn-related (randomly permuted) looking behavior using the same eye movement data.
-We created 5,000 permutations of the original turn transitions, thereby creating 5,000 anticipatory gaze datasets with randomly de-linked gaze data.
-Because the randomly shuffled turn transitions could occur anywhere in the stimulus (so long as they didn't overlap each other within a single iteration), the resulting turn-transition windows collectively covered the entire stimulus---during speech and silence, during speaker change and speaker continuation, and during all turn transitions in the stimulus, even those excluded in the original analyses (e.g., because they were overlapped).
-This technique crucially differs from that used by Keitel and colleagues (2013, 2015), which tests anticipatory gaze at turn transitions against anticipatory gaze during speech.
-Pooled together, our 5,000 anticipatory gaze datasets yielded an average anticipatory switch rate for each participant over all possible starting points in the stimuli: a random baseline.
-Using this technique we compared participants' anticipatory switches at turn transition windows to their anticipatory switches over the stimulus as a whole.
-If participants looked randomly back and forth between the speakers, we would have seen similar patterns in both cases.
-
-Rather than simply comparing participants' overall anticipatory switch rates with real and random transition windows, we estimated the likelihood that each of the predictor effects in the original data (e.g., the effect of language condition; Table \ref{tab:E1-models}) could have arisen with random gaze switching: we ran identical statistical models on the real and randomly permuted data sets.
-This tells us not only whether participants' switches were above chance, but whether the specific underlying effects of their anticipatory gaze patterns (e.g., the effect of language condition) were above that expected by chance.
-Because these analyses are complex and secondary to the main results, we report their full details in \ref{sec:permutation}.
-
-Our baseline analyses revealed that none of the significant predictors from models of the original, turn-related data can be explained by random looking.
-For the children's data, the original \textit{z}-values for language condition, transition type, the age-language condition interaction, the transition type-gap duration interaction, and the language condition-transition type interaction were all greater than 95\% of \textit{z}-values from models of the randomly permuted data (99.3\%, 99.1\%, 98.9\%, 97\%, and 96\%, respectively, all \textit{p}$<$.05).
-Similarly, the adults' data showed significant differentiation from the randomly permuted data for all three significant predictors from the real transition dataset.
-Transition type, the interaction between transition type and gap duration, and the interaction between language condition and transition type showed \textit{z}-values that exceeded 100\%, 99.8\%, and 95\% of random \textit{z}-values, respectively (all \textit{p}$\leq$.05).
-See \ref{sec:permutation} for more information on each predictor's random permutation distribution.\footnote{This baseline analysis tests ``random looking'' against ''turn-driven looking'', but it does not test subtypes of turn-driven looking.
-For example, children might switch their gaze from the current speaker to the addressee out of boredom with the ongoing speech rather than from active anticipation of an upcoming response.
-We address this hypothesis about ``boredom'' gaze switches vs.
-``turn-transition'' gaze switches in \ref{sec:boredlooks}}
-
-\subsubsection*{Developmental effects}
-
-The models reported above revealed a significant interaction of age and language condition (Table \ref{tab:E1-models}) that was unlikely be due to random gaze switching (Figure \ref{fig:E1-randvsreal}).
-To further explore this effect, we compared the effect of language condition across age groups: using the permuted datasets described above, we extracted the average difference score for the two language conditions (English minus non-English) for each participant, computing an overall average for each random permutation of the data.
-Then, within each permutation, we made pairwise comparisons of the average difference scores across participant age groups.
-This process yielded a distribution of random permutation-based difference scores that we could then compare to the difference score in the actual data.
-Details are given in \ref{sec:pairwisedev}.
-
-These analyses revealed that, while 3- and 4-year olds showed similarly-sized effects of language condition, 5-year-olds had a significantly smaller effect of language condition, compared to both younger age groups.
-The difference in the language condition effect between 5-year-olds and 3-year-olds was greater than would be expected by chance (99.52\% of the randomly permuted data sets; \textit{p}$<$.01).
-Similarly, the difference in the language condition effect between 5-year-olds and 4-year-olds was greater than would be expected by chance (99.96\% of the data sets; \textit{p}$<$.001).
-See Figure \ref{fig:E1-lgageinteraction} for each difference score distribution.
-
-When does spontaneous turn prediction emerge developmentally?
-We tested whether the youngest age group (3-year-olds) already exceeded chance in their anticipatory gaze switches by comparing children's real gaze rates to the random baseline in the English condition with two-tailed \textit{t}-tests.
-We used the English condition because we are most interested in finding out when children begin to make spontaneous turn predictions for natural speech.
-We found that three-year-olds made anticipatory gaze switches significantly above chance, when all transitions were considered (\textit{t}(22.824)$=$-4.147, \textit{p}$<$.001) as well as for question transitions alone (\textit{t}(21.677)$=$-5.268, \textit{p}$<$.001).
-
-\subsection*{Discussion}
-\label{sec:discussion1}
-
-Children and adults spontaneously tracked the turn structure of the conversations, making anticipatory gaze switches at an above-chance rate across all ages and conditions.
-Children's anticipatory gaze rates were affected by language condition, transition type, age, and gap duration (Table \ref{tab:E1-models}), none of which could be explained by a baseline of random gaze switching (\ref{sec:permutation}; Figure \ref{fig:E1-ChiTs}).
-These data show a number of important features that bear on our questions of interest.
-
-First, both adults' and children's anticipations were strongly affected by transition type.
-Both groups made more anticipatory switches after hearing questions, compared to non-questions, especially for the English stimuli compared to the non-English stimuli.
-Overall, participants made few anticipatory switches after non-questions, even in the English videos when they had full linguistic access.
-Prior work using online, metalinguistic tasks has shown that participants can use linguistic cues to accurately predict upcoming turn ends \citep{bogelstorreira2015, magyari2012, de-ruiter2006}.
-The current results add a new dimension to our understanding of how listeners make predictions about turn ends: both children and adults spontaneously monitor the linguistic structure of unfolding turns for cues to imminent responses.
-
-Second, children made more anticipatory switches overall in English videos, compared to non-English videos.
-This effect suggests that linguistic access is important for children's ability to anticipate upcoming turn structure, consistent with prior work on turn-end prediction in adults (\citealp{de-ruiter2006}; \citealp{magyari2012}) and children \citep{keitel2013}.
-
-Third, we saw that older children made anticipatory switches more reliably than younger children, but only in the non-English videos.
-In the English videos, children anticipated well at all ages, especially after hearing questions.
-This interaction between age and language condition suggests that the 5-year-olds were able to leverage anticipatory cues in the non-English videos in a way that 3- and 4-year-olds could not, possibly by shifting more attention to the non-English prosodic or non-verbal cues.
-Prior work on children's turn-structure anticipation has proposed that children's turn-end predictions rely primarily on lexicosemantic structure (and not, e.g., prosody) as they get older \citep{keitel2013}.
-The current results suggest more flexibility in children's predictions; when they do not have access to lexical information, older children and adults find alternative cues to turn taking behavior.
-
-Finally, children and adults made more anticipatory switches in transitions with longer inter-turn gaps, though this effect was limited to non-question turns (Table \ref{tab:E1-models}).
-This finding suggests that gap duration indeed serves as a cue to upcoming turn structure; while short gaps may be perceived as within-turn pauses \citep{mannel2009}, long gaps could instead be indicative of between-turn pauses (where speaker transition occurs).
-Participants might use long silences to retroactively assign turn boundaries and anticipate speaker switches that were otherwise not anticipated (in this case, because the preceding turn was not a question).
-An alternative explanation for effects of gap duration is that longer inter-turn gaps result in longer analysis windows, which gives participants more time to make an anticipatory gaze.
-However, if participants are generally more likely to make a switch at question transitions (as our results suggest), and if question-driven switches aren't already at ceiling when gaps are short, we would expect that longer gaps would benefit questions more than non-questions---the opposite pattern from what the data show here.
-We take this as evidence that inter-turn silence may be most useful when participants have limited ability to make predictions about upcoming speaker transitions.
-
-In Experiment 2, we followed up on these findings, improving on two aspects of the design: first, our language manipulation in this first experiment was too coarse to provide data regarding specific linguistic information channels (e.g., the effect of prosodic information alone).
-In Experiment 2, we compared lexicosyntactic and prosodic cues with phonetically altered speech and used puppets to eliminate non-verbal cues to turn taking.
-Second, we were not able to pinpoint the emergence of anticipatory switching because the youngest age group in our sample was already able to make anticipatory switches at above-chance rates.
-In Experiment 2, we explored a wider developmental range.
-
-% Adults' anticipatory gaze rates were also affected by transition type, language condition, and gap duration (Table \ref{tab:E1-models}), none of which could be easily explained by a baseline of random gaze switching (Figure \ref{fig:E1-AduTs}).
-Like children, adults made more anticipatory switches after hearing questions compared to non-questions, suggesting that anticipation mattered more to them when an immediate response was expected.
-Also like children, the advantage for questions was driven % by lexical access such that adults must have relied on lexicosyntactic cues to questionhood in picking out turns that potentially require an immediate response, though this effect was only marginally divergent from the distribution of randomly permuted data (\textit{p}$=$.053; Figure \ref{fig:E1-AduTs}).
-Finally, adults' anticipation rates were also affected by gap duration, but more so for questions than non-questions (Table \ref{tab:E1-models}), suggesting that adults were less likely overall to make switches at non-questions, and so did not benefit from extra time to do so.
-
-% Children and adults' predictions alike were benefited by access to lexical information (English) and speech act status (questionhood), suggesting that linguistic cues, particularly lexical ones, facilitate their spontaneous predictions about upcoming turn structure through the identification of turns with immediate responses.
-Children's anticipatory gaze rates for questions and non-questions in English was stable across ages and comparable to adult behavior (Figure \ref{fig:E1-randvsreal}), suggesting that they can identify questions in native stimuli with adult-like competence by age three.
-Although participants' ability to recognize questions was facilitated by lexical access (i.e., English vs.
-non-English), the prosody in the non-English videos was non-native, and so the experimental design can not conclusively show which linguistic cues children relied on in the English videos to identify question turns.
-Relatedly, though lexical access clearly facilitated participants' anticipatory gaze rate, it was not necessary for participants---especially adults---in order to exceed chance switching rates (Figure \ref{fig:E1-randvsreal}), suggesting that participants use non-lexical cues (e.g., prosody, non-verbal behavior) to make anticipatory eye movements at least some of the time.
-
-\section*{Experiment 2}
-\label{sec:exp2}
-
-Experiment 2 used English-only stimuli, controlled for lexical and prosodic information, eliminated non-verbal cues, and tested children from a wider age range.
-To tease apart the role of lexical and prosodic information, we phonetically manipulated the speech signal for pitch, syllable duration, and lexical access.
-By testing 1- to 6-year-olds we hoped to find the developmental onset of turn-predictive gaze.
-We also hoped to measure changes in the relative roles of prosody and lexicosyntax across development.
-
-Non-verbal gestural cues in Experiment 1 could have helped participants make predictions about upcoming turn structure \citep{rossano2009, stivers2010}.
-Since our focus here is on linguistic cues, we eliminated all gaze and gestural signals in Experiment 2 by replacing the videos of human actors with videos of puppets.
-Puppets are less realistic and expressive than human actors, but they create a natural context for having somewhat motionless talkers in the videos.
-Additionally, the prosody-controlled condition (described below) included small but global changes to syllable duration that would have required complex video manipulation or precise re-enactment with human talkers, neither of which was feasible.
-For these reasons, we decided to use puppet videos rather than human videos in the final stimuli.
-As in the first experiment, we recorded participants' eye movements as they watched six short videos of dyadic conversation, and then analyzed their anticipatory glances from the current speaker to the upcoming speaker at points of turn transition.
-
-\subsection*{Methods}
-\label{sec:methods2}
-
-\subsubsection*{Participants}
-We recruited 27 undergraduate adults and 129 children ages 1;0--6;11 to participate in our experiment.
-Adult participants were recruited again via the Stanford University Psychology participant database and were either paid or received course credit for their time.
-We recruited our child participants from the Children's Discovery Museum in San Jose, California\footnote{We ran Experiment 2 at a local children's museum because it gave us access to children with a wider range of ages.
-Participants were volunteers.}, targeting approximately 20 children for each of the six one-year age groups (range: 20--23).
-All participants were native English speakers, though some parents (N$=$27) reported that their child heard a second (and sometimes third) language at home.
-None of the adult participants reported fluency in a second language.
-
-\begin{figure}[t]
-\begin{center}
-\includegraphics[width=\textwidth]{figures/FIG-EN-stim.png}
-\end{center}
-\caption{The six puppet pairs (and associated audio conditions). Each pair was linked to three distinct conversations from the same condition across the three experiment versions.}
-\label{fig:puppets}
-\end{figure}
-
-\subsubsection*{Materials}
-We created 18 short videos of improvised, child-friendly conversation (Figure \ref{fig:puppets}).
-To eliminate non-verbal cues to turn transition and to control the types of linguistic information available in the stimuli we first audio-recorded improvised conversations, then phonetically manipulated those recordings to limit the availability of prosodic and lexical information, and finally recorded video to accompany the manipulated audio, featuring puppets as talkers.
-
-\textit{Audio recordings}.
-The recording session was set up in the same way as the first experiment, but with a shorter warm up period (5--10 minutes) and a pre-determined topic for the child-friendly improvisation (`riding bikes', `pets', `breakfast', `birthday cake', `rainy days', or `the library').
-All of the talkers were native English speakers, and were recorded in male-female pairs.
-As before, we asked talkers to speak ``as if they were on a children's television show'' and to ask at least a few questions during the improvisation.
-We cut each audio recording down to the (approximate) 20-second interval with the most turn activity.
-The 20-second clips were then phonetically manipulated and used in the final video stimuli.
-
-\textit{Audio Manipulation}.
-We created four versions of each audio conversation: \textit{normal}, \textit{words only}, \textit{prosody only}, and \textit{no speech}.
-That is, one version with a full linguistic signal (\textit{normal}), and three with incomplete linguistic information (hereafter ``partial cue'' conditions).
-The \textit{normal} conversations were the unmanipulated, original audio clips.
-
-The \textit{words only} conversations were manipulated to have robot-like speech: we flattened the intonation contours to each talker's average pitch (F$_{0}$) and we reset the duration of every nucleus and coda to each talker's average nucleus and coda duration.\footnote{We excluded hyper-lengthened words like [w\textipa{aU:}] `woooow!'.} We made duration and pitch manipulations using PSOLA resynthesis in Praat \citep{Praat}.
-Thus, the \textit{words only} versions of the conversations had no pitch or durational cues to upcoming turn boundaries, but did have intact lexicosyntactic cues (and some residual phonetic correlates of prosody, e.g., intensity).
-
-We created the \textit{prosody only} conversations by low-pass filtering the original recording at 500 Hz with a 50 Hz Hanning window (following de Ruiter et al., 2006).
-This manipulation creates a ``muffled speech'' effect because low-pass filtering removes most of the phonetic information used to distinguish between phonemes.
-The \textit{prosody only} versions of the conversations lacked lexical information, but retained their intonational and rhythmic cues to upcoming turn boundaries.
-
-The \textit{no speech} condition served as a non-linguistic baseline.
-For this condition, we replaced the original audio clip for the conversation with multi-talker babble: we overlaid multiple child-oriented conversations (excluding the original one), and then cropped the result to the duration of the original conversation clip.
-Thus, the \textit{no speech} conversation lacked any linguistic information to upcoming turn boundaries---the only cue to turn taking was the opening and closing of the puppets' mouths.
-
-Finally, because low-pass filtering removes significant acoustic energy, the \textit{prosody only} conversations were much quieter than the other three conditions.
-Our last step was to downscale the intensity of the audio tracks in the three other conditions to match the volume of the \textit{prosody only} clips.
-We referred to the conditions as ``normal'', ``robot'', ``mermaid'', and ``birthday party'' speech when interacting with participants.
-
-\textit{Video recordings}.
-We created puppet video recordings to match the manipulated 20-second audio clips.
-The puppets were minimally expressive; the puppeteer could only control the opening and closing of their mouths, and the puppets' heads, eyes, arms, and bodies stayed still.
-Puppets were positioned side-by-side, looking in the same direction to eliminate shared gaze as a cue to turn structure \citep{thorgrimsson2015}.
-We took care to match the puppets' mouth movements to the syllable onsets as closely as possible, specifically avoiding mouth movement before the onset of a turn.
-We then added the manipulated audio clips to the puppet video recordings by hand with video editing software.
-
-We used three pairs of puppets for the \textit{normal} condition---`red', `blue' and `yellow'---and one pair of puppets for each partial cue condition: `robots', `merpeople', and `party-goers' (Figure \ref{fig:puppets}).
-We randomly assigned half of the conversation topics (`birthday cake', `pets', and `breakfast') to the \textit{normal} condition, and half to the partial cue conditions (`riding bikes', `rainy days', and `the library').
-We then created three versions of the experiment, so that each of the six puppet pairs was associated with three different conversation topics across the different versions of the experiment (18 videos in total; 6 videos per experiment version).
-We ensured that the position of the talkers (left and right) was counterbalanced in each version by flipping the video and audio channels as needed.
-
-As before, the duration of turn transitions and the number of speaker changes across videos was variable because the conversations were recorded semi-spontaneously.
-We measured turn transitions from the audio signal of the \textit{normal}, \textit{words only}, and \textit{prosody only} conditions.
-There was no audio from the original conversation in the \textit{no speech} condition videos, so we measured turn transitions from puppets' mouth movements in the video signal, using ELAN video annotation software \citep{ELAN}.
-
-There were 85 turn transitions for analysis after excluding transitions longer than 550 msec and shorter than 90 msec.
-The remaining turn transitions had more questions than non-questions (N$=$47 and N$=$38, respectively), with transitions distributed somewhat evenly across conditions, keeping in mind that there were three \textit{normal} videos and only one video for each partial cue condition in each experiment version: \textit{normal} (N$=$36), \textit{words only} (N$=$13), \textit{prosody only} (N$=$17), and \textit{no speech} (N$=$19).
-Inter-turn gaps for questions (mean$=$366, median$=$438, stdev$=$138 msec) were longer than those for non-questions (mean$=$305, median$=$325, stdev$=$94 msec) on average, but gap duration was overall comparable across conditions: \textit{normal} (mean$=$334, median$=$321, stdev$=$130 msec), \textit{words only} (mean$=$347, median$=$369, stdev$=$ 115 msec), \textit{prosody only} (mean$=$365, median$=$369, stdev$=$104 msec), and \textit{no words} (mean$=$319, median$=$329, stdev$=$136 msec).
-
-\subsection*{Procedure}
-We used the same experimental apparatus and procedure as in the first experiment.
-Each participant watched six puppet videos in random order, with 15--30 second filler videos placed in-between (e.g., running puppies, moving balls, flying bugs).
-Three of the puppet videos had \textit{normal} audio while the other three had \textit{words only}, \textit{prosody only}, and \textit{no speech} audio.
-As before, the experimenter immediately began each session with calibration and then stimulus presentation.
-Participants were given no instruction about how to watch the videos or what their purpose was, they were simply encouraged to watch the ``(fun/nice) puppet videos''.
-The entire experiment took less than five minutes.
-
-\linespread{1}
-\begin{table}[t]
-\begin{center}
-  \begin{tabular}{llcccc}
-    \hline
-    Age group & Speaker & Addressee & Other onscreen & Offscreen\\
-    \hline
-    1 & 0.44 & 0.14 & 0.23 & 0.19 \\
-    2 & 0.50 & 0.13 & 0.24 & 0.14 \\
-    3 & 0.47 & 0.12 & 0.25 & 0.16 \\
-    4 & 0.48 & 0.11 & 0.29 & 0.12 \\
-    5 & 0.54 & 0.11 & 0.20 & 0.14 \\
-    6 & 0.60 & 0.12 & 0.18 & 0.10 \\
-    Adult & 0.69 & 0.12 & 0.09 & 0.10 \\
-%   Overall & 0.53 & 0.12 & 0.21 & 0.13 \\
-    \hline
-  \end{tabular}
-\end{center}
-  \caption{Average proportion of gaze to the current speaker and addressee during periods of talk across ages in Experiment 2.}
-\label{tab:look_e2}
-\end{table}
-%\linespread{2}
-
-\linespread{1}
-\begin{table}
-\begin{center}
-  \begin{tabular}{llcccc}
-    \hline
-    Condition & Speaker & Addressee & Other onscreen & Offscreen\\
-    \hline
-    Normal 			& 0.58 & 0.12 & 0.17 & 0.13 \\
-    Words only 		& 0.54 & 0.11 & 0.24 & 0.10 \\
-    Prosody only 	& 0.48 & 0.12 & 0.26 & 0.15 \\
-    No speech 		& 0.44 & 0.13 & 0.26 & 0.18 \\
-    \hline
-  \end{tabular}
-\end{center}
-  \caption{Average proportion of gaze to the current speaker and addressee during periods of talk across conditions in Experiment 2.}
-\label{tab:look_e2b}
-\end{table}
-%\linespread{2}
-
-% Average prop gaze to the current speaker across conditions for one-year-olds
-%                           speaker addressee other-screen offscreen
-%Normal   		  0.4737582 0.1366432    0.2003952 0.1892035
-%Words only    0.3977261 0.1283417    0.3112005 0.1627317
-%Prosody only 0.4290202 0.1611453    0.2558657 0.1539688
-%No speech    0.3786925 0.1228615    0.2356423 0.2628037
-
-\subsubsection*{Data preparation and coding}
-We coded each turn transition for its linguistic condition (\textit{normal}, \textit{words only}, \textit{prosody only}, and \textit{no speech}) and transition type (question/non-question)\footnote{We coded \textit{wh-}questions as ``non-questions'' for the \textit{prosody only} videos.
-Polar questions often have a final rising intonational contour, but \textit{wh-}questions do not  \citep{hedberg2010}.}, and identified anticipatory gaze switches to the upcoming speaker using the methods from Experiment 1.
-
-\begin{figure}[!ht]
-\begin{center}
-\includegraphics[width=\textwidth]{figures/E2-samples-by-lang-groups-trans-types.png}
-\end{center}
-\caption{Anticipatory gaze rates across language condition and transition type for the real and randomly permuted datasets. Vertical bars represent 95\% confidence intervals.}
-\label{fig:E2-randvsreal}
-\end{figure}
-
-\subsection*{Results}
-\label{sec:results2}
-
-Participants' pattern of gaze indicated that they performed basic turn tracking across all ages and in all conditions.
-Participants looked at the screen most of the time during video playback (82\% and 86\% average for children and adults, respectively), primarily looking at the person who was currently speaking (Tables \ref{tab:look_e2} and \ref{tab:look_e2b}).
-They tracked the current speaker in every condition---even one-year-olds looked more at the current speaker than at anything else in the three partial cue conditions (40\% for \textit{words only}, 43\% for \textit{prosody only}, and 39\% for \textit{no speech}).
-There was a steady overall increase in looks to the current speaker with age and added linguistic information (Tables \ref{tab:look_e2} and \ref{tab:look_e2b}).
-Looks to the addressee also decreased with age, but the change was minimal.
-Figure \ref{fig:E2-randvsreal} shows participants' anticipatory gaze rates across age, the four language conditions, and transition type.
-
-\subsubsection*{Statistical models}
-\label{sec:models2}
-
-We identified anticipatory gaze switches for all 85 usable turn transitions, and analyzed them for effects of language condition, transition type, and age with two mixed-effects logistic regressions.
-We again built separate models for children and adults because effects of age were only pertinent to the children's data.
-The child model included condition (\textit{normal}/\textit{prosody only}/\textit{words only}/\textit{no speech}; with \textit{no speech} as the reference level), transition type (question vs.
-non-question), age (1, 2, 3, 4, 5, 6; numeric, intercept as age=0), and duration of the inter-turn gap (in seconds) as predictors, with full interactions between language condition, transition type, and age and two-way interactions between gap duration and the other basic fixed effects (age, linguistic condition, and transition type).
-We also included random effects of participant and item (turn transition), with maximal random slopes of transition type for participant.
-The adult model included condition, transition type, their interactions, gap duration, and two-way interactions between gap duration and condition and transition type, with participant and item as random effects and maximal random slopes of condition and transition type for participant.
-
-Children's anticipatory gaze switches showed an effect of gap duration (\textit{$\beta$}=3.85, \textit{SE}=1.73, \textit{z}=2.22, \textit{p}$<$.05), a two-way interaction of age and language condition (for \textit{prosody only} speech compared to the \textit{no speech} reference level; \textit{$\beta$}=0.38, \textit{SE}=0.19, \textit{z}=1.97, \textit{p}$<$.05), a marginal two-way interaction of language condition and gap duration (for \textit{prosody only} speech compared to the \textit{no speech} reference level; \textit{$\beta$}=-4.77, \textit{SE}=2.63, \textit{z}=-1.82, \textit{p}$=$.07), and a three-way interaction of age, transition type, and language condition (for \textit{normal} speech compared to the \textit{no speech} reference level; \textit{$\beta$}=-0.35, \textit{SE}=0.17, \textit{z}=-2.05, \textit{p}$<$.05).
-There were no significant effects of age or transition type alone (Table \ref{tab:E2-chimodels}; \textit{$\beta$}=-0.05, \textit{SE}=0.14, \textit{z}=-0.38, \textit{p}$=$.7 and \textit{$\beta$}=-1.22, \textit{SE}=0.96, \textit{z}=-1.27, \textit{p}$=$.2, respectively)
-
-Adults' anticipatory gaze switches showed a significant effect of language condition (for \textit{words only} speech compared to the \textit{no speech} reference level; \textit{$\beta$}=3.79, \textit{SE}=1.62, \textit{z}=2.34, \textit{p}$<$.05) and a marginal two-way interaction between language condition and transition type (for \textit{words only} speech compared to the \textit{no speech} reference level; \textit{$\beta$}=-1.68, \textit{SE}=0.89, \textit{z}=-1.89, \textit{p}$=$.06).
-There was no significant effect of transition type alone (Table \ref{tab:E2-adumodels}; \textit{$\beta$}=-0.02, \textit{SE}=1.44, \textit{z}=-0.02, \textit{p}$=$.99).
-
-\subsubsection*{Random baseline comparison}
-\label{sec:randbaseline2}
-
-Using the same technique described in Experiment 1, we created and modeled random permutations of participants' anticipatory gaze switches.
-These analyses revealed that the significant predictors from models of the original, turn-related data were unlikely to be explained by random looking.
-In the children's data, the original model's \textit{z}-values for gap duration, the two-way interaction of age and language condition (\textit{prosody only}) and the three-way interaction of age, transition type, and language condition (\textit{normal} speech) were all greater than 93\% of the randomly permuted \textit{z}-values (95.6\%, 94\%, and 93.3\%, respectively, \textit{p}$=$.04, .06, and .07).
-Similarly, the adults' data showed significant differentiation from the randomly permuted data for the effect of language condition (\textit{words only} speech; greater than 98.3\% of random \textit{z}-values, \textit{p}$<$.02).
-See \ref{sec:permutation} for more information on each predictor's random permutation distribution.
-
-\linespread{1}
-\begin{minipage}[t]{0.95\linewidth}
-\begin{footnotesize}
-\begin{longtable}{lcccl}
-  \textbf{\textit{Children}} &&&& \\
-    \hline
-           &  Estimate & Std. Error & \textit{z} value & Pr($>$$|$\textit{z}$|$) \\
-    \hline
-    (Intercept)											& -3.452		& 0.76	& -4.543	& 5.55e-06 ***	\\
-    Age														& -0.054		& 0.143	& -0.379	& 0.705				\\
-    TType$=$\textit{non-Question}				& -1.217		& 0.958	& -1.27		& 0.204				\\
-    GapDuration											&  3.852		& 1.735	&  2.221	& 0.026 *			\\
-    Age*TType$=$\textit{non-Question}		&  0.152		& 0.141	&  1.081	& 0.28				\\
-    Age*GapDuration									&  0.214		& 0.266	&  0.805	& 0.421				\\
-    TType$=$\textit{non-Question}*			&  0.995		& 2.134	&  0.466	& 0.641				\\
-    \hspace*{5mm} GapDuration &&&& \\
-	\hline
-    Condition$=$\textit{normal}					&  0.54		& 0.742	&  0.728	& 0.467					\\
-    Age*Condition$=$\textit{normal}			&  0.125		& 0.103	&  1.221	& 0.222				\\
-    Condition$=$\textit{normal}*				&  0.908		& 0.748	&  1.215	& 0.224				\\
-    \hspace*{5mm} TType$=$\textit{non-Question} &&&& \\
-    Age*Condition$=$\textit{normal}*			& -0.355		& 0.173	& -2.051	& 0.04 *			\\
-    \hspace*{5mm} TType$=$\textit{non-Question} &&&& \\
-    Condition$=$\textit{normal}*				& -0.431		& 1.67	& -0.258	& 0.797				\\
-    \hspace*{5mm} GapDuration &&&& \\
-	\hline
-    Condition$=$\textit{prosody}				&  0.549		& 1.452	&  0.378	& 0.705				\\
-    Age*Condition$=$\textit{prosody}			&  0.375		& 0.191	&  1.967	& 0.049 *			\\
-    Condition$=$\textit{prosody}*				&  1.076		& 1.105	&  0.974	& 0.33				\\
-    \hspace*{5mm} TType$=$\textit{non-Question} &&&& \\
-    Age*Condition$=$\textit{prosody}*		& -0.296		& 0.235	& -1.257	& 0.209				\\
-    \hspace*{5mm} TType$=$\textit{non-Question} &&&& \\
-    Condition$=$\textit{prosody}*				& -4.767		& 2.625	& -1.816	& 0.069 (.)			\\
-    \hspace*{5mm} GapDuration &&&& \\
-	\hline
-    Condition$=$\textit{words}				&  0.684		& 1.06	&  0.645	& 0.519					\\
-    Age*Condition$=$\textit{words}		&  0.127		& 0.136	&  0.934	& 0.350					\\
-    Condition$=$\textit{words}*				& -1.244		& 1.031	& -1.207	& 0.228					\\
-    \hspace*{5mm} TType$=$\textit{non-Question} &&&& \\
-    Age*Condition$=$\textit{words}*		&  0.111		& 0.225	&  0.495	& 0.621					\\
-    \hspace*{5mm} TType$=$\textit{non-Question} &&&& \\
-    Condition$=$\textit{words}*				& -2.285		& 2.232	& -1.024	& 0.306					\\
-    \hspace*{5mm} GapDuration &&&& \\
-    \hline
-    &&&& \\
-  \caption{Model output for children's anticipatory gaze switches in Experiment 2.}
-\label{tab:E2-chimodels}
-\end{longtable}
-\end{footnotesize}
-\end{minipage}
-%\linespread{2}
-
-\linespread{1}
-\begin{minipage}[t]{0.95\linewidth}
-\begin{footnotesize}
-\begin{longtable}{lcccl}
-  \textbf{\textit{Adults}} &&&& \\
-    \hline
-           &  Estimate & Std. Error & \textit{z} value & Pr($>$$|$\textit{z}$|$) \\
-    \hline
-    (Intercept)										& -3.117		& 1.176	& -2.649	& 0.008 **				\\
-    TType$=$\textit{non-Question}			& -0.022		& 1.44	& -0.015	& 0.988					\\
-    GapDuration										&  4.073		& 2.947	&  1.382	& 0.167					\\
-    TType$=$\textit{non-Question}*		&  1.304		& 3.859	&  0.338	& 0.735					\\
-    \hspace*{5mm} GapDuration &&&& \\
-	\hline
-    Condition$=$\textit{normal}				&  0.39		& 1.316	&  0.296	& 0.767						\\
-    Condition$=$\textit{normal}*			& -0.709		& 0.754	& -0.94		& 0.347					\\
-   \hspace*{5mm} TType$=$\textit{non-Question} &&&& \\
-    Condition$=$\textit{normal}*			&  2.1			& 3.336	&  0.629	& 0.529					\\
-    \hspace*{5mm} GapDuration &&&& \\
-	\hline
-    Condition$=$\textit{prosody}			&  0.757		& 2.193	&  0.345	& 0.73					\\
-    Condition$=$\textit{prosody}*			&  0.386		& 1.065	&  0.362	& 0.717					\\
-    \hspace*{5mm} TType$=$\textit{non-Question} &&&& \\
-    Condition$=$\textit{prosody}*			& -1.118		& 4.543	& -0.246	& 0.805					\\
-    \hspace*{5mm} GapDuration &&&& \\
-	\hline
-    Condition$=$\textit{words}				&  3.792		& 1.621	&  2.338	& 0.019 *				\\
-    Condition$=$\textit{words}*				& -1.678		& 0.889	& -1.888	& 0.059 (.)				\\
-    \hspace*{5mm} TType$=$\textit{non-Question} &&&& \\
-    Condition$=$\textit{words}*				& -5.653		& 3.861	& -1.464	& 0.143					\\
-    \hspace*{5mm} GapDuration &&&& \\
-    \hline
-    &&&& \\
-  \caption{Model output for adults' anticipatory gaze switches in Experiment 2.}
-\label{tab:E2-adumodels}
-\end{longtable}
-\end{footnotesize}
-\end{minipage}
-%\linespread{2}
-
-\subsubsection*{Developmental effects}
-
-Our main goal in extending the age range to 1- and 2-year-olds in Experiment 2 was to find the age of emergence for spontaneous predictions about upcoming turn structure.
-As in Experiment 1, we used two-tailed \textit{t}-tests to compare children's real gaze rates to the random baseline rates in the \textit{normal} speech condition (in which the speech stimulus is most like what children hear every day).
-We tested real gaze rates against baseline rates for three age groups: one-, two-, and three-year-olds.
-Two- and three-year-old children made anticipatory gaze switches significantly above chance both when all transitions were considered (2-year-olds: \textit{t}(26.193)$=$-4.137, \textit{p}$<$.001; 3-year-olds: \textit{t}(22.757)$=$-2.662, \textit{p}$<$.05) and for question transitions alone (2-year-olds: \textit{t}(25.345)$=$-4.269, \textit{p}$<$.001; 3-year-olds: \textit{t}(21.555)$=$-3.03, \textit{p}$<$.01).
-One-year-olds, however, only made anticipatory gaze shifts marginally above chance for turn transitions overall and for question turns alone (overall: \textit{t}(24.784)$=$-2.049, \textit{p}$=$.051; questions: \textit{t}(25.009)$=$-2.03, \textit{p}$=$.053).
-
-We also tested the two baseline linguistic conditions against each other---\textit{no speech} and \textit{normal speech}---to find out when linguistic information made a difference in children's anticipations.
-Because, as we have seen, children primarily show linguistic effects in question-answer turn transitions, we investigated the use of linguistic cues across age by testing anticipation separately for question and non-question turns.
-Compared to the \textit{no speech} condition, children made significantly more anticipatory switches in the \textit{normal} speech condition for questions at ages 6, 4, and 3, and also marginally at age 2 (6-year-olds: \textit{t}(36.919)$=$3.8019, \textit{p}$<$.001; 4-year-olds: \textit{t}(41.449)$=$2.9777, \textit{p}$<$.01; 3-year-olds: \textit{t}(35.724)$=$2.4286, \textit{p}$<$.05; 2-year-olds: \textit{t}(41.078)$=$1.8018, \textit{p}$=$.079).
-Children's anticipatory switches for questions did not significantly differ in the \textit{no speech} and \textit{normal} speech conditions at ages 5 or 1 (5-year-olds: \textit{t}(29.406)$=$1.2783, \textit{p}$=$.211; 1-year-olds: \textit{t}(35.907)$=$0.4961, \textit{p}$=$.623).
-In contrast, children's anticipatory switch rates for non-question turns were not significantly different between the \textit{no speech} and \textit{normal} speech conditions at any age (all \textit{p}$>$0.09).
-Thus, consistent with the regression results, children were more likely to show an effect of linguistic content as they got older, but only for question transitions.
-
-The regression models for the children's data also revealed two significant interactions with age.
-The first was a significant interaction of age and language condition (for \textit{prosody only} compared to the \textit{no speech} reference level), suggesting a different age effect between the two linguistic conditions.
-As in Experiment 1, we explored each age interaction by extracting an average difference score over participants for the effect of language condition (\textit{no speech} vs.
-\textit{prosody only}) within each random permutation of the data, making pairwise comparisons between the six age groups.
-These tests revealed that children's anticipation in the \textit{prosody only} condition significantly improved at ages five and six compared to the \textit{no speech} baseline (with difference scores greater than 95\% of the random data scores; \textit{p}$<$.05).
-See Figure \ref{fig:E2-lgageinteraction} for these \textit{prosody only} difference score distributions.
-
-The second age-based interaction was a three-way interaction of age, transition type, and language condition (for \textit{normal} speech compared to the \textit{no speech} baseline).
-We again created pairwise comparisons of the average difference scores for the transition type-language condition interaction across age groups in each random permutation of the data, finding that the effect of transition type in the \textit{normal} speech condition became larger with age, with significant improvements by age 4 over ages 1 and 2 (99.9\% and 98.86\%, respectively), by age 5 over age 4 (97.54\%), and by age 6 over ages 1, 2, and 5 (99.5\%, 97.36\%, and 95.04\%), all significantly different from chance (\textit{p}$<$.05).
-See Figure \ref{fig:E2-lgagetypeinteraction} for these \textit{normal} speech difference score distributions.
-
-\subsection*{Discussion}
-\label{sec:discussion2}
-
-The core aims of Experiment 2 were to gain better traction on the individual roles of prosody and lexicosyntax in children's turn predictions, and to find the age of emergence for spontaneous turn anticipation.
-Many of our results replicate the findings from Experiment 1: participants often made more anticipatory switches when they had access to linguistic information and, when they did, tended to make more anticipatory switches for questions compared to non-questions.
-
-As in Experiment 1, children and adults spontaneously tracked the turn structure of the conversations.
-Participants made anticipatory gaze switches at above-chance rates starting at age two for both questions and non-questions.
-Longer gaps had a broader impact on participants' anticipations in this second experiment; we saw that, overall, longer inter-turn gaps resulted in more anticipatory switches, with the \textit{no speech} condition showing equal or stronger effects of gap duration than all other conditions.
-
-As before, participants made far more anticipations for questions than for non-question turns---at least for those two years old and older.
-But these effects were different for the conditions with partial linguistic information: \textit{prosody only} and \textit{words only}.
-In the \textit{prosody only} condition, performance was initially low for young children and increased significantly with age.
-In the \textit{words only} condition, children age two and older showed robust switching for questions (much like in \textit{normal} speech), but never rose above chance for non-question turns (Figure \ref{fig:E2-randvsreal}), with no significant differences from the \textit{no speech} baseline.
-These findings do not support an early role for prosody or lexical information alone in children's spontaneous predictions about turn structure.
-They also give no support for the idea that lexical information is sufficient on its own to support children's anticipatory switching.
-They do underscore the developing relationship between the online use of linguistic cues, inter-turn silence, and speech act in spontaneous predictions about upcoming turn structure.
-
-\section*{General Discussion}
-\label{sec:gendisc}
-
-Children begin to develop conversational turn-taking skills long before their first words emerge \citep{bateson1975, hilbrink2015, jaffe2001, snow1977}.
-As they become fast and knowledgeable language users, they also become able to make accurate predictions about upcoming turn structure.
-Until recently, we have had very little data on how children weave language into their already-existing turn-taking behaviors.
-In two experiments investigating children's anticipatory gaze to upcoming speakers, we found evidence that turn prediction develops early in childhood and that, when spontaneous predictions begin, they are primarily driven by participants' expectation of an immediate response in the next turn (e.g., after questions).
-In making predictions about upcoming turn structure, children used a combination of lexical and prosodic cues; neither signal alone was sufficient to support increased anticipatory gaze.
-We also found no early advantage for prosody over lexicosyntax; children's anticipatory switch rates in the \textit{prosody only} condition were initially low, but showed significant gains by age five.
-We discuss these findings with respect to the role of linguistic processing and inter-turn silence for predicting upcoming turn structure, the importance of questions in predictions about conversation, and children's developing competence as conversationalists.
-
-\subsection*{Predicting upcoming turn structure}
-
-Prior work with adults has found a consistent role for lexicosyntax in predicting upcoming turn structure \citep{de-ruiter2006, magyari2012}, whereas the role of prosody is still under debate \citep{duncan1972, ford1996, bogelstorreira2015}.
-Knowing that children comprehend more about prosody than lexicosyntax early on (see \citealp{speer2009} for a review), we thought it possible that young children would instead show an advantage for prosody in their predictions about turn structure in conversation.
-Our results suggest that, on the contrary, exclusively presenting prosodic information to children limits their spontaneous predictions about upcoming turn structure until age five.
-
-Thus, using prosody alone to accurately predict turn boundaries in conversation appears to be difficult for adults and children.
-Prosodic information is continuous, multidimensional, and can index multiple meanings at once---it encodes syntactic structure, speech act, and extralinguistic information without clear one-to-one mappings between form and meaning \citep{cutler1997, shriberg1998, lammertink2015}.
-For these reasons, prosodic information alone may not be enough for young children to easily make precise temporal predictions about turn structure, and identify question turns in unfolding speech.
-Therefore, although children show early facility with prosodic discrimination \citep{nazzi2003, soderstrom2003, johnson2001, jusczyk1995, morgan1995, mehler1988}, using prosodic knowledge for turn prediction may be difficult without additional information from lexical or syntactic cues.
-
-Our findings suggest that there is one prosodic cue that is an exception to this rule: inter-turn silence.
-Generally speaking, participants showed a greater anticipatory switches for longer inter-turn gaps, but the effect of inter-turn gap duration is strongest in our data when upcoming responses are less predictable, whether due to the asymmetrical response expectations for questions vs.
-non-questions (Experiment 1) or the lack of non-verbal cues and any linguistic information (Experiment 2).
-Notably, there were no significant interactions of gap duration with participant age.
-This pattern of results suggests that, when predictive information about upcoming responses is absent, long silences may increase participants' expectation for a speaker change and promote more anticipatory gaze switches.
-Pauses are detected and related to phrasal structure from early on; 5-month-old infants use pauses to parse intonational phrases \citep{mannel2009}.
-The lack of interactions between age and gap duration suggests that the use of inter-turn silence remains important for older speakers and the interactions between transition type and gap duration (Experiment 1) and condition and gap duration (Experiment 2; marginal), suggest that this effect is not simply the result of having more time to make a gaze switch.
-These findings thus suggest that silence is an early and lasting cue for identifying turn structure online when other predictive information is not adequate.
-
-Notably, many other non-linguistic cues encode information about transition type, including gaze and gesture.
-We did not systematically test those cues here but, like inter-turn silence, they may play a critical role in parsing and making predictions about turn structure when other linguistic information is not sufficient to make accurate predictions.
-
-%Because inter-turn silence works as a backwards-looking cue, it's predictive utility for speaker change is diminished compared to forward-looking linguistic cues (e.g., syntactic constituent completion).
-However, it is a strategy that can be usefully employed to predict upcoming responses some of the time, even without access to language.
-
-Perhaps surprisingly, we found no evidence that lexical information alone is equivalent to the full linguistic signal in driving children's predictions, as has been shown previously for adults \citep{magyari2012, de-ruiter2006} and as is replicated with adult participants in the current study.
-Unlike prosodic cues, lexicosyntactic cues are discreet and have much clearer form-to-meaning mappings, with clear lexicosyntactic cues to questionhood that occur early within turns (e.g., \textit{wh}-words, \textit{do}-insertion, and subject-auxiliary inversion).
-That said, children's lexical and syntactic knowledge is limited for quite some time (\citealp{tomasello1999}, but see also \citealp{bergelson2013, shi2010}).
-Although our stimuli were made in a child-friendly style, they are still other-directed and fairly complex, with 20--30 seconds of continuous conversational speech.
-
-It is perhaps for this reason that children's performance was always best with the full signal, where lexicosyntactic information was supported by prosodic information and vice versa.
-Even in adults, B\"{o}gels and Torreira (\citeyear{bogelstorreira2015}) showed that the trade-off in informativity between lexical and prosodic cues is more subtle in semi-natural speech.
-The present findings are the first to show evidence of a similar effect developmentally.
-
-%Finally, because most anticipatory switches were made following question turns, one could predict that the only lexical and prosodic cues that matter for spontaneous turn prediction are those that cue questionhood.
-Our linguistic manipulations here were focused on lexical and prosodic information globally, which may have affected lexical and prosodic cues to questionhood asymmetrically; while lexicosyntactic question cues were available on every instance of \textit{wh}- and \textit{yes/no} questions in our stimuli, prosodic question cues were only salient on \textit{yes/no} questions.
-
-\subsection*{The question effect}
-
-In both experiments, anticipatory looking was primarily driven by question transitions, a pattern that has not been previously reported in other anticipatory gaze studies, on children or adults (Keitel et al., 2013; Hirvenkari, 2013; Tice and Henetz, \citeyear{TiceHenetz11}).
-Questions make an upcoming speaker switch immediately relevant, helping the listener to predict with high certainty what will happen next (i.e., an answer from the addressee), and are often easily identifiable by overt prosodic and lexicosyntactic cues.
-
-Prior work on children's acquisition of questions indicates that they may already have some knowledge of question-answer sequences by the time they begin to speak: questions make up approximately one third of the utterances children hear, before and after the onset of speech, and even into their preschool years, though the type and complexity of questions changes throughout development \citep{casillas2016, fitneva2012, henning2005, shatz1979}.\footnote{There is substantial variation in question frequency by individual and socioeconomic class \citep{hart1992, weisleder2012}.} For the first few years, many of the questions directed to children are ``test'' questions---questions that the caregiver already has the answer to (e.g., ``What does a cat say?''), but this changes as children get older.
-Questions help caregivers to get their young children's attention and to ensure that information is in common ground, even if the responses are non-verbal or infelicitous \citep{bruner1985, fitneva2012, snow1977}.
-Moreover, because of their high frequency and relatively limited number of formats, questions, especially \textit{wh}-questions, may be more identifiable and predictable compared to other types of speech acts.
-So, in addition to having a special interactive status, questions are a frequent, predictable, and core characteristic of many caregiver-child interactions, motivating a general benefit for questions in turn structure anticipation.
-
-Two important routes for future work are then: (1) how does children's ability to monitor for questions in conversation relate to their prior experience with questions?
-and (2) what is it about questions that makes children and adults more likely to anticipatorily switch their gaze to addressees?
-If this ``question'' effect exists for all turns that require an immediate response (``adjacency pairs''; \citealp{schegloff2007}), other turn types, such as imperatives, compliments, and complaints should show similar patterns.
-If the effect is instead about overall predictability of the syntactic frame, children would instead show similar patterns for other frequent frames from child-directed speech (e.g., ``Look at the X''; \citealp{mintz2003}).
-The recognizability and predictability of syntactic frames is likely to play a role in turn prediction as children become more sophisticated language users, even if the effect is truly about adjacency pairs; for example, rhetorical and tag questions take a very similar form to prototypical polar questions, but usually do not require an answer.
-So, though it is clear that adults and children anticipate responses more often for questions than non-questions, we do not yet know whether their predictive action is limited to turns formatted as questions, turns with high recognizability and predictability, or turns that project an immediate response from the addressee.
-
-A question effect suggests that participants' spontaneous predictions may be driven by what lies \textit{beyond} the end of the current turn---not just by the upcoming end of the turn itself, as has been focused on in prior work \citep{bogelstorreira2015, keitel2013, magyari2012, de-ruiter2006}.
-In future work, it will be crucial to measure prediction from a first-person perspective to find out what kinds of predictions are most relevant to addressees in conversation.
-
-One possible scenario is that listeners in spontaneous, first-person conversation use multiple strategies to make predictions about upcoming turn structure: they could semi-passively attend to incoming speech for cues to upcoming speaker transition (e.g., questions and other adjacency pairs) and, when possible upcoming transition is detected, switch into a more precise turn-end prediction mode (\`{a} la \citealp{de-ruiter2006}).
-A flexible prediction system like this one allows listeners to continuously monitor ongoing conversation for turn-related cues at a low cost while still managing to plan their responses and come in quickly when needed.
-
-To test this hypothesis, we would need to look at prediction from a first-person perspective, which very little work so far has accomplished (present work included).
-Although third-party measures enable us to measure participants' predictions without any interference from language production, they also limit our knowledge about how the need to give a response might itself play an important role in addressees' prediction strategies.
-Recent work has shown that shifts in addressee gaze similar to those measured here indeed occur in spontaneous conversation \citep{holler2015}, but much more work is needed to determine how participants make predictions about turn structure in first-person contexts and whether those mechanisms shift at points of imminent speaker change.
-
-\subsection*{Early competence for turn taking?}
-
-One of the core aims of our study was to test whether children show an early competence for turn taking, as is proposed by studies of spontaneous mother-infant proto-conversation and theories about the mechanisms underlying human interaction in general \citep{hilbrink2015, levinson2006}.
-We found evidence that young children make spontaneous predictions about upcoming turn structure: definitely by age two and marginally by age one.
-
-These results contrast with Keitel and colleagues' \citeyearpar{keitel2013} finding that children cannot anticipate upcoming turn structure at above-chance rates until age three.
-The current study used an appreciably more conservative random baseline than the one used in Keitel and colleagues' study.
-Therefore, this difference in age of emergence more likely stems from our use of a more engaging speech style, stereo speech playback, and more typical turn transition durations.
-The child-friendly style of speech in particular may have helped in two ways: keeping children more engaged with the stimuli and using less syntactically complex and more prosodically exaggerated speech \citep{fernald1989, werker1989, snow1977} compared to what they would get with adult-adult conversation.
-
-To be clear, young children's ``above chance'' performance was often still far from adult-like predictive behavior---turn prediction (and the concurrent use of linguistic cues from unfolding speech) increased only gradually with age.
-Children at ages one and two were still very close to chance in their anticipations and, even at age six, children were not fully adult-like in their predictions.
-This indicates that young children may at first rely primarily on non-verbal cues, like inter-turn silence, to anticipate turn transitions but that, by adulthood, listeners use both verbal and non-verbal cues to make predictions.
-Relatedly, adult listeners may be more expert in flexibly adapting to the turn-relevant cues present at any moment, e.g., responding to non-English prosodic cues in Experiment 1.
-
-Taken together, our data suggest that turn-taking skills do begin to emerge in infancy, but that children cannot consistently make effective predictions until they can identify question turns in unfolding speech and react to them quickly.
-This finding leads us to wonder how participant role (first- instead of third-person) and differences in early interactional experience (e.g., frequent vs.
-infrequent question-asking from caregivers) feed into this early predictive skill.
-It also bridges prior work showing a predisposition for turn taking in infancy (e.g., \citealp{bateson1975, hilbrink2015, jaffe2001, snow1977}) with children's apparently \textit{late} acquisition of adult-like competence for turn taking in spontaneous conversation \citep{casillas2016, garvey1984, garvey1981, ervin-tripp1979}.
-It also reinforces the idea that it takes children several years to fully integrate linguistic information into their turn-taking systems \citep{casillas2016,garvey1981}.
-
-What makes the integration of linguistic information so gradual?
-We suspect that two slow-developing processes---children's linguistic knowledge (e.g., \textit{wh}-words, subject-auxiliary inversion) and their speed of processing for linguistic information (e.g., parsing and retrieval)---both contribute to their ability to make predictions about turn structure in unfolding speech.
-Children may be able to integrate predictive cues for turn taking from the start, but their knowledge of these cues and their speed in parsing and recognizing them may be too slow at first for use in online prediction.
-This account falls in line with the early and continued use of non-verbal cues  found in the current study, but more work is needed to tease these developmental threads apart.
-
-\subsection*{Limitations and future work}
-
-There are at least two major limitations to our work: speech naturalness and participant role.
-Following prior work \citep{de-ruiter2006, keitel2013}, we used phonetically manipulated speech in Experiment 2.
-This decision resulted in speech sounds that children don't usually hear in their natural environment.
-Many prior studies have used phonetically-altered speech with infants and young children \citep[cf.][]{jusczyk2000}, but few of them have done so in a conversational context.
-Future work could instead carefully script speech or cross-splice sub-parts of turns to control for the presence of linguistic cues for turn transition (see, e.g., \citealp{bogelstorreira2015}).
-
-The prediction measure used in our studies is based on an observer's view of conversation but, because participants' role in the interaction could affect their online predictions about turn taking, an ideal measure would instead capture first-person predictions.
-If conversational participants' predictions are partly shaped by their need to respond, first-person measures of spontaneous turn prediction will be key to revealing how participants distribute their attention over verbal and non-verbal cues while taking part in everyday interaction, the implications of which relate to theories of online language processing for both language learning and everyday talk.
-
-That said, the third-person paradigm used in the present study still has much to tell us about turn prediction.
-The task is natural and intuitive in that no instruction is required, which means that it captures spontaneous predictive behavior and can be used with participants of all ages.
-Frequencies of anticipatory gaze switching appear to be stable across language communities where similar tasks have been tested \citep{keitel2013, keitel2015, holler2015, hirvenkari2013}---even from a first-person perspective---so the task is one that measures robust predictive behavior relevant to conversational processing across languages.
-It also lends itself to many possibilities for controlling the presence of individual verbal and non-verbal cues and has a clear method for assessing random switching baselines across the entire stimulus.
-Also, if it is the case that response preparation interferes with our ability to see prediction at the ends of incoming turns \citep{levinson2016}, third-person paradigms are one of the only ways to measure prediction processes in isolation.
-
-The current findings also make predictions about what we would see in first-person paradigms.
-For example, a focus on possible upcoming speaker transitions is even more important when the participants themselves may need to respond; we would thus expect question-like effects to occur in first-person paradigms, and perhaps even be amplified compared to third-person paradigms.
-If so, participants' use of linguistic information would still subserve this goal, with prediction at a premium.
-Regarding development, the same facts about the complexity of prosody-based prediction and children's initial limited lexical inventories would still hold, as would the use of silence and non-verbal cues to assess and predict turn structure in the absence of clear predictive linguistic information.
-The paradigm presented here thus has important contributions to make in our understanding of how participants attend to and make predictions about conversational interaction.
-
-\subsection*{Conclusions}
-
-Conversation plays a central role in children's language learning.
-It is the driving force behind what children say and what they hear.
-Adults use linguistic information to accurately predict turn structure in conversation, which facilitates their online comprehension and allows them to respond relevantly and on time.
-The present study offers new findings regarding the role of speech acts and linguistic processing in online turn prediction, and has given evidence that turn prediction emerges by age two, increases with age, and is driven by the ability to identify and react to question turns in unfolding speech.
-However, children's successful integration of online linguistic processing and online predictions about upcoming turn structure develops gradually.
-When participants can't use predictive linguistic cues (because they are absent, unfamiliar, or are processed too late), children and adults alike rely on retroactive cues such as inter-turn silence to predict upcoming speaker change.
-Using language to make predictions about upcoming interactive content takes time to develop and, for participants of all ages appears to be primarily driven by participants' expectations about what will happen next, beyond the end of the current turn.
-
-\section*{Acknowledgements}
-
-We gratefully acknowledge the parents and children at Bing Nursery School and the Children's Discovery Museum of San Jose.
-This work was supported by an ERC Advanced Grant to Stephen C.
-Levinson (269484-INTERACT), an NSF Graduate Research Fellowship and NSF Dissertation Improvement Grant to MC, and a Merck Foundation fellowship to MCF.
-Earlier versions of these data and analyses were presented to conference audiences \citep{casillas2012, casillas2013}.
-We also thank Tania Henetz, Francisco Torreira, Stephen C.
-Levinson, and Eve V.
-Clark for their feedback on earlier versions of this work.
-The analysis code for this project can be found on GitHub at https://github.com/langcog/turn\_taking/.
-
-\bibliographystyle{model5-names}
-\biboptions{authoryear}
-\bibliography{anticip}
-\clearpage
 
 \appendix
+\renewcommand{\appendixname}{Material}
 
@@
 \setcounter{table}{0}
+\normalsize
 How can we be sure that our primary dependent measure (anticipatory gaze switching) actually relates to turn transitions?
@@
 
-We estimated the baseline probability of making an anticipatory switch by randomly permuting the placement of the transition windows within each stimulus (Figure \ref{fig:shuffling}).
+We estimated the baseline probability of making an anticipatory switch by randomly permuting the placement of the transition windows within each stimulus (Main paper; Figure 4).
 We then used the switch identification procedure from Experiments 1 and 2 to find out how often participants made ``anticipatory'' switches within these randomly permuted windows.
@@
 If the original effect of a predictor exceeded 95\% of the random model effects for that same predictor, we deemed that predictor's effect to be significantly different from the random baseline (i.e., \textit{p}$<$.05).
+
+\newpage
 
@@
 Although this same pattern has yet to be established for children's turn predictions, the looking behavior here is at least consistent with adult response patterns in previous work.
-Additionally, because our analysis windows in the main study only overlapped with the pre-gap utterance by 300 msec (Figure \ref{fig:criterion}), our primary results are unlikely to capture any of these very early or early boredom-driven gaze switches, which makes them unproblematic either way in the current analysis.
+Additionally, because our analysis windows in the main study only overlapped with the pre-gap utterance by 300 msec (Main paper; Figure 2), our primary results are unlikely to capture any of these very early or early boredom-driven gaze switches, which makes them unproblematic either way in the current analysis.
 
@@
 There is therefore a confound between linguistic condition and puppet pair; for example, children could have made fewer anticipatory switches in the \textit{prosody only} condition because the puppets were less interesting.
-To test whether puppet pair drove the condition-based differences found in Experiment 2, we ran a short follow-up study.
+To test whether puppet pair drove the condition-based differences found in Experiment 2, we ran a follow-up study.
 
-\newpage
+\medskip
 \noindent \textbf{Methods}
@@
 We created 6 short videos from the stimulus recordings made for Experiment 2.
-Each video featured a puppet pair (red/blue/yellow/robot/ merpeople/party-goer; Figure \ref{fig:puppets}).
+Each video featured a puppet pair (red/blue/yellow/robot/ merpeople/party-goer; Main paper; Figure 5).
 Puppets in all six videos performed the exact same conversation recording (`birthday party'; Experiment 2) with normal, unmanipulated speech.
@@
 
+\newpage
+
+\bibliographystyle{model5-names}
+\biboptions{authoryear}
+\bibliography{anticip}
+\clearpage
+
 \end{document}

### 2016-06-27 paper/CasillasFrank-Anticip.tex
separated supplementary materials per editor's request

@@
 The random shuffling procedure de-links participants' gaze data from the turn structure in the original stimulus, thereby allowing us to compare turn-related (original) and non-turn-related (randomly permuted) looking behavior using the same eye movement data.
-We created 5,000 permutations of the original turn transitions (step 1 above), thereby creating 5,000 anticipatory gaze datasets with randomly de-linked gaze data (step 2 above).
-Because the randomly shuffled turn transitions could occur anywhere in the stimulus (so long as they didn't overlap each other within a single iteration), the resulting turn-transition windows collectively covered the entire stimulus---during speech and silence, during speaker change and speaker continuation, and during all turn transitions in the stimulus, even those excluded in the original analyses (e.g., because they were overlapped).\footnote{This technique crucially differs from that used by Keitel and colleagues (2013, 2015), which tests anticipatory gaze at turn transitions against anticipatory gaze during speech, which maximized the possibility of finding a difference from the baseline measure.} Pooled together, the anticipatory gaze datasets yielded an average anticipatory switch rate for each participant over all possible starting points in the stimuli: a random baseline.
-
+We created 5,000 permutations of the original turn transitions, thereby creating 5,000 anticipatory gaze datasets with randomly de-linked gaze data.
+Because the randomly shuffled turn transitions could occur anywhere in the stimulus (so long as they didn't overlap each other within a single iteration), the resulting turn-transition windows collectively covered the entire stimulus---during speech and silence, during speaker change and speaker continuation, and during all turn transitions in the stimulus, even those excluded in the original analyses (e.g., because they were overlapped).
+This technique crucially differs from that used by Keitel and colleagues (2013, 2015), which tests anticipatory gaze at turn transitions against anticipatory gaze during speech.
+Pooled together, our 5,000 anticipatory gaze datasets yielded an average anticipatory switch rate for each participant over all possible starting points in the stimuli: a random baseline.
 Using this technique we compared participants' anticipatory switches at turn transition windows to their anticipatory switches over the stimulus as a whole.
 If participants looked randomly back and forth between the speakers, we would have seen similar patterns in both cases.
+
 Rather than simply comparing participants' overall anticipatory switch rates with real and random transition windows, we estimated the likelihood that each of the predictor effects in the original data (e.g., the effect of language condition; Table \ref{tab:E1-models}) could have arisen with random gaze switching: we ran identical statistical models on the real and randomly permuted data sets.
 This tells us not only whether participants' switches were above chance, but whether the specific underlying effects of their anticipatory gaze patterns (e.g., the effect of language condition) were above that expected by chance.
-Because these analyses are complex and secondary to the main results, we report their full details in \ref{sec:permutation}.
+Because these analyses are complex and secondary to the main results, we report their full details in Supplementary material A.
 
@@
 Transition type, the interaction between transition type and gap duration, and the interaction between language condition and transition type showed \textit{z}-values that exceeded 100\%, 99.8\%, and 95\% of random \textit{z}-values, respectively (all \textit{p}$\leq$.05).
-See Section \ref{sec:permutation} for more information on each predictor's random permutation distribution.\footnote{This baseline analysis tests ``random looking'' against ''turn-driven looking'', but it does not test subtypes of turn-driven looking.
-For example, children might switch their gaze from the current speaker to the addressee out of boredom with the ongoing speech rather than active anticipation of an upcoming response.
-We address this hypothesis about ``turn-transition'' gaze switches vs.
-``boredom'' gaze switches in \ref{sec:boredlooks}}
+See Supplementary material A for more information on each predictor's random permutation distribution.\footnote{This baseline analysis tests ``random looking'' against ''turn-driven looking'', but it does not test subtypes of turn-driven looking.
+For example, children might switch their gaze from the current speaker to the addressee out of boredom with the ongoing speech rather than from active anticipation of an upcoming response.
+We address this hypothesis about ``boredom'' gaze switches vs.
+``turn-transition'' gaze switches in Supplementary material C.}
 
@@
 This process yielded a distribution of random permutation-based difference scores that we could then compare to the difference score in the actual data.
-Details are given in \ref{sec:pairwisedev}.
+Details are given in Supplementary material B.
 
@@
 Similarly, the difference in the language condition effect between 5-year-olds and 4-year-olds was greater than would be expected by chance (99.96\% of the data sets; \textit{p}$<$.001).
-See Figure \ref{fig:E1-lgageinteraction} for each difference score distribution.
+See Supplementary material Figure B.1 for each difference score distribution.
 
@@
 Children and adults spontaneously tracked the turn structure of the conversations, making anticipatory gaze switches at an above-chance rate across all ages and conditions.
-Children's anticipatory gaze rates were affected by language condition, transition type, age, and gap duration (Table \ref{tab:E1-models}), none of which could be explained by a baseline of random gaze switching (\ref{sec:permutation}; Figure \ref{fig:E1-ChiTs}).
+Children's anticipatory gaze rates were affected by language condition, transition type, age, and gap duration (Table \ref{tab:E1-models}), none of which could be explained by a baseline of random gaze switching (see Supplementary material Figure A.1).
 These data show a number of important features that bear on our questions of interest.
@@
 Similarly, the adults' data showed significant differentiation from the randomly permuted data for the effect of language condition (\textit{words only} speech; greater than 98.3\% of random \textit{z}-values, \textit{p}$<$.02).
-See Section \ref{sec:permutation} for more information on each predictor's random permutation distribution.
+See Supplementary material A for more information on each predictor's random permutation distribution.
 
@@
 These tests revealed that children's anticipation in the \textit{prosody only} condition significantly improved at ages five and six compared to the \textit{no speech} baseline (with difference scores greater than 95\% of the random data scores; \textit{p}$<$.05).
-See Figure \ref{fig:E2-lgageinteraction} for these \textit{prosody only} difference score distributions.
+See Supplementary material Figure B.2 for these \textit{prosody only} difference score distributions.
 
@@
 We again created pairwise comparisons of the average difference scores for the transition type-language condition interaction across age groups in each random permutation of the data, finding that the effect of transition type in the \textit{normal} speech condition became larger with age, with significant improvements by age 4 over ages 1 and 2 (99.9\% and 98.86\%, respectively), by age 5 over age 4 (97.54\%), and by age 6 over ages 1, 2, and 5 (99.5\%, 97.36\%, and 95.04\%), all significantly different from chance (\textit{p}$<$.05).
-See Figure \ref{fig:E2-lgagetypeinteraction} for these \textit{normal} speech difference score distributions.
+See Supplementary material Figure B.3 for these \textit{normal} speech difference score distributions.
 
@@
 
+\section*{Appendix A. Supplementary material}
+
+Supplementary data associated with this article can be found, in the online version, at XX.
+
 \bibliographystyle{model5-names}
@@
 
-\appendix
-
-\section{Permutation Analyses}
-\label{sec:permutation}
-\setcounter{figure}{0}
-\setcounter{table}{0}
-How can we be sure that our primary dependent measure (anticipatory gaze switching) actually relates to turn transitions?
-Even if children were gazing back and forth randomly during the experiment, we would have still captured some false hits---switches that ended up in the turn-transition windows by chance.
-
-We estimated the baseline probability of making an anticipatory switch by randomly permuting the placement of the transition windows within each stimulus (Figure \ref{fig:shuffling}).
-We then used the switch identification procedure from Experiments 1 and 2 to find out how often participants made ``anticipatory'' switches within these randomly permuted windows.
-This procedure de-links participants' gaze data from turn structure by randomly re-assigning the onset time of each turn-transition in each permutation.
-We created 5,000 of these permutations for each experiment to get an anticipatory switch baselines over all possible starting points.
-
-Importantly, the randomized windows were not allowed to overlap with each other, keeping true to the original stimuli.
-We also made sure that the properties of each turn transition stayed constant across permutations.
-So, while ``transition window A'' might start 2 seconds into Random Permutation 1 and 17 seconds into Random Permutation 2, it maintained the same prior speaker identity, transition type, gap duration, language condition, etc., across both permutations.
-
-We then re-ran the statistical models from the original data on each of the random permutations, e.g., using Experiment 1's original model to analyze the anticipatory switches from each random permutation of the Experiment 1 looking data.
-We could then calculate the proportion of random data \textit{z}-values exceeded by the original \textit{z}-value for each predictor.
-We used the absolute value of all \textit{z}-values to conduct a two-tailed test.
-If the original effect of a predictor exceeded 95\% of the random model effects for that same predictor, we deemed that predictor's effect to be significantly different from the random baseline (i.e., \textit{p}$<$.05).
-
-For example, children's ``language condition'' effect from Experiment 1 had a \textit{z}-value of $|$3.65$|$, which is greater than 99.3\% of all $|$\textit{z}-value$|$ estimates from Experiment 1's random permutation models (i.e., \textit{p}$=$.007).
-It is therefore highly unlikely that the effect of language condition in the original model came from random gaze shifting.
-
-We used this procedure to derive the random-baseline comparison values in the main text (above).
-However, we ran into two issues along the way: first, we had to report \textit{z}-values rather than beta estimates of each effect.
-Second, we had to exclude a substantial portion of the models, especially in Experiment 2 because of model non-convergence.
-We address each of these issues below.
-
-\subsection{Beta, standard error, and z estimates}
-We reported \textit{z}-values in the main text rather than beta estimates because the standard errors in the randomly permuted data models were much higher than for the original data.
-The distributions for each predictor's beta estimate, standard error, and \textit{z}-value for adults and children in each experiment are shown in the graphs below (Figures \ref{fig:E1-ChiTs}--\ref{fig:E2-AduSEs}).
-In each plot, the gray dots represent the absolute value of the 5,000 randomly permuted model estimates for the estimate type plotted (beta, standard error, or \textit{z}), the white circles represent the model estimates from the original data, and the black triangles represent the 95th percentile for each random distribution.
-
-\begin{sidewaysfigure}[!htb]
-  \centering
-  \textbf{Experiment 1: \textit{z}-value estimates}\par\medskip
-  \subfloat[Children]{\includegraphics[width=0.44\textwidth]{figures/E1-chi-randrun-z-vals-absolute.png}\label{fig:E1-ChiTs}}
-  \hfill
-  \subfloat[Adults]{\includegraphics[width=0.44\textwidth]{figures/E1-adu-randrun-z-vals-absolute.png}\label{fig:E1-AduTs}}
-  \caption{Random-permutation and original $|$\textit{z}-values$|$ for predictors of anticipatory gaze rates in Experiment 1.}
-\label{fig:E1-Ts}
-\end{sidewaysfigure}
-
-\begin{sidewaysfigure}[!htb]
-  \centering
-  \textbf{Experiment 1: \textit{$\beta$} estimates}\par\medskip
-  \subfloat[Children]{\includegraphics[width=0.44\textwidth]{figures/E1-chi-randrun-betas-absolute.png}\label{fig:E1-ChiBs}}
-  \hfill
-  \subfloat[Adults]{\includegraphics[width=0.44\textwidth]{figures/E1-adu-randrun-betas-absolute.png}\label{fig:E1-AduBs}}
-  \caption{Random-permutation and original $|$\textit{$\beta$}-values$|$ for predictors of gaze rates in Experiment 1.}
-\label{fig:E1-Bs}
-\end{sidewaysfigure}
-
-\begin{sidewaysfigure}[!htb]
-  \centering
-  \textbf{Experiment 1: \textit{SE} estimates}\par\medskip
-  \subfloat[Children]{\includegraphics[width=0.44\textwidth]{figures/E1-chi-randrun-SEs-absolute.png}\label{fig:E1-ChiSEs}}
-  \hfill
-  \subfloat[Adults]{\includegraphics[width=0.44\textwidth]{figures/E1-adu-randrun-SEs-absolute.png}\label{fig:E1-AduSEs}}
-  \caption{Random-permutation and original \textit{SE}-values for predictors of anticipatory gaze rates in Experiment 1.}
-\label{fig:E1-SEs}
-\end{sidewaysfigure}
-
-\begin{sidewaysfigure}[!htb]
-  \centering
-  \textbf{Experiment 2: \textit{z} estimates}\par\medskip
-  \subfloat[Children]{\includegraphics[width=0.44\textwidth]{figures/E2-chi-randrun-z-vals-absolute.png}\label{fig:E2-ChiTs}}
-  \hfill
-  \subfloat[Adults]{\includegraphics[width=0.44\textwidth]{figures/E2-adu-randrun-z-vals-absolute.png}\label{fig:E2-AduTs}}
-  \caption{Random-permutation and original $|$\textit{z}-values$|$ for predictors of anticipatory gaze rates in Experiment 2.}
-\label{fig:E2-Ts}
-\end{sidewaysfigure}
-
-\begin{sidewaysfigure}[!htb]
-  \centering
-  \textbf{Experiment 2: \textit{$\beta$} estimates}\par\medskip
-  \subfloat[Children]{\includegraphics[width=0.44\textwidth]{figures/E2-chi-randrun-betas-absolute.png}\label{fig:E2-ChiBs}}
-  \hfill
-  \subfloat[Adults]{\includegraphics[width=0.44\textwidth]{figures/E2-adu-randrun-betas-absolute.png}\label{fig:E2-AduBs}}
-  \caption{Random-permutation and original $|$\textit{$\beta$}-values$|$ for predictors of anticipatory gaze rates in Experiment 2.}
-\label{fig:E2-Bs}
-\end{sidewaysfigure}
-
-\begin{sidewaysfigure}[!htb]
-  \centering
-  \textbf{Experiment 2: \textit{SE} estimates}\par\medskip
-  \subfloat[Children]{\includegraphics[width=0.44\textwidth]{figures/E2-chi-randrun-SEs-absolute.png}\label{fig:E2-ChiSEs}}
-  \hfill
-  \subfloat[Adults]{\includegraphics[width=0.44\textwidth]{figures/E2-adu-randrun-SEs-absolute.png}\label{fig:E2-AduSEs}}
-  \caption{Random-permutation and original \textit{SE}-values for predictors of anticipatory gaze rates in Experiment 2.}
-\label{fig:E2-SEs}
-\end{sidewaysfigure}
-
-\clearpage
-
-\subsection{Non-convergent models}
-In comparing the real and randomly permuted datasets, we excluded the output of random-permutation models that gave convergence warnings to remove erratic model estimates from our analyses.
-Non-convergent models made up 22.4--24.4\% of the random permutation models in Experiment 1 and 69--70\% of the random permutation models in Experiment 2.
-The \textit{z}-values for each predictor in the converging and non-converging models from Experiment 1 are shown in Table \ref{tab:nonconv_e1}.
-
-Although many of the non-converging models show estimates within range of the converging models (e.g., with a mean difference of only 0.09 in median \textit{z}-value across predictors), they also show many radically outlying estimates (e.g., showing a mean difference of 237.3 in mean \textit{z}-value across predictors).
-Similar patterns were obtained in the non-converging models for Experiment 2 and persisted across multiple attempts with different optimizers.
-
-We suspect that the issue derives from data sparsity in some of the random permutations.
-This problem is known to occur when there are limited numbers of binary observations in each of a design matrix's bins \citep{allison2004}.
-We could instead use zero-inflated poisson or negative binomial regression models to allow for overdispersion in our data \citep{allison2012}.
-However, these would give us baselines for the normal, convergent model, which is not the aim of this analysis.
-
-\linespread{1}
-\begin{sidewaystable}[!htbp]
-  \begin{scriptsize}
-\centering
-  \begin{tabular}{lllllllllll}
-    & Mean$_{C}$ & Mean$_{NC}$ & Median$_{C}$ & Median$_{NC}$ & \textit{SD}$_{C}$ & \textit{SD}$_{NC}$ & Min$_{C}$ & Min$_{NC}$ & Max$_{C}$ & Max$_{NC}$ \\
-    \hline
-	\textbf{\textit{Children}} &&&&& \\
-    \hline
-    (Intercept)					& -2.52	& -458.42	& -2.54	& -2.86	& 0.87	& 1319.22	& -5.53	& -8185.36		& 0.41	& 0.97 \\
-    Age								& -0.51	& -17.83	& -0.49	& -0.53	& 0.79	& 83.78		& -3.71	& -672.2		& 2.3		& 342.8 \\
-    LgCond							& -0.53	& -109.91	& -0.55	& -0.63	& 0.93	& 564.42	& -3.93	& -4418.74		& 3.23	& 2296.19 \\
-    Type							& -0.1	& -29.66	& -0.09	& -0.1	& 0.98	& 515.12	& -4.06	& -4383.92		& 3.36	& 3416.68 \\
-    Duration						& 0.99	& 345.53	& 0.98	& 1.15	& 1.07	& 1323.13	& -2.44	& -5048.24		& 5.78	& 9985.16 \\
-    Age*LgCond					& 0.19	& 10.64		& 0.2		& 0.18	& 0.9		& 109.6		& -3.31	& -581.61		& 3.59	& 946.81 \\
-    Age*Type						& 0.02	& -1.8		& 0.001	& -0.04	& 0.9		& 98.27		& -3.36	& -884.36		& 3.45	& 640.43 \\
-    LgCond*Type					& 0.2		& 45.32		& 0.2		& 0.27	& 0.96	& 691.3		& -3.12	& -4160.06		& 3.39	& 5107.64 \\
-    Age*LgCond*Type			& -0.12	& -14.23	& -0.12	& -0.15	& 0.93	& 156.72	& -2.98	& -1318.26		& 2.90	& 927.69 \\
-    \hline
-	\textbf{\textit{Adults}} &&&&& \\
-    \hline
-    (Intercept)    			& -1.63	& -126.14	& -1.71	& -1.73	& 0.97	& 713.39	& -4.08	& -12111.22	& 2.15	& 649.55		\\
-    LgCond  						& -0.26	& -679.6	& -0.3	& -0.53	& 1.02	& 15894.33	& -3.45	& -494979.7	& 3.35	& 88581.58	\\
-    Type 							& -0.11	& 6.29		& -0.13	& -0.04	& 1.11	& 501.5		& -3.85	& -6420.75		& 3.28	& 8177.88		\\
-    Duration  					& 0.25	& 84.09		& 0.27	& 0.26	& 1.1		& 1152.94	& -3.25	& -10864.51	& 3.46	& 18540.62	\\
-    LgCond*Type  				& 0.12	& -242.27	& 0.1		& 0.34	& 1.07	& 26836.7	& -3.41	& -622642.7	& 3.81	& 509198.4	\\
-    LgCond*Duration			& 0.15	& 780.03	& 0.16	& 0.39	& 1.04	& 44105.02	& -3.84	& -798498.6	& 3.55	& 1145951	\\
-    Type*Duration 				& 0.05	& -6.56		& 0.05	& 0.02	& 1.13	& 1389.9	& -3.54	& -15979.22	& 3.87	& 16419.46	\\
-    LgCond*Type*Duration	& -0.06	& 1083.63	& -0.08	& -0.21	& 1.1		& 63116.54	& -4.21	& -1201895		& 4.02	& 1284965	\\
-    \hline
-  \end{tabular}
-  \caption{Estimated \textit{z}-values for each predictor in converging (\textit{C}) and non-converging (\textit{NC}) child and adult models from Experiment 1. }
-\label{tab:nonconv_e1}
- \end{scriptsize}
-\end{sidewaystable}
-%\linespread{2}
-
-\section{Pairwise developmental tests}
-\label{sec:pairwisedev}
-\setcounter{figure}{0}
-Experiments 1 and 2 both showed effects of age in interaction with linguistic condition and transition type.
-To explore these effects in more depth, in each permutation we recorded the average difference score for each participant, for each predictor that interacted with age (e.g., English minus non-English anticipatory switches for each participant).
-We then used these values to compute an average difference score over the participants in each age group (e.g., age 3, 4, and 5) within each random permutation.
-This averaging process produces 5,000 baseline-derived difference scores for each age group.
-
-We then made pairwise age comparisons of the difference scores (e.g., the linguistic condition effect in 3-year-olds vs.
-4-year-olds), computing the percent of random-permutation difference scores exceeded by the real-data difference score.
-If the real-data difference score exceeded 95\% of the random-data age difference scores, we deemed it to be an age effect significantly different from chance, e.g., a significant difference between ages three and four in the effect of linguistic condition.
-This procedure is essentially a two-tailed \textit{t}-test, adapted for use with the randomly permuted baseline data.
-
-In each of the plots below, the black dot represents the real data value for the effect being shown.
-The effect sizes from the 5,000 randomly permuted data sets are shown as a distribution.
-The percentage displayed is the percentage of random permutation difference scores exceeded by the original data differences score (taking the absolute value of all data points for a two-tailed test).
-Comparisons marked with 95\% or higher are significant at the \textit{p}$<$0.05 level.
-
-\begin{sidewaysfigure}[!htb]
-\begin{center}
-\textbf{Experiment 1: Age and linguistic condition}\par\medskip
-\includegraphics[width=0.95\textwidth]{figures/E1-child-randvsreal-ttest-agebylg.png}
-\end{center}
-\caption{Pairwise comparisons of the language condition effect across ages in Experiment 1.}
-\label{fig:E1-lgageinteraction}
-\end{sidewaysfigure}
-
-\begin{sidewaysfigure}[!htb]
-\begin{center}
-\textbf{Experiment 2: Age and the \textit{prosody only} condition}\par\medskip
-\includegraphics[width=0.95\textwidth]{figures/E2-child-randvsreal-ttest-muffledages.png}
-\end{center}
-\caption{Significant pairwise comparisons of the \textit{prosody only}-\textit{no speech} linguisitic condition effect, across ages in Experiment 2. Non-significant comparisons are not shown here.}
-\label{fig:E2-lgageinteraction}
-\end{sidewaysfigure}
-
-\begin{sidewaysfigure}[!htb]
-\begin{center}
-\textbf{Experiment 2: Age, transition type, and \textit{normal} speech}\par\medskip
-\includegraphics[width=0.95\textwidth]{figures/E2-child-randvsreal-ttest-normaltypesages.png}
-\end{center}
-\caption{Significant pairwise comparisons of the \textit{normal speech}-\textit{no speech} language condition effect for transition type, across ages, in Experiment 2.
-Non-significant comparisons are not shown here.}
-\label{fig:E2-lgagetypeinteraction}
-\end{sidewaysfigure}
-
-\section{Boredom-driven anticipatory looking}
-\label{sec:boredlooks}
-\setcounter{figure}{0}
-
-One alternative hypothesis for children's anticipatory gazes is that they look at the current speaker at the start of each turn, but then grow bored and start looking away at a constant rate.
-Even though this alternative hypothesis does not predict the primary effects in our data (e.g., the difference between questions and non-questions), we cannot rule out the possibility that a portion of participants' saccades come from boredom.
-
-The data plotted here show a hypothetical group of boredom-driven participants (gray dots) and participants from the actual data in Experiment 2 (black dots).
-The hypothetical boredom-driven participants look away from the current speaker at a linear rate, beginning one second after the start of a turn.
-
-\begin{figure}[!htb]
-\begin{center}
-\includegraphics[width=0.95\textwidth]{figures/boredom-hypothesis.png}
-\end{center}
-\caption{Proportion of participants (hypothetical boredom-driven$=$gray; actual Experiment 2$=$black) looking at the current speaker, split by turn duration.
-Vertical bars indicate standard error in the experimental data.}
-\label{fig:boredomhypothesis}
-\end{figure}
-
-If children's switches away from the current speaker were purely driven by boredom, they would switch away equally quickly on long and short turns.
-Therefore, their crossover point---the point in time at which 50\% of the children have switched away from the current speaker---would be the same for all turns, no matter the length of the turn.
-This pattern is demonstrated in the hypothetical boredom-driven crossover points, which always occur 2.5 seconds after the start of speech (gray vertical lines; Figure \ref{fig:boredomhypothesis}).
-
-In children's \textit{actual} looking data we see that crossover points increase with turn duration: 2.0, 2.9, and 3.6 seconds after the start of speech for turns with durations of 1--2, 2--3, and 3--4 seconds, respectively (black vertical lines; Figure \ref{fig:boredomhypothesis}).
-This pattern suggests that, though children do look away as the turn is unfolding, their looks away are not simply driven by boredom.
-
-Are the looks away in Figure \ref{fig:boredomhypothesis} still too early to count as ``turn-transition'' anticipation?
-It is true that children start looking away after one second has passed, but then only gradually.
-Some of these early looks away may be boredom-driven, but it is equally plausible that some of them are turn-driven.
-Early predictive behavior is common in turn-taking studies with adults, in both constrained turn-taking tasks \citep{de-ruiter2006, gisladottir2015, bogels2015} and in spontaneous conversation \citep{holler2015, torreira2015}.
-Although this same pattern has yet to be established for children's turn predictions, the looking behavior here is at least consistent with adult response patterns in previous work.
-Additionally, because our analysis windows in the main study only overlapped with the pre-gap utterance by 300 msec (Figure \ref{fig:criterion}), our primary results are unlikely to capture any of these very early or early boredom-driven gaze switches, which makes them unproblematic either way in the current analysis.
-
-We therefore conclude that the boredom-driven effects in our data are unlikely to change our primary results, though we acknowledge that characterizing different gaze switching strategies in this kind of data is an important avenue for future work.
-
-\clearpage
-
-\section{Puppet pair and linguistic condition}
-\label{sec:puppetconfound}
-\setcounter{figure}{0}
-The design for Experiment 2 does not fully cross puppet pair (e.g., robots, blue puppets) with linguistic condition (e.g., \textit{words only} and \textit{no speech}).
-Even though each puppet pair is associated with different conversation clips across children (e.g., robots talking about kitties, birthday parties, and pancakes), the robot puppets themselves were exclusively associated with the \textit{words only} condition.
-Similarly, merpeople were exclusively associated with \textit{prosody only} speech, and the puppets wearing dressy clothes were exclusively associated with the \textit{no speech} condition.
-We designed the experiment this way to increase its pragmatic felicity for older children (i.e., robots make robot sounds, merpeople's voices are muffled under the water, the party-going puppets are in a `party' room with many other voices).
-There is therefore a confound between linguistic condition and puppet pair; for example, children could have made fewer anticipatory switches in the \textit{prosody only} condition because the puppets were less interesting.
-To test whether puppet pair drove the condition-based differences found in Experiment 2, we ran a short follow-up study.
-
-\newpage
-\noindent \textbf{Methods}
-\medskip
-
-\noindent We recruited 30 children between ages 3;0 and 5;11 from the Children's Discovery Museum of San Jose, California to participate in our experiment.
-All participants were native English speakers.
-Children were randomly assigned to one of six videos (five children per video).
-
-\medskip
-\noindent \textit{Materials}.
-We created 6 short videos from the stimulus recordings made for Experiment 2.
-Each video featured a puppet pair (red/blue/yellow/robot/ merpeople/party-goer; Figure \ref{fig:puppets}).
-Puppets in all six videos performed the exact same conversation recording (`birthday party'; Experiment 2) with normal, unmanipulated speech.
-This experiment therefore holds all things constant across stimuli except for the appearance of the puppets.
-
-\medskip
-\noindent \textit{Procedure}.
-We used the same experimental apparatus and procedure as in Experiments 1 and 2.
-Each participant was randomly assigned to watch only one of the six puppet videos.
-Five children watched each video.
-As in Experiment 2, the experimenter immediately began each session with calibration and then stimulus presentation because no special instructions were required.
-The entire experiment took less than three minutes.
-
-\medskip
-\noindent \textit{Data preparation}. We identified anticipatory gaze switches to the upcoming speaker using the same method as in Experiments 1 and 2.
-
-\bigskip
-\noindent \textbf{Results and discussion}
-\medskip
-
-\noindent We modeled children's anticipatory switches (yes or no at each transition) with mixed effects logistic regression, including puppet pair (robots/mer- people/party-goers/other-3) as a fixed effect and participant and turn transition as random effects.
-We grouped the red, blue, and yellow puppets together because they collectively represented the puppets used in the \textit{normal} speech condition---this follow-up experiment is meant to test whether the condition-based differences from Experiment 2 arose from the puppets used in each condition.
-
-\begin{figure}[!htb]
-\begin{center}
-\includegraphics[width=0.95\textwidth]{figures/all-puppetdyads-bars.png}
-\end{center}
-\caption{Proportion gaze switches across puppet pairs when linguistic condition and conversation are held constant.}
-\label{fig:pairconfound}
-\end{figure}
-
-\clearpage
-
-\linespread{1}
-\begin{table}
-\begin{center}
-  \begin{tabular}{llcccc}
-%    \hline
-           &  Estimate & Std. Error & \textit{z} value & Pr($>$$|$\textit{z}$|$) \\
-    \hline
-		 \footnotesize{\textit{Reference level: normal-condition puppets}} &&&& \\
-    \hline
-     (Intercept) 							& -0.148	& 0.328		& -0.451		& 0.652 \\
-     Puppets$=$\textit{mermaid} 	& -0.076	& 0.655		& -0.116		& 0.908 \\
-     Puppets$=$\textit{robot} 		& -0.071	& 0.653		& -0.109		& 0.913 \\
-     Puppets$=$\textit{party} 		& -0.782	& 0.687		& -1.138		& 0.255 \\
-    \hline
-		 \footnotesize{\textit{Reference level: mer-puppets}} &&&& \\
-    \hline
-     (Intercept) 							& -0.224	& 0.568		& -0.394		& 0.694 \\
-     Puppets$=$\textit{robot} 		& 0.0048	& 0.801		& 0.006			& 0.995 \\
-     Puppets$=$\textit{party} 		& -0.706	& 0.827		& -0.854		& 0.393 \\
-    \hline
-		 \footnotesize{\textit{Reference level: robot puppets}} &&&& \\
-    \hline
-     (Intercept) 							& -0.219	& 0.566		& -0.387		& 0.699 \\
-     Puppets$=$\textit{party} 		& -0.711	& 0.827		& -0.860		& 0.390 \\
-    \hline
-		 \footnotesize{\textit{Reference level: party-goer puppets}} &&&& \\
-    \hline
-     (Intercept) 							& -0.93		& 0.607		& -1.533		& 0.125 \\
-  \end{tabular}
-\end{center}
-  \caption{Model output for children's anticipatory gaze switches with reference levels varied to show all possible pairwise differences between puppet pairs.}
-\label{tab:control_exp}
-\end{table}
-%\linespread{2}
-
-In four versions of this model, we systematically varied the reference level of the puppet pair to check for any cross-condition differences.
-We found no significant effects of puppet pair on switching rate (all \textit{p}$>$0.25; Table \ref{tab:control_exp}).
-
-We take this finding as evidence that our decision to not fully cross puppet pairs and linguistic conditions in Experiment 2 was unlikely to have affected children's anticipatory gaze rates above and beyond the intended effects of linguistic condition.
-
 \end{document}
